### 2016-04-20 model.tex
added notes

\section{Single fluid description of well-coupled dusty gas}\label{setup}

\cite{laibe14}, Eq. 84 --- 85, 87:

\begin{align}
  &\frac{D\rho}{Dt} = -\rho\nabla\cdot\bm{v}, \label{masseq}\\ 
  &\frac{D\bm{v}}{Dt} = - \nabla\Phi - \frac{1}{\rho}\nabla
  P, \label{momeq}\\ 
  &\frac{D\epsilon}{Dt} = -\frac{1}{\rho} \nabla \cdot \left(\epsilon 
  \tstop \nabla P \right),\label{dusteq}  
\end{align}
where $D/Dt \equiv \p_t + \bm{v}\cdot\nabla$ is the Lagragian
derivative; 
\begin{align}
  \rho \equiv \rhog + \rhod
\end{align}
is the total density with $\rhog$ and $\rhod$ being the gas and dust
density, respectively; 
\begin{align}
  \bm{v} \equiv \frac{\rhog\bm{v}_\mathrm{g} +
    \rhod\bm{v}_\mathrm{d}}{\rho}
\end{align}
is the center-of-mass velocity with $\bm{v}_\mathrm{g}$ and
$\bm{v}_\mathrm{d}$ being the gas and dust velocities, respectively; 
and 
\begin{align}
  \epsilon = \frac{\rhod}{\rho} 
\end{align}
is the dust fraction; and $\tstop$ is the stopping time.

\subsection{Locally isothermal equation of state for the gas}

\begin{align}\label{eos}
  P = c_s^2(r,z)\rhog = c_s^{2}(1 - \epsilon)\rho,   
\end{align}

$c_s$ is a prescribed sound-speed profile fixed in time.

\subsection{Effective energy equation for the mixture with a
  stationary temperature profile}   
Eliminate $\epsilon$ in favor of $P$,

\begin{align}
  \epsilon = 1 - \frac{P}{c_s^2(r,z)\rho}, 
\end{align}

Eq. \ref{dusteq} becomes

\begin{align}
\frac{DP}{Dt} = - P \nabla\cdot\bm{v} + P\bm{v}\cdot\nabla\ln{c_s^2} +
c_s^2 \nabla\cdot\left[\tstop\left(1 - \frac{P}{c_s^2\rho}\right)\nabla
  P\right]. \label{eff_energy}
\end{align}

This is the energy equation for an ideal gas of adiabatic index
$\gamma=1$, with additional source terms due to the imposed
temperature profile, as well as dust-gas friction.

Eq. \ref{eff_energy} can be written in conservative form,

\begin{align*}
  \frac{\p P}{ \p t} + \nabla\cdot\left\{P\left[\bm{v} -
      \tstop\left(c_s^2 - \frac{P}{\rho}\right)\nabla\ln{P}\right]
    \right\}\\
  = \left[P\bm{v} - \tstop\left(c_s^2 - \frac{P}{\rho}\right)\nabla
    P\right]\cdot\nabla\ln{c_s^2}. 
\end{align*}
note: non-linear in $P$

\section{Equilibrium}

\begin{align}
  r\Omega^2 &= \frac{\p \Phi}{\p r} + \frac{1}{\rho}\frac{\p P}{\p
    r},\\
  0 & = \frac{\p\Phi}{\p z} + \frac{1}{\rho}\frac{\p P}{\p z}.
\end{align}

\subsection{Vertical shear}

\begin{align}\label{vshear}
  r\frac{\p \Omega^2}{\p z} = \frac{\p\ln{\rho}}{\p r}\frac{\p}{\p
    z}\left[c_s^2(1-\epsilon)\right] - \frac{\p\ln{\rho}}{\p z}
  \frac{\p}{\p r} \left[c_s^2(1-\epsilon)\right]. 
\end{align}

\section{Limiting behaviours}

\subsection{Strictly isothermal gas perfectly coupled to dust}  
When $c_s^2$ is a constant and $\tstop=0$, the dusty gas equations are
exactly equivalent to that for adiabatic hydrodyamics with unit adiabatic
index. Although the gas is strictly isothermal, the mixture behaves 
adiabatically because the dust fraction $\epsilon$ is advected with
the gas (i.e. evolves like an entropy when there is no heating or
cooling).

In this case one may immediately apply the Solberg-Hoiland
criteria to assess axisymmtric \emph{stability}:

\begin{align}
  \kappa^2 - \frac{1}{\rho}\nabla P \cdot \nabla s &> 0,\\
  -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, 
\end{align}  
where 
\begin{align}
  s \equiv \ln \frac{P}{\rho} = \ln{(1-\epsilon)} 
\end{align}
may be interpreted as the mixture entropy. The first equality is
similar to the usual defintion for gas entropy (with unit adiabatic
index). The second follows from the imposed equation of state,
Eq. \ref{eos}, with a constant sound-speed.

The stability conditions are thus
\begin{align}
  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\epsilon &> 0,  \\
  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\epsilon}{\p
    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \epsilon}{\p r} \right) & > 0. 
\end{align}
For accretion disks, $\rhog$ and $\epsilon$ are expected to decrease
outwards and away from the midplane, implying 
$\nabla\rhog\cdot\nabla\epsilon>0$, so the first criterion is
typically satisifed.

To evaulate the second criterion, we can consider $z>0$ without loss
of generality. Then $\p_z\rhog,\p_z\rhod <0$, so $\p_z\rho<0$. We
assume $\p_r\rho <0$ and $\p_z\epsilon, \p_r\epsilon <0$ as
well\footnote{Due to dust sedimentation and inwards migration.}. Then,
inserting the vertical shear profile, Eq. \ref{vshear}, we find the
second criterion becomes

\begin{align} 
  \kappa^2\left|\frac{\p\epsilon}{\p z}\right| + c_s^2\left|\frac{\p
    \epsilon}{\p r}\frac{\p\epsilon}{\p z}\frac{\p\ln{\rho}}{\p
    r}\right| - c_s^2 \left(\frac{\p\epsilon}{\p
    r}\right)^2\left|\frac{\p\ln{\rho}}{\p z}\right| > 0.  \label{criterion2}
\end{align}

This equation shows that the vertical gradient in dust-fraction (and
thus the dust-to-gas ratio $\rhod/\rhog$) has a stabilizing effect for
axisymmetric perturbations. Instead, the only way to destabilize the
system is a sufficiently large radial gradient in the dust-fraction
and/or vertical (total) density gradient (the last term).

Note that Eq. \ref{criterion2} is only applicable for the assume disk
structure described above: decreasing dust and gas densities, as well
as the dust-fraction, away from the star and away from the
midplane. For less typical disk structures, one must evalulate above the
critera seperately.

\subsection{Locally isothermal gas perfectly coupled to dust} 
$c_s(r,z)$ is non-uniform, but $\tstop=0$.

### 2016-04-20 paper.tex
added notes

%\documentclass[12pt]{aastex}
\documentclass[iop, numberedappendix]{emulateapj}
%\documentclass{article}
%\usepackage{draftwatermark}
%\usepackage{paralist}
\shorttitle{untitled}
\shortauthors{M-K.\ Lin}

\usepackage{amsmath}
%\usepackage{url}
\usepackage{bm}
%\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\newcommand{\p}{\partial}
\newcommand{\zmax}{z_\mathrm{max}}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\sbar}{\bar{\sigma}} 
\newcommand{\He}{\operatorname{He}}
\newcommand{\imgi}{\mathrm{i}} 
\newcommand{\ii}{\mathrm{i}}
\newcommand{\C}{\mathcal{C}^\lambda}
\newcommand{\wop}[1]{\mathcal{V}^{(#1)}}
\newcommand{\avg}[1]{\langle{#1}\rangle}
%\newcommand{\max}[1]{\mathrm{max}({#1})}
%\newcommand{\min}[1]{\mathrm{min}({#1})}
\newcommand{\zhat}{\hat{z}}
\newcommand{\khat}{\hat{k}}
\newcommand{\rhat}{\hat{r}}
\newcommand{\he}{\mathcal{H}}
\newcommand{\amp}{\mathcal{A}}
\newcommand{\dd}{\delta}
\newcommand{\real}{\operatorname{Re}}
\newcommand{\imag}{\operatorname{Im}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\wbar}{\tilde{W}}
\newcommand{\qbar}{\tilde{Q}}
\newcommand{\dbigH}{\frac{H^\prime}{H}}
\newcommand{\Rp}{\frac{\rho_0^\prime}{\rho_0}}
\newcommand{\Fp}{\frac{g^\prime}{g}}
\newcommand{\Rpp}{\frac{\rho_0^{\prime\prime}}{\rho_0}}
\newcommand{\Fpp}{\frac{g^{\prime\prime}}{g}}
\newcommand{\ils}{L_s^{-1}}
\newcommand{\ilp}{L_p^{-1}}
\newcommand{\ihp}{H_p^{-1}}
\newcommand{\ihs}{H_s^{-1}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\csmid}{c_{s0}} 
\newcommand{\hiso}{h_\mathrm{iso}}
\newcommand{\Hiso}{H_\mathrm{iso}}
\newcommand{\zeus}{\texttt{ZEUS-MP }}
\newcommand{\rsph}{r_\mathrm{sph}}
\newcommand{\dvx}{\dd v_x}
\newcommand{\dvy}{\dd v_y}
\newcommand{\dvz}{\dd v_z}
\newcommand{\dbx}{\dd B_x}
\newcommand{\dby}{\dd B_y}
\newcommand{\dbz}{\dd B_z}
\newcommand{\rhod}{\rho_\mathrm{d}}
\newcommand{\rhog}{\rho_\mathrm{g}}
\newcommand{\tstop}{t_\mathrm{s}}
\newcommand{\w}{ \widetilde{W}}
\newcommand{\dphi}{\dd \Phi}
\newcommand{\actaa}{Acta Astron.}
\defcitealias{lin12}{L12} 
\defcitealias{nelson13}{N13}
\defcitealias{barker15}{BL15}
\defcitealias{mcnally14}{MP14}
\defcitealias{goldreich67}{GS67}

\def \OmK {\Omega_{\rm K}}

\begin{document}

\title{untitled}
%\title{Vertical shear instability in astrophysical disks: vertically
%global linear calculations with thermal relaxation and applications
%to the Solar Nebula}  
\author{Min-Kai Lin\altaffilmark{1}}
\affil{Department of Astronomy and Steward Observatory, \\
University of
  Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA}
\altaffiltext{1}{Steward Theory Fellow}
\email{minkailin@email.arizona.edu}

\begin{abstract}
dusty vsi, or the lack thereof
\end{abstract}

\input{model}
%\input{linear_iso}
%\input{linear_adia}
%\input{numerical}
%\input{application} 
%\input{summary}

%\appendix
%\input{appendix}

\bibliographystyle{apj}
\bibliography{ref}

\end{document}

### 2016-04-21 appendix.tex
updated notes

\section{Variational principle for locally isothermal gas perfectly
  coupled to dust}\label{var_prin}

Consider axisymmetric disturbances of the form
\begin{align}
  \delta\rho(r,z)\exp{\left(-\ii\sigma\right)}. 
\end{align}
Linearized equations for dusty gas with $\tstop=0$, and using the
pressure equation in place of the dust-fraction
(Eq. \ref{masseq}---\ref{momeq}, Eq. \ref{eff_energy}):

\begin{align}
  \ii\sigma\frac{\delta\rho}{\rho} &= \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{\rho},\\
  \ii\sigma\frac{\delta P}{P} &= \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{c_s^2},\\
  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
  \delta\bm{F}\cdot\hat{\bm{r}},\\
  \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
    v_\phi}{\p z}\dd v_z,\\
  -\ii\sigma\dd v_z &=  \delta\bm{F}\cdot\hat{\bm{z}},
\end{align} 
with
\begin{align}
  \delta \bm{F} \equiv \frac{\dd\rho}{\rho^2}\nabla P -
  \frac{1}{\rho}\nabla\dd P.
\end{align}
Note that for the axisymmetric problem, $\dd\bm{F}$ is purely
meridional. 
From the linearized meridional momentum equations, we find 
\begin{align}
  \sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV = \int\left( \rho
  \kappa^2 |\dd v_r|^2 + \rho r\frac{\p \Omega^2}{\p z} \dd v_z \dd
  v_r^*  + \ii \rho \dd \bm{F}\cdot\dd\bm{v}^*
  \right)dV,
\end{align}
where the integral is taken over the volume of the fluid. Integrating
by parts and ignoring surface integrals, the last term is
\begin{align}
  \int \ii\sigma\rho \dd \bm{F}\cdot\dd\bm{v}^*dV &=\int\left( \ii\sigma \frac{\dd
    \rho}{\rho}\dd\bm{v}^*\cdot \nabla P - \ii\sigma \dd
  \bm{v}^*\cdot\nabla \dd P  \right)dV\notag\\
&= \int\left( \ii\sigma \frac{\dd
    \rho}{\rho}\dd\bm{v}^*\cdot\nabla P + \ii\sigma \dd
 P  \nabla\cdot \dd \bm{v}^*  \right)dV\notag\\
 & = \int\left[
   \left(
   \ii\sigma \frac{\dd P}{P} - \dd\bm{v}\cdot\nabla s + \dd
   \bm{v} \cdot \nabla \ln{c_s^2}
   \right)\dd\bm{v}^*\cdot\nabla P  + \ii\sigma \dd
   P  \nabla\cdot \dd \bm{v}^*
   \right]dV\notag\\
 & = \int\left[
   \ii\sigma\frac{\dd P}{P}\nabla\cdot\left(P\dd\bm{v}^*\right) +
   \left(\dd\bm{v}^*\cdot\nabla P\right) \left(\dd\bm{v} \cdot \nabla
   \ln{c_s^2}   -  \dd\bm{v}\cdot\nabla s\right) 
   \right]dV\notag\\
 & = \int\left\{
   \left[\frac{1}{P}\nabla\cdot\left(P\dd\bm{v}\right) -
   \dd\bm{v}\cdot\nabla\ln{c_s^2}\right]\nabla\cdot\left(P\dd\bm{v}^*\right)
   +\left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v} \cdot \nabla
   \ln{c_s^2}   -  \dd\bm{v}\cdot\nabla s \right)
   \right\}dV \notag\\
   &=\int\left[
     \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2 -
     \left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v}\cdot\nabla s
     \right) -
     P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right) 
     \right]dV,
\end{align}
where we recall $s\equiv\ln{(P/\rho)}$. Hence,
\begin{align}\label{integral_ex}
  &\sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV \notag\\
&=  \int\left\{
  \rho |\dd v_r|^2 \left(\kappa^2 - \frac{1}{\rho}\frac{\p P}{\p
    r}\frac{\p s}{\p r}\right)
  + \rho |\dd v_z|^2\left(-\frac{1}{\rho}\frac{\p P}{\p
    z}\frac{\p s}{\p z}\right)
   + \rho \dd v_z \dd v_r^*\left(r\frac{\p\Omega^2}{\p z} -
  \frac{1}{\rho}\frac{\p P}{\p
    r}\frac{\p s}{\p z}\right) 
  + \rho \dd v_z^*\dd v_r \left(-\frac{1}{\rho}\frac{\p P}{\p
    z}\frac{\p s}{\p r}\right)\right. \notag\\
&\phantom{==\int}\left.+
  \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2
  \right\}dV -\int P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV.
\end{align}
Note that the coefficients of the mixed terms are equal owing to the
equilibrium state.

%If $c_s$ is constant then the second integral vanishes, $\sigma^2$
%is real, and the first integral leads to the usual Solberg-Hoiland
%criteria. However, for a stationary but non-uniform temperature
%profile, the second

### 2016-04-21 model.tex
updated notes

@@
 \begin{align}\label{vshear}
-  r\frac{\p \Omega^2}{\p z} = \frac{\p\ln{\rho}}{\p r}\frac{\p}{\p
+  r\frac{\p \Omega^2}{\p z} &= \frac{\p\ln{\rho}}{\p r}\frac{\p}{\p
     z}\left[c_s^2(1-\epsilon)\right] - \frac{\p\ln{\rho}}{\p z}
-  \frac{\p}{\p r} \left[c_s^2(1-\epsilon)\right]. 
+  \frac{\p}{\p r} \left[c_s^2(1-\epsilon)\right]\\  
+  & = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p s}{\p z} -\frac{\p
+    P}{\p z}\frac{\p s}{\p r} \right),\notag
 \end{align}
+where
+\begin{align}
+   s \equiv \ln \frac{P}{\rho}
+\end{align}
+may be interpreted as the mixture entropy.
 
@@
   -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
-  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, 
+  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0. 
 \end{align}  
-where 
-\begin{align}
-  s \equiv \ln \frac{P}{\rho} = \ln{(1-\epsilon)} 
-\end{align}
-may be interpreted as the mixture entropy. The first equality is
-similar to the usual defintion for gas entropy (with unit adiabatic
-index). The second follows from the imposed equation of state,
-Eq. \ref{eos}, with a constant sound-speed.
+Here we may take $s=\ln{1-\epsilon}$ from the definition of $s$ and
+the equation of state (Eq. \ref{eos}) with constant sound-speed.
+
+%\begin{align}
+%  s \equiv \ln \frac{P}{\rho} = \ln{(1-\epsilon)} 
+%\end{align}
+% The first equality is
+%similar to the usual defintion for gas entropy (with unit adiabatic
+%index). The second follows from the imposed equation of state,
+%Eq. \ref{eos}, with a constant sound-speed.
 
@@
 \subsection{Locally isothermal gas perfectly coupled to dust} 
-$c_s(r,z)$ is non-uniform, but $\tstop=0$.
+$c_s(r,z)$ is non-uniform, but $\tstop=0$. Integral relation for
+linear, axisymmetric waves with frequency $\sigma$ (see Appendix \ref{var_prin})
+\begin{align}
+&  \sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV \notag\\
+&= \int\left[ \rho
+  |\dd v_r|^2A + \rho  \dd v_z \dd v_r^* B + \rho \dd v_z^*\dd v_r C +
+  \rho |\dd v_z|^2 D\phantom{\frac{1}{1}}\right. \notag\\
+&\phantom{===}  \left. + \frac{1}{P}\left|\nabla\cdot\left(P\dd
+  \bm{v}\right)\right|^2\right]dV - \int P
+  \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV.
+\end{align}
+The real coefficients $A,B,C,D$ can be read off
+Eq. \ref{integral_ex}. If $c_s$ is constant, $\sigma$ is real, and 
+the first integral leads to the usual Solberg-Hoiland criteria for
+axisymmetric stability. However, when $c_s$ is non-uniform, $\sigma$
+is generally complex, implying instability. This is the usual vertical
+shear instability.

### 2016-04-22 model.tex
order of mag estimate

@@
 \section{Single fluid description of well-coupled dusty gas}\label{setup} 
-
-\cite{laibe14}, Eq. 84 --- 85, 87:
+In the limit of strong dust-gas drag (stopping time $\tstop\to 0$) 
+\cite{laibe14} give a single fluid description of the mixture:  
+(their Eq. 84 --- 85 and Eq. 87):
 
@@
 \begin{align}
-  \epsilon = \frac{\rhod}{\rho} 
+  \epsilon \equiv \frac{\rhod}{\rho}  = \frac{\tepsilon}{1+\tepsilon} 
 \end{align}
-is the dust fraction; and $\tstop$ is the stopping time.
+is the dust-fraction and $\tepsilon=\rhod/\rhog$ is the usual
+dust-to-gas ratio. The gas pressure $P$ is given by the chosen 
+equation of state described below. The external potential $\Phi$ is
+that for a star of mass $M_*$ at the origin 
+\begin{align}
+  \Phi(r,z) =-\frac{GM_*}{\sqrt{r^2 + z^2}},
+\end{align}
+where $G$ is the gravitational constant.
 
 \subsection{Locally isothermal equation of state for the gas} 
-
+We adopt 
 \begin{align}\label{eos}
@@
 \end{align}
+where $c_s$ is a prescribed sound-speed profile fixed in time.
 
-$c_s$ is a prescribed sound-speed profile fixed in time.
+\subsection{Effective mixture energy equation}
+We show that for a prescribed temperature distribution, the mixture
+obeys an evolutionary energy equation, due to the advection of the
+dust-fraction.
 
-\subsection{Effective energy equation for the mixture with a
-  stationary temperature profile}   
-Eliminate $\epsilon$ in favor of $P$,
-
+We eliminate $\epsilon$ in favor of $P$, 
 \begin{align}
@@
 \end{align}
-
-Eq. \ref{dusteq} becomes
+then Eq. \ref{dusteq} becomes
 
@@
 This is the energy equation for an ideal gas of adiabatic index
-$\gamma=1$, with additional source terms due to the imposed
-temperature profile, as well as dust-gas friction.
+$\gamma=1$, i.e. isothermal evolution, but now with additional source 
+terms due to the imposed temperature profile, as well as dust-gas
+friction.
 
 Eq. \ref{eff_energy} can be written in conservative form,
-
 \begin{align*}
@@
 \end{align*}
-note: non-linear in $P$
+We may thus re-interpret $P$ as the mixture's energy density, but the
+energy flux has an additional contribution from the pressure
+gradient. The term on the right-hand-side, owing to the imposed
+temperature profile, can be interpreted as an external heat source.
 
 \section{Equilibrium}
-
+To obtain an axisymmetric steady state with $\rho(r,z)$ and 
+$\bm{v}=r\Omega(r,z)\hat{\bm{\phi}}$ where $\Omega = v_\phi/r$, 
+we need to solve 
 \begin{align}
   r\Omega^2 &= \frac{\p \Phi}{\p r} + \frac{1}{\rho}\frac{\p P}{\p
-    r},\\
-  0 & = \frac{\p\Phi}{\p z} + \frac{1}{\rho}\frac{\p P}{\p z}.
+    r},\label{steady_momr}\\
+  0 & = \frac{\p\Phi}{\p z} + \frac{1}{\rho}\frac{\p P}{\p z},\label{steady_momz}\\
+  0 & = \nabla\cdot\left(\epsilon\tstop\nabla P\right) \label{steady_dust}
 \end{align}
+for $\rho$, $\Omega$ and $\epsilon$ with $P=P(\epsilon,\rho)$ given by
+the equation of state (Eq. \ref{eos}).
+
+Eq. \ref{steady_dust} makes it difficult to obtain equilibrium
+solutions explicitly. However, if the diffusive process is slow and
+can be neglected, then one may just solve
+Eq. \ref{steady_momr}---\ref{steady_momz} with a prescribed (initial)
+distribution of the dust fraction $\epsilon(r,z)$.
 
 \subsection{Vertical shear}
-
+The mixture possess vertical shear. To see this, we eliminate $\Phi$
+between Eq. \ref{steady_momr}---\ref{steady_momz} to 
+obtain 
 \begin{align}\label{vshear}
@@
 \end{align}
-may be interpreted as the mixture entropy.
+may be consider as a pseudo-entropy for the mixture.
 
@@
 \end{align}  
-Here we may take $s=\ln{1-\epsilon}$ from the definition of $s$ and
-the equation of state (Eq. \ref{eos}) with constant sound-speed.
+These criteria are applicable to any adiabatic fluid when $s$
+is the usual definition of entropy. For the dusty problem, we may take 
+$s=\ln{(1-\epsilon)}+\mathrm{const.}$ from the definition of $s$ and  
+the equation of state (Eq. \ref{eos}) with uniform sound-speed.
 
@@
 \end{align}
-For accretion disks, $\rhog$ and $\epsilon$ are expected to decrease
-outwards and away from the midplane, implying 
+In accretion disks, $\rhog$, $\rhod$ and $\epsilon$ are expected to
+decrease outwards and away from the midplane, 
+\begin{align*}  
+  \frac{\p\rhog}{\p |z|} < 0, \quad \frac{\p \rhog}{\p r} < 0,   
+\end{align*}
+and similarly for $\rhod$ and $\epsilon$. Thus 
 $\nabla\rhog\cdot\nabla\epsilon>0$, so the first criterion is
@@
 To evaulate the second criterion, we can consider $z>0$ without loss
-of generality. Then $\p_z\rhog,\p_z\rhod <0$, so $\p_z\rho<0$. We
-assume $\p_r\rho <0$ and $\p_z\epsilon, \p_r\epsilon <0$ as
-well\footnote{Due to dust sedimentation and inwards migration.}. Then,
-inserting the vertical shear profile, Eq. \ref{vshear}, we find the
-second criterion becomes
+of generality. Then, inserting the vertical shear profile, 
+Eq. \ref{vshear}, we find the second criterion becomes
 
@@
 
+\subsubsection{Order-of-magnitude estimates}
+Let us assume the gas and dust distributions have a Gaussian
+distribution vertically,
+\begin{align}
+  \rho_\mathrm{g,d} = \rho_\mathrm{g0,d0} \exp
+      {\left(-\frac{z^2}{2H^2_\mathrm{g,d}}\right)}.
+\end{align}
+Then the dust-to-gas ratio is 
+\begin{align}
+  \tepsilon = \tepsilon_0
+  \exp{\left(-\frac{z^2}{2\widetilde{H}^2}\right)}, 
+\end{align}
+with $\tepsilon_0= \rho_\mathrm{d,0}/\rho_{g_0}$, and  
+\begin{align}
+  \frac{1}{\widetilde{H}^2} \equiv \frac{1}{H_\mathrm{d}^2} -
+  \frac{1}{H_\mathrm{g}^2}. 
+\end{align}
+We assume the dust layer is much thinner than the gas scale-height,
+$H_\mathrm{d}\ll H_\mathrm{g}$, so that 
+\begin{align}
+ \widetilde{H} \sim H_\mathrm{d}. 
+\end{align}
+
+We further assume the gas and dust
+distributions are smoothly varying in the radial direction so that  
+\begin{align}
+  \frac{\p}{\p r}\sim \frac{1}{r}, 
+\end{align}
+the disk is roughly Keplerian so that
+\begin{align}
+  \kappa \sim \OmK, 
+\end{align}
+and the gas sound-speed is 
+\begin{align}
+  c_s \sim h_\mathrm{g}r\OmK, 
+\end{align}
+where $h_\mathrm{g}\ll1$ is the aspect-ratio of the gas component.
+
+Inserting the above approximations into Eq. \ref{criterion2}, and
+evaluating it at $z = H_\mathrm{d}$, we find 
+\begin{align}
+1 + \frac{h_\mathrm{g}^2\tepsilon}{\left(1 + \tepsilon\right)^2}
+\left[1 - \left(\frac{H_\mathrm{d}}{H_\mathrm{g}}\right)^2 +
+  \frac{\tepsilon}{1+\tepsilon}\left(1 -
+  \frac{H_\mathrm{d}}{H_\mathrm{g}}\right)   
+\right]>0.
+\end{align}
+
 \subsection{Locally isothermal gas perfectly coupled to dust} 
-$c_s(r,z)$ is non-uniform, but $\tstop=0$. Integral relation for
-linear, axisymmetric waves with frequency $\sigma$ (see Appendix \ref{var_prin})
+When $c_s(r,z)$ is non-uniform but $\tstop=0$, simple stability
+criteria are not available. However, we may obtain an integral
+relation for linear, axisymmetric waves with frequency $\sigma$ (see
+Appendix \ref{var_prin}) in the form:  
 \begin{align}
@@
 The real coefficients $A,B,C,D$ can be read off
-Eq. \ref{integral_ex}. If $c_s$ is constant, $\sigma$ is real, and 
+Eq. \ref{integral_ex}. Note that $B=C$ from the assume equilibrium
+structure (Eq. \ref{vshear}).
+
+If $c_s$ is constant, then $\sigma^2$ is real, and 
 the first integral leads to the usual Solberg-Hoiland criteria for
-axisymmetric stability. However, when $c_s$ is non-uniform, $\sigma$
-is generally complex, implying instability. This is the usual vertical
-shear instability.
+axisymmetric stability. However, when $c_s$ is non-uniform, $\sigma^2$
+is generally complex, implying instability. This is responsible for
+the usual vertical shear instability in locally isothermal disks.
+
+\subsection{Strictly isothermal gas well-coupled to dust} 
+Now consider $c_s^2=\mathrm{const.}$ but $\tstop\neq0$. In this case
+the effective energy equation is 
+\begin{align}\label{dusty_cooling}
+  \frac{DP}{Dt} &= - P \nabla \cdot \bm{v} + \mathcal{C},\\
+  \mathcal{C} & = \nabla\cdot\left[\tstop\left(c_s^2 -
+    \frac{P}{\rho}\right)\nabla P\right].           
+\end{align} 
+The function $\mathcal{C}$ can be considered as a cooling term. By
+analogy with enabling VSI in an adiabatic gas through cooling
+\citep{nelson13,lin15}, we may expect a `dusty VSI' to develop 
+if this term is large enough (physically meaning diffusion of the
+dust-fraction is strong).
+
+Note that $\mathcal{C}$ is also responsible for the streaming
+instability  in the strong-drag limit \citep{laibe14}, though this
+does not require vertical stratification. In the form of
+Eq. \ref{dusty_cooling}, however, the streaming instability has a
+thermodynamical interpretation, similar to thermal instabilities in
+stars \citep{latter06}.

### 2016-04-23 model.tex
refining

@@
   -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
-  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0. 
+  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, 
 \end{align}  
+where $\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)$ is the
+square of the epicylic frequency. We assume $\kappa^2>0$. 
 These criteria are applicable to any adiabatic fluid when $s$
@@
 The stability conditions are thus
-\begin{align}
+\begin{align}\label{dusty_solberg}
   \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\epsilon &> 0,  \\
@@
 \end{align}
+
 In accretion disks, $\rhog$, $\rhod$ and $\epsilon$ are expected to
-decrease outwards and away from the midplane, 
+decrease away from the mid-plane
 \begin{align*}  
-  \frac{\p\rhog}{\p |z|} < 0, \quad \frac{\p \rhog}{\p r} < 0,   
+  \frac{\p\rhog}{\p |z|} < 0,
+%, \quad \frac{\p \rhog}{\p r} < 0,   
 \end{align*}
-and similarly for $\rhod$ and $\epsilon$. Thus 
-$\nabla\rhog\cdot\nabla\epsilon>0$, so the first criterion is
-typically satisifed.
-
-To evaulate the second criterion, we can consider $z>0$ without loss
-of generality. Then, inserting the vertical shear profile, 
-Eq. \ref{vshear}, we find the second criterion becomes
+and similarly for $\rhod$, $\epsilon$, and the dust-to-gas ratio $\tepsilon$. 
+We may thus evaluate the stability criteria at $z>0$ without loss of
+generality. We insert the vertical shear profile,
+Eq. \ref{vshear}, into Eq. \ref{dusty_solberg} and write the result 
+in terms of $\tepsilon$, to find
 
 \begin{align} 
-  \kappa^2\left|\frac{\p\epsilon}{\p z}\right| + c_s^2\left|\frac{\p
-    \epsilon}{\p r}\frac{\p\epsilon}{\p z}\frac{\p\ln{\rho}}{\p
-    r}\right| - c_s^2 \left(\frac{\p\epsilon}{\p
-    r}\right)^2\left|\frac{\p\ln{\rho}}{\p z}\right| > 0.  \label{criterion2}
+  &1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1+
+    \tepsilon\right)^2}\frac{\Omega^2}{\kappa^2}\left(\frac{\p\ln{\rhog}}{\p\ln{r}}
+  \frac{\p\ln{\tepsilon}}{\p\ln{r}} + r^2\frac{\p\ln{\rhog}}{\p z}
+  \frac{\p\ln{\tepsilon}}{\p z} \right)>0,   \\
+&-\frac{\p\ln{\tepsilon}}{\p z} \notag\\
+  &- \frac{\tepsilon
+  h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}\frac{\Omega^2}{\kappa^2}
+\frac{\p\ln{\tepsilon}}{\p\ln{r}}\left( \frac{\p\ln{\rhog}}{\p\ln{r}}\frac{\p\ln{\tepsilon}}{\p
+    z} -\frac{\p\ln{\rhog}}{\p z}\frac{\p\ln{\tepsilon}}{\p\ln{
+    r}}\right) > 0,
+%  \kappa^2\left|\frac{\p\epsilon}{\p z}\right| + c_s^2\left|\frac{\p
+%    \epsilon}{\p r}\frac{\p\epsilon}{\p z}\frac{\p\ln{\rho}}{\p
+%    r}\right| - c_s^2 \left(\frac{\p\epsilon}{\p
+%    r}\right)^2\left|\frac{\p\ln{\rho}}{\p z}\right| > 0.  \label{criterion2}
 \end{align}
+as the stability criteria, where 
+\begin{align}
+  h_\mathrm{g}\equiv c_s/\left(r\Omega\right)
+\end{align}
+is the characteristic aspect-ratio of the gas component.
 
-This equation shows that the vertical gradient in dust-fraction (and
-thus the dust-to-gas ratio $\rhod/\rhog$) has a stabilizing effect for
-axisymmetric perturbations. Instead, the only way to destabilize the
-system is a sufficiently large radial gradient in the dust-fraction
-and/or vertical (total) density gradient (the last term).
+%This equation shows that the vertical gradient in dust-fraction (and
+%thus the dust-to-gas ratio $\rhod/\rhog$) has a stabilizing effect for
+%axisymmetric perturbations. Instead, the only way to destabilize the
+%system is a sufficiently large radial gradient in the dust-fraction
+%and/or vertical (total) density gradient (the last term).
 
-Note that Eq. \ref{criterion2} is only applicable for the assume disk
-structure described above: decreasing dust and gas densities, as well
-as the dust-fraction, away from the star and away from the
-midplane. For less typical disk structures, one must evalulate above the
-critera seperately.
+%Note that Eq. \ref{criterion2} is only applicable for the assume disk
+%structure described above: decreasing dust and gas densities, as well
+%as the dust-fraction, away from the star and away from the
+%midplane. For less typical disk structures, one must evalulate above the
+%critera seperately.
+
+\begin{align} 
+  &1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1+
+    \tepsilon\right)^2}\frac{\p\ln{\rhog}}{\p\ln{r}}
+  \frac{\p\ln{\tepsilon}}{\p\ln{r}} +
+  \tepsilon\frac{z^2}{\widetilde{H}^2}>0,   \\ 
+&1 - \frac{\tepsilon
+  h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}
+  \frac{\p\ln{\tepsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\widetilde{H}^2}{H_\mathrm{g}^2}\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right) > 0,
+\end{align}
 

### 2016-04-23 model.tex
backup

--- 2016-04-22 model.tex
@@
   -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
-  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0. 
+  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, 
 \end{align}
-These criteria are applicable to any adiabatic fluid when $s$
-is the usual definition of entropy. For the dusty problem, we may take 
+with $s = \ln{P/\rho^\gamma}$. These criteria are applicable to any adiabatic fluid. 
+For the present dusty problem, we may take 
 $s=\ln{(1-\epsilon)}+\mathrm{const.}$ from the definition of $s$ and  
@@
 \begin{align}
-  \rho_\mathrm{g,d} = \rho_\mathrm{g0,d0} \exp
-      {\left(-\frac{z^2}{2H^2_\mathrm{g,d}}\right)}.
+  \rho_\mathrm{g,d}(r,z) = \rho_\mathrm{g0,d0}(r) \exp
+      {\left[-\frac{z^2}{2H^2_\mathrm{g,d}(r)}\right]}.
 \end{align}
@@
 \begin{align}
-  \tepsilon = \tepsilon_0
-  \exp{\left(-\frac{z^2}{2\widetilde{H}^2}\right)}, 
+  \tepsilon = \tepsilon_0(r)
+  \exp{\left[-\frac{z^2}{2\widetilde{H}^2(r)}\right]}, 
 \end{align}
@@
 \end{align}
-We assume the dust layer is much thinner than the gas scale-height,
-$H_\mathrm{d}\ll H_\mathrm{g}$, so that 
-\begin{align}
- \widetilde{H} \sim H_\mathrm{d}. 
-\end{align}
+%We assume the dust layer is much thinner than the gas scale-height,
+%$H_\mathrm{d}\ll H_\mathrm{g}$, so that 
+%\begin{align}
+% \widetilde{H} \sim H_\mathrm{d}. 
+%\end{align}
 
-We further assume the gas and dust
-distributions are smoothly varying in the radial direction so that  
-\begin{align}
-  \frac{\p}{\p r}\sim \frac{1}{r}, 
-\end{align}
+We further assume
+%the gas and dust
+%distributions are smoothly varying in the radial direction so that  
+%\begin{align}
+%  \frac{\p}{\p r}\sim \frac{p}{r}, 
+%\end{align}
 the disk is roughly Keplerian so that
@@
 
-Inserting the above approximations into Eq. \ref{criterion2}, and
-evaluating it at $z = H_\mathrm{d}$, we find 
-\begin{align}
-1 + \frac{h_\mathrm{g}^2\tepsilon}{\left(1 + \tepsilon\right)^2}
-\left[1 - \left(\frac{H_\mathrm{d}}{H_\mathrm{g}}\right)^2 +
-  \frac{\tepsilon}{1+\tepsilon}\left(1 -
-  \frac{H_\mathrm{d}}{H_\mathrm{g}}\right)   
-\right]>0.
+Inserting the above approximations into Eq. \ref{criterion2} and writing the result in terms of the
+dust-to-gas ratio $\tepsilon$,           
+we find 
+\begin{align}\label{ineq_est}
+  1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1 + \tepsilon\right)^2}
+  \left|\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right| 
+  \left[
+    \left|\frac{\p\ln{\rhog}}{\p\ln{r}}\right| - \frac{\widetilde{H}^2}{H^2_\mathrm{g}}\left|\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right|  
+    \right]>0
 \end{align}
+is required \emph{for stability.} 
+This inequality is easily satisifed in typical disk models where 
+the dust layer is much thinner than the gas scale-height, so
+that $\widetilde{H}\sim H_\mathrm{d}\ll H_\mathrm{g}$, and where the radial profiles smoothly vary on a scale of $r$.
+Then the magnitude the only destabilizing term (the third term) is negligible.
+According to Eq.
+\ref{ineq_est}, instability is only possible if there is a rapid variation the radial profile of the dust-to-gas ratio.
 

### 2016-04-23 model.tex
update

--- 2016-04-23 model.tex
@@
 \end{align}
+with $s = \ln{P/\rho^\gamma}$. These criteria are applicable to any adiabatic fluid. 
+For the present dusty problem, we may take 
 where $\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)$ is the
@@
 
-\begin{align} 
-  &1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1+
-    \tepsilon\right)^2}\frac{\p\ln{\rhog}}{\p\ln{r}}
-  \frac{\p\ln{\tepsilon}}{\p\ln{r}} +
-  \tepsilon\frac{z^2}{\widetilde{H}^2}>0,   \\ 
-&1 - \frac{\tepsilon
-  h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}
-  \frac{\p\ln{\tepsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\widetilde{H}^2}{H_\mathrm{g}^2}\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right) > 0,
-\end{align}
-
 \subsubsection{Order-of-magnitude estimates}
@@
 
-Inserting the above approximations into Eq. \ref{criterion2}, and
-evaluating it at $z = H_\mathrm{d}$, we find 
+%Inserting the above approximations into Eq. \ref{criterion2} and writing the result in terms of the
+%dust-to-gas ratio $\tepsilon$,           
+%we find 
+%\begin{align}\label{ineq_est}
+%  1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1 + \tepsilon\right)^2}
+%  \left|\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right| 
+%  \left[
+%    \left|\frac{\p\ln{\rhog}}{\p\ln{r}}\right| - \frac{\widetilde{H}^2}{H^2_\mathrm{g}}\left|\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right|  
+%    \right]>0
+%\end{align}
+%is required \emph{for stability.} 
+%This inequality is easily satisifed in typical disk models where 
+%the dust layer is much thinner than the gas scale-height, so
+%that $\widetilde{H}\sim H_\mathrm{d}\ll H_\mathrm{g}$, and where the radial profiles smoothly vary on a scale of $r$.
+%Then the magnitude the only destabilizing term (the third term) is negligible.
+According to Eq.
+\ref{ineq_est}, instability is on%ly possible if there is a rapid variation the radial profile of the dust-to-gas ratio.
+
 \begin{align} 
-1 + \frac{h_\mathrm{g}^2\tepsilon}{\left(1 + \tepsilon\right)^2}
-\left[1 - \left(\frac{H_\mathrm{d}}{H_\mathrm{g}}\right)^2 +
-  \frac{\tepsilon}{1+\tepsilon}\left(1 -
-  \frac{H_\mathrm{d}}{H_\mathrm{g}}\right)   
-\right]>0.
+  &1 + \frac{\tepsilon}{\left(1+
+    \tepsilon\right)^2}\left(h_\mathrm{g}^2\frac{\p\ln{\rhog}}{\p\ln{r}}
+  \frac{\p\ln{\tepsilon}}{\p\ln{r}} + 
+  \frac{z^2}{\widetilde{H}^2}\right)>0,   \\ 
+&1 - \frac{\tepsilon
+  h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}
+  \frac{\p\ln{\tepsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\widetilde{H}^2}{H_\mathrm{g}^2}\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right) > 0,
 \end{align}

### 2016-04-23 model.tex
update

@@
 \end{align}
-may be consider as a pseudo-entropy for the mixture.
+is a pseudo-entropy for the mixture.
 
@@
 \subsubsection{Order-of-magnitude estimates}
-Let us assume the gas and dust distributions have a Gaussian
+We evaluate the above stability criteria by assuming the gas and dust
+distributions have a Gaussian 
 distribution vertically, 
@@
 \end{align}
-Then the dust-to-gas ratio is 
+We note that $H_\mathrm{g}=h_gr$ by definition. The dust-to-gas ratio
+is  
 \begin{align}
@@
 \end{align}
-%We assume the dust layer is much thinner than the gas scale-height,
-%$H_\mathrm{d}\ll H_\mathrm{g}$, so that 
-%\begin{align}
-% \widetilde{H} \sim H_\mathrm{d}. 
-%\end{align}
-
-We further assume
-%the gas and dust
-%distributions are smoothly varying in the radial direction so that  
-%\begin{align}
-%  \frac{\p}{\p r}\sim \frac{p}{r}, 
-%\end{align}
-the disk is roughly Keplerian so that
+We further assume the disk is approximately Keplerian so that
 \begin{align}
-  \kappa \sim \OmK, 
+  \kappa \simeq \Omega. 
 \end{align}
-and the gas sound-speed is 
-\begin{align}
-  c_s \sim h_\mathrm{g}r\OmK, 
-\end{align}
-where $h_\mathrm{g}\ll1$ is the aspect-ratio of the gas component.
-
 %Inserting the above approximations into Eq. \ref{criterion2} and writing the result in terms of the
@@
 \ref{ineq_est}, instability is on%ly possible if there is a rapid variation the radial profile of the dust-to-gas ratio.
-
+Then the above criteria becomes
 \begin{align}
@@
   \frac{\p\ln{\tepsilon}}{\p\ln{r}} + 
-  \frac{z^2}{\widetilde{H}^2}\right)>0,   \\ 
+  \frac{z^2}{\widetilde{H}^2}\right)>0 \label{stability_est1},   \\ 
 &1 - \frac{\tepsilon
   h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}
-  \frac{\p\ln{\tepsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\widetilde{H}^2}{H_\mathrm{g}^2}\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right) > 0,
+  \frac{\p\ln{\tepsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\widetilde{H}^2}{H_\mathrm{g}^2}\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right)
+  > 0 \label{stability_est2},
 \end{align}
+\emph{for stability}.
+
+Eq. \ref{stability_est1}---\ref{stability_est2} are easily satisfied
+in typical accretion disk models where radial gradients are $O(1/r)$ and
+$h_\mathrm{g},\tepsilon, \widetilde{H}/H_\mathrm{g}\ll 1$, since then the
+magnitude of the second term on the left-hand-side  of either
+inequality is much less than unity. Such disks are stable against
+axisymmetric perturbations.
+
+\subsubsection{Possible instability at dust edges}
+Notice the left-hand-side of Eq. \ref{stability_est2} is a quadratic in
+$\p_r\tepsilon$. It is possible to
+violate this inequality for sufficiently large (in magnitude) radial
+gradients  in the dust-to-gas ratio, 
+\begin{align}
+  \frac{\p\ln{\tepsilon}}{\p\ln{r}} > S_+ \quad \mathrm{or} \quad 
+  \frac{\p\ln{\tepsilon}}{\p\ln{r}} < S_-,
+\end{align}
+\emph{for instability}, where
+\begin{align}\label{spm}
+S_\pm = \frac{1}{2}\frac{H_\mathrm{g}^2}{\widetilde{H}^2} 
+  \left[
+  \frac{\p\ln{\rhog}}{\p\ln{r}} \pm 
+  \sqrt{
+  \left(\frac{\p\ln{\rhog}}{\p\ln{r}}\right)^2 + 
+  4 \frac{\widetilde{H}^2}{H_\mathrm{g}^2}
+  \frac{\left(1+\tepsilon\right)^2}{\tepsilon h_\mathrm{g}^2}
+  }
+  \,\right]. 
+\end{align}
+
+Now, if $\p_r\rhog<0$, as typical in
+accretion models, then it is easier to achieve instability at a given
+radius for rapidly increasing dust-to-gas ratios ($\p_r\tepsilon >
+0$), and vice versa. 
+%This might occur, for example, when dust is
+%inserted in a finite radial range within a wider gaseous disk. 
+%Physically, one might expect
+%$\p_r\rhog\p_r\tepsilon > 0$ because
+
+\emph{If} the $\p_r\rhog$ can be neglected in Eq. \ref{spm},
+then instability requires
+\begin{align}
+  \left|\frac{\p\ln{\tepsilon}}{\p r}\right| \gtrsim
+  \frac{1}{\widetilde{H}}\frac{\left(1+\tepsilon\right)}{\sqrt{\tepsilon}}. 
+\end{align}
+That is, if the dust-to-gas ratio varies on a lengthscale less than
+$O(\widetilde{H}\sqrt{\tepsilon})$ for $\tepsilon\ll 1$, or less than
+$O(\widetilde{H}/\sqrt{\tepsilon})$ for $\tepsilon\gg 1$, then the system is
+potentially unstable.
 

### 2016-04-23 model.tex
typo

@@
 \end{align}
-with $s = \ln{P/\rho^\gamma}$. These criteria are applicable to any adiabatic fluid. 
-For the present dusty problem, we may take 
-where $\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)$ is the
+with $s = \ln{P/\rho^\gamma}$ and 
+$\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)$ is the
 square of the epicylic frequency. We assume $\kappa^2>0$. 

### 2016-04-30 appendix.tex
new appendices

@@
   \kappa^2 |\dd v_r|^2 + \rho r\frac{\p \Omega^2}{\p z} \dd v_z \dd
-  v_r^*  + \ii \rho \dd \bm{F}\cdot\dd\bm{v}^*
+  v_r^*  + \ii\sigma \rho \dd \bm{F}\cdot\dd\bm{v}^*
   \right)dV,
@@
 %profile, the second
+
+\section{Explicit expressions for pressure forces and its divergence}
+Define
+\begin{align}
+  \bm{F} \equiv - \frac{\nabla P}{\rho}. 
+\end{align}
+For a locally isothermal equation of state, Eq. \ref{eos}, this
+becomes
+\begin{align}
+  \bm{F} = c_s^2\nabla \epsilon - c_s^2(1-\epsilon)\nabla\ln{\rho} -
+  (1-\epsilon)\nabla c_s^2,
+\end{align} 
+and its linearized form is
+\begin{align}
+  \dd\bm{F} = c_s^2\nabla\dd\epsilon + c_s^2\dd\epsilon\nabla\ln{\rho}
+  - c_s^2(1-\epsilon)\nabla\left(\frac{\dd\rho}{\rho}\right) +
+  \dd\epsilon\nabla c_s^2. 
+\end{align}
+The divergence of the above expressions are
+\begin{align}
+  \nabla\cdot\bm{F} &= 2 \nabla c_s^2\cdot\nabla\epsilon +
+  c_s^2\nabla^2\epsilon  - \left[(1-\epsilon)\nabla c_s^2 -
+    c_s^2\nabla\epsilon\right]\cdot\nabla\ln{\rho}
+ - c_s^2(1-\epsilon)\nabla^2\ln{\rho} - (1-\epsilon)\nabla^2 c_s^2,\\
+ \nabla\cdot\dd\bm{F} &= 2\nabla c_s^2\cdot\nabla\dd\epsilon +
+ c_s^2\nabla^2\dd\epsilon + \left(\dd\epsilon \nabla c_s^2 +
+ c_s^2\nabla\dd\epsilon\right)\cdot\nabla\ln{\rho} + c_s^2 \dd\epsilon
+ \nabla^2\ln{\rho}   + \dd\epsilon\nabla^2 c_s^2 \notag\\
+&\phantom{=}-c_s^2(1-\epsilon)\nabla^2\left(\frac{\dd\rho}{\rho}\right) - \left[(1-\epsilon)\nabla c_s^2 -
+    c_s^2\nabla\epsilon\right]\cdot\nabla \left(\frac{\dd\rho}{\rho}\right).
+\end{align}
+For a power-law prescription for the radial temperature
+profile $c_s^2\propto r^q$, we have
+\begin{align}
+  \nabla c_s^2 &= \frac{qc_s^2}{r}\hat{\bm{r}},\\
+  \nabla^2 c_s^2 & = \left(\frac{q}{r}\right)^2 c_s^2. 
+\end{align}
+
+\section{Linearized dust evolution equation}
+We consider small grains in the Epstein regime, with fixed internal
+density and size, so that
+\begin{align}
+  \tstop \propto \frac{1}{\rho c_s}
+\end{align}
+\citep{price15}. 
+Then the evolutionary equation for the dust-fraction $\epsilon$ is 
+\begin{align}\label{dust_epstein}
+  \frac{D\epsilon}{Dt} = \tstop c_s \nabla \cdot
+  \left(\frac{\epsilon}{c_s}\bm{F}\right). 
+\end{align} 
+Linearizing Eq. \ref{dust_epstein},
+\begin{align}
+  s \dd\epsilon + \dd\bm{v}\cdot\nabla\epsilon &= \tstop
+  \left(\dd\epsilon\nabla\cdot\bm{F} + \epsilon\nabla\cdot\dd\bm{F} +  
+  \nabla\dd\epsilon\cdot\bm{F} + \nabla \epsilon\cdot\dd\bm{F} -
+  \frac{1}{2}\dd\epsilon\bm{F}\cdot\nabla\ln{c_s^2} -
+  \frac{1}{2}\epsilon\dd\bm{F}\cdot\nabla\ln{c_s^2} \right)
+  \notag\\
+  &\phantom{=} - \tstop \left(\epsilon\nabla\cdot\bm{F} +
+  \nabla\epsilon\cdot\bm{F}-\frac{1}{2}\epsilon\bm{F}\cdot\nabla\ln{c_s^2}\right)\frac{\dd\rho}{\rho}.
+\end{align}
+
+\section{Conversion formulae}
+The gas density $\rhog$ and dust-to-gas ratio $\tepsilon$ is related
+to the total density $\rho$ and dust-fraction $\epsilon$ by
+\begin{align}
+  \ln{\rho} &= \ln{\rhog} + \ln{\left(1 + \tepsilon\right)},\\
+  \nabla\ln{\rho} &= \nabla\ln{\rhog} + \frac{\nabla\tepsilon}{1+
+    \tepsilon},\\
+  \nabla^2\ln{\rho} &= \nabla^2\ln{\rhog} +
+  \frac{\nabla^2\tepsilon}{1+\tepsilon} -
+  \frac{\left|\nabla\tepsilon\right|^2}{\left(1+\tepsilon\right)^2}, 
+\end{align}
+and
+\begin{align}
+  \epsilon &= \frac{\tepsilon}{1+ \tepsilon},\\
+ \nabla\epsilon & =
+ \frac{\nabla\tepsilon}{\left(1+\tepsilon\right)^2},\\
+ \nabla^2\epsilon &=
+ \frac{\nabla^2\tepsilon}{\left(1+\tepsilon\right)^2} -
+ \frac{2\left|\nabla\tepsilon\right|^2}{\left(1+\tepsilon\right)^3}. 
+\end{align}

### 2016-05-03 code/fixed/basic.f90
code update

@@ -1,0 +1,325 @@
+real*8 function lnrhog(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+
+  lnrhog = -0.5d0*zhat**2d0 - dgratio*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) 
+end function lnrhog
+
+real*8 function dlnrhog_dr(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+ 
+  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) )
+end function dlnrhog_dr
+
+real*8 function d2lnrhog_dr2(zhat)
+  !this is (1/r)*(d/dr)( r*dlnrhog_dr )
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  
+  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) 
+end function d2lnrhog_dr2
+
+real*8 function dlnrhog_dz(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  
+  dlnrhog_dz = -zhat*(1d0 + dgratio*exp( - zhat**2d0/2d0/delta**2d0 ) )
+end function dlnrhog_dz
+
+real*8 function d2lnrhog_dz2(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  
+  d2lnrhog_dz2 = -1d0 - dgratio*exp(-zhat**2d0/2d0/delta**2d0)*(1d0 - zhat**2d0/delta**2d0)
+end function d2lnrhog_dz2
+
+real*8 function del2_rhog(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  real*8, external  :: d2lnrhog_dr2, d2lnrhog_dz2
+
+  del2_rhog = d2lnrhog_dr2(zhat) + d2lnrhog_dz2(zhat)
+end function del2_rhog
+
[…]

### 2016-05-03 code/fixed/dvsi.f90

@@ -1,0 +1,497 @@
+module global
+  implicit none 
+  integer, parameter :: nvar = 5 
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  complex*16, parameter :: ii = (0d0, 1d0), zero = (0d0, 0d0)
+
+  integer :: nz, bignz, Nzmid
+  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd
+  real*8 :: zmax 
+  real*8 :: dlnHg_dlnr
+  real*8 :: krad, bc_tol 
+
+  real*8, allocatable :: zaxis(:), lnrho_arr(:), eps_arr(:), dlnrhodr_arr(:), depsdr_arr(:),dlnrhodz_arr(:), depsdz_arr(:) 
+  real*8, allocatable :: del2lnrho_arr(:), del2eps_arr(:), dellnrho_dot_deleps_arr(:), tstop_arr(:)
+  real*8, allocatable :: Fr_arr(:), Fz_arr(:), F_dot_deleps_arr(:), divF_arr(:), F_dot_dellncs2_arr(:)
+  real*8, allocatable :: omega_arr(:), kappa2_arr(:), vshear_arr(:) 
+
+  complex*16, allocatable     :: T(:,:), Tp(:,:), Tpp(:,:)
+
+  complex*16, allocatable :: L11(:,:), L12(:,:), L13(:,:), L14(:,:), L15(:,:)
+  complex*16, allocatable :: L21(:,:), L22(:,:), L23(:,:), L24(:,:), L25(:,:)
+  complex*16, allocatable :: L31(:,:), L32(:,:), L33(:,:), L34(:,:), L35(:,:)
+  complex*16, allocatable :: L41(:,:), L42(:,:), L43(:,:), L44(:,:), L45(:,:)
+  complex*16, allocatable :: L51(:,:), L52(:,:), L53(:,:), L54(:,:), L55(:,:)
+  complex*16, allocatable :: R11(:,:), R12(:,:), R13(:,:), R14(:,:), R15(:,:) 
+  complex*16, allocatable :: R21(:,:), R22(:,:), R23(:,:), R24(:,:), R25(:,:)
+  complex*16, allocatable :: R31(:,:), R32(:,:), R33(:,:), R34(:,:), R35(:,:)
[…]

### 2016-05-03 code/iso/basic.f90

@@ -1,0 +1,314 @@
+real*8 function lnrhog(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+
+  lnrhog = -0.5d0*zhat**2d0 - dgratio*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) 
+end function lnrhog
+
+real*8 function dlnrhog_dr(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+ 
+  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) )
+end function dlnrhog_dr
+
+real*8 function d2lnrhog_dr2(zhat)
+  !this is (1/r)*(d/dr)( r*dlnrhog_dr )
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  
+  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) 
+end function d2lnrhog_dr2
+
+real*8 function dlnrhog_dz(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  
+  dlnrhog_dz = -zhat*(1d0 + dgratio*exp( - zhat**2d0/2d0/delta**2d0 ) )
+end function dlnrhog_dz
+
+real*8 function d2lnrhog_dz2(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  
+  d2lnrhog_dz2 = -1d0 - dgratio*exp(-zhat**2d0/2d0/delta**2d0)*(1d0 - zhat**2d0/delta**2d0)
+end function d2lnrhog_dz2
+
+real*8 function del2_rhog(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  real*8, external  :: d2lnrhog_dr2, d2lnrhog_dz2
+
+  del2_rhog = d2lnrhog_dr2(zhat) + d2lnrhog_dz2(zhat)
+end function del2_rhog
+
[…]

### 2016-05-03 code/iso/dvsi.f90

@@ -1,0 +1,423 @@
+module global
+  implicit none 
+  integer, parameter :: nvar = 5 
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  complex*16, parameter :: ii = (0d0, 1d0), zero = (0d0, 0d0)
+
+  integer :: nz, bignz, Nzmid
+  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop
+  real*8 :: zmax 
+  real*8 :: dlnHg_dlnr
+
+  real*8, allocatable :: zaxis(:), lnrho_arr(:), eps_arr(:), dlnrhodr_arr(:), depsdr_arr(:),dlnrhodz_arr(:), depsdz_arr(:) 
+  real*8, allocatable :: del2lnrho_arr(:), del2eps_arr(:), dellnrho_dot_deleps_arr(:), tstop_arr(:)
+  real*8, allocatable :: Fr_arr(:), Fz_arr(:), F_dot_deleps_arr(:), divF_arr(:)
+  real*8, allocatable :: omega_arr(:), kappa2_arr(:), vshear_arr(:) 
+
+  real*8, allocatable     :: T(:,:), Tp(:,:), Tpp(:,:)
+
+  complex*16, allocatable :: L11(:,:), L12(:,:), L13(:,:), L14(:,:), L15(:,:)
+  complex*16, allocatable :: L21(:,:), L22(:,:), L23(:,:), L24(:,:), L25(:,:)
+  complex*16, allocatable :: L31(:,:), L32(:,:), L33(:,:), L34(:,:), L35(:,:)
+  complex*16, allocatable :: L41(:,:), L42(:,:), L43(:,:), L44(:,:), L45(:,:)
+  complex*16, allocatable :: L51(:,:), L52(:,:), L53(:,:), L54(:,:), L55(:,:)
+  complex*16, allocatable :: R11(:,:), R12(:,:), R13(:,:), R14(:,:), R15(:,:) 
+  complex*16, allocatable :: R21(:,:), R22(:,:), R23(:,:), R24(:,:), R25(:,:)
+  complex*16, allocatable :: R31(:,:), R32(:,:), R33(:,:), R34(:,:), R35(:,:)
[…]

### 2016-05-03 code/press/basic.f90

@@ -1,0 +1,355 @@
+real*8 function lnrhog(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+
+  lnrhog = -0.5d0*zhat**2d0 - dgratio*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) 
+end function lnrhog
+
+real*8 function dlnrhog_dr(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+ 
+  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) )
+end function dlnrhog_dr
+
+real*8 function d2lnrhog_dr2(zhat)
+  !this is (1/r)*(d/dr)( r*dlnrhog_dr )
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  
+  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) 
+end function d2lnrhog_dr2
+
+real*8 function dlnrhog_dz(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  
+  dlnrhog_dz = -zhat*(1d0 + dgratio*exp( - zhat**2d0/2d0/delta**2d0 ) )
+end function dlnrhog_dz
+
+real*8 function d2lnrhog_dz2(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  
+  d2lnrhog_dz2 = -1d0 - dgratio*exp(-zhat**2d0/2d0/delta**2d0)*(1d0 - zhat**2d0/delta**2d0)
+end function d2lnrhog_dz2
+
+real*8 function del2_rhog(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  real*8, external  :: d2lnrhog_dr2, d2lnrhog_dz2
+
+  del2_rhog = d2lnrhog_dr2(zhat) + d2lnrhog_dz2(zhat)
+end function del2_rhog
+
[…]

### 2016-05-03 code/press/dvsi.f90

@@ -1,0 +1,480 @@
+module global
+  implicit none 
+  integer, parameter :: nvar = 5 
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  complex*16, parameter :: ii = (0d0, 1d0), zero = (0d0, 0d0)
+
+  integer :: nz, bignz, Nzmid
+  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd
+  real*8 :: zmax 
+  real*8 :: dlnHg_dlnr
+  real*8 :: krad, bc_tol 
+
+  real*8, allocatable :: zaxis(:), lnrho_arr(:), eps_arr(:), dlnrhodr_arr(:), depsdr_arr(:),dlnrhodz_arr(:), depsdz_arr(:) 
+  real*8, allocatable :: del2lnrho_arr(:), del2eps_arr(:), dellnrho_dot_deleps_arr(:), tstop_arr(:), d2lnrhodz2_arr(:)
+  real*8, allocatable :: Fr_arr(:), Fz_arr(:), F_dot_deleps_arr(:), divF_arr(:), F_dot_dellncs2_arr(:)
+  real*8, allocatable :: omega_arr(:), kappa2_arr(:), vshear_arr(:) 
+
+  complex*16, allocatable     :: T(:,:), Tp(:,:), Tpp(:,:)
+
+  complex*16, allocatable :: L11(:,:), L12(:,:), L13(:,:), L14(:,:), L15(:,:)
+  complex*16, allocatable :: L21(:,:), L22(:,:), L23(:,:), L24(:,:), L25(:,:)
+  complex*16, allocatable :: L31(:,:), L32(:,:), L33(:,:), L34(:,:), L35(:,:)
+  complex*16, allocatable :: L41(:,:), L42(:,:), L43(:,:), L44(:,:), L45(:,:)
+  complex*16, allocatable :: L51(:,:), L52(:,:), L53(:,:), L54(:,:), L55(:,:)
+  complex*16, allocatable :: R11(:,:), R12(:,:), R13(:,:), R14(:,:), R15(:,:) 
+  complex*16, allocatable :: R21(:,:), R22(:,:), R23(:,:), R24(:,:), R25(:,:)
+  complex*16, allocatable :: R31(:,:), R32(:,:), R33(:,:), R34(:,:), R35(:,:)
[…]

### 2016-05-03 code/test/dvsi.f90

@@ -1,0 +1,394 @@
+module global
+  implicit none 
+  integer, parameter :: nvar = 4 
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  complex*16, parameter :: ii = (0d0, 1d0), zero = (0d0, 0d0)
+
+  integer :: nz, bignz, Nzmid
+  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd
+  real*8 :: zmax 
+  real*8 :: dlnHg_dlnr
+  real*8 :: krad, bc_tol 
+
+  real*8, allocatable :: zaxis(:), lnrho_arr(:), eps_arr(:), dlnrhodr_arr(:), depsdr_arr(:),dlnrhodz_arr(:), depsdz_arr(:) 
+  real*8, allocatable :: del2lnrho_arr(:), del2eps_arr(:), dellnrho_dot_deleps_arr(:), tstop_arr(:)
+  real*8, allocatable :: Fr_arr(:), Fz_arr(:), F_dot_deleps_arr(:), divF_arr(:), F_dot_dellncs2_arr(:)
+  real*8, allocatable :: omega_arr(:), kappa2_arr(:), vshear_arr(:) 
+
+  complex*16, allocatable     :: T(:,:), Tp(:,:), Tpp(:,:)
+
+  complex*16, allocatable :: L11(:,:), L12(:,:), L13(:,:), L14(:,:), L15(:,:)
+  complex*16, allocatable :: L21(:,:), L22(:,:), L23(:,:), L24(:,:), L25(:,:)
+  complex*16, allocatable :: L31(:,:), L32(:,:), L33(:,:), L34(:,:), L35(:,:)
+  complex*16, allocatable :: L41(:,:), L42(:,:), L43(:,:), L44(:,:), L45(:,:)
+  complex*16, allocatable :: L51(:,:), L52(:,:), L53(:,:), L54(:,:), L55(:,:)
+  complex*16, allocatable :: R11(:,:), R12(:,:), R13(:,:), R14(:,:), R15(:,:) 
+  complex*16, allocatable :: R21(:,:), R22(:,:), R23(:,:), R24(:,:), R25(:,:)
+  complex*16, allocatable :: R31(:,:), R32(:,:), R33(:,:), R34(:,:), R35(:,:)
[…]

### 2016-05-04 result.tex
demo fig

\section{Results}

\begin{figure}
\includegraphics[width=\linewidth]{figures/eigenvalues}
\caption{do not believe}
\end{figure}

### 2016-05-06 appendix.tex
updating appendix:

--- 2016-04-21 appendix.tex
@@
-\section{Variational principle for locally isothermal gas perfectly
-  coupled to dust}\label{var_prin}
+\section{Variational principle for locally isothermal gas
+  well-coupled to dust}\label{var_prin}
 
@@
   \ii\sigma\frac{\delta P}{P} &= \nabla\cdot\dd\bm{v} +
-  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{c_s^2},\\
+  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P},\\
   -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
@@
   \delta \bm{F} \equiv \frac{\dd\rho}{\rho^2}\nabla P -
-  \frac{1}{\rho}\nabla\dd P.
+  \frac{1}{\rho}\nabla\dd P, 
 \end{align}
+and $\dd\mathcal{C}$ is the linearized dust-duffusion function (see Eq. \ref{dusty_cooling}).  
 Note that for the axisymmetric problem, $\dd\bm{F}$ is purely
 meridional.
+
 From the linearized meridional momentum equations, we find 
@@
    \ii\sigma \frac{\dd P}{P} - \dd\bm{v}\cdot\nabla s + \dd
-   \bm{v} \cdot \nabla \ln{c_s^2}
+   \bm{v} \cdot \nabla \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}
    \right)\dd\bm{v}^*\cdot\nabla P  + \ii\sigma \dd
@@
    \left(\dd\bm{v}^*\cdot\nabla P\right) \left(\dd\bm{v} \cdot \nabla
-   \ln{c_s^2}   -  \dd\bm{v}\cdot\nabla s\right) 
+   \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}   -  \dd\bm{v}\cdot\nabla s\right) 
    \right]dV\notag\\
@@
    \left[\frac{1}{P}\nabla\cdot\left(P\dd\bm{v}\right) -
-   \dd\bm{v}\cdot\nabla\ln{c_s^2}\right]\nabla\cdot\left(P\dd\bm{v}^*\right)
+   \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P}  \right]\nabla\cdot\left(P\dd\bm{v}^*\right)
    +\left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v} \cdot \nabla
-   \ln{c_s^2}   -  \dd\bm{v}\cdot\nabla s \right)
+   \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}   -  \dd\bm{v}\cdot\nabla s \right)
    \right\}dV \notag\\
@@
      \right) -
-     P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right) 
+     P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2} + \frac{\dd\mathcal{C}}{P} \right)  
      \right]dV,
@@
   \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2
-  \right\}dV -\int P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV.
+  \right\}dV -\int
+  P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV
+  -\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV.  
 \end{align}
@@
 %profile, the second
+
+\section{Linearized pressure forces and its divergence}
+Define
+\begin{align}
+  \bm{F} \equiv - \frac{\nabla P}{\rho}. 
+\end{align}
+Then its linearized form is 
+\begin{align}
+  \delta \bm{F} = - \frac{\dd\rho}{\rho}\bm{F} -
+  \frac{1}{\rho}\nabla\dd P,
+\end{align}
+with divergence
+\begin{align}
+  \nabla\cdot\dd\bm{F} = -
+  \nabla\left(\frac{\dd\rho}{\rho}\right)\cdot\bm{F} -
+  \frac{\dd\rho}{\rho}\nabla\cdot\bm{F} +
+  \nabla\ln{\rho}\cdot\frac{\nabla\dd P}{\rho} -
+  \frac{1}{\rho}\nabla^2\dd P. 
+\end{align}
+
+\subsection{Radially local approximation}
+We assume perturbations have small radial wavelengths compared to the
+local disk radius. We take a wave-like dependence $\sim \exp{\ii
+  k r}$ for the pertubations with $|kr|\gg 1$ so that curvature terms 
+may be neglected. We thus set $\p_r\to \ii k$ when, \emph{and only
+  when}, operating on the \emph{primitive} perturbations $\dd P,
+\dd\rho$ and $\dd\bm{v}$. %In the following expressions, we only
+%apply the radially local approximation at the last step.
+
+\subsection{Explicit expressions}
+Introducing
+\begin{align}
+  W \equiv \frac{\dd\rho}{\rho}, \quad Q \equiv \frac{\dd P}{\rho},
+\end{align}
+the explicit forms for $\dd\bm{F}$ and $\nabla\cdot\dd\bm{F}$, in the
+radially-local approximation, are
+\begin{align}
+  &\dd F_r = - W F_r - \ii k Q,\\
+  &\dd F_z = - W F_z - \left[Q^\prime + Q
+    \left(\ln{\rho}\right)^\prime\right]  = s\dd v_z, 
+\end{align} 
+where $^\prime = \p_z$ and the last equality is the linearized
+vertical momentum equation; and 
+\begin{align}
+  \nabla\cdot\dd\bm{F} &= F_r\left(\p_r\ln{\rho} - \ii k \right)W - F_z
+  W^\prime - W \nabla\cdot\bm{F} + \ii kQ \p_r\ln{\rho} +
+  \left(\ln{\rho}\right)^\prime\left[Q^\prime + Q
+    \left(\ln{\rho}\right)^\prime\right] \notag\\
+  & \phantom{=} + k^2 Q - \left\{Q^{\prime\prime} + 2
+    Q^\prime\left(\ln{\rho}\right)^\prime
+    + Q\left[\left(\ln{\rho}\right)^{\prime\prime}+\left(\ln{\rho}\right)^{\prime
+      2}\right]\right\}\notag\\
+    &= \left[\left(\p_r\ln{\rho}-\ii k\right)F_r - \nabla\cdot\bm{F} +
+      F_z^\prime\right]W + \left(\ii k \p_r\ln{\rho} + k^2\right)Q +
+    s\dd v_z^\prime.
+\end{align}
+
+\section{Linearized dust diffusion}
+We consider small grains in the Epstein regime, with fixed internal
+density and size, so that
+\begin{align}
+  \tstop  =  \frac{K}{\rho c_s}
+\end{align}
+\citep{price15}, where $K$ is a constant. Then the dust diffusion
+function becomes
+\begin{align}
+  \mathcal{C} \equiv c_s^2\nabla\cdot\left(\epsilon\tstop\nabla
+  P\right) = - K c_s^2 \nabla\cdot
+  \left(\frac{\epsilon}{c_s}\bm{F}\right) =
+  -Kc_s \left(\bm{F}\cdot\nabla\epsilon + \epsilon \nabla\cdot\bm{F}
+  - \frac{1}{2}\epsilon\bm{F}\cdot\nabla\ln{c_s^2}\right).  
+\end{align}
+Linearizing,
+\begin{align}
+  - \frac{\dd\mathcal{C}}{Kc_s} = \bm{F}\cdot\nabla\dd\epsilon +
+  \dd\bm{F}\cdot\nabla\epsilon + \dd\epsilon\nabla\cdot\bm{F} +
+  \epsilon\nabla\cdot\dd\bm{F} -
+  \frac{1}{2}\nabla\ln{c_s^2}\cdot\left(\bm{F}\dd\epsilon + \epsilon
+  \dd\bm{F}\right). 
+\end{align}
+The linearized dust-fraction $\dd\epsilon$ and its derivatives in the
+radially-local approximation are given by
+\begin{align}
+  \dd\epsilon      &= (1-\epsilon)W - \frac{Q}{c_s^2}, \\
+  \p_r\dd\epsilon &= \left[(1-\epsilon)\left(\ii k -
+    \p_r\ln{\rho}\right) - \p_r\epsilon\right]W + \left(\p_r\ln{c_s^2}
+  + \p_r\ln{\rho} - \ii k \right)\frac{Q}{c_s^2},\\
+  \dd\epsilon^\prime &= (1-\epsilon)W^\prime - \epsilon^\prime W -
+  \left(\frac{Q}{c_s^2}\right)^\prime. 
+\end{align}
+
+\section{Conversion formulae}
+The gas density $\rhog$ and dust-to-gas ratio $\tepsilon$ is related
+to the total density $\rho$ and dust-fraction $\epsilon$ by
+\begin{align}
+  \ln{\rho} &= \ln{\rhog} + \ln{\left(1 + \tepsilon\right)},\\
+  \nabla\ln{\rho} &= \nabla\ln{\rhog} + \frac{\nabla\tepsilon}{1+
+    \tepsilon},\\
+  \nabla^2\ln{\rho} &= \nabla^2\ln{\rhog} +
+  \frac{\nabla^2\tepsilon}{1+\tepsilon} -
+  \frac{\left|\nabla\tepsilon\right|^2}{\left(1+\tepsilon\right)^2}, 
+\end{align}
+and
+\begin{align}
+  \epsilon &= \frac{\tepsilon}{1+ \tepsilon},\\
+ \nabla\epsilon & =
+ \frac{\nabla\tepsilon}{\left(1+\tepsilon\right)^2},\\
+ \nabla^2\epsilon &=
+ \frac{\nabla^2\tepsilon}{\left(1+\tepsilon\right)^2} -
+ \frac{2\left|\nabla\tepsilon\right|^2}{\left(1+\tepsilon\right)^3}. 
+\end{align}

### 2016-05-06 appendix.tex
updated notes

@@
 
-Consider axisymmetric disturbances of the form
-\begin{align}
-  \delta\rho(r,z)\exp{\left(-\ii\sigma\right)}. 
-\end{align}
-Linearized equations for dusty gas with $\tstop=0$, and using the
-pressure equation in place of the dust-fraction
-(Eq. \ref{masseq}---\ref{momeq}, Eq. \ref{eff_energy}):
+The linearized equations for axisymmetric disturbances are
 
@@
   \ii\sigma\frac{\delta\rho}{\rho} &= \nabla\cdot\dd\bm{v} +
-  \dd\bm{v}\cdot\nabla\ln{\rho},\\
+  \dd\bm{v}\cdot\nabla\ln{\rho},\label{lin_mass_full}\\
  \ii\sigma\frac{\delta P}{P} &= \nabla\cdot\dd\bm{v} +
-  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P},\\
+  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
   -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
@@
     v_\phi}{\p z}\dd v_z,\\
-  -\ii\sigma\dd v_z &=  \delta\bm{F}\cdot\hat{\bm{z}},
+  -\ii\sigma\dd v_z &=  \delta\bm{F}\cdot\hat{\bm{z}}.
 \end{align}  
-with
-\begin{align}
-  \delta \bm{F} \equiv \frac{\dd\rho}{\rho^2}\nabla P -
-  \frac{1}{\rho}\nabla\dd P, 
-\end{align}
-and $\dd\mathcal{C}$ is the linearized dust-duffusion function (see Eq. \ref{dusty_cooling}).  
-Note that for the axisymmetric problem, $\dd\bm{F}$ is purely
-meridional.
+These equations do not assume the radially-local approximation used in
+the main text.
 
@@
 
-\section{Linearized pressure forces and its divergence}
+\section{Linearized pressure forces and its divergence}\label{lin_press}
 Define
@@
 
-\subsection{Radially local approximation}
-We assume perturbations have small radial wavelengths compared to the
-local disk radius. We take a wave-like dependence $\sim \exp{\ii
-  k r}$ for the pertubations with $|kr|\gg 1$ so that curvature terms 
-may be neglected. We thus set $\p_r\to \ii k$ when, \emph{and only
-  when}, operating on the \emph{primitive} perturbations $\dd P,
-\dd\rho$ and $\dd\bm{v}$. %In the following expressions, we only
-%apply the radially local approximation at the last step.
+%\subsection{Explicit expressions}
 
-\subsection{Explicit expressions}
-Introducing
-\begin{align}
-  W \equiv \frac{\dd\rho}{\rho}, \quad Q \equiv \frac{\dd P}{\rho},
-\end{align}
-the explicit forms for $\dd\bm{F}$ and $\nabla\cdot\dd\bm{F}$, in the
+The explicit forms for $\dd\bm{F}$ and $\nabla\cdot\dd\bm{F}$, in the
 radially-local approximation, are
@@
 \end{align} 
-where $^\prime = \p_z$ and the last equality is the linearized
+where the last equality is the linearized
 vertical momentum equation; and 
@@
 
-\section{Linearized dust diffusion}
+\section{Linearized dust diffusion}\label{lin_dust}
 We consider small grains in the Epstein regime, with fixed internal

### 2016-05-06 linear.tex
updated notes

\section{Linear problem} 
We consider radially-localized axisymmetric disturbances of the form 
\begin{align}
  \delta\rho(r, z)\exp{\left(\ii k r-\ii\sigma t\right)},
\end{align} 
and similarly for $\dd P$ and $\dd\bm{v}$. Here, $k$ is a real
wavenumber such that $|kr|\gg 1$, and the amplitude $\dd\rho(r,z)$ is
a slowly-varying function of $r$. Then 
$\p_r\to i k$ when acting on the above primitive perturbations, and we may
neglect curvature terms. 
We take  
$k>0$ without loss of  generality. The frequency $\sigma = \omega +
\ii s$ is generally complex, with $\omega$ being the real frequency
and $s$ is the real growth rate.

Introducing 
\begin{align}
  W \equiv \frac{\dd\rho}{\rho}, \quad Q \equiv \frac{\dd P}{\rho},
\end{align}
the linearized equations for
locally isothermal dusty gas with the pressure equation in place of
the dust-fraction (Eq. \ref{masseq}---\ref{momeq}, Eq. \ref{eff_energy}) are then:

\begin{align}
  \ii\sigma W &= \ii k \dd v_r + \dd v_z^\prime +
  \dd v_r \p_r\ln{\rho} + \dd v_z\p_z\ln{\rho},\label{lin_mass}\\
  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
  \delta\bm{F}\cdot\hat{\bm{r}},\label{lin_xmom}\\
  \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
    v_\phi}{\p z}\dd v_z, \label{lin_ymom}\\
  -\ii\sigma\dd v_z &=  \delta\bm{F}\cdot\hat{\bm{z}},\label{lin_zmom}\\
  \ii\sigma Q &= \frac{P}{\rho}\left(\ii k \dd v_r + \dd
               v_z^\prime\right) + \frac{1}{\rho}\left(\dd v_r\p_rP + \dd v_z \p_zP\right)\notag\\
                &\phantom{=}-\frac{P}{\rho} \dd v_r\p_r \ln{c_s^2} - \frac{\dd\mathcal{C}}{\rho},\label{lin_energy}
\end{align}  
where $^\prime \equiv \p_z$; $\dd\bm{F}$ is the linearized pressure force, given in Appendix
\ref{lin_press}; and 
%\begin{align}
%  \delta \bm{F} \equiv \frac{\dd\rho}{\rho^2}\nabla P -
%  \frac{1}{\rho}\nabla\dd P, 
%\end{align}
$\dd\mathcal{C}$ is the linearized dust-duffusion function, given in
Appendix \ref{lin_dust}.  We consider stopping times appropriate for
small grains in the Epstein regime. Note that for the axisymmetric
problem, $\dd\bm{F}$ is purely meridional.

Eq. \ref{lin_mass}---\ref{lin_energy} is a set of ordinary
differntial equations in $z$. All the coefficients are evaluated at a
fiducial radius $r=r_0$, but their full $z$-dependence is retained.  
\subsection{Boundary conditions}
Eq. \ref{lin_mass}---\ref{lin_energy} can be reduced to a set of
first-order differential equations for $W, Q$ and $\dd v_z$. We can
see this schematically as follows. Eq. \ref{lin_xmom} and
\ref{lin_ymom} may be combined to yield 
\begin{align*}
  \dd v_r = \dd v_r (W, Q, \dd v_z). 
\end{align*} 
We can then take the continuity equation as an equation for $\dd v_z$, 
\begin{align*}
\text{Eq. \ref{lin_mass}} \Rightarrow \dd v_z^\prime(z) = \dd
  v_z^\prime(W,Q,\dd v_z),
\end{align*}
and the vertical momentum equation as an equation for $Q$, 
\begin{align*}
\text{Eq. \ref{lin_zmom}} \Rightarrow Q^\prime(z) = 
 Q^\prime(W,Q,\dd v_z). 
\end{align*}

Now, for finite dust-gas coupling, $\tstop\neq0$, inspection of $\dd C$
(Appendix \ref{lin_dust}) shows that it involves $W^\prime$ (assuming
$\epsilon < 1$). Then Eq. \ref{lin_energy} may be taken as a
differential equation for $W$. However, for perfectly coupled dust,
$\tstop =\dd\mathcal{C}= 0$. In that case Eq. \ref{lin_energy} gives an algebraic
relation between $W$, $Q$ and $\dd v_z$. That is,
\begin{align*}
\text{Eq. \ref{lin_energy}}\Rightarrow \begin{cases}
  W^\prime(z) = W^\prime(W, Q, \dd v_z) & \text{if } \tstop \neq 0, \\
  W           (z) = W(Q, \dd v_z) & \text{if }\tstop = 0.
\end{cases}
\end{align*}

This means that for the perfectly coupled problem, we have a pair of
ODEs for $Q$ and $\dd v_z$, and two boundary conditions are needed. 
This is expected since this problem is exactly 
equivalent to that in adiabatic hydrodynamics \citep[e.g.]{lubow93}. When
$\tstop\neq0$, we require three boundary conditions. {\bf WHAT?? Why odd
  number of BCs? Shouldn't we have even number of BCs? We have
two boundaries? I've only seen problems like this involving even
number of BCs.} In this case we impose $\dd v_z =0$ at $z=\pm
z_\mathrm{max}$, and consider even modes about the midplane such that
$\dd v_z^\prime(0) = 0$.

### 2016-05-06 model.tex
updated notes

@@
 \begin{align}
-\frac{DP}{Dt} = - P \nabla\cdot\bm{v} + P\bm{v}\cdot\nabla\ln{c_s^2} +
-c_s^2 \nabla\cdot\left[\tstop\left(1 - \frac{P}{c_s^2\rho}\right)\nabla
-  P\right]. \label{eff_energy}
+\frac{DP}{Dt} &= - P \nabla\cdot\bm{v} + P\bm{v}\cdot\nabla\ln{c_s^2}
+                + \mathcal{C},  \label{eff_energy} \\
+\mathcal{C}&\equiv c_s^2 \nabla\cdot\left[\tstop\left(1 - \frac{P}{c_s^2\rho}\right)\nabla
+  P\right]
 \end{align}
 
-This is the energy equation for an ideal gas of adiabatic index
+Eq. \ref{eff_energy} is the energy equation for an ideal gas of adiabatic index
 $\gamma=1$, i.e. isothermal evolution, but now with additional source 
@@
 relation for linear, axisymmetric waves with frequency $\sigma$ (see
-Appendix \ref{var_prin}) in the form:  
+Appendix \ref{var_prin}, where we also account for finite $\tstop$) in
+the form:   
 \begin{align}
@@
 \begin{align}\label{dusty_cooling}
-  \frac{DP}{Dt} &= - P \nabla \cdot \bm{v} + \mathcal{C},\\
-  \mathcal{C} & = \nabla\cdot\left[\tstop\left(c_s^2 -
-    \frac{P}{\rho}\right)\nabla P\right].           
+  \frac{DP}{Dt} &= - P \nabla \cdot \bm{v} + \mathcal{C}
 \end{align} 

### 2016-05-07 model.tex
general edits

@@
 that for a star of mass $M_*$ at the origin 
-\begin{align}
-  \Phi(r,z) =-\frac{GM_*}{\sqrt{r^2 + z^2}},
+\begin{align}\label{thin_disk_potential}
+  \Phi(r,z) =-\frac{GM_*}{\sqrt{r^2 + z^2}}\simeq
+  -\frac{GM_*}{r}\left(1 - \frac{z^2}{2r^2}\right), 
 \end{align}
-where $G$ is the gravitational constant.
+where $G$ is the gravitational constant. The second equality is the 
+thin-disk approximation for the disk potential, appropriate for
+$|z|\ll r$. We shall adopt this approximate potential in order to
+obtain analytic expressions for the disk equilibria.
 
@@
 \mathcal{C}&\equiv c_s^2 \nabla\cdot\left[\tstop\left(1 - \frac{P}{c_s^2\rho}\right)\nabla
-  P\right]
+  P\right].
 \end{align}
-
 Eq. \ref{eff_energy} is the energy equation for an ideal gas of adiabatic index
@@
 
-\section{Equilibrium}
+\section{Equilibrium}\label{eqm}
 To obtain an axisymmetric steady state with $\rho(r,z)$ and 
@@
 
+\subsection{Disk structure with a prescribed dust distribution} 
+We assume a Gaussian profile in the dust-to-gas ratio, 
+\begin{align}\label{dust_gauss}
+  \tepsilon(r,z) = \tepsilon_0
+  \exp{\left(-\frac{z^2}{2\widetilde{H}^2}\right)}. 
+\end{align}
+For simplicity we assume the dust-to-gas ratio at the mid-plane
+$\tepsilon_0$, as well as its characteristic scale-height
+$\widetilde{H}$, are both constant.
+
+% We take the mid-plane dust-to-gas ratio
+%to be a power-law in radius,  
+%\begin{align}
+%  \tepsilon_0(r) = \tepsilon_{00}\left(\frac{r}{r_0}\right)^{-d},  
+%\end{align}
+%where $\tepsilon_{00}$ is the dust-to-gas ratio at the fiducial radius
+%$r_0$.
+
+Inserting Eq. \ref{dust_gauss} into vertical hydrostatic equilibrium,
+Eq. \ref{steady_momz} and integrating with the approximate
+gravitational potential (Eq. \ref{thin_disk_potential}) we obtain the
+gas density as
+\begin{align}
+  \rhog(r,z) = \rho_\mathrm{g0}(r)\exp{\left\{ - \frac{z^2}{2\Hgas^2}
+    -\tepsilon_0\frac{\Htilde^2}{\Hgas^2}\left[1 -
+      \exp{\left(-\frac{z^2}{2\Htilde^2}\right)}\right] \right\}}, 
+\end{align}
+where
+\begin{align}
+  \Hgas = \frac{c_s}{\OmK}, \quad \OmK \equiv \sqrt{\frac{GM_*}{r^3}},   
+\end{align}
+is the gas scale-height in the dust-free limit and $\OmK$ is the
+Keplerian frequency, respectively.
+
+In practice, we consider $\tepsilon_0 \ll 1$ and settled dust such that
+$\Htilde<\Hgas$. Then $\rhog(r,z)$ is effectively Gaussian, as in the
+dust-free case. The dust density is approximately 
+\begin{align}
+  \rhod \simeq \tepsilon_0\rho_\mathrm{g0}(r) \exp
+        {\left(-\frac{z}{2H_\mathrm{d}^2}\right)}, 
+\end{align}
+with 
+\begin{align}
+  \frac{1}{H_\mathrm{d}^2} = \frac{1}{\Htilde^2} + \frac{1}{\Hgas^2}, 
+\end{align}
+and $H_\mathrm{d}$ is the dust-scale height. In numerical
+calculations, we  we specify $H_\mathrm{d}< \Hgas$ to obtain 
+$\Htilde$ for input.
+
+\subsection{Orbital frequency} 
+In the thin-disk approximation, the disk orbital frequency is 
+\begin{align}
+  \Omega(r,z) = \OmK(r)\left[1 - \frac{3}{2}\frac{z^2}{r^2} +
+    \frac{h_\mathrm{g}^2}{\left(1+\tepsilon\right)}\frac{\p}{\p\ln{r}}\ln{\left(c_s^2\rhog\right)}
+    \right]^{1/2}, 
+\end{align}
+where 
+\begin{align}
+  h_\mathrm{g} \equiv \frac{\Hgas}{r}
+\end{align}
+is the characteristic disk aspect-ratio.
+
 \subsection{Vertical shear}
@@
 \begin{align}\label{vshear}
-  r\frac{\p \Omega^2}{\p z} &= \frac{\p\ln{\rho}}{\p r}\frac{\p}{\p
-    z}\left[c_s^2(1-\epsilon)\right] - \frac{\p\ln{\rho}}{\p z}
-  \frac{\p}{\p r} \left[c_s^2(1-\epsilon)\right]\\  
-  & = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p s}{\p z} -\frac{\p
-    P}{\p z}\frac{\p s}{\p r} \right),\notag
+  r\frac{\p \Omega^2}{\p z} 
+%&= \frac{\p\ln{\rho}}{\p r}\frac{\p}{\p
+%    z}\left[c_s^2(1-\epsilon)\right] - \frac{\p\ln{\rho}}{\p z}
+%  \frac{\p}{\p r} \left[c_s^2(1-\epsilon)\right]\\  
+   = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p s}{\p z} -\frac{\p
+    P}{\p z}\frac{\p s}{\p r} \right),
 \end{align}
@@
 \begin{align}
-   s \equiv \ln \frac{P}{\rho}
+   s \equiv \ln \frac{P}{\rho} = \ln{\left[c_s^2(1-\epsilon)\right]} 
 \end{align}
@@
 square of the epicylic frequency. We assume $\kappa^2>0$. 
-These criteria are applicable to any adiabatic fluid when $s$
-is the usual definition of entropy. For the dusty problem, we may take 
-$s=\ln{(1-\epsilon)}+\mathrm{const.}$ from the definition of $s$ and  
-the equation of state (Eq. \ref{eos}) with uniform sound-speed.
-
+These criteria are applicable to any adiabatic fluid. For 
+isothermal gas perfectly coupled to dust, we may take 
+$s=\ln{(1-\epsilon)}$. 
+%from the definition of $s$ and  
+%the equation of state (Eq. \ref{eos}) with uniform sound-speed. 
 %\begin{align}
@@
 %Eq. \ref{eos}, with a constant sound-speed. 
-
 The stability conditions are thus
-\begin{align}\label{dusty_solberg}
-  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\epsilon &> 0,  \\
+\begin{align}
+  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\epsilon &> 0,\label{dusty_solberg1}  \\
   -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\epsilon}{\p
-    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \epsilon}{\p r} \right) & > 0. 
+    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \epsilon}{\p r} \right) & > 0. \label{dusty_solberg2}
 \end{align}
 
-In accretion disks, $\rhog$, $\rhod$ and $\epsilon$ are expected to
-decrease away from the mid-plane
-\begin{align*}  
-  \frac{\p\rhog}{\p |z|} < 0,
+%In accretion disks, $\rhog$, $\rhod$ and $\epsilon$ are expected to
+%decrease away from the mid-plane
+%\begin{align*}  
+%  \frac{\p\rhog}{\p |z|} < 0,
 %, \quad \frac{\p \rhog}{\p r} < 0,   
-\end{align*}
-and similarly for $\rhod$, $\epsilon$, and the dust-to-gas ratio $\tepsilon$. 
-We may thus evaluate the stability criteria at $z>0$ without loss of
-generality. We insert the vertical shear profile,
-Eq. \ref{vshear}, into Eq. \ref{dusty_solberg} and write the result 
-in terms of $\tepsilon$, to find
+%\end{align*}
+%and similarly for $\rhod$, $\epsilon$, and the dust-to-gas ratio $\tepsilon$. 
+%We may thus evaluate the stability criteria at $z>0$ without loss of
+%generality.
 
-\begin{align} 
-  &1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1+
-    \tepsilon\right)^2}\frac{\Omega^2}{\kappa^2}\left(\frac{\p\ln{\rhog}}{\p\ln{r}}
-  \frac{\p\ln{\tepsilon}}{\p\ln{r}} + r^2\frac{\p\ln{\rhog}}{\p z}
-  \frac{\p\ln{\tepsilon}}{\p z} \right)>0,   \\
-&-\frac{\p\ln{\tepsilon}}{\p z} \notag\\
-  &- \frac{\tepsilon
-  h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}\frac{\Omega^2}{\kappa^2}
-\frac{\p\ln{\tepsilon}}{\p\ln{r}}\left( \frac{\p\ln{\rhog}}{\p\ln{r}}\frac{\p\ln{\tepsilon}}{\p
-    z} -\frac{\p\ln{\rhog}}{\p z}\frac{\p\ln{\tepsilon}}{\p\ln{
-    r}}\right) > 0,
-%  \kappa^2\left|\frac{\p\epsilon}{\p z}\right| + c_s^2\left|\frac{\p
-%    \epsilon}{\p r}\frac{\p\epsilon}{\p z}\frac{\p\ln{\rho}}{\p
-%    r}\right| - c_s^2 \left(\frac{\p\epsilon}{\p
-%    r}\right)^2\left|\frac{\p\ln{\rho}}{\p z}\right| > 0.  \label{criterion2}
-\end{align}
-as the stability criteria, where 
-\begin{align}
-  h_\mathrm{g}\equiv c_s/\left(r\Omega\right)
-\end{align}
-is the characteristic aspect-ratio of the gas component.
-
-%This equation shows that the vertical gradient in dust-fraction (and
-%thus the dust-to-gas ratio $\rhod/\rhog$) has a stabilizing effect for
-%axisymmetric perturbations. Instead, the only way to destabilize the
-%system is a sufficiently large radial gradient in the dust-fraction
-%and/or vertical (total) density gradient (the last term).
-
-%Note that Eq. \ref{criterion2} is only applicable for the assume disk
-%structure described above: decreasing dust and gas densities, as well
-%as the dust-fraction, away from the star and away from the
-%midplane. For less typical disk structures, one must evalulate above the
-%critera seperately.
+%We insert the vertical shear profile,
+%Eq. \ref{vshear}, into Eq. \ref{dusty_solberg} and write the result 
+%in terms of $\tepsilon$, to find
 
 \subsubsection{Order-of-magnitude estimates} 
-We evaluate the above stability criteria by assuming the gas and dust
-distributions have a Gaussian 
-distribution vertically, 
-\begin{align}
-  \rho_\mathrm{g,d}(r,z) = \rho_\mathrm{g0,d0}(r) \exp
-      {\left[-\frac{z^2}{2H^2_\mathrm{g,d}(r)}\right]}.
-\end{align}
-We note that $H_\mathrm{g}=h_gr$ by definition. The dust-to-gas ratio
-is  
-\begin{align}
-  \tepsilon = \tepsilon_0(r)
-  \exp{\left[-\frac{z^2}{2\widetilde{H}^2(r)}\right]}, 
-\end{align}
-with $\tepsilon_0= \rho_\mathrm{d,0}/\rho_{g_0}$, and  
-\begin{align}
-  \frac{1}{\widetilde{H}^2} \equiv \frac{1}{H_\mathrm{d}^2} -
-  \frac{1}{H_\mathrm{g}^2}. 
-\end{align}
-We further assume the disk is approximately Keplerian so that
-\begin{align}
-  \kappa \simeq \Omega. 
-\end{align}
-%Inserting the above approximations into Eq. \ref{criterion2} and writing the result in terms of the
-%dust-to-gas ratio $\tepsilon$,           
-%we find 
-%\begin{align}\label{ineq_est}
-%  1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1 + \tepsilon\right)^2}
-%  \left|\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right| 
-%  \left[
-%    \left|\frac{\p\ln{\rhog}}{\p\ln{r}}\right| - \frac{\widetilde{H}^2}{H^2_\mathrm{g}}\left|\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right|  
-%    \right]>0
-%\end{align}
-%is required \emph{for stability.} 
-%This inequality is easily satisifed in typical disk models where 
-%the dust layer is much thinner than the gas scale-height, so
-%that $\widetilde{H}\sim H_\mathrm{d}\ll H_\mathrm{g}$, and where the radial profiles smoothly vary on a scale of $r$.
-%Then the magnitude the only destabilizing term (the third term) is negligible.
-According to Eq.
-\ref{ineq_est}, instability is on%ly possible if there is a rapid variation the radial profile of the dust-to-gas ratio.
-Then the above criteria becomes
+We evaluate the above stability criteria for the vertically-Gaussian
+distributions for the gas density and the dust-to-gas ratio as 
+described in \S\ref{eqm}. However, here we permit 
+$\tepsilon_0=\tepsilon_0(r)$ and $\Htilde=\Htilde(r)$ to vary with
+radius. We assume the disk is approximately Keplerian so that
+$\kappa\simeq\OmK$. In terms of the dust-to-gas
+ratio $\tepsilon$, Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2}
+become   
 \begin{align}
@@
 Eq. \ref{stability_est1}---\ref{stability_est2} are easily satisfied
-in typical accretion disk models where radial gradients are $O(1/r)$ and
+in thin accretion disks where radial gradients are $O(1/r)$ and
 $h_\mathrm{g},\tepsilon, \widetilde{H}/H_\mathrm{g}\ll 1$, since then the
 magnitude of the second term on the left-hand-side  of either
-inequality is much less than unity. Such disks are stable against
+inequality is much less than unity. Thus radially smooth, strictly
+isothermal disks perfectly coupled with dust are stable against
 axisymmetric perturbations.
+
+%Such disks are stable against
+%axisymmetric perturbations.
 
@@
 \end{align} 
+In typical accretion disks where $\p_r\rhog<0$, it is easier to
+achieve instability at a given radius for increasing dust-to-gas
+ratios in outwards ($\p_r\tepsilon > 0$), and vice versa.
 
-Now, if $\p_r\rhog<0$, as typical in
-accretion models, then it is easier to achieve instability at a given
-radius for rapidly increasing dust-to-gas ratios ($\p_r\tepsilon >
-0$), and vice versa. 
-%This might occur, for example, when dust is
-%inserted in a finite radial range within a wider gaseous disk. 
-%Physically, one might expect
-%$\p_r\rhog\p_r\tepsilon > 0$ because
-
-\emph{If} the $\p_r\rhog$ can be neglected in Eq. \ref{spm},
-then instability requires
+If the $\p_r\rhog$ can be neglected in Eq. \ref{spm}, then instability
+requires 
 \begin{align}
@@
 The real coefficients $A,B,C,D$ can be read off
-Eq. \ref{integral_ex}. Note that $B=C$ from the assume equilibrium
-structure (Eq. \ref{vshear}).
+Eq. \ref{integral_ex}. Note that $B=C$ from dynamical equilibrium
+(Eq. \ref{vshear}).
 
 If $c_s$ is constant, then $\sigma^2$ is real, and 
-the first integral leads to the usual Solberg-Hoiland criteria for
+the first integral leads to the previous Solberg-Hoiland criteria for 
 axisymmetric stability. However, when $c_s$ is non-uniform, $\sigma^2$ 

### 2016-05-10 linear.tex
sample results

@@
                v_z^\prime\right) + \frac{1}{\rho}\left(\dd v_r\p_rP + \dd v_z \p_zP\right)\notag\\
-                &\phantom{=}-\frac{P}{\rho} \dd v_r\p_r \ln{c_s^2} - \frac{\dd\mathcal{C}}{\rho},\label{lin_energy}
+                &\phantom{=}-\frac{P}{\rho}\left( \dd v_r\p_r
+               \ln{c_s^2} + \dd v_z \p_z\ln{c_s^2}\right)-
+               \frac{\dd\mathcal{C}}{\rho},\label{lin_energy} 
 \end{align}  
@@
 number of BCs.} In this case we impose $\dd v_z =0$ at $z=\pm
-z_\mathrm{max}$, and consider even modes about the midplane such that
-$\dd v_z^\prime(0) = 0$.
+z_\mathrm{max}$, and classify modes according to their symmetry about
+the midplane: `even' modes with $\dd v_z^\prime(0) = 0$, and `odd' 
+modes with $\dd v_z(0)=0$.

### 2016-05-10 result.tex
sample results

@@
 \section{Results}
+
+We consider a globally isothermal disk, $c_s=\mathrm{const.}$, by
+setting $q=0$ in the equation of state, Eq. \ref{eos}, and a flat mid-plane 
+gas density, $\rhog(0) = \mathrm{const.}$, so there is no global
+radial pressure gradient at the disk midplane. The stopping time is
+prescribed as 
+\begin{align}
+  \tstop = t_\mathrm{s0}\frac{\rho(z=0)}{\rho(z)}. 
+\end{align}
+
+As an example, we set the dust-to-gas ratio $\tepsilon_0=0.025$, dust 
+scale height $H_\mathrm{d}=0.4\Hgas$, and stopping time
+$t_\mathrm{s0}=0.02$. We consider modes with radial wavenumber
+$k_x\Hdust=2\pi$. The vertical domain size is set to 
+$\zmax=2\Hgas$ and we use $N=513$ grid points. Fig. \ref{eigen_map}
+shows a map of the eigenvalues in the $(\omega, s)$ plane.
 
@@
   \includegraphics[width=\linewidth]{figures/eigenvalues}
-\caption{do not believe}
+  \caption{Eigenvalues found for the fiducial isothermal dusty
+    problem.
+  \label{eigen_map}} 
 \end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/eigenvec}
+  \caption{Example eigenvectors from the fiducial case. 
+  \label{eigen_vec}} 
+\end{figure}

### 2016-05-10 code/fixed/basic.f90
code update

@@ -11,7 +11,8 @@ real*8 function dlnrhog_dr(zhat)
   implicit none 
   real*8, intent(in) :: zhat 
  
-  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) )
+  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
+              + smalld*dgratio*smallh_g*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) )
 end function dlnrhog_dr
 
 real*8 function d2lnrhog_dr2(zhat)
@@ -20,7 +21,9 @@ real*8 function d2lnrhog_dr2(zhat)
   implicit none 
   real*8, intent(in) :: zhat 
   
-  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) 
+  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
+                 -4d0*smalld*dgratio*smallh_g**2d0*delta**2d0*dlnHg_dlnr*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) &
+                 -(smalld*smallh_g*delta)**2d0*dgratio*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) )
 end function d2lnrhog_dr2
 
 real*8 function dlnrhog_dz(zhat)
@@ -141,6 +144,16 @@ real*8 function dlnrho_dz(zhat)
   dlnrho_dz =dlnrhog_dz(zhat) + deps_tilde_dz(zhat)/(1d0 + eps_tilde(zhat))
 end function dlnrho_dz
 
+
+real*8 function d2lnrho_dz2(zhat)
+  use global
+  implicit none
+  real*8, intent(in) :: zhat 
[…]

### 2016-05-10 code/fixed/dvsi.f90

@@ -202,7 +202,7 @@ subroutine fill_bigmatrix(knum)
      !continuity equation 
      L11(i,:) = (0d0, 0d0)
      L12(i,:) = (0d0, 0d0)
-     L13(i,:) = -( ii*kx + dlnrhodr_arr(i)*0d0)*T(i,:)
+     L13(i,:) = -( ii*kx + dlnrhodr_arr(i))*T(i,:)
      L14(i,:) = (0d0, 0d0)
      L15(i,:) =( -Tp(i,:) - dlnrhodz_arr(i)*T(i,:) )
      
@@ -213,48 +213,48 @@ subroutine fill_bigmatrix(knum)
      R15(i,:) = zero
 
      !dust continuity equation (equiv. energy equation) 
-!!$     L21(i,:) = -tstop_arr(i)*(eps_arr(i)*divF_arr(i) + F_dot_deleps_arr(i) - 0.5d0*eps_arr(i)*F_dot_dellncs2_arr(i) )*T(i,:) 
-!!$     L21(i,:) =  L21(i,:) &
-!!$          + tstop_arr(i)*eps_arr(i)*( &
-!!$          -smallh_g*smallq*(1d0 - eps_arr(i))*ii*kx*T(i,:) &
-!!$          + ii*kx*depsdr_arr(i)*T(i,:) &
-!!$          + depsdz_arr(i)*Tp(i,:) &
-!!$          - (1d0 - eps_arr(i))*( -kx*kx*T(i,:) + Tpp(i,:) ) &
-!!$          )
-!!$     L21(i,:) =  L21(i,:) &
-!!$          + tstop_arr(i)*( &
-!!$          -(1d0 - eps_arr(i))*(ii*kx*depsdr_arr(i)*T(i,:) + depsdz_arr(i)*Tp(i,:) ) &
-!!$          )
-!!$     L21(i,:) =  L21(i,:) &
-!!$          + tstop_arr(i)*( &
-!!$          0.5*eps_arr(i)*smallh_g*smallq*(1d0 - eps_arr(i))*ii*kx*T(i,:) &
-!!$          )
-!!$    
-!!$
-!!$     L22(i,:) =  tstop_arr(i)*divF_arr(i)*T(i,:)
-!!$     L22(i,:) =  L22(i,:) + tstop_arr(i)*( Fr_arr(i)*ii*kx*T(i,:) + Fz_arr(i)*Tp(i,:) )
-!!$     L22(i,:) =  L22(i,:) &
[…]

### 2016-05-10 code/iso/basic.f90

@@ -1,314 +1,0 @@
-real*8 function lnrhog(zhat)
-  use global
-  implicit none 
-  real*8, intent(in) :: zhat 
-
-  lnrhog = -0.5d0*zhat**2d0 - dgratio*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) 
-end function lnrhog
-
-real*8 function dlnrhog_dr(zhat)
-  use global
-  implicit none 
-  real*8, intent(in) :: zhat 
- 
-  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) )
-end function dlnrhog_dr
-
-real*8 function d2lnrhog_dr2(zhat)
-  !this is (1/r)*(d/dr)( r*dlnrhog_dr )
-  use global
-  implicit none 
-  real*8, intent(in) :: zhat 
-  
-  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) 
-end function d2lnrhog_dr2
-
-real*8 function dlnrhog_dz(zhat)
-  use global
-  implicit none 
-  real*8, intent(in) :: zhat 
-  
-  dlnrhog_dz = -zhat*(1d0 + dgratio*exp( - zhat**2d0/2d0/delta**2d0 ) )
-end function dlnrhog_dz
-
-real*8 function d2lnrhog_dz2(zhat)
-  use global
-  implicit none 
-  real*8, intent(in) :: zhat 
-  
-  d2lnrhog_dz2 = -1d0 - dgratio*exp(-zhat**2d0/2d0/delta**2d0)*(1d0 - zhat**2d0/delta**2d0)
-end function d2lnrhog_dz2
-
-real*8 function del2_rhog(zhat)
-  use global
-  implicit none 
-  real*8, intent(in) :: zhat 
-  real*8, external  :: d2lnrhog_dr2, d2lnrhog_dz2
-
-  del2_rhog = d2lnrhog_dr2(zhat) + d2lnrhog_dz2(zhat)
-end function del2_rhog
-
[…]

### 2016-05-10 code/iso/dvsi.f90

@@ -1,423 +1,0 @@
-module global
-  implicit none 
-  integer, parameter :: nvar = 5 
-  real*8, parameter :: pi = 2d0*acos(0d0)
-  complex*16, parameter :: ii = (0d0, 1d0), zero = (0d0, 0d0)
-
-  integer :: nz, bignz, Nzmid
-  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop
-  real*8 :: zmax 
-  real*8 :: dlnHg_dlnr
-
-  real*8, allocatable :: zaxis(:), lnrho_arr(:), eps_arr(:), dlnrhodr_arr(:), depsdr_arr(:),dlnrhodz_arr(:), depsdz_arr(:) 
-  real*8, allocatable :: del2lnrho_arr(:), del2eps_arr(:), dellnrho_dot_deleps_arr(:), tstop_arr(:)
-  real*8, allocatable :: Fr_arr(:), Fz_arr(:), F_dot_deleps_arr(:), divF_arr(:)
-  real*8, allocatable :: omega_arr(:), kappa2_arr(:), vshear_arr(:) 
-
-  real*8, allocatable     :: T(:,:), Tp(:,:), Tpp(:,:)
-
-  complex*16, allocatable :: L11(:,:), L12(:,:), L13(:,:), L14(:,:), L15(:,:)
-  complex*16, allocatable :: L21(:,:), L22(:,:), L23(:,:), L24(:,:), L25(:,:)
-  complex*16, allocatable :: L31(:,:), L32(:,:), L33(:,:), L34(:,:), L35(:,:)
-  complex*16, allocatable :: L41(:,:), L42(:,:), L43(:,:), L44(:,:), L45(:,:)
-  complex*16, allocatable :: L51(:,:), L52(:,:), L53(:,:), L54(:,:), L55(:,:)
-  complex*16, allocatable :: R11(:,:), R12(:,:), R13(:,:), R14(:,:), R15(:,:) 
-  complex*16, allocatable :: R21(:,:), R22(:,:), R23(:,:), R24(:,:), R25(:,:)
-  complex*16, allocatable :: R31(:,:), R32(:,:), R33(:,:), R34(:,:), R35(:,:)
[…]

### 2016-05-10 code/parity/basic.f90

@@ -11,7 +11,8 @@ real*8 function dlnrhog_dr(zhat)
   implicit none 
   real*8, intent(in) :: zhat 
  
-  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) )
+  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
+              + smalld*dgratio*smallh_g*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) )
 end function dlnrhog_dr
 
 real*8 function d2lnrhog_dr2(zhat)
@@ -20,7 +21,9 @@ real*8 function d2lnrhog_dr2(zhat)
   implicit none 
   real*8, intent(in) :: zhat 
   
-  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) 
+  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
+                 -4d0*smalld*dgratio*smallh_g**2d0*delta**2d0*dlnHg_dlnr*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) &
+                 -(smalld*smallh_g*delta)**2d0*dgratio*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) )
 end function d2lnrhog_dr2
 
 real*8 function dlnrhog_dz(zhat)
@@ -141,6 +144,16 @@ real*8 function dlnrho_dz(zhat)
   dlnrho_dz =dlnrhog_dz(zhat) + deps_tilde_dz(zhat)/(1d0 + eps_tilde(zhat))
 end function dlnrho_dz
 
+
+real*8 function d2lnrho_dz2(zhat)
+  use global
+  implicit none
+  real*8, intent(in) :: zhat 
[…]

### 2016-05-10 code/parity/dvsi.f90

@@ -1,6 +1,6 @@
 module global
   implicit none 
-  integer, parameter :: nvar = 4 
+  integer, parameter :: nvar = 5 
   real*8, parameter :: pi = 2d0*acos(0d0)
   complex*16, parameter :: ii = (0d0, 1d0), zero = (0d0, 0d0)
 
@@ -9,13 +9,14 @@ module global
   real*8 :: zmax 
   real*8 :: dlnHg_dlnr
   real*8 :: krad, bc_tol 
+  real*8 :: max_rate 
 
   real*8, allocatable :: zaxis(:), lnrho_arr(:), eps_arr(:), dlnrhodr_arr(:), depsdr_arr(:),dlnrhodz_arr(:), depsdz_arr(:) 
-  real*8, allocatable :: del2lnrho_arr(:), del2eps_arr(:), dellnrho_dot_deleps_arr(:), tstop_arr(:)
+  real*8, allocatable :: del2lnrho_arr(:), del2eps_arr(:), dellnrho_dot_deleps_arr(:), tstop_arr(:), d2lnrhodz2_arr(:)
   real*8, allocatable :: Fr_arr(:), Fz_arr(:), F_dot_deleps_arr(:), divF_arr(:), F_dot_dellncs2_arr(:)
   real*8, allocatable :: omega_arr(:), kappa2_arr(:), vshear_arr(:) 
 
-  complex*16, allocatable     :: T(:,:), Tp(:,:), Tpp(:,:)
+  complex*16, allocatable     :: T(:,:), Tp(:,:), Tpp(:,:), T_odd(:,:), Tp_odd(:,:),  Tpp_odd(:,:)
 
   complex*16, allocatable :: L11(:,:), L12(:,:), L13(:,:), L14(:,:), L15(:,:)
   complex*16, allocatable :: L21(:,:), L22(:,:), L23(:,:), L24(:,:), L25(:,:)
@@ -38,12 +39,12 @@ end module global
 program dvsi
   use global
   implicit none 
-  integer :: i, j, k, lmax 
-  real*8 :: z, zbar, m, T_l, Tp_l, Tpp_l, dT_l, d2T_l
+  integer :: i, j, k, lmax, jmid
+  real*8 :: z, zbar, m, T_l, Tp_l, Tpp_l, dT_l, d2T_l, n 
[…]

### 2016-05-10 code/press/basic.f90

@@ -11,7 +11,8 @@ real*8 function dlnrhog_dr(zhat)
   implicit none 
   real*8, intent(in) :: zhat 
  
-  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) )
+  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
+              + smalld*dgratio*smallh_g*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) )
 end function dlnrhog_dr
 
 real*8 function d2lnrhog_dr2(zhat)
@@ -20,7 +21,9 @@ real*8 function d2lnrhog_dr2(zhat)
   implicit none 
   real*8, intent(in) :: zhat 
   
-  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) 
+  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
+                 -4d0*smalld*dgratio*smallh_g**2d0*delta**2d0*dlnHg_dlnr*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) &
+                 -(smalld*smallh_g*delta)**2d0*dgratio*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) )
 end function d2lnrhog_dr2
 
 real*8 function dlnrhog_dz(zhat)

### 2016-05-10 code/press/dvsi.f90

@@ -249,12 +249,13 @@ subroutine fill_bigmatrix(knum)
           - divF_arr(i)*T(i,:) & 
           + eps_arr(i)*(&
           ii*kx*dlnrhodr_arr(i)*T(i,:) + dlnrhodz_arr(i)*( Tp(i,:) + dlnrhodz_arr(i)*T(i,:) ) &
-          + kx*kx*T(i,:) - ( Tpp(i,:) + 2d0*Tp(i,:)*dlnrhodz_arr(i) + T(i,:)*(d2lnrhodz2_arr(i) + dlnrhodz_arr(i)**2d0) ) &
+          + kx*kx*T(i,:) &
+          - ( Tpp(i,:) + 2d0*Tp(i,:)*dlnrhodz_arr(i) + T(i,:)*(d2lnrhodz2_arr(i) + dlnrhodz_arr(i)**2d0) ) &
           )&
           +0.5d0*Fr_arr(i)*smallh_g*smallq*T(i,:) &
           +0.5d0*eps_arr(i)*smallh_g*smallq*ii*kx*T(i,:) &
           )
-     
+    
      L23(i,:) =-gr_at_z*T(i,:) - ii*kx*(1d0 - eps_arr(i))*T(i,:) + (1d0 - eps_arr(i))*smallh_g*smallq*T(i,:)
      L24(i,:) = (0d0, 0d0)
      L25(i,:) =-gz_at_z*T(i,:) -       (1d0 - eps_arr(i))*Tp(i,:)
@@ -321,24 +322,53 @@ subroutine fill_bigmatrix(knum)
      R54(i,:) = zero
      R55(i,:) = zero 
   enddo
-  
-  !delta eps = 0 , if stopping time is non zero
-  if(tstop.gt.0d0) then
-     do i = 1, nz, nz-1
-        L21(i,:) = (1d0 - eps_arr(i))*T(i,:)
-        L22(i,:) =-T(i,:)
-        L23(i,:) = zero 
-        L24(i,:) = zero 
-        L25(i,:) = zero
-        
-        R21(i,:) = zero 
-        R22(i,:) = zero 
-        R23(i,:) = zero 
-        R24(i,:) = zero
-        R25(i,:) = zero 
-     enddo
-  endif
 
+!!$  !deps = 0 
+!!$  do i = 1, nz, nz-1 
+!!$     L21(i,:) = (1d0 - eps_arr(i))*T(i,:)
[…]

### 2016-05-11 model.tex
sample results
notes and code update

@@
 where $c_s$ is a prescribed sound-speed profile fixed in time.  
+%For analytic discussion we consider arbitrary functions of
+%$c_s(r,z)$. 
+In numerical calculations we will consider vertically  
+isothermal disks with 
+\begin{align}
+  c_s^2(r) \propto r^{q},
+\end{align}
+where $q$ is the power-law index for the disk temperature.
 
-\subsection{Effective mixture energy equation}
+\subsection{Effective energy equation}
 We show that for a prescribed temperature distribution, the mixture
 obeys an evolutionary energy equation, due to the advection of the
-dust-fraction.
-
-We eliminate $\epsilon$ in favor of $P$, 
-\begin{align}
+dust-fraction. Eliminate $\epsilon$ in favor of $P$, 
+\begin{align*}
   \epsilon = 1 - \frac{P}{c_s^2(r,z)\rho}, 
-\end{align}
+\end{align*}
 then Eq. \ref{dusteq} becomes
-
 \begin{align}
@@
                 + \mathcal{C},  \label{eff_energy} \\
-\mathcal{C}&\equiv c_s^2 \nabla\cdot\left[\tstop\left(1 - \frac{P}{c_s^2\rho}\right)\nabla
+\mathcal{C}&\equiv c_s^2 \nabla\cdot\left[\tstop\left(1 -
+  \frac{P}{c_s^2\rho}\right)\nabla 
   P\right].

### 2016-05-11 result.tex
notes and code update

--- 2016-05-04 result.tex
@@
 \section{Results}
+
+We consider a globally isothermal disk, $c_s=\mathrm{const.}$, by
+setting $q=0$ in the equation of state (Eq. \ref{eos}), and a flat
+mid-plane gas density, $\rhog(0) = \mathrm{const.}$, so there is no
+global radial pressure gradient at the disk midplane. This setup
+eliminates classic vertical shear instability and streaming
+instabilities. The stopping time is
+prescribed as 
+\begin{align}
+  \tstop = t_\mathrm{s0}\frac{\rho(z=0)}{\rho(z)}. 
+\end{align}
+
+As an example, we set the midplane dust-to-gas ratio 
+$\tepsilon_0=0.02$, dust scale-height $H_\mathrm{d}=0.6\Hgas$, and
+stopping time $t_\mathrm{s0}\OmK=0.02$. We consider modes with radial
+wavenumber $k_x\Hdust=2\pi$. The vertical domain size is set to  
+$\zmax=2\Hgas$. The aspect-ratio of the gas layer is
+$h_\mathrm{g}=0.05$.
+
+We use a pseudo-spectral method to solve  
+Eq. \ref{lin_mass}---\ref{lin_energy} as a generalized eigenvalue
+problem with $N=513$ grid points. Fig. \ref{eigen_map}   
+shows a map of the eigenvalues in the $(\omega, s)$ plane. Note that
+we restrict attention to modes with $|\sigma|\leq \OmK$. 
+Eigenvalues on the `Y' branch should not be trusted. The 
+high growth-rate modes are associated with disturbances concentrated
+at the domain boundaries. Modes with $\omega\simeq 0$, i.e. purely
+growing modes, are likely spurious because we generally expect
+eigenvalues to be complex (see the integral relation in Appendix
+\ref{var_prin}).
 
@@
   \includegraphics[width=\linewidth]{figures/eigenvalues}
-\caption{do not believe}
+  \caption{Eigenvalues found for the fiducial isothermal dusty
+    problem. 
+  \label{eigen_map}} 
 \end{figure}
+
+We thus consider the slower-growing modes near the horizontal
+axis. Fig. \ref{eigen_vec} show example eigenfunctions for a case with
+frequency $(\omega,s)=(-2.79,0.20)\hgas\OmK$.
+
+In Fig. \ref{eigen_2d}
+we visualize this mode in the $(x,z)$ plane: colors show the
+perturbation to the dust-fraction ($\dd\epsilon$), while vectors show
+the perturbed meridional momenta. This mode has a growth
+timescale of $\simeq 16$ orbits. This mode is unlikely driven by
+vertical shear: the maximum vertical 
+shear rate, $\mathrm{max}\left|r\p_z\Omega\right|\simeq10^{-2}\hgas\OmK$,
+is much smaller than the mode growth rate $s\simeq 0.2\hgas\OmK$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/eigenvec}
+  \caption{Example eigenfunctions from the fiducial case. 
+    \label{eigen_vec}} 
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/result2d}
+  \caption{Example eigenvectors from the fiducial case. The eigenvalue
+    is $(\omega,s)=(-2.79,0.20)\hgas\OmK$.  
+    \label{eigen_2d}} 
+\end{figure}

### 2016-05-11 code/parity/dvsi.f90

@@ -58,7 +58,7 @@ program dvsi
   close(7)
 
   open(10,file='params.dat') 
-  write(10,fmt='(2(e22.15,x))'), dble(nz), smallh_g 
+  write(10,fmt='(4(e22.15,x))'), dble(nz), smallh_g, krad, Hd
   close(10)
 
   !check we have odd grid points so one grid point coincides with midplane 
@@ -342,12 +342,6 @@ subroutine fill_bigmatrix(knum)
 
 
 
-
-
-
-
-
-
   !fill big matrices 
   ibeg = 1
   iend = nz
@@ -428,11 +422,16 @@ subroutine eigenvalue_problem
   integer :: i, j, info, loc(1), cnt 
   integer :: lwork 
   complex*16, allocatable :: work(:)
-  complex*16 :: apha(bignz), bta(bignz), eigen(bignz), dP(nz), ddP(nz), dbigW(nz), ddfrac(nz) 
+  complex*16 :: apha(bignz), bta(bignz), eigen(bignz), bigQ(nz), dbigQ(nz), d2bigQ(nz), dbigW(nz), dvz(nz)
+  complex*16 :: err1(nz), err2(nz), err3(nz), err4(nz), err5(nz)
+  complex*16 :: dFx(nz), dFz(nz), ddivF(nz), ep(nz), depdx(nz), depdz(nz), dC(nz), divv(nz)
   complex*16 :: vl(bignz,bignz), vr(bignz,bignz)
-  real*8, parameter :: min_rate = 1d-6, max_eigen = 1d0 
+  real*8, parameter :: min_rate = 1d-6  
   real*8 :: rwork(8*bignz), eigen_re, eigen_im, eigen_mag, vz_max, error_in, error_out 
+  real*8 :: kx
   
+  kx = krad/Hd 
+
   lwork = 4*bignz
   allocate(work(lwork))
 
@@ -443,11 +442,11 @@ subroutine eigenvalue_problem
 
   !compute eigenvalues, output data 
 
-  bc_tol = 1d-4 
-
-
   open(20,file='eigenvalues.dat')
   open(30,file='eigenvectors.dat')
+  open(40,file='error.dat')
+  open(50,file='nonadia.dat')
+  
   cnt = 0 
   do i=1, bignz
      eigen(i) = apha(i)/bta(i)
[…]

### 2016-05-11 code/press/dvsi.f90

@@ -57,7 +57,7 @@ program dvsi
   close(7)
 
   open(10,file='params.dat') 
-  write(10,fmt='(2(e22.15,x))'), dble(nz), smallh_g 
+  write(10,fmt='(4(e22.15,x))'), dble(nz), smallh_g, krad, Hd
   close(10)
 
   !check we have odd grid points so one grid point coincides with midplane 
@@ -368,6 +368,7 @@ subroutine fill_bigmatrix(knum)
 !!$     R25(i,:) = zero 
 !!$  enddo
 
+!note: no lag pert in rhod give crazy results 
   
   !fill big matrices 
   ibeg = 1

### 2016-07-28 model.tex
text and code update

@@
 isothermal disks with 
-\begin{align}
+\begin{align}\label{power_temp}
   c_s^2(r) \propto r^{q},
@@
 
-\subsection{Disk structure with a prescribed dust distribution} 
+\subsection{Disk structure with a prescribed dust distribution}\label{steady_state}  
 We assume a Gaussian profile in the dust-to-gas ratio, 
 \begin{align}\label{dust_gauss}
-  \tepsilon(r,z) = \tepsilon_0
-  \exp{\left(-\frac{z^2}{2\widetilde{H}^2}\right)}. 
+  \tepsilon(r,z) = \tepsilon_0(r)
+  \exp{\left[-\frac{z^2}{2\widetilde{H}^2(r)}\right]}. 
 \end{align}
-For simplicity we assume the dust-to-gas ratio at the mid-plane
-$\tepsilon_0$, as well as its characteristic scale-height
-$\widetilde{H}$, are both constant.
-
+%For simplicity we assume the dust-to-gas ratio at the mid-plane
+%$\tepsilon_0$, as well as its characteristic scale-height
+%$\widetilde{H}$, are both constant.
 % We take the mid-plane dust-to-gas ratio
@@
 %$r_0$.  
-
 Inserting Eq. \ref{dust_gauss} into vertical hydrostatic equilibrium,
@@
 is a pseudo-entropy for the mixture.
+
+Writing Eq. \ref{vshear} in terms of the gas density and dust-to-gas
+ratio with a power-law temperature profil (Eq. \ref{power_temp}) e gives 
+\begin{align*}
+  &r\frac{\p \Omega^2}{\p z}  =
+  \frac{c_s^2(r)}{\left(1+\tepsilon\right)^2}\left\{
+  \frac{\p\tepsilon}{\p r}\frac{\p\ln{\rhog}}{\p z}
+  -\frac{\p\tepsilon}{\p z}\frac{\p\ln{\rhog}}{\p r}\right.\notag\\
+  &\phantom{ r\frac{\p \Omega^2}{\p z}  =
+    \frac{c_s^2(r)}{\left(1+\tepsilon\right)^2}\left\{\right\} }
+  \left. +\frac{q}{r}\left[\frac{\p\tepsilon}{\p z} + \left(1+\tepsilon\right)\frac{\p\ln{\rhog}}{\p z}\right]
+  \right\} 
+\end{align*}
+The first two terms correspond to vertical shear caused by spatial
+variations in the dust-to-gas ratio; while the second term
+proportional to $q$ corresponds to vertical shear due to the 
+radial temperature gradient.
+
+We can compare these sources by
+evaluating them, to order-of-magnitude, using the equilibrium
+solutions in \S\ref{steady_state}. We assume the disk is radially
+smooth so that $\p_r\sim 1/r$, and the dust-to-gas ratio
+$\tepsilon\ll1$. This gives
+
+\begin{align}\label{vshear_split}
+  \mathcal{V} =  \frac{\left|r\p_z\Omega\right|_{\text{
+        dust/gas gradient}}}{\left|r\p_z\Omega\right|_{\text{
+        temp. gradient}}} \sim
+  \tepsilon \frac{\mathrm{max}\left(\delta^2,
+    1\right)}{\left|q\right|\left(\tepsilon + \delta^2\right)},
+\end{align}
+where $\delta\equiv \Htilde/\Hgas$. 
+Since $|q|=O(1)$ in PPDs, Eq. \ref{vshear_split} indicates that
+vertical shear due to variations in the dust-to-gas ratio can be at
+most comparable to that due to the radial temperature gradient; and
+this only occurs for thin dust layers with $\delta^2\ll \tepsilon$ (so
+that $\mathcal{V}\sim 1/|q|$). Otherwise, vertical shear is
+predominantly due to $\p_rc_s^2$.
 

### 2016-07-28 code/press/basic.f90

@@ -11,7 +11,8 @@ real*8 function dlnrhog_dr(zhat)
   implicit none 
   real*8, intent(in) :: zhat 
  
-  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
+  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( &
+              zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
               + smalld*dgratio*smallh_g*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) )
 end function dlnrhog_dr
 
@@ -21,7 +22,8 @@ real*8 function d2lnrhog_dr2(zhat)
   implicit none 
   real*8, intent(in) :: zhat 
   
-  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
+  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 &
+                 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
                  -4d0*smalld*dgratio*smallh_g**2d0*delta**2d0*dlnHg_dlnr*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) &
                  -(smalld*smallh_g*delta)**2d0*dgratio*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) )
 end function d2lnrhog_dr2
@@ -151,7 +153,8 @@ real*8 function d2lnrho_dz2(zhat)
   real*8, intent(in) :: zhat 
   real*8, external  :: eps_tilde, deps_tilde_dz, d2eps_tilde_dz2, d2lnrhog_dz2
   
[…]

### 2016-07-28 code/press_loop/basic.f90

@@ -1,0 +1,646 @@
+real*8 function lnrhog(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+
+if(vstruct.eq.'mlin') lnrhog = -0.5d0*zhat**2d0 - dgratio*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) 
+if(vstruct.eq.'tl02') lnrhog  = -0.5d0*zhat**2d0
+end function lnrhog
+
+real*8 function dlnrhog_dr(zhat)
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+ 
+if(vstruct.eq.'mlin') then 
+  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
+              + smalld*dgratio*smallh_g*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) )
+endif
+
+if(vstruct.eq.'tl02')  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*zhat**2d0
+end function dlnrhog_dr
+
+real*8 function d2lnrhog_dr2(zhat)
+  !this is (1/r)*(d/dr)( r*dlnrhog_dr )
+  use global
+  implicit none 
+  real*8, intent(in) :: zhat 
+  
+if(vstruct.eq.'mlin') then
+  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
+                 -4d0*smalld*dgratio*smallh_g**2d0*delta**2d0*dlnHg_dlnr*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) &
+                 -(smalld*smallh_g*delta)**2d0*dgratio*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) )
+endif  
+
+if(vstruct.eq.'tl02') d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*zhat**2d0
+end function d2lnrhog_dr2
+
+real*8 function dlnrhog_dz(zhat)
+  use global
+  implicit none 
[…]

### 2016-07-28 code/press_loop/dvsi.f90

@@ -1,14 +1,16 @@
 module global
   implicit none 
+  character*4 :: vstruct 
   integer, parameter :: nvar = 5 
   real*8, parameter :: pi = 2d0*acos(0d0)
   complex*16, parameter :: ii = (0d0, 1d0), zero = (0d0, 0d0)
 
-  integer :: nz, bignz, Nzmid
-  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd
+  integer :: nz, bignz, Nzmid, nk, ndgr
+  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd, beta
   real*8 :: zmax 
-  real*8 :: dlnHg_dlnr
-  real*8 :: krad, bc_tol 
+  real*8 :: dlnHg_dlnr, dlogk, dlogdgr 
+  real*8 :: kmin, kmax, krad, dgrmin, dgrmax, bc_tol, growth, freq
+  real*8, allocatable :: kaxis(:), dgraxis(:), rates(:)
 
   real*8, allocatable :: zaxis(:), lnrho_arr(:), eps_arr(:), dlnrhodr_arr(:), depsdr_arr(:),dlnrhodz_arr(:), depsdz_arr(:) 
   real*8, allocatable :: del2lnrho_arr(:), del2eps_arr(:), dellnrho_dot_deleps_arr(:), tstop_arr(:), d2lnrhodz2_arr(:)
@@ -28,6 +30,8 @@ module global
   complex*16, allocatable :: R41(:,:), R42(:,:), R43(:,:), R44(:,:), R45(:,:) 
   complex*16, allocatable :: R51(:,:), R52(:,:), R53(:,:), R54(:,:), R55(:,:) 
   complex*16, allocatable :: bigW(:), dfrac(:), vx(:), vy(:), vz(:)
+  complex*16, allocatable :: err1(:), err2(:), err3(:), err4(:), err5(:)
+  complex*16, allocatable :: divv(:), dC(:)
 
   complex*16, allocatable :: bigmatrix_lhs(:,:)
   complex*16, allocatable :: bigmatrix_rhs(:,:)
@@ -38,16 +42,16 @@ end module global
 program dvsi
   use global
   implicit none 
-  integer :: i, j, k, lmax 
+  integer :: i, j, k, lmax, loc(1) 
   real*8 :: z, zbar, m, T_l, Tp_l, Tpp_l, dT_l, d2T_l
[…]

### 2016-07-29 model.tex
polishing up model section

@@
 Eq. \ref{eff_energy} is the energy equation for an ideal gas of adiabatic index
-$\gamma=1$, i.e. isothermal evolution, but now with additional source 
-terms due to the imposed temperature profile, as well as dust-gas
-friction.
+$\gamma=1$, i.e. isothermal evolution, but now with additional energy source 
+terms due to the imposed temperature profile via $\nabla c_s^2$, as
+well as dust-gas drag via $\mathcal{C}$. 
+%In this form, the function $\mathcal{C}$ can be interpreted
+%as a cooling term.
 
@@
 energy flux has an additional contribution from the pressure
-gradient. The term on the right-hand-side, owing to the imposed
+gradient and dust-gas friction. 
+The term on the right-hand-side, owing to the imposed
 temperature profile, can be interpreted as an external heat source.
 
-\section{Equilibrium}\label{eqm}
-To obtain an axisymmetric steady state with $\rho(r,z)$ and 
-$\bm{v}=r\Omega(r,z)\hat{\bm{\phi}}$ where $\Omega = v_\phi/r$, 
-we need to solve 
+\section{Dynamical equilibria}\label{eqm}
+
+For a given distribution of the dust-fraction $\epsilon$ (or
+dust-to-gas ratio $\tepsilon$), the 
+mass and momentum Eqs. \ref{masseq}---\ref{momeq} admit     
+axisymmetric steady states with $\rho(r,z)$ and 
+$\bm{v}=r\Omega(r,z)\hat{\bm{\phi}}$ where $\Omega = v_\phi/r$, which satisfy 
 \begin{align}
@@
     r},\label{steady_momr}\\
-  0 & = \frac{\p\Phi}{\p z} + \frac{1}{\rho}\frac{\p P}{\p z},\label{steady_momz}\\
-  0 & = \nabla\cdot\left(\epsilon\tstop\nabla P\right) \label{steady_dust}
+  0 & = \frac{\p\Phi}{\p z} + \frac{1}{\rho}\frac{\p P}{\p z},\label{steady_momz}
+%  0 & = \nabla\cdot\left(\epsilon\tstop\nabla P\right) \label{steady_dust}
 \end{align}
-for $\rho$, $\Omega$ and $\epsilon$ with $P=P(\epsilon,\rho)$ given by
-the equation of state (Eq. \ref{eos}).
+with $P=P(\epsilon,\rho)$ given by the equation of state
+(Eq. \ref{eos}). An explicit solution is presented in
+\S\ref{steady_state}.  
+However, disk structures obtained this way generally do not satisfy 
+the steady state effective energy Eq. \ref{eff_energy} 
+except in the limit of perfectly coupled dust ($\tstop\to0$).
 
-Eq. \ref{steady_dust} makes it difficult to obtain equilibrium
-solutions explicitly. However, if the diffusive process is slow and
-can be neglected, then one may just solve
-Eq. \ref{steady_momr}---\ref{steady_momz} with a prescribed (initial)
-distribution of the dust fraction $\epsilon(r,z)$.
+True steady states may be calculated by invoking an underlying gas
+turbulence, which leads to dust diffusion \citep{takeuchi02, youdin07, 
+  lyra13}. This complication is beyond the scope of this paper and we 
+defer it to a future work. Here, we will either consider $\tstop = 0$,
+or assume that $\tstop$ is sufficiently small such that the background
+disk (as obtained from Eq. \ref{steady_momr}---\ref{steady_momz}) does 
+not evolve significantly.  %relative to perturbations
+
+%Eq. \ref{steady_dust} makes it difficult to obtain equilibrium
+%solutions explicitly. However, if the diffusive process is slow and
+%can be neglected, then one may just solve
+%Eq. \ref{steady_momr}---\ref{steady_momz} with a prescribed (initial)
+%distribution of the dust fraction $\epsilon(r,z)$.
 
 \subsection{Disk structure with a prescribed dust distribution}\label{steady_state}  
-We assume a Gaussian profile in the dust-to-gas ratio, 
+To obtain an actual disk structure for numerical computations, we
+assume a Gaussian profile in the dust-to-gas ratio,   
 \begin{align}\label{dust_gauss}
@@
 \begin{align}
-  \rhog(r,z) = \rho_\mathrm{g0}(r)\exp{\left\{ - \frac{z^2}{2\Hgas^2}
+  &\rhog(r,z)= \notag\\
+&\rho_\mathrm{g0}(r)\exp{\left\{ - \frac{z^2}{2\Hgas^2}
     -\tepsilon_0\frac{\Htilde^2}{\Hgas^2}\left[1 -
@@
 Writing Eq. \ref{vshear} in terms of the gas density and dust-to-gas
-ratio with a power-law temperature profil (Eq. \ref{power_temp}) e gives 
+ratio with a power-law temperature profile (Eq. \ref{power_temp}) gives 
 \begin{align*}
@@
 
+In this section we review the basic properties of the 
+single-fluid mixture using Eqs.\ref{masseq}--\ref{momeq} with
+Eq. \ref{eff_energy} in place of the dust continuity equation. Our
+discussion is based on the corresponding linearized equations about an
+axisymmetric steady state. These lead to the integral relation,
+
+\begin{align}
+&  \sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV \notag\\
+&= \int\left[ \rho
+  |\dd v_r|^2A + \rho  \dd v_z \dd v_r^* B + \rho \dd v_z^*\dd v_r C +
+  \rho |\dd v_z|^2 D\phantom{\frac{1}{1}}\right. \notag\\
+&\phantom{===}  \left. + \frac{1}{P}\left|\nabla\cdot\left(P\dd
+  \bm{v}\right)\right|^2\right]dV  -\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV \notag\\
+&\phantom{===}
+- \int P
+  \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV,\label{int_rel}
+\end{align}
+between the (complex) mode frequency $\sigma$ and the spatial
+structure of the disturbances which are prefixed by deltas. 
+The real coefficients $A,B,C,D$ can be read off
+Eq. \ref{integral_ex}. Note that $B=C$ from dynamical equilibrium
+(Eq. \ref{vshear}).
+
 \subsection{Strictly isothermal gas perfectly coupled to dust}   
-When $c_s^2$ is a constant and $\tstop=0$, the dusty gas equations are
+When $c_s^2$ is a constant and $\tstop=0$, the dusty-gas equations are
 exactly equivalent to that for adiabatic hydrodyamics with unit adiabatic
@@
 adiabatically because the dust fraction $\epsilon$ is advected with 
-the gas (i.e. evolves like an entropy when there is no heating or
-cooling).
+the gas. This is similar to entropy being conserved following an
+adiabatic gas when there is no heating or cooling.
 
-In this case one may immediately apply the Solberg-Hoiland
-criteria to assess axisymmtric \emph{stability}:
+In this case, the last two integrals in Eq. \ref{int_rel} vanish. Then 
+requiring $\sigma^2>0$  gives the Solberg-Hoiland criteria 
+for the axisymmtric \emph{stability} of an isothermal gas perfectly 
+coupled to dust,  
+\begin{align}
+  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\epsilon &> 0,\label{dusty_solberg1}  \\
+  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\epsilon}{\p
+    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \epsilon}{\p r} \right) & > 0, \label{dusty_solberg2}
+\end{align}
+%\begin{align}
+%  \kappa^2 - \frac{1}{\rho}\nabla P \cdot \nabla s &> 0,\\
+%  -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
+%  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, 
+%\end{align} 
+where %$s = \ln{P/\rho}$, 
+$\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)$ is the
+square of the epicylic frequency, and we assume $\kappa^2>0$. 
+Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} are equivalent to the
+Solberg-Hoiland criteria for adiabatic gas if one writes the 
+entropy $s=\ln{\left[c_s^2\left(1 - \epsilon\right)\right]}$.
 
-\begin{align}
-  \kappa^2 - \frac{1}{\rho}\nabla P \cdot \nabla s &> 0,\\
-  -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
-  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, 
-\end{align}
-with $s = \ln{P/\rho^\gamma}$ and 
-$\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)$ is the
-square of the epicylic frequency. We assume $\kappa^2>0$. 
-These criteria are applicable to any adiabatic fluid. For 
-isothermal gas perfectly coupled to dust, we may take 
-$s=\ln{(1-\epsilon)}$. 
+%These criteria are applicable to any adiabatic fluid. 
+%For an isothermal mixture, we may take $s=\ln{(1-\epsilon)}$.   
 %from the definition of $s$ and  
@@
 %Eq. \ref{eos}, with a constant sound-speed. 
-The stability conditions are thus
-\begin{align}
-  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\epsilon &> 0,\label{dusty_solberg1}  \\
-  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\epsilon}{\p
-    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \epsilon}{\p r} \right) & > 0. \label{dusty_solberg2}
-\end{align}
-
+%The stability conditions are thus
 %In accretion disks, $\rhog$, $\rhod$ and $\epsilon$ are expected to
@@
 distributions for the gas density and the dust-to-gas ratio as 
-described in \S\ref{eqm}. However, here we permit 
-$\tepsilon_0=\tepsilon_0(r)$ and $\Htilde=\Htilde(r)$ to vary with
-radius. We assume the disk is approximately Keplerian so that
+described in \S\ref{eqm}. 
+%However, here we permit 
+%$\tepsilon_0=\tepsilon_0(r)$ and $\Htilde=\Htilde(r)$ to vary with
+%radius. 
+We assume the disk is approximately Keplerian so that
 $\kappa\simeq\OmK$. In terms of the dust-to-gas
@@
 Eq. \ref{stability_est1}---\ref{stability_est2} are easily satisfied
-in thin accretion disks where radial gradients are $O(1/r)$ and
-$h_\mathrm{g},\tepsilon, \widetilde{H}/H_\mathrm{g}\ll 1$, since then the
+in thin accretion disks where radial gradients are $O(1/r)$, 
+$h_\mathrm{g},\tepsilon,\ll 1$ and $
+\widetilde{H}/H_\mathrm{g}\lesssim 1$. Then the
 magnitude of the second term on the left-hand-side  of either
@@
 \subsection{Locally isothermal gas perfectly coupled to dust} 
-When $c_s(r,z)$ is non-uniform but $\tstop=0$, simple stability
-criteria are not available. However, we may obtain an integral
-relation for linear, axisymmetric waves with frequency $\sigma$ (see
-Appendix \ref{var_prin}, where we also account for finite $\tstop$) in
-the form:   
+When $c_s(r,z)$ is non-uniform but $\tstop=0$, Eq. \ref{int_rel}
+indicates that $\sigma^2$ is generally complex, implying instability.  
+Writing $\sigma = 
+\omega + \ii s$ where $\omega$ and $s$ are the real frequency and
+growth rates, respectively, we have 
 \begin{align}
-&  \sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV \notag\\
-&= \int\left[ \rho
-  |\dd v_r|^2A + \rho  \dd v_z \dd v_r^* B + \rho \dd v_z^*\dd v_r C +
-  \rho |\dd v_z|^2 D\phantom{\frac{1}{1}}\right. \notag\\
-&\phantom{===}  \left. + \frac{1}{P}\left|\nabla\cdot\left(P\dd
-  \bm{v}\right)\right|^2\right]dV - \int P
-  \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV.
+  s = -\frac{\imag\int P
+  \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV}{2\omega\int\rho\left(|\dd
+    v_r|^2 + |\dd v_z|^2\right)dV}. 
 \end{align} 
-The real coefficients $A,B,C,D$ can be read off
-Eq. \ref{integral_ex}. Note that $B=C$ from dynamical equilibrium
-(Eq. \ref{vshear}).
+This is classic VSI caused by vertical shear arising from a radial
+temperature gradient \citep{nelson13,barker15,lin15}.
 
-If $c_s$ is constant, then $\sigma^2$ is real, and 
-the first integral leads to the previous Solberg-Hoiland criteria for 
-axisymmetric stability. However, when $c_s$ is non-uniform, $\sigma^2$ 
-is generally complex, implying instability. This is responsible for 
-the usual vertical shear instability in locally isothermal disks.
+%This is responsible for 
+%the usual vertical shear instability in locally isothermal disks.
+%However, we may obtain an integral
+%relation for linear, axisymmetric waves with frequency $\sigma$ (see
+%Appendix \ref{var_prin}, where we also account for finite $\tstop$) in
+%the form:   
+%If $c_s$ is constant, then $\sigma^2$ is real, and 
+%the first integral leads to the previous Solberg-Hoiland criteria for 
+%axisymmetric stability. However, when $c_s$ is non-uniform,
 
 \subsection{Strictly isothermal gas well-coupled to dust} 
-Now consider $c_s^2=\mathrm{const.}$ but $\tstop\neq0$. In this case 
-the effective energy equation is 
-\begin{align}\label{dusty_cooling}
-  \frac{DP}{Dt} &= - P \nabla \cdot \bm{v} + \mathcal{C}
+Now consider constant $c_s$ but $\tstop\neq0$. Then Eq. \ref{int_rel}
+gives 
+\begin{align}
+  s = -\frac{\imag\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV}{2\omega\int\rho\left(|\dd
+    v_r|^2 + |\dd v_z|^2\right)dV}, 
 \end{align}
-The function $\mathcal{C}$ can be considered as a cooling term. By
-analogy with enabling VSI in an adiabatic gas through cooling
-\citep{nelson13,lin15}, we may expect a `dusty VSI' to develop 
-if this term is large enough (physically meaning diffusion of the
-dust-fraction is strong).
-
-Note that $\mathcal{C}$ is also responsible for the streaming
-instability  in the strong-drag limit \citep{laibe14}, though this
-does not require vertical stratification. In the form of
-Eq. \ref{dusty_cooling}, however, the streaming instability has a
-thermodynamical interpretation, similar to thermal instabilities in
-stars \citep{latter06}.
+%In this case 
+%the effective energy equation is 
+%\begin{align}\label{dusty_cooling}
+%  \frac{DP}{Dt} &= - P \nabla \cdot \bm{v} + \mathcal{C}
+%\end{align} 
+%The function $\mathcal{C}$ can be considered as a cooling term. 
+%By
+%analogy with enabling VSI in an adiabatic gas through cooling
+%\citep{nelson13,lin15}, we may expect a `dusty VSI' to develop 
+%if this term is large enough (physically meaning diffusion of the
+%dust-fraction is strong). 
+%Note that $\mathcal{C}$ is also responsible for the streaming
+%instability  in the strong-drag limit \citep{laibe14}, though this
+%does not require vertical stratification. In the form of
+%Eq. \ref{dusty_cooling}, however, the streaming instability has a
+%thermodynamical interpretation, similar to thermal instabilities in
+%stars \citep{latter06}.  
+which represent all instabilities due to large-but-finite dust-gas
+friction. The formalism we have adopted give these instabilities a 
+thermodynamical interpretation \citep[thermal instabilities in
+  stars][]{latter06} because the dust-gas friction enters the problem
+as an energy source term. This includes the streaming instability in
+the strong-drag limit \citep{youdin05a, jacquet11, laibe14}.

### 2016-07-29 model.tex
polishing up model section

@@
 friction. The formalism we have adopted give these instabilities a 
-thermodynamical interpretation \citep[thermal instabilities in
-  stars][]{latter06} because the dust-gas friction enters the problem
-as an energy source term. This includes the streaming instability in
+thermodynamical interpretation because the dust-gas friction enters the problem
+as an energy source term \citep[cf. thermal instabilities in
+  stars][]{latter06}. 
+This includes the streaming instability in
 the strong-drag limit \citep{youdin05a, jacquet11, laibe14}.

### 2016-07-30 linear.tex
adding results

--- 2016-05-06 linear.tex
@@
-\section{Linear problem} 
-We consider radially-localized axisymmetric disturbances of the form 
+\section{Linear problem for perfectly coupled dust}\label{linear_problem}
+In this paper we calculate explicit solutions to the linearized
+equations in the limit of perfectly coupled dust ($\tstop=0$), so that
+the equilibria defined in \S\ref{eqm} are exact steady states; and we
+consider radially-localized axisymmetric disturbances of the form  
 \begin{align}
-  \delta\rho(r, z)\exp{\left(\ii k r-\ii\sigma t\right)},
+  \delta f (r, z) = \delta f_1(r,z)\exp{(\ii k r)},
 \end{align} 
-and similarly for $\dd P$ and $\dd\bm{v}$. Here, $k$ is a real
-wavenumber such that $|kr|\gg 1$, and the amplitude $\dd\rho(r,z)$ is
+%and similarly for $\dd P$ and $\dd\bm{v}$. 
+where $k$ is a real wavenumber such that $|kr|\gg 1$, and the
+amplitude $\dd f_1(r,z)$ is 
 a slowly-varying function of $r$. Then 
 $\p_r\to i k$ when acting on the above primitive perturbations, and we may
-neglect curvature terms. 
-We take  
-$k>0$ without loss of  generality. The frequency $\sigma = \omega +
-\ii s$ is generally complex, with $\omega$ being the real frequency
-and $s$ is the real growth rate.
+neglect curvature terms. We take  
+$k>0$ without loss of generality. Hereafter, we drop the subscript 1
+on the amplitudes. 
+%The frequency $\sigma = \omega +
+%\ii s$ is generally complex, with $\omega$ being the real frequency
+%and $s$ is the real growth rate.
 
@@
 the linearized equations for 
-locally isothermal dusty gas with the pressure equation in place of
-the dust-fraction (Eq. \ref{masseq}---\ref{momeq}, Eq. \ref{eff_energy}) are then:
+locally isothermal, perfectly-coupled dusty gas with the pressure
+equation in place of the dust-fraction
+(Eq. \ref{masseq}---\ref{momeq}, Eq. \ref{eff_energy}) are then:
 
@@
                v_z^\prime\right) + \frac{1}{\rho}\left(\dd v_r\p_rP + \dd v_z \p_zP\right)\notag\\
-                &\phantom{=}-\frac{P}{\rho} \dd v_r\p_r \ln{c_s^2} - \frac{\dd\mathcal{C}}{\rho},\label{lin_energy}
+                &\phantom{=}-\frac{P}{\rho} \dd v_r\p_r
+               \ln{c_s^2}, \label{lin_energy} 
+%+ \dd v_z \p_z\ln{c_s^2}\right),\label{lin_energy} 
+%               - \frac{\dd\mathcal{C}}{\rho},\label{lin_energy} 
 \end{align}  
-where $^\prime \equiv \p_z$; $\dd\bm{F}$ is the linearized pressure force, given in Appendix
-\ref{lin_press}; and 
+where $^\prime \equiv \p_z$ and $\dd\bm{F}$ is the linearized pressure
+force, given in Appendix \ref{lin_press}. Note that we have assumed a
+temperature profile that only depends on $r$.
+
+%; and 
 %\begin{align}
@@
 %\end{align}
-$\dd\mathcal{C}$ is the linearized dust-duffusion function, given in
-Appendix \ref{lin_dust}.  We consider stopping times appropriate for
-small grains in the Epstein regime. Note that for the axisymmetric
-problem, $\dd\bm{F}$ is purely meridional.
+%$\dd\mathcal{C}$ is the linearized dust-duffusion function, given in
+%Appendix \ref{lin_dust}. We consider stopping times appropriate for
+%small grains in the Epstein regime. Note that for the axisymmetric
+%problem, $\dd\bm{F}$ is purely meridional.
 
 Eq. \ref{lin_mass}---\ref{lin_energy} is a set of ordinary
-differntial equations in $z$. All the coefficients are evaluated at a
-fiducial radius $r=r_0$, but their full $z$-dependence is retained.  
-\subsection{Boundary conditions}
-Eq. \ref{lin_mass}---\ref{lin_energy} can be reduced to a set of
-first-order differential equations for $W, Q$ and $\dd v_z$. We can
-see this schematically as follows. Eq. \ref{lin_xmom} and
-\ref{lin_ymom} may be combined to yield 
-\begin{align*}
-  \dd v_r = \dd v_r (W, Q, \dd v_z). 
-\end{align*} 
-We can then take the continuity equation as an equation for $\dd v_z$, 
-\begin{align*}
-\text{Eq. \ref{lin_mass}} \Rightarrow \dd v_z^\prime(z) = \dd
-  v_z^\prime(W,Q,\dd v_z),
-\end{align*}
-and the vertical momentum equation as an equation for $Q$, 
-\begin{align*}
-\text{Eq. \ref{lin_zmom}} \Rightarrow Q^\prime(z) = 
- Q^\prime(W,Q,\dd v_z). 
-\end{align*}
+differntial equations in $z$. All coefficients and amplitudes are
+evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
+is retained.  Two boundary conditions are needed for the $\tstop=0$
+problem. This is expected since this problem is exactly equivalent to
+that in adiabatic hydrodynamics \citep[e.g.][]{lubow93}. For simplicity
+we impose solid boundaries so that $\delta v_z(\pm\zmax)=0$ and
+$\zmax=5\Hgas$.
 
-Now, for finite dust-gas coupling, $\tstop\neq0$, inspection of $\dd C$
-(Appendix \ref{lin_dust}) shows that it involves $W^\prime$ (assuming
-$\epsilon < 1$). Then Eq. \ref{lin_energy} may be taken as a
-differential equation for $W$. However, for perfectly coupled dust,
-$\tstop =\dd\mathcal{C}= 0$. In that case Eq. \ref{lin_energy} gives an algebraic
-relation between $W$, $Q$ and $\dd v_z$. That is,
-\begin{align*}
-\text{Eq. \ref{lin_energy}}\Rightarrow \begin{cases}
-  W^\prime(z) = W^\prime(W, Q, \dd v_z) & \text{if } \tstop \neq 0, \\
-  W           (z) = W(Q, \dd v_z) & \text{if }\tstop = 0.
-\end{cases}
-\end{align*}
+We solve the linearized equations as a generalized eigenvalue problem 
+using a pseudo-spectral code adapted from \cite{lin15}. Amplitudes 
+are expanded in Chebyshev polynomials up to order $N_z-1$. Our 
+standard resolution is $N_z=513$. We check results using
+Eq. \ref{vsi_check}.
 
-This means that for the perfectly coupled problem, we have a pair of
-ODEs for $Q$ and $\dd v_z$, and two boundary conditions are needed. 
-This is expected since this problem is exactly 
-equivalent to that in adiabatic hydrodynamics \citep[e.g.]{lubow93}. When
-$\tstop\neq0$, we require three boundary conditions. {\bf WHAT?? Why odd
-  number of BCs? Shouldn't we have even number of BCs? We have
-two boundaries? I've only seen problems like this involving even
-number of BCs.} In this case we impose $\dd v_z =0$ at $z=\pm
-z_\mathrm{max}$, and consider even modes about the midplane such that
-$\dd v_z^\prime(0) = 0$.
+%\subsection{Boundary conditions}
+
+%Eq. \ref{lin_mass}---\ref{lin_energy} can be reduced to a set of
+%first-order differential equations for $W, Q$ and $\dd v_z$. We can
+%see this schematically as follows. Eq. \ref{lin_xmom} and
+%\ref{lin_ymom} may be combined to yield 
+%\begin{align*}
+%  \dd v_r = \dd v_r (W, Q, \dd v_z). 
+%\end{align*} 
+%We can then take the continuity equation as an equation for $\dd v_z$, 
+%\begin{align*}
+%\text{Eq. \ref{lin_mass}} \Rightarrow \dd v_z^\prime(z) = \dd
+%  v_z^\prime(W,Q,\dd v_z),
+%\end{align*}
+%and the vertical momentum equation as an equation for $Q$, 
+%\begin{align*}
+%\text{Eq. \ref{lin_zmom}} \Rightarrow Q^\prime(z) = 
+% Q^\prime(W,Q,\dd v_z). 
+%\end{align*}
+
+%Now, for finite dust-gas coupling, $\tstop\neq0$, inspection of $\dd C$
+%(Appendix \ref{lin_dust}) shows that it involves $W^\prime$ (assuming
+%$\epsilon < 1$). Then Eq. \ref{lin_energy} may be taken as a
+%differential equation for $W$. However, for perfectly coupled dust,
+%$\tstop =\dd\mathcal{C}= 0$. In that case Eq. \ref{lin_energy} gives an algebraic
+%relation between $W$, $Q$ and $\dd v_z$. That is,
+%\begin{align*}
+%\text{Eq. \ref{lin_energy}}\Rightarrow \begin{cases}
+%  W^\prime(z) = W^\prime(W, Q, \dd v_z) & \text{if } \tstop \neq 0, \\
+%  W           (z) = W(Q, \dd v_z) & \text{if }\tstop = 0.
+%\end{cases}
+%\end{align*}
+
+%This means that for the perfectly coupled problem, we have a pair of
+%ODEs for $Q$ and $\dd v_z$, and 
+%When
+%$\tstop\neq0$, we require three boundary conditions. {\bf WHAT?? Why odd
+%  number of BCs? Shouldn't we have even number of BCs? We have
+%two boundaries? I've only seen problems like this involving even
+%number of BCs.} In this case we impose $\dd v_z =0$ at $z=\pm
+%z_\mathrm{max}$, and classify modes according to their symmetry about
+%the midplane: `even' modes with $\dd v_z^\prime(0) = 0$, and `odd' 
+%modes with $\dd v_z(0)=0$.

### 2016-07-30 model.tex
adding results

@@
 Eq. \ref{eff_energy} in place of the dust continuity equation. Our
-discussion is based on the corresponding linearized equations about an
-axisymmetric steady state. These lead to the integral relation,
-
+discussion is based on the corresponding axisymmetric linearized
+equations about an axisymmetric steady state. A variable $f$ is
+subject to perturbations of the form 
+\begin{align}
+  \delta f(r,z)\exp{\left(-\ii\sigma t\right)},
+\end{align}
+where $\sigma$ is the complex mode frequency. In Appendix
+\ref{var_prin} we derive the integral relation,  
 \begin{align}
@@
 - \int P
-  \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV,\label{int_rel}
+  \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV.\label{int_rel}
 \end{align} 
-between the (complex) mode frequency $\sigma$ and the spatial
-structure of the disturbances which are prefixed by deltas. 
 The real coefficients $A,B,C,D$ can be read off 
 Eq. \ref{integral_ex}. Note that $B=C$ from dynamical equilibrium
-(Eq. \ref{vshear}).
+(Eq. \ref{vshear}). We now consider various limits of
+Eq. \ref{int_rel}.
 
@@
 
-\subsection{Locally isothermal gas perfectly coupled to dust} 
+\subsection{Locally isothermal gas perfectly coupled to dust}\label{dusty_vsi_int}
 When $c_s(r,z)$ is non-uniform but $\tstop=0$, Eq. \ref{int_rel}
@@
 Writing $\sigma = 
-\omega + \ii s$ where $\omega$ and $s$ are the real frequency and
+-\omega + \ii s$ where $\omega$ and $s$ are the real frequency and
 growth rates, respectively, we have 
 \begin{align}
-  s = -\frac{\imag\int P
+  s = \frac{\imag\int P
   \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV}{2\omega\int\rho\left(|\dd
-    v_r|^2 + |\dd v_z|^2\right)dV}. 
+    v_r|^2 + |\dd v_z|^2\right)dV}. \label{vsi_check}
 \end{align} 
 This is classic VSI caused by vertical shear arising from a radial
-temperature gradient \citep{nelson13,barker15,lin15}.
+temperature gradient \citep{nelson13,barker15,lin15}. We compute
+explicit solutions in this limit in
+\S\ref{linear_problem}---\ref{results}.
 
@@
 \begin{align}
-  s = -\frac{\imag\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV}{2\omega\int\rho\left(|\dd
+  s = \frac{\imag\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV}{2\omega\int\rho\left(|\dd
     v_r|^2 + |\dd v_z|^2\right)dV}, 

### 2016-07-30 result.tex
adding results

--- 2016-05-04 result.tex
@@
-\section{Results}
+\section{Perfectly coupled
+  dusty vertical shear instability}\label{results}
+
+When $\tstop=0$ we may construct exact dusty equilibra by specifying
+the dust-to-gas ratio, as described in \S\ref{eqm}. We use this to
+examine the effect of dust on classic VSI, defined to be driven by the
+temperature gradient (\S\ref{dusty_vsi_int}). For simplicity we
+consider constant input values of the midplane dust-to-gas ratio
+$\tepsilon_0$ and the characteristic dust thickness $\Hd$. These,
+together with the perturbation wavenumber $k$, comprises main the
+parameters of the linear problem. We fix the radial power-law index
+for the midplane gas density to $p = -1.5$, that for the
+temperature profile to $q=-1$; and set the gas disk aspect-ratio
+$h_\mathrm{g}=0.05$. These are fiducial values used in \citetalias{lin15}.
+
+\subsection{Effect of dust-loading}
+We first vary the dust-to-gas ratio $\tepsilon_0\in[10^{-3},1]$,
+fixing the dust thickness to $\Hd=0.99\Hg$ (so that $\tepsilon$ is
+roughly constant with height) and perturbation
+wavenumber to $k\Hg = 30$.
+
+Fig. \ref{vsi_dust_loading} show unstable
+modes found for different values of $\tepsilon_0$. The eigenvalue
+distributions for $\tepsilon_0 \leq 10^{-2}$ are similar to the
+dust-free fiducial case considered by \citetalias{lin15}, consisting
+of the roughly horizontal `body modes' and the nearly-vertical
+`surface modes'. The latter is due to the imposed vertical boundaries
+\citep{barker15}.
+
+We find that increasing the dust-to-gas ratio reduce VSI growth
+rates. Notably `surface modes', which are typically fastest growing in
+the dust-free case, are surpressed in dusty disks for $\tepsilon_0\geq
+0.1$. The body modes' growth rates remain $\sim O(h\OmK)$ but their
+oscillation frequency increases with $\tepsilon_0$.
+
+In Fig. \ref{vsi_dust_loading2d} we compare the
 
 \begin{figure}
-\includegraphics[width=\linewidth]{figures/eigenvalues}
-\caption{do not believe}
+  \includegraphics[width=\linewidth]{figures/compare_eigenvals_kx30Hd1} 
+  \caption{Unstable modes in a locally isothermal, perfectly coupled
+    dusty disk with fiducial parameters
+    $(p,q,h_\mathrm{g}, \Hd/\Hg )=(-1.5,-1,0.05, 0.99)$. The real
+    frequency $\omega$ and growth rates $s$ are shown for a range of
+    midplane dust-to-gas ratios $\tepsilon_0=\rhod/\rhog$. 
+    \label{vsi_dust_loading}
+    }
 \end{figure}
+
+\begin{figure}
+  \includegraphics[scale=0.54, clip=true, trim=0cm 2.5cm 0cm 0cm]{figures/result2d_dg1d-3.ps}\\
+  \includegraphics[scale=0.54]{figures/result2d_dg1.ps} 
+  \caption{Fundamental dusty VSI mode in real space for midplane dust-to-gas
+    ratio $\tepsilon_0=10^{-3}$ (top) and $\tepsilon_0=1$
+    (bottom). The color scale shows the perturbation to the
+    dust-to-gas ratio, $\delta\tepsilon$; and the arrows show
+    $\sqrt{\rho}\left(\dd v_x, \dd v_z\right)$. 
+    \label{vsi_dust_loading2d}
+    }
+\end{figure}
+
+\subsection{Effect of dust thickness}

### 2016-08-01 appendix.tex
general edits
regular update

@@
 \begin{align}
-  \mathcal{C} \equiv c_s^2\nabla\cdot\left(\epsilon\tstop\nabla
+  \mathcal{C} \equiv c_s^2\nabla\cdot\left(\tepsilon\tstop\nabla
   P\right) = - K c_s^2 \nabla\cdot
-  \left(\frac{\epsilon}{c_s}\bm{F}\right) =
-  -Kc_s \left(\bm{F}\cdot\nabla\epsilon + \epsilon \nabla\cdot\bm{F}
-  - \frac{1}{2}\epsilon\bm{F}\cdot\nabla\ln{c_s^2}\right).  
+  \left(\frac{\tepsilon}{c_s}\bm{F}\right) =
+  -Kc_s \left(\bm{F}\cdot\nabla\tepsilon + \tepsilon \nabla\cdot\bm{F}
+  - \frac{1}{2}\tepsilon\bm{F}\cdot\nabla\ln{c_s^2}\right).  
 \end{align}
@@
 \begin{align}
-  - \frac{\dd\mathcal{C}}{Kc_s} = \bm{F}\cdot\nabla\dd\epsilon +
-  \dd\bm{F}\cdot\nabla\epsilon + \dd\epsilon\nabla\cdot\bm{F} +
-  \epsilon\nabla\cdot\dd\bm{F} -
-  \frac{1}{2}\nabla\ln{c_s^2}\cdot\left(\bm{F}\dd\epsilon + \epsilon
+  - \frac{\dd\mathcal{C}}{Kc_s} = \bm{F}\cdot\nabla\dd\tepsilon +
+  \dd\bm{F}\cdot\nabla\tepsilon + \dd\tepsilon\nabla\cdot\bm{F} +
+  \tepsilon\nabla\cdot\dd\bm{F} -
+  \frac{1}{2}\nabla\ln{c_s^2}\cdot\left(\bm{F}\dd\tepsilon + \tepsilon
   \dd\bm{F}\right). 
 \end{align}
-The linearized dust-fraction $\dd\epsilon$ and its derivatives in the
+The linearized dust-fraction $\dd\tepsilon$ and its derivatives in the
 radially-local approximation are given by
 \begin{align}
-  \dd\epsilon      &= (1-\epsilon)W - \frac{Q}{c_s^2}, \\
-  \p_r\dd\epsilon &= \left[(1-\epsilon)\left(\ii k -
-    \p_r\ln{\rho}\right) - \p_r\epsilon\right]W + \left(\p_r\ln{c_s^2}
+  \dd\tepsilon      &= (1-\tepsilon)W - \frac{Q}{c_s^2}, \\
+  \p_r\dd\tepsilon &= \left[(1-\tepsilon)\left(\ii k -
+    \p_r\ln{\rho}\right) - \p_r\tepsilon\right]W + \left(\p_r\ln{c_s^2}
   + \p_r\ln{\rho} - \ii k \right)\frac{Q}{c_s^2},\\
-  \dd\epsilon^\prime &= (1-\epsilon)W^\prime - \epsilon^\prime W -
+  \dd\tepsilon^\prime &= (1-\tepsilon)W^\prime - \tepsilon^\prime W -
   \left(\frac{Q}{c_s^2}\right)^\prime. 
@@
 \section{Conversion formulae}
-The gas density $\rhog$ and dust-to-gas ratio $\tepsilon$ is related
-to the total density $\rho$ and dust-fraction $\epsilon$ by
+The gas density $\rhog$ and dust-to-gas ratio $\epsilon$ is related
+to the total density $\rho$ and dust-fraction $\tepsilon$ by
 \begin{align}
-  \ln{\rho} &= \ln{\rhog} + \ln{\left(1 + \tepsilon\right)},\\
-  \nabla\ln{\rho} &= \nabla\ln{\rhog} + \frac{\nabla\tepsilon}{1+
-    \tepsilon},\\
+  \ln{\rho} &= \ln{\rhog} + \ln{\left(1 + \epsilon\right)},\\
+  \nabla\ln{\rho} &= \nabla\ln{\rhog} + \frac{\nabla\epsilon}{1+
+    \epsilon},\\
   \nabla^2\ln{\rho} &= \nabla^2\ln{\rhog} +
-  \frac{\nabla^2\tepsilon}{1+\tepsilon} -
-  \frac{\left|\nabla\tepsilon\right|^2}{\left(1+\tepsilon\right)^2}, 
+  \frac{\nabla^2\epsilon}{1+\epsilon} -
+  \frac{\left|\nabla\epsilon\right|^2}{\left(1+\epsilon\right)^2}, 
 \end{align}
@@
 \begin{align}
-  \epsilon &= \frac{\tepsilon}{1+ \tepsilon},\\
- \nabla\epsilon & =
- \frac{\nabla\tepsilon}{\left(1+\tepsilon\right)^2},\\
- \nabla^2\epsilon &=
- \frac{\nabla^2\tepsilon}{\left(1+\tepsilon\right)^2} -
- \frac{2\left|\nabla\tepsilon\right|^2}{\left(1+\tepsilon\right)^3}. 
+  \tepsilon &= \frac{\epsilon}{1+ \epsilon},\\
+ \nabla\tepsilon & =
+ \frac{\nabla\epsilon}{\left(1+\epsilon\right)^2},\\
+ \nabla^2\tepsilon &=
+ \frac{\nabla^2\epsilon}{\left(1+\epsilon\right)^2} -
+ \frac{2\left|\nabla\epsilon\right|^2}{\left(1+\epsilon\right)^3}. 
 \end{align}

### 2016-08-01 linear.tex
regular update

@@
 \section{Linear problem for perfectly coupled dust}\label{linear_problem}
-In this paper we calculate explicit solutions to the linearized
+We calculate explicit solutions to the linearized
 equations in the limit of perfectly coupled dust ($\tstop=0$), so that
-the equilibria defined in \S\ref{eqm} are exact steady states; and we
+the equilibria defined in \S\ref{eqm} are exact steady states. We 
 consider radially-localized axisymmetric disturbances of the form  
@@
 problem. This is expected since this problem is exactly equivalent to
-that in adiabatic hydrodynamics \citep[e.g.][]{lubow93}. For simplicity
+ adiabatic hydrodynamics \citep[e.g.][]{lubow93}. For simplicity
 we impose solid boundaries so that $\delta v_z(\pm\zmax)=0$ and
@@
 %(Appendix \ref{lin_dust}) shows that it involves $W^\prime$ (assuming
-%$\epsilon < 1$). Then Eq. \ref{lin_energy} may be taken as a
+%$\tepsilon < 1$). Then Eq. \ref{lin_energy} may be taken as a
 %differential equation for $W$. However, for perfectly coupled dust,

### 2016-08-01 model.tex
regular update

@@
-\section{Single fluid description of well-coupled dusty gas}\label{setup} 
+\section{Single fluid description of dusty gas}\label{setup} 
 In the limit of strong dust-gas drag (stopping time $\tstop\to 0$) 
@@
   P, \label{momeq}\\ 
-  &\frac{D\epsilon}{Dt} = -\frac{1}{\rho} \nabla \cdot \left(\epsilon 
+  &\frac{D\tepsilon}{Dt} = -\frac{1}{\rho} \nabla \cdot \left(\tepsilon 
   \tstop \nabla P \right),\label{dusteq}  
@@
 \begin{align}
-  \epsilon \equiv \frac{\rhod}{\rho}  = \frac{\tepsilon}{1+\tepsilon} 
+  \tepsilon \equiv \frac{\rhod}{\rho}  = \frac{\epsilon}{1+\epsilon} 
 \end{align}
-is the dust-fraction and $\tepsilon=\rhod/\rhog$ is the usual
+is the dust-fraction and $\epsilon=\rhod/\rhog$ is the usual
 dust-to-gas ratio. The gas pressure $P$ is given by the chosen 
@@
 \begin{align}\label{eos}
-  P = c_s^2(r,z)\rhog = c_s^{2}(1 - \epsilon)\rho,   
+  P = c_s^2(r,z)\rhog = c_s^{2}(1 - \tepsilon)\rho,   
 \end{align}
 where $c_s$ is a prescribed sound-speed profile fixed in time. 
-%For analytic discussion we consider arbitrary functions of
-%$c_s(r,z)$. 
 In numerical calculations we will consider vertically 
-isothermal disks with 
-\begin{align}\label{power_temp}
+isothermal disks with \begin{align}\label{power_temp}
   c_s^2(r) \propto r^{q},
@@
 
+If we define the effective sound-speed $c_{s,\mathrm{eff}} = c_s\sqrt{\left(1 -
+    \tepsilon\right)}$, then the equation of state has the standard
+form $P=c_{s,\mathrm{eff}}^2\rho$. Since the dust-fraction $\tepsilon$
+typically decrease away from the midplane, we expect vertically
+isothermal dusty disks to in fact behave as if it
+had a temperture increasing away from $z=0$.
+
 \subsection{Effective energy equation}
@@
 obeys an evolutionary energy equation, due to the advection of the
-dust-fraction. Eliminate $\epsilon$ in favor of $P$, 
+dust-fraction. Eliminate $\tepsilon$ in favor of $P$, 
 \begin{align*}
-  \epsilon = 1 - \frac{P}{c_s^2(r,z)\rho}, 
+  \tepsilon = 1 - \frac{P}{c_s^2(r,z)\rho}, 
 \end{align*}
@@
 
-For a given distribution of the dust-fraction $\epsilon$ (or
-dust-to-gas ratio $\tepsilon$), the 
+For a given distribution of the dust-fraction $\tepsilon$ (or
+dust-to-gas ratio $\epsilon$), the 
 mass and momentum Eqs. \ref{masseq}---\ref{momeq} admit     
@@
   0 & = \frac{\p\Phi}{\p z} + \frac{1}{\rho}\frac{\p P}{\p z},\label{steady_momz}
-%  0 & = \nabla\cdot\left(\epsilon\tstop\nabla P\right) \label{steady_dust}
+%  0 & = \nabla\cdot\left(\tepsilon\tstop\nabla P\right) \label{steady_dust}
 \end{align}
-with $P=P(\epsilon,\rho)$ given by the equation of state
+with $P=P(\tepsilon,\rho)$ given by the equation of state
 (Eq. \ref{eos}). An explicit solution is presented in
@@
 %Eq. \ref{steady_momr}---\ref{steady_momz} with a prescribed (initial)
-%distribution of the dust fraction $\epsilon(r,z)$.
+%distribution of the dust fraction $\tepsilon(r,z)$.
 
@@
 \begin{align}\label{dust_gauss}
-  \tepsilon(r,z) = \tepsilon_0(r)
-  \exp{\left[-\frac{z^2}{2\widetilde{H}^2(r)}\right]}. 
+  \epsilon(r,z) = \epsilon_0(r)
+  \exp{\left[-\frac{z^2}{2\Htilde^2(r)}\right]}. 
 \end{align}
 %For simplicity we assume the dust-to-gas ratio at the mid-plane
-%$\tepsilon_0$, as well as its characteristic scale-height
+%$\epsilon_0$, as well as its characteristic scale-height
 %$\widetilde{H}$, are both constant.
@@
 %\begin{align}
-%  \tepsilon_0(r) = \tepsilon_{00}\left(\frac{r}{r_0}\right)^{-d},  
+%  \epsilon_0(r) = \epsilon_{00}\left(\frac{r}{r_0}\right)^{-d},  
 %\end{align}
-%where $\tepsilon_{00}$ is the dust-to-gas ratio at the fiducial radius
+%where $\epsilon_{00}$ is the dust-to-gas ratio at the fiducial radius
 %$r_0$.  
@@
 &\rho_\mathrm{g0}(r)\exp{\left\{ - \frac{z^2}{2\Hgas^2}
-    -\tepsilon_0\frac{\Htilde^2}{\Hgas^2}\left[1 -
+    -\epsilon_0\frac{\Htilde^2}{\Hgas^2}\left[1 -
       \exp{\left(-\frac{z^2}{2\Htilde^2}\right)}\right] \right\}}, 
@@
 
-In practice, we consider $\tepsilon_0 \ll 1$ and settled dust such that
+In practice, we consider $\epsilon_0 \ll 1$ and settled dust such that
 $\Htilde<\Hgas$. Then $\rhog(r,z)$ is effectively Gaussian, as in the
@@
 \begin{align}
-  \rhod \simeq \tepsilon_0\rho_\mathrm{g0}(r) \exp
+  \rhod \simeq \epsilon_0\rho_\mathrm{g0}(r) \exp
         {\left(-\frac{z}{2H_\mathrm{d}^2}\right)}, 
@@
   \Omega(r,z) = \OmK(r)\left[1 - \frac{3}{2}\frac{z^2}{r^2} +
-    \frac{h_\mathrm{g}^2}{\left(1+\tepsilon\right)}\frac{\p}{\p\ln{r}}\ln{\left(c_s^2\rhog\right)}
+    \frac{h_\mathrm{g}^2}{\left(1+\epsilon\right)}\frac{\p}{\p\ln{r}}\ln{\left(c_s^2\rhog\right)}
     \right]^{1/2}, 
@@
 %&= \frac{\p\ln{\rho}}{\p r}\frac{\p}{\p
-%    z}\left[c_s^2(1-\epsilon)\right] - \frac{\p\ln{\rho}}{\p z}
-%  \frac{\p}{\p r} \left[c_s^2(1-\epsilon)\right]\\  
+%    z}\left[c_s^2(1-\tepsilon)\right] - \frac{\p\ln{\rho}}{\p z}
+%  \frac{\p}{\p r} \left[c_s^2(1-\tepsilon)\right]\\  
    = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p s}{\p z} -\frac{\p
@@
 \begin{align}
-   s \equiv \ln \frac{P}{\rho} = \ln{\left[c_s^2(1-\epsilon)\right]} 
+   s \equiv \ln \frac{P}{\rho} = \ln{\left[c_s^2(1-\tepsilon)\right]} 
 \end{align}
@@
   &r\frac{\p \Omega^2}{\p z}  =
-  \frac{c_s^2(r)}{\left(1+\tepsilon\right)^2}\left\{
-  \frac{\p\tepsilon}{\p r}\frac{\p\ln{\rhog}}{\p z}
-  -\frac{\p\tepsilon}{\p z}\frac{\p\ln{\rhog}}{\p r}\right.\notag\\
+  \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\left\{
+  \frac{\p\epsilon}{\p r}\frac{\p\ln{\rhog}}{\p z}
+  -\frac{\p\epsilon}{\p z}\frac{\p\ln{\rhog}}{\p r}\right.\notag\\
   &\phantom{ r\frac{\p \Omega^2}{\p z}  =
-    \frac{c_s^2(r)}{\left(1+\tepsilon\right)^2}\left\{\right\} }
-  \left. +\frac{q}{r}\left[\frac{\p\tepsilon}{\p z} + \left(1+\tepsilon\right)\frac{\p\ln{\rhog}}{\p z}\right]
+    \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\left\{\right\} }
+  \left. +\frac{q}{r}\left[\frac{\p\epsilon}{\p z} + \left(1+\epsilon\right)\frac{\p\ln{\rhog}}{\p z}\right]
   \right\} 
@@
 smooth so that $\p_r\sim 1/r$, and the dust-to-gas ratio
-$\tepsilon\ll1$. This gives
+$\epsilon\ll1$. This gives
 
@@
         temp. gradient}}} \sim
-  \tepsilon \frac{\mathrm{max}\left(\delta^2,
-    1\right)}{\left|q\right|\left(\tepsilon + \delta^2\right)},
+  \epsilon \frac{\mathrm{max}\left(\delta^2,
+    1\right)}{\left|q\right|\left(\epsilon + \delta^2\right)},
 \end{align}
@@
 most comparable to that due to the radial temperature gradient; and
-this only occurs for thin dust layers with $\delta^2\ll \tepsilon$ (so
+this only occurs for thin dust layers with $\delta^2\ll \epsilon$ (so
 that $\mathcal{V}\sim 1/|q|$). Otherwise, vertical shear is
 predominantly due to $\p_rc_s^2$.
+
+\subsection{Vertical buoyancy}
+Having identified the entropy of an isothermal dusty gas, we find the
+vertical buoyancy frequency $N_z$ of the mixture is given by 
+\begin{align}
+  N_z^2 \equiv - \frac{1}{\rho}\frac{\p P}{\p z}\frac{\p s}{\p z} &=
+  \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\frac{\p\ln\rhog}{\p 
+  z}\frac{\p\epsilon}{\p z} \\ &
+                                  =
+  \frac{\epsilon}{\left(1+\epsilon\right)^2}\left(\frac{z}{\Htilde}\right)^2\OmK^2\notag,  
+\end{align}
+where the second equality assumes $c_s=c_s(r)$ and the final equality
+applies to the equilibria in \S\ref{eqm}. We find 
+\begin{align*}
+N_z\lesssim
+O\left(\sqrt{\epsilon_0}\OmK\right). 
+\end{align*}
+Thus in PPDs where $\epsilon_0 = O(10^{-1}$---$10^{-2})$, 
+vertical buoyancy only has a modestly stabilizing effect.
+
+%\begin{align}
+ % N_z^2 
+%\end{align}
 
@@
 index. Although the gas is strictly isothermal, the mixture behaves 
-adiabatically because the dust fraction $\epsilon$ is advected with 
+adiabatically because the dust fraction $\tepsilon$ is advected with 
 the gas. This is similar to entropy being conserved following an
@@
 \begin{align}
-  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\epsilon &> 0,\label{dusty_solberg1}  \\
-  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\epsilon}{\p
-    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \epsilon}{\p r} \right) & > 0, \label{dusty_solberg2}
+  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\tepsilon &> 0,\label{dusty_solberg1}  \\
+  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\tepsilon}{\p
+    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \tepsilon}{\p r} \right) & > 0, \label{dusty_solberg2}
 \end{align}
@@
 Solberg-Hoiland criteria for adiabatic gas if one writes the 
-entropy $s=\ln{\left[c_s^2\left(1 - \epsilon\right)\right]}$.
+entropy $s=\ln{\left[c_s^2\left(1 - \tepsilon\right)\right]}$.
 
 %These criteria are applicable to any adiabatic fluid. 
-%For an isothermal mixture, we may take $s=\ln{(1-\epsilon)}$.   
+%For an isothermal mixture, we may take $s=\ln{(1-\tepsilon)}$.   
 %from the definition of $s$ and  
@@
 %\begin{align}
-%  s \equiv \ln \frac{P}{\rho} = \ln{(1-\epsilon)} 
+%  s \equiv \ln \frac{P}{\rho} = \ln{(1-\tepsilon)} 
 %\end{align}
@@
 %The stability conditions are thus
-%In accretion disks, $\rhog$, $\rhod$ and $\epsilon$ are expected to
+%In accretion disks, $\rhog$, $\rhod$ and $\tepsilon$ are expected to
 %decrease away from the mid-plane
@@
 %\end{align*}
-%and similarly for $\rhod$, $\epsilon$, and the dust-to-gas ratio $\tepsilon$. 
+%and similarly for $\rhod$, $\tepsilon$, and the dust-to-gas ratio $\epsilon$. 
 %We may thus evaluate the stability criteria at $z>0$ without loss of
@@
 %Eq. \ref{vshear}, into Eq. \ref{dusty_solberg} and write the result 
-%in terms of $\tepsilon$, to find
+%in terms of $\epsilon$, to find
 
@@
 %However, here we permit 
-%$\tepsilon_0=\tepsilon_0(r)$ and $\Htilde=\Htilde(r)$ to vary with
+%$\epsilon_0=\epsilon_0(r)$ and $\Htilde=\Htilde(r)$ to vary with
 %radius. 
@@
 $\kappa\simeq\OmK$. In terms of the dust-to-gas
-ratio $\tepsilon$, Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2}
+ratio $\epsilon$, Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2}
 become   
 \begin{align}
-  &1 + \frac{\tepsilon}{\left(1+
-    \tepsilon\right)^2}\left(h_\mathrm{g}^2\frac{\p\ln{\rhog}}{\p\ln{r}}
-  \frac{\p\ln{\tepsilon}}{\p\ln{r}} + 
-  \frac{z^2}{\widetilde{H}^2}\right)>0 \label{stability_est1},   \\ 
-&1 - \frac{\tepsilon
-  h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}
-  \frac{\p\ln{\tepsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\widetilde{H}^2}{H_\mathrm{g}^2}\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right)
+  &1 + \frac{\epsilon}{\left(1+
+    \epsilon\right)^2}\left(h_\mathrm{g}^2\frac{\p\ln{\rhog}}{\p\ln{r}}
+  \frac{\p\ln{\epsilon}}{\p\ln{r}} + 
+  \frac{z^2}{\Htilde^2}\right)>0 \label{stability_est1},   \\ 
+&1 - \frac{\epsilon
+  h_\mathrm{g}^2}{\left(1+\epsilon\right)^2}
+  \frac{\p\ln{\epsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\Htilde^2}{H_\mathrm{g}^2}\frac{\p\ln{\epsilon}}{\p\ln{r}}\right)
   > 0 \label{stability_est2},
@@
 in thin accretion disks where radial gradients are $O(1/r)$, 
-$h_\mathrm{g},\tepsilon,\ll 1$ and $
-\widetilde{H}/H_\mathrm{g}\lesssim 1$. Then the
+$h_\mathrm{g},\epsilon,\ll 1$ and $
+\Htilde/H_\mathrm{g}\lesssim 1$. Then the
 magnitude of the second term on the left-hand-side  of either
@@
 Notice the left-hand-side of Eq. \ref{stability_est2} is a quadratic in
-$\p_r\tepsilon$. It is possible to
+$\p_r\epsilon$. It is possible to
 violate this inequality for sufficiently large (in magnitude) radial
@@
 \begin{align}
-  \frac{\p\ln{\tepsilon}}{\p\ln{r}} > S_+ \quad \mathrm{or} \quad 
-  \frac{\p\ln{\tepsilon}}{\p\ln{r}} < S_-,
+  \frac{\p\ln{\epsilon}}{\p\ln{r}} > S_+ \quad \mathrm{or} \quad 
+  \frac{\p\ln{\epsilon}}{\p\ln{r}} < S_-,
 \end{align}
@@
 \begin{align}\label{spm}
-S_\pm = \frac{1}{2}\frac{H_\mathrm{g}^2}{\widetilde{H}^2} 
+S_\pm = \frac{1}{2}\frac{H_\mathrm{g}^2}{\Htilde^2} 
   \left[
@@
   \left(\frac{\p\ln{\rhog}}{\p\ln{r}}\right)^2 + 
-  4 \frac{\widetilde{H}^2}{H_\mathrm{g}^2}
-  \frac{\left(1+\tepsilon\right)^2}{\tepsilon h_\mathrm{g}^2}
+  4 \frac{\Htilde^2}{H_\mathrm{g}^2}
+  \frac{\left(1+\epsilon\right)^2}{\epsilon h_\mathrm{g}^2}
   }
@@
 achieve instability at a given radius for increasing dust-to-gas
-ratios in outwards ($\p_r\tepsilon > 0$), and vice versa.
+ratios in outwards ($\p_r\epsilon > 0$), and vice versa.
 
@@
 \begin{align}
-  \left|\frac{\p\ln{\tepsilon}}{\p r}\right| \gtrsim
-  \frac{1}{\widetilde{H}}\frac{\left(1+\tepsilon\right)}{\sqrt{\tepsilon}}. 
+  \left|\frac{\p\ln{\epsilon}}{\p r}\right| \gtrsim
+  \frac{1}{\Htilde}\frac{\left(1+\epsilon\right)}{\sqrt{\epsilon}}. 
 \end{align}
-That is, if the dust-to-gas ratio varies on a lengthscale less than
-$O(\widetilde{H}\sqrt{\tepsilon})$ for $\tepsilon\ll 1$, or less than
-$O(\widetilde{H}/\sqrt{\tepsilon})$ for $\tepsilon\gg 1$, then the system is
-potentially unstable.
+That is, if the dust-to-gas ratio varies on a radial lengthscale less
+than $O(\Htilde)$, i.e. its vertical lengthscale,  
+%$O(\widetilde{H}\sqrt{\epsilon})$ for $\epsilon\ll 1$, or less than
+%$O(\widetilde{H}/\sqrt{\epsilon})$ for $\epsilon\gg 1$, 
+then the system is potentially unstable.
 
@@
 %stars \citep{latter06}.  
-which represent all instabilities due to large-but-finite dust-gas
-friction. The formalism we have adopted give these instabilities a 
+which represent instabilities due to large-but-finite dust-gas
+friction. Our formalism gives these dusty instabilities a 
 thermodynamical interpretation because the dust-gas friction enters the problem
 as an energy source term \citep[cf. thermal instabilities in
-  stars][]{latter06}. 
+  stars, see also][]{latter06}. 
 This includes the streaming instability in

### 2016-08-01 paper.tex
updated notes
order of mag estimate
demo fig
general edits
sample results
notes and code update
title change
text and code update
adding results
regular update

@@
 \usepackage{amsmath}
+\usepackage{paralist}
 %\usepackage{url}
@@
 \newcommand{\dbz}{\dd B_z}
+\newcommand{\Htilde}{H_\epsilon}
+\newcommand{\Hdust}{H_\mathrm{d}}
+\newcommand{\Hgas}{H_\mathrm{g}}
+\newcommand{\hgas}{h_\mathrm{g}}
+\newcommand{\tepsilon}{\tilde{\epsilon}}
 \newcommand{\rhod}{\rho_\mathrm{d}}
@@
 \newcommand{\dphi}{\dd \Phi}
+\newcommand{\Hg}{H_\mathrm{g}}
+\newcommand{\Hd}{H_\mathrm{d}} 
 \newcommand{\actaa}{Acta Astron.}
 \defcitealias{lin12}{L12} 
+\defcitealias{lin15}{LY15}
 \defcitealias{nelson13}{N13}
@@
 
-\title{untitled}
-%\title{Vertical shear instability in astrophysical disks: vertically
-%global linear calculations with thermal relaxation and applications
-%to the Solar Nebula}  
+\title{On the entropy of 
+  isothermal dusty gas: application to protoplanetary disks}  
 \author{Min-Kai Lin\altaffilmark{1}}
@@
 \begin{abstract}
-dusty vsi, or the lack thereof
+  Small dust particles embedded in gaseous protoplanetary disks are subject
+  to strong dust-gas friction, causing them to almost follow the gas
+  flow. This is similar to the conservation of entropy following an
+  adiabatic fluid.  We show that in the strong drag
+  limit that an isothermal dusty-gas indeed behaves like an  
+  adiabatic pure gas, while finite dust-gas coupling may be
+  interpreted as an effective 
+  heating/cooling. We exploit this equivalence 
+  to 
+\begin{inparaenum}[1)]
+\item
+  define the entropy of isothermal dusty gas; and 
+\item 
+  generalize previous, well-known gaseous instabilities in
+  protoplanetary disks to include dust dynamics. 
+\end{inparaenum}
+This includes the axisymmetric stability of strictly isothermal dusty
+gas, from which we place quantitative constraints on the sharpness of 
+dust edges that can persist in protoplanetary disks. We also extend
+the vertical shear instability (VSI) in locally isothermal disks to include
+dust. We find that  dust-loading reduces VSI growth rates. The presence of
+strongly-coupled dust particles preferentially stabilize VSI modes with small radial
+wavelengths, as well as `surface modes' concentrated near the
+disk boundaries. We briefly discuss other applications of the 
+equivalence between isothermal dusty gas and adiabatic hydrodynamics
+to understand other dusty instabilities in protoplanetary disks. 
 \end{abstract}
 
+\section{Introduction}
+
 \input{model}
-%\input{linear_iso}
+\input{linear}
+\input{result}
 %\input{linear_adia}
@@
 
-%\appendix
-%\input{appendix}
+\appendix
+\input{appendix}
 

### 2016-08-01 result.tex
regular update

@@
-\section{Perfectly coupled
-  dusty vertical shear instability}\label{results}
+\section{
+  Dusty vertical shear instability}\label{results}
 
@@
 consider constant input values of the midplane dust-to-gas ratio
-$\tepsilon_0$ and the characteristic dust thickness $\Hd$. These,
+$\epsilon_0$ and the characteristic dust thickness $\Hd$. These,
 together with the perturbation wavenumber $k$, comprises main the
@@
 \subsection{Effect of dust-loading}
-We first vary the dust-to-gas ratio $\tepsilon_0\in[10^{-3},1]$,
-fixing the dust thickness to $\Hd=0.99\Hg$ (so that $\tepsilon$ is
+We first vary the dust-to-gas ratio $\epsilon_0\in[10^{-3},1]$,
+fixing the dust thickness to $\Hd=0.99\Hg$ (so that $\epsilon$ is
 roughly constant with height) and perturbation
@@
 Fig. \ref{vsi_dust_loading} show unstable
-modes found for different values of $\tepsilon_0$. The eigenvalue
-distributions for $\tepsilon_0 \leq 10^{-2}$ are similar to the
+modes for different values of $\epsilon_0$. The eigenvalue
+distributions for $\epsilon_0 \leq 10^{-2}$ are similar to the
 dust-free fiducial case considered by \citetalias{lin15}, consisting
@@
 rates. Notably `surface modes', which are typically fastest growing in
-the dust-free case, are surpressed in dusty disks for $\tepsilon_0\geq
-0.1$. The body modes' growth rates remain $\sim O(h\OmK)$ but their
-oscillation frequency increases with $\tepsilon_0$.
-
-In Fig. \ref{vsi_dust_loading2d} we compare the
+the dust-free case, are surpressed in dusty disks for $\epsilon_0\geq
+0.1$. The body modes' growth rates remain $\sim O(h_\mathrm{g}\OmK)$ but their
+oscillation frequency increases with $\epsilon_0$. The lowest
+frequency `fundamental' body mode is not the fastest growing, but they
+are energetically dominant because the entire disk column is perturbed
+\citep[cf. surface modes which only disturb the disk boundaries,][]{umurhan16c}.
 
@@
     frequency $\omega$ and growth rates $s$ are shown for a range of
-    midplane dust-to-gas ratios $\tepsilon_0=\rhod/\rhog$. 
+    midplane dust-to-gas ratios $\epsilon_0=\rhod/\rhog$. 
     \label{vsi_dust_loading}
@@
 \end{figure}
+
+In Fig. \ref{vsi_dust_loading2d}
+we compare the fundamental mode between the nearly 
+dust-free case $\epsilon_0=10^{-3}$ an a dusty disk with
+$\epsilon_0=1$. Dust-loading preferentially
+stabilizes the disk atmosphere against the VSI, restricting
+meridional motions to $|z|\lesssim 2\Hg$, where most of the disk mass
+reside. Notice also the perturbed dust-to-gas ratio,
+$\delta\epsilon$, shifts from being maximized at the disk boundaries
+to the midplane.
 
@@
   \caption{Fundamental dusty VSI mode in real space for midplane dust-to-gas
-    ratio $\tepsilon_0=10^{-3}$ (top) and $\tepsilon_0=1$
+    ratio $\epsilon_0=10^{-3}$ (top) and $\epsilon_0=1$
     (bottom). The color scale shows the perturbation to the
-    dust-to-gas ratio, $\delta\tepsilon$; and the arrows show
+    dust-to-gas ratio, $\delta\epsilon$; and the arrows show
     $\sqrt{\rho}\left(\dd v_x, \dd v_z\right)$. 
@@
 
+In Fig. \ref{vsi_dust_loading_vareps} we plot the growth rates as a
+function of $\epsilon_0$ for different perturbation wavenumbers $k$. We
+find that dust-loading stabilizes the VSI more effectively for shorter
+wavelength perturbations. This is because for high wavenumbers the
+dominant modes are surface modes, which are stabilized by
+dust-loading. The figure suggest that VSI becomes much less efficient
+for $\epsilon_0\gtrsim 0.1$ and $k\Hg\gtrsim 50$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_eigenvals_vareps} 
+  \caption{Maximum growth rate of the dusty VSI as a function of the
+    midplane dust-to-gas ratio $\epsilon_0$ for perturbations with
+    different radial wavenumbers $k$. The dust layer thickness is
+    fixed to $\Hd\simeq \Hg$. 
+    \label{vsi_dust_loading_vareps}
+    }
+\end{figure}
+
 \subsection{Effect of dust thickness}
+
+\subsection{Fixed metalicity}

### 2016-08-01 paper.tex
regular update

@@
   adiabatic fluid.  We show that in the strong drag
-  limit that an isothermal dusty-gas indeed behaves like an  
-  adiabatic pure gas, while finite dust-gas coupling may be
-  interpreted as an effective 
+  limit that an isothermal dusty gas indeed behaves like an  
+  adiabatic pure gas, and finite dust-gas coupling may be
+  regarded as an effective 
   heating/cooling. We exploit this equivalence 
@@
 wavelengths, as well as `surface modes' concentrated near the
-disk boundaries. We briefly discuss other applications of the 
-equivalence between isothermal dusty gas and adiabatic hydrodynamics
-to understand other dusty instabilities in protoplanetary disks. 
+disk boundaries. We briefly discuss potential applications of the 
+equivalence between isothermal dusty gas and standard hydrodynamics
+to provide thermodynamical interpretations of  other dusty instabilities in protoplanetary disks. 
 \end{abstract}

### 2016-08-02 paper.tex
adding fixZ results

@@
   isothermal dusty gas: application to protoplanetary disks}  
-\author{Min-Kai Lin\altaffilmark{1}}
+\author{Min-Kai Lin\altaffilmark{1} and Andrew N. Youdin}
 \affil{Department of Astronomy and Steward Observatory, \\
@@
 \begin{abstract}
-  Small dust particles embedded in gaseous protoplanetary disks are subject
-  to strong dust-gas friction, causing them to almost follow the gas
-  flow. This is similar to the conservation of entropy following an
-  adiabatic fluid.  We show that in the strong drag
+  Small solids embedded in gaseous protoplanetary disks are subject
+  to strong dust-gas friction. Tightly-coupled dust particles
+  therefore almost follow the gas flow. This is analogous to the
+  conservation of entropy 
+  following an adiabatic fluid.  We show that in the strong drag
   limit that an isothermal dusty gas indeed behaves like an  
@@
 This includes the axisymmetric stability of strictly isothermal dusty
-gas, from which we place quantitative constraints on the sharpness of 
+gas, allowing us to constrain the sharpness of 
 dust edges that can persist in protoplanetary disks. We also extend
 the vertical shear instability (VSI) in locally isothermal disks to include
-dust. We find that  dust-loading reduces VSI growth rates. The presence of
-strongly-coupled dust particles preferentially stabilize VSI modes with small radial
+dust. We find that  dust-loading reduces VSI growth rates, and
+preferentially stabilizes VSI modes with small radial 
 wavelengths, as well as `surface modes' concentrated near the
@@
 equivalence between isothermal dusty gas and standard hydrodynamics
-to provide thermodynamical interpretations of  other dusty instabilities in protoplanetary disks. 
+to provide thermodynamical interpretations of  other dusty
+instabilities in protoplanetary disks.  
 \end{abstract}
@@
 %\input{application} 
-%\input{summary}
+\input{summary}
 

### 2016-08-02 result.tex
adding fixZ results

@@
 the dust-free case, are surpressed in dusty disks for $\epsilon_0\geq
-0.1$. The body modes' growth rates remain $\sim O(h_\mathrm{g}\OmK)$ but their
-oscillation frequency increases with $\epsilon_0$. The lowest
-frequency `fundamental' body mode is not the fastest growing, but they
-are energetically dominant because the entire disk column is perturbed
-\citep[cf. surface modes which only disturb the disk boundaries,][]{umurhan16c}.
+0.1$. The body modes' growth rates remain $\sim O(h_\mathrm{g}\OmK)$
+but their oscillation frequency increases with
+$\epsilon_0$. The total number of modes do not change
+significantly. This is in contrast with the effect of increasing
+cooling times, where \citetalias{lin15} find fewer unstable modes.
+%until the system is eventually completely stable.
 
@@
     frequency $\omega$ and growth rates $s$ are shown for a range of
-    midplane dust-to-gas ratios $\epsilon_0=\rhod/\rhog$. 
+    midplane dust-to-gas ratios $\epsilon_0=\rho_\mathrm{g0}/\rho_\mathrm{d0}$. 
     \label{vsi_dust_loading}
@@
 
-In Fig. \ref{vsi_dust_loading2d}
+The lowest frequency `fundamental' body mode is not the fastest
+growing, but they are energetically dominant because the entire disk
+column is perturbed \citep[cf. surface modes which only disturb the
+  disk boundaries,][]{umurhan16c}. In Fig. \ref{vsi_dust_loading2d}
 we compare the fundamental mode between the nearly 
@@
 \subsection{Fixed metalicity} 
+Here we vary the dust layer thickness $\Hd$ but hold 
+\begin{align}\label{metalicity_def}
+Z \equiv \epsilon_0 \Htilde = \text{constant,}
+\end{align}
+so that $\Sigma_\mathrm{d}/\Sigma_\mathrm{g}$ is
+fixed. Eq. \ref{metalicity_def} then gives the required midplane
+dust-to-gas ratio.
+
+%zmax = 2 because we have very thin dust layer
+
+We set $Z=0.03$ and consider two disks with $\Hd=0.1\Hg$ and
+$\Hd=0.99\Hg$. Fig. \ref{compare_vshear_fixZ} compares the
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_vshear_fixZ} 
+  \caption{Vertical shear rate in a locally isothermal, dusty disk
+    with fixed dust-to-gas mass and dust thickness $\Hd=0.1\Hg$
+    (black) and $\Hd=0.99\Hg$ (blue). 
+    \label{compare_vshear_fixZ}
+    }
+\end{figure}
+
+\begin{figure}
+  \includegraphics[scale=0.33, clip=true, trim=0.5cm 0cm 3cm 0cm]{figures/result2d_Hd1.ps}\includegraphics[scale=0.33, clip=true, trim=2cm 0cm 0cm 0cm]{figures/result2d_Hd0d1.ps} 
+  \caption{
+%dusty VSI mode in real space for midplane dust-to-gas
+%    ratio $\epsilon_0=10^{-3}$ (top) and $\epsilon_0=1$
+%    (bottom). The color scale shows the perturbation to the
+%    dust-to-gas ratio, $\delta\epsilon$; and the arrows show
+%    $\sqrt{\rho}\left(\dd v_x, \dd v_z\right)$. 
+    stuff
+    \label{result2d_fixZ}
+    }
+\end{figure}

### 2016-08-02 model.tex
adding fixZ results
new vshear figs

@@
 
-In practice, we consider $\epsilon_0 \ll 1$ and settled dust such that
+We usually consider $\epsilon_0 \ll 1$ and settled dust such that
 $\Htilde<\Hgas$. Then $\rhog(r,z)$ is effectively Gaussian, as in the
@@
   \rhod \simeq \epsilon_0\rho_\mathrm{g0}(r) \exp
-        {\left(-\frac{z}{2H_\mathrm{d}^2}\right)}, 
+        {\left(-\frac{z^2}{2H_\mathrm{d}^2}\right)}, 
 \end{align}
@@
 $\Htilde$ for input.
+
+Finally, we define 
+\begin{align}
+  Z \equiv \epsilon_0\frac{\Hd}{\Hg} \simeq
+  \frac{\Sigma_\mathrm{d}}{\Sigma_\mathrm{g}} 
+\end{align}
+as a proxy of the ratio between the dust surface density
+$\Sigma_\mathrm{d}$ and the gas surface density $\Sigma_\mathrm{g}$.
 

### 2016-08-02 result.tex
adding fixZ results
new vshear figs

@@
 wavenumber to $k\Hg = 30$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_vshear_Nz2_fixHd} 
+  \caption{Vertical shear rate (solid) compared to vertical buoyancy (dotted) in a locally isothermal, dusty disk
+    with dust thickness $\Hd=0.99\Hg$ and midplane dust-to-gas ratio
+    of $\epsilon_0=10^{-3}$ (black) and $\epsilon_0=1$ (blue). 
+    \label{compare_vshear_fixHd}
+    }
+\end{figure}
 
@@
 \begin{figure}
-  \includegraphics[scale=0.54, clip=true, trim=0cm 2.5cm 0cm 0cm]{figures/result2d_dg1d-3.ps}\\
-  \includegraphics[scale=0.54]{figures/result2d_dg1.ps} 
+%  \includegraphics[scale=0.54, clip=true, trim=0cm 2.5cm 0cm 0cm]{figures/result2d_dg1d-3.ps}\\
+%  \includegraphics[scale=0.54]{figures/result2d_dg1.ps} 
+\includegraphics[scale=0.32, clip=true, trim=0.5cm 0cm 3cm 0cm]{figures/result2d_dg1d-3.ps}\includegraphics[scale=0.32, clip=true, trim=1.8cm 0cm 0cm 0cm]{figures/result2d_dg1.ps}
   \caption{Fundamental dusty VSI mode in real space for midplane dust-to-gas
@@
 Here we vary the dust layer thickness $\Hd$ but hold 
-\begin{align}\label{metalicity_def}
-Z \equiv \epsilon_0 \Htilde = \text{constant,}
-\end{align}
+$Z \equiv \epsilon_0 \Htilde = \text{constant,}$ 
 so that $\Sigma_\mathrm{d}/\Sigma_\mathrm{g}$ is
-fixed. Eq. \ref{metalicity_def} then gives the required midplane
-dust-to-gas ratio.
+fixed. This gives the required midplane dust-to-gas ratio.
 
@@
 \begin{figure}
-  \includegraphics[width=\linewidth]{figures/compare_vshear_fixZ} 
-  \caption{Vertical shear rate in a locally isothermal, dusty disk
-    with fixed dust-to-gas mass and dust thickness $\Hd=0.1\Hg$
+  \includegraphics[width=\linewidth]{figures/compare_vshear_Nz2_fixZ} 
+  \caption{Vertical shear rate (solid) compared to vertical buoyancy (dotted) in a locally isothermal, dusty disk
+    with fixed dust-to-gas mass ratio $Z=0.03$ and dust thickness $\Hd=0.1\Hg$
     (black) and $\Hd=0.99\Hg$ (blue). 
@@
 \begin{figure}
-  \includegraphics[scale=0.33, clip=true, trim=0.5cm 0cm 3cm 0cm]{figures/result2d_Hd1.ps}\includegraphics[scale=0.33, clip=true, trim=2cm 0cm 0cm 0cm]{figures/result2d_Hd0d1.ps} 
+  \includegraphics[scale=0.32, clip=true, trim=0.5cm 0cm 3cm 0cm]{figures/result2d_Hd1.ps}\includegraphics[scale=0.32, clip=true, trim=1.8cm 0cm 0cm 0cm]{figures/result2d_Hd0d1.ps} 
   \caption{

### 2016-08-03 paper.tex
new vshear figs
started summary

@@
   Small solids embedded in gaseous protoplanetary disks are subject
-  to strong dust-gas friction. Tightly-coupled dust particles
-  therefore almost follow the gas flow. This is analogous to the
+  to strong dust-gas friction. Consequently, tightly-coupled dust
+  particles almost follows the gas flow. 
+  %Tightly-coupled dust particles
+  % in protoplanetary disks almost follow the gas flow. 
+  This is analogous to the 
   conservation of entropy 
@@
   limit that an isothermal dusty gas indeed behaves like an  
-  adiabatic pure gas, and finite dust-gas coupling may be
+  adiabatic pure gas, and that finite dust-gas coupling may be
   regarded as an effective 
@@
 \item
-  define the entropy of isothermal dusty gas; and 
+  identify the entropy of isothermal dusty gas; and 
 \item 
@@
 gas, allowing us to constrain the sharpness of 
-dust edges that can persist in protoplanetary disks. We also extend
-the vertical shear instability (VSI) in locally isothermal disks to include
-dust. We find that  dust-loading reduces VSI growth rates, and
+dust edges that can persist in protoplanetary disks. We also develop 
+the dusty vertical shear instability (VSI) in locally isothermal
+disks. We find that  dust-loading reduces VSI growth rates, and
 preferentially stabilizes VSI modes with small radial 
 wavelengths, as well as `surface modes' concentrated near the
-disk boundaries. We briefly discuss potential applications of the 
+disk boundaries. These results are consistent with our definition of
+the buoyancy frequency of a isothermal dusty disk. 
+%We show that vertical shear arising from
+%dust settling does not lead to axisymmetric instability. 
+We briefly discuss potential applications of the 
 equivalence between isothermal dusty gas and standard hydrodynamics
-to provide thermodynamical interpretations of  other dusty
+in providing thermodynamical interpretations of other dusty
 instabilities in protoplanetary disks.  

### 2016-08-03 result.tex
started summary

@@
 \subsection{Effect of dust-loading}
-We first vary the dust-to-gas ratio $\epsilon_0\in[10^{-3},1]$,
-fixing the dust thickness to $\Hd=0.99\Hg$ (so that $\epsilon$ is
-roughly constant with height) and perturbation
-wavenumber to $k\Hg = 30$.
+We first vary the midplane dust-to-gas ratio 
+$\epsilon_0\in[10^{-3},1]$, fixing the dust thickness to 
+$\Hd=0.99\Hg$. Then  $\epsilon$ is roughly constant with height. We
+set the vertical domain to $\zmax=5\Hg$.
+
+We compare in Fig. \ref{compare_vshear_fixHd} the basic state
+vertical shear rate, which is destabilizing, and the vertical buoyancy
+frequency, which is stabilizing. For the nearly dust-free disk
+$\epsilon_0=10^{-3}$ the vertical shear dominates buoyancy for all
+$|z|>0$. However, a heavy dust-load with $\epsilon_0=1$ renders the
+buoyancy to dominate over vertical shear in the disk atmosphere
+$|z|\gtrsim 2.5$. We thus expect instability all heights for
+$\epsilon_0=10^{-3}$, but to be restricted to the midplane for
+$\epsilon_0=1$.
 
@@
   \includegraphics[width=\linewidth]{figures/compare_vshear_Nz2_fixHd} 
-  \caption{Vertical shear rate (solid) compared to vertical buoyancy (dotted) in a locally isothermal, dusty disk
-    with dust thickness $\Hd=0.99\Hg$ and midplane dust-to-gas ratio
+  \caption{Vertical shear rate (solid) compared to vertical buoyancy
+    (dotted) in a locally isothermal, dusty disl with midplane dust-to-gas ratio
     of $\epsilon_0=10^{-3}$ (black) and $\epsilon_0=1$ (blue). 
+    The dust layer thickness is fixed to $\Hd=0.99\Hg$ so the 
+    dust-to-gas ratio $\epsilon$ is approximately constant with
+    height. 
     \label{compare_vshear_fixHd}
@@
 
-Fig. \ref{vsi_dust_loading} show unstable
-modes for different values of $\epsilon_0$. The eigenvalue
-distributions for $\epsilon_0 \leq 10^{-2}$ are similar to the
+Fig. \ref{vsi_dust_loading} show unstable modes for different values
+of $\epsilon_0$ with fixed perturbation wavenumber  $k\Hg = 30$. The
+eigenvalue distributions for $\epsilon_0 \leq 10^{-2}$ are similar to the
 dust-free fiducial case considered by \citetalias{lin15}, consisting
@@
 
-The lowest frequency `fundamental' body mode is not the fastest
-growing, but they are energetically dominant because the entire disk
-column is perturbed \citep[cf. surface modes which only disturb the
-  disk boundaries,][]{umurhan16c}. In Fig. \ref{vsi_dust_loading2d}
+The lowest frequency `fundamental' body mode is energetically dominant
+because the entire disk column is perturbed \citep[cf. surface modes
+  which only disturb the disk boundaries,][]{umurhan16c}. In Fig. \ref{vsi_dust_loading2d}
 we compare the fundamental mode between the nearly 
@@
 meridional motions to $|z|\lesssim 2\Hg$, where most of the disk mass
-reside. Notice also the perturbed dust-to-gas ratio,
+reside. This is consistent with the above comparison of the basic state vertical
+shear and buoyancy. Notice also the perturbed dust-to-gas ratio,
 $\delta\epsilon$, shifts from being maximized at the disk boundaries
@@
   \caption{Fundamental dusty VSI mode in real space for midplane dust-to-gas
-    ratio $\epsilon_0=10^{-3}$ (top) and $\epsilon_0=1$
-    (bottom). The color scale shows the perturbation to the
+    ratio $\epsilon_0=10^{-3}$ (left) and $\epsilon_0=1$
+    (right). The color scale shows the perturbation to the
     dust-to-gas ratio, $\delta\epsilon$; and the arrows show
@@
 \subsection{Effect of dust thickness} 
+We now vary the dust layer thickness $\Hd$ but hold 
+$Z \equiv \epsilon_0 \Hd/\Hg = \text{constant,}$ 
+so that $\Sigma_\mathrm{d}/\Sigma_\mathrm{g}$ is
+fixed. This then gives the required midplane dust-to-gas ratio.    
+Since we will consider thin dust layers, here we use a smaller
+domain with $\zmax=2\Hg$ so that $\epsilon$ does not become
+too small.
 
-\subsection{Fixed metalicity} 
-Here we vary the dust layer thickness $\Hd$ but hold 
-$Z \equiv \epsilon_0 \Htilde = \text{constant,}$ 
-so that $\Sigma_\mathrm{d}/\Sigma_\mathrm{g}$ is
-fixed. This gives the required midplane dust-to-gas ratio.
-
-%zmax = 2 because we have very thin dust layer
-
-We set $Z=0.03$ and consider two disks with $\Hd=0.1\Hg$ and
-$\Hd=0.99\Hg$. Fig. \ref{compare_vshear_fixZ} compares the
+We set $Z=0.03$ and analyse two disks with $\Hd=0.1\Hg$ and 
+$\Hd=0.99\Hg$. Fig. \ref{compare_vshear_fixZ} compares the vertical
+vhear rate and buoyancy frequency. For $|z|\gtrsim 0.4\Hg$ the two
+disks have the same profile with vertical shear dominating over
+buoyancy. We thus expect instabilities away from the disk midplane in 
+both cases. For $|z|\lesssim 0.4\Hg$, a thin dust 
+layer with $\Hd=0.1\Hg$ boosts the vertical shear rate, but the
+associated buoyancy is larger still. Thus we expect the mid-plane of
+the disk with $\Hd=0.1\Hg$ to be stable.
 
@@
   \includegraphics[width=\linewidth]{figures/compare_vshear_Nz2_fixZ} 
-  \caption{Vertical shear rate (solid) compared to vertical buoyancy (dotted) in a locally isothermal, dusty disk
-    with fixed dust-to-gas mass ratio $Z=0.03$ and dust thickness $\Hd=0.1\Hg$
+  \caption{Vertical shear rate (solid) compared to vertical buoyancy
+    (dotted) in a locally isothermal, dusty disk 
+    with dust-to-gas mass ratio $Z=0.03$ and dust thickness $\Hd=0.1\Hg$
     (black) and $\Hd=0.99\Hg$ (blue). 
@@
 
+Fig. \ref{result2d_fixZ} compares the fastest-growing VSI body modes  
+with $k\Hg=30$ for the two cases above.\citepalias[The thinner domain
+  adopted here eliminates surface modes, ][]{lin15}.   
+We find very similar mode 
+frequencies 
+\begin{align*}
+  \sigma = \begin{cases}
+    \left(0.3053\ii - 0.8142\right)h_\mathrm{g}\OmK & \Hd=0.99\Hg, \\
+    \left(0.3178\ii - 1.2237\right)h_\mathrm{g}\OmK & \Hd=0.1\Hg,
+  \end{cases}
+\end{align*}
+since the vertical shear profile is similar throughout most of the
+disk. However, meridional motions are surpressed near the midplane of
+the $\Hd=0.1\Hg$ disk, as expected from the larger buoyancy frequency
+as compared to vertical shear there. This leads to a structure
+analogous to PPD dead zones: a quiescent midplane between active
+surface layers \citep{gammie96}.
+
 \begin{figure}
   \includegraphics[scale=0.32, clip=true, trim=0.5cm 0cm 3cm 0cm]{figures/result2d_Hd1.ps}\includegraphics[scale=0.32, clip=true, trim=1.8cm 0cm 0cm 0cm]{figures/result2d_Hd0d1.ps} 
-  \caption{
-%dusty VSI mode in real space for midplane dust-to-gas
-%    ratio $\epsilon_0=10^{-3}$ (top) and $\epsilon_0=1$
-%    (bottom). The color scale shows the perturbation to the
-%    dust-to-gas ratio, $\delta\epsilon$; and the arrows show
-%    $\sqrt{\rho}\left(\dd v_x, \dd v_z\right)$. 
-    stuff
+  \caption{Fastest-growing dusty VSI mode in real space for midplane
+    dust layer thickness $\Hd=0.99\Hg$ (left) and $\Hd=0.1\Hg$
+    (right). The dust content is fixed to
+    $\Sigma_\mathrm{d}=0.03\Sigma_\mathrm{g}$. 
+    The color scale shows the perturbation to the
+    dust-to-gas ratio, $\delta\epsilon$; and the arrows show
+    $\sqrt{\rho}\left(\dd v_x, \dd v_z\right)$.
     \label{result2d_fixZ}

### 2016-08-03 summary.tex
started summary

\section{Summary and discussion}
In this paper we highlight the similarity between an isothermal 
dusty gas and an adiabatic, pure gas. 
In the limit of perfect dust-gas coupling, the dust content of a gas 
parcel $\rhod/\rhog$ is conserved. This is analogous to entropy being 
conserved following a adiabatic, pure gas. %s = s(\rhod/\rhog)
We explicitly show that for 
a locally isothermal, dusty gas that the evolutionary equation of 
$\rhod/\rhog$ may be recast into an effective energy equation identical
in form to that in standard adiabatic hydrodynamics. We find the 
natural definition of the entropy of the dusty gas is 
\begin{align*}
  s  = \ln{\left(\frac{c_s^2\rhog}{\rhog+\rhod}\right)}.  
\end{align*}
This allows us to define the buoyancy frequency of an isothermal 
dusty gas.

%
%finite coupling/heat source

Finite dust-gas friction causes a relative drift between the two
components of the mixture. Then $\rhod/\rhog $ is no
longer conserved following a fluid parcel, as it is allowed to
exchange dust particles with neighboring parcels owing to a non-zero
stopping time $\tstop$. This is analogous to heat exchange between a 
pure gas parcel and its surroundings. Thus, finite dust-gas friction
may be interepreted as an energy/entropy source term for the dusty
gas.

%
In this first study, we apply this equivalence principle to study the
axisymmetric stability of locally isothermal, protoplanetary disks
perfectly coupled to dust ($\tstop=0$). This is so that we may 
compute exact, steady equilibria for analyses. Furthermore,
in this limit the dusty problem is \emph{identical} to adiabatic, pure
gas dynamics.

We obtain the Solberg-Hoiland criteria to assess
axisymmetric stability when the equation of state is strictly
isothermal. We show that {\bf dust-settling cannot lead to
  axisymmetric instability, however large the associated vertical
  shear} \citep[cf. \emph{non-axisymmetric} Kelvin-Helmholtz instabilities
  induced by  dust-settling, ][]{lee10}. We  also show that
protoplanetary disks cannot have sharp edges in the dust-to-gas ratio,
as this implies a sharp entropy gradient and may thus render the disk
unstable.

%does not preclude non-axisymmetric instabilities. but if there is a
%global radial temp gradient, then vertical shear arising from
%it will always be comparable to dust-settling. but there is vsi with
%global rad temp grad

For locally isothermal disks $c_s=c_s(r)$ we generalize the vertical
shear instability (VSI) to include dust. We find that dust-loading
generally has a stabilizing effect. In our disk models
dust-loading does not affect VSI growth rates significantly, but
meridional motions may be suppressed at heights where vertical 
buoyancy dominates over vertical shear. This confirms our 
identification of the entropy and buoyancy of an isothermal, dusty
gas.

%growth rates not sig changed 
%v shear from thin dust layers are overwhelmed by

\subsection{Dusty analogs of other gaseous instabilities} 
%gi
%rwi

%\subsection{Implication for numerical simulations}

\subsection{Thermodynamic interpretation of other dusty instabilities}

%\subsection{Caveats and future directions}
%gi

### 2016-08-04 model.tex
regular update

@@
 energy flux has an additional contribution from the pressure
-gradient and dust-gas friction. 
-The term on the right-hand-side, owing to the imposed
-temperature profile, can be interpreted as an external heat source.
+gradient and dust-gas friction. The term on the right-hand-side, owing
+to the imposed temperature profile, can be interpreted as an external
+heat source.
+
+%We comment that
 
@@
 \end{align}
-as a proxy of the ratio between the dust surface density
-$\Sigma_\mathrm{d}$ and the gas surface density $\Sigma_\mathrm{g}$.
+as a measure of the local metalicity. The second approximation holds
+for $\epsilon_0\ll1$. 
+%\Sigma_\mathrm{d}$ and the gas
+%surface density $\Sigma_\mathrm{g}$.
 
@@
     \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\left\{\right\} }
-  \left. +\frac{q}{r}\left[\frac{\p\epsilon}{\p z} + \left(1+\epsilon\right)\frac{\p\ln{\rhog}}{\p z}\right]
+  \left. -\frac{q}{r}\left[\frac{\p\epsilon}{\p z} + \left(1+\epsilon\right)\frac{\p\ln{\rhog}}{\p z}\right]
   \right\} 
@@
 In this case, the last two integrals in Eq. \ref{int_rel} vanish. Then 
-requiring $\sigma^2>0$  gives the Solberg-Hoiland criteria 
+requiring $\sigma^2>0$  gives the criteria 
 for the axisymmtric \emph{stability} of an isothermal gas perfectly 
 coupled to dust,  
+%\begin{align}
+%  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\tepsilon &> 0,\label{dusty_solberg1}  \\
+%  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\tepsilon}{\p
+%    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \tepsilon}{\p r} \right) & > 0, \label{dusty_solberg2}
+%\end{align} 
 \begin{align}
-  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\tepsilon &> 0,\label{dusty_solberg1}  \\
-  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\tepsilon}{\p
-    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \tepsilon}{\p r} \right) & > 0, \label{dusty_solberg2}
+  \kappa^2 - \frac{1}{\rho}\nabla P \cdot \nabla s &> 0, \label{dusty_solberg1}    \\
+  -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
+  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, \label{dusty_solberg2}
 \end{align} 
-%\begin{align}
-%  \kappa^2 - \frac{1}{\rho}\nabla P \cdot \nabla s &> 0,\\
-%  -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
-%  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, 
-%\end{align} 
 where %$s = \ln{P/\rho}$, 
@@
 square of the epicylic frequency, and we assume $\kappa^2>0$.  
-Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} are equivalent to the
-Solberg-Hoiland criteria for adiabatic gas if one writes the 
-entropy $s=\ln{\left[c_s^2\left(1 - \tepsilon\right)\right]}$.
+Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} has the identical form
+to the Solberg-Hoiland criteria for adiabatic gas dynamics 
+\citep[e.g.][]{tassoul78}, except here the entropy $s$ is really a
+representation of the dust distribution via $s=\ln{\left[c_s^2\left(1
+    - \tepsilon\right)\right]}$.
 

### 2016-08-04 result.tex
regular update

@@
 \begin{figure}
-  \includegraphics[width=\linewidth]{figures/compare_eigenvals_vareps} 
+  \includegraphics[width=\linewidth]{figures/compare_eigenvals_vareps2} 
   \caption{Maximum growth rate of the dusty VSI as a function of the
@@
 \subsection{Effect of dust thickness} 
-We now vary the dust layer thickness $\Hd$ but hold 
-$Z \equiv \epsilon_0 \Hd/\Hg = \text{constant,}$ 
-so that $\Sigma_\mathrm{d}/\Sigma_\mathrm{g}$ is
-fixed. This then gives the required midplane dust-to-gas ratio.    
-Since we will consider thin dust layers, here we use a smaller
+We now vary the dust layer thickness $\Hd$ but fix the metalicity 
+$Z \equiv \epsilon_0 \Hd/\Hg = \text{constant}$. This then gives the
+required midplane dust-to-gas ratio. Since we will consider thin dust
+layers, here we use a smaller 
 domain with $\zmax=2\Hg$ so that $\epsilon$ does not become
@@
     (dotted) in a locally isothermal, dusty disk 
-    with dust-to-gas mass ratio $Z=0.03$ and dust thickness $\Hd=0.1\Hg$
+    with metalicity $Z=0.03$ and dust thickness $\Hd=0.1\Hg$
     (black) and $\Hd=0.99\Hg$ (blue). 
@@
 \end{figure}
+
+Fig. \ref{compare_eigenvals_fixZ} shows the maximum VSI growth rates
+as a function of the dust layer thickness. As before, we find growth
+rates are most affected by the vertical structure of the dust layer
+when the perturbation wavenumer is large. VSI growth rates converge as
+$\Hd\to 0$. %to dust free values?    
+Thus a thin dust layer, however large its associated vertical shear,
+does not affect the VSI. The non-monotonic behavior for $\Hd\gtrsim
+0.5\Hg$ arises because the vertical buoyancy frequency 
+\begin{align*}
+N_z^2(H_d;z,Z) \simeq &Z\Hg z^2
+\exp{\left(-\frac{z^2}{2\Hg^2}\right)}\OmK^2\notag\\
+&\times 
+\frac{1}{\Hd}\left(\frac{1}{\Hd^2} -
+\frac{1}{\Hg^2}\right)\exp{\left(-\frac{z^2}{2\Hd^2}\right)}  
+\end{align*}
+is a non-monotonic function of $\Hd$ at fixed $z$. At $z=\Hg$ and $z=2\Hg$
+the buoyancy frequency is maximized for $\Hd\simeq0.5\Hg$ and
+$\Hd\simeq 0.8\Hg$, respectively. This is consistent with the abscissa
+of minima in growth rates in Fig. \ref{compare_eigenvals_fixZ}.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_eigenvals_fixZ} 
+  \caption{Maximum VSI growth rate for different perturbation
+    wavenumbers $k$ as a function of the dust layer
+    thickness $\Hd$ at fixed metalicity $Z=0.03$. 
+    \label{compare_eigenvals_fixZ}
+    }
+\end{figure}

### 2016-08-04 summary.tex
regular update

@@
 
+\subsection{Generalization to locally polytropic disks}
+We can extend the dusty/adiabatic gas correspondence to 
+other fixed equations of state. As an example, consider the locally
+polytropic disk 
+\begin{align}
+  P = K(r,z)\rhog^{\Gamma}, 
+\end{align}
+where $K$ is a prescribed function and $\Gamma$ is the constant
+adiabatic index. Then eliminating $\tepsilon$ from the dust equation
+\ref{dusteq} gives 
+\begin{align}
+  \frac{D P}{D t} = - \Gamma P\nabla\cdot\bm{v}  + P \bm{v}\cdot\nabla \ln{K}
+  + \frac{\Gamma P}{\rhog}\nabla\cdot\left(\tepsilon\tstop\nabla
+  P\right).
+\end{align}
+Thus the dusty gas behaves like a pure gas with adiabatic index
+$\Gamma$. The entropy is given by 
+\begin{align}
+  s = \ln{\left[K(r,z)\left(1 - \tepsilon\right)^\Gamma\right]}.  
+\end{align}
+When $K$ is constant, the corresponding Solberg-Hoiland criteria for
+axisymmetric stability may be obtained from that in standard adiabatic 
+hydrodynamics \citep[e.g.][]{tassoul78} by
+using this definition of entropy in
+Eqs. \ref{dusty_solberg1}---\ref{dusty_solberg2}.
+
 \subsection{Dusty analogs of other gaseous instabilities}
+
 %gi
@@
 
-\subsection{Thermodynamic interpretation of other dusty instabilities}
+\subsection{Thermodynamic interpretation of dusty instabilities}
 

### 2016-08-04 model.tex
minor edits at home

@@
 
-\subsection{Locally isothermal equation of state for the gas} 
+\subsection{Locally isothermal equation of state} 
 We adopt 
@@
 \end{align}
-where $q$ is the power-law index for the disk temperature.
+where $q$ is the power-law index for the disk temperature. For $q=0$
+the disk is strictly isothermal.
 
@@
 isothermal dusty disks to in fact behave as if it
-had a temperture increasing away from $z=0$.
+had a temperture \emph{increasing} away from $z=0$.
 
@@
 %We comment that
+
+\subsection{Entropy of an isothermal dusty gas }
+%Having made the connection between the dust fraction (or dust-to-gas
+%ratio) of the mixture and an energy equation, we now define 
+We define the (dimensionless) entropy of the mixture as 
+\begin{align}
+   s \equiv \ln \frac{P}{\rho} = \ln{\left[c_s^2(1-\tepsilon)\right]}.  
+\end{align}
+We shall see that by defining the entropy this way, many of the 
+results concerning the stability of dusty gas will have identical form
+and interpretations to that in standard adiabatic hydrodynamics.
 
@@
 \end{align}
-as a measure of the local metalicity. The second approximation holds
-for $\epsilon_0\ll1$. 
+as a measure of the local metalicity, where $\Sigma_\mathrm{d}$ and
+$\Sigma_\mathrm{g}$ are the dust and gas surface densities,
+respectively. The second equality holds for $\epsilon_0\ll1$.
+
 %\Sigma_\mathrm{d}$ and the gas
@@
    = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p s}{\p z} -\frac{\p
-    P}{\p z}\frac{\p s}{\p r} \right),
+    P}{\p z}\frac{\p s}{\p r} \right). 
 \end{align}
-where
-\begin{align}
-   s \equiv \ln \frac{P}{\rho} = \ln{\left[c_s^2(1-\tepsilon)\right]} 
-\end{align}
-is a pseudo-entropy for the mixture.
-
 Writing Eq. \ref{vshear} in terms of the gas density and dust-to-gas
@@
 We can compare these sources by
-evaluating them, to order-of-magnitude, using the equilibrium
+evaluating them using the equilibrium
 solutions in \S\ref{steady_state}. We assume the disk is radially
@@
 where the second equality assumes $c_s=c_s(r)$ and the final equality
-applies to the equilibria in \S\ref{eqm}. We find 
+applies to the equilibria in \S\ref{eqm}. Thus,  
 \begin{align*}
@@
 \end{align*}
-Thus in PPDs where $\epsilon_0 = O(10^{-1}$---$10^{-2})$, 
-vertical buoyancy only has a modestly stabilizing effect.
+However, for thin dust layers such that $H_\epsilon \ll Hg$, 
+$\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.
+
+We see that $N_z^2\sim \epsilon \OmK^2$, while $r\p_z\Omega^2\sim
+h_\mathrm{g}\OmK^2$ for a thin disk where vertical shear is mainly due to the
+radial temperature gradient. Since buoyancy is stabilizing, we expect
+the dust-induced buoyancy forces to stabilize the disk against the VSI
+where $\epsilon \gtrsim h_\mathrm{g}$.
 

### 2016-08-04 paper.tex
minor edits at home

@@
 %dust settling does not lead to axisymmetric instability. 
-We briefly discuss potential applications of the 
-equivalence between isothermal dusty gas and standard hydrodynamics
-in providing thermodynamical interpretations of other dusty
+We briefly discuss potential applications of the dusty/adiabatic gas
+correspondence to provide thermodynamical interpretations of dusty
 instabilities in protoplanetary disks.  

### 2016-08-04 summary.tex
minor edits at home

@@
 dusty gas and an adiabatic, pure gas. 
-In the limit of perfect dust-gas coupling, the dust content of a gas 
+In the limit of perfect dust-gas coupling, the dust content of a fluid 
 parcel $\rhod/\rhog$ is conserved. This is analogous to entropy being 
@@
   induced by  dust-settling, ][]{lee10}. We also show that
-protoplanetary disks cannot have sharp edges in the dust-to-gas ratio,
-as this implies a sharp entropy gradient and may thus render the disk
-unstable.
+sharp edges in the dust-to-gas ratio cannot persist in protoplanetary
+disks, as these imply a sharp entropy gradient and may thus destablize
+the disk.
 
@@
 
-For locally isothermal disks $c_s=c_s(r)$ we generalize the vertical
+For locally isothermal disks $c^2_s=c_s^2(r)$ we generalize the vertical
 shear instability (VSI) to include dust. We find that dust-loading
@@
 \subsection{Dusty analogs of other gaseous instabilities}  
-
 %gi

### 2016-08-04 model.tex
minor edits at home

@@
 \begin{align}\label{vshear_split}
-  \mathcal{V} =  \frac{\left|r\p_z\Omega\right|_{\text{
+\frac{\left|r\p_z\Omega\right|_{\text{
         dust/gas gradient}}}{\left|r\p_z\Omega\right|_{\text{
@@
 vertical shear due to variations in the dust-to-gas ratio can be at
-most comparable to that due to the radial temperature gradient; and
-this only occurs for thin dust layers with $\delta^2\ll \epsilon$ (so
-that $\mathcal{V}\sim 1/|q|$). Otherwise, vertical shear is
-predominantly due to $\p_rc_s^2$.
+most comparable to that due to the radial temperature gradient. 
+This only occurs for thin dust layers with $\delta^2\ll \epsilon$ (so 
+the above ratio $\sim 1/|q|$). Otherwise, vertical shear is associated
+with $\p_rc_s^2$.
 
@@
 \end{align*}
-However, for thin dust layers such that $H_\epsilon \ll Hg$, 
+However, for thin dust layers such that $H_\epsilon \ll \Hg$, 
 $\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.
-
-We see that $N_z^2\sim \epsilon \OmK^2$, while $r\p_z\Omega^2\sim
-h_\mathrm{g}\OmK^2$ for a thin disk where vertical shear is mainly due to the
-radial temperature gradient. Since buoyancy is stabilizing, we expect
-the dust-induced buoyancy forces to stabilize the disk against the VSI
-where $\epsilon \gtrsim h_\mathrm{g}$.
 

### 2016-08-04 result.tex
minor edits at home

@@
 $h_\mathrm{g}=0.05$. These are fiducial values used in \citetalias{lin15}.
+
+We see that $N_z^2\sim \epsilon \OmK^2$, while $r\p_z\Omega^2\sim
+h_\mathrm{g}\OmK^2$ for a thin disk where vertical shear is mainly due to the
+radial temperature gradient. Since buoyancy is stabilizing, we expect
+the dust-induced buoyancy forces to stabilize the disk against the VSI
+where $\epsilon \gtrsim h_\mathrm{g}$.
 

### 2016-08-05 model.tex
minor changes

@@
 
-\subsection{Vertical shear}
+\subsection{Vertical shear}\label{vertshear}
 The mixture possess vertical shear. To see this, we eliminate $\Phi$
@@
 Since $|q|=O(1)$ in PPDs, Eq. \ref{vshear_split} indicates that
-vertical shear due to variations in the dust-to-gas ratio can be at
+vertical shear due to variations in the dust-to-gas ratio is at 
 most comparable to that due to the radial temperature gradient. 
 This only occurs for thin dust layers with $\delta^2\ll \epsilon$ (so
-the above ratio $\sim 1/|q|$). Otherwise, vertical shear is associated
-with $\p_rc_s^2$.
+the above ratio $\sim 1/|q|$). Otherwise, vertical shear is 
+predominantly associated with  $\p_rc_s^2$.
 
-\subsection{Vertical buoyancy}
+\subsection{Vertical buoyancy}\label{vbuoyancy}
 Having identified the entropy of an isothermal dusty gas, we find the
@@
 \end{align*}
-However, for thin dust layers such that $H_\epsilon \ll \Hg$, 
+However, for thin dust layers such that $H_\epsilon \ll Hg$, 
 $\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.
 
-%\begin{align}
  % N_z^2 

### 2016-08-05 result.tex
minor changes

--- 2016-08-04 result.tex
@@
 $h_\mathrm{g}=0.05$. These are fiducial values used in \citetalias{lin15}.
+
+\subsection{Qualitative expectations}
+\citetalias{lin15} find the appropriate way to compare 
+destabilizing vertical shear to stabilizing vertical buoyancy is
+$r\p_z\Omega/\OmK$ against $N_z^2/\OmK^2$. We see from \S\ref{vertshear}
+and \S\ref{vbuoyancy} that $|r\p_z\Omega|\sim q h_\mathrm{g}\OmK$ for a
+thin, gas dominated disk; while $N_z^2\sim \epsilon\OmK^2$. Thus we
+expect dust-induced buoyancy forces to stabilize the disk against the
+VSI where $\epsilon \gtrsim h_\mathrm{g}$.
+
+%\begin{align}
 
@@
 \end{figure}
+
+\subsection{Validity}
+We have considered $\tstop=0$ for simplicity. When $\tstop\neq0$, a 
+disk initialized as in \S\ref{eqm} will evolve 
+due to dust-gas relative drift. In particular, in the
+absence of turbulent/diffusive processes, dust particles settle to the
+midplane on a timescale $t_\mathrm{settle}\OmK \sim 1/\tstop\OmK$
+\citep{takeuchi02}. For the previous results to be valid (i.e. the
+basic state does not evolve significantly) we require VSI growth
+timescales $t_\mathrm{grow}\ll t_\mathrm{settle}$. Since the maximum
+VSI growth rate is $\sim h_\mathrm{g}\OmK$, we need stopping times
+such that 
+\begin{align}
+  \tstop\OmK \ll h_\mathrm{g}. 
+\end{align} 
+For thin PPD with $h_\mathrm{g}\sim 0.05$, our dusty VSI results are
+valid for $\tstop\OmK < O(10^{-2})$.

### 2016-08-05 linear.tex
started summary

@@
  adiabatic hydrodynamics \citep[e.g.][]{lubow93}. For simplicity
-we impose solid boundaries so that $\delta v_z(\pm\zmax)=0$ and
-$\zmax=5\Hgas$.
+we impose solid boundaries so that $\delta v_z(\pm\zmax)=0$.
 
@@
 using a pseudo-spectral code adapted from \cite{lin15}. Amplitudes 
-are expanded in Chebyshev polynomials up to order $N_z-1$. Our 
-standard resolution is $N_z=513$. We check results using
-Eq. \ref{vsi_check}.
+are expanded in Chebyshev polynomials up to order $N_z=512$. We check
+results using Eq. \ref{vsi_check}.
 

### 2016-08-05 model.tex

--- 2016-08-04 model.tex
@@
 
-\subsection{Vertical buoyancy}
+\subsection{Vertical buoyancy}\label{vbuoyancy}
 Having identified the entropy of an isothermal dusty gas, we find the
@@
 
-%\begin{align}
  % N_z^2 

### 2016-08-05 summary.tex
minor changes
more discussions

@@
 \subsection{Dusty analogs of other gaseous instabilities}
+
 %gi
 %rwi
+
+\subsubsection{Gravitational instability} %vertical structure 
+The addition of dust enhances gravitational
+instability (GI). This is because dust particles contribute to the
+total disk mass but not thermal pressure, which effectively lowers the
+disk temperature \citep[][]{thompson88}. For typical dust-loading
+$\epsilon, Z\ll1$, this effect is unimportant. However, if $\epsilon$ is
+large (e.g. due to dust settling) then the effective temperature may 
+be lowered to enable instability.
+
+As noted in \S\ref{loc_iso_eos}, dust settling causes the effective
+temperature to \emph{increase} away from the midplane. This contrasts
+to previous studies of GI in vertically stratified disks
+\citep[e.g.][]{mamat10, kim12,lin14c} where the temperature decreases
+from the midplane. While we expect only the total surface density and
+characteristic temperatures are relevant to stability 
+\citep{toomre64}, having a non-trivial vertical temperature
+structure, induced by dust, may modify the vertical structure of 3D
+waves and unstable modes. We will examine this problem in a following
+study.
+
+%see also
+%\S\ref{loc_iso_eos}
+%We thus expect the standard Toomre
+%parameter to be modified to $Q = c_s\sqrt{1-\tepsilon}\Omega/\pi
+%G\Sigma$.
+
+\subsubsection{Rossby wave instability}
+The Rossby wave instability (RWI) is a non-axisymmetric, 2D shear
+instability that operates in thin disks when it has radial structure
+\citep{lovelace99,li00}. These studies consider adiabatic pure gas and 
+show instability is possible if there is an extrema in the quantity
+\begin{align}
+ \eta = \frac{\kappa^2}{2\Omega\Sigma}\mathcal{S}^{2/\gamma},  
+\end{align} 
+where $\mathcal{S} = \Pi/\Sigma^\gamma$. Here, $\Sigma$ is the total
+surface density and $\Pi$ is the vertically integrated pressure.
+
+From the dusty/adiabatic gas equivalence, we deduce the corresponding
+condition for RWI in a (locally isothermal or polytroipc) dusty gas is
+also given by extrema in $\eta$ with $\mathcal{S} = \exp{s}$ (and
+$\tepsilon\to \Sigma_\mathrm{d}/\Sigma$). Since
+the dusty entropy $s$ reflects the dust distribution, we expect that
+RWI may also be triggered by extrema in  the dust-to-gas ratio.
+
+%gap edges. thin dust rings.
 

### 2016-08-05 paper.tex
more discussions
starting intro

@@
 %\usepackage{paralist}
-\shorttitle{untitled}
-\shortauthors{M-K.\ Lin}
+\shorttitle{Isothermal dusty gas dynamics}
+\shortauthors{M-K.\ Lin, A.N. Youdin}
 
@@
 \section{Introduction}
+%si, sgi as examples
+Protoplanetary disks are comprised of a mixture of gas and dust
+\citep{chiang10}. Although the overall dust content is typically small
+($\sim 1\%$ by mass), their presence can have profound impacts on the
+gas dynamics and vice versa.  
+%phenomena that depends on gas drag 
+%planet formation in dusty disks 
+%dust component is building block of planets
+%understanding dusty gas dynamics 
+%small amount of dust can have significant consequences
+
+%particles 
+Current state-of-the-art models of dusty protoplanetary disks
+directly simulate pure gas dynamics with explicit Lagragian dust
+particles.    
+%refs
+
+A simpler framework is to model the dust population as a pressureless
+fluid. The advantage is that standard hydrodynamics codes may be
+modified by adding a 
+%The two-fluid model is also suitable for analytic studies.
+
+%how to model dusty gas?
+
+%two fluid
+%one fluid
+
+We draw the physical and mathematical parallel between dusty gas and
+adiabatic pure gas. 
+%suitable for 
+%analytic studies
 
@@
 \input{result}
-%\input{linear_adia}
-%\input{numerical}
-%\input{application} 
 \input{summary}

### 2016-08-05 summary.tex
starting intro

@@
 
+\subsection{Implications for numerical simulations}
+The dusty/adiabatic gas equivalence means that pure
+gas dynamics codes can be applied to simulate locally
+isothermal/polytropic dusty gas with minimal modification. In fact, no
+modification is needed  
+for strictly isothermal or polytropic gas ($c_s$ or $K$ being
+constant, respectively) perfectly-coupled to dust. The usual 
+adiabatic energy equation is effectively a dust continuity
+equation. When $c_s$ or $K$ is not constant and/or $\tstop\neq0$ one
+can simply add corresponding source terms in energy equation. In the
+large but finite coupling case, this source term is analogous to
+radiative diffusion \citep{price15}, which is also standard in many
+hydrodynamics codes.
+
 \subsection{Dusty analogs of other gaseous instabilities}
@@
 total disk mass but not thermal pressure, which effectively lowers the
-disk temperature \citep[][]{thompson88}. For typical dust-loading
+disk temperature \citep[][]{thompson88,shi13}. For typical dust-loading 
 $\epsilon, Z\ll1$, this effect is unimportant. However, if $\epsilon$ is
@@
 structure, induced by dust, may modify the vertical structure of 3D
-waves and unstable modes. We will examine this problem in a following
+waves and unstable modes. We will examine this linear problem in a following 
 study.

### 2016-08-08 paper.tex

--- 2016-08-04 paper.tex
@@
 \section{Introduction}
+%si, sgi as examples
+Protoplanetary disks are comprised of a mixture of gas and dust
+\citep{chiang10}. Although the overall dust content is small
+($\sim 1\%$ by mass), their presence can have profound impacts on the 
+gas dynamics and vice versa. Dust-gas drag introduces a 
+wide range of new phenomena that are absent in pure gas
+disks. Important examples include the streaming instability
+\citep{youdin05a,youdin07b,johansen07} and secular gravitational
+instabilities \citep{ward00,youdin11,takahashi14} that lead to
+planetesimal formation. More recently, new dusty instabilities have
+also been uncovered \citep{loren15,loren16,lamb16}
+
+%Understanding dusty gas dynamics is
+%therefore essential to any planet formation theory. 
+%phenomena that depends on gas drag 
+%planet formation in dusty disks 
+%dust component is building block of planets
+%understanding dusty gas dynamics 
+%small amount of dust can have significant consequences
+
+%particles 
+Current state-of-the-art models of dusty protoplanetary
+disks directly simulate gas dynamics with explicit
+Lagragian dust particles \citep{nelson10,bai10,johansen11,yang14,zhu14,gibbons15,simon16,baruteau16}. In this
+approach, the equations of motion for each particle is solved
+including dust-gas drag, the strength of which is measured 
+by a stopping time $\tstop$ --- the characteristic timescale for the
+relative velocity between gas and dust to decay. Thus small $\tstop$
+corresponds to tightly-coupled particles.
+
+%refs
+
+Another approach is to model the dust population as a continous, pressureless
+fluid
+\citep{paardekooper06b,meheut12,laibe12,loren14,fu14b,surville16}. The equations
+hydrodynamics are solved for two fluids: the gas and dust with density and
+velocity $\rhog$, $\bm{v}_\mathrm{g}$ and $\rhod$,
+$\bm{v}_\mathrm{d}$, respectively. Source terms are introduced into
+the momentum/energy equations to model dust-gas drag. This approach
+can take advantage of, and adapt, already-developed methods/codes for
+simulating pure gas dynamics.
+
+%refs
+
+It is possible to re-formulate the two-fluid equations into that for a
+single fluid of for the total density $\rho$ and center-of-mass
+velocity $\bm{v}$ of the dusty gas mixture
+\citep{laibe14}. The standard hydrodynamic equations for
+$\rho$ and $\bm{v}$ are supplemented by evolutionary equations for
+the dust-to-gas ratio $\rhod/\rhog$ and the dust-gas velocity
+difference $\Delta\bm{v}$. The latter also contributes to
+additional source terms in the momentum and energy equations.  
+The one-fluid approach is appropriate for 
+simulating gas tightly coupled to dust particles with small $\tstop$.  
+This can make the two-fluid or Lagragian particles approach
+numerically challanging; but is straight forward to model in the
+one-fluid framework \citep{price15}.
+
+The single fluid formalism also has the important advantage that it is
+amenable to standard analytical methods
+\citep{youdin05a,jacquet11}. This is the necessary  approach to gain physical insight.  
+In this work, we are motivated to develop a general analytical
+framework to  study the stability of dusty protoplanetary disks by
+drawing on the physical and mathematical parallel between dusty gas and standard
+hydrodynamics. This is possible in certain limits of astrophysical
+interest.
+
+We consider tightly coupled dust particles with $\tstop\Omega\ll
+1$, where $\Omega$ is the disk orbital frequency. We may then use the
+`terminal velocity approximation' to set $\Delta \bm{v} = \tstop\nabla
+P/\rhog$, where $P$ is the gas pressure \citep{youdin05a,
+  jacquet11,laibe14}. We also consider irradiated protoplanetary disks 
+in which the temperature is set externally \citep{chiang97,stam08}. We
+can then adopt a locally isothermal equation of state $P = c_s^2\rhog$
+where $c_s$ is a prescribed sound-speed profile.  This physically
+represents short cooling times \citep{lin15}.
+
+Under these approximations, we show that the one-fluid equations of 
+isothermal dusty gas dynamics are equivalent to that of adiabatic, pure gas
+dynamics. This is because tightly-coupled dust particles are almost
+entrained in the gas; and this property is analogous to the advection
+of entropy following an adiabatic fluid. We will formally relate the
+entropy of isothermal dusty gas to its dust content. Finite 
+dust-gas drag can be re-interpreted as heating/cooling: the relative
+motion between dust and gas causes a fluid parcel to exchange dust
+particles with its surroundings; an effect similar to heat exchange
+between a gas parcel and its surroundings.
+
+%suitable for 
+%analytic studies 
+%more importantly, can understood
+
+By re-casting the dusty problem into identical form to standard
+hydrodynamics, we can effortlessly generalize previous stability
+results for pure gas disks to the dusty case. We demonstrate this for
+perfectly coupled dust in strictly and locally isothermal disks. We
+show that sharp dust edges may destabilize protoplanetary disks.  
+We also study the effect of dust-loading on the vertical shear
+instability \citep[VSI,][]{nelson13,lin15,barker15}. We provide 
+thermodynamical interpretations of our dusty stability calculations.
+
+This paper is organized as follows. In \S\ref{setup} we list the
+equations describing tightly-coupled dusty gas. We show they are in
+fact the usual hydrodynamic equations, including a precise 
+definition of the entropy isothermal dusty gas. We describe approximate 
+equilibrium solutions in \S\ref{eqm}. In \S\ref{limits} we discuss
+general stability properties of the dusty problem in various
+limits. We use these results to constraint the sharpness of dust edges
+in isothermal disks. In \S\ref{linear_problem} we formulate an
+explicit stability calcultion to study the dusty VSI, for which we
+present numerical results \S\ref{results}. We summarize and discuss
+future applications of the dusty/adiabatic gas equivalence in
+\S\ref{summary}.
 

### 2016-08-08 summary.tex
rough intro

@@
-\section{Summary and discussion}
+\section{Summary and discussion}\label{summary}
 In this paper we highlight the similarity between an isothermal 
@@
 
-%\subsection{Implication for numerical simulations}
+\subsection{Thermodynamic view of dusty instabilities}
 
-\subsection{Thermodynamic interpretation of dusty instabilities}
+%streaming instability 
+Finite dust-gas drag can lead to the streaming instability (SI) in the
+presence of a background radial pressure gradient \citep{youdin05a}.    
+SI can also be recovered from the one-fluid framework
+\citep{laibe14}. \cite{jacquet11} explains the essence of SI as dust
+accumulation at pressure maxima (from terminal velocity
+approximation), which back-reacts on the gas flow in 
+such a way to strengthen said maxima.
 
-%\subsection{Caveats and future directions}
-%gi
+% Now consider our energy formulation of the
+% dusty problem, Eq. \ref{eff_energy}. Assume constant $c_s$
+% and that the dominant contribution to the energy source
+% $\mathcal{C}$ is associated with $\nabla^2P$ \citep[which is
+% responsible for SI, ][]{jacquet11}. Then 
+% \begin{align*}
+% \frac{DP}{Dt} \simeq \frac{P}{\rho}\frac{D\rho}{D t} + c_s^2\tepsilon \tstop
+%   \nabla^2 P.
+% \end{align*} 
+% At pressure maxima, dust-drag acts as a cooling term. However, if
+% dust accumulation into this maximum leads to a sufficiently large
+% compression, there may be a net heating, enhancing the pressure
+% maximum, and a positive feedback established.
+
+% This picture relies on unstable correlations between pressure
+% perturbations and fluid compression, as pointed out by
+% \cite{jacquet11}. Indeed, we see from Eq. \ref{thermal_instabilities}
+% that instabilities due to drag are attributed to  
+% \begin{align*}
+%   \left(\nabla\cdot\dd\bm{v}^*\right) \delta \mathcal{C}.
+% \end{align*} 
+% Thus instabilities may arise if $\nabla\cdot\delta \bm{v}$ is suitably
+% correlated to the perturbed drag force $\delta \mathcal{C}$ (which for
+% SI is due to pressure perturbations). This is analogous  
+% to thermal instabilities in gas dynamics where $\mathcal{C}$
+% represents non-adiabatic heating/cooling and may lead to
+
+% \citep{latter06},
+
+%loren's instability

### 2016-08-09 paper.tex
minor intro edits

@@
 disks. Important examples include the streaming instability
-\citep{youdin05a,youdin07b,johansen07} and secular gravitational
-instabilities \citep{ward00,youdin11,takahashi14} that lead to
-planetesimal formation. More recently, new dusty instabilities have
-also been uncovered \citep{loren15,loren16,lamb16}
+\citep{youdin05a,youdin07b,johansen07}, secular gravitational
+instabilities \citep{ward00,youdin11,michikoshi12,takahashi14} and
+Kelvin-Helmholtz instabilities
+\citep{chiang08,barranco09,lee10}. These lead to structure and
+planetesimal formation, as well as turbulence. More recently, new
+dusty instabilities have also been uncovered
+\citep{loren15,loren16,lamb16}
 
@@
 including dust-gas drag, the strength of which is measured 
-by a stopping time $\tstop$ --- the characteristic timescale for the
-relative velocity between gas and dust to decay. Thus small $\tstop$
+by a stopping time $\tstop$ --- the decay timescale for the 
+relative velocity between gas and dust. Thus small $\tstop$
 corresponds to tightly-coupled particles.
@@
 simulating gas tightly coupled to dust particles with small $\tstop$.  
-This can make the two-fluid or Lagragian particles approach
-numerically challanging; but is straight forward to model in the
+This can be numerically challenging for the two-fluid or Lagragian
+particles approach; but is straight forward to model in the
 one-fluid framework \citep{price15}.
@@
 in which the temperature is set externally \citep{chiang97,stam08}. We
-can then adopt a locally isothermal equation of state $P = c_s^2\rhog$
+may then adopt a locally isothermal equation of state $P = c_s^2\rhog$
 where $c_s$ is a prescribed sound-speed profile.  This physically
@@
 equations describing tightly-coupled dusty gas. We show they are in 
-fact the usual hydrodynamic equations, including a precise 
-definition of the entropy isothermal dusty gas. We describe approximate 
+fact the usual hydrodynamic equations. This leads to a precise 
+definition of the entropy of an isothermal dusty gas. We describe approximate 
 equilibrium solutions in \S\ref{eqm}. In \S\ref{limits} we discuss
@@
 explicit stability calcultion to study the dusty VSI, for which we
-present numerical results \S\ref{results}. We summarize and discuss
+present numerical results in \S\ref{results}. We summarize and discuss
 future applications of the dusty/adiabatic gas equivalence in

### 2016-08-12 appendix.tex
started summary

@@
 
-\section{Conversion formulae}
-The gas density $\rhog$ and dust-to-gas ratio $\epsilon$ is related
-to the total density $\rho$ and dust-fraction $\tepsilon$ by
+%\section{Conversion formulae}
+%The gas density $\rhog$ and dust-to-gas ratio $\epsilon$ is related
+%to the total density $\rho$ and dust-fraction $\tepsilon$ by
+%\begin{align}
+%  \ln{\rho} &= \ln{\rhog} + \ln{\left(1 + \epsilon\right)},\\
+%  \nabla\ln{\rho} &= \nabla\ln{\rhog} + \frac{\nabla\epsilon}{1+
+%    \epsilon},\\
+%  \nabla^2\ln{\rho} &= \nabla^2\ln{\rhog} +
+%  \frac{\nabla^2\epsilon}{1+\epsilon} -
+%  \frac{\left|\nabla\epsilon\right|^2}{\left(1+\epsilon\right)^2}, 
+%\end{align}
+%and
+%\begin{align}
+%  \tepsilon &= \frac{\epsilon}{1+ \epsilon},\\
+% \nabla\tepsilon & =
+% \frac{\nabla\epsilon}{\left(1+\epsilon\right)^2},\\
+% \nabla^2\tepsilon &=
+% \frac{\nabla^2\epsilon}{\left(1+\epsilon\right)^2} -
+% \frac{2\left|\nabla\epsilon\right|^2}{\left(1+\epsilon\right)^3}. 
+%\end{align}
+
+\section{Dispersion relation for unstratified disks}
+Unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
+0$. The background $\p_rP/\rho$ and $\tepsilon$ are constant. 
+Fourier analyze in $r$ and $z$ so that $\p_z\to \ii k_z$ and
+$\p_r\to \ii k_x$ when acting on perturbations. Denote $|\bm{k}|^2
+\equiv k_x^2 + k_z^2$. Neglect variations in
+background when compared with that of perturbations. Linearized
+equations are, after eliminating the azimuthal velocity:
+
 \begin{align}
-  \ln{\rho} &= \ln{\rhog} + \ln{\left(1 + \epsilon\right)},\\
-  \nabla\ln{\rho} &= \nabla\ln{\rhog} + \frac{\nabla\epsilon}{1+
-    \epsilon},\\
-  \nabla^2\ln{\rho} &= \nabla^2\ln{\rhog} +
-  \frac{\nabla^2\epsilon}{1+\epsilon} -
-  \frac{\left|\nabla\epsilon\right|^2}{\left(1+\epsilon\right)^2}, 
+  \ii \sigma W &=\nabla\cdot \bm{\dd v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\\
+    \sigma^2 \dd v_x &= \kappa^2 \dd v_x - \ii \sigma F_r W +
+    k_x\sigma Q,\\
+  -\ii\sigma\dd v_z &= -\ii k_zQ,\\
+\ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \bm{\dd v}   -
+  \frac{\dd \mathcal{C}}{\rho}. 
 \end{align}
-and
+The artifical factor $\zeta = 1$  is inserted to keep track of the
+left-hand-side of the energy equation. The linearized dust diffusion
+function (Appendix \ref{lin_dust}), 
+under the above approximations, is 
 \begin{align}
-  \tepsilon &= \frac{\epsilon}{1+ \epsilon},\\
- \nabla\tepsilon & =
- \frac{\nabla\epsilon}{\left(1+\epsilon\right)^2},\\
- \nabla^2\tepsilon &=
- \frac{\nabla^2\epsilon}{\left(1+\epsilon\right)^2} -
- \frac{2\left|\nabla\epsilon\right|^2}{\left(1+\epsilon\right)^3}. 
+-\frac{\dd \mathcal{C}}{\rho}  = \tstop c_s^2 \left[\ii k_x F_r\left( 1-2\tepsilon\right)
+  W + \tepsilon |\bm{k}|^2 Q\right]. 
 \end{align}
+
+Eliminate the velocity perturbations to obtain
+\begin{align}
+  \left(\ii \zeta \sigma - \tstop c_s^2 \tepsilon |\bm{k}|^2\right)Q &=
+  \ii \left[
+  \frac{P}{\rho}\sigma + k_x\tstop c_s^2 F_r\left(1 -
+  2\tepsilon\right)\right]W. \\
+   \sigma^2 \left( \kappa^2 - \sigma^2 - \ii k_x F_r\right)W &=
+    \left(k_z^2\kappa^2 - \sigma^2|\bm{k}|^2\right)Q. 
+\end{align}
+Futher eliminate $Q$ and $W$ to obtain the dispersion relation
+\begin{align}
+  &\frac{\zeta}{c_s^2|\bm{k}|^2}\sigma^5 + \ii \tepsilon \tstop
+  \sigma^4 - \left[ \frac{\zeta}{c_s^2|\bm{k}|^2} \left(\kappa^2 - \ii
+  k_x F_r\right) + \left(1 - \tepsilon\right)\right]\sigma^3 - \ii
+  \tstop \left[\tepsilon \kappa^2 - \ii k_x F_r\left( 1 -
+  \tepsilon\right)\right]\sigma^2 \notag\\ 
+  &+ \left(1 -
+  \tepsilon\right)\left(\frac{k_z\kappa}{|\bm{k}|}\right)^2\sigma +
+  k_x \tstop F_r\left(1 - 2\tepsilon\right)
+  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0,
+\end{align}
+where the equation of state $P = c_s^2\left( 1 - \tepsilon\right)\rho$
+was used. 
+If we now set $\zeta = 0$ or consider $c_s^2\to \infty$ to obtain the
+incompressible limit, and define $\tau_\mathrm{s} \equiv
+\tstop/\left(1 - \tepsilon\right)$, we obtain: 
+\begin{align*}
+\ii \tepsilon \tstop
+  \sigma^4 - \sigma^3 - 
+  \tau_\mathrm{s} \left[ \ii \tepsilon \kappa^2 + k_x F_r\left( 1 -
+  \tepsilon\right)\right]\sigma^2 
+  + \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2\sigma + 
+  k_x \tau_\mathrm{s} F_r\left(1 - 2\tepsilon\right)
+  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0,
+\end{align*}
+as derived by \cite{jacquet11} and \cite{laibe14} for the streaming
+instability.

### 2016-08-12 paper.tex

@@
 By re-casting the dusty problem into identical form to standard
-hydrodynamics, we can effortlessly generalize previous stability
+hydrodynamics, we can follow standard procedures to generalize previous stability 
 results for pure gas disks to the dusty case. We demonstrate this for

### 2016-08-13 linear.tex
rough cut. still missing last discussion section and finite ts VSI results

@@
-\section{Linear problem for perfectly coupled dust}\label{linear_problem}
-We calculate explicit solutions to the linearized
-equations in the limit of perfectly coupled dust ($\tstop=0$), so that
-the equilibria defined in \S\ref{eqm} are exact steady states. We 
-consider radially-localized axisymmetric disturbances of the form  
+\section{Quasi-global linear problem}\label{linear_problem}
+We consider radially-localized axisymmetric disturbances of the form  
 \begin{align}
@@
 the linearized equations for 
-locally isothermal, perfectly-coupled dusty gas with the pressure
+locally isothermal dusty gas with the pressure
 equation in place of the dust-fraction
@@
   \dd v_r \p_r\ln{\rho} + \dd v_z\p_z\ln{\rho},\label{lin_mass}\\
-  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
-  \delta\bm{F}\cdot\hat{\bm{r}},\label{lin_xmom}\\
+  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi 
+% +  \delta\bm{F}\cdot\hat{\bm{r}}
+- W F_r - \ii k Q,\label{lin_xmom}\\
   \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
     v_\phi}{\p z}\dd v_z, \label{lin_ymom}\\
-  -\ii\sigma\dd v_z &=  \delta\bm{F}\cdot\hat{\bm{z}},\label{lin_zmom}\\
+  -\ii\sigma\dd v_z &= - W F_z - \left[Q^\prime + Q
+    \left(\ln{\rho}\right)^\prime\right]  %\delta\bm{F}\cdot\hat{\bm{z}} 
+,\label{lin_zmom}\\
   \ii\sigma Q &= \frac{P}{\rho}\left(\ii k \dd v_r + \dd
@@
                 &\phantom{=}-\frac{P}{\rho} \dd v_r\p_r
-               \ln{c_s^2}, \label{lin_energy} 
+               \ln{c_s^2} %, \label{lin_energy} 
 %+ \dd v_z \p_z\ln{c_s^2}\right),\label{lin_energy} 
-%               - \frac{\dd\mathcal{C}}{\rho},\label{lin_energy} 
+               - \frac{\dd\mathcal{C}}{\rho},\label{lin_energy} 
 \end{align}  
-where $^\prime \equiv \p_z$ and $\dd\bm{F}$ is the linearized pressure
-force, given in Appendix \ref{lin_press}. Note that we have assumed a
-temperature profile that only depends on $r$.
+where $^\prime \equiv \p_z$ and $\bm{F} \equiv -\nabla P/\rho$. 
+The linearized dust-duffusion function 
+$\dd\mathcal{C}$ is given in  Appendix \ref{lin_dust}. 
+% and $\dd\bm{F}$ is the linearized pressure
+%force, given in Appendix \ref{lin_press}. 
+Note that we have assumed a temperature profile that only depends on $r$.
 
@@
 evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
-is retained.  Two boundary conditions are needed for the $\tstop=0$
-problem. This is expected since this problem is exactly equivalent to
- adiabatic hydrodynamics \citep[e.g.][]{lubow93}. For simplicity
-we impose solid boundaries so that $\delta v_z(\pm\zmax)=0$.
+is retained. Numerical solutions are generally required for the 
+verticall-stratified problem, as we pursue below.
 
-We solve the linearized equations as a generalized eigenvalue problem 
-using a pseudo-spectral code adapted from \cite{lin15}. Amplitudes 
-are expanded in Chebyshev polynomials up to order $N_z=512$. We check
-results using Eq. \ref{vsi_check}.
+However, in the special case of an unstratified disk (neglecting the vertical
+component of the stellar gravity), we may also Fourier analyze in $z$
+to obtain a local dispersion relation. This is done in 
+Appendix \ref{compressible_streaming}. There we show the above linearized
+equations give the streaming instability in the appropriate limit 
+\citep{youdin05a,jacquet11}.
 

### 2016-08-13 model.tex
more discussions
rough intro
doing last section
rough cut. still missing last discussion section and finite ts VSI results

--- 2016-08-05 model.tex
@@
 
-\subsection{Locally isothermal equation of state} 
+\subsection{Locally isothermal equation of state}\label{loc_iso_eos}
 We adopt 
@@
 
-\section{Limiting behaviours}
+\section{Limiting behaviours}\label{limits}
 
@@
   \rho |\dd v_z|^2 D\phantom{\frac{1}{1}}\right. \notag\\
-&\phantom{===}  \left. + \frac{1}{P}\left|\nabla\cdot\left(P\dd
-  \bm{v}\right)\right|^2\right]dV  -\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV \notag\\
+&\phantom{===}  \left. + \frac{1}{P}\Bigl\lvert \nabla\cdot\left(P\dd
+  \bm{v}\right)\Bigr\rvert^2\right]dV  -\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV \notag\\
 &\phantom{===}
@@
 coupled to dust,  
+\begin{align}
+  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\tepsilon &> 0,\label{dusty_solberg1}  \\
+  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\tepsilon}{\p
+    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \tepsilon}{\p r} \right) & > 0, \label{dusty_solberg2}
+\end{align} 
 %\begin{align}
-%  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\tepsilon &> 0,\label{dusty_solberg1}  \\
-%  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\tepsilon}{\p
-%    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \tepsilon}{\p r} \right) & > 0, \label{dusty_solberg2}
+%  \kappa^2 - \frac{1}{\rho}\nabla P \cdot \nabla s &> 0, \label{dusty_solberg1}    \\
+%  -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
+%  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, \label{dusty_solberg2}
 %\end{align} 
-\begin{align}
-  \kappa^2 - \frac{1}{\rho}\nabla P \cdot \nabla s &> 0, \label{dusty_solberg1}    \\
-  -\frac{1}{\rho}\frac{\p P}{\p z} \left(\kappa^2 \frac{\p s}{\p z} -
-  r\frac{\p\Omega^2}{\p z}\frac{\p s}{\p r}\right) &>0, \label{dusty_solberg2}
-\end{align} 
 where %$s = \ln{P/\rho}$, 
 $\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)$ is the 
-square of the epicylic frequency, and we assume $\kappa^2>0$.  
-Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} has the identical form
-to the Solberg-Hoiland criteria for adiabatic gas dynamics 
-\citep[e.g.][]{tassoul78}, except here the entropy $s$ is really a
-representation of the dust distribution via $s=\ln{\left[c_s^2\left(1
-    - \tepsilon\right)\right]}$.
+square of the epicylic frequency. We assume $\kappa^2>0$.  
+Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} are in fact the 
+the Solberg-Hoiland criteria for axisymmetric stability of adiabatic
+gas if one inserts $C_p\ln{\left(1 -
+    \tepsilon\right)}$ for the entropy in the standard
+expression for said criteria \citep[e.g.][ where $C_p$ is the heat
+  capacity]{tassoul78}.
+
+%except here the entropy $s$ is really a
+%representation of the dust distribution via $s=\ln{\left[c_s^2\left(1
+%    - \tepsilon\right)\right]}$.
 
@@
 
-\subsection{Locally isothermal gas perfectly coupled to dust}\label{dusty_vsi_int}
-When $c_s(r,z)$ is non-uniform but $\tstop=0$, Eq. \ref{int_rel}
-indicates that $\sigma^2$ is generally complex, implying instability.  
-Writing $\sigma = 
--\omega + \ii s$ where $\omega$ and $s$ are the real frequency and
-growth rates, respectively, we have 
-\begin{align}
-  s = \frac{\imag\int P
-  \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV}{2\omega\int\rho\left(|\dd
-    v_r|^2 + |\dd v_z|^2\right)dV}. \label{vsi_check}
-\end{align} 
-This is classic VSI caused by vertical shear arising from a radial
-temperature gradient \citep{nelson13,barker15,lin15}. We compute
-explicit solutions in this limit in
-\S\ref{linear_problem}---\ref{results}.
-
 %This is responsible for 
@@
 
-\subsection{Strictly isothermal gas well-coupled to dust} 
-Now consider constant $c_s$ but $\tstop\neq0$. Then Eq. \ref{int_rel}
-gives 
+\subsection{Thermodynamic view of dusty instabilities}\label{dust_work}
+Now consider constant $c_s$ but $\tstop\neq0$. 
+%When $c_s(r,z)$ is non-uniform but $\tstop=0$, 
+Eq. \ref{int_rel} indicates that $\sigma^2$ is generally complex, so
+we may have growing oscillations. Writing $\sigma = -\omega + \ii s$
+where $\omega$ and $s$ are the real frequency and growth rates, 
+respectively, the imaginary part of Eq. \ref{int_rel} gives 
 \begin{align}
   s = \frac{\imag\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV}{2\omega\int\rho\left(|\dd
-    v_r|^2 + |\dd v_z|^2\right)dV}, 
+    v_r|^2 + |\dd v_z|^2\right)dV}, \label{thermal_instability}
 \end{align}
@@
 %stars \citep{latter06}.  
-which represent instabilities due to large-but-finite dust-gas
-friction. Our formalism gives these dusty instabilities a 
-thermodynamical interpretation because the dust-gas friction enters the problem
-as an energy source term \citep[cf. thermal instabilities in
-  stars, see also][]{latter06}. 
-This includes the streaming instability in
-the strong-drag limit \citep{youdin05a, jacquet11, laibe14}.
+which represent instabilities due to dust-gas friction. 
+%Our formalism
+%gives these dusty instabilities a thermodynamical interpretation. 
+% We further discuss this concept in \S\ref{dust_as_thermo}. 
+%In \S\ref{dust_work} we showed that instabilities may 
+%arise from the term proportional $\imag\left( \nabla\cdot\dd\bm{v}^*
+%\delta C\right)$. 
+ Since drag ($\delta \mathcal{C}$) appears as a source term in our
+effective energy equation, the quantity 
+$\imag\left[\left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}\right]$
+may be thought of as a correlation between compression/expansion and  
+heating/cooling.
+
+It is well-known that such correlations can lead to pulsational
+instabilities in stars \citep{cox67}. We thus suggest an interpretation
+of dust-drag instabilities in disks in a similar way. We give a qualitative
+description below, based on the more rigorous treatments of stellar 
+pulsational instabilities by 
+\cite{cox67} and lecture notes from \cite{samadi15} and
+J. Christensen-Dalsgaard\footnote{\url{http://astro.phys.au.dk/$\sim$jcd/oscilnotes/}}.
+
+%and disks
+%\cite{kato78}.
+
+Consider the Lagragian perturbation $\Delta$ of a variable $f$, 
+related to the Eulerian perturbation by $\Delta f = \delta f + 
+\bm{\xi}\cdot\nabla f$, where $\bm{\xi}$ is the Lagragian  
+displacement ($\dd v_{x,z} = -\ii\sigma\xi_{x,z}$). %Note that $\Delta
+%\mathcal{C} = \delta \mathcal{C}$ for a uniform background $C$ as
+%required for equilibrium. 
+Then Eq. \ref{lin_mass_full}---\ref{lin_energy_full} imply  
+\begin{align} 
+  \imag \left(\delta \mathcal{C}
+  \nabla\cdot\dd \bm{v}^*\right) = 
+  -|\sigma|^2\imag \left(\Delta P 
+  \frac{\Delta\rho^*}{\rho}\right). 
+\end{align}
+This reflects `$TdS  =  PdV$' from thermodynamics, if the 
+internal energy is constant. That is, dust-gas drag provides 
+heating ($TdS$) to do $PdV$ work. Whether positive or negative work is 
+done depends on the correlation between the pressure and density
+perturbations, as follows.
+
+The full energy equation is 
+\begin{align*}
+  \frac{DP}{Dt} = \frac{P}{\rho}\frac{D\rho}{D t} + \mathcal{C}. 
+\end{align*}
+Dust-drag causes a phase difference between the 
+pressure and density evolution following a parcel. Now, consider a 
+fluid element at maximum density during an oscillation cycle, 
+so that $D\rho/ D t = 0$. If at this point dust-gas drag provides an
+effective heating ($\mathcal{C}>0$), then the fluid element is
+experiencing increasing pressure, $DP/D t>0$. It will attain
+pressure maximum \emph{after} reaching density maximum. This implies 
+pressure maximizes while the parcel is expanding: it will do 
+positive work (the enclosed area by its `$P$-$V$' curve  is 
+positive). This will increase the oscillation amplitude.
+
+Indeed, \cite{jacquet11} explains the essence of the streaming 
+instability in dusty protoplanetary disks as dust accumulation
+(and hence compression) at a pressure bump which back-reacts on the
+gas flow in such a way to strengthen the pressure bump. A phase
+difference was also noted in the numerical calculations of
+\cite[][although they consider the full two-fluid system]{youdin07b}.
+
+We suggest that a useful way to understand  
+dust-drag instabilities in protoplanetary disks may be to re-cast the  
+problem into that of thermodynamics, since such instabilities have 
+already been well-studied in the stellar context.
+
+\subsection{Locally isothermal gas perfectly coupled to dust}\label{dusty_vsi_int}
+If $c_s(r,z)$ is non-uniform but $\tstop=0$, Eq. \ref{int_rel} 
+gives  
+\begin{align}
+  s = \frac{\imag\int P
+  \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV}{2\omega\int\rho\left(|\dd
+    v_r|^2 + |\dd v_z|^2\right)dV}. \label{vsi_check}
+\end{align} 
+This is classic VSI caused by vertical shear arising from a radial
+temperature gradient \citep{nelson13,barker15,lin15}. We compute
+explicit solutions in this limit in \S\ref{results}.

### 2016-08-13 paper.tex
rough cut. still missing last discussion section and finite ts VSI results

@@
 \usepackage{paralist}
-%\usepackage{url}
+\usepackage{url}
 \usepackage{bm}
+%\usepackage{hyperref}
 %\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
@@
 
-\title{On the entropy of 
-  isothermal dusty gas: application to protoplanetary disks}  
+%\title{On the entropy of 
+%  isothermal dusty gas: application to protoplanetary disks}  
+\title{A thermodynamic view of dusty protoplanetary disks:
+  axisymmetric stability}  
 \author{Min-Kai Lin\altaffilmark{1} and Andrew N. Youdin}
@@
   particles almost follows the gas flow. 
-  %Tightly-coupled dust particles
-  % in protoplanetary disks almost follow the gas flow. 
   This is analogous to the 
@@
   to 
-\begin{inparaenum}[1)]
-\item
-  identify the entropy of isothermal dusty gas; and 
-\item 
-  generalize previous, well-known gaseous instabilities in
-  protoplanetary disks to include dust dynamics. 
-\end{inparaenum}
-This includes the axisymmetric stability of strictly isothermal dusty
-gas, allowing us to constrain the sharpness of 
-dust edges that can persist in protoplanetary disks. We also develop 
+%\begin{inparaenum}[1)]
+%\item
+%  identify the entropy of isothermal dusty gas; and 
+%\item 
+  generalize previous, well-known gaseous instabilities
+  to dusty protoplanetary disks.
+%\end{inparaenum}
+%This includes the axisymmetric stability of strictly isothermal dusty
+%gas, allowing us to constrain the sharpness of 
+%dust edges that can persist in protoplanetary disks. 
+We show that strictly isothermal, non-self-gravitating, axisymmetric
+disks perfectly coupled to dust are generally stable, however thin the
+dust layer is.  %except possibily in the presence of sharp 
+%edges in radial dependence of the dust-to-gas ratio. 
+We also develop 
 the dusty vertical shear instability (VSI) in locally isothermal
@@
 wavelengths, as well as `surface modes' concentrated near the
-disk boundaries. These results are consistent with our definition of
+disk boundaries. These results can be understood in terms of 
 the buoyancy frequency of a isothermal dusty disk. 
@@
 %dust settling does not lead to axisymmetric instability. 
-We briefly discuss potential applications of the dusty/adiabatic gas
-correspondence to provide thermodynamical interpretations of dusty
-instabilities in protoplanetary disks.  
+We briefly discuss potential thermodynamical interpretations of 
+dust-gas drag dynamics, for example the streaming instability 
+as a manifestation of pulsational instabilities in stars. 
 \end{abstract}
@@
 explicit stability calcultion to study the dusty VSI, for which we
-present numerical results in \S\ref{results}. We summarize and discuss
-future applications of the dusty/adiabatic gas equivalence in
-\S\ref{summary}.
+present numerical results in \S\ref{results}. We 
+discuss future applications of the dusty/adiabatic gas equivalence in 
+\S\ref{discussion} before summarizing in \S\ref{summary}.
 

### 2016-08-13 result.tex
rough cut. still missing last discussion section and finite ts VSI results

@@
 
-When $\tstop=0$ we may construct exact dusty equilibra by specifying
-the dust-to-gas ratio, as described in \S\ref{eqm}. We use this to
+%When $\tstop=0$ we may construct exact dusty equilibra by specifying
+%the dust-to-gas ratio, as described in \S\ref{eqm}.
+
+We calculate explicit solutions to the linearized
+equations in the limit of perfectly coupled dust ($\tstop=0$), so that
+the equilibria defined in \S\ref{eqm} are exact steady states. 
+We use this to
 examine the effect of dust on classic VSI, defined to be driven by the
@@
 $h_\mathrm{g}=0.05$. These are fiducial values used in \citetalias{lin15}.
+
+Two boundary conditions are needed for the $\tstop=0$
+problem. This is expected since this problem is exactly equivalent to
+adiabatic hydrodynamics \citep[e.g.][]{lubow93}. For simplicity we
+impose solid boundaries so that $\delta v_z(\pm\zmax)=0$.  
+We solve the linearized equations as a generalized eigenvalue problem 
+using a pseudo-spectral code adapted from \cite{lin15}. Amplitudes 
+are expanded in Chebyshev polynomials up to order $N_z=512$. We check
+results using Eq. \ref{vsi_check}.
 

### 2016-08-13 summary.tex
rough cut. still missing last discussion section and finite ts VSI results

--- 2016-08-05 summary.tex
@@
-\section{Summary and discussion}
-In this paper we highlight the similarity between an isothermal 
-dusty gas and an adiabatic, pure gas. 
-In the limit of perfect dust-gas coupling, the dust content of a fluid 
-parcel $\rhod/\rhog$ is conserved. This is analogous to entropy being 
-conserved following a adiabatic, pure gas. %s = s(\rhod/\rhog)
-We explicitly show that for 
-a locally isothermal, dusty gas that the evolutionary equation of 
-$\rhod/\rhog$ may be recast into an effective energy equation identical
-in form to that in standard adiabatic hydrodynamics. We find the 
-natural definition of the entropy of the dusty gas is 
-\begin{align*}
-  s  = \ln{\left(\frac{c_s^2\rhog}{\rhog+\rhod}\right)}.  
-\end{align*}
-This allows us to define the buoyancy frequency of an isothermal 
-dusty gas.
-
-%
-%finite coupling/heat source
-
-Finite dust-gas friction causes a relative drift between the two
-components of the mixture. Then $\rhod/\rhog $ is no
-longer conserved following a fluid parcel, as it is allowed to
-exchange dust particles with neighboring parcels owing to a non-zero
-stopping time $\tstop$. This is analogous to heat exchange between a 
-pure gas parcel and its surroundings. Thus, finite dust-gas friction
-may be interepreted as an energy/entropy source term for the dusty
-gas.
-
-%
-In this first study, we apply this equivalence principle to study the
-axisymmetric stability of locally isothermal, protoplanetary disks
-perfectly coupled to dust ($\tstop=0$). This is so that we may 
-compute exact, steady equilibria for analyses. Furthermore,
-in this limit the dusty problem is \emph{identical} to adiabatic, pure
-gas dynamics.
-
-We obtain the Solberg-Hoiland criteria to assess
-axisymmetric stability when the equation of state is strictly
-isothermal. We show that {\bf dust-settling cannot lead to
-  axisymmetric instability, however large the associated vertical
-  shear} \citep[cf. \emph{non-axisymmetric} Kelvin-Helmholtz instabilities
-  induced by  dust-settling, ][]{lee10}. We also show that
-sharp edges in the dust-to-gas ratio cannot persist in protoplanetary
-disks, as these imply a sharp entropy gradient and may thus destablize
-the disk.
-
-%does not preclude non-axisymmetric instabilities. but if there is a
-%global radial temp gradient, then vertical shear arising from
-%it will always be comparable to dust-settling. but there is vsi with
-%global rad temp grad
-
-For locally isothermal disks $c^2_s=c_s^2(r)$ we generalize the vertical
-shear instability (VSI) to include dust. We find that dust-loading
-generally has a stabilizing effect. In our disk models
-dust-loading does not affect VSI growth rates significantly, but
-meridional motions may be suppressed at heights where vertical 
-buoyancy dominates over vertical shear. This confirms our 
-identification of the entropy and buoyancy of an isothermal, dusty
-gas.
-
 %growth rates not sig changed 
 %v shear from thin dust layers are overwhelmed by 
+\section{Discussion}\label{discussion}
 
@@
 \ref{dusteq} gives 
-\begin{align}
+\begin{align}\label{poly_energy}
   \frac{D P}{D t} = - \Gamma P\nabla\cdot\bm{v}  + P \bm{v}\cdot\nabla \ln{K}
@@
 \begin{align}
-  s = \ln{\left[K(r,z)\left(1 - \tepsilon\right)^\Gamma\right]}.  
+  s = \ln{\left[K^\frac{1}{\Gamma}(r,z)\left(1 - \tepsilon\right)\right]}.  
 \end{align}
-When $K$ is constant, the corresponding Solberg-Hoiland criteria for
-axisymmetric stability may be obtained from that in standard adiabatic 
-hydrodynamics \citep[e.g.][]{tassoul78} by
-using this definition of entropy in
-Eqs. \ref{dusty_solberg1}---\ref{dusty_solberg2}.
+%When $K$ is constant, the corresponding Solberg-Hoiland criteria for
+%axisymmetric stability may be obtained from that in standard adiabatic 
+%hydrodynamics \citep[e.g.][]{tassoul78} by
+%using this definition of entropy in
+%Eqs. \ref{dusty_solberg1}---\ref{dusty_solberg2}.
 
@@
 
-%\subsection{Implication for numerical simulations}
+\subsection{Understanding dust-drag
+  instabilities}\label{dust_as_thermo}
 
-\subsection{Thermodynamic interpretation of dusty instabilities}
+\section{Summary}\label{summary}
+In this paper we highlight the similarity between an isothermal 
+dusty gas and an adiabatic, pure gas. 
+In the limit of perfect dust-gas coupling, the dust content of a fluid 
+parcel $\rhod/\rhog$ is conserved. This is analogous to entropy being 
+conserved following a adiabatic, pure gas. %s = s(\rhod/\rhog)
+We explicitly show that for 
+a locally isothermal, dusty gas that the evolutionary equation of 
+$\rhod/\rhog$ may be recast into an effective energy equation identical
+in form to that in standard adiabatic hydrodynamics. We find the 
+natural definition of the entropy of the dusty gas is 
+\begin{align*}
+  s  = \ln{\left(\frac{c_s^2\rhog}{\rhog+\rhod}\right)}.  
+\end{align*}
+This allows us to define the buoyancy frequency of an isothermal 
+dusty gas.
 
-%\subsection{Caveats and future directions}
-%gi
+%
+%finite coupling/heat source
+
+Finite dust-gas friction causes a relative drift between the two
+components of the mixture. Then $\rhod/\rhog $ is no
+longer conserved following a fluid parcel, as it is allowed to
+exchange dust particles with neighboring parcels owing to a non-zero
+stopping time $\tstop$. This is analogous to heat exchange between a 
+pure gas parcel and its surroundings. Thus, finite dust-gas friction
+may be interepreted as an energy/entropy source term for the dusty
+gas.
+
+%
+In this first study, we apply this equivalence principle to study the
+axisymmetric stability of locally isothermal, protoplanetary disks
+perfectly coupled to dust ($\tstop=0$). This is so that we may 
+compute exact, steady equilibria for analyses. Furthermore,
+in this limit the dusty problem is \emph{identical} to adiabatic, pure
+gas dynamics.
+
+We obtain the Solberg-Hoiland criteria to assess
+axisymmetric stability when the equation of state is strictly
+isothermal. We show that {\bf dust-settling cannot lead to
+  axisymmetric instability, however large the associated vertical
+  shear} \citep[cf. \emph{non-axisymmetric} Kelvin-Helmholtz instabilities
+  induced by  dust-settling, ][]{lee10}. We also show that
+sharp edges in the dust-to-gas ratio cannot persist in protoplanetary
+disks, as these imply a sharp entropy gradient and may thus destablize
+the disk.
+
+%does not preclude non-axisymmetric instabilities. but if there is a
+%global radial temp gradient, then vertical shear arising from
+%it will always be comparable to dust-settling. but there is vsi with
+%global rad temp grad
+
+For locally isothermal disks $c^2_s=c_s^2(r)$ we generalize the vertical
+shear instability (VSI) to include dust. We find that dust-loading
+generally has a stabilizing effect. In our disk models
+dust-loading does not affect VSI growth rates significantly, but
+meridional motions may be suppressed at heights where vertical 
+buoyancy dominates over vertical shear. This confirms our 
+identification of the entropy and buoyancy of an isothermal, dusty
+gas.

### 2016-08-13 appendix.tex

@@
-\section{Variational principle for locally isothermal gas
-  well-coupled to dust}\label{var_prin}
+\section{Variational principle}\label{var_prin}
 
-The linearized equations for axisymmetric disturbances are
+Here we consider the more general equation of state for the gas with
+$P = K \rhog^{\Gamma}$ where $K$ is a prescribed function of position
+and $\Gamma$ is a constant. The effective energy equation is then
+Eq. \ref{poly_energy}. The linearized equations for axisymmetric
+disturbances are
 
@@
   \dd\bm{v}\cdot\nabla\ln{\rho},\label{lin_mass_full}\\
- \ii\sigma\frac{\delta P}{P} &= \nabla\cdot\dd\bm{v} +
+%  \ii\sigma\frac{\delta P}{P} &= \nabla\cdot\dd\bm{v} +
+   \ii\sigma\frac{\delta P}{P} &= \Gamma \nabla\cdot\dd\bm{v} +
   \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
@@
     v_\phi}{\p z}\dd v_z,\\
-  -\ii\sigma\dd v_z &=  \delta\bm{F}\cdot\hat{\bm{z}}.
+  -\ii\sigma\dd v_z &=  \delta\bm{F}\cdot\hat{\bm{z}}, 
 \end{align}  
-These equations do not assume the radially-local approximation used in
-the main text.
-
+where the linearizd pressure force $\delta \bm{F}$ is given in
+Appendix \ref{lin_press}. These equations do not assume the
+radially-local approximation used in the main text. 
 From the linearized meridional momentum equations, we find 
@@
  & = \int\left[
+%   \left(
+%   \ii\sigma \frac{\dd P}{P} - \dd\bm{v}\cdot\nabla s + \dd
+%   \bm{v} \cdot \nabla \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}
+%   \right)\dd\bm{v}^*\cdot\nabla P  
    \left(
-   \ii\sigma \frac{\dd P}{P} - \dd\bm{v}\cdot\nabla s + \dd
-   \bm{v} \cdot \nabla \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}
-   \right)\dd\bm{v}^*\cdot\nabla P  + \ii\sigma \dd
-   P  \nabla\cdot \dd \bm{v}^*
+   \ii\sigma \frac{\dd P}{\Gamma P} - \dd\bm{v}\cdot\nabla s + \frac{\dd
+   \bm{v}}{\Gamma} \cdot \nabla \ln{K} +
+   \frac{\dd\mathcal{C}}{\Gamma P} \right)\dd\bm{v}^*\cdot\nabla P
+   + \ii\sigma \dd P  \nabla\cdot \dd \bm{v}^*
    \right]dV\notag\\
  & = \int\left[
-   \ii\sigma\frac{\dd P}{P}\nabla\cdot\left(P\dd\bm{v}^*\right) +
-   \left(\dd\bm{v}^*\cdot\nabla P\right) \left(\dd\bm{v} \cdot \nabla
-   \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}   -  \dd\bm{v}\cdot\nabla s\right) 
+%   \ii\sigma\frac{\dd P}{P}\nabla\cdot\left(P\dd\bm{v}^*\right) +
+%   \left(\dd\bm{v}^*\cdot\nabla P\right) \left(\dd\bm{v} \cdot \nabla
+%   \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}   -  \dd\bm{v}\cdot\nabla s\right) 
+%   \right]dV\notag\\
+   \ii\sigma\frac{\dd P}{\Gamma P}\left(\bm{\dd v}^*\cdot\nabla P +
+   \Gamma P\nabla\cdot\bm{\delta v}^*   \right)
+   +\left(\dd\bm{v}^*\cdot\nabla P\right) \left( \frac{\dd\bm{v}}{\Gamma} \cdot \nabla
+   \ln{K} + \frac{\dd\mathcal{C}}{\Gamma P}   -  \dd\bm{v}\cdot\nabla s\right) 
    \right]dV\notag\\
  & = \int\left\{
-   \left[\frac{1}{P}\nabla\cdot\left(P\dd\bm{v}\right) -
-   \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P}  \right]\nabla\cdot\left(P\dd\bm{v}^*\right)
-   +\left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v} \cdot \nabla
-   \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}   -  \dd\bm{v}\cdot\nabla s \right)
+%   \left[\frac{1}{P}\nabla\cdot\left(P\dd\bm{v}\right) -
+%   \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P}  \right]\nabla\cdot\left(P\dd\bm{v}^*\right)
+%   +\left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v} \cdot \nabla
+%   \ln{c_s^2} + \frac{\dd\mathcal{C}}{P}   -  \dd\bm{v}\cdot\nabla s \right)
+%   \right\}dV \notag\\
+ \left[
+   \nabla\cdot\dd\bm{v} +
+  \frac{\dd\bm{v}}{\Gamma}\cdot\nabla\ln{P} -
+  \frac{\dd\bm{v}}{\Gamma}\cdot\nabla\ln{K} -
+  \frac{\dd\mathcal{C}}{\Gamma P}. 
+  \right]
+ \left(\bm{\dd v}^*\cdot\nabla P +
+ \Gamma P\nabla\cdot\bm{\delta v}^*   \right)\right. \notag\\
+ &\phantom{=\int\left\{\right\}}\left. 
+ + \left(\dd\bm{v}^*\cdot\nabla P\right)\left(\frac{\dd\bm{v}}{\Gamma} \cdot \nabla
+ \ln{K}
+ + \frac{\dd\mathcal{C}}{\Gamma P}   -  \dd\bm{v}\cdot\nabla s \right)
  \right\}dV \notag\\
  &=\int\left[
-     \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2 -
+%     \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2 -
+%     \left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v}\cdot\nabla s
+%     \right) -
+%     P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2} + \frac{\dd\mathcal{C}}{P} \right)  
+    \frac{1}{\Gamma P} \Bigl\lvert \bm{\dd v}\cdot\nabla P +
+ \Gamma P\nabla\cdot\bm{\delta v}    \Bigr\rvert ^2 - 
      \left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v}\cdot\nabla s
      \right) -
-     P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2} + \frac{\dd\mathcal{C}}{P} \right)  
+      P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{K} + \frac{\dd\mathcal{C}}{P} \right) 
      \right]dV,
 \end{align}
-where we recall $s\equiv\ln{(P/\rho)}$. Hence,
+where  $s\equiv\ln{(P^{1/\Gamma}/\rho)}$. Hence,
 \begin{align}\label{integral_ex}
@@
 &\phantom{==\int}\left.+
-  \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2
-  \right\}dV -\int
-  P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV
+%  \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2
+ \frac{1}{\Gamma P}\Bigl\lvert \bm{\dd v}\cdot\nabla P +
+ \Gamma P\nabla\cdot\bm{\delta v}     \Bigr\rvert^2
+  \right\}dV 
+%  P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV
+-\int    P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{K}\right)dV
   -\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV.  
 \end{align}
-Note that the coefficients of the mixed terms are equal owing to the
-equilibrium state.
+Note that the coefficient of $\delta v_z\delta v_r^*$ and $\delta
+v_z^*\delta v_r$ are equal owing to the
+equilibrium state (Eq. \ref{vshear}). 
+The special case $\Gamma=1,\, K = c_s^2$ gives
+Eq. \ref{int_rel}. A radially-local, vertically-global, viscous version of
+Eq. \ref{integral_ex} is given by \cite{latter06}. See also
+\cite{kato78} .
 
@@
 
-\section{Dispersion relation for unstratified disks}
+\section{Dispersion relation for unstratified disks}\label{compressible_streaming}
 Unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
@@
 \end{align}
+We note that the incompressible condition used by \cite{jacquet11} is
+obtained by setting $\zeta$ to zero.
 
@@
 as derived by \cite{jacquet11} and \cite{laibe14} for the streaming
-instability.
+instability. If $|\sigma| = O(\tau_\mathrm{s})$ then the quartic term
+is small and may be neglected. In that case we obtain the cubic
+dispersion relation of \citet{youdin05a}.

### 2016-08-13 model.tex

@@
 Since $|q|=O(1)$ in PPDs, Eq. \ref{vshear_split} indicates that
-vertical shear due to variations in the dust-to-gas ratio is at 
+vertical shear due to variations in the dust-to-gas ratio can be at
 most comparable to that due to the radial temperature gradient. 
 This only occurs for thin dust layers with $\delta^2\ll \epsilon$ (so 
-the above ratio $\sim 1/|q|$). Otherwise, vertical shear is 
-predominantly associated with  $\p_rc_s^2$.
+the above ratio $\sim 1/|q|$). Otherwise, vertical shear is associated
+with $\p_rc_s^2$.
 
@@
 \end{align*}
-However, for thin dust layers such that $H_\epsilon \ll Hg$, 
+However, for thin dust layers such that $H_\epsilon \ll \Hg$, 
 $\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.

### 2016-08-13 appendix.tex
minor edits

@@
 Appendix \ref{lin_press}. These equations do not assume the
-radially-local approximation used in the main text. 
-From the linearized meridional momentum equations, we find 
+radially-local approximation used in the main text. From the
+linearized meridional momentum equations, we find
+
 \begin{align}
@@
 Note that the coefficient of $\delta v_z\delta v_r^*$ and $\delta
-v_z^*\delta v_r$ are equal owing to the
-equilibrium state (Eq. \ref{vshear}). 
-The special case $\Gamma=1,\, K = c_s^2$ gives
-Eq. \ref{int_rel}. A radially-local, vertically-global, viscous version of
-Eq. \ref{integral_ex} is given by \cite{latter06}. See also
-\cite{kato78} .
+v_z^*\delta v_r$ are equal owing to the equilibrium state
+(Eq. \ref{vshear}). The special case $\Gamma=1,\, K = c_s^2$ gives
+Eq. \ref{int_rel}. Similar integral relations are given by
+\cite{kato78, kley93,latter06}.
 
@@
 \section{Linearized pressure forces and its divergence}\label{lin_press}
-Define
-\begin{align}
-  \bm{F} \equiv - \frac{\nabla P}{\rho}. 
-\end{align}
-Then its linearized form is 
+%Define
+%\begin{align}
+ % \bm{F} \equiv - \frac{\nabla P}{\rho}. 
+%\end{align}
+The linearized form of the pressure force $\bm{F} = -\nabla P/\rho$ is  
 \begin{align}
@@
   &\dd F_z = - W F_z - \left[Q^\prime + Q
-    \left(\ln{\rho}\right)^\prime\right]  = s\dd v_z, 
+    \left(\ln{\rho}\right)^\prime\right]  = - \ii\sigma \dd v_z, 
 \end{align} 
@@
     &= \left[\left(\p_r\ln{\rho}-\ii k\right)F_r - \nabla\cdot\bm{F} +
-      F_z^\prime\right]W + \left(\ii k \p_r\ln{\rho} + k^2\right)Q +
-    s\dd v_z^\prime.
+      F_z^\prime\right]W + \left(\ii k \p_r\ln{\rho} + k^2\right)Q  -
+      \ii\sigma \dd v_z^\prime.
 \end{align}
@@
 
-\section{Dispersion relation for unstratified disks}\label{compressible_streaming}
-Unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
-0$. The background $\p_rP/\rho$ and $\tepsilon$ are constant. 
-Fourier analyze in $r$ and $z$ so that $\p_z\to \ii k_z$ and
-$\p_r\to \ii k_x$ when acting on perturbations. Denote $|\bm{k}|^2
-\equiv k_x^2 + k_z^2$. Neglect variations in
-background when compared with that of perturbations. Linearized
+\section{Compressible streaming instability}\label{compressible_streaming}
+We consider an unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
+0$. The background $\p_rP/\rho$ and $\tepsilon$ are constants. 
+We Fourier analyze in $r$ and $z$ so that $\p_z\to \ii k_z$ and
+$\p_r\to \ii k_x$ when acting on perturbations, and denote $|\bm{k}|^2
+\equiv k_x^2 + k_z^2$. We neglect 
+background gradients when compared with that of perturbations. The linearized
 equations are, after eliminating the azimuthal velocity:
@@
 The artifical factor $\zeta = 1$  is inserted to keep track of the
-left-hand-side of the energy equation. The linearized dust diffusion
+left-hand-side of the energy equation. We note that the incompressible
+condition used by \cite{jacquet11} is obtained by setting $\zeta$ to zero. 
+The linearized dust diffusion
 function (Appendix \ref{lin_dust}), 
@@
 \end{align}
-We note that the incompressible condition used by \cite{jacquet11} is
-obtained by setting $\zeta$ to zero.
 
-Eliminate the velocity perturbations to obtain
+We eliminate the velocity perturbations to obtain
 \begin{align}
@@
   \frac{P}{\rho}\sigma + k_x\tstop c_s^2 F_r\left(1 -
-  2\tepsilon\right)\right]W. \\
+  2\tepsilon\right)\right]W, \\
    \sigma^2 \left( \kappa^2 - \sigma^2 - \ii k_x F_r\right)W &=
-    \left(k_z^2\kappa^2 - \sigma^2|\bm{k}|^2\right)Q. 
+    \left(k_z^2\kappa^2 - \sigma^2|\bm{k}|^2\right)Q, 
 \end{align}
-Futher eliminate $Q$ and $W$ to obtain the dispersion relation
+which yields the dispersion relation
 \begin{align}
@@
   k_x \tstop F_r\left(1 - 2\tepsilon\right)
-  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0,
+  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0.
 \end{align}
-where the equation of state $P = c_s^2\left( 1 - \tepsilon\right)\rho$
-was used. 
-If we now set $\zeta = 0$ or consider $c_s^2\to \infty$ to obtain the
+The equation of state $P = c_s^2\left( 1 - \tepsilon\right)\rho$
+was used. If we now set $\zeta = 0$ or consider $c_s^2\to \infty$ to
+obtain the 
 incompressible limit, and define $\tau_\mathrm{s} \equiv

### 2016-08-13 model.tex
minor edits

@@
   &\frac{D\rho}{Dt} = -\rho\nabla\cdot\bm{v}, \label{masseq}\\ 
-  &\frac{D\bm{v}}{Dt} = - \nabla\Phi - \frac{1}{\rho}\nabla
-  P, \label{momeq}\\ 
+  &\frac{D\bm{v}}{Dt} = - \nabla\Phi - \frac{1}{\rho}\nabla  P, \label{momeq}\\ 
   &\frac{D\tepsilon}{Dt} = -\frac{1}{\rho} \nabla \cdot \left(\tepsilon 
@@
 \end{align}
-where $G$ is the gravitational constant. The second equality is the 
+where $G$ is the gravitational constant, and $(r,\phi, z)$ are
+cylindrical co-ordinates centered on the star. 
+The second equality is the 
 thin-disk approximation for the disk potential, appropriate for
@@
 where $c_s$ is a prescribed sound-speed profile fixed in time. 
-In numerical calculations we will consider vertically 
+In practice we consider vertically 
 isothermal disks with \begin{align}\label{power_temp}
@@
 %as a cooling term.    
+If we denote 
+\begin{align}
+  \bm{F} \equiv  - \frac{\nabla P}{\rho}
+\end{align}
+for the pressure forces, then
+\begin{align*}
+  \mathcal{C} = - c_s^2\nabla \left( \tepsilon \tstop \rho \bm{F}
+  \right), 
+\end{align*}
+which is in the same form as radiative diffusion. In protoplanetary
+disks the heat flux, proportional to $\bm{F}$, is directly radially
+outwards and vertically upwards. This is simply a reflection of
+particle drift (inwards and downwards). Particle flux into a region contributes
+to `cooling' of that region because the effective temperature is
+lowered.
 
@@
 where %$s = \ln{P/\rho}$, 
-$\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)$ is the 
-square of the epicylic frequency. We assume $\kappa^2>0$.  
+$\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)>0$ is the 
+square of the epicylic frequency. %We assume $\kappa^2>0$.  
 Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} are in fact the 
@@
 It is well-known that such correlations can lead to pulsational
-instabilities in stars \citep{cox67}. We thus suggest an interpretation
-of dust-drag instabilities in disks in a similar way. We give a qualitative
+instabilities in stars \citep{cox67}. One may thus interpret 
+dust-drag instabilities in a similar way. We give a
+qualitative 
 description below, based on the more rigorous treatments of stellar 
-pulsational instabilities by 
+pulsations by 
 \cite{cox67} and lecture notes from \cite{samadi15} and
@@
 
-Consider the Lagragian perturbation $\Delta$ of a variable $f$, 
-related to the Eulerian perturbation by $\Delta f = \delta f + 
-\bm{\xi}\cdot\nabla f$, where $\bm{\xi}$ is the Lagragian  
-displacement ($\dd v_{x,z} = -\ii\sigma\xi_{x,z}$). %Note that $\Delta
+The Lagragian perturbation $\Delta$ of a variable $f$ is 
+related to its Eulerian perturbation by $\Delta f = \delta f + 
+\bm{\xi}\cdot\nabla f$; where $\bm{\xi}$ is the Lagragian  
+displacement ($\dd v_{x,z} = -\ii\sigma\xi_{x,z}$). 
+%Note that $\Delta
 %\mathcal{C} = \delta \mathcal{C}$ for a uniform background $C$ as
@@
 Dust-drag causes a phase difference between the 
-pressure and density evolution following a parcel. Now, consider a 
+pressure and density evolution of a fluid element. Now, consider a 
 fluid element at maximum density during an oscillation cycle, 
-so that $D\rho/ D t = 0$. If at this point dust-gas drag provides an
-effective heating ($\mathcal{C}>0$), then the fluid element is
-experiencing increasing pressure, $DP/D t>0$. It will attain
+so that $D\rho/ D t = 0$. If at this point dust-gas drag (or particle flux) provides an
+effective heating to the fluid element ($\mathcal{C}>0$), then it will 
+experience increasing pressure, $DP/D t>0$. It attains 
 pressure maximum \emph{after} reaching density maximum. This implies 
 pressure maximizes while the parcel is expanding: it will do 
-positive work (the enclosed area by its `$P$-$V$' curve  is 
-positive). This will increase the oscillation amplitude.
+positive work  --- the area enclosed by its path in the `$P$-$V$' plane  is 
+positive. This will increase the oscillation amplitude.
 
@@
 instability in dusty protoplanetary disks as dust accumulation
-(and hence compression) at a pressure bump which back-reacts on the
-gas flow in such a way to strengthen the pressure bump. A phase
+(and hence compression) at a pressure bump, which then drags the 
+gas towards it (and hence heating) to strengthen the pressure bump. A phase
 difference was also noted in the numerical calculations of
-\cite[][although they consider the full two-fluid system]{youdin07b}.
+\citet[][although they consider the full two-fluid system]{youdin07b}.
 
-We suggest that a useful way to understand  
-dust-drag instabilities in protoplanetary disks may be to re-cast the  
-problem into that of thermodynamics, since such instabilities have 
-already been well-studied in the stellar context.
+We suggest a useful way to understand dust-drag instabilities in
+protoplanetary disks is to re-cast the problem into that of
+thermodynamics. Then such instabilities appear in the equations in the
+same way as pulsational instabities in stars, which has already been
+well studied.
 

### 2016-08-13 paper.tex
minor edits

@@
 We briefly discuss potential thermodynamical interpretations of 
-dust-gas drag dynamics, for example the streaming instability 
-as a manifestation of pulsational instabilities in stars. 
+dust-gas drag dynamics, such as the streaming instability in
+protoplanetary disks as a manifestation of pulsational instabilities
+in stars.  
 \end{abstract}
@@
 
-By re-casting the dusty problem into identical form to standard
-hydrodynamics, we can follow standard procedures to generalize previous stability 
+By re-casting the dusty problem into identical form to 
+hydrodynamics, we can apply standard methods to generalize previous stability 
 results for pure gas disks to the dusty case. We demonstrate this for

### 2016-08-18 model.tex
minor changes

@@
 dust-to-gas ratio. The gas pressure $P$ is given by the chosen 
-equation of state described below. The external potential $\Phi$ is
-that for a star of mass $M_*$ at the origin 
+equation of state described below. Note that Eq. \ref{dusteq} is
+equivalent to Eq. 46 of \cite{jacquet11}.
+
+The external potential $\Phi$ is 
+that for a star of mass $M_*$, 
 \begin{align}\label{thin_disk_potential}
@@
 \end{align*}
-which is in the same form as radiative diffusion. In protoplanetary
+which is in the same form as cooling by radiative diffusion. In protoplanetary
 disks the heat flux, proportional to $\bm{F}$, is directly radially
@@
 
-\subsubsection{Possible instability at dust edges}
+\subsubsection{Instability at dust edges}
 Notice the left-hand-side of Eq. \ref{stability_est2} is a quadratic in
@@
 fluid element at maximum density during an oscillation cycle, 
-so that $D\rho/ D t = 0$. If at this point dust-gas drag (or particle flux) provides an
-effective heating to the fluid element ($\mathcal{C}>0$), then it will 
-experience increasing pressure, $DP/D t>0$. It attains 
+so that $D\rho/ D t = 0$. If at this point dust-gas drag provides an
+effective heating to the fluid element ($\mathcal{C}>0$, i.e. a particle
+flux \emph{away} from it), then it will 
+experience increasing pressure, $DP/D t>0$. It will attain  
 pressure maximum \emph{after} reaching density maximum. This implies 
 pressure maximizes while the parcel is expanding: it will do 
-positive work  --- the area enclosed by its path in the `$P$-$V$' plane  is 
+positive work  --- the area enclosed by the parcel's path in the `$P$-$V$' plane  is 
 positive. This will increase the oscillation amplitude.

### 2016-08-18 paper.tex
minor changes

@@
   limit that an isothermal dusty gas indeed behaves like an  
-  adiabatic pure gas, and that finite dust-gas coupling may be
-  regarded as an effective 
+  adiabatic pure gas; and that finite dust-gas coupling may be
+  considered as an effective 
   heating/cooling. We exploit this equivalence 
@@
 %dust settling does not lead to axisymmetric instability. 
-We briefly discuss potential thermodynamical interpretations of 
+We briefly discuss thermodynamical interpretations of 
 dust-gas drag dynamics, such as the streaming instability in

### 2016-08-19 paper.tex
minor changes

@@
 $\rho$ and $\bm{v}$ are supplemented by evolutionary equations for
-the dust-to-gas ratio $\rhod/\rhog$ and the dust-gas velocity
-difference $\Delta\bm{v}$. The latter also contributes to
+the dust-to-gas ratio $\rhod/\rhog$ and the relative dust-gas velocity
+$\Delta\bm{v}\equiv \bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$. The latter also contributes to 
 additional source terms in the momentum and energy equations.  

### 2016-08-19 summary.tex
minor changes

@@
 
+As described in \S\ref{dust_work}, a phase lag between the   
+pressure and density response in a dusty gas is the
+physical cause of dust-drag instabilities. %Whether the phase
+%difference is positive or negative depends on the cooling function
+%$\mathcal{C}$, but we may expect a system to generally have both
+%modes, and that with a phase lag are unstable. 
+This may be used to re-interpret the streaming instability or to
+understand other dusty instabilities such as that recently discovered
+by \cite{loren15,loren16}. These authors find an axisymmetric
+instability as dust settles to the disk midplane.
+
+%give numerical example
+
+%expect the thermodynamic interpretation to hold even when tstop not
+%small 
+In fact, we expect a phase difference between pressure and the total
+density  whenever $\tstop\neq0$; not just $\tstop\OmK\ll 1$ as 
+considered in this work. This is simply because of the finite
+relaxation time required for dust to respond to gas, and vice versa. 
+An effective energy equation in the form of Eq. \ref{eff_energy} can
+still be obtained for general $\tstop$. The
+heat flux $\bm{F}$ in the cooling function $\mathcal{C}$ is determined
+through solution to the relative velocity $\Delta \bm{v} =
+\bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$ \citep{youdin05,laibe14}.  We
+thus expect the `pressure-density lag' interpretation of dusty
+instabilities to be generally applicable.
+
 \section{Summary}\label{summary}

### 2016-08-20 model.tex
added loren's instability

@@
 
-We suggest a useful way to understand dust-drag instabilities in
-protoplanetary disks is to re-cast the problem into that of
-thermodynamics. Then such instabilities appear in the equations in the
+We suggest understanding dusty instabilities in
+protoplanetary disks by re-casting the problem into that of
+thermodynamics. Then dust-drag instabilities appear in the equations in the 
 same way as pulsational instabities in stars, which has already been 
-well studied.
+well studied. Whether or not the dusty disk is actually unstable
+depends on details of the dust-drag or `cooling' function
+$\mathcal{C}$. However, we might generally expect the system to
+support a range of modes. On physical grounds, unstable modes 
+\emph{must} be so by causing a phase lag between pressure and total
+density.
+
+ %Whether the phase
+%difference is positive or negative depends on the cooling function
+%$\mathcal{C}$, but we may expect a system to generally have both
+%modes, and that with a phase lag are unstable.
 

### 2016-08-20 paper.tex
added loren's instability

@@
   heating/cooling. We exploit this equivalence 
-  to 
-%\begin{inparaenum}[1)]
-%\item
-%  identify the entropy of isothermal dusty gas; and 
-%\item 
-  generalize previous, well-known gaseous instabilities
+  to generalize previous, well-known gaseous instabilities
   to dusty protoplanetary disks.
-%\end{inparaenum}
-%This includes the axisymmetric stability of strictly isothermal dusty
-%gas, allowing us to constrain the sharpness of 
-%dust edges that can persist in protoplanetary disks. 
-We show that strictly isothermal, non-self-gravitating, axisymmetric
-disks perfectly coupled to dust are generally stable, however thin the
-dust layer is.  %except possibily in the presence of sharp 
-%edges in radial dependence of the dust-to-gas ratio. 
+%We show that strictly isothermal, non-self-gravitating, axisymmetric
+%disks perfectly coupled to dust are generally stable, however thin the
+%dust layer is. 
+We show in the limit of perfect coupling, that  dust-settling cannot lead to axisymmetric instabilities, however thin the dust layer become.     
+%
 We also develop 
@@
 wavelengths, as well as `surface modes' concentrated near the
-disk boundaries. These results can be understood in terms of 
+disk boundaries. 
+These results can be understood in terms of 
 the buoyancy frequency of a isothermal dusty disk. 

### 2016-08-20 summary.tex
added loren's instability

--- 2016-08-13 summary.tex
@@
 
-\subsection{Understanding dust-drag
+\subsection{Other dust-drag
   instabilities}\label{dust_as_thermo}
+
+As described in \S\ref{dust_work}, 
+the physical cause of dust-drag instabilities is a phase lag between
+the pressure and density response in a dusty gas\footnote{
+In fact, we expect a phase difference between pressure and the total
+density  whenever $\tstop\neq0$; not just $\tstop\OmK\ll 1$ as 
+considered in this work. This is simply because of the finite
+relaxation time required for dust to respond to gas, and vice versa. 
+An effective energy equation in the form of Eq. \ref{eff_energy} can
+still be obtained for general $\tstop$. The
+heat flux $\bm{F}$ in the cooling function $\mathcal{C}$ is determined
+through solution to the relative velocity $\Delta \bm{v} =
+\bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$ \citep{youdin05a,laibe14}.  We
+thus expect the `pressure-density lag' interpretation of dust-drag 
+instabilities to be generally applicable.}. 
+This may be used to re-interpret the streaming instability or to
+understand other dusty instabilities. For example,
+\cite{loren15,loren16} have recently discovered a new axisymmetric 
+instability in their two-fluid numerical simulations of dusty
+protoplanetary disks. They find it leads to the formation of
+`toroidal vortices'.
+
+%give numerical example 
+We ran our one-fluid stability code to find unstable modes  
+similar to that described by \cite{loren15}. We use the same
+parameters reported by \citeauthor{loren15}, as follows. 
+%we set $p=q=0$, which eliminates VSI and minimizes streaming
+%instabilities as there is no radial pressure gradient at the disk
+%midplane.   
+The total metalicity $Z=0.01$ and dust scale-height $H_\mathrm{d} =
+0.4H_\mathrm{g}$, giving a maximum dust-to-gas ratio
+$\mathrm{max}(\epsilon) = 0.025$. The midplane stopping time is 
+$t_\mathrm{s0}=0.02\OmK^{-1}$. The disk parameters are $p=-1.5,\, q=-1$, and  
+$\hgas=0.05$. We set $k = \pi/H_\mathrm{d}$.
+
+Fig. \ref{result2d_loren} shows an unstable mode found for this setup,
+with  
+frequency 
+\begin{align*}
+\sigma = (0.3631\ii - 0.1251)\hgas\OmK. 
+\end{align*}
+This growth rate corresponds to $8.8$ orbits, comparable to the $12$
+orbits reported by \citeauthor{loren15}. This is not a VSI mode, as we
+find it persists when $q=0$. We estimate a much smaller growth
+rate for the streaming instability, of 
+$O(10^{-3}\hgas\OmK)$, for the above parameters. We thus also rule out
+SI, as did \citeauthor{loren15}.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/result2d_loren.ps}
+  \caption{Possible eigenmode of the dusty instability discovered by 
+    \cite{loren15} in their numerical simulations. The growth rate and 
+    radial spacing of the `toroidal vortices' are similar to that
+    reported by \citeauthor{loren15}. The instability is driven by
+    dust-gas drag. The color scale shows the
+    perturbed  
+    dust-to-gas ratio, $\delta\epsilon$; and the arrows show
+    $\sqrt{\rho}\left(\dd v_x, \dd v_z\right)$.
+    \label{result2d_loren}
+  }
+\end{figure}
+
+This preliminary calculation only serves to highlight the possibility
+of new axisymmetric dust-drag instabilities in stratified disks. One 
+issue is that the base state for the above calculation is not a true
+equilibrium (although \citeauthor {loren15} find the dust settling
+time is much longer than the instability growth time). We will study
+this instability in more detail with a proper basic
+state in a follow up work.
 

### 2016-08-20 code/press_loop/dvsi.f90
code update

@@ -161,9 +161,13 @@ program dvsi
   allocate(kaxis(nk))
   allocate(rates(nk))
   dlogk = 0d0 
-  if(nk .gt. 1) dlogk = log10(kmax/kmin)/(nk - 1d0)!(kmax-kmin)/(nk-1d0)
+  if(nk .gt. 1) then 
+!   dlogk = log10(kmax/kmin)/(nk - 1d0)
+    dlogk = (kmax-kmin)/(nk-1d0)
+  endif
   do j=1, nk 
-     kaxis(j) = 10d0**(log10(kmin) + dlogk*(j-1d0))!kmin + dlogk*(j-1d0) 
+!     kaxis(j) = 10d0**(log10(kmin) + dlogk*(j-1d0)) 
+      kaxis(j) = kmin + dlogk*(j-1d0)
   enddo
 
   !set up dgratio axis (output every result for every value)
@@ -324,7 +328,7 @@ subroutine eigenvalue_problem
 
      !filter out modes that grow too slow or too fast or too low frequency 
      
-     if((eigen_re.ge.min_rate).and.(eigen_mag.lt.max_rate).and.(abs(eigen_im).gt.smallh_g)) then
+     if((eigen_re.ge.min_rate).and.(eigen_mag.lt.max_rate)) then !.and.(abs(eigen_im).gt.smallh_g)) then
         
         !compute eigenfunc in real space for error tests 
         call get_eigenvec_real(i, eigen, vr, error)

### 2016-08-20 code/press_loopHd/dvsi.f90

@@ -6,11 +6,12 @@ module global
   real*8, parameter :: pi = 2d0*acos(0d0)
   complex*16, parameter :: ii = (0d0, 1d0), zero = (0d0, 0d0)
 
-  integer :: nz, bignz, Nzmid
-  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd
-  real*8 :: zmax, Zmetal 
-  real*8 :: dlnHg_dlnr, beta 
-  real*8 :: krad, bc_tol 
+  integer :: nz, bignz, Nzmid, nk, nHd
+  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd, beta
+  real*8 :: zmax, Hdmin, Hdmax, Zmetal 
+  real*8 :: dlnHg_dlnr, dlogk, dlogdgr 
+  real*8 :: kmin, kmax, krad, dgrmin, dgrmax, bc_tol, growth, freq
+  real*8, allocatable :: kaxis(:), Hdaxis(:), rates(:)
 
   real*8, allocatable :: zaxis(:), lnrho_arr(:), eps_arr(:), dlnrhodr_arr(:), depsdr_arr(:),dlnrhodz_arr(:), depsdz_arr(:) 
   real*8, allocatable :: del2lnrho_arr(:), del2eps_arr(:), dellnrho_dot_deleps_arr(:), tstop_arr(:), d2lnrhodz2_arr(:)
@@ -30,6 +31,8 @@ module global
   complex*16, allocatable :: R41(:,:), R42(:,:), R43(:,:), R44(:,:), R45(:,:) 
   complex*16, allocatable :: R51(:,:), R52(:,:), R53(:,:), R54(:,:), R55(:,:) 
   complex*16, allocatable :: bigW(:), dfrac(:), vx(:), vy(:), vz(:)
+  complex*16, allocatable :: err1(:), err2(:), err3(:), err4(:), err5(:)
+  complex*16, allocatable :: divv(:), dC(:)
 
   complex*16, allocatable :: bigmatrix_lhs(:,:)
   complex*16, allocatable :: bigmatrix_rhs(:,:)
@@ -40,16 +43,16 @@ end module global
 program dvsi
   use global
   implicit none 
-  integer :: i, j, k, lmax 
+  integer :: i, j, k, lmax, loc(1) 
   real*8 :: z, zbar, m, T_l, Tp_l, Tpp_l, dT_l, d2T_l
[…]

### 2016-08-20 summary.tex
edited summary

@@
 
-\subsection{Other dust-drag
+\subsection{New dust-drag
   instabilities}\label{dust_as_thermo}
@@
 %give numerical example 
-We ran our one-fluid stability code to find unstable modes  
+We have applied our one-fluid stability framework to find unstable modes  
 similar to that described by \cite{loren15}. We use the same
@@
 In this paper we highlight the similarity between an isothermal 
-dusty gas and an adiabatic, pure gas. 
-In the limit of perfect dust-gas coupling, the dust content of a fluid 
-parcel $\rhod/\rhog$ is conserved. This is analogous to entropy being 
-conserved following a adiabatic, pure gas. %s = s(\rhod/\rhog)
-We explicitly show that for 
-a locally isothermal, dusty gas that the evolutionary equation of 
-$\rhod/\rhog$ may be recast into an effective energy equation identical
-in form to that in standard adiabatic hydrodynamics. We find the 
+dusty gas and an adiabatic, pure gas. The correspondence arises
+because drag forces reduce the relative velocity between gas and
+dust. In the limit of perfect dust-gas coupling, $tstop \to 0$,  
+ dust is entrained in 
+the gas, and $\rhod/\rhog$ is conserved following the dusty flow. 
+This is analogous to entropy being conserved following a adiabatic,
+pure gas.
+
+For $\tstop\neq 0$, the dust content of a
+fluid parcel of the dusty mixture no longer conserved. The parcel is
+is allowed to exchange dust particles with neighboring parcels owing
+to finite dust-gas friction. This is analogous to heat exchange between a 
+pure gas parcel and its surroundings. %Thus, finite dust-gas friction
+%may be interepreted as an energy source for the dusty
+%gas.
+
+We explicitly show that for a locally isothermal, dusty gas that the
+evolutionary equation of  $\rhod/\rhog$ may be replaced by an 
+effective energy equation. We find the 
 natural definition of the entropy of the dusty gas is 
@@
 \end{align*}
-This allows us to define the buoyancy frequency of an isothermal 
-dusty gas.
-
-%
-%finite coupling/heat source
-
-Finite dust-gas friction causes a relative drift between the two
-components of the mixture. Then $\rhod/\rhog $ is no
-longer conserved following a fluid parcel, as it is allowed to
-exchange dust particles with neighboring parcels owing to a non-zero
-stopping time $\tstop$. This is analogous to heat exchange between a 
-pure gas parcel and its surroundings. Thus, finite dust-gas friction
-may be interepreted as an energy/entropy source term for the dusty
-gas.
+%This allows us to define the buoyancy frequency of an isothermal 
+%dusty gas. 
+In our framework, finite dust-gas friction appears as an energy
+source term.  This analogy with standard adiabatic
+hydrodynamics allow us to obtain dusty analogs of gaseous
+instabilities, and provide thermodynamical interpretations of  
+dust-drag instabilities.
 
@@
 In this first study, we apply this equivalence principle to study the
-axisymmetric stability of locally isothermal, protoplanetary disks
+axisymmetric stability of isothermal, protoplanetary disks
 perfectly coupled to dust ($\tstop=0$). This is so that we may 
@@
 axisymmetric stability when the equation of state is strictly
-isothermal. We show that {\bf dust-settling cannot lead to
+isothermal. We show that {\bf dust settling, by itself, cannot lead to
   axisymmetric instability, however large the associated vertical
   shear} \citep[cf. \emph{non-axisymmetric} Kelvin-Helmholtz instabilities
-  induced by  dust-settling, ][]{lee10}. We also show that
+  in dusty disks][]{lee10}. We show that
 sharp edges in the dust-to-gas ratio cannot persist in protoplanetary
@@
 meridional motions may be suppressed at heights where vertical 
-buoyancy dominates over vertical shear. This confirms our 
-identification of the entropy and buoyancy of an isothermal, dusty
-gas.
+buoyancy dominates over vertical shear. These results are readily
+understood by comparing the vertical shear rate with the buoyancy of
+an isothermal, dusty gas which we have correctly identified.
+
+We also discuss future applications of our thermodynamic framework to 
+study dusty protoplanetary disks. In particular, we emphasize  that the 
+physical origin for any instability associated with strong dust-gas
+drag is a phase lag between the pressure and total density evolution
+of fluid elements. This leads to growing oscillations as positive work
+is done, provided by dust-gas drag, which is equivalent to a `heat'
+source. %We demonstrate a potentially new dust-drag instability in
+%stratified disks that may explain
+
+%We also applied
+%our framework to find unstable modes in dusty, stratified disks
+%similar to that reported recently by \cite{loren15}.

### 2016-08-20 summary.tex
edited summary

@@
 dust-loading does not affect VSI growth rates significantly, but
-meridional motions may be suppressed at heights where vertical 
-buoyancy dominates over vertical shear. These results are readily
-understood by comparing the vertical shear rate with the buoyancy of
-an isothermal, dusty gas which we have correctly identified.
+meridional motions may be suppressed at heights where dust-induced
+vertical buoyancy dominates over vertical shear, consistent with our
+previous study \citep{lin15}.
+
+%These results are readily
+%understood by comparing the vertical shear rate with the buoyancy of
+%an isothermal, dusty gas which we have correctly identified.
 

### 2016-08-20 summary.tex
edited summary

@@
 because drag forces reduce the relative velocity between gas and
-dust. In the limit of perfect dust-gas coupling, $tstop \to 0$,  
+dust. In the limit of perfect dust-gas coupling, $\tstop \to 0$,  
  dust is entrained in 
@@
 axisymmetric stability when the equation of state is strictly
-isothermal. We show that {\bf dust settling, by itself, cannot lead to
+isothermal. We show that {\bf dust settling, in itself, cannot lead to
   axisymmetric instability, however large the associated vertical

### 2016-08-20 paper.tex
edited summary

@@
   particles almost follows the gas flow. 
-  This is analogous to the 
-  conservation of entropy 
-  following an adiabatic fluid.  We show that in the strong drag
-  limit that an isothermal dusty gas indeed behaves like an  
+  This property is similar to 
+  entropy  conservation
+  following an adiabatic fluid.  We show that an isothermal dusty gas indeed behaves like an  
   adiabatic pure gas; and that finite dust-gas coupling may be
@@
 %dust layer is. 
-We show in the limit of perfect coupling, that  dust-settling cannot lead to axisymmetric instabilities, however thin the dust layer become.     
+We show in the limit of perfect coupling, that  dust settling in
+itself cannot lead to axisymmetric instabilities, however thin the dust layer becomes.     
 % 
@@
 These results can be understood in terms of 
-the buoyancy frequency of a isothermal dusty disk. 
+buoyancy forces of a isothermal dusty disk. 
 %We show that vertical shear arising from
 %dust settling does not lead to axisymmetric instability. 
-We briefly discuss thermodynamical interpretations of 
-dust-gas drag dynamics, such as the streaming instability in
-protoplanetary disks as a manifestation of pulsational instabilities
+We discuss thermodynamical interpretations of 
+dust-gas drag instabilities in disks, for example the streaming
+instability, as a phase lag between the pressure 
+and density of a dusty gas, which is   
+analogous to pulsational instabilities 
 in stars.  

### 2016-08-20 summary.tex
edited summary

@@
 %give numerical example 
-We have applied our one-fluid stability framework to find unstable modes  
+We have applied our one-fluid stability analysis to find unstable modes  
 similar to that described by \cite{loren15}. We use the same
-parameters reported by \citeauthor{loren15}, as follows. 
+parameters reported by them as follows. 
 %we set $p=q=0$, which eliminates VSI and minimizes streaming
@@
 %midplane.   
-The total metalicity $Z=0.01$ and dust scale-height $H_\mathrm{d} =
+The total metalicity is $Z=0.01$ and the dust scale-height $H_\mathrm{d} =
 0.4H_\mathrm{g}$, giving a maximum dust-to-gas ratio
@@
 This growth rate corresponds to $8.8$ orbits, comparable to the $12$
-orbits reported by \citeauthor{loren15}. This is not a VSI mode, as we
-find it persists when $q=0$. We estimate a much smaller growth
-rate for the streaming instability, of 
-$O(10^{-3}\hgas\OmK)$, for the above parameters. We thus also rule out
+orbits reported by \citeauthor{loren15}. We find this mode 
+persists when $q=0$, so it is not the VSI.  For the above parameters
+we also estimate a much smaller growth
+rate for the streaming instability (using
+Eq. \ref{streaming_dispersion}) of 
+$O(10^{-3}\hgas\OmK)$, We thus also rule out
 SI, as did \citeauthor{loren15}.
@@
 axisymmetric stability when the equation of state is strictly
-isothermal. We show that {\bf dust settling, in itself, cannot lead to
-  axisymmetric instability, however large the associated vertical
-  shear} \citep[cf. \emph{non-axisymmetric} Kelvin-Helmholtz instabilities
-  in dusty disks][]{lee10}. We show that
+isothermal. We show that {\bf the vertical shear associated with thin dust
+ layers, however large it is, cannot lead to axisymmetric
+  instabilities} \citep[cf. \emph{non-axisymmetric} Kelvin-Helmholtz instabilities
+  in dusty disks, ][]{lee10}. We show that
 sharp edges in the dust-to-gas ratio cannot persist in protoplanetary

### 2016-08-21 paper.tex
minor intro edits

@@
   adiabatic pure gas; and that finite dust-gas coupling may be
-  considered as an effective 
+  regarded as an effective 
   heating/cooling. We exploit this equivalence 
   to generalize previous, well-known gaseous instabilities
-  to dusty protoplanetary disks.
+  to dusty accretion disks.
 %We show that strictly isothermal, non-self-gravitating, axisymmetric
@@
 %dust layer is. 
-We show in the limit of perfect coupling, that  dust settling in
+We show that in the limit of perfect coupling, dust settling in
 itself cannot lead to axisymmetric instabilities, however thin the dust layer becomes.     
 % 
-We also develop 
-the dusty vertical shear instability (VSI) in locally isothermal
+We also extend 
+the vertical shear instability (VSI) to locally isothermal, dusty 
 disks. We find that  dust-loading reduces VSI growth rates, and
-preferentially stabilizes VSI modes with small radial 
+preferentially stabilize modes with small radial 
 wavelengths, as well as `surface modes' concentrated near the
@@
 These results can be understood in terms of 
-buoyancy forces of a isothermal dusty disk. 
+buoyancy forces in an isothermal dusty disk. 
 %We show that vertical shear arising from
@@
 We discuss thermodynamical interpretations of 
-dust-gas drag instabilities in disks, for example the streaming
-instability, as a phase lag between the pressure 
+dust-gas drag instabilities, for example the streaming
+instability in protoplanetary disks, as a phase lag between the pressure 
 and density of a dusty gas, which is   
@@
 ($\sim 1\%$ by mass), their presence can have profound impacts on the 
-gas dynamics and vice versa. Dust-gas drag introduces a 
-wide range of new phenomena that are absent in pure gas
-disks. Important examples include the streaming instability
+gas dynamics and vice versa. Dust-gas friction introduces phenomena
+that are absent in pure gas disks. Important examples include the
+streaming instability 
 \citep{youdin05a,youdin07b,johansen07}, secular gravitational
@@
 \citep{chiang08,barranco09,lee10}. These lead to structure and
-planetesimal formation, as well as turbulence. More recently, new
-dusty instabilities have also been uncovered
-\citep{loren15,loren16,lamb16}
-
-%Understanding dusty gas dynamics is
-%therefore essential to any planet formation theory. 
-%phenomena that depends on gas drag 
-%planet formation in dusty disks 
-%dust component is building block of planets
-%understanding dusty gas dynamics 
-%small amount of dust can have significant consequences
+planetesimal formation, as well as turbulence. 
+More recently, new 
+dusty instabilities have also been uncovered through numerical
+simulations \citep{loren15,loren16,lamb16}, but these are not yet well
+understood.
 
@@
 Current state-of-the-art models of dusty protoplanetary
-disks directly simulate gas dynamics with explicit
-Lagragian dust particles \citep{nelson10,bai10,johansen11,yang14,zhu14,gibbons15,simon16,baruteau16}. In this
-approach, the equations of motion for each particle is solved
-including dust-gas drag, the strength of which is measured 
+disks directly simulate gas dynamics coupled to explicit
+Lagragian dust particles
+\citep{nelson10,bai10,johansen11,yang14,zhu14,gibbons15,simon16,baruteau16}. 
+The equation of motion for each particle is solved 
+including dust-gas drag; the strength of which is measured 
 by a stopping time $\tstop$ --- the decay timescale for the 
@@
 fluid 
-\citep{paardekooper06b,meheut12,laibe12,loren14,fu14b,surville16}. The equations
-hydrodynamics are solved for two fluids: the gas and dust with density and
+\citep{paardekooper06b,meheut12,laibe12,loren14,fu14b,surville16}. The 
+hydrodynamic equations are evolved for two fluids: the gas and dust with density and
 velocity $\rhog$, $\bm{v}_\mathrm{g}$ and $\rhod$, 
@@
 the momentum/energy equations to model dust-gas drag. This approach
-can take advantage of, and adapt, already-developed methods/codes for
+can take advantage of, and adapt, already-developed numerical methods/codes for
 simulating pure gas dynamics.
@@
 It is possible to re-formulate the two-fluid equations into that for a
-single fluid of for the total density $\rho$ and center-of-mass
+single fluid with total density $\rho$ and center-of-mass
 velocity $\bm{v}$ of the dusty gas mixture
@@
 additional source terms in the momentum and energy equations.
+
 The one-fluid approach is appropriate for 
 simulating gas tightly coupled to dust particles with small $\tstop$.  
-This can be numerically challenging for the two-fluid or Lagragian
-particles approach; but is straight forward to model in the
-one-fluid framework \citep{price15}.
+This can be numerically challenging for the two-fluid and Lagragian
+particles framework; but is straight forward to simulate in the
+one-fluid formulation \citep{price15}.  The single fluid formalism
+also has the important advantage that it is 
+amenable to standard analytical methods
+\citep{youdin05a,jacquet11}. This is the necessary  approach to gain
+physical insight and explain some of the above newer dusty
+instabilities.
 
-The single fluid formalism also has the important advantage that it is
-amenable to standard analytical methods
-\citep{youdin05a,jacquet11}. This is the necessary  approach to gain physical insight.  
 In this work, we are motivated to develop a general analytical
-framework to  study the stability of dusty protoplanetary disks by
-drawing on the physical and mathematical parallel between dusty gas and standard
+framework to  study the stability of dusty protoplanetary
+disks by drawing on the physical and mathematical parallel between dusty gas and standard
 hydrodynamics. This is possible in certain limits of astrophysical
@@
 P/\rhog$, where $P$ is the gas pressure \citep{youdin05a,
-  jacquet11,laibe14}. We also consider irradiated protoplanetary disks 
-in which the temperature is set externally \citep{chiang97,stam08}. We
-may then adopt a locally isothermal equation of state $P = c_s^2\rhog$
-where $c_s$ is a prescribed sound-speed profile.  This physically
-represents short cooling times \citep{lin15}.
+  jacquet11,laibe14}. %This eliminates the evolutionary equation for
+%$\Delta \bm{v}$. 
+ We also consider irradiated protoplanetary disks 
+in which the temperature is set externally
+\citep{chiang97,stam08}. This allows us to adopt a locally isothermal
+equation of state, $P = c_s^2\rhog$, where $c_s$ is a prescribed
+sound-speed profile. This physically represents short cooling times
+\citep{lin15}.
 
@@
 dynamics. This is because tightly-coupled dust particles are almost
-entrained in the gas; and this property is analogous to the advection
+entrained in the gas, which is is analogous to the advection
 of entropy following an adiabatic fluid. We will formally relate the 
 entropy of isothermal dusty gas to its dust content. Finite  
-dust-gas drag can be re-interpreted as heating/cooling: the relative
+dust-gas drag can be interpreted as heating/cooling: the relative
 motion between dust and gas causes a fluid parcel to exchange dust
@@
 
-By re-casting the dusty problem into identical form to 
-hydrodynamics, we can apply standard methods to generalize previous stability 
-results for pure gas disks to the dusty case. We demonstrate this for
+By re-casting the dusty problem into identical form as standard  
+hydrodynamics, we can apply well-known methods to generalize previous stability 
+analyses for pure gas disks to the dusty case. We explicitly demonstrate this for
 perfectly coupled dust in strictly and locally isothermal disks. We 
-show that sharp dust edges may destabilize protoplanetary disks.  
-We also study the effect of dust-loading on the vertical shear
+show that strictly isothermal protoplanetary disks are generally
+stable to axisymmetric perturbations, regardless of how  thin the dust
+layer is. We also study the effect of dust-loading on the vertical shear
 instability \citep[VSI,][]{nelson13,lin15,barker15}. We provide 
@@
 equations describing tightly-coupled dusty gas. We show they are in 
-fact the usual hydrodynamic equations. This leads to a precise 
-definition of the entropy of an isothermal dusty gas. We describe approximate 
-equilibrium solutions in \S\ref{eqm}. In \S\ref{limits} we discuss
-general stability properties of the dusty problem in various
-limits. We use these results to constraint the sharpness of dust edges
-in isothermal disks. In \S\ref{linear_problem} we formulate an
-explicit stability calcultion to study the dusty VSI, for which we
-present numerical results in \S\ref{results}. We 
+fact the usual hydrodynamic equations in disguise. 
+We describe approximate  equilibrium solutions in \S\ref{eqm}. In \S\ref{limits} we discuss
+general stability properties of the dusty disk in various
+limits. There, we give a thermodynamical interpretation of
+instabilities caused by directly by dust-gas drag. In
+\S\ref{linear_problem} we list the linearized equations, which we
+solve numerically in \S\ref{results} to study the dusty VSI. We 
 discuss future applications of the dusty/adiabatic gas equivalence in 

### 2016-08-23 summary.tex
final fig in place

@@
 
-Fig. \ref{result2d_loren} shows an unstable mode found for this setup,
-with  
+Fig. \ref{compare_eigenvals_ts0d02} shows the unstable modes found for
+this setup. We find the usual VSI modes, here with $\omega> 3,\, s<
+0.2$. However, we also find the mode marked by the blue asterisk with
 frequency 
@@
 This growth rate corresponds to $8.8$ orbits, comparable to the $12$
-orbits reported by \citeauthor{loren15}. We find this mode 
-persists when $q=0$, so it is not the VSI.  For the above parameters
+orbits reported by \citeauthor{loren15}. We visualize this mode in
+Fig. \ref{result2d_loren}. We find this mode 
+persists when $q=0$, so it is not the VSI. 
+Using Eq. \ref{streaming_dispersion} with the above parameters,  
 we also estimate a much smaller growth 
-rate for the streaming instability (using
-Eq. \ref{streaming_dispersion}) of 
+rate for the streaming instability of 
 $O(10^{-3}\hgas\OmK)$, We thus also rule out
-SI, as did \citeauthor{loren15}.
+SI, as did \citeauthor{loren15}. However, the mode disappears with
+$\tstop=0$, so it is associated with finite dust-gas drag.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_eigenvals_ts0d02}
+  \caption{Unstable modes in a locally isothermal, dusty disk
+    with stopping time  
+    $\tstop=0.02\OmK^{-1}$. Modes with $\omega>3$ are due to
+    VSI. The mode marked by the blue asterisk, visualized in
+    Fig. \ref{result2d_loren}, is associated with finite dust-gas
+    drag.   
+    \label{compare_eigenvals_ts0d02}
+  }
+\end{figure}
 
@@
   \includegraphics[width=\linewidth]{figures/result2d_loren.ps}
-  \caption{Possible eigenmode of the dusty instability discovered by 
-    \cite{loren15} in their numerical simulations. The growth rate and 
+  \caption{The mode marked by the blue asterisk in
+    Fig. \ref{compare_eigenvals_ts0d02}. This is a possible eigenmode
+    of the dusty instability discovered by  
+    \cite{loren15} in their numerical simulations. The instability growth timescale and 
     radial spacing of the `toroidal vortices' are similar to that 
-    reported by \citeauthor{loren15}. The instability is driven by
+    reported by \citeauthor{loren15}. This instability is driven by
     dust-gas drag. The color scale shows the 

### 2016-08-24 model.tex
edited summary
minor changes

@@
 
-If the $\p_r\rhog$ can be neglected in Eq. \ref{spm}, then instability
-requires 
+We can neglect density gradients in Eq. \ref{spm} 
+if $r\p_r\ln{\rhog}\sim O(1)$ and $H_\epsilon/\Hgas\gg
+\sqrt{\epsilon}\hgas$. For example, if $\epsilon\simeq 0.01$ and
+$\hgas\simeq 0.05$, then we require $H_\epsilon/\Hg\gg
+5\times10^{-3}$. (So the dust layer thickness $\Hd\gg 5
+\times10^{-3}\Hgas$ as well.) Then 
+instability requires 
 \begin{align}
@@
 
-\subsection{Thermodynamic view of dusty instabilities}\label{dust_work}
+\subsection{Thermodynamics of dusty instabilities}\label{dust_work}
 Now consider constant $c_s$ but $\tstop\neq0$. 

### 2016-08-24 summary.tex
minor changes

@@
 this setup. We find the usual VSI modes, here with $\omega> 3,\, s<
-0.2$. However, we also find the mode marked by the blue asterisk with
+0.2$. However, we also find modes with $\omega<s$. One of these,
+marked by the blue asterisk, has 
 frequency 
@@
     VSI. The mode marked by the blue asterisk, visualized in
-    Fig. \ref{result2d_loren}, is associated with finite dust-gas
+    Fig. \protect\ref{result2d_loren}
+    , is associated with finite dust-gas
     drag. 
-    \label{compare_eigenvals_ts0d02}
-  }
+  }\label{compare_eigenvals_ts0d02}
 \end{figure}
@@
   \includegraphics[width=\linewidth]{figures/result2d_loren.ps}
-  \caption{The mode marked by the blue asterisk in
-    Fig. \ref{compare_eigenvals_ts0d02}. This is a possible eigenmode
-    of the dusty instability discovered by  
+  \caption{Visualization of the mode marked by the blue asterisk in
+    Fig. \protect\ref{compare_eigenvals_ts0d02}. 
+    This is a possible eigenmode of the dusty instability discovered by  
     \cite{loren15} in their numerical simulations. The instability growth timescale and 
@@
     $\sqrt{\rho}\left(\dd v_x, \dd v_z\right)$. 
-    \label{result2d_loren}
-  }
+  }\label{result2d_loren}
 \end{figure}
@@
 equilibrium (although \citeauthor {loren15} find the dust settling
-time is much longer than the instability growth time). We will study
+time is longer than the instability growth time). We will study
 this instability in more detail with a proper basic

### 2016-08-24 linear.tex
minor edits
minor changes
general edits

@@
-\section{Quasi-global linear problem}\label{linear_problem}
-We consider radially-localized axisymmetric disturbances of the form  
+\section{Linearized equations}\label{linear_problem}
+To compute explicit solutions to the linear problem, 
+we consider radially-localized axisymmetric disturbances of the form  
 \begin{align}
@@
 \end{align}  
-where $^\prime \equiv \p_z$ and $\bm{F} \equiv -\nabla P/\rho$. 
+where $^\prime \equiv \p_z$ and recall $\bm{F} \equiv -\nabla P/\rho$. 
 The linearized dust-duffusion function 

### 2016-08-24 model.tex
general edits

@@
 
-\subsection{Strictly isothermal gas perfectly coupled to dust}   
+\subsection{Strictly isothermal gas perfectly coupled to dust}\label{iso_perfect}
 When $c_s^2$ is a constant and $\tstop=0$, the dusty-gas equations are
@@
 \bm{\xi}\cdot\nabla f$; where $\bm{\xi}$ is the Lagragian  
-displacement ($\dd v_{x,z} = -\ii\sigma\xi_{x,z}$). 
+displacement ($    \xi_{x,z} =  \ii \dd v_{x,z}/\sigma$). 
 %Note that $\Delta
@@
 depends on details of the dust-drag or `cooling' function
-$\mathcal{C}$. However, we might generally expect the system to
-support a range of modes. On physical grounds, unstable modes 
+$\mathcal{C}$. However, given the complexity of $\mathcal{C}$ (
+Appendix \ref{lin_dust}), we might generally expect the system to
+support a range of modes, stable and unstable. 
+On physical grounds, growing modes  
 \emph{must} be so by causing a phase lag between pressure and total
@@
 \end{align} 
-This is classic VSI caused by vertical shear arising from a radial
-temperature gradient \citep{nelson13,barker15,lin15}. We compute
-explicit solutions in this limit in \S\ref{results}.
+This is VSI caused by vertical shear arising from a radial
+temperature gradient \citep{nelson13,barker15,lin15}. We present 
+numerical solutions of the VSI with perfectly-coupled dust in \S\ref{results}.

### 2016-08-24 paper.tex
general edits

@@
   particles almost follows the gas flow. 
-  This property is similar to 
+  This property is analogous to 
   entropy  conservation
@@
 instability in protoplanetary disks, as a phase lag between the pressure 
-and density of a dusty gas, which is   
-analogous to pulsational instabilities 
+and density of a dusty gas, much like pulsational instabilities 
 in stars.  

### 2016-08-24 result.tex
minor edits
general edits

@@
 
-We calculate explicit solutions to the linearized
-equations in the limit of perfectly coupled dust ($\tstop=0$), so that
-the equilibria defined in \S\ref{eqm} are exact steady states. 
-We use this to
-examine the effect of dust on classic VSI, defined to be driven by the
-temperature gradient (\S\ref{dusty_vsi_int}). For simplicity we
-consider constant input values of the midplane dust-to-gas ratio
+We present numerical solutions to the linearized
+equations to examine the effect of dust on the VSI. 
+We consider perfectly coupled dust ($\tstop=0$) so that the 
+equilibria defined in \S\ref{eqm} are exact steady states.
+
+For simplicity we input constant values of the midplane dust-to-gas ratio
 $\epsilon_0$ and the characteristic dust thickness $\Hd$. These,
@@
 temperature profile to $q=-1$; and set the gas disk aspect-ratio
-$h_\mathrm{g}=0.05$. These are fiducial values used in \citetalias{lin15}.
+$h_\mathrm{g}=0.05$. These are fiducial values used by 
+\citet[][\citetalias{lin15} in this section]{lin15}.
 
 Two boundary conditions are needed for the $\tstop=0$
-problem. This is expected since this problem is exactly equivalent to
-adiabatic hydrodynamics \citep[e.g.][]{lubow93}. For simplicity we
+problem. This is expected since this problem is equivalent to
+adiabatic hydrodynamics \citep[e.g.][]{lubow93}. We 
 impose solid boundaries so that $\delta v_z(\pm\zmax)=0$.  
@@
 
-\subsection{Qualitative expectations}
-\citetalias{lin15} find the appropriate way to compare 
+\subsection{Qualitative expectations}\label{vsi_est}
+\cite{lin15} found the appropriate way to compare 
 destabilizing vertical shear to stabilizing vertical buoyancy is
@@
 
-We compare in Fig. \ref{compare_vshear_fixHd} the basic state
+Fig. \ref{compare_vshear_fixHd} compares the basic state
 vertical shear rate, which is destabilizing, and the vertical buoyancy
@@
 eigenvalue distributions for $\epsilon_0 \leq 10^{-2}$ are similar to the
-dust-free fiducial case considered by \citetalias{lin15}, consisting
+dust-free fiducial case considered by \citetalias{lin15}. This is
+expected since $\epsilon_0 < \hgas$ (\S\ref{vsi_est}). Eigenvalues 
+consists 
 of the roughly horizontal `body modes' and the nearly-vertical
@@
 We find that increasing the dust-to-gas ratio reduce VSI growth
-rates. Notably `surface modes', which are typically fastest growing in
+rates. Notably, `surface modes', which are typically fastest growing in
 the dust-free case, are surpressed in dusty disks for $\epsilon_0\geq
-0.1$. The body modes' growth rates remain $\sim O(h_\mathrm{g}\OmK)$
+0.1$ (i.e. $\epsilon_0> \hgas$).  
+The body modes' growth rates remain $ O(h_\mathrm{g}\OmK)$ 
 but their oscillation frequency increases with
-$\epsilon_0$. The total number of modes do not change
+$\epsilon_0$. %i.e. increases with buoyancing freq 
+The total number of modes do not change
 significantly. This is in contrast with the effect of increasing
@@
 stabilizes the disk atmosphere against the VSI, restricting
-meridional motions to $|z|\lesssim 2\Hg$, where most of the disk mass
-reside. This is consistent with the above comparison of the basic state vertical
+meridional motions to $|z|\lesssim 2\Hg$. 
+This is consistent with Fig. \ref{compare_vshear_fixHd} comparing the basic state vertical
 shear and buoyancy. Notice also the perturbed dust-to-gas ratio,
@@
 wavelength perturbations. This is because for high wavenumbers the
-dominant modes are surface modes, which are stabilized by
-dust-loading. The figure suggest that VSI becomes much less efficient
+dominant modes are surface modes, which are effectively stabilized by
+dust-loading since bouyancy forces are larges near the disk
+boundaries. The figure suggest that VSI becomes much less efficient
 for $\epsilon_0\gtrsim 0.1$ and $k\Hg\gtrsim 50$.
@@
 We now vary the dust layer thickness $\Hd$ but fix the metalicity 
-$Z \equiv \epsilon_0 \Hd/\Hg = \text{constant}$. This then gives the
+$Z \equiv \epsilon_0 \Hd/\Hg = 0.03$. This then gives the
 required midplane dust-to-gas ratio. Since we will consider thin dust
@@
 
-We set $Z=0.03$ and analyse two disks with $\Hd=0.1\Hg$ and 
+We analyze two disks with $\Hd=0.1\Hg$ and 
 $\Hd=0.99\Hg$. Fig. \ref{compare_vshear_fixZ} compares the vertical
@@
 disks have the same profile with vertical shear dominating over
-buoyancy. We thus expect instabilities away from the disk midplane in 
+buoyancy. We thus expect perturbations away from the disk midplane in 
 both cases. For $|z|\lesssim 0.4\Hg$, a thin dust 
 layer with $\Hd=0.1\Hg$ boosts the vertical shear rate, but the
-associated buoyancy is larger still. Thus we expect the mid-plane of
-the disk with $\Hd=0.1\Hg$ to be stable.
+associated buoyancy is larger still, implying the mid-plane should be
+stable. 
+%Thus we expect the mid-plane of 
+%the disk with $\Hd=0.1\Hg$ to have limited vertical motions.
 
@@
 
-Fig. \ref{result2d_fixZ} compares the fastest-growing VSI body modes  
+Fig. \ref{result2d_fixZ} compares the fastest growing VSI body modes  
 with $k\Hg=30$ for the two cases above.\citepalias[The thinner domain
@@
 Fig. \ref{compare_eigenvals_fixZ} shows the maximum VSI growth rates
-as a function of the dust layer thickness. As before, we find growth
+as a function of $\Hd$. As before, we find growth
 rates are most affected by the vertical structure of the dust layer
-when the perturbation wavenumer is large. VSI growth rates converge as
+when the perturbation wavenumer is large. Notice VSI growth rates converge as
 $\Hd\to 0$. %to dust free values?    
@@
 \subsection{Validity}
-We have considered $\tstop=0$ for simplicity. When $\tstop\neq0$, a 
+We considered $\tstop=0$ in order to perform stability analysis on an
+exact steady state. When $\tstop\neq0$, a 
 disk initialized as in \S\ref{eqm} will evolve 
@@
 absence of turbulent/diffusive processes, dust particles settle to the
-midplane on a timescale $t_\mathrm{settle}\OmK \sim 1/\tstop\OmK$
-\citep{takeuchi02}. For the previous results to be valid (i.e. the
-basic state does not evolve significantly) we require VSI growth
-timescales $t_\mathrm{grow}\ll t_\mathrm{settle}$. Since the maximum
-VSI growth rate is $\sim h_\mathrm{g}\OmK$, we need stopping times
-such that 
+midplane on a timescale $t_\mathrm{settle}\sim 1/\tstop\OmK^2$
+\citep{takeuchi02}. However, we expect the above results to remain
+valid provided the VSI growth
+timescale $t_\mathrm{grow}\ll t_\mathrm{settle}$. Since 
+VSI growth rates are $O(h_\mathrm{g}\OmK)$, we require 
 \begin{align}
@@
 \end{align} 
-For thin PPD with $h_\mathrm{g}\sim 0.05$, our dusty VSI results are
-valid for $\tstop\OmK < O(10^{-2})$.
+For thin PPDs with $h_\mathrm{g}\sim 0.05$, this is satisfied 
+for $\tstop\OmK \ll O(10^{-2})$.

### 2016-08-24 summary.tex
general edits

@@
 \begin{align}\label{poly_energy}
-  \frac{D P}{D t} = - \Gamma P\nabla\cdot\bm{v}  + P \bm{v}\cdot\nabla \ln{K}
-  + \frac{\Gamma P}{\rhog}\nabla\cdot\left(\tepsilon\tstop\nabla
+  \frac{D P}{D t} &= - \Gamma P\nabla\cdot\bm{v}  + P \bm{v}\cdot\nabla
+  \ln{K} + \mathcal{C},\\
+  \mathcal{C}& = \frac{\Gamma P}{\rhog}\nabla\cdot\left(\tepsilon\tstop\nabla
   P\right).
@@
 \end{align}
+The results of \S\ref{iso_perfect}---\ref{dust_work} remain valid for the strictly 
+polytropic disk with $K=$constant, while one can set $c_s^2\to K$ in
+\S\ref{dusty_vsi_int}. 
 %When $K$ is constant, the corresponding Solberg-Hoiland criteria for
@@
 modification is needed  
-for strictly isothermal or polytropic gas ($c_s$ or $K$ being
-constant, respectively) perfectly-coupled to dust. The usual 
-adiabatic energy equation is effectively a dust continuity
+for strictly isothermal or polytropic gas perfectly-coupled to dust:  
+the standard adiabatic energy equation substitutes for the dust continuity
 equation. When $c_s$ or $K$ is not constant and/or $\tstop\neq0$ one
-can simply add corresponding source terms in energy equation. In the
-large but finite coupling case, this source term is analogous to
-radiative diffusion \citep{price15}, which is also standard in many
-hydrodynamics codes.
+should add corresponding source terms in energy equation. 
+This source term associated with dust-gas drag, $\mathcal{C}$, is 
+analogous to radiative diffusion \citep{price15}, which is also
+standard in many hydrodynamics codes.
 

### 2016-08-24 summary.tex
general edits

@@
 The results of \S\ref{iso_perfect}---\ref{dust_work} remain valid for the strictly 
-polytropic disk with $K=$constant, while one can set $c_s^2\to K$ in
+polytropic disk with $K=$constant, while one sets $c_s^2\to K$ in
 \S\ref{dusty_vsi_int}. 
@@
 equation. When $c_s$ or $K$ is not constant and/or $\tstop\neq0$ one
-should add corresponding source terms in energy equation. 
+should add corresponding source terms in the energy equation. 
 This source term associated with dust-gas drag, $\mathcal{C}$, is 

### 2016-08-26 result.tex
tiny change

@@
 For simplicity we input constant values of the midplane dust-to-gas ratio
-$\epsilon_0$ and the characteristic dust thickness $\Hd$. These,
-together with the perturbation wavenumber $k$, comprises main the
+$\epsilon_0$ and its characteristic height $H_\epsilon$, the latter being 
+specified via the dust thickness $\Hd$. 
+These, together with the perturbation wavenumber $k$, comprises main the
 parameters of the linear problem. We fix the radial power-law index

### 2016-08-26 model.tex
section on radially varying epsilon

@@
 ratio with a power-law temperature profile (Eq. \ref{power_temp}) gives 
-\begin{align*}
+\begin{align}\label{vshear2}
   &r\frac{\p \Omega^2}{\p z}  =
@@
   \right\} 
-\end{align*}
+\end{align}
 The first two terms correspond to vertical shear caused by spatial

### 2016-08-26 result.tex
section on radially varying epsilon

--- 2016-08-24 result.tex
@@
 
-For simplicity we input constant values of the midplane dust-to-gas ratio
-$\epsilon_0$ and the characteristic dust thickness $\Hd$. These,
-together with the perturbation wavenumber $k$, comprises main the
+For simplicity, we mostly consider constant midplane
+dust-to-gas ratios $\epsilon_0$ and its characteristic height
+$H_\epsilon$. The latter is specified via the dust layer thickness
+$\Hd(r_0)$. (We will allow for radial dependices in $\epsilon_0$ and
+$H_\epsilon$ in \S\ref{varHd}.)   
+These, together with the perturbation wavenumber $k$, comprises main the
 parameters of the linear problem. We fix the radial power-law index
@@
 
+\subsection{Effect of a radially-varying dust-to-gas ratio}\label{varHd}
+{\bf move to appendix?}
+
+Here we allow the dust-to-gas ratio $\epsilon$ to depend on
+$r$. Specifically we let $\epsilon_0\propto r^{-1}$ and
+$H_\epsilon \propto \Hg$ (cf. constant values in the previous
+calculations). Then
+\begin{align*}
+  \frac{\p\epsilon}{\p r} =
+  \left(\frac{z^2}{H_\epsilon^2}\frac{d\ln{\Hg}}{dr} -
+    \frac{1}{r}\right)\epsilon.   
+\end{align*}
+We fix $Z=0.01$, $\Hd=0.8\Hg$ at $r=r_0$, and
+$k\Hg=30$.
+
+Fig. \ref{compare_vshear_varHd} shows that a radially-varying
+dust-to-gas ratio increases the magnitude of the vertical shear rate
+away from the midplane (see also Eq. \ref{vshear2}). Thus we find
+higher VSI growth rates, as shown in Fig. \ref{vsi_dust_varHd}. 
+Surfaces mode are more effectively enhanced by the additional vertical 
+shear induced by $\p_r\epsilon$. However, all growth rates remain
+$O(\hgas\OmK)$. Importantly, the increase in the growth rate of the
+fundamental mode is small. Thus we do not expect the radial
+dependence in $\epsilon$ to significantly affect the VSI in a dusty
+disk.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_vshear_Nz2_varHd} 
+  \caption{Vertical shear rate (solid) compared to vertical buoyancy
+    (dotted) in a locally isothermal, dusty disk 
+    with metalicity $Z=0.01$ and dust thickness $\Hd=0.8\Hg$.
+    Black curves assume a dust-to-gas ratio that only depends on
+    height; whereas the blue curve also allows a radial dependence in
+    $\epsilon$. See text for details.  
+    \label{compare_vshear_varHd}
+    }
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_eigenvals_varHd} 
+  \caption{Unstable VSI modes in the disk models of
+    Fig. \protect\ref{compare_vshear_varHd}.  
+    \label{vsi_dust_varHd}
+    }
+\end{figure}
+
 \subsection{Validity}

### 2016-08-26 summary.tex
general edits
section on radially varying epsilon

@@
 should add corresponding source terms in the energy equation. 
-This source term associated with dust-gas drag, $\mathcal{C}$, is 
+The source term associated with dust-gas drag, $\mathcal{C}$, is 
 analogous to radiative diffusion \citep{price15}, which is also
@@
 $\mathrm{max}(\epsilon) = 0.025$. The midplane stopping time is 
-$t_\mathrm{s0}=0.02\OmK^{-1}$. The disk parameters are $p=-1.5,\, q=-1$, and  
-$\hgas=0.05$. We set $k = \pi/H_\mathrm{d}$.
+$t_\mathrm{s0}=0.02\OmK^{-1}$. The gas disk parameters are $(p,q,
+\hgas) =(-1.5, -1,0.05)$.  We set $k = \pi/H_\mathrm{d}$.
 
 Fig. \ref{compare_eigenvals_ts0d02} shows the unstable modes found for
-this setup. We find the usual VSI modes, here with $\omega> 3,\, s<
-0.2$. However, we also find modes with $\omega<s$. One of these,
+this setup. We find the usual VSI modes, here with $\omega> 3\hgas\OmK,\, s<
+0.2\hgas\OmK$. However, we also find modes with $\omega<s$. One of these,
 marked by the blue asterisk, has 

### 2016-08-26 code/press/dvsi.f90
code update

@@ -1,13 +1,13 @@
 module global
   implicit none 
-  logical :: fixZ 
+  logical :: fixZ
   character*4 :: vstruct 
   integer, parameter :: nvar = 5 
   real*8, parameter :: pi = 2d0*acos(0d0)
   complex*16, parameter :: ii = (0d0, 1d0), zero = (0d0, 0d0)
 
   integer :: nz, bignz, Nzmid
-  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd
+  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd, varHd
   real*8 :: zmax, Zmetal 
   real*8 :: dlnHg_dlnr, beta 
   real*8 :: krad, bc_tol 
@@ -47,7 +47,7 @@ program dvsi
   real*8, external  :: rotation, kappa_sq, vertical_shear
   real*8, external  :: Fr, Fz, F_dot_deleps, div_F, F_dot_dellncs2, gr, gz
   real*8, external  :: stopping_time 
-  namelist /params/ smallh_g, rhog0_power, smallq, vstruct, Hd, Zmetal, fixZ, smalld, tstop 
+  namelist /params/ smallh_g, rhog0_power, smallq, vstruct, Hd, varHd, Zmetal, fixZ, smalld, tstop 
   namelist /grid/ zmax, nz 
   namelist /mode/ krad
 

### 2016-08-26 code/press_loop/basic.f90

@@ -80,7 +80,7 @@ real*8 function deps_tilde_dr(zhat)
   real*8, intent(in) :: zhat 
   real*8, external :: eps_tilde
 
-if(vstruct.eq.'mlin')  deps_tilde_dr = -smalld*dgratio*smallh_g*exp(-zhat**2d0/2d0/delta**2d0)
+if(vstruct.eq.'mlin')  deps_tilde_dr =(-smalld*dgratio + varHd*(zhat**2d0/delta**2d0)*dlnHg_dlnr)*smallh_g*exp(-zhat**2d0/2d0/delta**2d0)
 if(vstruct.eq.'tl02')  deps_tilde_dr = eps_tilde(zhat)*( -smalld*smallh_g + beta**2d0*zhat**2d0*smallh_g*dlnHg_dlnr*exp(zhat**2d0/2d0) )
 end function deps_tilde_dr
 
@@ -90,10 +90,15 @@ real*8 function d2eps_tilde_dr2(zhat)
    implicit none
   real*8, intent(in) :: zhat 
   real*8, external :: eps_tilde, deps_tilde_dr
-  real*8 :: temp 
-
-if(vstruct.eq.'mlin')  d2eps_tilde_dr2 = (smalld*smallh_g)**2d0*dgratio*exp(-zhat**2d0/2d0/delta**2d0) 
+  real*8 :: temp, delta2 
   
+if(vstruct.eq.'mlin')  then 
+!   d2eps_tilde_dr2 = (smalld*smallh_g)**2d0*dgratio*exp(-zhat**2d0/2d0/delta**2d0) 
+   d2eps_tilde_dr2  = smalld**2d0*dgratio - varHd*(zhat**2d0/delta**2d0)*dlnHg_dlnr**2d0
+   d2eps_tilde_dr2  = d2eps_tilde_dr2 + varHd*( -smalld*dgratio + (zhat**2d0/delta**2d0)*dlnHg_dlnr )*zhat**2d0*dlnHg_dlnr/delta**2d0 
+   d2eps_tilde_dr2  = d2eps_tilde_dr2*smallh_g**2d0*exp(-zhat**2d0/2d0/delta**2d0) 
+endif
+
 if(vstruct.eq.'tl02') then 
   temp = beta**2d0*zhat**2d0*smallh_g**2d0*(-2d0 - zhat**2d0)*dlnHg_dlnr**2d0*exp(zhat**2d0/2d0)
[…]

### 2016-08-26 code/press_loop/dvsi.f90

@@ -6,7 +6,7 @@ module global
   complex*16, parameter :: ii = (0d0, 1d0), zero = (0d0, 0d0)
 
   integer :: nz, bignz, Nzmid, nk, ndgr
-  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd, beta
+  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd, varHd, beta
   real*8 :: zmax 
   real*8 :: dlnHg_dlnr, dlogk, dlogdgr 
   real*8 :: kmin, kmax, krad, dgrmin, dgrmax, bc_tol, growth, freq
@@ -49,7 +49,7 @@ program dvsi
   real*8, external  :: rotation, kappa_sq, vertical_shear
   real*8, external  :: Fr, Fz, F_dot_deleps, div_F, F_dot_dellncs2, gr, gz
   real*8, external  :: stopping_time 
-  namelist /params/ smallh_g, rhog0_power, smallq, vstruct, Hd, dgrmin, dgrmax, ndgr, smalld, tstop 
+  namelist /params/ smallh_g, rhog0_power, smallq, vstruct, Hd, varHd, dgrmin, dgrmax, ndgr, smalld, tstop 
   namelist /grid/ zmax, nz 
   namelist /mode/ kmin, kmax, nk 
 

### 2016-08-26 code/press_loopHd/dvsi.f90

@@ -7,7 +7,7 @@ module global
   complex*16, parameter :: ii = (0d0, 1d0), zero = (0d0, 0d0)
 
   integer :: nz, bignz, Nzmid, nk, nHd
-  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd, beta
+  real*8 :: smallh_g, rhog0_power, smallq, dgratio, delta, smalld, tstop, Hd, varHd, beta
   real*8 :: zmax, Hdmin, Hdmax, Zmetal 
   real*8 :: dlnHg_dlnr, dlogk, dlogdgr 
   real*8 :: kmin, kmax, krad, dgrmin, dgrmax, bc_tol, growth, freq
@@ -50,7 +50,7 @@ program dvsi
   real*8, external  :: rotation, kappa_sq, vertical_shear
   real*8, external  :: Fr, Fz, F_dot_deleps, div_F, F_dot_dellncs2, gr, gz
   real*8, external  :: stopping_time 
-  namelist /params/ smallh_g, rhog0_power, smallq, vstruct, Zmetal, fixZ, Hdmin, Hdmax, nHd, smalld, tstop 
+  namelist /params/ smallh_g, rhog0_power, smallq, vstruct, Zmetal, fixZ, Hdmin, Hdmax, varHd, nHd, smalld, tstop 
   namelist /grid/ zmax, nz 
   namelist /mode/ kmin, kmax, nk 
 

### 2016-08-26 model.tex
minor edits

@@
 (and hence compression) at a pressure bump, which then drags the 
-gas towards it (and hence heating) to strengthen the pressure bump. A phase
-difference was also noted in the numerical calculations of
-\citet[][although they consider the full two-fluid system]{youdin07b}.
+gas towards it (and hence heating) to strengthen the pressure
+bump. That is, heating occurs upon compression, as occurs for stellar 
+pulsational instabilities. A phase difference was also noted in the 
+numerical calculations of the streaming instability by 
+\citet{youdin07b}.
 
@@
 depends on details of the dust-drag or `cooling' function
-$\mathcal{C}$. However, given the complexity of $\mathcal{C}$ (
+$\mathcal{C}$. However, given the complexity of $\mathcal{C}$ (see
 Appendix \ref{lin_dust}), we might generally expect the system to 
-support a range of modes, stable and unstable. 
-On physical grounds, growing modes  
-\emph{must} be so by causing a phase lag between pressure and total
-density.
+support a range of stable and unstable modes. On physical grounds,
+growing modes \emph{must} be so by causing a phase lag between
+pressure and total density.
 

### 2016-08-26 summary.tex
minor edits

@@
 to previous studies of GI in vertically stratified disks
-\citep[e.g.][]{mamat10, kim12,lin14c} where the temperature decreases
-from the midplane. While we expect only the total surface density and
+ where the temperature decreases
+from the midplane \citep[e.g.][]{mamat10, kim12,lin14c}. 
+While we expect only the total surface density and
 characteristic temperatures are relevant to stability  
@@
  dust is entrained in 
-the gas, and $\rhod/\rhog$ is conserved following the dusty flow. 
-This is analogous to entropy being conserved following a adiabatic,
-pure gas.
+the gas, and the dust-to-gas ratio $\rhod/\rhog$ is conserved
+following the dusty flow. This is analogous to entropy being conserved
+following a adiabatic, pure gas.
 
-For $\tstop\neq 0$, the dust content of a
-fluid parcel of the dusty mixture no longer conserved. The parcel is
-is allowed to exchange dust particles with neighboring parcels owing
-to finite dust-gas friction. This is analogous to heat exchange between a 
-pure gas parcel and its surroundings. %Thus, finite dust-gas friction
-%may be interepreted as an energy source for the dusty
-%gas.
+For finite dust-gas friction, $\tstop\neq 0$, the dust content of a
+fluid parcel of the dusty mixture no longer conserved. A fluid parcel 
+can exchange dust particles with neighboring parcels owing 
+to dust-gas relative drift. This is analogous to heat exchange between
+a  pure gas parcel and its surroundings.
 
@@
 
-We obtain the Solberg-Hoiland criteria to assess 
-axisymmetric stability when the equation of state is strictly
-isothermal. We show that {\bf the vertical shear associated with thin dust
+We obtain the Solberg-Hoiland criteria for axisymmetric stability for
+strictly isothermal, perfectly-coupled dusty gas.  
+We show that {\bf the vertical shear associated with thin dust
  layers, however large it is, cannot lead to axisymmetric
-  instabilities} \citep[cf. \emph{non-axisymmetric} Kelvin-Helmholtz instabilities
-  in dusty disks, ][]{lee10}. We show that
-sharp edges in the dust-to-gas ratio cannot persist in protoplanetary
-disks, as these imply a sharp entropy gradient and may thus destablize
-the disk.
+  instabilities}. This is consistent with the fact that only 
+\emph{non-axisymmetric} Kelvin-Helmholtz instabilities have been 
+observed in numerical simulations in the perfectly-coupled limit. 
+\citep{chiang08,barranco09,lee10}. We show that 
+sharp radial edges in the dust-to-gas ratio cannot persist in
+protoplanetary disks, as these imply a sharp entropy gradient and may
+thus destablize the disk.
 
@@
 
-For locally isothermal disks $c^2_s=c_s^2(r)$ we generalize the vertical
-shear instability (VSI) to include dust. We find that dust-loading
-generally has a stabilizing effect. In our disk models
+For locally isothermal disks, $c^2_s=c_s^2(r)$, we generalize the 
+vertical shear instability (VSI) to dusty disks, which is still 
+driven by the radial temperature gradient. We find that 
+dust-loading is generally stabilizing. In our disk models 
 dust-loading does not affect VSI growth rates significantly, but
@@
 We also discuss future applications of our thermodynamic framework to 
-study dusty protoplanetary disks. In particular, we emphasize  that the 
-physical origin for any instability associated with strong dust-gas
+study dusty protoplanetary disks. In particular, we emphasize the 
+physical origin of any instability driven by strong dust-gas
 drag is a phase lag between the pressure and total density evolution
-of fluid elements. This leads to growing oscillations as positive work
-is done, provided by dust-gas drag, which is equivalent to a `heat'
-source. %We demonstrate a potentially new dust-drag instability in
+of the dusty gas mixture. This leads to growing modes because 
+positive work is done during an oscillation cycle, provided by
+dust-gas drag, which is equivalent to a `heat' 
+source. This interpretation is entirely analogous to pulsational
+instabilities in stars.
+
+%We demonstrate a potentially new dust-drag instability in
 %stratified disks that may explain

### 2016-08-27 code/press_loopHd/basic.f90
tiny tex and code updates

@@ -12,9 +12,15 @@ real*8 function dlnrhog_dr(zhat)
   implicit none 
   real*8, intent(in) :: zhat 
  
-if(vstruct.eq.'mlin') then 
-  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
-              + smalld*dgratio*smallh_g*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) )
+  if(vstruct.eq.'mlin') then 
+     if(varHd.eq.0d0) then 
+        dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
+             + smalld*dgratio*smallh_g*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) )
+     else 
+         dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*zhat**2d0 &
+             + smalld*dgratio*smallh_g*delta**2d0*(1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) &
+             + smallh_g*dgratio*zhat**2d0*dlnHg_dlnr*exp(-zhat**2d0/2d0/delta**2d0) 
+     endif
 endif
 
 if(vstruct.eq.'tl02')  dlnrhog_dr = smallh_g*rhog0_power + dlnHg_dlnr*smallh_g*zhat**2d0
@@ -27,9 +33,15 @@ real*8 function d2lnrhog_dr2(zhat)
   real*8, intent(in) :: zhat 
   
 if(vstruct.eq.'mlin') then
-  d2lnrhog_dr2 = -2d0*dlnHg_dlnr**2d0*smallh_g**2d0*( zhat**2d0 + 2d0*dgratio*delta**2d0*( 1d0 - exp(-zhat**2d0/2d0/delta**2d0) ) ) &
[…]

### 2016-08-28 model.tex
tiny tex and code updates
tiny corrections

@@
 
-We usually consider $\epsilon_0 \ll 1$ and settled dust such that
-$\Htilde<\Hgas$. Then $\rhog(r,z)$ is effectively Gaussian, as in the
+We usually consider gas-dominated disks with $\epsilon_0 \ll 1$.  
+Then $\rhog(r,z)$ is effectively Gaussian, as in the 
 dust-free case. The dust density is approximately 
@@
 \end{align*}
-However, for thin dust layers such that $H_\epsilon \ll \Hg$, 
+However, for thick dust layers such that $H_\epsilon \gg \Hg$, 
 $\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.

### 2016-08-28 result.tex
minor edits
tiny corrections

@@
 These, together with the perturbation wavenumber $k$, comprises main the
-parameters of the linear problem. We fix the radial power-law index
-for the midplane gas density to $p = -1.5$, that for the
-temperature profile to $q=-1$; and set the gas disk aspect-ratio
+parameters of the linear problem. We choose a power-law 
+midplane gas density profile, $\rho_\mathrm{g0}\propto r^q$ with $p =
+-1.5$. The power-law index for the 
+temperature profile is $q=-1$; and we set the gas disk aspect-ratio
 $h_\mathrm{g}=0.05$. These are fiducial values used by 
@@
 higher VSI growth rates, as shown in Fig. \ref{vsi_dust_varHd}. 
-Surfaces mode are more effectively enhanced by the additional vertical 
+Surfaces modes are more effectively enhanced by the additional vertical 
 shear induced by $\p_r\epsilon$. However, all growth rates remain

### 2016-08-28 model.tex
tiny corrections

@@
 discussion is based on the corresponding axisymmetric linearized
-equations about an axisymmetric steady state. A variable $f$ is
-subject to perturbations of the form 
+equations. A variable $f$ is
+subject to Eulerian perturbations of the form 
 \begin{align}
@@
 in thin accretion disks where radial gradients are $O(1/r)$, 
-$h_\mathrm{g},\epsilon,\ll 1$ and $
-\Htilde/H_\mathrm{g}\lesssim 1$. Then the
+$h_\mathrm{g},\epsilon,\ll 1$ and $\Htilde/H_\mathrm{g}$ not too
+large.  
+Then the
 magnitude of the second term on the left-hand-side  of either

### 2016-08-28 model.tex
additional discussion in dust edge instability

@@
 instability requires 
-\begin{align}
+\begin{align}\label{dust_edge}
   \left|\frac{\p\ln{\epsilon}}{\p r}\right| \gtrsim
   \frac{1}{\Htilde}\frac{\left(1+\epsilon\right)}{\sqrt{\epsilon}}.
+%\simeq
+ % \frac{1}{\sqrt{\epsilon}H_\epsilon}, 
 \end{align}
-That is, if the dust-to-gas ratio varies on a radial lengthscale less
+%where the second equality applies to $\epsilon\ll 1$. 
+That is, if the dust-to-gas ratio varies on a radial lengthscale much less
 than $O(\Htilde)$, i.e. its vertical lengthscale,  
@@
 then the system is potentially unstable.  
+%For $\epsilon\ll 1$ the
+%dust-to-gas ratio should vary on a lengthscale shorter than
+%$\sqrt{\epsilon}H_\epsilon$. We apply this below.
+
+PPDs typically have $\epsilon$ of $O(10^{-2})$.  Then 
+for thin dust layers, $H_\epsilon \simeq \Hd\ll \Hg$, instability
+requires $|\p_r\ln{\epsilon}|^{-1}\ll O(0.1\Hg)$, i.e. a very short
+radial length-scale in the dust-to-gas ratio. This is unlikely to
+occur. On the other hand, if dust is well-mixed, say $\Hd\simeq
+0.99\Hg$, then instability only requires
+$|\p_r\ln{\epsilon}|^{-1}\lesssim O(10\Hg)$, which is more realistic.
 

### 2016-08-28 model.tex
additional discussion in dust edge instability

--- 2016-08-28 model.tex
@@
 instability requires 
-\begin{align}
+\begin{align}\label{dust_edge}
 \left|\frac{\p\ln{\epsilon}}{\p r}\right| \gtrsim
   \frac{1}{\Htilde}\frac{\left(1+\epsilon\right)}{\sqrt{\epsilon}}.
+%\simeq
+ % \frac{1}{\sqrt{\epsilon}H_\epsilon}, 
 \end{align}
-That is, if the dust-to-gas ratio varies on a radial lengthscale less
-than $O(\Htilde)$, i.e. its vertical lengthscale,  
-%$O(\widetilde{H}\sqrt{\epsilon})$ for $\epsilon\ll 1$, or less than
-%$O(\widetilde{H}/\sqrt{\epsilon})$ for $\epsilon\gg 1$, 
+%where the second equality applies to $\epsilon\ll 1$. 
+That is, if the radial lengthscale of the dust-to-gas ratio,
+$L_\epsilon\ll O(\Htilde)$, i.e. its vertical lengthscale,  
 then the system is potentially unstable.
+
+Taking $\epsilon\sim 0.01$,  we find that 
+for thin dust layers with $H_\epsilon \simeq \Hd\ll \Hg$,  instability
+requires $L_\epsilon\ll O(0.1\Hg)$, i.e. the dust-to-gas ratio must
+vary on an extremely short lengthscale. 
+%an extremely short  
+%radial length-scale in the dust-to-gas ratio. 
+This is unlikely to
+occur. On the other hand, if dust is well-mixed, say $\Hd\simeq
+0.99\Hg$, then instability only requires 
+$L_\epsilon\lesssim O(10\Hg)$, which is more realistic.
 
@@
 
-\subsection{Thermodynamics of dusty instabilities}\label{dust_work}
+\subsection{Thermodynamics of dust-drag instabilities}\label{dust_work}
 Now consider constant $c_s$ but $\tstop\neq0$. 

### 2016-09-04 summary.tex
tiny changes

@@
 strictly isothermal, perfectly-coupled dusty gas.  
-We show that {\bf the vertical shear associated with thin dust
- layers, however large it is, cannot lead to axisymmetric
-  instabilities}. This is consistent with the fact that only 
+We show that the vertical shear associated with thin dust
+ layers \emph{cannot} lead to axisymmetric
+  instabilities, however large it is.  
+This is consistent with the fact that only 
 \emph{non-axisymmetric} Kelvin-Helmholtz instabilities have been 
-observed in numerical simulations in the perfectly-coupled limit. 
+observed in numerical simulations in the perfectly-coupled limit 
 \citep{chiang08,barranco09,lee10}. We show that 

### 2016-10-13 model.tex
revising with youdin's comments

@@
 \section{Single fluid description of dusty gas}\label{setup} 
-In the limit of strong dust-gas drag (stopping time $\tstop\to 0$) 
-\cite{laibe14} give a single fluid description of the mixture:  
-(their Eq. 84 --- 85 and Eq. 87):
+We model an accretion disk as a mixture of gas with dust treated as a 
+pressureless fluid. We denote their density and
+velocity field as $(\rhog,\bm{v}_\mathrm{g})$ and
+$(\rhod,\bm{v}_\mathrm{d})$, respectively. The two fluids 
+interact via a drag force parameterized by a stopping time $\tstop$
+such that 
+\begin{align}  
+  \rhod\left.\frac{\p \bm{v}_\mathrm{d}}{\p t}\right|_\mathrm{drag}= -
+  \rhog\left.\frac{\p \bm{v}_\mathrm{g}}{\p t}\right|_\mathrm{drag}=
+  - \frac{\rhog\rhod}{\left(\rhod+\rhog\right)}\frac{\left(\bm{v}_\mathrm{d} -
+    \bm{v}_\mathrm{g}\right)}{\tstop}. 
+\end{align}
+is the dust-gas friction force per unit volume. Note that $\tstop$ 
+differs slightly from the stopping time $\tau_\mathrm{s} =
+\tstop\rhog/\left(\rhog+\rhod\right)$ defined in previous studies
+\citep[e.g.][]{youdin05a,jacquet11}.
+
+We consider the limit of tightly-coupled dust particles with 
+$\tstop\OmK\ll 1 $, where $\OmK$ is the Keplerian orbital
+frequency. This allows us to simplify the problem by using the
+`terminal velocity approximation' so that the dust-gas velocity
+difference is 
+\begin{align}
+  \bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} = \frac{\nabla
+    P}{\rhog}\tstop, 
+\end{align}
+\citep{youdin05a,jacquet11}.
+
+%In the limit of strong dust-gas drag (stopping time $\tstop\to 0$) 
+In this limit of strong dust-gas drag, \cite{laibe14} give a single
+fluid description of the mixture (their Eq. 84 --- 85 and Eq. 87):
 
@@
 where $D/Dt \equiv \p_t + \bm{v}\cdot\nabla$ is the Lagragian
-derivative; 
+derivative \emph{following the mixture} with center-of-mass velocity  
+\begin{align}
+  \bm{v} \equiv \frac{\rhog\bm{v}_\mathrm{g} + 
+    \rhod\bm{v}_\mathrm{d}}{\rho}, 
+\end{align}
+and 
 \begin{align}
@@
 \end{align}
-is the total density with $\rhog$ and $\rhod$ being the gas and dust
-density, respectively; 
-\begin{align}
-  \bm{v} \equiv \frac{\rhog\bm{v}_\mathrm{g} + 
-    \rhod\bm{v}_\mathrm{d}}{\rho}
-\end{align}
-is the center-of-mass velocity with $\bm{v}_\mathrm{g}$ and
-$\bm{v}_\mathrm{d}$ being the gas and dust velocities, respectively; 
-and 
+is the total mixture density.
+
+In Eq. \ref{dusteq}, 
 \begin{align}
@@
 
-The external potential $\Phi$ is 
+For the disk problem the external potential $\Phi$ is  
 that for a star of mass $M_*$, 
@@
 \subsection{Locally isothermal equation of state}\label{loc_iso_eos}
-We adopt 
+To complete the system of equations we need to specify a gas equation
+of state. We start with the 
+energy equation for an ideal gas of adiabatic index $\gamma$: 
+\begin{align}
+  \frac{\p P}{\p t} + \bm{v}_\mathrm{g}\cdot\nabla P = - \gamma P
+  \nabla\cdot \bm{v}_\mathrm{g}  + \mathcal{H} - \Lambda 
+%\frac{\rho \mathcal{R}}{\mu}\frac{T
+%  - T_\mathrm{ref}}{t_\mathrm{cool}}, \label{realenergy}
+\end{align}
+where $\mathcal{H} = 
+(\gamma-1)\tepsilon|\rhog\left(\bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}\right)|^2$
+accounts for heating due to dust-gas drag \citep{laibe14}; and 
+\begin{align}
+  \Lambda = \frac{\rho \mathcal{R}}{\mu}\frac{T
+  - T_\mathrm{ref}}{t_\mathrm{cool}}, \label{realenergy}
+\end{align}
+represents radiative cooling on a characteristic timescale
+$\tcool$ that relaxes the gas back to a prescribed temperature profile  
+$T_\mathrm{ref}$, where $\mathcal{R},\, \mu$ are  the gas constant and
+mean molecular weight of the gas, respectively.
+
+We now simplify the gas energy equation by considering short cooling 
+times $\tcool\to 0$, appropriate for the outer parts of an irradiated
+protoplanetary disk \citep{chiang97,lin15}. Then the disk temperature
+$T = T_\mathrm{ref}$ at all times, and so we may
+adopt a locally isothermal equation of state 
 \begin{align}\label{eos}
@@
 \end{align}
-where $c_s$ is a prescribed sound-speed profile fixed in time. 
+where $c_s(r,z)= \sqrt{\mathcal{R}T_\mathrm{ref}/\mu}$ is a prescribed
+sound-speed profile fixed in time.
+
 In practice we consider vertically 
@@
 \subsection{Effective energy equation}
-We show that for a prescribed temperature distribution, the mixture
-obeys an evolutionary energy equation, due to the advection of the
-dust-fraction. Eliminate $\tepsilon$ in favor of $P$, 
+Although we have eliminated the need of an evolutionary energy
+equation (Eq. \ref{realenergy}) by choosing a locally isothermal
+equation of state (Eq. \ref{eos}), we show that the mixture 
+nevertheless obeys an effective evolutionary energy equation. This is
+because advection of the dust-fraction, described by
+Eq. \ref{dusteq}, can be transformed into an energy-like 
+equation. We 
+%We show that for a prescribed temperature distribution, the mixture
+%obeys an evolutionary energy equation, due to the advection of the
+%dust-fraction. 
+eliminate $\tepsilon$ in favor of $P$, 
 \begin{align*}
@@
 \begin{align}
-\frac{DP}{Dt} &= - P \nabla\cdot\bm{v} + P\bm{v}\cdot\nabla\ln{c_s^2}
+%\frac{DP}{Dt} 
+\frac{\p P}{\p t} + \bm{v}\cdot\nabla P  
+&= - P \nabla\cdot\bm{v} + P\bm{v}\cdot\nabla\ln{c_s^2}
                 + \mathcal{C},  \label{eff_energy} \\
@@
 \end{align}
-Eq. \ref{eff_energy} is the energy equation for an ideal gas of adiabatic index
-$\gamma=1$, i.e. isothermal evolution, but now with additional energy source 
-terms due to the imposed temperature profile via $\nabla c_s^2$, as
-well as dust-gas drag via $\mathcal{C}$. 
+Comparing with Eq. \ref{realenergy}, we see that 
+Eq. \ref{eff_energy} is the energy equation for an ideal gas of
+adiabatic index $\gamma=1$, but now with the imposed temperature
+gradient $\nabla c_s^2$ and dust-gas drag $\mathcal{C}$ playing the
+role of radiative cooling ($\Lambda$ in 
+Eq. \ref{realenergy}).
+
 %In this form, the function $\mathcal{C}$ can be interpreted

### 2016-10-13 model.tex
edits in response to youdin's comments

@@
 \subsection{Entropy of an isothermal dusty gas }
-%Having made the connection between the dust fraction (or dust-to-gas
-%ratio) of the mixture and an energy equation, we now define 
-We define the (dimensionless) entropy of the mixture as 
+
+The gas component of the mixture has an entropy that is not conserved 
+because of the (locally) isothermal equation of state 
+adopted. However, we can define an effective (dimensionless)
+entropy for the \emph{entire mixture} as 
 \begin{align}
-   s \equiv \ln \frac{P}{\rho} = \ln{\left[c_s^2(1-\tepsilon)\right]}.  
+   \seff \equiv \ln \frac{P}{\rho} = \ln{\left[c_s^2(1-\tepsilon)\right]}.  
 \end{align} 
-We shall see that by defining the entropy this way, many of the 
+We shall see that by defining an effective entropy this way, many of the 
 results concerning the stability of dusty gas will have identical form
-and interpretations to that in standard adiabatic hydrodynamics.
+and interpretations to that in standard adiabatic hydrodynamics of 
+pure gas. For example, if $c_s^2$ is constant and $\tstop=0$, then
+$D\seff/ D t = 0$ since $D\tepsilon/D t=0$ in that case.
+
+The physical reason for this analogy is that with strong drag, the dust is almost
+perfectly entrained in the gas flow, but there is some gain/loss of 
+dust particles between different gas parcels if dust-gas
+friction if finite. This property is analogous to the entropy of an adiabatic
+pure gas subject to heating/cooling: entropy is conserved following
+a gas parcel, except if there is heat exchange between a fluid parcel
+and the surrounding.
 
@@
 the steady state effective energy Eq. \ref{eff_energy} 
-except in the limit of perfectly coupled dust ($\tstop\to0$).
+except in the limit of perfectly coupled dust ($\tstop\to0$).  In
+practice the dust-component would settle on a timescale
+$t_\mathrm{settle}\sim 1/\OmK^2\tstop$, but for tightly-coupled dust
+particles this is much longer than that to establish vertical
+hydrostatic equilibrium in the gas component ($\sim 1/\OmK$).
 
@@
 %  \frac{\p}{\p r} \left[c_s^2(1-\tepsilon)\right]\\  
-   = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p s}{\p z} -\frac{\p
-    P}{\p z}\frac{\p s}{\p r} \right). 
+   = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p \seff}{\p z} -\frac{\p
+    P}{\p z}\frac{\p \seff}{\p r} \right). 
 \end{align}
@@
   \frac{\p\epsilon}{\p r}\frac{\p\ln{\rhog}}{\p z}
-  -\frac{\p\epsilon}{\p z}\frac{\p\ln{\rhog}}{\p r}\right.\notag\\
+%  -\frac{\p\epsilon}{\p z}\left[\frac{\p\ln{\rhog}}{\p r} + \frac{q}{r}\right]\right.\notag\\
+ -\frac{\p\epsilon}{\p z}\frac{\p\ln{P}}{\p r}\right.\notag\\
   &\phantom{ r\frac{\p \Omega^2}{\p z}  =
     \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\left\{\right\} }
-  \left. -\frac{q}{r}\left[\frac{\p\epsilon}{\p z} + \left(1+\epsilon\right)\frac{\p\ln{\rhog}}{\p z}\right]
-  \right\} 
+  \left. -\frac{q}{r} \left(1+\epsilon\right)\frac{\p\ln{\rhog}}{\p z} \right\} 
 \end{align}
 The first two terms correspond to vertical shear caused by spatial
-variations in the dust-to-gas ratio; while the second term
+variations in the dust-to-gas ratio; while the third term
 proportional to $q$ corresponds to vertical shear due to the 
-radial temperature gradient.
+radial temperature gradient. The last term is present in the dust-free
+limit.
 
@@
         temp. gradient}}} \sim
+ % \epsilon \frac{\mathrm{max}\left(\delta^2,
+  %  1\right)}{\left|q\right|\left(\epsilon + \delta^2\right)},
  \epsilon \frac{\mathrm{max}\left(\delta^2,
-    1\right)}{\left|q\right|\left(\epsilon + \delta^2\right)},
+    1\right)}{\left|q\right|\left(1+\epsilon\right)\delta^2},
 \end{align}
@@
 Since $|q|=O(1)$ in PPDs, Eq. \ref{vshear_split} indicates that
-vertical shear due to variations in the dust-to-gas ratio can be at
-most comparable to that due to the radial temperature gradient. 
-This only occurs for thin dust layers with $\delta^2\ll \epsilon$ (so 
-the above ratio $\sim 1/|q|$). Otherwise, vertical shear is associated
-with $\p_rc_s^2$.
+vertical shear due to variations in the dust-to-gas ratio dominates 
+over that due to the radial temperature gradient for thin dust layers
+such that $\delta^2\ll \epsilon$. Otherwise, vertical shear is
+associated with $\p_rc_s^2$.
 
@@
 \begin{align}
-  N_z^2 \equiv - \frac{1}{\rho}\frac{\p P}{\p z}\frac{\p s}{\p z} &=
+  N_z^2 \equiv - \frac{1}{\rho}\frac{\p P}{\p z}\frac{\p \seff}{\p z} &=
   \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\frac{\p\ln\rhog}{\p 

### 2016-10-14 model.tex
more setup

@@
 velocity field as $(\rhog,\bm{v}_\mathrm{g})$ and
-$(\rhod,\bm{v}_\mathrm{d})$, respectively. The two fluids 
-interact via a drag force parameterized by a stopping time $\tstop$
+$(\rhod,\bm{v}_\mathrm{d})$, respectively. The mixture has a single 
+temperature $T$, and the mixture pressure $P$ arise solely from the 
+gas component.
+
+The two fluids interact via a drag force parameterized by a stopping
+time $\tstop$ 
 such that 
@@
 In this limit of strong dust-gas drag, \cite{laibe14} give a single
-fluid description of the mixture (their Eq. 84 --- 85 and Eq. 87):
-
+fluid description of the mixture: 
 \begin{align}
@@
   &\frac{D\tepsilon}{Dt} = -\frac{1}{\rho} \nabla \cdot \left(\tepsilon 
-  \tstop \nabla P \right),\label{dusteq}  
+  \tstop \nabla P \right),\label{dusteq}\\
+  &\frac{D T}{D t} = - \left(\gamma-1\right)T\nabla\cdot\bm{v} +
+  \mathcal{H}_\mathrm{drag} + \mathcal{H}_\mathrm{eff}  - \Lambda, \label{tempeq} 
 \end{align}
@@
 \end{align}
-is the total mixture density.
+is the total mixture density. For an ideal gas the pressure is given
+by $P = \mathcal{R}\rhog T/\mu $, where $\mathcal{R}$ is
+the gas constant and $\mu$ is the mean molecular weight.
 
@@
 is the dust-fraction and $\epsilon=\rhod/\rhog$ is the usual
-dust-to-gas ratio. The gas pressure $P$ is given by the chosen 
-equation of state described below. Note that Eq. \ref{dusteq} is
+dust-to-gas ratio. Note that Eq. \ref{dusteq} is 
 equivalent to Eq. 46 of \cite{jacquet11}.
+
+For the mixture's temperature evolution, Eq. \ref{tempeq}, $\gamma$ is
+the adiabatic index; $\mathcal{H}_\mathrm{drag}$ represents heating
+due to dust-gas drag and $\mathcal{H}_\mathrm{eff}$ is an effective
+energy source arising from transforming from the two-fluid to
+one-fluid description \citep[see][ for details]{laibe14}. We also
+include a simple model of radiative cooling,
+\begin{align}
+  \Lambda = \frac{T -
+    T_\mathrm{ref}}{t_\mathrm{cool}}, \label{realenergy} 
+\end{align}
+which relaxes the mixture back to a prescribed temperature profile   
+$T_\mathrm{ref}$ on a timescale of $\tcool$. We will shortly simplify
+the problem by considering rapid cooling, $\tcool\to0$.
 
@@
 \subsection{Locally isothermal equation of state}\label{loc_iso_eos}
-To complete the system of equations we need to specify a gas equation
-of state. We start with the 
-energy equation for an ideal gas of adiabatic index $\gamma$: 
+For later comparison it is convenient to re-write the temperature
+evolution, Eq. \ref{tempeq}, to that for the pressure: 
 \begin{align}
   \frac{\p P}{\p t} + \bm{v}_\mathrm{g}\cdot\nabla P = - \gamma P
-  \nabla\cdot \bm{v}_\mathrm{g}  + \mathcal{H} - \Lambda 
+  \nabla\cdot \bm{v}_\mathrm{g}  + \mathcal{H} 
 %\frac{\rho \mathcal{R}}{\mu}\frac{T
@@
 \end{align}
-where $\mathcal{H} = 
-(\gamma-1)\tepsilon|\rhog\left(\bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}\right)|^2$
-accounts for heating due to dust-gas drag \citep{laibe14}; and 
-\begin{align}
-  \Lambda = \frac{\rho \mathcal{R}}{\mu}\frac{T
-  - T_\mathrm{ref}}{t_\mathrm{cool}}, \label{realenergy}
-\end{align}
-represents radiative cooling on a characteristic timescale
-$\tcool$ that relaxes the gas back to a prescribed temperature profile  
-$T_\mathrm{ref}$, where $\mathcal{R},\, \mu$ are  the gas constant and
-mean molecular weight of the gas, respectively.
+where $\mathcal{H}$ collects together dust-drag heating and radiative
+cooling.
 

### 2016-10-16 model.tex
edit tex

@@
 \begin{align}
-  \delta f(r,z)\exp{\left(-\ii\sigma t\right)},
+ \real\left[ \delta f(r,z)\exp{\left(-\ii\sigma t\right)},\right]
 \end{align}
@@
 
-%except here the entropy $s$ is really a
-%representation of the dust distribution via $s=\ln{\left[c_s^2\left(1
-%    - \tepsilon\right)\right]}$.
-
-%These criteria are applicable to any adiabatic fluid. 
-%For an isothermal mixture, we may take $s=\ln{(1-\tepsilon)}$.   
-%from the definition of $s$ and  
-%the equation of state (Eq. \ref{eos}) with uniform sound-speed. 
-%\begin{align}
-%  s \equiv \ln \frac{P}{\rho} = \ln{(1-\tepsilon)} 
-%\end{align}
-% The first equality is
-%similar to the usual defintion for gas entropy (with unit adiabatic
-%index). The second follows from the imposed equation of state,
-%Eq. \ref{eos}, with a constant sound-speed. 
-%The stability conditions are thus
-%In accretion disks, $\rhog$, $\rhod$ and $\tepsilon$ are expected to
-%decrease away from the mid-plane
-%\begin{align*}  
-%  \frac{\p\rhog}{\p |z|} < 0,
-%, \quad \frac{\p \rhog}{\p r} < 0,   
-%\end{align*}
-%and similarly for $\rhod$, $\tepsilon$, and the dust-to-gas ratio $\epsilon$. 
-%We may thus evaluate the stability criteria at $z>0$ without loss of
-%generality.
-
-%We insert the vertical shear profile,
-%Eq. \ref{vshear}, into Eq. \ref{dusty_solberg} and write the result 
-%in terms of $\epsilon$, to find
-
 \subsubsection{Order-of-magnitude estimates} 
@@
 in thin accretion disks where radial gradients are $O(1/r)$, 
-$h_\mathrm{g},\epsilon,\ll 1$ and $\Htilde/H_\mathrm{g}$ not too
-large.  
-Then the
+$h_\mathrm{g}\ll 1$ and $\Htilde/H_\mathrm{g}$ not too
+large.  Then the
 magnitude of the second term on the left-hand-side  of either
@@
 \end{align}
-%In this case 
-%the effective energy equation is 
-%\begin{align}\label{dusty_cooling}
-%  \frac{DP}{Dt} &= - P \nabla \cdot \bm{v} + \mathcal{C}
-%\end{align} 
-%The function $\mathcal{C}$ can be considered as a cooling term. 
-%By
-%analogy with enabling VSI in an adiabatic gas through cooling
-%\citep{nelson13,lin15}, we may expect a `dusty VSI' to develop 
-%if this term is large enough (physically meaning diffusion of the
-%dust-fraction is strong). 
-%Note that $\mathcal{C}$ is also responsible for the streaming
-%instability  in the strong-drag limit \citep{laibe14}, though this
-%does not require vertical stratification. In the form of
-%Eq. \ref{dusty_cooling}, however, the streaming instability has a
-%thermodynamical interpretation, similar to thermal instabilities in
-%stars \citep{latter06}.  
 which represent instabilities due to dust-gas friction. 
-%Our formalism
-%gives these dusty instabilities a thermodynamical interpretation. 
-% We further discuss this concept in \S\ref{dust_as_thermo}. 
-%In \S\ref{dust_work} we showed that instabilities may 
-%arise from the term proportional $\imag\left( \nabla\cdot\dd\bm{v}^*
-%\delta C\right)$. 
 Since drag ($\delta \mathcal{C}$) appears as a source term in our
@@
 qualitative 
-description below, based on the more rigorous treatments of stellar 
+description below, adapted from the more rigorous treatments of stellar 
 pulsations by 
-\cite{cox67} and lecture notes from \cite{samadi15} and
+\cite{cox67} and lecture notes by \cite{samadi15} and
 J. Christensen-Dalsgaard\footnote{\url{http://astro.phys.au.dk/$\sim$jcd/oscilnotes/}}.
@@
 %\cite{kato78}.
+
+\subsubsection{Work done by dusty gas}
 
@@
   -|\sigma|^2\imag \left(\Delta P 
-  \frac{\Delta\rho^*}{\rho}\right). 
+  \frac{\Delta\rho^*}{\rho}\right), \label{pdv}
 \end{align}
-This reflects `$TdS  =  PdV$' from thermodynamics, if the 
-internal energy is constant. That is, dust-gas drag provides 
-heating ($TdS$) to do $PdV$ work. Whether positive or negative work is 
-done depends on the correlation between the pressure and density
-perturbations, as follows.
+which states that `PdV' work (right hand side) is
+provided by dust-drag friction. This can be seen explicitly as
+follows.
+
+The average rate of work done by the dusty-gas is
+\begin{align}
+  \mathcal{W} = \frac{1}{T_p}\int^{t+T_p}_{t}dt^\prime\int_M P
+  \frac{Dv}{Dt^\prime} dm 
+\end{align}
+\citep{cox67}, 
+where $T_p$ is the oscillation period, $v=1/\rho$ is the specific
+volume of the mixture, and the spatial integral is taken over the
+total mass $M$ of the mixture and $dm$ is the mass element.
+
+Our aim is to relate the work done by the mixture, $\mathcal{W}$, 
+to the growth rate of oscillation amplitudes, $s$. To do this, we
+first evaluate $\mathcal{W}$ assuming pure oscillations (perturbations
+with time dependence $e^{\ii\omega t}$ and real $\omega$), then show
+that in fact the amplitudes grow if the work done is positive.
+
+Noting that only products in the perturbations contribute to 
+$\mathcal{W}$, we find that for real $\omega$: 
+\begin{align}
+  \mathcal{W} = - \frac{\omega}{2} \int \imag\left(\Delta P
+  \frac{\Delta \rho^*}{\rho}\right)dV,\label{work_real}
+\end{align}
+where the integral is taken over the volume of the
+mixture. Eq. \ref{work_real} show that a phase difference between
+pressure and density leads to work done.
+
+Now combining Eq. \ref{work_real}, \ref{pdv} and \ref{thermal_instability}
+gives  
+\begin{align}
+s = \frac{2}{\mathcal{I}^2}\mathcal{W}, \label{growth_work}
+\end{align}
+where $\mathcal{I}^2>0$  is the integral in the denominator of
+Eq. \ref{thermal_instability}, and we have set $\sigma= - \omega$ in
+Eq. \ref{pdv} for pure oscillations.  Thus growing modes in fact arise if
+positive work is done by the mixture during an oscillation. This in
+turn requires pressure to lag density, discussed next.
+
+\subsubsection{Physical picture of dust-drag instabilities}
 

### 2016-10-16 model.tex
improved pdv interpret

@@
 \subsubsection{Physical picture of dust-drag instabilities} 
-
 The full energy equation is 
@@
 \end{align*}
-Dust-drag causes a phase difference between the 
-pressure and density evolution of a fluid element. Now, consider a 
-fluid element at maximum density during an oscillation cycle, 
-so that $D\rho/ D t = 0$. If at this point dust-gas drag provides an
-effective heating to the fluid element ($\mathcal{C}>0$, i.e. a particle
-flux \emph{away} from it), then it will 
-experience increasing pressure, $DP/D t>0$. It will attain  
-pressure maximum \emph{after} reaching density maximum. This implies 
-pressure maximizes while the parcel is expanding: it will do 
-positive work  --- the area enclosed by the parcel's path in the `$P$-$V$' plane  is 
-positive. This will increase the oscillation amplitude.
+Dust-drag causes a phase difference between the pressure and density 
+evolution of a fluid element. In Fig. \ref{pdv_cartoon}, we show that
+if pressure lags behind density, then positive
+work done ($\mathcal{W}>0$) and would lead to growing oscillations
+($s>0$).
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/drag}\\
+\includegraphics[width=\linewidth]{figures/pdv}
+  \caption{Thermodynamic interpretation of overstable modes caused by
+    dust-gas drag. Top: pressure and density evolution in time. 
+    Bottom: oscillation cycle in the $P$-$V$ plane. In the case shown,
+    dust-drag causes 
+    pressure to lag behind density, which results in a clockwise path in
+    the $P$-$V$ plane, implying positive work done by the fluid and
+    hence would lead to growing oscillation amplitudes.  
+    \label{pdv_cartoon}
+  }
+\end{figure}
+
+Consider a fluid element at maximum density during an oscillation cycle
+(point $A$), so that $D\rho/ D t = 0$. If dust-gas drag provides an
+effective heating to the fluid element at $A$ (i.e. gas influx)  then
+it will experience increasing pressure, $DP/D t>0$. It attains
+pressure maximum (point $B$) \emph{after} reaching density maximum.   
+This implies pressure maximizes while the parcel is expanding: 
+the parcel's path in the `$P$-$V$' plane  is clockwise, as shown in
+the bottom panel of Fig. \ref{pdv_cartoon}. This corresponds to
+positive work done.
+
+Our thermodynamic interpretation does not rigorously explain
+\emph{why} dust drag causes a pressure lag behind density, but merely shows
+that this \emph{must} be the case for any growing modes associated the
+dust drag. To understand how dust-drag causes this lag requires an
+explicit solution to the linearized equations with detailed treatment
+of the function $\mathcal{C}$. However, given the complexity of
+$\mathcal{C}$ (see Appendix \ref{lin_dust}), we might generally expect
+the system to  support a range of stable and unstable modes, with the
+latter being associated with pressure-density lag.
 
@@
 \citet{youdin07b}.
-
-We suggest understanding dusty instabilities in
-protoplanetary disks by re-casting the problem into that of
-thermodynamics. Then dust-drag instabilities appear in the equations in the 
-same way as pulsational instabities in stars, which has already been 
-well studied. Whether or not the dusty disk is actually unstable
-depends on details of the dust-drag or `cooling' function
-$\mathcal{C}$. However, given the complexity of $\mathcal{C}$ (see
-Appendix \ref{lin_dust}), we might generally expect the system to 
-support a range of stable and unstable modes. On physical grounds,
-growing modes \emph{must} be so by causing a phase lag between
-pressure and total density.
 

### 2016-10-17 model.tex
rough edits for youdin's comments

@@
 \begin{align}
- \real\left[ \delta f(r,z)\exp{\left(-\ii\sigma t\right)},\right]
+ \real\left[ \delta f(r,z)\exp{\left(-\ii\sigma t\right)}\right], 
 \end{align}
@@
 
-Our thermodynamic interpretation does not rigorously explain
+Our thermodynamic interpretation does not explain
 \emph{why} dust drag causes a pressure lag behind density, but merely shows
 that this \emph{must} be the case for any growing modes associated the
-dust drag. To understand how dust-drag causes this lag requires an
+dust drag. To rigorously understand how dust-drag causes this lag requires an
 explicit solution to the linearized equations with detailed treatment
@@
 $\mathcal{C}$ (see Appendix \ref{lin_dust}), we might generally expect
-the system to  support a range of stable and unstable modes, with the
+the dusty disks to  support a range of stable and unstable modes, with the
 latter being associated with pressure-density lag.

### 2016-10-17 result.tex
tiny changes
rough edits for youdin's comments

@@
 buoyancy to dominate over vertical shear in the disk atmosphere 
-$|z|\gtrsim 2.5$. We thus expect instability all heights for 
+$|z|\gtrsim 2.5\Hg$. We thus expect instability all heights for 
 $\epsilon_0=10^{-3}$, but to be restricted to the midplane for
@@
 
+\subsection{Axisymmetric stability of ultra-thin dust layers}
+In \S\ref{iso_perfect} we asserted that in the limit of perfectly
+coupled dust (as considered here) that radially smooth, strictly
+isothermal disks are stable against axisymmetric perturbations, no
+matter how thin the dust layer is. We now demonstrate this
+numerically.
+
+For comparison with previous studies of dusty disks, we also use 
+the Richardson number $\rich \equiv N_z^2/\left(r\p_z\Omega\right)^2$
+to label the calculations here \citep{youdin02}.   
+Numerical simulations show   
+non-axisymmetric instabilities develops when $\rich\lesssim0.1$, brought about
+by very thin dust layers \citep{chiang08, lee10}. We show that \emph{axisymmetric}
+instabilities never develop, however small $\rich$.
+
+We shall consider ultra thin dust layers with $\Hd\leq 0.01\Hg$ and thus
+restrict the vertical domain to $\zmax = 0.02\Hg$. We fix the 
+metalicity $Z=0.01$ so the midpane dust-to-gas ratio $\epsilon_0$ is 
+$O(1)$.  We consider radial wavenumbers such $k\Hd=1$.
+
+Fig. \ref{ultra_thin} shows the maximum growth rate  as a function of
+the radial temperature gradient, $q$. For all cases the vertical
+shear is dominated by that due to the dust layer (see
+\S\ref{vertshear}). However, we see that growth rates $s\propto
+|q|$. In particular, this holds for $\rich < 0.1$, the critical value
+for non-axisymmetric instabilities.
+
+Axisymmetric instability is associated with the thermal
+contribution to vertical shear: as $q\to0$,  $\p_z\Omega$ becomes
+entirely due to dust-settling and there is no instability ($s\to
+0$). This result is independent of $\rich$, so the Richardson number
+is not useful in characterizing the axisymmetric stability of dust
+layers.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_eigenvals_thindust} 
+  \caption{Maximum VSI growth rate for ultra-thin dust layers $\Hd
+    \leq 0.01\Hg$. Vertical shear is dominated by that due to
+    dust-settling, but axisymmetric instability is associated with the
+    radial temperature gradient $q$.
+    \label{ultra_thin}
+    }
+\end{figure}
+
 \subsection{Effect of a radially-varying dust-to-gas ratio}\label{varHd}

### 2016-10-17 appendix.tex
tiny change
added loren's instability
section on radially varying epsilon
tiny typo fix
edits in response to youdin's comments
added self-gravity in appendix 1

@@
 and $\Gamma$ is a constant. The effective energy equation is then
-Eq. \ref{poly_energy}. The linearized equations for axisymmetric
+Eq. \ref{poly_energy}. We also include the self-gravitational
+potential $\psi$ of the gas
+plus dust mixture, which satisfies 
+%\begin{align}
+$  \nabla^2 \psi  = 4 \pi G \rho. $
+%\end{align} 
+The linearized equations for axisymmetric
 disturbances are
@@
   -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
-  \delta\bm{F}\cdot\hat{\bm{r}},\\
+  \hat{\bm{r}} \cdot \delta\bm{F} -  \hat{\bm{r}} \cdot \nabla\dd\psi ,\\
   \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
     v_\phi}{\p z}\dd v_z,\\
-  -\ii\sigma\dd v_z &=  \delta\bm{F}\cdot\hat{\bm{z}}, 
+  -\ii\sigma\dd v_z &=  \hat{\bm{z}} \cdot \delta\bm{F}  -  \hat{\bm{z}} \cdot \nabla\dd\psi, \\ 
+\nabla^2 \delta\psi & = 4\pi G \dd\rho. 
 \end{align}  
@@
   v_r^*  + \ii\sigma \rho \dd \bm{F}\cdot\dd\bm{v}^*
-  \right)dV,
+  - \ii \sigma \rho \nabla\dd\psi\cdot  \dd\bm{v}^* 
+  \right)dV,\label{meridional}
 \end{align}
@@
    \left(
-   \ii\sigma \frac{\dd P}{\Gamma P} - \dd\bm{v}\cdot\nabla s + \frac{\dd
+   \ii\sigma \frac{\dd P}{\Gamma P} - \dd\bm{v}\cdot\nabla \seff + \frac{\dd
    \bm{v}}{\Gamma} \cdot \nabla \ln{K} +
@@
 %   \right]dV\notag\\
-   \ii\sigma\frac{\dd P}{\Gamma P}\left(\bm{\dd v}^*\cdot\nabla P +
-   \Gamma P\nabla\cdot\bm{\delta v}^*   \right)
+   \ii\sigma\frac{\dd P}{\Gamma P}\left(\dd\bm{v}^*\cdot\nabla P +
+   \Gamma P\nabla\cdot\dd\bm{v}^*   \right)
    +\left(\dd\bm{v}^*\cdot\nabla P\right) \left( \frac{\dd\bm{v}}{\Gamma} \cdot \nabla
-   \ln{K} + \frac{\dd\mathcal{C}}{\Gamma P}   -  \dd\bm{v}\cdot\nabla s\right) 
+   \ln{K} + \frac{\dd\mathcal{C}}{\Gamma P}   -  \dd\bm{v}\cdot\nabla \seff\right) 
    \right]dV\notag\\
@@
   \right]
- \left(\bm{\dd v}^*\cdot\nabla P +
- \Gamma P\nabla\cdot\bm{\delta v}^*   \right)\right. \notag\\
+ \left(\dd \bm{v}^*\cdot\nabla P +
+ \Gamma P\nabla\cdot\dd\bm{v}^*   \right)\right. \notag\\
  &\phantom{=\int\left\{\right\}}\left. 
@@
  \ln{K}
- + \frac{\dd\mathcal{C}}{\Gamma P}   -  \dd\bm{v}\cdot\nabla s \right)
+ + \frac{\dd\mathcal{C}}{\Gamma P}   -  \dd\bm{v}\cdot\nabla \seff \right)
  \right\}dV \notag\\
@@
 %     P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2} + \frac{\dd\mathcal{C}}{P} \right)  
-    \frac{1}{\Gamma P} \Bigl\lvert \bm{\dd v}\cdot\nabla P +
- \Gamma P\nabla\cdot\bm{\delta v}    \Bigr\rvert ^2 - 
-     \left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v}\cdot\nabla s
+    \frac{1}{\Gamma P} \Bigl\lvert \dd \bm{v}\cdot\nabla P +
+ \Gamma P\nabla\cdot\dd \bm{v}    \Bigr\rvert ^2 - 
+     \left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v}\cdot\nabla \seff
      \right) -
@@
 \end{align}
-where  $s\equiv\ln{(P^{1/\Gamma}/\rho)}$. Hence,
+where  $\seff\equiv\ln{(P^{1/\Gamma}/\rho)}$. 
+The self-gravitational part of Eq. \ref{meridional} is, again
+nelgecting surface terms when integrating by parts, 
+\begin{align}
+-\int \ii\sigma \rho \nabla\dd\psi \cdot\dd  \bm{v}^* &= \int \ii\sigma
+  \dd\psi \nabla\cdot\left(\rho\dd\bm{v}^*\right) dV
+ = \int \left|\sigma\right|^2 \dd\psi \dd \rho^*dV 
+  = -\frac{1}{4\pi G}\int   \left|\sigma\right|^2
+   \left|\nabla\psi\right|^2 dV,  
+\end{align}
+where the linearized continuity and Poisson equations have been used.
+
+Hence,
 \begin{align}\label{integral_ex}
@@
   \rho |\dd v_r|^2 \left(\kappa^2 - \frac{1}{\rho}\frac{\p P}{\p
-    r}\frac{\p s}{\p r}\right)
+    r}\frac{\p \seff}{\p r}\right)
   + \rho |\dd v_z|^2\left(-\frac{1}{\rho}\frac{\p P}{\p
-    z}\frac{\p s}{\p z}\right)
+    z}\frac{\p \seff}{\p z}\right)
    + \rho \dd v_z \dd v_r^*\left(r\frac{\p\Omega^2}{\p z} -
   \frac{1}{\rho}\frac{\p P}{\p
-    r}\frac{\p s}{\p z}\right) 
+    r}\frac{\p \seff}{\p z}\right) 
   + \rho \dd v_z^*\dd v_r \left(-\frac{1}{\rho}\frac{\p P}{\p
-    z}\frac{\p s}{\p r}\right)\right. \notag\\
+    z}\frac{\p \seff}{\p r}\right)\right. \notag\\
 &\phantom{==\int}\left.+
 %  \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2
- \frac{1}{\Gamma P}\Bigl\lvert \bm{\dd v}\cdot\nabla P +
- \Gamma P\nabla\cdot\bm{\delta v}     \Bigr\rvert^2
+ \frac{1}{\Gamma P}\Bigl\lvert \dd\bm{v}\cdot\nabla P +
+ \Gamma P\nabla\cdot\dd\bm{v}     \Bigr\rvert^2
+ - \frac{1}{4\pi G}\left|\sigma\nabla\dd\psi\right|^2 
   \right\}dV 
@@
 v_z^*\delta v_r$ are equal owing to the equilibrium state
-(Eq. \ref{vshear}). The special case $\Gamma=1,\, K = c_s^2$ gives
+(Eq. \ref{vshear}). The non-self-gravitating case with $\Gamma=1,\, K = c_s^2$ gives
 Eq. \ref{int_rel}. Similar integral relations are given by 
@@
   k_x \tstop F_r\left(1 - 2\tepsilon\right)
-  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0.
+  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0.\label{streaming_dispersion}
 \end{align}
@@
 \begin{align*}
-\ii \tepsilon \tstop
+\ii \tepsilon \tau_\mathrm{s}
   \sigma^4 - \sigma^3 - 
@@
 as derived by \cite{jacquet11} and \cite{laibe14} for the streaming
-instability. If $|\sigma| = O(\tau_\mathrm{s})$ then the quartic term
+instability. If $|\sigma|/\OmK = O(\tau_\mathrm{s}\OmK)$ and
+$\tau_\mathrm{s}\OmK\ll 1$, then the quartic term
 is small and may be neglected. In that case we obtain the cubic

### 2016-10-18 appendix.tex
adding SI to main text

@@
 
-\section{Compressible streaming instability}\label{compressible_streaming}
+\section{Dispersion relation in an unstratified isothermal dusty disk}\label{compressible_streaming}
 We consider an unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =

### 2016-10-18 linear.tex
adding SI to main text

@@
 is retained. Numerical solutions are generally required for the 
-verticall-stratified problem, as we pursue below.
+verticall-stratified problem. We pursue this in \S\ref{results}
 
-However, in the special case of an unstratified disk (neglecting the vertical
+\section{Compressible streaming instability}  
+In the special case of an unstratified disk (neglecting the vertical
 component of the stellar gravity), we may also Fourier analyze in $z$
-to obtain a local dispersion relation. This is done in 
-Appendix \ref{compressible_streaming}. There we show the above linearized
+to obtain a local dispersion relation of the form 
+$\sum_{j=1}^{5}c_j\sigma^j = 0$. 
+This is done in 
+Appendix \ref{compressible_streaming}.
+
+There we show the above linearized
 equations give the streaming instability in the appropriate limit 
@@
 
+\begin{deluxetable*}{llrrrrr}
+  \tablecolumns{7}
+  \tablecaption{Selected eigenfrequencies for the streaming
+    instability. \label{bstar_compare}
+  }
+  \tablehead{
+    \colhead{Mode} & 
+    \colhead{$\tau_\mathrm{s}\OmK$} &
+    \colhead{$\epsilon$} &
+     \colhead{$K_x = K_z$} &
+      \colhead{$\sigma/\OmK$ (two-fluid)} &
+    \colhead{$\sigma/\OmK$ (one-fluid, Eq. \ref{streaming_dispersion})}   &
+    \colhead{$\mathcal{W}$ (arbitrary units) } 
+  }
+\startdata
+ linA, \cite{youdin07} &  $0.1$       & 3.0 & 30    & $-0.3480 +
+ 0.4190\ii$ & $0.3647 + 0.4251\ii$ & $0.196$\\
+
+linB, \cite{youdin07} & $0.1$        &  0.2 & 6 & $0.4999 +
+0.0154\ii$&   $-0.4999 + 0.0054\ii$  & $6.95$ \\
+
+linC,  \cite{bai10b}  & $10^{-2}$   &  2.0 & 1500&   $0.1049 +
+0.5981\ii$   &  $0.1344 + 0.6665\ii$  & $0.0198$\\
+
+linD, \cite{bai10b} &  $10^{-3}$   &  2.0 & 2000 & $0.3225 +
+0.3154\ii$& $0.3228 + 0.3155\ii$ &  $1.77\times10^{-3}$
+\enddata
+\end{deluxetable*}
+
 %\subsection{Boundary conditions}
-
 %Eq. \ref{lin_mass}---\ref{lin_energy} can be reduced to a set of

### 2016-10-18 appendix.tex
notation change (k to kx), fig update, SI in main text

--- 2016-10-17 appendix.tex
@@
 \begin{align}
-  &\dd F_r = - W F_r - \ii k Q,\\
+  &\dd F_r = - W F_r - \ii k_x Q,\\
   &\dd F_z = - W F_z - \left[Q^\prime + Q
@@
 \begin{align}
-  \nabla\cdot\dd\bm{F} &= F_r\left(\p_r\ln{\rho} - \ii k \right)W - F_z
-  W^\prime - W \nabla\cdot\bm{F} + \ii kQ \p_r\ln{\rho} +
+  \nabla\cdot\dd\bm{F} &= F_r\left(\p_r\ln{\rho} - \ii k_x \right)W - F_z
+  W^\prime - W \nabla\cdot\bm{F} + \ii k_x Q \p_r\ln{\rho} +
   \left(\ln{\rho}\right)^\prime\left[Q^\prime + Q
     \left(\ln{\rho}\right)^\prime\right] \notag\\
-  & \phantom{=} + k^2 Q - \left\{Q^{\prime\prime} + 2
+  & \phantom{=} + k_x^2 Q - \left\{Q^{\prime\prime} + 2
     Q^\prime\left(\ln{\rho}\right)^\prime
@@
       2}\right]\right\}\notag\\
-    &= \left[\left(\p_r\ln{\rho}-\ii k\right)F_r - \nabla\cdot\bm{F} +
-      F_z^\prime\right]W + \left(\ii k \p_r\ln{\rho} + k^2\right)Q  -
+    &= \left[\left(\p_r\ln{\rho}-\ii k_x\right)F_r - \nabla\cdot\bm{F} +
+      F_z^\prime\right]W + \left(\ii k_x \p_r\ln{\rho} + k_x^2\right)Q  -
       \ii\sigma \dd v_z^\prime.
@@
   \dd\tepsilon      &= (1-\tepsilon)W - \frac{Q}{c_s^2}, \\
-  \p_r\dd\tepsilon &= \left[(1-\tepsilon)\left(\ii k -
+  \p_r\dd\tepsilon &= \left[(1-\tepsilon)\left(\ii k_x -
     \p_r\ln{\rho}\right) - \p_r\tepsilon\right]W + \left(\p_r\ln{c_s^2}
-  + \p_r\ln{\rho} - \ii k \right)\frac{Q}{c_s^2},\\
+  + \p_r\ln{\rho} - \ii k_x \right)\frac{Q}{c_s^2},\\
   \dd\tepsilon^\prime &= (1-\tepsilon)W^\prime - \tepsilon^\prime W -
@@
 
-\section{Compressible streaming instability}\label{compressible_streaming}
+\section{One-fluid dispersion relation for the streaming instability}\label{compressible_streaming}
 We consider an unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
@@
 \begin{align}
-  \ii \sigma W &=\nabla\cdot \bm{\dd v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\\
+  \ii \sigma W &=\nabla\cdot \dd \bm{ v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\\
     \sigma^2 \dd v_x &= \kappa^2 \dd v_x - \ii \sigma F_r W +
@@
   -\ii\sigma\dd v_z &= -\ii k_zQ,\\
-\ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \bm{\dd v}   -
+\ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \dd\bm{v}   -
   \frac{\dd \mathcal{C}}{\rho}. 

### 2016-10-18 linear.tex
notation change (k to kx), fig update, SI in main text

--- 2016-08-24 linear.tex
@@
 \begin{align}
-  \delta f (r, z) = \delta f_1(r,z)\exp{(\ii k r)},
+  \delta f (r, z) = \delta f_1(r,z)\exp{(\ii k_x r)},
 \end{align} 
 %and similarly for $\dd P$ and $\dd\bm{v}$. 
-where $k$ is a real wavenumber such that $|kr|\gg 1$, and the
+where $k_x$ is a real wavenumber such that $|k_xr|\gg 1$, and the 
 amplitude $\dd f_1(r,z)$ is 
 a slowly-varying function of $r$. Then 
-$\p_r\to i k$ when acting on the above primitive perturbations, and we may
+$\p_r\to i k_x$ when acting on the above primitive perturbations, and we may
 neglect curvature terms. We take  
-$k>0$ without loss of generality. Hereafter, we drop the subscript 1
+$k_x>0$ without loss of generality. Hereafter, we drop the subscript 1
 on the amplitudes. 
@@
 \begin{align}
-  \ii\sigma W &= \ii k \dd v_r + \dd v_z^\prime +
+  \ii\sigma W &= \ii k_x \dd v_r + \dd v_z^\prime +
   \dd v_r \p_r\ln{\rho} + \dd v_z\p_z\ln{\rho},\label{lin_mass}\\
@@
 % +  \delta\bm{F}\cdot\hat{\bm{r}}
-- W F_r - \ii k Q,\label{lin_xmom}\\
+- W F_r - \ii k_x Q,\label{lin_xmom}\\
   \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
@@
 ,\label{lin_zmom}\\
-  \ii\sigma Q &= \frac{P}{\rho}\left(\ii k \dd v_r + \dd
+  \ii\sigma Q &= \frac{P}{\rho}\left(\ii k_x \dd v_r + \dd
                v_z^\prime\right) + \frac{1}{\rho}\left(\dd v_r\p_rP + \dd v_z \p_zP\right)\notag\\
@@
 evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
-is retained. Numerical solutions are generally required for the 
-verticall-stratified problem, as we pursue below.
+is retained. We now discuss explicit solutions to these equations. 
+% Numerical solutions are generally required for the 
+%vertically-stratified problem.
 
-However, in the special case of an unstratified disk (neglecting the vertical
+\section{Compressible streaming instability} 
+In the special case of an unstratified disk (neglecting the vertical
 component of the stellar gravity), we may also Fourier analyze in $z$
-to obtain a local dispersion relation. This is done in 
-Appendix \ref{compressible_streaming}. There we show the above linearized
-equations give the streaming instability in the appropriate limit 
-\citep{youdin05a,jacquet11}.
+to obtain a local dispersion relation of the form  
+$\sum_{j=1}^{5}c_j(k_x,k_z)\sigma^j = 0$, where $k_z$ is a real
+vertical wavenumber. The coefficients $c_j$ can be read 
+off Eq. \ref{streaming_dispersion} in  Appendix \ref{compressible_streaming}. 
+There we also show that this dispersion relation reduces to that for
+the streaming instability in the limit of incompressible gas and small
+$\tstop$ \citep{youdin05a,jacquet11}.
+
+We solve Eq. \ref{streaming_dispersion} for selected cases where the  
+growth rates of the streaming instability have been 
+calculated analytically from full two-fluid disk models, as well as
+numerically from particle-gas simulations 
+\citep[namely][]{youdin07,bai10b}. These authors use normalized
+wavenumbers $K_x = \eta r k_x$ and similarly for $K_z$, where
+\begin{align} 
+  \eta \equiv -\frac{1}{2\rhog r\OmK^2}\frac{\p P}{\p r} = 
+  \frac{1}{2\left(1-\tepsilon\right)}\frac{F_r}{r\OmK^2}, 
+\end{align} 
+measures the pressure offset of Keplerian rotation. We use
+$\eta=0.05c_s/r\OmK$. They also parametrize the stopping time with
+$\tau_\mathrm{s}=\tstop/(1-\tepsilon)$.
+
+Table \ref{si_compare} compares the eigenfrequencies obtained from the
+one-fluid dispersion relation to that from the above studies. As
+expected eigenfrequencies agree better with decreasing $\tstop$ since
+in that limit the mixture behaves more like a single fluid. Most
+importantly, we find the work done $\mathcal{W}>0$ in all cases, and
+hence find growing oscillations. Thus the streaming instability stems
+from the pressure evolution lagging behind that of the total density.
+
+\begin{deluxetable*}{llrrrrr}
+  \tablecolumns{7}
+  \tablecaption{Selected eigenfrequencies for the streaming
+    instability. \label{si_compare}
+  }
+  \tablehead{
+    \colhead{Mode} & 
+    \colhead{$\tau_\mathrm{s}\OmK$} &
+    \colhead{$\epsilon$} &
+     \colhead{$K_x = K_z$} &
+      \colhead{$\sigma/\OmK$ (two-fluid)} &
+    \colhead{$\sigma/\OmK$ (one-fluid, Eq. \ref{streaming_dispersion})}   &
+    \colhead{$\mathcal{W}$ (arbitrary units) } 
+  }
+\startdata
+ linA, \cite{youdin07} &  $0.1$       & 3.0 & 30    & $-0.3480 +
+ 0.4190\ii$ & $0.3647 + 0.4251\ii$ & $0.026$\\
+
+linB, \cite{youdin07} & $0.1$        &  0.2 & 6 & $0.4999 +
+0.0154\ii$&   $-0.4999 + 0.0054\ii$  & $1.74$ \\
+
+linC,  \cite{bai10b}  & $10^{-2}$   &  2.0 & 1500&   $0.1049 +
+0.5981\ii$   &  $0.1344 + 0.6665\ii$  & $3.58\times10^{-4}$\\
+
+linD, \cite{bai10b} &  $10^{-3}$   &  2.0 & 2000 & $0.3225 +
+0.3154\ii$& $0.3228 + 0.3155\ii$ &  $1.84\times10^{-4}$
+\enddata
+\end{deluxetable*}
 
 %\subsection{Boundary conditions}
-
 %Eq. \ref{lin_mass}---\ref{lin_energy} can be reduced to a set of

### 2016-10-18 result.tex
notation change (k to kx), fig update, SI in main text

@@
 Fig. \ref{vsi_dust_loading} show unstable modes for different values
-of $\epsilon_0$ with fixed perturbation wavenumber  $k\Hg = 30$. The
+of $\epsilon_0$ with fixed perturbation wavenumber  $k_x\Hg = 30$. The
 eigenvalue distributions for $\epsilon_0 \leq 10^{-2}$ are similar to the
@@
 boundaries. The figure suggest that VSI becomes much less efficient
-for $\epsilon_0\gtrsim 0.1$ and $k\Hg\gtrsim 50$.
+for $\epsilon_0\gtrsim 0.1$ and $k_x\Hg\gtrsim 50$.
 
@@
 Fig. \ref{result2d_fixZ} compares the fastest growing VSI body modes  
-with $k\Hg=30$ for the two cases above.\citepalias[The thinner domain
+with $k_x\Hg=30$ for the two cases above.\citepalias[The thinner domain
   adopted here eliminates surface modes, ][]{lin15}.   
@@
 metalicity $Z=0.01$ so the midpane dust-to-gas ratio $\epsilon_0$ is 
-$O(1)$.  We consider radial wavenumbers such $k\Hd=1$.
+$O(1)$.  We consider radial wavenumbers such $k_x\Hd=1$.
 
@@
 \subsection{Effect of a radially-varying dust-to-gas ratio}\label{varHd}
-{\bf move to appendix?}
+{\bf bonus stuff, maybe delete in final version}
 
@@
 We fix $Z=0.01$, $\Hd=0.8\Hg$ at $r=r_0$, and
-$k\Hg=30$.
+$k_x\Hg=30$.
 

### 2016-10-19 linear.tex
re-written thermo interpret section

@@
 
-\section{Compressible streaming instability} 
+\section{Compressible streaming instability}\label{si}
 In the special case of an unstratified disk (neglecting the vertical

### 2016-10-19 model.tex
minor updates
re-written thermo interpret section

@@
 
-The gas component of the mixture has an entropy that is not conserved 
-because of the (locally) isothermal equation of state 
-adopted. However, we can define an effective (dimensionless)
+The gas component of the mixture has a specific entropy
+$S = C_p\ln{P/\rhog^\gamma}$, where $C_p$ is the heat capacity at constant
+pressure. However, this is not a useful quantity for (locally)
+isothermal gas.
+
+Nevertheless, since we have shown that an isothermal dusty gas
+effectively has $\gamma=1$ we can define an effective (dimensionless)  
 entropy for the \emph{entire mixture} as 
@@
 \begin{align}
-&  \sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV \notag\\
+%&  \sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV \notag\\
+  \sigma^2\mathcal{I}^2
 &= \int\left[ \rho
@@
 \end{align} 
-The real coefficients $A,B,C,D$ can be read off 
-Eq. \ref{integral_ex}. Note that $B=C$ from dynamical equilibrium
-(Eq. \ref{vshear}). We now consider various limits of
-Eq. \ref{int_rel}.
+The real integral $\mathcal{I}^2>0$ and coefficients $A,B,C,D$ can be
+read off Eq. \ref{integral_ex}. %Note that $B=C$ from dynamical equilibrium
+%(Eq. \ref{vshear}). 
+We now consider various limits of Eq. \ref{int_rel}.
 
@@
     \tepsilon\right)}$ for the entropy in the standard
-expression for said criteria \citep[e.g.][ where $C_p$ is the heat
-  capacity]{tassoul78}.
+expression for said criteria \citep[e.g.]{tassoul78}.
 
@@
 Now consider constant $c_s$ but $\tstop\neq0$. 
-%When $c_s(r,z)$ is non-uniform but $\tstop=0$, 
 Eq. \ref{int_rel} indicates that $\sigma^2$ is generally complex, so
-we may have growing oscillations. Writing $\sigma = -\omega + \ii s$
-where $\omega$ and $s$ are the real frequency and growth rates, 
-respectively, the imaginary part of Eq. \ref{int_rel} gives 
+we may have growing oscillations (overstability). Writing $\sigma =
+-\omega + \ii s$ where $\omega$ and $s$ are the real frequency and
+growth rates, respectively, the imaginary part of Eq. \ref{int_rel}
+gives  
 \begin{align}
-  s = \frac{\imag\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV}{2\omega\int\rho\left(|\dd
-    v_r|^2 + |\dd v_z|^2\right)dV}, \label{thermal_instability}
+%  s = \frac{\imag\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV}{2\omega\int\rho\left(|\dd
+%    v_r|^2 + |\dd v_z|^2\right)dV}, \label{thermal_instability}
+  s = \frac{\imag\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV}{2\omega\mathcal{I}^2}. \label{thermal_instability}
 \end{align}
-which represent instabilities due to dust-gas friction. 
+This represents instability due to dust-gas friction. 
 Since drag ($\delta \mathcal{C}$) appears as a source term in our
@@
 It is well-known that such correlations can lead to pulsational
-instabilities in stars \citep{cox67}. One may thus interpret 
-dust-drag instabilities in a similar way. We give a
-qualitative 
-description below, adapted from the more rigorous treatments of stellar 
-pulsations by 
-\cite{cox67} and lecture notes by \cite{samadi15} and
+instabilities in stars \citep{cox67}. We thus interpret  
+dust-drag instabilities in a similar way, 
+adapting from the treatment of stellar 
+pulsations by \cite{cox67} and lecture notes by \cite{samadi15} and 
 J. Christensen-Dalsgaard\footnote{\url{http://astro.phys.au.dk/$\sim$jcd/oscilnotes/}}.
@@
 
-The Lagragian perturbation $\Delta$ of a variable $f$ is 
-related to its Eulerian perturbation by $\Delta f = \delta f + 
-\bm{\xi}\cdot\nabla f$; where $\bm{\xi}$ is the Lagragian  
-displacement ($    \xi_{x,z} =  \ii \dd v_{x,z}/\sigma$). 
-%Note that $\Delta
-%\mathcal{C} = \delta \mathcal{C}$ for a uniform background $C$ as
-%required for equilibrium. 
-Then Eq. \ref{lin_mass_full}---\ref{lin_energy_full} imply  
-\begin{align} 
-  \imag \left(\delta \mathcal{C}
-  \nabla\cdot\dd \bm{v}^*\right) = 
-  -|\sigma|^2\imag \left(\Delta P 
-  \frac{\Delta\rho^*}{\rho}\right), \label{pdv}
-\end{align}
-which states that `PdV' work (right hand side) is
-provided by dust-drag friction. This can be seen explicitly as
-follows.
+The physical interpretation of Eq. \ref{thermal_instability} is that 
+work done by pressure forces in the dusty gas leads to growth in
+oscillation amplitudes. To see this, we first compute the work done
+assuming periodic oscillations. We then show that if the work done is positive, then the
+oscillation amplitude in fact grows: the work done leads the
+system to `overshoot' its equilibrium.
 
-The average rate of work done by the dusty-gas is
+The average rate of work done is 
 \begin{align}
   \mathcal{W} = \frac{1}{T_p}\int^{t+T_p}_{t}dt^\prime\int_M P
-  \frac{Dv}{Dt^\prime} dm 
+  \frac{Dv}{Dt^\prime} dm \label{work_def} 
 \end{align}
@@
 total mass $M$ of the mixture and $dm$ is the mass element. 
-
-Our aim is to relate the work done by the mixture, $\mathcal{W}$, 
-to the growth rate of oscillation amplitudes, $s$. To do this, we
-first evaluate $\mathcal{W}$ assuming pure oscillations (perturbations
-with time dependence $e^{\ii\omega t}$ and real $\omega$), then show
-that in fact the amplitudes grow if the work done is positive.
-
-Noting that only products in the perturbations contribute to 
-$\mathcal{W}$, we find that for real $\omega$: 
+Noting that only products of perturbations contribute to
+$\mathcal{W}$, we find that for periodic oscillations (time dependence 
+$e^{\ii\omega t}$ and real $\omega$): 
 \begin{align}
@@
 \end{align}
-where the integral is taken over the volume of the
-mixture. Eq. \ref{work_real} show that a phase difference between
-pressure and density leads to work done.
+where %the integral is taken over the volume of the
+%mixture; and 
+ the Lagragian perturbation $\Delta$ of a variable $f$ is 
+$\Delta f = \delta f + \bm{\xi}\cdot\nabla f$; where $\bm{\xi}$ is the
+Lagragian displacement ($    \xi_{x,z} =  \ii \dd v_{x,z}/\sigma$).  
+Eq. \ref{work_real} show that a phase difference between gas pressure and
+the total density leads to work done
+($\mathcal{W}\neq0$).
 
-Now combining Eq. \ref{work_real}, \ref{pdv} and \ref{thermal_instability}
+Now, quite generally, Eq. \ref{lin_mass_full}---\ref{lin_energy_full} 
+imply the integrand of the numerator in Eq. \ref{thermal_instability} 
+is 
+\begin{align} 
+  \imag \left(\delta \mathcal{C}
+  \nabla\cdot\dd \bm{v}^*\right) = 
+  -|\sigma|^2\imag \left(\Delta P 
+  \frac{\Delta\rho^*}{\rho}\right). \label{pdv}
+\end{align}
+%This states that `PdV' work (the right-hand-side) is 
+%provided by dust-drag friction, and can be seen explicitly as 
+%follows. 
+Then combining Eq. \ref{work_real}, \ref{pdv} and
+\ref{thermal_instability} 
 gives  
 \begin{align}
-s = \frac{2}{\mathcal{I}^2}\mathcal{W}, \label{growth_work}
+s = \frac{1}{\mathcal{I}^2}\mathcal{W}, \label{growth_work}
 \end{align}
-where $\mathcal{I}^2>0$  is the integral in the denominator of
-Eq. \ref{thermal_instability}, and we have set $\sigma= - \omega$ in
-Eq. \ref{pdv} for pure oscillations.  Thus growing modes in fact arise if
-positive work is done by the mixture during an oscillation. This in
-turn requires pressure to lag density, discussed next.
+where we have set $\sigma= - \omega$ in 
+Eq. \ref{pdv}. Eq. \ref{growth_work} shows that if the average work
+done is positive during a periodic oscillation, $\mathcal{W}>0$, then
+then its amplitude would actually grow ($s>0$).
+
+To see when/where positive work is done, 
+consider a point in the volume. We set $\Delta\rho = 1$ here 
+and take $\omega>0$ without loss of generality. (Thus the phase of
+perturbations increase in time.) This point contributes to positive
+work done if $-\omega \imag\left(\Delta
+P\Delta\rho^*\right)>0$, which requires $\imag\left(\Delta
+P\right)<0$. This corresponds to pressure lagging behind density since
+$\imag\left(\Delta\rho\right)=0$.
 
 \subsubsection{Physical picture of dust-drag instabilities}
-The full energy equation is 
-\begin{align*}
-  \frac{DP}{Dt} = \frac{P}{\rho}\frac{D\rho}{D t} + \mathcal{C}. 
-\end{align*}
-Dust-drag causes a phase difference between the pressure and density 
-evolution of a fluid element. In Fig. \ref{pdv_cartoon}, we show that
-if pressure lags behind density, then positive
-work done ($\mathcal{W}>0$) and would lead to growing oscillations
-($s>0$).
+
+%We now use physical arguments to show that $\mathcal{W}>0$ if pressure
+%perturbations (from the gas) lags behind the total density of the gas
+%plus dust mixture. 
+%The full energy equation is  
+%\begin{align*}
+%  \frac{DP}{Dt} = \frac{P}{\rho}\frac{D\rho}{D t} + \mathcal{C}. 
+%\end{align*} 
+%Dust-drag causes a phase difference between the pressure and density 
+%evolution of a fluid element.
+
+Fig. \ref{pdv_cartoon} gives a graphical demonstration that 
+if gas pressure lags behind total density, then positive work done 
+($\mathcal{W}>0$) because it leads to a clockwise path in the
+`P-V' plane. Consider a fluid element at maximum density during an
+oscillation cycle (point $A$), so that $D\rho/ D t = 0$. If dust-gas
+drag provides an effective heating to the fluid element at $A$
+(i.e. gas influx)  then it will experience increasing pressure, $DP/D
+t>0$. It attains pressure maximum (point $B$) \emph{after} reaching
+density maximum.    
+From $A$ to $B$, pressure forces increase while the parcel is already
+expanding. This corresponds to positive work done.
 
@@
     dust-gas drag. Top: pressure and density evolution in time. 
-    Bottom: oscillation cycle in the $P$-$V$ plane. In the case shown,
+    Bottom: oscillation cycle in the P-V plane. In the case shown,
     dust-drag causes 
     pressure to lag behind density, which results in a clockwise path in
-    the $P$-$V$ plane, implying positive work done by the fluid and
+    the P-V plane, implying positive work done by the fluid and
     hence would lead to growing oscillation amplitudes.  
@@
 
-Consider a fluid element at maximum density during an oscillation cycle
-(point $A$), so that $D\rho/ D t = 0$. If dust-gas drag provides an
-effective heating to the fluid element at $A$ (i.e. gas influx)  then
-it will experience increasing pressure, $DP/D t>0$. It attains
-pressure maximum (point $B$) \emph{after} reaching density maximum.   
-This implies pressure maximizes while the parcel is expanding: 
-the parcel's path in the `$P$-$V$' plane  is clockwise, as shown in
-the bottom panel of Fig. \ref{pdv_cartoon}. This corresponds to
-positive work done.
-
-Our thermodynamic interpretation does not explain
-\emph{why} dust drag causes a pressure lag behind density, but merely shows
+This thermodynamic interpretation does not explain
+\emph{why} dust drag causes pressure to lag behind density, but merely shows
 that this \emph{must} be the case for any growing modes associated the
-dust drag. To rigorously understand how dust-drag causes this lag requires an
-explicit solution to the linearized equations with detailed treatment
-of the function $\mathcal{C}$. However, given the complexity of
+dust drag. To rigorously understand how dust-drag causes this
+lag requires an explicit solution to the linearized equations with
+detailed treatment of the function $\mathcal{C}$. However, given the complexity of
 $\mathcal{C}$ (see Appendix \ref{lin_dust}), we might generally expect
@@
 gas towards it (and hence heating) to strengthen the pressure
-bump. That is, heating occurs upon compression, as occurs for stellar 
-pulsational instabilities. A phase difference was also noted in the 
-numerical calculations of the streaming instability by 
-\citet{youdin07b}.
+bump. That is, heating occurs upon compression, as in stellar  
+pulsational instabilities. In \S\ref{si} we check that the
+streaming instability fits into this thermodynamic interpretation in
+the strong drag limit.
+
+%A phase difference was also noted in the 
+%numerical calculations of the streaming instability by 
+%\citet{youdin07b}.
 
@@
 \begin{align}
+%  s = \frac{\imag\int P
+%  \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV}{2\omega\int\rho\left(|\dd 
+%    v_r|^2 + |\dd v_z|^2\right)dV}. \label{vsi_check} 
   s = \frac{\imag\int P
-  \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV}{2\omega\int\rho\left(|\dd
-    v_r|^2 + |\dd v_z|^2\right)dV}. \label{vsi_check}
+    \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV}{2\omega\mathcal{I}^2}. \label{vsi_check} 
 \end{align} 

### 2016-10-19 linear.tex
re-written thermo interpret section

--- 2016-10-18 linear.tex
@@
 to obtain a local dispersion relation of the form  
-$\sum_{j=1}^{5}c_j(k_x,k_z)\sigma^j = 0$, where $k_z$ is a real
+$\sum_{j=0}^{5}c_j(k_x,k_z)\sigma^j = 0$, where $k_z$ is a real
 vertical wavenumber. The coefficients $c_j$ can be read 

### 2016-10-19 linear.tex
re-written thermo interpret section
final changes before break

--- 2016-10-18 linear.tex
@@
 
-\section{Compressible streaming instability} 
+\section{Compressible streaming instability}\label{si}
 In the special case of an unstratified disk (neglecting the vertical
 component of the stellar gravity), we may also Fourier analyze in $z$
-to obtain a local dispersion relation of the form  
-$\sum_{j=1}^{5}c_j(k_x,k_z)\sigma^j = 0$, where $k_z$ is a real
+to obtain algebraic dispersion relation of the form  
+$\sum_{j=0}^{5}c_j(k_x,k_z)\sigma^j = 0$, where $k_z$ is a real
 vertical wavenumber. The coefficients $c_j$ can be read 
@@
 growth rates of the streaming instability have been 
-calculated analytically from full two-fluid disk models, as well as
-numerically from particle-gas simulations 
-\citep[namely][]{youdin07,bai10b}. These authors use normalized
+calculated analytically from full two-fluid disk models and measured 
+from direct numerical particle-gas simulations 
+\citep[namely][]{youdin07b,bai10b}. These authors use normalized
 wavenumbers $K_x = \eta r k_x$ and similarly for $K_z$, where
@@
 measures the pressure offset of Keplerian rotation. We use
-$\eta=0.05c_s/r\OmK$. They also parametrize the stopping time with
+$\eta=0.05c_s/r\OmK$. They also define the stopping time as 
 $\tau_\mathrm{s}=\tstop/(1-\tepsilon)$.
@@
 from the pressure evolution lagging behind that of the total density.
+
+Notice also there is better agreement between eigenfrequencies for
+smaller $\mathcal{W}$, i.e. smaller phase pressure-density phase
+lags. While this is generally achieved with decreasing $\tstop$, other
+problem parameters may affect the magnitude of this phase lag.
+For example, `linA' and `linB' have the same $\tstop$ but linA shows
+better agreement with the one-fluid growth rate. This suggest that how
+well dust and gas are effectively coupled is problem-dependent.
+
+%"generalized stokes number" would include other problem parameters,
+%e.g. wavenumber
 
@@
 \startdata
- linA, \cite{youdin07} &  $0.1$       & 3.0 & 30    & $-0.3480 +
+ linA, \cite{youdin07b} &  $0.1$       & 3.0 & 30    & $-0.3480 +
  0.4190\ii$ & $0.3647 + 0.4251\ii$ & $0.026$\\
 
-linB, \cite{youdin07} & $0.1$        &  0.2 & 6 & $0.4999 +
+linB, \cite{youdin07b} & $0.1$        &  0.2 & 6 & $0.4999 +
 0.0154\ii$&   $-0.4999 + 0.0054\ii$  & $1.74$ \\

### 2016-10-19 model.tex
final changes before break

@@
 \begin{align}
-s = \frac{1}{\mathcal{I}^2}\mathcal{W}, \label{growth_work}
+s = \frac{\mathcal{W}}{\mathcal{I}^2}, \label{growth_work}
 \end{align}
@@
 
-To see when/where positive work is done, 
-consider a point in the volume. We set $\Delta\rho = 1$ here 
+Under what condition is positive work is done? 
+Consider a point in the volume where $\Delta\rho = 1$  
 and take $\omega>0$ without loss of generality. (Thus the phase of
@@
 
-\subsubsection{Physical picture of dust-drag instabilities}
+\subsubsection{Physical picture of dust-gas drag instabilities}
 
@@
 Fig. \ref{pdv_cartoon} gives a graphical demonstration that 
-if gas pressure lags behind total density, then positive work done 
+if gas pressure lags behind density, then positive work done 
 ($\mathcal{W}>0$) because it leads to a clockwise path in the
-`P-V' plane. Consider a fluid element at maximum density during an
-oscillation cycle (point $A$), so that $D\rho/ D t = 0$. If dust-gas
-drag provides an effective heating to the fluid element at $A$
-(i.e. gas influx)  then it will experience increasing pressure, $DP/D
-t>0$. It attains pressure maximum (point $B$) \emph{after} reaching
-density maximum.    
-From $A$ to $B$, pressure forces increase while the parcel is already
-expanding. This corresponds to positive work done.
+`P-V' plane. From $A$ to $B$, a fluid parcel is expanding to return to 
+equilbrium, but pressure (i.e. restoring) forces is still
+increasing. This over-compensation corresponds to positive work done. 
+Similarly, from $C$ to $D$ the parcel is contracting but pressure 
+(restoring) forces are still dropping, which allows the 
+parcel's contraction to over-shoot its equilibrium density. The
+overall positive work done leads to growth in the oscillation amplitude.
+
+%Consider a fluid element at maximum density during an
+%oscillation cycle (point $A$), so that $D\rho/ D t = 0$. If dust-gas
+%drag provides an effective heating to the fluid element at $A$
+%(i.e. gas influx)  then it will experience increasing pressure, $DP/D
+%t>0$. It attains pressure maximum (point $B$) \emph{after} reaching
+%density maximum.
+
+When applied to dusty gas, the work done is attributed to finite 
+dust-gas drag, which allows relative drift between gas and dust. 
+A pressure-density lag of $A$ to $B$ is achieved if, 
+after a region reaches maximum total density, its 
+gas component is still increasing. This in turn implies a sufficiently
+large particle flux \emph{out} of that region.
 

### 2016-10-19 result.tex
fixed some typos, corrected kappa^2 dependence on dust

@@
 
-%When $\tstop=0$ we may construct exact dusty equilibra by specifying
-%the dust-to-gas ratio, as described in \S\ref{eqm}.
-
-We present numerical solutions to the linearized
-equations to examine the effect of dust on the VSI. 
-We consider perfectly coupled dust ($\tstop=0$) so that the 
-equilibria defined in \S\ref{eqm} are exact steady states.
+We now present numerical solutions to the linearized equations in a 
+vertically stratified disk. We consider perfectly coupled dust 
+($\tstop=0$) so that the equilibria defined in \S\ref{eqm} are exact
+steady states. In this limit any growing modes must be associated with
+the imposed temperature gradient (\S\ref{limits}). We thus 
+examine the effect of dust-loading on the VSI, previously studied in 
+pure gas disks \citep[][\citetalias{lin15} in this section]{lin15}.
 
@@
 parameters of the linear problem. We choose a power-law 
-midplane gas density profile, $\rho_\mathrm{g0}\propto r^q$ with $p =
--1.5$. The power-law index for the 
+midplane gas density profile, $\rho_\mathrm{g0}\propto r^p$ with $p =
+-1.5$. The fiducial power-law index for the 
 temperature profile is $q=-1$; and we set the gas disk aspect-ratio
-$h_\mathrm{g}=0.05$. These are fiducial values used by 
-\citet[][\citetalias{lin15} in this section]{lin15}.
+$h_\mathrm{g}=0.05$. These values were also used by 
+\citetalias{lin15}.
 
 Two boundary conditions are needed for the $\tstop=0$
-problem. This is expected since this problem is equivalent to
-adiabatic hydrodynamics \citep[e.g.][]{lubow93}. We 
-impose solid boundaries so that $\delta v_z(\pm\zmax)=0$.  
-We solve the linearized equations as a generalized eigenvalue problem 
-using a pseudo-spectral code adapted from \cite{lin15}. Amplitudes 
-are expanded in Chebyshev polynomials up to order $N_z=512$. We check
-results using Eq. \ref{vsi_check}.
+problem \citep[e.g.][]{lubow93}. We impose solid boundaries so that
+$\delta v_z(\pm\zmax)=0$. We solve the linearized equations as a
+generalized eigenvalue problem using a pseudo-spectral code adapted
+from \citetalias{lin15}. Amplitudes are expanded in Chebyshev
+polynomials up to order $N_z=512$. We check results using
+Eq. \ref{vsi_check}.
 
 \subsection{Qualitative expectations}\label{vsi_est}
-\cite{lin15} found the appropriate way to compare 
+\citetalias{lin15} found the appropriate way to compare 
 destabilizing vertical shear to stabilizing vertical buoyancy is

### 2016-10-20 linear.tex
fixed some typos, corrected kappa^2 dependence on dust
corrected kappa^2 dependence on dust
added pressure phase lag
tiny edits

@@
 evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
-is retained. We now discuss explicit solutions to these equations. 
+is retained. We now discuss solutions to these equations in an unstratified disk 
+ for the streaming instability (\S\ref{si}), and a stratified disk for the VSI (\S\ref{results}).
+
  % Numerical solutions are generally required for the 
@@
 
+{\bf note: kappa2 accounts for dust effect assuming smallh=0.05}
 We solve Eq. \ref{streaming_dispersion} for selected cases where the  
@@
 importantly, we find the work done $\mathcal{W}>0$ in all cases, and
-hence find growing oscillations. Thus the streaming instability stems
-from the pressure evolution lagging behind that of the total density.
+hence find growing oscillations. This implies the streaming instability stems
+from the gas pressure evolution lagging behind that of the total (and hence dust) density.
 
-Notice also there is better agreement between eigenfrequencies for
-smaller $\mathcal{W}$, i.e. smaller phase pressure-density phase
-lags. While this is generally achieved with decreasing $\tstop$, other
-problem parameters may affect the magnitude of this phase lag.
-For example, `linA' and `linB' have the same $\tstop$ but linA shows
-better agreement with the one-fluid growth rate. This suggest that how
-well dust and gas are effectively coupled is problem-dependent.
+Notice also eigenfrequencies agree better for 
+smaller pressure-density phase lags. While this is generally true with decreasing $\tstop$, other 
+problem parameters may affect the magnitude of this phase lag, and hence the validity of the one-fluid model.  
+For example, `linA' has a larger $\tstop$ than `linB', but linA's one and 
+two-fluid growth rates differ by $\sim 1\%$; compare to $\sim 10\%$ error for linB. 
+However, linB has a larger phase lag. This means that the effective coupling strength between dust and gas 
+is problem-dependent.
 
@@
 
-\begin{deluxetable*}{llrrrrr}
-  \tablecolumns{7}
+\begin{deluxetable*}{llrrrrrr}
+  \tablecolumns{8}
   \tablecaption{Selected eigenfrequencies for the streaming
@@
     \colhead{$\epsilon$} &
-     \colhead{$K_x = K_z$} &
+     \colhead{$K_{x,z}$} &
       \colhead{$\sigma/\OmK$ (two-fluid)} &
-    \colhead{$\sigma/\OmK$ (one-fluid, Eq. \ref{streaming_dispersion})}   &
-    \colhead{$\mathcal{W}$ (arbitrary units) } 
+    \colhead{$\sigma/\OmK$ (one-fluid)} &
+    \colhead{$\mathcal{W}$ (arbitrary units)} &
+      \colhead{Pressure lag}
   }
@@
  linA, \cite{youdin07b} &  $0.1$       & 3.0 & 30    & $-0.3480 +
- 0.4190\ii$ & $0.3647 + 0.4251\ii$ & $0.026$\\
+ 0.4190\ii$ & $0.3640 + 0.4249\ii$ & $0.026$ & $42\degr$\\
 
 linB, \cite{youdin07b} & $0.1$        &  0.2 & 6 & $0.4999 +
-0.0154\ii$&   $-0.4999 + 0.0054\ii$  & $1.74$ \\
+0.0155\ii$&   $-0.4981 + 0.0054\ii$  & $1.75$ & $88\degr$ \\
 
 linC,  \cite{bai10b}  & $10^{-2}$   &  2.0 & 1500&   $0.1049 +
-0.5981\ii$   &  $0.1344 + 0.6665\ii$  & $3.58\times10^{-4}$\\
+0.5981\ii$   &  $0.1338 + 0.6650\ii$  & $3.57\times10^{-4}$& $78\degr$ \\
 
 linD, \cite{bai10b} &  $10^{-3}$   &  2.0 & 2000 & $0.3225 + 
-0.3154\ii$& $0.3228 + 0.3155\ii$ &  $1.84\times10^{-4}$
+0.3154\ii$& $0.3219 + 0.3154\ii$ &  $1.85\times10^{-4}$ & $21\degr$ 
 \enddata

### 2016-10-20 linear.tex
tiny edits

@@
 problem parameters may affect the magnitude of this phase lag, and hence the validity of the one-fluid model.  
-For example, `linA' has a larger $\tstop$ than `linB', but linA's one and 
-two-fluid growth rates differ by $\sim 1\%$; compare to $\sim 10\%$ error for linB. 
-However, linB has a larger phase lag. This means that the effective coupling strength between dust and gas 
+For example, `linA' has a larger $\tstop$ than `linC', but linA's one and 
+two-fluid growth rates differ by $\sim 1\%$; compare to $\sim 10\%$ error for linC. 
+However, linC has a larger phase lag. This means that the effective coupling strength between dust and gas 
 is problem-dependent.

### 2016-10-20 linear.tex
fixed SI results. need to use largrangian pert properly

@@
 There we also show that this dispersion relation reduces to that for
-the streaming instability in the limit of incompressible gas and small
+the streaming instability (SI) in the limit of incompressible gas and small
 $\tstop$ \citep{youdin05a,jacquet11}.
@@
 {\bf note: kappa2 accounts for dust effect assuming smallh=0.05}
-We solve Eq. \ref{streaming_dispersion} for selected cases where the  
-growth rates of the streaming instability have been 
+We solve Eq. \ref{streaming_dispersion} for selected cases where SI  
+growth rate have been 
 calculated analytically from full two-fluid disk models and measured 
@@
 importantly, we find the work done $\mathcal{W}>0$ in all cases, and
-hence find growing oscillations. This implies the streaming instability stems
-from the gas pressure evolution lagging behind that of the total (and hence dust) density.
+hence find growing oscillations.
 
-Notice also eigenfrequencies agree better for 
-smaller pressure-density phase lags. While this is generally true with decreasing $\tstop$, other 
-problem parameters may affect the magnitude of this phase lag, and hence the validity of the one-fluid model.  
-For example, `linA' has a larger $\tstop$ than `linC', but linA's one and 
-two-fluid growth rates differ by $\sim 1\%$; compare to $\sim 10\%$ error for linC. 
-However, linC has a larger phase lag. This means that the effective coupling strength between dust and gas 
-is problem-dependent.
+We also checked the growth rates $s$ satisfy 
+\begin{align} 
+  s = \frac{\left|\sigma\right|^2\imag\left(\Delta P
+    \Delta\rho^*/\rho\right)}{2\real\left(\sigma\right)\mathcal{I}^2}, \label{si_check}
+\end{align}
+as implied by Eq. \ref{thermal_instability} and
+Eq. \ref{pdv}. Interestingly, we find that for cases with
+$\epsilon>1$, Eq. \ref{si_check} can be satisified with $\Delta P \to
+\Delta P - \dd P$, i.e. the advection part of the Lagragian
+pressure perturbation is dominant. Conversely, for `linB' with $\epsilon
+< 1$, one can take $\Delta P \simeq \dd P$ in Eq. \ref{si_check}. This
+suggests that for SI in gas-dominated disks the distinction between
+Eulerian and Lagragian presure perturbations is unimportant.
+
+{\bf note: important to use largrangian pressure pert}
+In Table \ref{si_compare} we also calculate the phase difference
+between the Lagragian pressure perturbation and density perturbations
+as    
+\begin{align*} 
+\varphi = -\sgn(\omega)\arg\left(\Delta P\Delta\rho^*\right), 
+\end{align*}
+so that $\varphi > 0 $ indicates pressure lagging behind density, which
+is true for all the cases. In particular, `linB' indicates that growth
+rates vanish as $\varphi \to 0$.  
+Thus the streaming instability stems  
+from the gas pressure evolution lagging behind that of the total
+density (and hence behind the dust density for isothermal gas).
 
@@
     \colhead{$\mathcal{W}$ (arbitrary units)} &
-      \colhead{Pressure lag}
+      \colhead{$\Delta P$ lag}
   }
@@
  linA, \cite{youdin07b} &  $0.1$       & 3.0 & 30    & $-0.3480 +
- 0.4190\ii$ & $0.3640 + 0.4249\ii$ & $0.026$ & $42\degr$\\
+ 0.4190\ii$ & $0.3640 + 0.4249\ii$ & $0.90$ & $30\degr$\\
 
 linB, \cite{youdin07b} & $0.1$        &  0.2 & 6 & $0.4999 +
-0.0155\ii$&   $-0.4981 + 0.0054\ii$  & $1.75$ & $88\degr$ \\
+0.0155\ii$&   $-0.4981 + 0.0054\ii$  & $1.54$ & $1.2\degr$ \\
 
 linC,  \cite{bai10b}  & $10^{-2}$   &  2.0 & 1500&   $0.1049 +
-0.5981\ii$   &  $0.1338 + 0.6650\ii$  & $3.57\times10^{-4}$& $78\degr$ \\
+0.5981\ii$   &  $0.1338 + 0.6650\ii$  & $0.15$& $11\degr$ \\
 
 linD, \cite{bai10b} &  $10^{-3}$   &  2.0 & 2000 & $0.3225 + 
-0.3154\ii$& $0.3219 + 0.3154\ii$ &  $1.85\times10^{-4}$ & $21\degr$ 
+0.3154\ii$& $0.3219 + 0.3154\ii$ &  $1.28$ & $22\degr$ 
 \enddata

### 2016-10-21 linear.tex
tiny correction in entropy def

@@
 \begin{align*} 
-\varphi = -\sgn(\omega)\arg\left(\Delta P\Delta\rho^*\right), 
+\varphi &= -\sgn(\omega)\arg\left(\Delta P\Delta\rho^*\right)
+ &       = -\sgn(\omega)\arg\left(\Delta\rhog\Delta\rhod^*\right),\notag
 \end{align*}
-so that $\varphi > 0 $ indicates pressure lagging behind density, which
+where we have applied the isothermal equation of state for the second equality.
+Then $\varphi > 0 $ indicates pressure lagging behind density, which
 is true for all the cases. In particular, `linB' indicates that growth
@@
 Thus the streaming instability stems  
-from the gas pressure evolution lagging behind that of the total
-density (and hence behind the dust density for isothermal gas).
+from the gas pressure evolution lagging behind that of the dust 
+density.
 

### 2016-10-21 model.tex
tiny correction in entropy def
last update until november, for real

@@
 The gas component of the mixture has a specific entropy
-$S = C_p\ln{P/\rhog^\gamma}$, where $C_p$ is the heat capacity at constant
+$S = C_p\ln{\left(P^{1/\gamma}/\rhog\right)}$, where $C_p$ is the heat capacity at constant
 pressure. However, this is not a useful quantity for (locally)
@@
 \end{align}
-where $\sigma$ is the complex mode frequency. In Appendix
+where $\sigma$ is the complex mode frequency. For later convenience we
+write 
+\begin{align}
+  \sigma = \ii s - \omega,
+\end{align}
+where $s$ and $\omega$ are the growth rate and real frequencies,
+respectively. Thus perturbations have time dependence $e^{st +
+  \ii\omega t}$.
+
+In Appendix
 \ref{var_prin} we derive the integral relation,  
@@
 Eq. \ref{int_rel} indicates that $\sigma^2$ is generally complex, so
-we may have growing oscillations (overstability). Writing $\sigma =
--\omega + \ii s$ where $\omega$ and $s$ are the real frequency and
-growth rates, respectively, the imaginary part of Eq. \ref{int_rel}
+we may have growing oscillations (overstability). %Writing $\sigma =
+%-\omega + \ii s$ where $\omega$ and $s$ are the real frequency and
+%growth rates, respectively, 
+The imaginary part of Eq. \ref{int_rel}
 gives  

### 2016-10-22 _region_.tex
head to toe edits

--- 2016-10-19 result.tex
@@
+\message{ !name(result.tex)}
+\message{ !name(result.tex) !offset(-2) }
 \section{
@@
 
-We now present numerical solutions to the linearized equations in a 
-vertically stratified disk. We consider perfectly coupled dust 
-($\tstop=0$) so that the equilibria defined in \S\ref{eqm} are exact
+We present numerical solutions for  
+vertically stratified disks assuming perfectly coupled dust 
+($\tstop=0$). This is so that the equilibria defined in \S\ref{eqm} are exact
 steady states. In this limit any growing modes must be associated with
-the imposed temperature gradient (\S\ref{limits}). We thus 
-examine the effect of dust-loading on the VSI, previously studied in 
+the imposed temperature gradient (\S\ref{limits}). We are thus  
+examining the effect of dust-loading on the VSI, previously studied in 
 pure gas disks \citep[][\citetalias{lin15} in this section]{lin15}.
@@
 for $\tstop\OmK \ll O(10^{-2})$.
+
+\message{ !name(result.tex) !offset(-360) }

### 2016-10-22 appendix.tex
fixed some typos, corrected kappa^2 dependence on dust
head to toe edits

@@
 \begin{align}
-  \tstop  =  \frac{K}{\rho c_s}
+  \tstop  =  \frac{K}{\rho c_s}\label{epstein}
 \end{align}
@@
 We consider an unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
-0$. The background $\p_rP/\rho$ and $\tepsilon$ are constants. 
+0$. The background $\p_rP/\rho$, $\tepsilon$ and $c_s^2$ are constant
+input parameters. Then for the Epstein drag law, Eq. \ref{epstein}, we
+have $\mathcal{C}=0$ in the background state. 
 We Fourier analyze in $r$ and $z$ so that $\p_z\to \ii k_z$ and
@@
 \equiv k_x^2 + k_z^2$. We neglect 
-background gradients when compared with that of perturbations. The linearized
+background gradients when compared with that of
+perturbations. The linearized
 equations are, after eliminating the azimuthal velocity:
@@
 \begin{align}
-  \ii \sigma W &=\nabla\cdot \dd \bm{ v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\\
+  \ii \sigma W &=\nabla\cdot \dd \bm{ v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\label{streaming_mass}\\
     \sigma^2 \dd v_x &= \kappa^2 \dd v_x - \ii \sigma F_r W +
-    k_x\sigma Q,\\
-  -\ii\sigma\dd v_z &= -\ii k_zQ,\\
+    k_x\sigma Q,\label{streaming_vx}\\
+  -\ii\sigma\dd v_z &= -\ii k_zQ,\label{streaming_vz}\\
 \ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \dd\bm{v}   -
@@
 \end{align}
+%{\bf bg pressure gradient neglected in energy eq (is this ok? could
+%  fix, but get more complicated dispersion}
 The artifical factor $\zeta = 1$  is inserted to keep track of the
@@
 \end{align}
+%{\bf only the second term contributes to work done? dont think so}
 

### 2016-10-22 linear.tex
last update until november, for real
head to toe edits

--- 2016-10-20 linear.tex
@@
 component of the stellar gravity), we may also Fourier analyze in $z$
-to obtain algebraic dispersion relation of the form  
+to obtain an algebraic dispersion relation of the form  
 $\sum_{j=0}^{5}c_j(k_x,k_z)\sigma^j = 0$, where $k_z$ is a real
@@
 We solve Eq. \ref{streaming_dispersion} for selected cases where SI  
-growth rate have been 
+growth rates have been 
 calculated analytically from full two-fluid disk models and measured 
@@
 
+{\bf note: important to use largrangian pressure pert properly}
 We also checked the growth rates $s$ satisfy 
@@
   s = \frac{\left|\sigma\right|^2\imag\left(\Delta P
-    \Delta\rho^*/\rho\right)}{2\real\left(\sigma\right)\mathcal{I}^2}, \label{si_check}
+    \Delta\rho^*/\rho\right)}{2\real\left(\sigma\right)\rho\left(\left|\dd
+  v_x\right|^2+\left|\dd
+  v_z\right|^2\right)}, \label{si_check}
 \end{align}
 as implied by Eq. \ref{thermal_instability} and
-Eq. \ref{pdv}. Interestingly, we find that for cases with
-$\epsilon>1$, Eq. \ref{si_check} can be satisified with $\Delta P \to
-\Delta P - \dd P$, i.e. the advection part of the Lagragian
-pressure perturbation is dominant. Conversely, for `linB' with $\epsilon
-< 1$, one can take $\Delta P \simeq \dd P$ in Eq. \ref{si_check}. This
-suggests that for SI in gas-dominated disks the distinction between
-Eulerian and Lagragian presure perturbations is unimportant.
+Eq. \ref{pdv}. %, or directly from
+%Eq. \ref{streaming_mass}---\ref{streaming_vz}. 
+Interestingly, we find that for cases with
+$\epsilon>1$, Eq. \ref{si_check} can be satisified with $\Delta
+P\simeq \ii\dd v_x \p_rP/\sigma$, i.e. the radial pressure gradient
+is responsible for growth. Conversely, for `linB' with $\epsilon
+< 1$, one can approximate $\Delta P \simeq \dd P$ in Eq. \ref{si_check}. This
+suggests that for SI in gas-dominated disks, the distinction between
+Eulerian and Lagragian pressure perturbations is unimportant. 
+{\bf but weak growth in this case}
+%However
+%in that case the gro
 
-{\bf note: important to use largrangian pressure pert}
 In Table \ref{si_compare} we also calculate the phase difference
@@
 \begin{align*} 
-\varphi = -\sgn(\omega)\arg\left(\Delta P\Delta\rho^*\right), 
+\varphi \equiv -\sgn(\omega)\arg\left(\Delta P\Delta\rho^*\right)
+        = \sgn(\omega)\arg\left(\Delta\rhog^*\Delta\rhod\right),\notag
 \end{align*}
-so that $\varphi > 0 $ indicates pressure lagging behind density, which
+%where we have applied the isothermal equation of state for the second equality.
+then $\varphi > 0 $ indicates pressure lagging behind density, which
 is true for all the cases. In particular, `linB' indicates that growth
 rates vanish as $\varphi \to 0$.  
-Thus the streaming instability stems  
-from the gas pressure evolution lagging behind that of the total
-density (and hence behind the dust density for isothermal gas).
+Thus the streaming instability indeed stems  
+from the gas pressure evolution lagging behind that of the dust 
+density.
 
@@
 %e.g. wavenumber
+
+{\bf suppose large scale pressure grad drops outwards. 
+pert the sys by kicking more dust inwards. gas moves out, into region 
+  of lower pressure in the bg. i.e. create pressure bump at larger
+  radius. it attracts the first dust particles back out, plus some more
+  because bump is larger. (move gas from higher density to lower
+  density produce asymmetric bump/trough.}
 

### 2016-10-22 model.tex
head to toe edits

@@
 However, disk structures obtained this way generally do not satisfy 
-the steady state effective energy Eq. \ref{eff_energy} 
-except in the limit of perfectly coupled dust ($\tstop\to0$).  In
-practice the dust-component would settle on a timescale
-$t_\mathrm{settle}\sim 1/\OmK^2\tstop$, but for tightly-coupled dust
+the steady state effective energy Eq. \ref{eff_energy} because
+$\mathcal{C}\neq0$ for general solutions to
+Eq. \ref{steady_momr}---\ref{steady_momz}.
+
+In practice the dust particles settle vertically on a timescale 
+$t_\mathrm{settle}\sim 1/\OmK^2\tstop$, although for tightly-coupled dust
 particles this is much longer than that to establish vertical
 hydrostatic equilibrium in the gas component ($\sim 1/\OmK$).  
-
 True steady states may be calculated by invoking an underlying gas
 turbulence, which leads to dust diffusion \citep{takeuchi02, youdin07, 
-  lyra13}. This complication is beyond the scope of this paper and we 
-defer it to a future work. Here, we will either consider $\tstop = 0$,
-or assume that $\tstop$ is sufficiently small such that the background
-disk (as obtained from Eq. \ref{steady_momr}---\ref{steady_momz}) does 
-not evolve significantly.  %relative to perturbations
+ lyra13}. This complication is beyond the scope of this paper.
+
+In this work, we will assume that $\tstop$  is sufficiently small so
+that settling occurs on much longer timescales  than that of interest; or
+consider equilibria with $\mathcal{C}\equiv 0$, namely   
+\begin{inparaenum}[1)] 
+\item 
+  unstratified disks with finite dust-gas drag (\S\ref{si});  
+\item 
+  perfectly-coupled dust with $\tstop=0$ (\S\ref{results}). 
+\end{inparaenum} 
+Vertical settling is absent in both cases.
 
@@
 proportional to $q$ corresponds to vertical shear due to the 
-radial temperature gradient. The last term is present in the dust-free
+radial temperature gradient. The last term survives in the dust-free
 limit.
@@
   \mathcal{W} = \frac{1}{T_p}\int^{t+T_p}_{t}dt^\prime\int_M P
-  \frac{Dv}{Dt^\prime} dm \label{work_def} 
+  \frac{D\upsilon}{Dt^\prime} dm \label{work_def} 
 \end{align}
 \citep{cox67}, 
-where $T_p$ is the oscillation period, $v=1/\rho$ is the specific
+where $T_p$ is the oscillation period, $\upsilon=1/\rho$ is the specific
 volume of the mixture, and the spatial integral is taken over the
@@
 where we have set $\sigma= - \omega$ in 
-Eq. \ref{pdv}. Eq. \ref{growth_work} shows that if the average work
+Eq. \ref{pdv}. Eq. \ref{growth_work} states that if the average work
 done is positive during a periodic oscillation, $\mathcal{W}>0$, then
-then its amplitude would actually grow ($s>0$).
+then its amplitude would actually grow ($s>0$).  %and not be periodic
 
-Under what condition is positive work is done? 
-Consider a point in the volume where $\Delta\rho = 1$  
-and take $\omega>0$ without loss of generality. (Thus the phase of
-perturbations increase in time.) This point contributes to positive 
-work done if $-\omega \imag\left(\Delta 
-P\Delta\rho^*\right)>0$, which requires $\imag\left(\Delta
-P\right)<0$. This corresponds to pressure lagging behind density since
-$\imag\left(\Delta\rho\right)=0$.
+Positive work done is done by a parcel if $-\omega \imag\left(\Delta  
+P\Delta\rho^*\right)>0$. Without loss of generality, take $\omega>0$
+and consider a mass element with $\Delta\rho = 1$. Then positive work 
+requires $\imag\left(\Delta
+P\right)<0$.  This corresponds to pressure perturbations \emph{lagging}
+behind that in density (or displacements), since the phase of
+the assumed perturbations increase in time.
+
+Fig. \ref{pdv_cartoon} gives a graphical demonstration that 
+if gas pressure lags behind density, then %positive work done 
+$\mathcal{W}>0$ because it leads to a clockwise path in the
+`P-V' plane. From $A$ to $B$, a fluid parcel is expanding to return to 
+equilbrium, but pressure (i.e. restoring) forces is still
+increasing. This over-compensation causes the parcel to expand
+beyond its equilibrium volume. 
+%corresponds to positive work done. 
+Similarly, from $C$ to $D$ the parcel is already contracting towards
+equilibrium, but pressure
+forces are still dropping. This allows the  
+parcel's contraction to over-shoot its equilibrium density. The
+overall positive work done leads to growth in the oscillation amplitude.
+
+%$\imag\left(\Delta\rho\right)=0$.
 
@@
 
-Fig. \ref{pdv_cartoon} gives a graphical demonstration that 
-if gas pressure lags behind density, then positive work done 
-($\mathcal{W}>0$) because it leads to a clockwise path in the
-`P-V' plane. From $A$ to $B$, a fluid parcel is expanding to return to 
-equilbrium, but pressure (i.e. restoring) forces is still
-increasing. This over-compensation corresponds to positive work done. 
-Similarly, from $C$ to $D$ the parcel is contracting but pressure 
-(restoring) forces are still dropping, which allows the 
-parcel's contraction to over-shoot its equilibrium density. The
-overall positive work done leads to growth in the oscillation amplitude.
-
 %Consider a fluid element at maximum density during an
@@
 
-When applied to dusty gas, the work done is attributed to finite 
-dust-gas drag, which allows relative drift between gas and dust. 
-A pressure-density lag of $A$ to $B$ is achieved if, 
-after a region reaches maximum total density, its 
-gas component is still increasing. This in turn implies a sufficiently
-large particle flux \emph{out} of that region.
+%Applying the above argument to isothermal dusty gas, 
+For dusty gas, the work done during oscillations 
+ is attributed to finite dust-gas drag. The relative
+drift between gas and dust causes a phase difference between the two
+components.
+
+A parcel of the isothermal dusty-gas mixture does 
+postive work if  
+\begin{align*}
+-\sgn\left(\omega\right)\imag\left(\Delta\rhog\Delta\rhod^*\right)>0,
+\end{align*}
+meaning that gas follows dust. Instabilities are thus not possible if the
+gas does not respond to dust-drag (i.e. no back-reaction). The
+pressure-density lag shown in Fig. \ref{pdv_cartoon} is achieved if, 
+after a parcel reaches maximum total density, its 
+gas content is still increasing ($A$ to $B$). This in turn implies a sufficiently
+large particle flux \emph{out} of the parcel.
+
+{\bf lagrangian derivative follows the mixture. this highlights one
+  perk of the one-fluid framework. in two-fluid models there are two
+  types of lagragian deriv, not clear which fluid to follow.}
 
@@
 This thermodynamic interpretation does not explain 
-\emph{why} dust drag causes pressure to lag behind density, but merely shows
+\emph{why} drag forces causes gas pressure to lag behind the dust
+density, but merely shows 
 that this \emph{must} be the case for any growing modes associated the
-dust drag. To rigorously understand how dust-drag causes this
+dust-gas drag. To rigorously understand how dust-drag causes this 
 lag requires an explicit solution to the linearized equations with 
@@
 bump. That is, heating occurs upon compression, as in stellar  
-pulsational instabilities. In \S\ref{si} we check that the
+pulsational instabilities \citep{cox67}. In \S\ref{si} we check that the
 streaming instability fits into this thermodynamic interpretation in

### 2016-10-22 paper.tex
additional discussion in dust edge instability
revising with youdin's comments
edits in response to youdin's comments
rough edits for youdin's comments
head to toe edits

@@
 \newcommand{\dbz}{\dd B_z}
+\newcommand{\rich}{\mathrm{Ri}}
 \newcommand{\Htilde}{H_\epsilon}
@@
 \newcommand{\hgas}{h_\mathrm{g}}
-\newcommand{\tepsilon}{\tilde{\epsilon}}
+\newcommand{\tepsilon}{f_\mathrm{d}}
 \newcommand{\rhod}{\rho_\mathrm{d}}
@@
 \newcommand{\Hd}{H_\mathrm{d}}
+\newcommand{\seff}{S_\mathrm{eff}}
+\newcommand{\tcool}{t_\mathrm{cool}} 
 \newcommand{\actaa}{Acta Astron.}
@@
 itself cannot lead to axisymmetric instabilities, however thin the dust layer becomes.     
-% 
+However, sufficiently sharp radial gradients in the dust-to-gas ratio
+can destabilize the disk.  
 We also extend 
@@
 %dust settling does not lead to axisymmetric instability. 
-We discuss thermodynamical interpretations of 
-dust-gas drag instabilities, for example the streaming
-instability in protoplanetary disks, as a phase lag between the pressure 
-and density of a dusty gas, much like pulsational instabilities 
+We also duscuss thermodynamical interpretations of 
+dust-gas drag instabilities, and show that the streaming
+instability in protoplanetary disks occur because drag causes 
+a phase lag between the gas pressure  
+and dust density, much like pulsational instabilities 
 in stars.  
@@
 simulating gas tightly coupled to dust particles with small $\tstop$.  
-This can be numerically challenging for the two-fluid and Lagragian
-particles framework; but is straight forward to simulate in the
+This can be numerically challenging for two-fluid models and 
+following Lagragian particles; but is straight forward to simulate in the 
 one-fluid formulation \citep{price15}.  The single fluid formalism
 also has the important advantage that it is 
-amenable to standard analytical methods
-\citep{youdin05a,jacquet11}. This is the necessary  approach to gain
-physical insight and explain some of the above newer dusty
-instabilities.
+amenable to standard analytical methods \citep{youdin05a,jacquet11}.  
+This is the necessary  approach to gain physical insight.
 
@@
 framework to study the stability of dusty protoplanetary 
-disks by drawing on the physical and mathematical parallel between dusty gas and standard
-hydrodynamics. This is possible in certain limits of astrophysical
-interest.
+disks by drawing on the physical and mathematical parallel between
+dusty gas and standard hydrodynamics. This is possible in certain limits of astrophysical
+interest, allowing us to interpret dusty phenomenon through standard
+hydrodynamics.
 
@@
 of entropy following an adiabatic fluid. We will formally relate the 
-entropy of isothermal dusty gas to its dust content. Finite  
-dust-gas drag can be interpreted as heating/cooling: the relative
+effective entropy of isothermal dusty gas to its dust content. Finite  
+dust-gas drag act as effective heating/cooling: the relative
 motion between dust and gas causes a fluid parcel to exchange dust
@@
 stable to axisymmetric perturbations, regardless of how  thin the dust
-layer is. We also study the effect of dust-loading on the vertical shear
-instability \citep[VSI,][]{nelson13,lin15,barker15}. We provide 
-thermodynamical interpretations of our dusty stability calculations.
+layer is. We put forward a thermodynamical thermodynamical interpretation of
+dust-gas drag instabilities, and apply it to the
+streaming instability \citep[SI, ][]{youdin05a,youdin07b,jacquet11}. 
+We also study the effect of dust-loading on the vertical shear
+instability \citep[VSI,][]{nelson13,lin15,barker15}.
 
 This paper is organized as follows. In \S\ref{setup} we list the
-equations describing tightly-coupled dusty gas. We show they are in 
+one-fluid equations describing tightly-coupled dusty gas. We show that for a
+fixed equation of state, these are in 
 fact the usual hydrodynamic equations in disguise. 
-We describe approximate  equilibrium solutions in \S\ref{eqm}. In \S\ref{limits} we discuss
-general stability properties of the dusty disk in various
-limits. There, we give a thermodynamical interpretation of
-instabilities caused by directly by dust-gas drag. In
-\S\ref{linear_problem} we list the linearized equations, which we
-solve numerically in \S\ref{results} to study the dusty VSI. We 
+We describe equilibrium solutions in \S\ref{eqm} which are used in
+explicit calculations. In \S\ref{limits} we discuss
+general stability properties of dusty disks. We then move onto
+explicit stablity analyses. In  
+\S\ref{linear_problem} we list the linearized equations, and solve 
+them for the SI in \S\ref{si}, and in 
+ \S\ref{results} study dusty VSI. We 
 discuss future applications of the dusty/adiabatic gas equivalence in 

### 2016-10-22 summary.tex
edits in response to youdin's comments
fixed some typos, corrected kappa^2 dependence on dust
tiny correction in entropy def
head to toe edits

@@
 \begin{align}
-  s = \ln{\left[K^\frac{1}{\Gamma}(r,z)\left(1 - \tepsilon\right)\right]}.  
+  \seff = \ln{\left[K^{1/\Gamma}(r,z)\left(1 - \tepsilon\right)\right]}.  
 \end{align}
@@
 equation. When $c_s$ or $K$ is not constant and/or $\tstop\neq0$ one
-should add corresponding source terms in the energy equation. 
+should add corresponding source terms in the energy equation
+(e.g. Eq. \ref{poly_energy}). 
 The source term associated with dust-gas drag, $\mathcal{C}$, is 
 analogous to radiative diffusion \citep{price15}, which is also
-standard in many hydrodynamics codes.
+featured in many hydrodynamics codes.
 
@@
 characteristic temperatures are relevant to stability  
-\citep{toomre64}, having a non-trivial vertical temperature
+\citep{toomre64}, a non-trivial effective vertical temperature
 structure, induced by dust, may modify the vertical structure of 3D
@@
 \citep{lovelace99,li00}. These studies consider adiabatic pure gas and 
-show instability is possible if there is an extrema in the quantity
+show instability is possible if there is an extrema in the generalized
+potential vorticity 
 \begin{align}
- \eta = \frac{\kappa^2}{2\Omega\Sigma}\mathcal{S}^{2/\gamma},  
+ \mathcal{V} = \frac{\kappa^2}{2\Omega\Sigma}\mathcal{S}^{2/\gamma},  
 \end{align} 
@@
 condition for RWI in a (locally isothermal or polytroipc) dusty gas is
-also given by extrema in $\eta$ with $\mathcal{S} = \exp{s}$ (and
-$\tepsilon\to \Sigma_\mathrm{d}/\Sigma$). Since
-the dusty entropy $s$ reflects the dust distribution, we expect that
-RWI may also be triggered by extrema in  the dust-to-gas ratio.
+also given by extrema in $\mathcal{V}$ with $\mathcal{S} = \exp{\left(\Gamma
+S_\mathrm{eff}\right)}$ and $\tepsilon = \Sigma_\mathrm{d}/\Sigma$. Since
+the effective entropy reflects the dust distribution, we expect that
+RWI may also be triggered by extrema in the dust-to-gas ratio, e.g. 
+narrow dust rings. This would lead to direction formation of dusty
+vortices.
+
+%{\bf so may it's not vortex formation then dust-trapping. could just
+%  form dusty vortices in one go}
 
@@
   instabilities}\label{dust_as_thermo}
+
+%{\bf vsi enabled by drag? slipping between dust and gas reduces
+ % effective buoyancy}
 
@@
 $O(10^{-3}\hgas\OmK)$, We thus also rule out
-SI, as did \citeauthor{loren15}. However, the mode disappears with
-$\tstop=0$, so it is associated with finite dust-gas drag.
+SI, as did \citeauthor{loren15}. However, the mode disappears with $\tstop=0$, so it \emph{is}
+associated with finite dust-gas drag.
 
@@
 
-This preliminary calculation only serves to highlight the possibility
-of new axisymmetric dust-drag instabilities in stratified disks. One 
-issue is that the base state for the above calculation is not a true
-equilibrium (although \citeauthor {loren15} find the dust settling
-time is longer than the instability growth time). We will study
-this instability in more detail with a proper basic
-state in a follow up work.
+One possible interpretation of this mode is `drag-driven VSI', to be 
+compared with `cooling-driven VSI' studied by \cite{lin15}. In the
+latter case, buoyancy forces are eliminated by fast
+cooling, allowing vertical shear to destabilize the disk. Now, in an
+isothermal dusty disk if $\tstop=0$ then mixture is 
+effectively adiabatic: dust-induced buoyancy forces are
+strongest. However, having $\tstop\neq0$ allow dust particles to slip
+past the gas, reducing buoyancy. That is, dust-gas drag
+provies an effective cooling.
+
+We stress that this example only serves to highlight
+the possibility of new axisymmetric dust-drag instabilities in
+stratified disks. A major caveat of the above result is that 
+%the base state is not an exact equilibrium, because dust-settling would occur
+%in practice. (
+ the background dust settling timescale is comparable to the instability
+growth timescale. It is essential to construct
+exact equilibria for a meaningful stability analysis. We defer this
+to future work.
+
+%also unclear what are "proper" vertical bc for dusty disks
 
 \section{Summary}\label{summary}
-In this paper we highlight the similarity between an isothermal 
+In this paper we describe the similarity between an isothermal 
 dusty gas and an adiabatic, pure gas. The correspondence arises
@@
 
-For finite dust-gas friction, $\tstop\neq 0$, the dust content of a
-fluid parcel of the dusty mixture no longer conserved. A fluid parcel 
+For finite dust-gas drag, $\tstop\neq 0$, the dust content of a 
+parcel of the mixture no longer conserved. The parcel 
 can exchange dust particles with neighboring parcels owing 
@@
 evolutionary equation of  $\rhod/\rhog$ may be replaced by an 
-effective energy equation. We find the 
-natural definition of the entropy of the dusty gas is 
+effective energy equation. This leads to a 
+natural definition of the effective entropy of isothermal dusty gas as  
 \begin{align*}
-  s  = \ln{\left(\frac{c_s^2\rhog}{\rhog+\rhod}\right)}.  
+  \seff  = \ln{\left(\frac{c_s^2\rhog}{\rhog+\rhod}\right)}.  
 \end{align*}
@@
 %dusty gas. 
-In our framework, finite dust-gas friction appears as an energy
-source term.  This analogy with standard adiabatic
-hydrodynamics allow us to obtain dusty analogs of gaseous
+Finite dust-gas friction then appears as an energy
+source term.  This analogy with standard 
+hydrodynamics with cooling/heating allow us to obtain dusty analogs of gaseous
 instabilities, and provide thermodynamical interpretations of  
@@
 
-%
-In this first study, we apply this equivalence principle to study the
-axisymmetric stability of isothermal, protoplanetary disks
-perfectly coupled to dust ($\tstop=0$). This is so that we may 
-compute exact, steady equilibria for analyses. Furthermore, 
-in this limit the dusty problem is \emph{identical} to adiabatic, pure
-gas dynamics.
-
 We obtain the Solberg-Hoiland criteria for axisymmetric stability for
 strictly isothermal, perfectly-coupled dusty gas.  
-We show that the vertical shear associated with thin dust
+Applying this to typical protoplanetary disks, we find that 
+the vertical shear associated with dust 
  layers \emph{cannot} lead to axisymmetric  
-  instabilities, however large it is.  
-This is consistent with the fact that only 
-\emph{non-axisymmetric} Kelvin-Helmholtz instabilities have been 
-observed in numerical simulations in the perfectly-coupled limit 
-\citep{chiang08,barranco09,lee10}. We show that 
-sharp radial edges in the dust-to-gas ratio cannot persist in
-protoplanetary disks, as these imply a sharp entropy gradient and may
-thus destablize the disk.
+  instabilities, however thin the dust layer is.   
+%This is consistent with the fact that only 
+%\emph{non-axisymmetric} Kelvin-Helmholtz instabilities have been 
+%observed in numerical simulations in the perfectly-coupled limit 
+%\citep{chiang08,barranco09,lee10}. 
+Instead, sharp radial  edges in the dust-to-gas ratio could destabilized the
+disk, as these imply sharp gradients in the disk's effective entropy
+profile. % {\bf presumably this limits how sharp dust rings can be}
 
-%does not preclude non-axisymmetric instabilities. but if there is a
-%global radial temp gradient, then vertical shear arising from
-%it will always be comparable to dust-settling. but there is vsi with
-%global rad temp grad
+We  apply our thermodynamic framework to interpret
+the streaming instability \citep[SI, ][]{youdin05a,jacquet11} and to generalize the vertical shear
+instability \citep[VSI, ][]{nelson13,lin15} to dusty disks. We
+explicitly show that the SI occurs because the evolution of gas
+pressure lags behind  dust density; a general property of \emph{all}
+instabilities driven by strong dust-gas drag. 
+It takes a finite time for the gas to respond to
+the dust motion. A lag implies a time interval when
+gas pressure is still increasing while dust is already being expelled. 
+The dusty gas then does positive work, which goes into amplifying
+oscillation amplitudes. This interpretation is analogous to stellar pulsational
+instabilities \citep{cox67}.
 
-For locally isothermal disks, $c^2_s=c_s^2(r)$, we generalize the 
-vertical shear instability (VSI) to dusty disks, which is still 
-driven by the radial temperature gradient. We find that 
-dust-loading is generally stabilizing. In our disk models 
+For the VSI we find dust-loading is generally stabilizing. In our disk models 
 dust-loading does not affect VSI growth rates significantly, but
-meridional motions may be suppressed at heights where dust-induced
+meridional motions may be suppressed where the dust-induced
 vertical buoyancy dominates over vertical shear, consistent with our
@@
 
-%These results are readily
-%understood by comparing the vertical shear rate with the buoyancy of
-%an isothermal, dusty gas which we have correctly identified.
+We also discuss future applications of our thermodynamic framework to 
+study dusty protoplanetary disks. Because isothermal dusty gas has an
+effective entropy, we suggest that purely hydrodynamic processes, such
+as the Rossby Wave Instability \citep{li00} or the `Zombie Vortex
+Instability' \citep{marcus15} could have dusty counter-parts.   
+Furthermore, hydrodynamic instabilities driven by thermal cooling, such as VSI in
+stably-stratified disks \citep{lin15} or the `Convective
+Overstability' \citep{klahr14, lyra14},  may also find dusty analogs 
+because finite dust-gas drag is equivalent to a heat sink/source.
 
-We also discuss future applications of our thermodynamic framework to 
-study dusty protoplanetary disks. In particular, we emphasize the 
-physical origin of any instability driven by strong dust-gas
-drag is a phase lag between the pressure and total density evolution
-of the dusty gas mixture. This leads to growing modes because 
-positive work is done during an oscillation cycle, provided by
-dust-gas drag, which is equivalent to a `heat' 
-source. This interpretation is entirely analogous to pulsational
-instabilities in stars.
+%In particular, we emphasize the 
+%physical origin of any instability driven by strong dust-gas
+%drag is a phase lag between the pressure and total density evolution
+%of the dusty gas mixture. This leads to growing modes because 
+%positive work is done during an oscillation cycle, provided by
+%dust-gas drag, which is equivalent to a `heat' 
+%source.
 

### 2016-10-23 model.tex
head to toe edits

@@
 oscillation amplitude in fact grows: the work done leads the
-system to `overshoot' its equilibrium.
+system to `overshoot' its equilibrium. {\bf should be overshoot the max displacement}
 
@@
 increasing. This over-compensation causes the parcel to expand
-beyond its equilibrium volume. 
+beyond its equilibrium volume. {\bf should be: expand beyond the max
+  vol of the previous cycle}
 %corresponds to positive work done. 
@@
 forces are still dropping. This allows the  
-parcel's contraction to over-shoot its equilibrium density. The
+parcel's contraction to over-shoot its equilibrium density. {\bf
+  should be contract beyond the min vol of prev cycle}
+The
 overall positive work done leads to growth in the oscillation amplitude.

### 2016-10-24 model.tex
tiny fixes

--- 2016-10-22 model.tex
@@
 oscillation amplitude in fact grows: the work done leads the
-system to `overshoot' its equilibrium.
+system to `overshoot' past the amplitude of the preceeding oscillation
+cycle.
+
+%its equilibrium. {\bf should be overshoot the max displacement}
 
@@
 increasing. This over-compensation causes the parcel to expand
-beyond its equilibrium volume. 
+beyond the maximum volume of the previous cycle. 
+%{\bf should be: expand beyond the max
+%  vol of the previous cycle}
 %corresponds to positive work done. 
@@
 forces are still dropping. This allows the  
-parcel's contraction to over-shoot its equilibrium density. The
+parcel's contraction to over-shoot the maximum density attained in the
+previous cycle. 
+% {\bf
+  %should be contract beyond the min vol of prev cycle}
+The 
 overall positive work done leads to growth in the oscillation amplitude.

### 2016-11-01 linear.tex
adding kz=0 special case in response to inutsuka-san's query

@@
 
-\section{Compressible streaming instability}\label{si}
+\section{Dust-drag instabilities}\label{si}
 In the special case of an unstratified disk (neglecting the vertical
@@
 
+\subsection{Overstable dusty epicycles}
+In the special case of $k_z=0$, the incompressible dispersion
+relation, Eq. \ref{streaming_incompressible} reduces to 
+\begin{align}      %streaming_incompressible
+\ii \sigma^2 - \frac{\sigma}{\tepsilon \taus}  - \left(\ii \kappa^2 +
+  \frac{k_xF_r}{\epsilon}\right) = 0. \label{kz_zero_si}
+\end{align}
+If the quadratic term is neglected, then $\imag{\sigma}<0$, implying
+stability, as concluded by \cite{youdin05a}.
+
+However, retaining all the terms (but still in the incompressible
+limit) can yield instability, as follows. Eq. \ref{kz_zero_si} admit
+pure oscillations with $\sigma = -\omega$ provided that 
+\begin{align*}
+  \omega &= \pm \kappa,\\
+  \taus & = \frac{\epsilon\kappa}{\tepsilon k _x\left|F_r\right|}
+         \equiv \tau_{\mathrm{s}*}.
+\end{align*}
+If we linearize about this state of marginal stability by writing
+$\sigma = - \kappa + \ii \dd s$ and $\taus = \tau_{\mathrm{s}*} +
+\dd\taus$ and assume $|\dd s|\ll \kappa$ then 
+\begin{align*} 
+\dd s = \frac{2\tepsilon }{1 +
+  4\tepsilon^2\tau_{\mathrm{s}*}^2}\dd \taus. 
+\end{align*}
+Thus given $\tepsilon$ and $k_x$, there are overstable modes with
+stopping times slightly larger than $\tau_{\mathrm{s}*}$.
+
+%{\bf note: several approx used}
+
+\subsection{Streaming instability}
+
 {\bf note: kappa2 accounts for dust effect assuming smallh=0.05}
@@
 \citep[namely][]{youdin07b,bai10b}. These authors use normalized
-wavenumbers $K_x = \eta r k_x$ and similarly for $K_z$, where
+wavenumbers $K_{x,z} = \eta r k_{x,z}$ where
 \begin{align} 

### 2016-11-01 linear.tex
adding kz=0 case

@@
 
-\subsection{Overstable dusty epicycles}
-In the special case of $k_z=0$, the incompressible dispersion
-relation, Eq. \ref{streaming_incompressible} reduces to 
+\subsection{Overstable epicycles}
+If we further consider the limit $k_z=0$, the incompressible dispersion
+relation, Eq. \ref{streaming_incompressible}, reduces to 
 \begin{align}      %streaming_incompressible
@@
 \end{align} 
+Here we use $\tau_\mathrm{s}=\tstop/(1-\tepsilon)$ for convenience. 
 If the quadratic term is neglected, then $\imag{\sigma}<0$, implying
@@
 limit) can yield instability, as follows. Eq. \ref{kz_zero_si} admit
-pure oscillations with $\sigma = -\omega$ provided that 
+pure oscillations with $\sigma = -\omega$ when 
 \begin{align*}
-  \omega &= \pm \kappa,\\
-  \taus & = \frac{\epsilon\kappa}{\tepsilon k _x\left|F_r\right|}
-         \equiv \tau_{\mathrm{s}*}.
+  \omega &= \pm \kappa, \\
+  \taus & = \frac{\epsilon\omega}{\tepsilon k _x F_r }
+         \equiv \tau_{\mathrm{s}0}.
 \end{align*} 
-If we linearize about this state of marginal stability by writing
-$\sigma = - \kappa + \ii \dd s$ and $\taus = \tau_{\mathrm{s}*} +
-\dd\taus$ and assume $|\dd s|\ll \kappa$ then 
+Since $\taus\geq 0$, we require $k_xF_r > 0$ if $\omega = \kappa$;
+and $k_xF_r < 0$ if $\omega = -\kappa$. For $k_x>0$ these marginally
+stable epicycles require radial pressure gradients decreasing and increasing
+outwards, respectively.
+
+%We require $\omega/k_xF_r> 0$
+
+We now perturb about this state of marginal stability by writing
+$\sigma = - \omega + \dd \sigma$ with $\dd\sigma = \ii\dd s -
+\dd\omega$, $\taus = \tau_{\mathrm{s}0} +
+\dd\taus$, and assume the $\dd$ quantities are small in
+magnitude. Then 
 \begin{align*} 
-\dd s = \frac{2\tepsilon }{1 +
-  4\tepsilon^2\tau_{\mathrm{s}*}^2}\dd \taus. 
+\dd s = \frac{2\tepsilon\kappa^2 }{1 +
+  4\tepsilon^2\tau_{\mathrm{s}0}^2\kappa^2}\dd \taus. 
 \end{align*}
-Thus given $\tepsilon$ and $k_x$, there are overstable modes with
-stopping times slightly larger than $\tau_{\mathrm{s}*}$.
+Thus given $\tepsilon$ and $k_x$, there exist growing modes
+when the stopping time is slightly larger than
+$\tau_{\mathrm{s}c}$. These correspond to slowly-growing epicycles
+\citep[cf. the `convective overstability', ][]{klahr14,lyra14,latter16}.
 
@@
 {\bf note: kappa2 accounts for dust effect assuming smallh=0.05}
+
+We now consider the streaming instability with $k_z\neq 0$. 
 We solve Eq. \ref{streaming_dispersion} for selected cases where SI  
@@
 measures the pressure offset of Keplerian rotation. We use
-$\eta=0.05c_s/r\OmK$. They also define the stopping time as 
-$\tau_\mathrm{s}=\tstop/(1-\tepsilon)$.
+$\eta=0.05c_s/r\OmK$.
 

### 2016-11-02 linear.tex
more expansion on dust-drag numerical section

@@
 \begin{align}      %streaming_incompressible
-\ii \sigma^2 - \frac{\sigma}{\tepsilon \taus}  - \left(\ii \kappa^2 +
+\ii \sigma^2 - \frac{\sigma}{\epsilon \tstop}  - \left(\ii \kappa^2 +
   \frac{k_xF_r}{\epsilon}\right) = 0. \label{kz_zero_si}
 \end{align} 
-Here we use $\tau_\mathrm{s}=\tstop/(1-\tepsilon)$ for convenience. 
+%Here we use $\tau_\mathrm{s}=\tstop/(1-\tepsilon)$ for convenience. 
 If the quadratic term is neglected, then $\imag{\sigma}<0$, implying
@@
   \omega &= \pm \kappa, \\
-  \taus & = \frac{\epsilon\omega}{\tepsilon k _x F_r }
-         \equiv \tau_{\mathrm{s}0}.
+  \tstop & = \frac{\omega}{ k _x F_r }
+         \equiv t_{\mathrm{s}0}.
 \end{align*} 
-Since $\taus\geq 0$, we require $k_xF_r > 0$ if $\omega = \kappa$;
+Since $\tstop\geq 0$, we require $k_xF_r > 0$ if $\omega = \kappa$;
 and $k_xF_r < 0$ if $\omega = -\kappa$. For $k_x>0$ these marginally
@@
 We now perturb about this state of marginal stability by writing
-$\sigma = - \omega + \dd \sigma$ with $\dd\sigma = \ii\dd s -
-\dd\omega$, $\taus = \tau_{\mathrm{s}0} +
-\dd\taus$, and assume the $\dd$ quantities are small in
+$\sigma = \dd \sigma - \omega $ with $\dd\sigma = \ii\dd s -
+\dd\omega$, $\tstop = t_{\mathrm{s}0} +
+\dd\tstop $, and assume the $\dd$ quantities are small in
 magnitude. Then 
 \begin{align*} 
-\dd s = \frac{2\tepsilon\kappa^2 }{1 +
-  4\tepsilon^2\tau_{\mathrm{s}0}^2\kappa^2}\dd \taus. 
+\dd s = \frac{2\epsilon \kappa^2 }{1 +
+  4\epsilon^2 t_{\mathrm{s}0}^2\kappa^2}\dd \tstop. 
 \end{align*}
-Thus given $\tepsilon$ and $k_x$, there exist growing modes
+Thus given $k_xF_r$, there exist growing modes 
 when the stopping time is slightly larger than
-$\tau_{\mathrm{s}c}$. These correspond to slowly-growing epicycles
-\citep[cf. the `convective overstability', ][]{klahr14,lyra14,latter16}.
+$t_{\mathrm{s}0}$.
 
 %{\bf note: several approx used}
+
+If we instead fix $\tstop$ and vary $k_xF_r$ about marginal stability,
+a similar exercise yields  
+\begin{align*} 
+  \dd s = \frac{\epsilon \tstop^3}{1 +
+  4\epsilon^2 \tstop^2 \kappa^2}\dd \left(k_x^2F_r^2\right).  
+\end{align*}
+Thus it is possible to destabilize the system by slightly increasing
+$|k_xF_r|>\kappa/ \tstop$ (i.e. shorter radial wavelengths and/or
+stronger pressure gradients).
+
+\citep[cf. the `convective overstability', ][]{klahr14,lyra14,latter16}.
 
@@
 measures the pressure offset of Keplerian rotation. We use
-$\eta=0.05c_s/r\OmK$.
+$\eta=0.05c_s/r\OmK$. They also define the stopping time as
+$\tau_\mathrm{s}=\tstop/(1-\tepsilon)$.
 
@@
 
+\subsection{Numerical example}
+
+To demonstrate the above instabilities we numerically solve the
+full dispersion relation, Eq. \ref{streaming_dispersion} at fixed $K_x
+= 1500$ and $\tepsilon=2$. Growth rates are shown in 
+Fig. \ref{dusty_growth} as a function of $K_z$ at fixed $\taus\OmK=0.01$
+(top panel, which includes `linC' in Table \ref{si_compare}); and as a
+function of $\taus\OmK$ at fixed $K_z=50$ (bottom panel).
+
+The top panel shows that the streaming instability vanishes as
+$K_z\to0$, but it dominates most of $K_z$
+-space. The overstable epicycles become important for 
+$K_z\lesssim 100$, but their growth rates are weak. We thus
+expect streaming instabilities to dominate, provided small
+lengthscales can be resolved.
+
+In the bottom panel we confirm that the overstable epicycles require a
+sufficiently large stopping time at fixed wavenumbers. The vertical
+dashed line is the critical stopping time 
+\begin{align*}
+  \tau_{\mathrm{s}0}\OmK = \frac{1}{2K_x\left(1-\tepsilon\right)^2} 
+\end{align*}
+in the above notation. For this example, $\tau_{\mathrm{s}0}\OmK\simeq
+0.003$, as observed. On the other hand, the streaming instability
+shows weak dependence on $\taus$ for this range of stopping times.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/streaming2}
+  \includegraphics[width=\linewidth]{figures/streaming3}
+  \caption{Growth rates of dust-drag instabilities with fixed radial
+    wavenumber and dust-to-gas ratio, as a function of
+    the vertical wavenumber at fix stopping time (top); and as a
+    function of stopping time at fixed vertical wavenumber (bottom).  
+\label{dusty_growth}
+  }
+\end{figure}
+
 %\subsection{Boundary conditions}

### 2016-11-02 linear.tex
regular update
work/home sync

@@
 
-However, retaining all the terms (but still in the incompressible
-limit) can yield instability, as follows. Eq. \ref{kz_zero_si} admit
-pure oscillations with $\sigma = -\omega$ when 
+Now, writing $\sigma = \ii s - \omega$, the real and imaginary parts of 
+Eq. \ref{kz_zero_si}  are
+\begin{align}
+\omega^2 - s^2 - \frac{s}{\epsilon\tstop}  - \kappa^2 &= 0, \label{epi_overs1}\\
+\omega\left(2 s + \frac{1}{\epsilon\tstop}\right) - \frac{k_xF_r}{\epsilon} & =
+                                                                     0.  \label{epi_overs2}
+\end{align}
+The growth rate satisfies
 \begin{align*}
-  \omega &= \pm \kappa, \\
-  \tstop & = \frac{\omega}{ k _x F_r }
-         \equiv t_{\mathrm{s}0}.
+\left( \frac{k_xF_r\tstop}{1 + 2 s\epsilon\tstop}\right)^2 - s^2
+  - \frac{s}{\epsilon \tstop} = \kappa^2. %\omega = \frac{k_xF_r\tstop}{1 + 2 s\epsilon\tstop}
 \end{align*}
-Since $\tstop\geq 0$, we require $k_xF_r > 0$ if $\omega = \kappa$;
-and $k_xF_r < 0$ if $\omega = -\kappa$. For $k_x>0$ these marginally
-stable epicycles require radial pressure gradients decreasing and increasing
-outwards, respectively.
+
+Assuming slow growth such that $\epsilon s \tstop \ll 1$, we find 
+\begin{align}
+  s \simeq \epsilon \tstop \left(k_xF_r\tstop\right)^2 - \kappa^2,
+\end{align}
+Growth is possible for
+sufficiently large stopping times, and/or large radial pressure
+gradients, and/or sufficiently small radial wavelengths. 
+%For small growth rates the real frequency $|\omega| =  
+%O(\kappa)$, so these are overstable epicycles.
+
+It is interesting to note that Eq. \ref{epi_overs1}---\ref{epi_overs2}
+admit pure oscillations with $s \equiv 0$, $\omega = 
+\pm\kappa$ if 
+\begin{align}
+  \tstop = \frac{\kappa}{k_x\left|F_r\right|}. 
+\end{align}
+This is the condition for marginal stability.
+
+% However, retaining all the terms (but still in the incompressible
+% limit) can yield instability, as follows. Eq. \ref{kz_zero_si} admit
+% pure oscillations with $\sigma = -\omega$ when 
+% \begin{align*}
+%   \omega &= \pm \kappa, \\
+%   \tstop & = \frac{\omega}{ k _x F_r }
+%          \equiv t_{\mathrm{s}0}.
+% \end{align*} 
+% Since $\tstop\geq 0$, we require $k_xF_r > 0$ if $\omega = \kappa$;
+% and $k_xF_r < 0$ if $\omega = -\kappa$. For $k_x>0$ these marginally
+% stable epicycles require radial pressure gradients decreasing and increasing
+% outwards, respectively.
 
@@
 
-We now perturb about this state of marginal stability by writing
-$\sigma = \dd \sigma - \omega $ with $\dd\sigma = \ii\dd s -
-\dd\omega$, $\tstop = t_{\mathrm{s}0} +
-\dd\tstop $, and assume the $\dd$ quantities are small in
-magnitude. Then 
-\begin{align*} 
-\dd s = \frac{2\epsilon \kappa^2 }{1 +
-  4\epsilon^2 t_{\mathrm{s}0}^2\kappa^2}\dd \tstop. 
-\end{align*}
-Thus given $k_xF_r$, there exist growing modes 
-when the stopping time is slightly larger than
-$t_{\mathrm{s}0}$.
+% We now perturb about this state of marginal stability by writing
+% $\sigma = \dd \sigma - \omega $ with $\dd\sigma = \ii\dd s -
+% \dd\omega$, $\tstop = t_{\mathrm{s}0} +
+% \dd\tstop $, and assume the $\dd$ quantities are small in
+% magnitude. Then 
+% \begin{align*} 
+% \dd s = \frac{2\epsilon \kappa^2 }{1 +
+%   4\epsilon^2 t_{\mathrm{s}0}^2\kappa^2}\dd \tstop. 
+% \end{align*}
+% Thus given $k_xF_r$, there exist growing modes 
+% when the stopping time is slightly larger than
+% $t_{\mathrm{s}0}$.
 
-%{\bf note: several approx used}
+% %{\bf note: several approx used}
 
-If we instead fix $\tstop$ and vary $k_xF_r$ about marginal stability,
-a similar exercise yields  
-\begin{align*} 
-  \dd s = \frac{\epsilon \tstop^3}{1 +
-  4\epsilon^2 \tstop^2 \kappa^2}\dd \left(k_x^2F_r^2\right).  
-\end{align*}
-Thus it is possible to destabilize the system by slightly increasing
-$|k_xF_r|>\kappa/ \tstop$ (i.e. shorter radial wavelengths and/or
-stronger pressure gradients).
+% If we instead fix $\tstop$ and vary $k_xF_r$ about marginal stability,
+% a similar exercise yields  
+% \begin{align*} 
+%   \dd s = \frac{\epsilon \tstop^3}{1 +
+%   4\epsilon^2 \tstop^2 \kappa^2}\dd \left(k_x^2F_r^2\right).  
+% \end{align*}
+% Thus it is possible to destabilize the system by slightly increasing
+% $|k_xF_r|>\kappa/ \tstop$ (i.e. shorter radial wavelengths and/or
+% stronger pressure gradients).
 
@@
 full dispersion relation, Eq. \ref{streaming_dispersion} at fixed $K_x
-= 1500$ and $\tepsilon=2$. Growth rates are shown in 
+= 1500$ and $\epsilon=2$. Growth rates are shown in 
 Fig. \ref{dusty_growth} as a function of $K_z$ at fixed $\taus\OmK=0.01$
@@
 shows weak dependence on $\taus$ for this range of stopping times.
+
+\begin{align*}
+  s = \epsilon \left(1 - \tepsilon
+  \right)\taus\OmK^2\left[4K_x^2\left(1-\tepsilon\right)^4\left(\taus\OmK\right)^2
+  -1 \right]
+\end{align*}
 

### 2016-11-03 appendix.tex
adding kz=0 special case in response to inutsuka-san's query
more notes on kz=0 case (spurious?)

@@
 We consider an unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
-0$. The background $\p_rP/\rho$, $\tepsilon$ and $c_s^2$ are constant
+0$. The background $\p_rP/\rho$, $\tepsilon$ are constant
 input parameters. Then for the Epstein drag law, Eq. \ref{epstein}, we
@@
 $\p_r\to \ii k_x$ when acting on perturbations, and denote $|\bm{k}|^2
-\equiv k_x^2 + k_z^2$. We neglect 
-background gradients when compared with that of
-perturbations. The linearized
+\equiv k_x^2 + k_z^2$. We consider large $k_x$ and 
+%background gradients when compared with that of 
+%perturbations. 
+neglect the radial density gradient compared to that of pertuburbations 
+(formally setting $\p_r\rho=0$). This is also done in most local studies of
+dusty disks \citep[e.g.][]{youdin07b}.  
+The linearized
 equations are, after eliminating the azimuthal velocity:
@@
 \tstop/\left(1 - \tepsilon\right)$, we obtain: 
-\begin{align*}
+\begin{align}
 \ii \tepsilon \tau_\mathrm{s}
@@
   k_x \tau_\mathrm{s} F_r\left(1 - 2\tepsilon\right)
-  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0,
-\end{align*}
+  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0,\label{streaming_incompressible}
+\end{align}
 as derived by \cite{jacquet11} and \cite{laibe14} for the streaming

### 2016-11-03 linear.tex
more notes on kz=0 case (spurious?)

@@
 is retained. We now discuss solutions to these equations in an unstratified disk 
- for the streaming instability (\S\ref{si}), and a stratified disk for the VSI (\S\ref{results}).
+in \S\ref{si} and a stratified disk in \S\ref{results}.
 
@@
 
-\subsection{Overstable epicycles}
+\subsection{Overstable dusty epicycles}
 If we further consider the limit $k_z=0$, the incompressible dispersion
@@
 %Here we use $\tau_\mathrm{s}=\tstop/(1-\tepsilon)$ for convenience. 
-If the quadratic term is neglected, then $\imag{\sigma}<0$, implying
-stability, as concluded by \cite{youdin05a}.
+Neglecting the quadratic term leads to stability, $\imag{\sigma}<0$,
+as concluded by \cite{youdin05a}.
 
-Now, writing $\sigma = \ii s - \omega$, the real and imaginary parts of 
-Eq. \ref{kz_zero_si}  are
+However, the real and imaginary parts of Eq. \ref{kz_zero_si}  are
 \begin{align}
-\omega^2 - s^2 - \frac{s}{\epsilon\tstop}  - \kappa^2 &= 0, \label{epi_overs1}\\
 \omega\left(2 s + \frac{1}{\epsilon\tstop}\right) - \frac{k_xF_r}{\epsilon} & =
-                                                                     0.  \label{epi_overs2}
+                                                                     0, \label{epi_overs1} \\
+\omega^2 - s^2 - \frac{s}{\epsilon\tstop}  - \kappa^2 &= 0, \label{epi_overs2}
 \end{align}
-The growth rate satisfies
+so the growth rate satisfies 
 \begin{align*}
@@
 \begin{align}
-  s \simeq \epsilon \tstop \left(k_xF_r\tstop\right)^2 - \kappa^2,
+  &|\omega| \simeq \left|k_xF_r\right| \tstop, \\
+  &s \simeq \epsilon \tstop \left[\left(k_xF_r\tstop\right)^2 - \kappa^2\right],\label{epi_grow}
 \end{align}
-Growth is possible for
-sufficiently large stopping times, and/or large radial pressure
-gradients, and/or sufficiently small radial wavelengths. 
+Notice the $\tstop$ and $\tstop^3$ dependence for 
+$\omega$ and $s$. This is also true for the streaming 
+instability for sufficiently small $\tstop$ \citep{jacquet11}. Unlike
+SI though, here there can be growth when $\epsilon=1$, whereas the SI
+is severly suppressed at equal dust-to-gas ratio.
+
+According to Eq. \ref{epi_grow}, growth is possible for sufficiently
+large stopping times, and/or large radial pressure gradients, and/or
+sufficiently small radial wavelengths. For example, given $k_xF_r$
+instability requires 
+\begin{align}
+\tstop > \frac{\kappa}{k_x\left|F_r\right|}\equiv t_{\mathrm{s}0}.\label{epi_grow_criterion}
+\end{align}
+The mode is marginally stable if $\tstop = t_{\mathrm{s}0}$.
+
+These slowly-growing modes have $|\omega|\simeq \kappa$, so they correspond to
+overstable epicycles. This property is similar to the `convective 
+overstability' discussed in dust-free disks 
+\citep{klahr14,lyra14,latter16}. In that case, the combination of an 
+unstable radial entropy gradient ($N_r^2  \propto \p_rP\p_rS < 0 $) and
+finite thermal relaxation time leads to growing epicycles. Indeed, the
+effective radial buoyancy frequency in our dusty disk 
+with negligble background density gradient is unstable, 
+\begin{align}
+  N_r^2 \equiv F_r\frac{\p S_\mathrm{eff}}{\p r} =
+  -\frac{\rho}{P}F_r^2<0.  
+\end{align}
+
+Convective overstability cannot exist in our locally isothermal disks,
+which have zero cooling times. On the other hand, dust-gas drag 
+is equivalent to heating/cooling in our framework. Thus it may not be 
+surprising to find a dusty analog of the convective 
+overstability. Both growth rates scale with 
+$|N_r^2|$, and both require finite effective cooling times.
+
+However, the structure of the modes differ. The dusty epicycles 
+have no vertical structure and require $k_x\neq0$; whereas the
+convective overstability has no radial structure but require 
+$k_z\neq0$. Dusty epicycles may thus operate in  
+razor-thin disks, but convective overstability requires the vertical
+dimension.
+
+An important caveat is that the one-fluid approximation for dusty gas 
+may not hold well when $|\omega| \gtrsim 
+\Omega$ \citep{youdin05a}, as is the case for dusty epicycles. Their
+existence should be checked with full dust-plus-gas 
+simulations. Several 2D, razor-thin disk simulations do suggest a
+dusty instability that destory vortices
+\citep{fu14b,raettig15,surville16}. These setups are significantly
+more complex than our models, but \citeauthor{raettig15} did find 
+particle clumping in a test simulation without a vortex. These results
+might be related to the overstable dusty epicycles discussed here.
+
+%While there is some uncertainty on the reality of the overstable dusty
+%epicyclic modes, 
+Real disks are three-dimensional, so the importance of dusty 
+epicycles should be compared with $k_z\neq0$ modes, which
+includes the classic streaming instability \citep{youdin05a},
+discussed next.
+
 %For small growth rates the real frequency $|\omega| =  
@@
 
-It is interesting to note that Eq. \ref{epi_overs1}---\ref{epi_overs2}
-admit pure oscillations with $s \equiv 0$, $\omega = 
-\pm\kappa$ if 
-\begin{align}
-  \tstop = \frac{\kappa}{k_x\left|F_r\right|}. 
-\end{align}
-This is the condition for marginal stability.
+%We finally note that Eq. \ref{epi_overs1}---\ref{epi_overs2} 
+%admit pure oscillations with $s \equiv 0$, $\omega = 
+%\pm\kappa$ if 
+%\begin{align}
+%  \tstop = \frac{\kappa}{k_x\left|F_r\right|}. 
+%\end{align}
+%This is the condition for marginal stability.
 
@@
 
-\citep[cf. the `convective overstability', ][]{klahr14,lyra14,latter16}.
-
 \subsection{Streaming instability}
@@
 {\bf note: important to use largrangian pressure pert properly}
+
 We also checked the growth rates $s$ satisfy 
@@
 
-{\bf suppose large scale pressure grad drops outwards. 
-pert the sys by kicking more dust inwards. gas moves out, into region 
-  of lower pressure in the bg. i.e. create pressure bump at larger
-  radius. it attracts the first dust particles back out, plus some more
-  because bump is larger. (move gas from higher density to lower
-  density produce asymmetric bump/trough.}
+%{\bf suppose large scale pressure grad drops outwards. 
+%pert the sys by kicking more dust inwards. gas moves out, into region 
+%  of lower pressure in the bg. i.e. create pressure bump at larger
+%  radius. it attracts the first dust particles back out, plus some more
+%  because bump is larger. (move gas from higher density to lower
+%  density produce asymmetric bump/trough.}
 
@@
 
-To demonstrate the above instabilities we numerically solve the
+We numerically solve the
 full dispersion relation, Eq. \ref{streaming_dispersion} at fixed $K_x
 = 1500$ and $\epsilon=2$. Growth rates are shown in 
-Fig. \ref{dusty_growth} as a function of $K_z$ at fixed $\taus\OmK=0.01$
-(top panel, which includes `linC' in Table \ref{si_compare}); and as a
-function of $\taus\OmK$ at fixed $K_z=50$ (bottom panel).
+Fig. \ref{dusty_growth1} as a function of $K_z$ at fixed $\taus\OmK=0.01$
+(which includes `linC' in Table \ref{si_compare}); and in
+Fig. \ref{dusty_growth2} as a function of $\taus\OmK$ at fixed
+$K_z=50$.
 
-The top panel shows that the streaming instability vanishes as
-$K_z\to0$, but it dominates most of $K_z$
--space. The overstable epicycles become important for 
-$K_z\lesssim 100$, but their growth rates are weak. We thus
-expect streaming instabilities to dominate, provided small
-lengthscales can be resolved.
-
-In the bottom panel we confirm that the overstable epicycles require a
-sufficiently large stopping time at fixed wavenumbers. The vertical
-dashed line is the critical stopping time 
-\begin{align*}
-  \tau_{\mathrm{s}0}\OmK = \frac{1}{2K_x\left(1-\tepsilon\right)^2} 
-\end{align*}
-in the above notation. For this example, $\tau_{\mathrm{s}0}\OmK\simeq
-0.003$, as observed. On the other hand, the streaming instability
-shows weak dependence on $\taus$ for this range of stopping times.
-
+Fig. \ref{dusty_growth1} show the overstable dusty epicycles' growth
+rates are weakly depedent on $K_z$. They are well approximated by
+that in the $K_z=0$ limit, Eq. \ref{epi_grow}, which becomes 
 \begin{align*}
@@
 \end{align*}
+in the above notation. 
+For the case considerd in Fig. \ref{dusty_growth1} we find
+$s\simeq0.067\OmK$, as observed. By contrast, the streaming
+instability requires $K_z>0$, and dominates when
+$K_z\gtrsim 100$.
 
@@
   \includegraphics[width=\linewidth]{figures/streaming2}
+  \caption{Growth rates of dust-drag instabilities with fixed radial
+    wavenumber and dust-to-gas ratio, as a function of
+    the vertical wavenumber at fix stopping time. \label{dusty_growth1}
+  }
+\end{figure}
+
+In Fig. \ref{dusty_growth2} we fix the wavenumbers and vary the
+stopping time. The vertical dashed line is the critical stopping time
+for the overstable dusty epicycles, Eq. \ref{epi_grow_criterion},
+which is  
+\begin{align*}
+  \tau_{\mathrm{s,min}}\OmK = \frac{1}{2K_x\left(1-\tepsilon\right)^2} 
+\end{align*}
+in the above notation. For this example, $\tau_{\mathrm{s,min}}\OmK\simeq 
+0.003$, as observed. On the other hand, the streaming instability
+shows weak dependence on $\taus$ for this range of stopping times 
+(it ultimated vanishes as $\taus\to0$).
+
+\begin{figure}
   \includegraphics[width=\linewidth]{figures/streaming3}
   \caption{Growth rates of dust-drag instabilities with fixed radial
-    wavenumber and dust-to-gas ratio, as a function of
-    the vertical wavenumber at fix stopping time (top); and as a
-    function of stopping time at fixed vertical wavenumber (bottom).  
-\label{dusty_growth}
+    wavenumber and dust-to-gas ratio, as a
+    function of stopping time at fixed vertical wavenumber. The
+    vertical dashed line correspond to the minimum stopping time for
+    growth of the dusty epicycle, given by
+    Eq. \protect\ref{epi_grow_criterion}. \label{dusty_growth2}
   }
 \end{figure}
+
+These calculations suggest that in a realistic 3D dusty disk where
+vertical motions are allowed, the classic streaming
+instability is dominant because there exists $K_z$ values for which it
+dominates. This also renders the uncertainties in the  
+approximations used to describe the dusty epicycle modes
+irrelevant. On the other hand, the one-fluid framework can capture the
+streaming instability in the appropriate limit, as seen in Table
+\ref{si_compare}.
 

### 2016-11-03 appendix.tex
dusty epicycles are not real, move to appendix

@@
 The equation of state $P = c_s^2\left( 1 - \tepsilon\right)\rho$
-was used. If we now set $\zeta = 0$ or consider $c_s^2\to \infty$ to
-obtain the 
-incompressible limit, and define $\tau_\mathrm{s} \equiv
+was used.
+
+\subsection{Incompressible limit}
+
+ We can set $\zeta = 0$ or consider $c_s^2\to \infty$ to
+obtain the incompressible limit. Using $\tau_\mathrm{s} \equiv
 \tstop/\left(1 - \tepsilon\right)$, we obtain: 
@@
 as derived by \cite{jacquet11} and \cite{laibe14} for the streaming
-instability. If $|\sigma|/\OmK = O(\tau_\mathrm{s}\OmK)$ and
+instability. 
+If $|\sigma|/\OmK = O(\tau_\mathrm{s}\OmK)$ and
 $\tau_\mathrm{s}\OmK\ll 1$, then the quartic term
@@
 dispersion relation of \citet{youdin05a}.
+
+\subsection{Spuriously growing epicycles}
+
+A caveat of the dispersion relations 
+Eq. \ref{streaming_dispersion} and \ref{streaming_incompressible} is
+that they admit spurious unstable modes with $|\sigma|\gtrsim \Omega$
+that violates the one-fluid framework used in this
+paper. We demonstrate this below by considering the
+incompressible dispersion relation. (We checked numerically that
+compressibilty has negligible effects on the modes examined.)
+
+Consider the limit $k_z=0$. Then Eq. \ref{streaming_incompressible}
+becomes 
+\begin{align}      %streaming_incompressible
+\ii \sigma^2 - \frac{\sigma}{\epsilon \tstop}  - \left(\ii \kappa^2 +
+  \frac{k_xF_r}{\epsilon}\right) = 0. \label{kz_zero_si}
+\end{align} 
+Neglecting the quadratic term leads to stability, $\imag{\sigma}<0$, 
+as found by \cite{youdin05a} in their full two-fluid analysis as 
+$k_z\to0$.  
+% However, the real and imaginary parts of Eq. \ref{kz_zero_si}  are
+% \begin{align}
+% \omega\left(2 s + \frac{1}{\epsilon\tstop}\right) - \frac{k_xF_r}{\epsilon} & =
+%                                                                      0, \label{epi_overs1} \\
+% \omega^2 - s^2 - \frac{s}{\epsilon\tstop}  - \kappa^2 &= 0, \label{epi_overs2}
+% \end{align}
+% so the growth rate satisfies 
+% \begin{align*}
+% \left( \frac{k_xF_r\tstop}{1 + 2 s\epsilon\tstop}\right)^2 - s^2
+%   - \frac{s}{\epsilon \tstop} = \kappa^2. %\omega = \frac{k_xF_r\tstop}{1 + 2 s\epsilon\tstop}
+% \end{align*} 
+However, solving Eq. \ref{kz_zero_si} explicitly assuming $\epsilon s \tstop \ll 1$ would yield 
+\begin{align}
+  |\omega| &\simeq \left|k_xF_r\right| \tstop \sim \kappa, \\
+  s &\simeq \epsilon \tstop \left[\left(k_xF_r\tstop\right)^2 - \kappa^2\right],\label{epi_grow}
+\end{align}
+Accordingly, given $k_xF_r$, growth is possible if 
+\begin{align}
+\tstop > \frac{\kappa}{k_x\left|F_r\right|}\equiv t_{\mathrm{s}0}.\label{epi_grow_criterion}
+\end{align}
+The mode is marginally stable when $\tstop =
+t_{\mathrm{s}0}$. Alternatively, for fixed $\tstop$ growth is enabled
+by a sufficiently large $k_xF_r$. Unlike the streaming instability,
+which is strongly surpressed when $\epsilon=1$, these modes can still
+grow at equal dust-to-gas ratio.
+
+These growing epicycles have $|\omega|\gtrsim \kappa$. 
+However, these modes are absent in full two-fluid
+models \citep{youdin05a}. The discrepency lies in the fact that the
+first-order  one-fluid equations are only valid for low-frequency waves with
+$|\omega|\lesssim \Omega$. Thus only 
+low-frequency modes should be retained from analyses based on the
+first-order one-fluid approximation.
+
+%When this is violated,
+%terms of $O(\tstop^2)$ should be retained in deriving the one-fluid
+%equations (Eq. \ref{masseq}---\ref{tempeq}).
+
+Fig. \ref{dusty_growth1} show unstable modes found from
+Eq. \ref{streaming_dispersion} as a function of $K_z$ at fixed $K_x 
+= 1500$, $\epsilon=2$ and $\taus\OmK=0.01$.  
+The (spurious) overstable dusty epicycles' growth
+rates are weakly depedent on $K_z$ and are well approximated by
+that in the $K_z=0$ limit, Eq. \ref{epi_grow}, which becomes 
+\begin{align*} 
+  s = \epsilon \left(1 - \tepsilon
+  \right)\taus\OmK^2\left[4K_x^2\left(1-\tepsilon\right)^4\left(\taus\OmK\right)^2
+  -1 \right]
+\end{align*}
+in the above notation. 
+For the case considerd in Fig. \ref{dusty_growth1} we find
+$s\simeq0.067\OmK$, as observed. By contrast, the streaming
+instability requires $K_z>0$, and dominates when
+$K_z\gtrsim 100$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/streaming2}
+  \caption{Growth rates of dust-drag instabilities with fixed radial
+    wavenumber and dust-to-gas ratio, as a function of
+    the vertical wavenumber at fix stopping time. \label{dusty_growth1}
+  }
+\end{figure}
+
+Caution must be taken if the first-order one-fluid equations are used
+to simulate dusty gas. 
+%One should ensure that the physical dust-gas instabilty
+%(which should have low-frequency in order for the first-order one-fluid
+%approximation) has a larger growth rate than that of the spurious
+%modes, which is given by Eq. \ref{epi_growth}. 
+%Fig. \ref{dusty_growth1} suggests that the
+%streaming instability would dominate, provided there is sufficient
+%resolution to resolve high $K_z$ modes.  
+For example, 2D, razor-thin disks would allow the spurious epicycles
+to dominate, since in that case the streaming instability cannot
+operate. Fig. \ref{dusty_growth1} suggest that in a realistic 3D disks
+where a range of $K_z$ is allowed, the streaming instability should
+dominate.
+
+Thus simulations based on the first-order one-fluid equations should
+be set up to surpress these spurious epicycles. This might be
+achieved, for example, through physical or numerical viscosity to
+eliminate high-$k_x$ modes, since for fixed disk/dust parameters these
+spurious epicycles only operate at sufficiently small radial
+wavelengths. Alternatively, one needs to ensure the physical
+instabilities of interest have larger growth rates than the spurious
+epicycles.
+
+%An important caveat is that the one-fluid approximation for dusty gas 
+%may not hold well when $|\omega| \gtrsim 
+%\Omega$ \citep{youdin05a}, as is the case for dusty epicycles.
+
+% Their
+% existence should be checked with full dust-plus-gas 
+% simulations. Several 2D, razor-thin disk simulations do suggest a
+% dusty instability that destory vortices
+% \citep{fu14b,raettig15,surville16}. These setups are significantly
+% more complex than our models, but \citeauthor{raettig15} did find 
+% particle clumping in a test simulation without a vortex. These results
+% might be related to the overstable dusty epicycles discussed here.
+
+% Real disks are three-dimensional, so the importance of dusty 
+% epicycles should be compared with $k_z\neq0$ modes, which
+% includes the classic streaming instability \citep{youdin05a},
+% discussed next.
+
+%\subsection{Numerical example}
+
+% In Fig. \ref{dusty_growth2} we fix the wavenumbers and vary the
+% stopping time. The vertical dashed line is the critical stopping time
+% for the overstable dusty epicycles, Eq. \ref{epi_grow_criterion},
+% which is  
+% \begin{align*}
+%   \tau_{\mathrm{s,min}}\OmK = \frac{1}{2K_x\left(1-\tepsilon\right)^2} 
+% \end{align*}
+% in the above notation. For this example, $\tau_{\mathrm{s,min}}\OmK\simeq 
+% 0.003$, as observed. On the other hand, the streaming instability
+% shows weak dependence on $\taus$ for this range of stopping times 
+% (it ultimated vanishes as $\taus\to0$).
+
+% \begin{figure}
+%   \includegraphics[width=\linewidth]{figures/streaming3}
+%   \caption{Growth rates of dust-drag instabilities with fixed radial
+%     wavenumber and dust-to-gas ratio, as a
+%     function of stopping time at fixed vertical wavenumber. The
+%     vertical dashed line correspond to the minimum stopping time for
+%     growth of the dusty epicycle, given by
+%     Eq. \protect\ref{epi_grow_criterion}. \label{dusty_growth2}
+%   }
+% \end{figure}
+
+% These calculations suggest that in a realistic 3D dusty disk where
+% vertical motions are allowed, the classic streaming
+% instability is dominant because there exists $K_z$ values for which it
+% dominates. This also renders the uncertainties in the  
+% approximations used to describe the dusty epicycle modes
+% irrelevant. On the other hand, the one-fluid framework can capture the
+% streaming instability in the appropriate limit, as seen in Table
+% \ref{si_compare}.
+
+%For small growth rates the real frequency $|\omega| =  
+%O(\kappa)$, so these are overstable epicycles.
+
+%We finally note that Eq. \ref{epi_overs1}---\ref{epi_overs2} 
+%admit pure oscillations with $s \equiv 0$, $\omega = 
+%\pm\kappa$ if 
+%\begin{align}
+%  \tstop = \frac{\kappa}{k_x\left|F_r\right|}. 
+%\end{align}
+%This is the condition for marginal stability.
+
+% However, retaining all the terms (but still in the incompressible
+% limit) can yield instability, as follows. Eq. \ref{kz_zero_si} admit
+% pure oscillations with $\sigma = -\omega$ when 
+% \begin{align*}
+%   \omega &= \pm \kappa, \\
+%   \tstop & = \frac{\omega}{ k _x F_r }
+%          \equiv t_{\mathrm{s}0}.
+% \end{align*} 
+% Since $\tstop\geq 0$, we require $k_xF_r > 0$ if $\omega = \kappa$;
+% and $k_xF_r < 0$ if $\omega = -\kappa$. For $k_x>0$ these marginally
+% stable epicycles require radial pressure gradients decreasing and increasing
+% outwards, respectively.
+
+%We require $\omega/k_xF_r> 0$
+
+% We now perturb about this state of marginal stability by writing
+% $\sigma = \dd \sigma - \omega $ with $\dd\sigma = \ii\dd s -
+% \dd\omega$, $\tstop = t_{\mathrm{s}0} +
+% \dd\tstop $, and assume the $\dd$ quantities are small in
+% magnitude. Then 
+% \begin{align*} 
+% \dd s = \frac{2\epsilon \kappa^2 }{1 +
+%   4\epsilon^2 t_{\mathrm{s}0}^2\kappa^2}\dd \tstop. 
+% \end{align*}
+% Thus given $k_xF_r$, there exist growing modes 
+% when the stopping time is slightly larger than
+% $t_{\mathrm{s}0}$.
+
+% %{\bf note: several approx used}
+
+% If we instead fix $\tstop$ and vary $k_xF_r$ about marginal stability,
+% a similar exercise yields  
+% \begin{align*} 
+%   \dd s = \frac{\epsilon \tstop^3}{1 +
+%   4\epsilon^2 \tstop^2 \kappa^2}\dd \left(k_x^2F_r^2\right).  
+% \end{align*}
+% Thus it is possible to destabilize the system by slightly increasing
+% $|k_xF_r|>\kappa/ \tstop$ (i.e. shorter radial wavelengths and/or
+% stronger pressure gradients).

### 2016-11-03 linear.tex
dusty epicycles are not real, move to appendix

--- 2016-11-02 linear.tex
@@
 
-\section{Dust-drag instabilities}\label{si}
+\section{Streaming instability}\label{si}
 In the special case of an unstratified disk (neglecting the vertical
@@
 
-\subsection{Overstable epicycles}
-If we further consider the limit $k_z=0$, the incompressible dispersion
-relation, Eq. \ref{streaming_incompressible}, reduces to 
-\begin{align}      %streaming_incompressible
-\ii \sigma^2 - \frac{\sigma}{\epsilon \tstop}  - \left(\ii \kappa^2 +
-  \frac{k_xF_r}{\epsilon}\right) = 0. \label{kz_zero_si}
-\end{align} 
-%Here we use $\tau_\mathrm{s}=\tstop/(1-\tepsilon)$ for convenience. 
-If the quadratic term is neglected, then $\imag{\sigma}<0$, implying
-stability, as concluded by \cite{youdin05a}.
-
-However, retaining all the terms (but still in the incompressible
-limit) can yield instability, as follows. Eq. \ref{kz_zero_si} admit
-pure oscillations with $\sigma = -\omega$ when 
-\begin{align*}
-  \omega &= \pm \kappa, \\
-  \tstop & = \frac{\omega}{ k _x F_r }
-         \equiv t_{\mathrm{s}0}.
-\end{align*} 
-Since $\tstop\geq 0$, we require $k_xF_r > 0$ if $\omega = \kappa$;
-and $k_xF_r < 0$ if $\omega = -\kappa$. For $k_x>0$ these marginally
-stable epicycles require radial pressure gradients decreasing and increasing
-outwards, respectively.
-
-%We require $\omega/k_xF_r> 0$
-
-We now perturb about this state of marginal stability by writing
-$\sigma = \dd \sigma - \omega $ with $\dd\sigma = \ii\dd s -
-\dd\omega$, $\tstop = t_{\mathrm{s}0} +
-\dd\tstop $, and assume the $\dd$ quantities are small in
-magnitude. Then 
-\begin{align*} 
-\dd s = \frac{2\epsilon \kappa^2 }{1 +
-  4\epsilon^2 t_{\mathrm{s}0}^2\kappa^2}\dd \tstop. 
-\end{align*}
-Thus given $k_xF_r$, there exist growing modes 
-when the stopping time is slightly larger than
-$t_{\mathrm{s}0}$.
-
-%{\bf note: several approx used}
-
-If we instead fix $\tstop$ and vary $k_xF_r$ about marginal stability,
-a similar exercise yields  
-\begin{align*} 
-  \dd s = \frac{\epsilon \tstop^3}{1 +
-  4\epsilon^2 \tstop^2 \kappa^2}\dd \left(k_x^2F_r^2\right).  
-\end{align*}
-Thus it is possible to destabilize the system by slightly increasing
-$|k_xF_r|>\kappa/ \tstop$ (i.e. shorter radial wavelengths and/or
-stronger pressure gradients).
-
-\citep[cf. the `convective overstability', ][]{klahr14,lyra14,latter16}.
-
-\subsection{Streaming instability}
+\subsection{Numerical examples}
 
@@
 
-We now consider the streaming instability with $k_z\neq 0$. 
-We solve Eq. \ref{streaming_dispersion} for selected cases where SI  
+%We now consider the streaming instability with $k_z\neq 0$. 
+We solve the full dispersion relation, Eq. \ref{streaming_dispersion} for selected cases where SI  
 growth rates have been 
@@
 
-\subsection{Numerical example}
+\subsection{Comparison to the convective overstability}
+%These slowly-growing modes have $|\omega|\simeq \kappa$, so they correspond to
+%overstable epicycles. 
+The streaming
+instability bears some similarity  to the `convective 
+overstability' in dust-free, non-isothermal disks 
+\citep{klahr14,lyra14,latter16}. In that case, the combination of an 
+unstable radial entropy gradient ($N_r^2  \propto \p_rP\p_rS < 0 $) and
+finite thermal relaxation time leads to epicyclics with
+growth rate $\propto  -N_r^2$
 
-To demonstrate the above instabilities we numerically solve the
-full dispersion relation, Eq. \ref{streaming_dispersion} at fixed $K_x
-= 1500$ and $\tepsilon=2$. Growth rates are shown in 
-Fig. \ref{dusty_growth} as a function of $K_z$ at fixed $\taus\OmK=0.01$
-(top panel, which includes `linC' in Table \ref{si_compare}); and as a
-function of $\taus\OmK$ at fixed $K_z=50$ (bottom panel).
+Indeed, the effective radial buoyancy frequency in our dusty disk 
+with negligble background density gradient is unstable, 
+\begin{align}
+  N_r^2 \equiv F_r\frac{\p S_\mathrm{eff}}{\p r} =
+  -\frac{\rho}{P}F_r^2<0.  
+\end{align}
+Convective overstability cannot exist in our locally isothermal disks,
+which have zero cooling times. On the other hand, dust-gas drag 
+is equivalent to heating/cooling in our framework.
 
-The top panel shows that the streaming instability vanishes as
-$K_z\to0$, but it dominates most of $K_z$
--space. The overstable epicycles become important for 
-$K_z\lesssim 100$, but their growth rates are weak. We thus
-expect streaming instabilities to dominate, provided small
-lengthscales can be resolved.
+One might then think of the streaming instability as a dusty relative
+to the convective  overstability. Explicit solutions in the small
+$\tstop$ limit show that streaming instability growth rates are
+also $\propto |N_r^2|$  \citep{jacquet11}, as is for the convective
+overstability \citep{lyra14}. Furthermore, both instabilities require
+finite effective cooling times, and both require $k_z\neq0$. 
+{\bf note: si osc freq not too small compared to kappa}
 
-In the bottom panel we confirm that the overstable epicycles require a
-sufficiently large stopping time at fixed wavenumbers. The vertical
-dashed line is the critical stopping time 
-\begin{align*}
-  \tau_{\mathrm{s}0}\OmK = \frac{1}{2K_x\left(1-\tepsilon\right)^2} 
-\end{align*}
-in the above notation. For this example, $\tau_{\mathrm{s}0}\OmK\simeq
-0.003$, as observed. On the other hand, the streaming instability
-shows weak dependence on $\taus$ for this range of stopping times.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/streaming2}
-  \includegraphics[width=\linewidth]{figures/streaming3}
-  \caption{Growth rates of dust-drag instabilities with fixed radial
-    wavenumber and dust-to-gas ratio, as a function of
-    the vertical wavenumber at fix stopping time (top); and as a
-    function of stopping time at fixed vertical wavenumber (bottom).  
-\label{dusty_growth}
-  }
-\end{figure}
+However, the convective overstability is most unstable at $k_x=0$
+\citep{latter14}  whereas the streaming instability growth rate is maximized at finite $k_x$
+\citep{youdin05a}. Nevertheless, they share the same underlying
+physical mechanism, i.e. pressure-density phase lags.
 

### 2016-11-06 summary.tex
adding kz=0 special case in response to inutsuka-san's query
home/office sync

@@
 
-\subsection{Generalization to locally polytropic disks}
+\subsection{Generalizations}
+
+\subsubsection{Locally polytropic disks}
 We can extend the dusty/adiabatic gas correspondence to 
@@
 %Eqs. \ref{dusty_solberg1}---\ref{dusty_solberg2}.
+
+\subsubsection{Higher order corrections as additional stresses}
+\begin{align}
+\left.\frac{D\bm{v}}{Dt}\right|_\mathrm{add} = -\frac{1}{\rho}\nabla\cdot\left(\frac{\rhod\rhog}{\rho}\Delta\bm{v}\otimes\Delta\bm{v}\right). 
+\end{align}
 

### 2016-11-07 appendix.tex
dusty epicycles are not real, move to appendix
home/work sync

@@
 $\p_r\to \ii k_x$ when acting on perturbations, and denote $|\bm{k}|^2
-\equiv k_x^2 + k_z^2$. We consider large $k_x$ and 
-%background gradients when compared with that of 
-%perturbations. 
-neglect the radial density gradient compared to that of pertuburbations 
-(formally setting $\p_r\rho=0$). This is also done in most local studies of
+\equiv k_x^2 + k_z^2$. We consider large $k_x$ and background
+gradients when compared with that of  perturbations. 
+%neglect the radial density gradient compared to that of pertuburbations 
+%(formally setting $\p_r\rho=0$). 
+ This is also done in most local studies of
 dusty disks \citep[e.g.][]{youdin07b}.  
@@
 
-\subsection{Spuriously growing epicycles}
+\subsection{Spuriously growing epicycles}\label{spurious_epi}
 
@@
 first-order  one-fluid equations are only valid for low-frequency waves with
-$|\omega|\lesssim \Omega$. Thus only 
+$|\sigma|\lesssim \Omega$. Thus only 
 low-frequency modes should be retained from analyses based on the

### 2016-11-07 summary.tex
home/work sync

--- 2016-10-22 summary.tex
@@
 \subsection{Generalization to locally polytropic disks}
+
+%\subsubsection{Locally polytropic disks}
 We can extend the dusty/adiabatic gas correspondence to 
@@
 %Eqs. \ref{dusty_solberg1}---\ref{dusty_solberg2}.
+
+\subsection{Higher order corrections}
+{\bf notes for self. maybe exclude in final. }
+
+As discussed in \S\ref{spurious_epi}, the first-order one-fluid 
+formulation (Eqs. \ref{masseq}---\ref{tempeq}) can produce
+spuriously-growing modes with frequency $|\sigma|\gtrsim \Omega$. This
+results from neglecting terms of $O(|\Delta\bm{v}|^2)$. The full set
+of one-fluid equations is significantly more complex than our
+model \cite[e.g.][]{laibe14}. However, it is possible to improve upon
+the first-order equations whilst retaining similarilty to
+standard hydrodynamics.
+
+The full one-fluid equations of motion includes the acceleration term 
+\begin{align}
+%\left.\frac{D\bm{v}}{Dt}\right|_\mathrm{add} = 
+\bm{a} =  
+-\frac{1}{\rho}\nabla\cdot\left(\frac{\rhod\rhog}{\rho}\Delta\bm{v}\otimes\Delta\bm{v}\right) 
+\equiv  \frac{1}{\rho}\nabla\cdot \bm{T},
+\end{align}
+where $\bm{T}$ is an effective stress tensor. %This term is neglected
+%when working to $O(\left|\Delta\bm{v}\right|)$, as we have done
+%thusfar.  % In 
+% general $\Delta\bm{v}$ obeys an evolutionary equation. In that case, the stress 
+% tensor $\bm{T}$ is reminenscent of magnetic tension in
+% magneto-hydrodynamics. 
+In the terminal velocity approximation with $\Delta\bm{v} = \tstop
+\nabla P/\rhog$, we have 
+\begin{align}
+\bm{T} = - \frac{\epsilon\tstop^2}{\rho}\nabla P \otimes \nabla P.
+\end{align}
+If we define 
+\begin{align}
+  \bm{g} \equiv \sqrt{4\pi G \epsilon \rho} \frac{\nabla P}{\rho} \tstop,
+\end{align}
+then
+\begin{align}
+\bm{T} = - \frac{1}{4\pi G} \bm{g}\otimes\bm{g}. \label{effective_g}
+\end{align}
+
+{\bf tensor virial theorems?}
+
+However, Eq. \ref{effective_g} is just the 
+gravitational stress tensor in self-gravitating hydrodynamics
+%\footnote{Technically there is an isotropic part, which} 
+\citep{lynden-bell72}.
+
+ It is
 

### 2016-11-08 summary.tex
added notes on future work

--- 2016-10-22 summary.tex
@@
 %Eqs. \ref{dusty_solberg1}---\ref{dusty_solberg2}.
+
+\subsection{Higher order corrections}
+%{\bf notes for self. maybe exclude in final. }
+
+The simplified one-fluid formulation,
+Eqs. \ref{masseq}---\ref{tempeq}, can 
+produce spuriously-growing modes with frequency $|\sigma|\gtrsim
+\Omega$ (see Appendix \S\ref{spurious_epi}). This 
+results from neglecting terms of $O(|\Delta\bm{v}|^2)$ when converting
+the two-fluid equations to the one-fluid formulation. The full
+one-fluid equations is significantly more complex than our 
+model \cite[e.g.][]{laibe14}. However, it is possible to improve upon
+the first-order formulation whilst retaining similarilty to  
+standard hydrodynamics.
+
+The full one-fluid equations of motion includes the acceleration term 
+\begin{align}
+%\left.\frac{D\bm{v}}{Dt}\right|_\mathrm{add} = 
+\bm{a} =  
+-\frac{1}{\rho}\nabla\cdot\left(\frac{\rhod\rhog}{\rho}\Delta\bm{v}\otimes\Delta\bm{v}\right) 
+\equiv  \frac{1}{\rho}\nabla\cdot \bm{T},
+\end{align}
+where $\bm{T}$ is an effective stress tensor. 
+In the terminal velocity approximation with $\Delta\bm{v} = \tstop
+\nabla P/\rhog$, we have 
+\begin{align}
+\bm{T} = - \frac{\epsilon\tstop^2}{\rho}\nabla P \otimes \nabla P.
+\end{align}
+If we define 
+\begin{align}
+  \bm{g} \equiv \sqrt{4\pi G \epsilon \rho} \frac{\nabla P}{\rho}
+  \tstop,\label{effective_g} 
+\end{align}
+then
+\begin{align}
+\bm{T} = - \frac{1}{4\pi G} \bm{g}\otimes\bm{g}, \label{effective_gstress}
+\end{align}
+%{\bf tensor virial theorems?}
+which is just the anisotropic part of the  
+gravitational stress tensor in self-gravitating hydrodynamics 
+\citep{lynden-bell72}. However, unlike self-gravity, where $\bm{g}$ is
+determined through the Poisson equation; here $\bm{g}$ is directly
+related to $P$ (Eq. \ref{effective_g}), which itself obeys an 
+evolutionary equation for isothermal gas (Eq. \ref{eff_energy}).
+
+Nevertheless, Eq. \ref{effective_g}---\ref{effective_gstress}
+indicates that pressure gradients provide torques for angular momentum
+transport of the mixture as a whole. This is only possible for
+non-axisymmetric pressure forces so that $T_{r\phi}\neq0$.
+
+In a subsequent paper we will extend the one-fluid formulation by
+including $\bm{T}$ in the equation of motion. 
+In particular, we will explore whether this can surpress the suprious
+modes discussed in Appendix \S\ref{spurious_epi}.
+
+%stratified case require more boundary conditions
 

### 2016-11-08 appendix.tex
general clean up

@@
 $\p_r\to \ii k_x$ when acting on perturbations, and denote $|\bm{k}|^2
-\equiv k_x^2 + k_z^2$. We consider large $k_x$ and background
-gradients when compared with that of  perturbations. 
+\equiv k_x^2 + k_z^2$. We consider large $k_x$ and thus neglect background
+gradients when compared to that of perturbations.  
 %neglect the radial density gradient compared to that of pertuburbations 
@@
 Neglecting the quadratic term leads to stability, $\imag{\sigma}<0$, 
-as found by \cite{youdin05a} in their full two-fluid analysis as 
-$k_z\to0$.  
+as found by \cite{youdin05a} in their full two-fluid analysis. 
 % However, the real and imaginary parts of Eq. \ref{kz_zero_si}  are
@@
 t_{\mathrm{s}0}$. Alternatively, for fixed $\tstop$ growth is enabled
-by a sufficiently large $k_xF_r$. Unlike the streaming instability,
+by a sufficiently large $\left|k_xF_r\right|$. Unlike the streaming instability,
 which is strongly surpressed when $\epsilon=1$, these modes can still
@@
 
-These growing epicycles have $|\omega|\gtrsim \kappa$. 
-However, these modes are absent in full two-fluid
-models \citep{youdin05a}. The discrepency lies in the fact that the
-first-order  one-fluid equations are only valid for low-frequency waves with
-$|\sigma|\lesssim \Omega$. Thus only 
-low-frequency modes should be retained from analyses based on the
-first-order one-fluid approximation.
+These growing epicycles with $|\omega|\gtrsim \kappa$ are absent in
+full two-fluid models \citep{youdin05a}. The discrepency lies in the
+fact that the one-fluid equations, to first order in $\tstop$, are
+only valid for low-frequency waves with $|\sigma|\lesssim
+\Omega$. Thus only low-frequency modes should be retained from
+analyses based on Eqs. \ref{masseq}---\ref{tempeq}.
 
@@
 to dominate, since in that case the streaming instability cannot
-operate. Fig. \ref{dusty_growth1} suggest that in a realistic 3D disks
-where a range of $K_z$ is allowed, the streaming instability should
-dominate.
+operate. However, Fig. \ref{dusty_growth1} show that in realistic 3D
+disks where a range of $K_z$ is allowed, the streaming instability
+should dominate.
 

### 2016-11-08 linear.tex
general clean up

@@
 
-{\bf note: kappa2 accounts for dust effect assuming smallh=0.05}
+%{\bf note: kappa2 accounts for dust effect assuming smallh=0.05}
 
-%We now consider the streaming instability with $k_z\neq 0$. 
 We solve the full dispersion relation, Eq. \ref{streaming_dispersion} for selected cases where SI  
@@
 
-{\bf note: important to use largrangian pressure pert properly}
+%{\bf note: important to use largrangian pressure pert properly}
 
@@
 Eulerian and Lagragian pressure perturbations is unimportant. 
-{\bf but weak growth in this case}
+%{\bf but weak growth in this case}
 %However
@@
 \citep{klahr14,lyra14,latter16}. In that case, the combination of an 
-unstable radial entropy gradient ($N_r^2  \propto \p_rP\p_rS < 0 $) and
-finite thermal relaxation time leads to epicyclics with
-growth rate $\propto  -N_r^2$
+unstable radial buoyancy frequency ($N_r^2 < 0 $) and
+finite thermal relaxation time leads to growing epicyclics. %with
+%growth rate $\propto  -N_r^2$
 
@@
 \begin{align}
-  N_r^2 \equiv F_r\frac{\p S_\mathrm{eff}}{\p r} =
+  N_r^2 \equiv -\frac{1}{\rho}\frac{\p P}{\p r}\frac{\p S_\mathrm{eff}}{\p r} =
   -\frac{\rho}{P}F_r^2<0.  
@@
 also $\propto |N_r^2|$  \citep{jacquet11}, as is for the convective
-overstability \citep{lyra14}. Furthermore, both instabilities require
+overstability \citep{lyra14,latter16}. Furthermore, both instabilities require
 finite effective cooling times, and both require $k_z\neq0$. 
-{\bf note: si osc freq not too small compared to kappa}
+%{\bf note: si osc freq not too small compared to kappa}
 
 However, the convective overstability is most unstable at $k_x=0$
-\citep{latter14}  whereas the streaming instability growth rate is maximized at finite $k_x$
+\citep{lyra14}  whereas the streaming instability growth rate is maximized at finite $k_x$
 \citep{youdin05a}. Nevertheless, they share the same underlying

### 2016-11-08 model.tex
more expansion on dust-drag numerical section
home/office sync
general clean up

@@
 differs slightly from the stopping time $\tau_\mathrm{s} =
-\tstop\rhog/\left(\rhog+\rhod\right)$ defined in previous studies
+\tstop\rhog/\left(\rhog+\rhod\right)$ used in some studies
 \citep[e.g.][]{youdin05a,jacquet11}.
 
-We consider the limit of tightly-coupled dust particles with 
+We consider tightly-coupled dust particles with 
 $\tstop\OmK\ll 1 $, where $\OmK$ is the Keplerian orbital
@@
 \begin{align}
-  \bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} = \frac{\nabla
+  \Delta\bm{v}\equiv\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} = \frac{\nabla
     P}{\rhog}\tstop, 
 \end{align}
-\citep{youdin05a,jacquet11}.
+\citep{youdin05a,jacquet11}. More generally, though, $\Delta\bm{v}$
+obeys an evolutionary
 
@@
 In this limit of strong dust-gas drag, \cite{laibe14} give a single
-fluid description of the mixture: 
+fluid description of the mixture, to $O\left(\tstop\right)$, as:  
 \begin{align} 
@@
 The gas component of the mixture has a specific entropy 
-$S = C_p\ln{\left(P^{1/\gamma}/\rhog\right)}$, where $C_p$ is the heat capacity at constant
-pressure. However, this is not a useful quantity for (locally)
-isothermal gas.
+$S = C_p\ln{\left(P^{1/\gamma}/\rhog\right)}$, where $C_p$ is the heat 
+capacity at constant pressure. However, this is not a useful quantity
+for (locally) isothermal gas.
 
@@
 effectively has $\gamma=1$ we can define an effective (dimensionless)  
-entropy for the \emph{entire mixture} as 
+entropy for the \emph{mixture} as 
 \begin{align}
@@
 \end{align} 
-We shall see that by defining an effective entropy this way, many of the 
+This follows naturally by combining the effective 
+energy Eq. \ref{eff_energy} and Eq. \ref{masseq}. 
+With an effective entropy defined this way, many of the 
 results concerning the stability of dusty gas will have identical form
@@
 
-The physical reason for this analogy is that with strong drag, the dust is almost
-perfectly entrained in the gas flow, but there is some gain/loss of 
-dust particles between different gas parcels if dust-gas
-friction if finite. This property is analogous to the entropy of an adiabatic
-pure gas subject to heating/cooling: entropy is conserved following
-a gas parcel, except if there is heat exchange between a fluid parcel
-and the surrounding.
+The physical reason for this analogy is that with strong drag, the
+dust is almost perfectly entrained in the gas flow, but there is some
+gain/loss of dust particles between different gas parcels. This
+property is analogous to the entropy of an adiabatic pure gas subject
+to heating/cooling: entropy is conserved following a gas parcel,
+except if there is heat exchange between a fluid parcel and the
+surrounding.
 
@@
 
-In practice the dust particles settle vertically on a timescale 
+In practice, dust particles settle to the disk midplane on a timescale 
 $t_\mathrm{settle}\sim 1/\OmK^2\tstop$, although for tightly-coupled dust
@@
 True steady states may be calculated by invoking an underlying gas
-turbulence, which leads to dust diffusion \citep{takeuchi02, youdin07, 
- lyra13}. This complication is beyond the scope of this paper.
+turbulence, which leads to dust diffusion counter-acting vertical
+settling \citep{takeuchi02, youdin07}. This complication is beyond the
+scope of this paper.
 
-In this work, we will assume that $\tstop$  is sufficiently small so
+In this work, we assume that $\tstop$  is sufficiently small so
 that settling occurs on much longer timescales  than that of interest; or
@@
 \item 
-  unstratified disks with finite dust-gas drag (\S\ref{si});  
+  unstratified disks with finite dust-gas drag; %(\S\ref{si});  
 \item 
-  perfectly-coupled dust with $\tstop=0$ (\S\ref{results}). 
+  perfectly-coupled dust with $\tstop=0$. %(\S\ref{results}). 
 \end{inparaenum} 
@@
 \subsection{Disk structure with a prescribed dust distribution}\label{steady_state}  
-To obtain an actual disk structure for numerical computations, we
-assume a Gaussian profile in the dust-to-gas ratio,   
+For numerical computations, we assume a Gaussian profile in the
+dust-to-gas ratio,    
 \begin{align}\label{dust_gauss}
@@
 The first two terms correspond to vertical shear caused by spatial
-variations in the dust-to-gas ratio; while the third term
+variations in the dust-to-gas ratio. The third term
 proportional to $q$ corresponds to vertical shear due to the 
-radial temperature gradient. The last term survives in the dust-free
-limit.
+radial temperature gradient, and survives in the dust-free limit.
 
@@
 
-{\bf lagrangian derivative follows the mixture. this highlights one
-  perk of the one-fluid framework. in two-fluid models there are two
-  types of lagragian deriv, not clear which fluid to follow.}
+%{\bf lagrangian derivative follows the mixture.}
+% this highlights one
+%  perk of the one-fluid framework. in two-fluid models there are two
+%  types of lagragian deriv, not clear which fluid to follow.}
 

### 2016-11-08 result.tex
general clean up

@@
 
-We now present numerical solutions to the linearized equations in a 
-vertically stratified disk. We consider perfectly coupled dust 
-($\tstop=0$) so that the equilibria defined in \S\ref{eqm} are exact
+We present numerical solutions for  
+vertically stratified disks assuming perfectly coupled dust 
+($\tstop=0$). This is so that the equilibria defined in \S\ref{eqm} are exact
 steady states. In this limit any growing modes must be associated with
-the imposed temperature gradient (\S\ref{limits}). We thus 
-examine the effect of dust-loading on the VSI, previously studied in 
+the imposed temperature gradient (\S\ref{limits}). We are thus  
+examining the effect of dust-loading on the VSI, previously studied in 
 pure gas disks \citep[][\citetalias{lin15} in this section]{lin15}.
 
-For simplicity, we mostly consider constant midplane
+Except in  \S\ref{varHd}, we consider constant midplane
 dust-to-gas ratios $\epsilon_0$ and its characteristic height
 $H_\epsilon$. The latter is specified via the dust layer thickness
-$\Hd(r_0)$. (We will allow for radial dependices in $\epsilon_0$ and
-$H_\epsilon$ in \S\ref{varHd}.)   
-These, together with the perturbation wavenumber $k$, comprises main the
+$\Hd(r_0)$. 
+These, together with the perturbation radial wavenumber $k_x$, comprises main the
 parameters of the linear problem. We choose a power-law 
@@
 consists 
-of the roughly horizontal `body modes' and the nearly-vertical
-`surface modes'. The latter is due to the imposed vertical boundaries
-\citep{barker15}.
+of the roughly horizontal `body modes', and the nearly-vertical
+`surface modes' \citep[which are associated with the imposed vertical
+boundaries][]{barker15}.
 
 We find that increasing the dust-to-gas ratio reduce VSI growth
-rates. Notably, `surface modes', which are typically fastest growing in
+rates. Notably, surface modes, which are typically fastest growing in
 the dust-free case, are surpressed in dusty disks for $\epsilon_0\geq
@@
 This is consistent with Fig. \ref{compare_vshear_fixHd} comparing the basic state vertical
-shear and buoyancy. Notice also the perturbed dust-to-gas ratio,
-$\delta\epsilon$, shifts from being maximized at the disk boundaries
+shear and buoyancy. Notice also maxium perturbtions to the dust-to-gas ratio,
+$\delta\epsilon$, shifts from disk boundaries 
 to the midplane.
@@
 In Fig. \ref{vsi_dust_loading_vareps} we plot the growth rates as a
-function of $\epsilon_0$ for different perturbation wavenumbers $k$. We
-find that dust-loading stabilizes the VSI more effectively for shorter
+function of $\epsilon_0$ for different perturbation wavenumbers
+$k_x$. We find dust-loading stabilizes the VSI more effectively for shorter
 wavelength perturbations. This is because for high wavenumbers the
 dominant modes are surface modes, which are effectively stabilized by
-dust-loading since bouyancy forces are larges near the disk
+dust-loading since buoyancy forces are larges near the disk
 boundaries. The figure suggest that VSI becomes much less efficient
@@
 In \S\ref{iso_perfect} we asserted that in the limit of perfectly
-coupled dust (as considered here) that radially smooth, strictly
+coupled dust: radially smooth, strictly
 isothermal disks are stable against axisymmetric perturbations, no
@@
 \S\ref{vertshear}). However, we see that growth rates $s\propto
-|q|$. In particular, this holds for $\rich < 0.1$, the critical value
+|q|$, i.e .growth rates vanish if the disk was strictly isothermal. 
+In particular, this holds for $\rich < 0.1$, the critical value
 for non-axisymmetric instabilities.
@@
 0$). This result is independent of $\rich$, so the Richardson number
-is not useful in characterizing the axisymmetric stability of dust
+is does not characterize the axisymmetric stability of dust
 layers.
@@
     dust-settling, but axisymmetric instability is associated with the
-    radial temperature gradient $q$.
+    radial temperature gradient $q$. Here, $\rich_0$ is the minium
+    Richardson number in the domain in the limit $q\to0$. 
     \label{ultra_thin}
@@
 \subsection{Effect of a radially-varying dust-to-gas ratio}\label{varHd}
-{\bf bonus stuff, maybe delete in final version}
+%{\bf bonus stuff, maybe delete in final version}
 
@@
 $O(\hgas\OmK)$. Importantly, the increase in the growth rate of the
-fundamental mode is small. Thus we do not expect the radial
+fundamental mode is small. We do not expect the radial
 dependence in $\epsilon$ to significantly affect the VSI in a dusty
@@
 We considered $\tstop=0$ in order to perform stability analysis on an
-exact steady state. When $\tstop\neq0$, a 
-disk initialized as in \S\ref{eqm} will evolve 
-due to dust-gas relative drift. In particular, in the
-absence of turbulent/diffusive processes, dust particles settle to the
+exact steady state. When $\tstop\neq0$, a stratified dusty 
+disk initialized as in \S\ref{eqm} would evolve 
+%due to dust-gas relative drift. In particular, in the
+%absence of turbulent/diffusive processes, dust particles 
+as dust settles to the
 midplane on a timescale $t_\mathrm{settle}\sim 1/\tstop\OmK^2$

### 2016-11-08 summary.tex
general clean up

@@
 
-The simplified one-fluid formulation,
+The one-fluid formulation, 
 Eqs. \ref{masseq}---\ref{tempeq}, can 
@@
 results from neglecting terms of $O(|\Delta\bm{v}|^2)$ when converting
-the two-fluid equations to the one-fluid formulation. The full
-one-fluid equations is significantly more complex than our 
-model \cite[e.g.][]{laibe14}. However, it is possible to improve upon
+the full two-fluid equations to the one-fluid formulation. The full
+one-fluid equations 
+%is significantly more complex than our 
+%model \cite[e.g.][]{laibe14}. It 
+ includes an evolutionary equation for $\Delta\bm{v}$, which has no
+ hydrodynamic counter-part \citep{laibe14}. However, it is
+ still possible to improve upon 
  the first-order formulation whilst retaining similarilty to  
  standard hydrodynamics.  
+%without having the complexity of the full equations
 
@@
 %\left.\frac{D\bm{v}}{Dt}\right|_\mathrm{add} = 
-\bm{a} =  
+%\bm{a} =  
 -\frac{1}{\rho}\nabla\cdot\left(\frac{\rhod\rhog}{\rho}\Delta\bm{v}\otimes\Delta\bm{v}\right) 
-\equiv  \frac{1}{\rho}\nabla\cdot \bm{T},
+\equiv  \frac{1}{\rho}\nabla\cdot \bm{T} 
 \end{align}
+\citep{youdin05a}, 
 where $\bm{T}$ is an effective stress tensor. 
-In the terminal velocity approximation with $\Delta\bm{v} = \tstop
+In the terminal velocity
+approximation with $\Delta\bm{v} = \tstop 
 \nabla P/\rhog$, we have 
@@
 gravitational stress tensor in self-gravitating hydrodynamics 
-\citep{lynden-bell72}. However, unlike self-gravity, where $\bm{g}$ is
+\citep{lynden-bell72}.
+
+ However, unlike self-gravity, where $\bm{g}$ is
 determined through the Poisson equation; here $\bm{g}$ is directly
 related to $P$ (Eq. \ref{effective_g}), which itself obeys an 
-evolutionary equation for isothermal gas (Eq. \ref{eff_energy}).
+evolutionary equation (Eq. \ref{eff_energy}). 
+Nevertheless, Eq. \ref{effective_g}---\ref{effective_gstress}
+show that correlations in  pressure gradients lead to 
+momentum transport of the mixture as a whole. This is similar to Reynolds stresses in
+a turbulent disk \citep{balbus99}. In particular, 
+outwards angular momentum transport requires $\p_rP\p_\phi P > 0$.
 
-Nevertheless, Eq. \ref{effective_g}---\ref{effective_gstress}
-indicates that pressure gradients provide torques for angular momentum
-transport of the mixture as a whole. This is only possible for
-non-axisymmetric pressure forces so that $T_{r\phi}\neq0$.
+In a following paper we will include $\bm{T}$ in the one-fluid
+equations and futher explore the analogy between dusty-gas and viscous
+hydrodynamics.
 
-In a subsequent paper we will extend the one-fluid formulation by
-including $\bm{T}$ in the equation of motion. 
-In particular, we will explore whether this can surpress the suprious
-modes discussed in Appendix \S\ref{spurious_epi}.
+%In particular, we will explore whether this can surpress the suprious
+%modes discussed in Appendix \S\ref{spurious_epi}.
 

### 2016-11-09 summary.tex
tiny edits

@@
 \section{Summary}\label{summary}
-In this paper we describe the similarity between an isothermal 
-dusty gas and an adiabatic, pure gas. The correspondence arises
+In this paper we develop an analogy between isothermal 
+dusty-gas and adiabatic, pure gas. The correspondence arises
 because drag forces reduce the relative velocity between gas and
@@
  dust is entrained in 
-the gas, and the dust-to-gas ratio $\rhod/\rhog$ is conserved
-following the dusty flow. This is analogous to entropy being conserved
+the gas. Then the dust-to-gas ratio $\rhod/\rhog$ is conserved
+following the flow. This is analogous to entropy being conserved
 following a adiabatic, pure gas.
@@
 For finite dust-gas drag, $\tstop\neq 0$, the dust content of a 
-parcel of the mixture no longer conserved. The parcel 
-can exchange dust particles with neighboring parcels owing 
-to dust-gas relative drift. This is analogous to heat exchange between
-a pure gas parcel and its surroundings.
+parcel of the mixture is no longer conserved. The parcel 
+can exchange dust particles with neighboring parcels. % owing 
+%to dust-gas relative drift. 
+This is analogous to heat exchange between a parcel of pure gas and its surroundings.
 
-We explicitly show that for a locally isothermal, dusty gas that the 
-evolutionary equation of  $\rhod/\rhog$ may be replaced by an 
+We explicitly show that for a fixed gas equation of state, the  
+evolutionary equation for $\rhod/\rhog$ may be replaced by an 
 effective energy equation. This leads to a 
@@
 It takes a finite time for the gas to respond to
-the dust motion. A lag implies a time interval when
-gas pressure is still increasing while dust is already being expelled. 
+the dust motion. A lag implies there exists a time interval where 
+the gas pressure of a parcel of the mixture is increasing whilst dust is already being expelled.  
 The dusty gas then does positive work, which goes into amplifying
@@
 previous study \citep{lin15}.
+Since the dust-induced buoyancy forces increase away from the midplane, we find dust-loading can stabilize `surface modes' of the VSI, that would otherwise have the largest growth rates.
 

### 2016-11-11 model.tex
tiny changes

@@
 $(\rhod,\bm{v}_\mathrm{d})$, respectively. The mixture has a single 
-temperature $T$, and the mixture pressure $P$ arise solely from the 
-gas component.
+temperature $T$, and its pressure $P$ arise solely from the 
+gas component. %neglect turbulence, viscosity, self-g and b fields.
 
@@
   \rhog\left.\frac{\p \bm{v}_\mathrm{g}}{\p t}\right|_\mathrm{drag}=
-  - \frac{\rhog\rhod}{\left(\rhod+\rhog\right)}\frac{\left(\bm{v}_\mathrm{d} -
-    \bm{v}_\mathrm{g}\right)}{\tstop}. 
+  - \frac{\rhog\rhod}{\rho}\frac{\Delta\bm{v}}{\tstop}. 
 \end{align}
-is the dust-gas friction force per unit volume. Note that $\tstop$ 
+is the dust-gas friction force per unit volume, where
+\begin{align}
+  \Delta\bm{v} \equiv\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} 
+\end{align}
+is the dust-gas velocity difference; and 
+\begin{align}
+  \rho \equiv \rhog + \rhod
+\end{align}
+is the total density. 
+Note that $\tstop$ 
 differs slightly from the stopping time $\tau_\mathrm{s} =
-\tstop\rhog/\left(\rhog+\rhod\right)$ used in some studies
+\tstop\rhog/\rho$ used in some studies
 \citep[e.g.][]{youdin05a,jacquet11}.
 
-We consider tightly-coupled dust particles with 
-$\tstop\OmK\ll 1 $, where $\OmK$ is the Keplerian orbital
-frequency. This allows us to simplify the problem by using the
-`terminal velocity approximation' so that the dust-gas velocity
-difference is 
+In general $\Delta\bm{v}$ obeys a complicated evolutionary equation
+\citep[see, e.g.][]{youdin05a}. However, for tightly-coupled dust
+particles with $\tstop\OmK\ll 1 $, where $\OmK$ is the Keplerian orbital
+frequency, we can use the 
+`terminal velocity approximation' to set %The dust-gas velocity difference 
+%Then so that the dust-gas velocity
+%difference is 
 \begin{align}
-  \Delta\bm{v}\equiv\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} = \frac{\nabla
-    P}{\rhog}\tstop, 
+  \Delta\bm{v} = \frac{\nabla
+    P}{\rhog}\tstop 
 \end{align}
-\citep{youdin05a,jacquet11}. More generally, though, $\Delta\bm{v}$
-obeys an evolutionary
+\citep{youdin05a,jacquet11}. This equation implies that dust particles
+drift towards pressure maxima.
 
 %In the limit of strong dust-gas drag (stopping time $\tstop\to 0$) 
-In this limit of strong dust-gas drag, \cite{laibe14} give a single
-fluid description of the mixture, to $O\left(\tstop\right)$, as:  
+In the limit of strong dust-gas drag ($\tstop\to0$), \cite{laibe14}
+give a single fluid description of the mixture, to
+$O\left(\tstop\right)$, as:   
 \begin{align} 
@@
   \bm{v} \equiv \frac{\rhog\bm{v}_\mathrm{g} + 
-    \rhod\bm{v}_\mathrm{d}}{\rho}, 
+    \rhod\bm{v}_\mathrm{d}}{\rho}. 
 \end{align}
-and 
-\begin{align}
-  \rho \equiv \rhog + \rhod
-\end{align}
-is the total mixture density. For an ideal gas the pressure is given
+For an ideal gas the pressure is given
 by $P = \mathcal{R}\rhog T/\mu $, where $\mathcal{R}$ is
@@
 \end{align} 
-This follows naturally by combining the effective 
-energy Eq. \ref{eff_energy} and Eq. \ref{masseq}. 
-With an effective entropy defined this way, many of the 
-results concerning the stability of dusty gas will have identical form
-and interpretations to that in standard adiabatic hydrodynamics of 
-pure gas. For example, if $c_s^2$ is constant and $\tstop=0$, then
-$D\seff/ D t = 0$ since $D\tepsilon/D t=0$ in that case.
+%This follows naturally by combining the effective 
+%energy Eq. \ref{eff_energy} and Eq. \ref{masseq}. 
+Then 
+\begin{align*}
+  \frac{D \seff}{D t} = \bm{v}\cdot\nabla\ln{c_s^2} +
+  \frac{c_s^2}{P}\nabla\cdot\left(\tstop\tepsilon\nabla P\right), 
+\end{align*}
+which is just the statement of entropy evolution in standard  
+hydrodynamics, albeit with source terms. With an effective
+entropy defined this way, many of the  
+results concerning the stability of (locally) isothermal dusty gas
+will have identical form and interpretations to that for pure gas.
 
-The physical reason for this analogy is that with strong drag, the
-dust is almost perfectly entrained in the gas flow, but there is some
-gain/loss of dust particles between different gas parcels. This
-property is analogous to the entropy of an adiabatic pure gas subject
-to heating/cooling: entropy is conserved following a gas parcel,
-except if there is heat exchange between a fluid parcel and the
-surrounding.
+%If $c_s^2$ is constant and $\tstop=0$, then 
+%$D\seff/ D t = D\tepsilon/D t=0$. In this case the effective 
+%entropy is conserved because the dust-fraction is `frozen in' the
+%flow.
+
+The physical reason for this analogy is that with strong drag
+($\tstop\to0$), dust is almost perfectly entrained in the gas, but
+there is some gain/loss of dust particles between different  
+gas parcels. This is analogous to the entropy of an ideal 
+pure gas subject to heating/cooling: entropy is conserved following
+a gas parcel, except if there is heat exchange between a fluid parcel and the
+surrounding. For strictly isothermal gas perfectly coupled to dust
+(constant $c_s^2$ and $\tstop=0$) we have 
+$D\seff/ D t = D\tepsilon/D t=0$. In this case the effective 
+entropy is exactly conserved because the dust-fraction is `frozen in' the
+flow.
 
@@
 
-We usually consider gas-dominated disks with $\epsilon_0 \ll 1$.  
+We usually consider gas-dominated disks with $\epsilon_0 < 1$.  
 Then $\rhog(r,z)$ is effectively Gaussian, as in the 
@@
     \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\left\{\right\} }
-  \left. -\frac{q}{r} \left(1+\epsilon\right)\frac{\p\ln{\rhog}}{\p z} \right\} 
+  \left. -\frac{q}{r} \left(1+\epsilon\right)\frac{\p\ln{\rhog}}{\p z}
+  \right\}. 
 \end{align}
@@
 \end{align*}
-However, for thick dust layers such that $H_\epsilon \gg \Hg$, 
+However, for well-mixed dust layers such that $H_\epsilon \gg \Hg$, 
 $\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.

### 2016-11-11 summary.tex
notation change for dg vel diff

@@
 \Omega$ (see Appendix \S\ref{spurious_epi}). This 
-results from neglecting terms of $O(|\Delta\bm{v}|^2)$ when converting
+results from neglecting terms of $O(|\widetilde{\bm{v}}|^2)$ when converting
 the full two-fluid equations to the one-fluid formulation. The full
@@
 %model \cite[e.g.][]{laibe14}. It 
- includes an evolutionary equation for $\Delta\bm{v}$, which has no
+ includes an evolutionary equation for $\widetilde{\bm{v}}$, which has no
  hydrodynamic counter-part \citep{laibe14}. However, it is
@@
 %\bm{a} =  
--\frac{1}{\rho}\nabla\cdot\left(\frac{\rhod\rhog}{\rho}\Delta\bm{v}\otimes\Delta\bm{v}\right) 
+-\frac{1}{\rho}\nabla\cdot\left(\frac{\rhod\rhog}{\rho}\widetilde{\bm{v}}\otimes\widetilde{\bm{v}}\right) 
 \equiv  \frac{1}{\rho}\nabla\cdot \bm{T} 
@@
 In the terminal velocity
-approximation with $\Delta\bm{v} = \tstop 
+approximation with $\widetilde{\bm{v}} = \tstop 
 \nabla P/\rhog$, we have 

### 2016-11-12 model.tex
notation change for dg vel diff
home/work sync

@@
   \rhog\left.\frac{\p \bm{v}_\mathrm{g}}{\p t}\right|_\mathrm{drag}=
-  - \frac{\rhog\rhod}{\rho}\frac{\Delta\bm{v}}{\tstop}. 
+  - \frac{\rhog\rhod}{\rho}\frac{\widetilde{\bm{v}}}{\tstop}. 
 \end{align}
@@
 \begin{align}
-  \Delta\bm{v} \equiv\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} 
+  \widetilde{\bm{v}} \equiv\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} 
 \end{align}
@@
 differs slightly from the stopping time $\tau_\mathrm{s} =
-\tstop\rhog/\rho$ used in some studies
+\tstop\rho/\rhog$ used in some studies
 \citep[e.g.][]{youdin05a,jacquet11}.
 
-In general $\Delta\bm{v}$ obeys a complicated evolutionary equation
+In general $\widetilde{\bm{v}}$ obeys a complicated evolutionary equation
 \citep[see, e.g.][]{youdin05a}. However, for tightly-coupled dust
@@
 \begin{align}
-  \Delta\bm{v} = \frac{\nabla
+  \widetilde{\bm{v}} = \frac{\nabla
     P}{\rhog}\tstop 
 \end{align}
-\citep{youdin05a,jacquet11}. This equation implies that dust particles
+\citep{jacquet11}. This equation implies that dust particles
 drift towards pressure maxima.
@@
 
-In Eq. \ref{dusteq}, 
+Eq. \ref{dusteq} results from mass conservation of dust, where 
 \begin{align}
@@
 
-The gas component of the mixture has a specific entropy 
+The specific entropy of an ideal gas is 
 $S = C_p\ln{\left(P^{1/\gamma}/\rhog\right)}$, where $C_p$ is the heat 
-capacity at constant pressure. However, this is not a useful quantity
-for (locally) isothermal gas.
+capacity at constant pressure.
 
-Nevertheless, since we have shown that an isothermal dusty gas
-effectively has $\gamma=1$ we can define an effective (dimensionless)  
+Since we have shown that a locally isothermal dusty gas effectively has $\gamma=1$ we can define an effective (dimensionless)   
 entropy for the \emph{mixture} as 
@@
 pulsations by \cite{cox67} and lecture notes by \cite{samadi15} and 
-J. Christensen-Dalsgaard\footnote{\url{http://astro.phys.au.dk/$\sim$jcd/oscilnotes/}}.
+J. Christensen-Dalsgaard\footnote{\url{http://astro.phys.au.dk/$\sim$jcd/oscilnotes/print-chap-full.pdf}}.
 
@@
 
-\subsubsection{Physical picture of dust-gas drag instabilities}
+\subsubsection{Physical requirement of dust-gas drag instabilities}
 

### 2016-11-12 linear.tex
minor structural change and edits

@@
 
-\subsection{Numerical examples}
+%\subsection{Numerical examples}
 
@@
 
-\subsection{Comparison to the convective overstability}
-%These slowly-growing modes have $|\omega|\simeq \kappa$, so they correspond to
-%overstable epicycles. 
-The streaming
-instability bears some similarity  to the `convective 
-overstability' in dust-free, non-isothermal disks 
-\citep{klahr14,lyra14,latter16}. In that case, the combination of an 
-unstable radial buoyancy frequency ($N_r^2 < 0 $) and
-finite thermal relaxation time leads to growing epicyclics. %with
-%growth rate $\propto  -N_r^2$
-
-Indeed, the effective radial buoyancy frequency in our dusty disk 
-with negligble background density gradient is unstable, 
-\begin{align}
-  N_r^2 \equiv -\frac{1}{\rho}\frac{\p P}{\p r}\frac{\p S_\mathrm{eff}}{\p r} =
-  -\frac{\rho}{P}F_r^2<0.  
-\end{align}
-Convective overstability cannot exist in our locally isothermal disks,
-which have zero cooling times. On the other hand, dust-gas drag 
-is equivalent to heating/cooling in our framework.
-
-One might then think of the streaming instability as a dusty relative
-to the convective  overstability. Explicit solutions in the small
-$\tstop$ limit show that streaming instability growth rates are
-also $\propto |N_r^2|$  \citep{jacquet11}, as is for the convective
-overstability \citep{lyra14,latter16}. Furthermore, both instabilities require
-finite effective cooling times, and both require $k_z\neq0$. 
-%{\bf note: si osc freq not too small compared to kappa}
-
-However, the convective overstability is most unstable at $k_x=0$
-\citep{lyra14}  whereas the streaming instability growth rate is maximized at finite $k_x$
-\citep{youdin05a}. Nevertheless, they share the same underlying
-physical mechanism, i.e. pressure-density phase lags.
-
 %\subsection{Boundary conditions}

### 2016-11-12 model.tex
minor structural change and edits

@@
 - \int P
-  \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV.\label{int_rel}
+  \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV,\label{int_rel}
 \end{align} 
-The real integral $\mathcal{I}^2>0$ and coefficients $A,B,C,D$ can be
+%The real integral $\mathcal{I}^2>0$ 
+where
+\begin{align}
+  \mathcal{I}^2 \equiv \int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV
+\end{align}
+is the meridional kinetic energy, 
+and coefficients $A,B,C,D$ can be
 read off Eq. \ref{integral_ex}. %Note that $B=C$ from dynamical equilibrium
@@
 work done by pressure forces in the dusty gas leads to growth in
-oscillation amplitudes. To see this, we first compute the work done
-assuming periodic oscillations. We then show that if the work done is positive, then the
-oscillation amplitude in fact grows: the work done leads the
-system to `overshoot' past the amplitude of the preceeding oscillation
-cycle.
+oscillation amplitudes. To demonstrate this, we calculate the average
+work done during a periodic oscillation, then show that if the work
+done is positive, then the growth rate is positive, i.e. the
+oscillation amplitude would actually grow.
 
-%its equilibrium. {\bf should be overshoot the max displacement}
-
+Consider oscillations in the dusty gas with period $T_p$. 
 The average rate of work done is 
@@
 \citep{cox67}, 
-where $T_p$ is the oscillation period, $\upsilon=1/\rho$ is the specific
-volume of the mixture, and the spatial integral is taken over the
-total mass $M$ of the mixture and $dm$ is the mass element. 
+where $\upsilon=1/\rho$ is the specific volume of the mixture, $D/Dt$
+is the Lagragian derivative, and the 
+spatial integral is taken over the total mass $M$ of the mixture and
+$dm$ is a mass element.
+
 Noting that only products of perturbations contribute to
@@
 
-Fig. \ref{pdv_cartoon} gives a graphical demonstration that 
-if gas pressure lags behind density, then %positive work done 
-$\mathcal{W}>0$ because it leads to a clockwise path in the
-`P-V' plane. From $A$ to $B$, a fluid parcel is expanding to return to 
-equilbrium, but pressure (i.e. restoring) forces is still
-increasing. This over-compensation causes the parcel to expand
-beyond the maximum volume of the previous cycle. 
-%{\bf should be: expand beyond the max
-%  vol of the previous cycle}
-%corresponds to positive work done. 
-Similarly, from $C$ to $D$ the parcel is already contracting towards
-equilibrium, but pressure
-forces are still dropping. This allows the  
-parcel's contraction to over-shoot the maximum density attained in the
-previous cycle. 
-% {\bf
-  %should be contract beyond the min vol of prev cycle}
-The 
-overall positive work done leads to growth in the oscillation amplitude.
-
 %$\imag\left(\Delta\rho\right)=0$.
@@
 %Applying the above argument to isothermal dusty gas,
+
 For dusty gas, the work done during oscillations 
@@
 %  types of lagragian deriv, not clear which fluid to follow.}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/drag}\\
-\includegraphics[width=\linewidth]{figures/pdv}
-  \caption{Thermodynamic interpretation of overstable modes caused by
-    dust-gas drag. Top: pressure and density evolution in time. 
-    Bottom: oscillation cycle in the P-V plane. In the case shown,
-    dust-drag causes 
-    pressure to lag behind density, which results in a clockwise path in
-    the P-V plane, implying positive work done by the fluid and
-    hence would lead to growing oscillation amplitudes.  
-    \label{pdv_cartoon}
-  }
-\end{figure}
 

### 2016-11-12 paper.tex
adding kz=0 special case in response to inutsuka-san's query
general clean up
minor structural change and edits

@@
 \newcommand{\tstop}{t_\mathrm{s}}
+\newcommand{\taus}{\tau_\mathrm{s}}
 \newcommand{\w}{ \widetilde{W}}
@@
 %  isothermal dusty gas: application to protoplanetary disks}  
-\title{A thermodynamic view of dusty protoplanetary disks:
+\title{A thermodynamic view of dusty protoplanetary disks I:
 axisymmetric stability}  
@@
   particles almost follows the gas flow. 
-  This property is analogous to 
-  entropy  conservation
-  following an adiabatic fluid.  We show that an isothermal dusty gas indeed behaves like an  
+  This is analogous to entropy conservation
+  following an adiabatic gas.  
+  Motivated by this similarity, we develop a thermodynamical
+  framework to study dusty gas dynamics in protoplanetary disks. We  
+ show that an isothermal dusty gas behaves like an  
   adiabatic pure gas; and that finite dust-gas coupling may be
@@
   heating/cooling. We exploit this equivalence 
-  to generalize previous, well-known gaseous instabilities
-  to dusty accretion disks.
-%We show that strictly isothermal, non-self-gravitating, axisymmetric
-%disks perfectly coupled to dust are generally stable, however thin the
-%dust layer is. 
-We show that in the limit of perfect coupling, dust settling in
-itself cannot lead to axisymmetric instabilities, however thin the dust layer becomes.     
-However, sufficiently sharp radial gradients in the dust-to-gas ratio
-can destabilize the disk.  
-We also extend 
-the vertical shear instability (VSI) to locally isothermal, dusty 
-disks. We find that  dust-loading reduces VSI growth rates, and
-preferentially stabilize modes with small radial 
-wavelengths, as well as `surface modes' concentrated near the
-disk boundaries. 
-These results can be understood in terms of 
-buoyancy forces in an isothermal dusty disk. 
+  to show that 
+\begin{inparaenum}[1)] 
+\item 
+ the settling of perfectly coupled dust cannot cause axisymmetric 
+ instabilties, however thin the dust layer becomes; 
+\item 
+  the streaming instability neccessarily involves a pressure response
+  that lags dust density; 
+\item dust-loading generally stabilizes the vertical shear
+  instability, and can be understood as a result of dust-induced
+  buoyancy forces. 
+\end{inparaenum}  
+%growth rates, and 
+%preferentially stabilize modes with small radial 
+%wavelengths, as well as `surface modes' concentrated near the
+%disk boundaries. 
+%These results can be understood in terms of 
+%buoyancy forces in an isothermal dusty disk. 
 %We show that vertical shear arising from
 %dust settling does not lead to axisymmetric instability. 
-We also duscuss thermodynamical interpretations of 
-dust-gas drag instabilities, and show that the streaming
-instability in protoplanetary disks occur because drag causes 
-a phase lag between the gas pressure  
-and dust density, much like pulsational instabilities 
-in stars.  
+We also discuss future applications of the dusty/adiabatic gas
+equivalence in finding dusty analogs of other hydrodynamic
+instabilities, as well as a way to simulate dusty protoplanetary disks
+using purely hydrodynamic codes with minimal effort.
+
 \end{abstract}
@@
 
-This paper is organized as follows. In \S\ref{setup} we list the
+This paper is organized as follows. We first explain in  
+\S\ref{grow_osc} how a pressure-density phase lag in fluids can lead to
+instability. This motives us to find an analogy between dusty gas and
+pure hydrodynamics, as it would allow us to understand dust dynamics
+within the framework of standard hydrodynamics.
+
+We begin our formal discussion in in \S\ref{setup} by listing the 
 one-fluid equations describing tightly-coupled dusty gas. We show that for a
@@
 
+\section{Growing oscillations by doing work}\label{grow_osc}
+A general result in fluid dynamics is that work is done whenever the
+pressure and density evolution of the fluid are not in phase. If 
+the average work done is positive, then oscillation amplitudes 
+grow: the work done allows the system to `overshoot' beyond
+the amplitude of preceeding oscillations.
+
+%To see this, we first compute the work done
+%assuming periodic oscillations. We then show that if the work done is positive, then the
+%oscillation amplitude in fact grows: the work done leads the
+%system to `overshoot' past the amplitude of the preceeding oscillation
+%cycle. 
+%its equilibrium. {\bf should be overshoot the max displacement}
+
+Fig. \ref{pdv_cartoon} gives a graphical demonstration that 
+if pressure lags behind density, then positive is work done 
+because it leads to a clockwise path in the 
+`P-V' plane. The annotations consider the particular case where the
+phase lag arises because the fluid has two components (gas and dust,
+see below) but the following description is general.
+
+From $A$ to $B$, a fluid parcel is expanding to return to 
+equilbrium, but pressure (i.e. restoring) forces is still 
+increasing. This over-compensation causes the parcel to expand
+beyond the maximum volume of the previous cycle. 
+Similarly, from $C$ to $D$ the parcel is already contracting towards
+equilibrium, but pressure 
+forces are still dropping. This allows the  
+parcel's contraction to over-shoot the maximum density attained in the
+previous cycle. 
+The overall positive work done leads to growth in the oscillation amplitude. 
+This effect is similar to being pushed downwards on a swing on the way
+down.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/drag}\\
+\includegraphics[width=\linewidth]{figures/pdv}
+  \caption{Thermodynamic interpretation of growing oscillations in a
+    fluid. Top: pressure and density evolution in time. 
+    Bottom: oscillation cycle in the P-V plane. In the case shown, 
+    dust-drag causes 
+    pressure to lag behind density, which results in a clockwise path in
+    the P-V plane, implying positive work done by the fluid and
+    hence would lead to growing oscillation amplitudes.  
+    \label{pdv_cartoon}
+  }
+\end{figure}
+
+There are several situations where the pressure and density of a fluid
+are not in phase. The obvious case is if the fluid is subject to
+external heating/cooling. For example, in strongly irradiated
+protoplanetary disks the
+disk temperature is time-independent but varies with radius
+\citep{chiang97}. Then the pressure of a fluid parcel varies with 
+position as well. There is no reason to expect $P$ and $\rho$ to
+be in phase. In fact, we will show that this is responsible for 
+the vertical shear instability \citep{lin15}.
+
+%temperature of a fluid parcel changes with position
+
+Another possibility, as annotated in Fig. \ref{pdv_cartoon}, is a dusty gas. The
+relevant density here is the total density $\rho = \rhod + \rhog$, 
+but the fluid pressure $P$ is due to gas only. If dust were 
+perfectly-coupled to the gas, then $P$ and $\rho$ would be in phase, 
+and there is no work done. However, for finite dust-gas drag, $\rhod$ and
+$\rhog$ are not neccessarily in phase. If this causes $P$ to lag
+$\rho$, then the positive work done would lead to growing
+oscillations. Indeed, we show that this is true for the streaming
+instability.
+
+In order to apply this thermodynamic interpretation of dust-gas drag
+instabilities, we need to develop an analogy between dusty-gas
+and pure gas dynamics. We show this is possible in the limit of
+strong drag and a fixed gas equation of state.
+
 \input{model}
@@
 \input{summary}
+
+%{\bf acknowledgements. s. inutsuka for asking about kz=0 limit}
 

### 2016-11-12 summary.tex
minor structural change and edits

@@
 %gap edges. thin dust rings.
+
+\subsubsection{Convective overstability}
+The `convective overstability' (ConO) operates in dust-free, non-adiabatic disks 
+when there is an unstable radial buoyancy frequency ($N_r^2 < 0 $)
+combined with finite thermal relaxation time, $t_\mathrm{cool}\sim
+\OmK^{-1}$. This lead to growing epicycles
+\citep{klahr14,lyra14,latter16}.  
+%with 
+%growth rate $\propto  -N_r^2$ 
+% Indeed, the effective radial buoyancy frequency in our dusty disk 
+% with negligble background density gradient is unstable, 
+% \begin{align}
+%   N_r^2 \equiv -\frac{1}{\rho}\frac{\p P}{\p r}\frac{\p S_\mathrm{eff}}{\p r} =
+%   -\frac{\rho}{P}F_r^2<0.  
+% \end{align}
+
+The conditions required for ConO might be realized in a locally
+isothermal dusty disk. Since $\p_rP < 0$ in typical disk models, 
+if the dust-induced radial entropy gradient is negative, then
+$N_r^2  =  F_r \p_r\seff < 0$. In 
+fact, typical disk models used for the streaming
+instability have negligible background radial density gradient, then  
+$N_r^2= -\rho F_r^2/P$.
+
+Locally isothermal disks have $\tcool=0$, so the pure gas ConO cannot
+exist. However, we have shown that dust-gas drag provides an effective
+energy source/sink to the mixture. It would be
+interesting to explore whether or not dust-gas drag can provide the
+required `cooling' to enable a dusty version of the convective
+overstability.
+
+%Furthermore, 
+%the effective radial buoyancy frequency in our dusty disk  
+%with negligble background density gradient is unstable, 
+%\begin{align}
+%   N_r^2 \equiv -\frac{1}{\rho}\frac{\p P}{\p r}\frac{\p S_\mathrm{eff}}{\p r} =
+%   -\frac{\rho}{P}F_r^2<0.  
+% \end{align}
+
+% One might then think of the streaming instability as a dusty relative
+% to the convective  overstability. Explicit solutions in the small
+% $\tstop$ limit show that streaming instability growth rates are
+% also $\propto |N_r^2|$  \citep{jacquet11}, as is for the convective
+% overstability \citep{lyra14,latter16}. Furthermore, both instabilities require
+% finite effective cooling times, and both require $k_z\neq0$. 
+% %{\bf note: si osc freq not too small compared to kappa}
+
+% However, the convective overstability is most unstable at $k_x=0$
+% \citep{lyra14}  whereas the streaming instability growth rate is maximized at finite $k_x$
+% \citep{youdin05a}.
+
+\subsubsection{Zombie vortices}
+
+The `zombie vortex instability' (ZVI) is a non-axisymmetric,
+non-linear (or finite-amplitude) instability in gas disks
+\citep{marcus15}. It operates by exciting `critical layers' which role
+up into vortices, exciting further critical layers. The result is an
+array of vortices. These critical
+layers are locations where the intrinsic wave frequency of a
+disturbance matches the local vertical buoyancy frequency
+\citep{marcus13}.
+
+Since buoyancy is critical for the ZVI, it requires an almost
+purely adiabatic gas ($\tcool\to\infty$). This limits its
+applicability to the inner most parts of typical protoplanetary disks
+\citep{lesur16}.
+
+However, we have show that dust-loading an isothermal disk also
+induces an effective buoyancy. This effect is maximized when dust is
+perfectly coupled to the gas. It is thus natural to ask whether or not
+the required adiabatic conditions for ZVI can be realized through
+dust-loading, and thus produce a `dusty ZVI'.
 

### 2016-11-13 linear.tex
various minor edits

@@
 growth rates have been 
-calculated analytically from full two-fluid disk models and measured 
-from direct numerical particle-gas simulations 
+calculated analytically from full two-fluid disk models and confirmed
+with particle-gas numerical simulations 
 \citep[namely][]{youdin07b,bai10b}. These authors use normalized

### 2016-11-13 model.tex
various minor edits

@@
 
-%In the limit of strong dust-gas drag (stopping time $\tstop\to 0$) 
-In the limit of strong dust-gas drag ($\tstop\to0$), \cite{laibe14}
-give a single fluid description of the mixture, to
-$O\left(\tstop\right)$, as:   
+In the limit of strong dust-gas drag, $\tstop\to0$, the mixture obeys 
+the first-order (in $\tstop$) single-fluid equations: 
 \begin{align} 
@@
   &\frac{D T}{D t} = - \left(\gamma-1\right)T\nabla\cdot\bm{v} +
-  \mathcal{H}_\mathrm{drag} + \mathcal{H}_\mathrm{eff}  - \Lambda, \label{tempeq} 
+  \mathcal{H}_\mathrm{drag} + \mathcal{H}_\mathrm{eff}  - \Lambda  \label{tempeq} 
 \end{align}
-where $D/Dt \equiv \p_t + \bm{v}\cdot\nabla$ is the Lagragian
+\citep{laibe14}, where $D/Dt \equiv \p_t + \bm{v}\cdot\nabla$ is the Lagragian
 derivative \emph{following the mixture} with center-of-mass velocity  
@@
 \begin{align}\label{eos}
-  P = c_s^2(r,z)\rhog = c_s^{2}(1 - \tepsilon)\rho,   
+  P = c_s^2(r,z)\rhog = c_s^{2}\left(1 - \tepsilon\right)\rho,   
 \end{align}
 where $c_s(r,z)= \sqrt{\mathcal{R}T_\mathrm{ref}/\mu}$ is a prescribed
-sound-speed profile fixed in time.
-
-In practice we consider vertically 
+sound-speed profile fixed in time. In practice we consider vertically 
 isothermal disks with \begin{align}\label{power_temp}
@@
 
-If we define the effective sound-speed $c_{s,\mathrm{eff}} = c_s\sqrt{\left(1 -
-    \tepsilon\right)}$, then the equation of state has the standard
-form $P=c_{s,\mathrm{eff}}^2\rho$. Since the dust-fraction $\tepsilon$
-typically decrease away from the midplane, we expect vertically
-isothermal dusty disks to in fact behave as if it
-had a temperture \emph{increasing} away from $z=0$.
+Notice Eq. \ref{eos} resembles an ideal gas equation of state but with  
+effective temperature $T_\mathrm{eff} = 
+T_\mathrm{ref}\left(1-\tepsilon\right)$. %sound-speed $c_{s,\mathrm{eff}} = c_s\sqrt{\left(1 -
+This effective mixture  
+temperature is lowered with dust-loading. 
+%   \tepsilon\right)}$. 
+%then the equation of state has the standard
+%form $P=c_{s,\mathrm{eff}}^2\rho$. 
+Since $\tepsilon$ typically decrease away from the midplane, we expect
+vertically isothermal dusty disks to in fact behave as if it had a
+temperture \emph{increasing} away from $z=0$.
 
@@
 
-In this section we review the basic properties of the 
-single-fluid mixture using Eqs.\ref{masseq}--\ref{momeq} with 
+In this section we discuss the basic properties of the 
+dusty-gas mixture using Eqs.\ref{masseq}--\ref{momeq} with 
 Eq. \ref{eff_energy} in place of the dust continuity equation. Our
@@
     \tepsilon\right)}$ for the entropy in the standard
-expression for said criteria \citep[e.g.]{tassoul78}.
+expression for said criteria \citep{tassoul78}.
 
@@
 isothermal disks perfectly coupled with dust are stable against
-axisymmetric perturbations.
+axisymmetric perturbations. This result holds for arbitrarily thin
+dust layers.
 
@@
 oscillation amplitudes. To demonstrate this, we calculate the average
-work done during a periodic oscillation, then show that if the work
-done is positive, then the growth rate is positive, i.e. the
+work done during a periodic oscillation, and show that if the work
+done is positive, then the 
 oscillation amplitude would actually grow.
@@
  the Lagragian perturbation $\Delta$ of a variable $f$ is 
-$\Delta f = \delta f + \bm{\xi}\cdot\nabla f$; where $\bm{\xi}$ is the
+$\Delta f = \delta f + \bm{\xi}\cdot\nabla f$ and $\bm{\xi}$ is the
 Lagragian displacement ($    \xi_{x,z} =  \ii \dd v_{x,z}/\sigma$).  
@@
 Eq. \ref{pdv}. Eq. \ref{growth_work} states that if the average work
-done is positive during a periodic oscillation, $\mathcal{W}>0$, then
+done is positive during an oscillation, $\mathcal{W}>0$, then 
 then its amplitude would actually grow ($s>0$).  %and not be periodic
 
-Positive work done is done by a parcel if $-\omega \imag\left(\Delta  
+Positive work is done by a fluid parcel if $-\omega \imag\left(\Delta  
 P\Delta\rho^*\right)>0$. Without loss of generality, take $\omega>0$
@@
 drift between gas and dust causes a phase difference between the two
-components.
+components, and hence between pressure and total density.
 
@@
 \emph{why} drag forces causes gas pressure to lag behind the dust
-density, but merely shows 
+density, but shows 
 that this \emph{must} be the case for any growing modes associated the
@@
 $\mathcal{C}$ (see Appendix \ref{lin_dust}), we might generally expect
-the dusty disks to  support a range of stable and unstable modes, with the
+ dusty disks to  support a range of stable and unstable modes, with the
 latter being associated with pressure-density lag.

### 2016-11-13 paper.tex
various minor edits

@@
 axisymmetric stability}  
-\author{Min-Kai Lin\altaffilmark{1} and Andrew N. Youdin}
-\affil{Department of Astronomy and Steward Observatory, \\
+\author{Min-Kai Lin\altaffilmark{1,2,3}}
+\author{Andrew N. Youdin\altaffilmark{1}}
+%\affil{Department of Astronomy and Steward Observatory, 
+%University of
+ % Arizona, Tucson, AZ 85721, USA\\
+%Institute of Astronomy and Astrophysics, Academia Sinica, 
+%Taipei 10617, Taiwan}
+\altaffiltext{1}{Department of Astronomy and Steward Observatory, 
 University of
-  Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA}
-\altaffiltext{1}{Steward Theory Fellow}
-\email{minkailin@email.arizona.edu}
+Arizona, Tucson, AZ 85721, USA}
+\altaffiltext{2}{Steward Theory Fellow}
+\altaffiltext{3}{Institute of Astronomy and Astrophysics, Academia Sinica, 
+Taipei 10617, Taiwan
+}
+\email{mklin@asiaa.sinica.edu.tw}
 
@@
 \item dust-loading generally stabilizes the vertical shear
-  instability, and can be understood as a result of dust-induced
+  instability, which can be understood as a effect of dust-induced
   buoyancy forces. 
@@
 equivalence in finding dusty analogs of other hydrodynamic
-instabilities, as well as a way to simulate dusty protoplanetary disks
-using purely hydrodynamic codes with minimal effort.
+phenomena (e.g. Rossby wave instability, convective overstability, and 
+zombie vortices), and to simulate dusty protoplanetary disks
+using purely hydrodynamic codes.
 
@@
 instability. This motives us to find an analogy between dusty gas and
-pure hydrodynamics, as it would allow us to understand dust dynamics
-within the framework of standard hydrodynamics.
+pure hydrodynamics, as it allows us to understand dusty-gas dynamics
+within a thermodynamic framework.
 
-We begin our formal discussion in in \S\ref{setup} by listing the 
+We begin our formal discussion in \S\ref{setup} by listing the 
 one-fluid equations describing tightly-coupled dusty gas. We show that for a
@@
 fact the usual hydrodynamic equations in disguise. 
-We describe equilibrium solutions in \S\ref{eqm} which are used in
+We describe equilibrium solutions in \S\ref{eqm} that are used in
 explicit calculations. In \S\ref{limits} we discuss 
 general stability properties of dusty disks. We then move onto
-explicit stablity analyses. In  
+numerical stability analyses. In  
 \S\ref{linear_problem} we list the linearized equations, and solve 
 them for the SI in \S\ref{si}, and in 
- \S\ref{results} study dusty VSI. We 
+ \S\ref{results} to study the dusty VSI. We 
 discuss future applications of the dusty/adiabatic gas equivalence in 
@@
 The overall positive work done leads to growth in the oscillation amplitude. 
-This effect is similar to being pushed downwards on a swing on the way
-down.
+This effect is similar to being pushed downwards on a swing on
+descent.
 

### 2016-11-13 summary.tex
various minor edits

@@
 \Omega$ (see Appendix \S\ref{spurious_epi}). This 
-results from neglecting terms of $O(|\widetilde{\bm{v}}|^2)$ when converting
+results from neglecting terms of $O(|\widetilde{\bm{v}}|^2)$ after converting
 the full two-fluid equations to the one-fluid formulation. The full
@@
 $N_r^2  =  F_r \p_r\seff < 0$. In 
-fact, typical disk models used for the streaming
-instability have negligible background radial density gradient, then  
-$N_r^2= -\rho F_r^2/P$.
+fact, the typical disk models used to study the streaming
+instability have zero background radial density gradient, in which
+case  $N_r^2= -\rho F_r^2/P$.
 
@@
 Since buoyancy is critical for the ZVI, it requires an almost
-purely adiabatic gas ($\tcool\to\infty$). This limits its
+purely adiabatic gas ($\tcool\OmK\gg 1$). This limits its
 applicability to the inner most parts of typical protoplanetary disks
@@
 $t_\mathrm{s0}=0.02\OmK^{-1}$. The gas disk parameters are $(p,q,
-\hgas) =(-1.5, -1,0.05)$.  We set $k = \pi/H_\mathrm{d}$.
+\hgas) =(-1.5, -1,0.05)$.  We set $k_x = \pi/H_\mathrm{d}$.
 

### 2016-11-13 model.tex
tiny edits

@@
 %We comment that
+
+An effective energy equation can also be derived for other 
+fixed gas equations of state, such as polytropic gas. We discuss this
+in \S\ref{gen_poly}.
 
@@
 
+The physical origin for this vertical buoyancy, which exists even in vertically isothermal
+disk, lies in the fact that coupling gas to dust particles increases
+the gas' inertia, but pressure (i.e. restoring) forces are
+unchanged. The increased weight of the fluid resists vertical
+oscillations and hence there is an associated buoyancy force. However,
+if $\tstop\neq0$ so the gas-dust coupling is imperfect, then gas is no
+longer `weighed down' by the dust, and can slip past it. This
+diminishes the dust-induced buoyancy.
+
  % N_z^2 
@@
 \begin{align}
-  \frac{\p\ln{\epsilon}}{\p\ln{r}} > S_+ \quad \mathrm{or} \quad 
-  \frac{\p\ln{\epsilon}}{\p\ln{r}} < S_-,
+  \frac{\p\ln{\epsilon}}{\p\ln{r}} > \chi_+ \quad \mathrm{or} \quad 
+  \frac{\p\ln{\epsilon}}{\p\ln{r}} < \chi_-,
 \end{align}
@@
 \begin{align}\label{spm}
-S_\pm = \frac{1}{2}\frac{H_\mathrm{g}^2}{\Htilde^2} 
+\chi_\pm = \frac{1}{2}\frac{H_\mathrm{g}^2}{\Htilde^2} 
   \left[

### 2016-11-13 paper.tex
tiny edits

@@
   heating/cooling. We exploit this equivalence 
-  to show that 
+  to deduce that   
 \begin{inparaenum}[1)] 
@@
 \item dust-loading generally stabilizes the vertical shear
-  instability, which can be understood as a effect of dust-induced
-  buoyancy forces. 
+  instability, because dust-gas coupling introduces buoyancy. 
 \end{inparaenum}  
@@
 We also discuss future applications of the dusty/adiabatic gas
-equivalence in finding dusty analogs of other hydrodynamic
+correspondence in finding dusty analogs of other hydrodynamic
 phenomena (e.g. Rossby wave instability, convective overstability, and 
@@
 dusty gas and standard hydrodynamics. This is possible in certain limits of astrophysical
-interest, allowing us to interpret dusty phenomenon through standard
+interest, allowing us to interpret dusty phenomena through standard
 hydrodynamics.
@@
 
-Under these approximations, we show that the one-fluid equations of 
+Under these approximations, we show that the equations of 
 isothermal dusty gas dynamics are equivalent to that of adiabatic, pure gas
@@
 entrained in the gas, which is is analogous to the advection
-of entropy following an adiabatic fluid. We will formally relate the 
-effective entropy of isothermal dusty gas to its dust content. Finite  
-dust-gas drag act as effective heating/cooling: the relative
-motion between dust and gas causes a fluid parcel to exchange dust
+of entropy following an adiabatic fluid. We will find that the   
+effective entropy of isothermal dusty gas scales with its gas
+fraction. Finite dust-gas drag act as effective heating/cooling: the relative
+drift between dust and gas causes a fluid parcel to exchange dust
 particles with its surroundings; an effect similar to heat exchange
@@
 hydrodynamics, we can apply well-known methods to generalize previous stability 
-analyses for pure gas disks to the dusty case. We explicitly demonstrate this for
-perfectly coupled dust in strictly and locally isothermal disks. We 
-show that strictly isothermal protoplanetary disks are generally
+analyses for pure gas disks to the dusty case. %We explicitly demonstrate this for
+%perfectly coupled dust in strictly and locally isothermal disks. 
+We show that strictly isothermal protoplanetary disks with
+perfectly-coupled dust are generally
 stable to axisymmetric perturbations, regardless of how  thin the dust
@@
 
-This paper is organized as follows. We first explain in  
-\S\ref{grow_osc} how a pressure-density phase lag in fluids can lead to
-instability. This motives us to find an analogy between dusty gas and
+This paper is organized as follows. We first
+explain in \S\ref{grow_osc} a general result of fluid dynamics that
+underpins the rest of our study. We describe how 
+a pressure-density phase lag in fluids can lead to 
+instability.
+
+This motives us to find an analogy between dusty gas and
 pure hydrodynamics, as it allows us to understand dusty-gas dynamics

### 2016-11-14 paper.tex
home/work sync

@@
 zombie vortices), and to simulate dusty protoplanetary disks
-using purely hydrodynamic codes.
+using purely hydrodynamic codes. %We briefly describe extending this
+%equivalence principle to polytropic and adiabatic gas.
 
@@
 a pressure-density phase lag in fluids can lead to 
-instability.
-
-This motives us to find an analogy between dusty gas and
-pure hydrodynamics, as it allows us to understand dusty-gas dynamics
-within a thermodynamic framework.
+instability. Thus finding an analogy between dusty gas and
+pure hydrodynamics allow us to interpret dusty-gas dynamics
+within a thermodynamics framework.
 
 We begin our formal discussion in \S\ref{setup} by listing the 
-one-fluid equations describing tightly-coupled dusty gas. We show that for a
+basic equations describing tightly-coupled dusty gas. We show that for a
 fixed equation of state, these are in 

### 2016-11-18 model.tex
home/work sync
edits on the plane

@@
 $|z|\ll r$. We shall adopt this approximate potential in order to
-obtain analytic expressions for the disk equilibria.
+obtain explicit expressions for disk equilibria.
 
@@
 solutions in \S\ref{steady_state}. We assume the disk is radially
-smooth so that $\p_r\sim 1/r$, and the dust-to-gas ratio
-$\epsilon\ll1$. This gives
+smooth so that $\p_r\sim 1/r$. % and the dust-to-gas ratio
+% $\epsilon\ll1$. 
+This gives
 
@@
 such that $\delta^2\ll \epsilon$. Otherwise, vertical shear is
-associated with $\p_rc_s^2$.
+associated with the radial temperature gradient. %However, axisymmtric
+%\emph{instability} always caused by $\p_rc_s^2$.
 
@@
 The physical origin for this vertical buoyancy, which exists even in vertically isothermal
-disk, lies in the fact that coupling gas to dust particles increases
-the gas' inertia, but pressure (i.e. restoring) forces are
+disks, is that coupling gas to dust particles increases
+the fluid's inertia, but pressure (i.e. restoring) forces are
 unchanged. The increased weight of the fluid resists vertical
@@
 if $\tstop\neq0$ so the gas-dust coupling is imperfect, then gas is no
-longer `weighed down' by the dust, and can slip past it. This
-diminishes the dust-induced buoyancy.
+longer `weighed down' by the dust, and can slip past it. Thus finite
+drag diminishes the dust-induced buoyancy.
 
@@
 discussion is based on the corresponding axisymmetric linearized
-equations. A variable $f$ is
+equations. A variable $X$ is
 subject to Eulerian perturbations of the form 
 \begin{align}
- \real\left[ \delta f(r,z)\exp{\left(-\ii\sigma t\right)}\right], 
+ \real\left[ \delta X(r,z)\exp{\left(-\ii\sigma t\right)}\right], 
 \end{align}
@@
 P\right)<0$.  This corresponds to pressure perturbations \emph{lagging}
-behind that in density (or displacements), since the phase of
-the assumed perturbations increase in time.
-
+behind that in density (or displacements).%, since the phase of
+%the assumed perturbations increase in time.  
 %$\imag\left(\Delta\rho\right)=0$.
@@
 
-For dusty gas, the work done during oscillations 
- is attributed to finite dust-gas drag. The relative
+The discussion above applies to any single fluid with a pressure and
+density. In the case of interest --- dusty gas --- the work done 
+is attributed to finite dust-gas drag. 
+%For dusty gas, the work done during oscillations 
+ %is attributed to finite dust-gas drag. 
+The relative 
 drift between gas and dust causes a phase difference between the two

### 2016-11-18 paper.tex
edits on the plane

@@
 axisymmetric stability}
+ % \title{Dust-free modeling of dusty protoplanetary disks I:
+%axisymmetric stability}
 \author{Min-Kai Lin\altaffilmark{1,2,3}}
@@
 dynamics. This is because tightly-coupled dust particles are almost
-entrained in the gas, which is is analogous to the advection
+entrained in the gas, which is analogous to the advection
 of entropy following an adiabatic fluid. We will find that the   
@@
 stable to axisymmetric perturbations, regardless of how  thin the dust
-layer is. We put forward a thermodynamical thermodynamical interpretation of
+layer is. We put forward a thermodynamical interpretation of
 dust-gas drag instabilities, and apply it to the
@@
 protoplanetary disks the
-disk temperature is time-independent but varies with radius
-\citep{chiang97}. Then the pressure of a fluid parcel varies with 
+disk temperature  $T(r)$ is time-independent but varies with radius
+\citep{chiang97}. Since $P\propto \rho T$ for an ideal gas, the pressure of a fluid parcel varies with 
 position as well. There is no reason to expect $P$ and $\rho$ to

### 2016-11-18 summary.tex
tiny edits
edits on the plane

@@
 
-\subsection{Generalization to locally polytropic disks}
+\subsection{Generalization to locally polytropic disks}\label{gen_poly}
 
@@
 where $K$ is a prescribed function and $\Gamma$ is the constant
-adiabatic index. Then eliminating $\tepsilon$ from the dust equation
+polytropic index. Then eliminating $\tepsilon$ from the dust equation
 \ref{dusteq} gives 
@@
 gas dynamics codes can be applied to simulate locally
-isothermal/polytropic dusty gas with minimal modification. In fact, no
-modification is needed  
-for strictly isothermal or polytropic gas perfectly-coupled to dust:  
-the standard adiabatic energy equation substitutes for the dust continuity
+isothermal/polytropic dusty gas. One need not add any new 
+equations to evolve the dust component: 
+%In fact, no
+%modification is needed  
+%for strictly isothermal or polytropic gas perfectly-coupled to dust:  
+ the standard energy equation substitutes for the dust continuity
 equation. When $c_s$ or $K$ is not constant and/or $\tstop\neq0$ one

### 2016-11-25 model.tex
deleted second energy equation in sec 3.2

@@
 
-Eq. \ref{eff_energy} can be written in conservative form,
-\begin{align*}
-  \frac{\p P}{ \p t} + \nabla\cdot\left\{P\left[\bm{v} -
-      \tstop\left(c_s^2 - \frac{P}{\rho}\right)\nabla\ln{P}\right]
-    \right\}\\
-  = \left[P\bm{v} - \tstop\left(c_s^2 - \frac{P}{\rho}\right)\nabla
-    P\right]\cdot\nabla\ln{c_s^2}. 
-\end{align*} 
-We may thus re-interpret $P$ as the mixture's energy density, but the
-energy flux has an additional contribution from the pressure
-gradient and dust-gas friction. The term on the right-hand-side, owing
-to the imposed temperature profile, can be interpreted as an external
-heat source.
+% Eq. \ref{eff_energy} can be written in conservative form,
+% \begin{align*}
+%   \frac{\p P}{ \p t} + \nabla\cdot\left\{P\left[\bm{v} -
+%       \tstop\left(c_s^2 - \frac{P}{\rho}\right)\nabla\ln{P}\right]
+%     \right\}\\
+%   = \left[P\bm{v} - \tstop\left(c_s^2 - \frac{P}{\rho}\right)\nabla
+%     P\right]\cdot\nabla\ln{c_s^2}. 
+% \end{align*} 
+% We may thus re-interpret $P$ as the mixture's energy density, but the
+% energy flux has an additional contribution from the pressure
+% gradient and dust-gas friction. The term on the right-hand-side, owing
+% to the imposed temperature profile, can be interpreted as an external
+% heat source.
 

### 2017-01-04 summary.tex
note to self in 9.4.1 on possibility of dusty gravito-turbulence, because dust-gas drift is like energy sink

@@
 study.
+
+Another potential connection to previous results for gas disks is  
+gravito-turbulence. Cooling, self-gravitating gaseous disks 
+sustain a turbulent state where shock heating due to gravitational 
+instabilities are balanced by thermal losses
+\citep{gammie01}. Now, dust-gas drift behaves like thermal diffusion
+or cooling (Eq. \ref{eff_energy}), similar to how radiative cooling 
+appears in models of gravito-turbulent disks. Thus it may be
+conceivable that a massive dusty disk admit `dusty
+gravito-turbulence'.
 

### 2017-01-05 linear.tex
edits on the plane
new section 7 on secular GI. rest of paper might need adjusting. mostly doing non-SG, but one sec and one appendix has SG

@@
 \begin{align}
-  \delta f (r, z) = \delta f_1(r,z)\exp{(\ii k_x r)},
+  \delta X (r, z) = \delta X_1(r,z)\exp{(\ii k_x r)},
 \end{align} 
@@
 where $k_x$ is a real wavenumber such that $|k_xr|\gg 1$, and the 
-amplitude $\dd f_1(r,z)$ is 
+amplitude $\dd X_1(r,z)$ is 
 a slowly-varying function of $r$. Then 
@@
 % +  \delta\bm{F}\cdot\hat{\bm{r}}
-- W F_r - \ii k_x Q,\label{lin_xmom}\\
+- W F_r - \ii k_x Q - \ii k_x\dd\psi,\label{lin_xmom}\\
   \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
@@
   -\ii\sigma\dd v_z &= - W F_z - \left[Q^\prime + Q
-    \left(\ln{\rho}\right)^\prime\right]  %\delta\bm{F}\cdot\hat{\bm{z}} 
+    \left(\ln{\rho}\right)^\prime\right] - \dd\psi^\prime  %\delta\bm{F}\cdot\hat{\bm{z}} 
 ,\label{lin_zmom}\\
@@
 %+ \dd v_z \p_z\ln{c_s^2}\right),\label{lin_energy} 
-               - \frac{\dd\mathcal{C}}{\rho},\label{lin_energy} 
+               - \frac{\dd\mathcal{C}}{\rho},\label{lin_energy}\\
+\dd\psi^{\prime\prime}  &= 4\pi G \rho W + k_x^2\dd\psi, \label{lin_sg}
 \end{align}  
@@
 
-Eq. \ref{lin_mass}---\ref{lin_energy} is a set of ordinary
+Eq. \ref{lin_mass}---\ref{lin_sg} is a set of ordinary
 differntial equations in $z$. All coefficients and amplitudes are
 evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
-is retained. We now discuss solutions to these equations in an unstratified disk 
-in \S\ref{si} and a stratified disk in \S\ref{results}.
+is retained. We now discuss solutions to these equations for a
+razor-thin, self-gravitating disk in \S\ref{sgi}; and for 3D
+non-self-gravitating, unstratified and stratified disks 
+in \S\ref{si} and \S\ref{results}, respectively.
 
@@
 
+\section{Secular gravitational instability}\label{sgi}
+We first consider a razor-thin disk so that $\rho =
+\Sigma\delta(z)$. Here $\delta$ is the delta function and $\Sigma$ is
+the total surface density. Note that $\epsilon$ and 
+$\tepsilon$ now refers to the global dust-to-gas ratio and dust 
+fraction, respectively. We neglect the vertical dimension   
+in Eq. \ref{lin_mass}---\ref{lin_energy}. The background disk is
+uniform. The linearized dust-gas drag term is then
+$-\delta\mathcal{C}/\rho = \tstop\tepsilon c_s^2k_x^2 Q$. The
+thin-disk solution to Eq. \ref{lin_sg} is $\dd\psi = -2\pi G
+\Sigma W/\left|k_x\right|$
+
+These simplifications yield the dispersion relation
+\begin{align*}
+  \left(\ii\sigma - \tstop\tepsilon c_s^2k_x^2\right)\left( 2 \pi G
+    \Sigma \left|k_x\right|  - \kappa^2 + \sigma^2 \right) = \ii
+  \sigma c_s^2 k_x^2\left(1 - \tepsilon\right). 
+\end{align*}
+Searching for slowly and purely growing modes, $\sigma = \ii s$ with 
+$|s|\ll \kappa$, we find 
+\begin{align}  
+s = \frac{\tstop\tepsilon c_s^2 k_x^2 \left( 2 \pi \Sigma G
+    \left|k_x\right| - \kappa^2\right)}{\kappa^2 -2 \pi \Sigma G
+    \left|k_x\right| + c_s^2k_x^2\left(1-\tepsilon\right) }. \label{sgi_disp}
+\end{align}
+This is just the secular gravitational instability in the absence of
+turbulent dust diffusion but strong drag
+\citep[][ their Eq. 13 becomes Eq. \ref{sgi_disp} in this limit with
+a change of variables]{takahashi14}. This is gravitational instability
+mediated by dust-gas drag. A similar effect occurs in viscous
+self-gravitating gas disks \citep{gammie96,lin16}. In fact, if we
+identify $\nu \equiv \tstop c_s^2$ as a kinematic viscosity, then
+Eq. \ref{sgi_disp} is effectively 
+\citeauthor{gammie96}'s Eq. 18.
+
+This exercise shows that the one-fluid framework, futher simplified by
+the terminal velocity approximation, is sufficient to capture the SGI
+in the strong drag limit.
+
 \section{Streaming instability}\label{si}
+We now consider 3D but non-self-gravitating disks.  
 In the special case of an unstratified disk (neglecting the vertical
@@
 
+% \subsection{Eulerian interpretation}
+
+% We can also consider SI in the Eulerian sense \citep[as
+% did][]{jacquet11}. Eq. \ref{thermal_instability} imply SI requires
+% \begin{align*}
+%   \frac{\imag\left[\delta\mathcal{C}\nabla\cdot\dd\bm{v}^*\right]}{\real(\sigma)}<0. 
+% \end{align*}
+% The linearized dust diffusion or `cooling' term $\delta\mathcal{C}$
+% for this problem is given by Eq. \ref{lin_drag_si}. Thus the
+% requirement is 
+% \begin{align*}
+% \tepsilon |\bm{k}|^2 
+%   \frac{\imag\left(\ii \sigma^* QW^*\right)}{\real(\sigma)}
+% -k_x F_r\left( 1-2\tepsilon\right) 
+%  | W |^2\frac{\imag\left(\sigma^*\right)}{\real(\sigma)} < 0, 
+% \end{align*}
+% where we have used Eq. \ref{streaming_mass}. 
+% For slow growth and/or large $\left|\bm{k}\right|^2$ the second term 
+% is small compared to the first. In this case instablity is due to
+
 %\subsection{Boundary conditions}

### 2017-01-05 model.tex
new section 7 on secular GI. rest of paper might need adjusting. mostly doing non-SG, but one sec and one appendix has SG

@@
   &\frac{D\rho}{Dt} = -\rho\nabla\cdot\bm{v}, \label{masseq}\\ 
-  &\frac{D\bm{v}}{Dt} = - \nabla\Phi - \frac{1}{\rho}\nabla  P, \label{momeq}\\ 
+  &\frac{D\bm{v}}{Dt} = - \nabla\Phi_\mathrm{tot} - \frac{1}{\rho}\nabla  P, \label{momeq}\\ 
   &\frac{D\tepsilon}{Dt} = -\frac{1}{\rho} \nabla \cdot \left(\tepsilon 
@@
 
-For the disk problem the external potential $\Phi$ is  
-that for a star of mass $M_*$, 
+The total gravitational potential $\Phi_\mathrm{tot}=\Phi + \psi$,
+where $\Phi$ originates from a star of mass $M_*$, 
 \begin{align}\label{thin_disk_potential}
@@
 \end{align}
-where $G$ is the gravitational constant, and $(r,\phi, z)$ are
+and $\psi$ is the disk potential satisfying
+\begin{align}
+  \nabla^2\psi = 4 \pi G \rho.
+\end{align}
+Here, $G$ is the gravitational constant, and $(r,\phi, z)$ are 
 cylindrical co-ordinates centered on the star.  
-The second equality is the 
+The second equality in Eq. \ref{thin_disk_potential} is the 
 thin-disk approximation for the disk potential, appropriate for
 $|z|\ll r$. We shall adopt this approximate potential in order to
-obtain explicit expressions for disk equilibria.
+obtain explicit expressions for disk equilibria (\S\ref{eqm}).
+
+We include the gravitational potential of the mixture 
+$\psi$ for completeness but will neglect it in most of this
+work. Henceforth we consider non-self-gravitating disks ($\psi \equiv
+0$) unless explicitly stated otherwise.
 

### 2017-01-05 paper.tex
new section 7 on secular GI. rest of paper might need adjusting. mostly doing non-SG, but one sec and one appendix has SG

@@
 \citep{youdin05a,youdin07b,johansen07}, secular gravitational
-instabilities \citep{ward00,youdin11,michikoshi12,takahashi14} and
+instabilities \citep[SGI,][]{ward00,youdin11,michikoshi12,takahashi14} and
 Kelvin-Helmholtz instabilities
@@
 general stability properties of dusty disks. We then move onto
-numerical stability analyses. In  
-\S\ref{linear_problem} we list the linearized equations, and solve 
-them for the SI in \S\ref{si}, and in 
- \S\ref{results} to study the dusty VSI. We 
+explicit stability analyses. In  
+\S\ref{linear_problem} we list the linearized equations. We solve them 
+to recover SGI in \S\ref{sgi}, to interpret the SI in \S\ref{si}, and
+in  \S\ref{results} to extend the VSI to dusty disks. We  
 discuss future applications of the dusty/adiabatic gas equivalence in 

### 2017-01-06 linear.tex
minor edits on SGI stuff

@@
 \end{align}  
-where $^\prime \equiv \p_z$ and recall $\bm{F} \equiv -\nabla P/\rho$. 
-The linearized dust-duffusion function 
+where $^\prime \equiv \p_z$ and recall $\bm{F} \equiv -\nabla
+P/\rho$. Self-gravity has been restored temporarily for the next
+section (only). The linearized dust-duffusion function 
 $\dd\mathcal{C}$ is given in  Appendix \ref{lin_dust}. 
@@
 evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
-is retained. We now discuss solutions to these equations for a
-razor-thin, self-gravitating disk in \S\ref{sgi}; and for 3D
-non-self-gravitating, unstratified and stratified disks 
-in \S\ref{si} and \S\ref{results}, respectively.
+is retained. We next discuss solutions to these equations.  We first
+show that the above equations yield the secular
+gravitational instability and streaming instability in the strong
+drag regime in \S\ref{sgi} and \S\ref{si}, respectively. We then
+consider 3D, stratified disks in \S\ref{results} to study the dusty
+vertical shear instability.
+
+%for a
+%razor-thin, self-gravitating disk in \S\ref{sgi}; and for 3D
+%non-self-gravitating, unstratified and stratified disks 
+%in \S\ref{si} and \S\ref{results}, respectively.
 
@@
 
-\section{Secular gravitational instability}\label{sgi}
-We first consider a razor-thin disk so that $\rho =
-\Sigma\delta(z)$. Here $\delta$ is the delta function and $\Sigma$ is
-the total surface density. Note that $\epsilon$ and 
-$\tepsilon$ now refers to the global dust-to-gas ratio and dust 
-fraction, respectively. We neglect the vertical dimension   
-in Eq. \ref{lin_mass}---\ref{lin_energy}. The background disk is
-uniform. The linearized dust-gas drag term is then
-$-\delta\mathcal{C}/\rho = \tstop\tepsilon c_s^2k_x^2 Q$. The
-thin-disk solution to Eq. \ref{lin_sg} is $\dd\psi = -2\pi G
+%\section{Connection to previous results}
+
+\subsection{Secular gravitational instability}\label{sgi}
+We first consider a razor-thin, self-gravitating disk so that $\rho =
+\Sigma\delta(z)$. Here $\delta$ is the Dirac delta function and $\Sigma$ is
+the total surface density. Now $\epsilon$ and 
+$\tepsilon$ refers to the global dust-to-gas ratio and dust 
+fraction, respectively. The background disk is
+uniform and we neglect the vertical dimension   
+in Eq. \ref{lin_mass}---\ref{lin_energy}. The linearized dust-gas drag
+term is then $-\delta\mathcal{C}/\rho = \tstop\tepsilon c_s^2k_x^2
+Q$. The thin-disk solution to Eq. \ref{lin_sg} is $\dd\psi = -2\pi G
 \Sigma W/\left|k_x\right|$
@@
 \end{align}
-This is just the secular gravitational instability in the absence of
-turbulent dust diffusion but strong drag
-\citep[][ their Eq. 13 becomes Eq. \ref{sgi_disp} in this limit with
-a change of variables]{takahashi14}. This is gravitational instability
-mediated by dust-gas drag. A similar effect occurs in viscous
+This is secular gravitational instability mediated by strong
+dust-gas drag with negligible turbulent dust diffusion 
+\citep[][ their Eq. 13 becomes our Eq. \ref{sgi_disp} in this limit
+after 
+a change of variables]{takahashi14}. A similar effect occurs in viscous
 self-gravitating gas disks \citep{gammie96,lin16}. In fact, if we
-identify $\nu \equiv \tstop c_s^2$ as a kinematic viscosity, then
-Eq. \ref{sgi_disp} is effectively 
-\citeauthor{gammie96}'s Eq. 18.
+identify $\nu \equiv \tstop \tepsilon c_s^2$ as a kinematic viscosity,
+then Eq. \ref{sgi_disp} is identical to \citeauthor{gammie96}'s Eq. 18.
 
 This exercise shows that the one-fluid framework, futher simplified by
-the terminal velocity approximation, is sufficient to capture the SGI
+the terminal velocity approximation, is sufficient to capture SGI
 in the strong drag limit.
 
-\section{Streaming instability}\label{si}
+\subsection{Streaming instability}\label{si}
 We now consider 3D but non-self-gravitating disks.  

### 2017-01-06 model.tex
minor edits on SGI stuff

@@
 
-The total gravitational potential $\Phi_\mathrm{tot}=\Phi + \psi$,
+The total gravitational potential is $\Phi_\mathrm{tot}=\Phi + \psi$,
 where $\Phi$ originates from a star of mass $M_*$, 
@@
 The second equality in Eq. \ref{thin_disk_potential} is the 
-thin-disk approximation for the disk potential, appropriate for
+thin-disk approximation for the stellar potential, appropriate for
 $|z|\ll r$. We shall adopt this approximate potential in order to
@@
 We include the gravitational potential of the mixture 
-$\psi$ for completeness but will neglect it in most of this
-work. Henceforth we consider non-self-gravitating disks ($\psi \equiv
-0$) unless explicitly stated otherwise.
+$\psi$ for completeness but will neglect it throughout most of this
+work ($\psi \equiv 0$) unless explicitly stated otherwise.
 
@@
 flow.
+
+{\bf
+\subsection{Physical disk conditions}
+physical disk conditions underwhich the above approx
+(small/well-coupled dust and isothermal gas) is applicable. smaller
+than mm at few tens of au. give dimensional numbers here
+}
 
@@
 oscillation amplitudes. To demonstrate this, we calculate the average
-work done during a periodic oscillation, and show that if the work
-done is positive, then the 
+work done assuming periodic oscillations, and show that if the average
+work done is positive, then the 
 oscillation amplitude would actually grow.

### 2017-01-06 paper.tex
minor edits on SGI stuff

@@
 explicit stability analyses. In  
-\S\ref{linear_problem} we list the linearized equations. We solve them 
-to recover SGI in \S\ref{sgi}, to interpret the SI in \S\ref{si}, and
-in  \S\ref{results} to extend the VSI to dusty disks. We  
+\S\ref{linear_problem} we list the linearized equations and show that 
+they reproduce previous results on SGI and SI. In \S\ref{results} we
+extend the VSI to dusty disks. We   
 discuss future applications of the dusty/adiabatic gas equivalence in 
@@
 The overall positive work done leads to growth in the oscillation amplitude. 
-This effect is similar to being pushed downwards on a swing on
-descent.
+This effect is similar to being pushed downwards on a swing when 
+descending.
 
@@
 disk temperature  $T(r)$ is time-independent but varies with radius
-\citep{chiang97}. Since $P\propto \rho T$ for an ideal gas, the pressure of a fluid parcel varies with 
+$r$ from the star \citep{chiang97}. Since $P\propto \rho T$ for an ideal gas, the pressure of a fluid parcel varies with 
 position as well. There is no reason to expect $P$ and $\rho$ to

### 2017-01-06 model.tex
tiny edits

@@
 Eq. \ref{int_rel} indicates that $\sigma^2$ is generally complex, so
-we may have growing oscillations (overstability). %Writing $\sigma =
+we may have growing oscillations. %Writing $\sigma =
 %-\omega + \ii s$ where $\omega$ and $s$ are the real frequency and
@@
 %    v_r|^2 + |\dd v_z|^2\right)dV}, \label{thermal_instability}
-  s = \frac{\imag\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV}{2\omega\mathcal{I}^2}. \label{thermal_instability}
+  s = \frac{\imag\int \left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}dV}{2\omega\mathcal{I}^2}, \label{thermal_instability}
 \end{align} 
-This represents instability due to dust-gas friction. 
+assuming $\omega\neq0$. This represents overstability due to dust-gas friction. 
 Since drag ($\delta \mathcal{C}$) appears as a source term in our
@@
 instabilities in stars \citep{cox67}. We thus interpret  
-dust-drag instabilities in a similar way, 
+dust-drag overstabilities in a similar way, 
 adapting from the treatment of stellar 
@@
 
-\subsubsection{Physical requirement of dust-gas drag instabilities}
+\subsubsection{Physical requirement of dust-gas drag overstabilities}
 
@@
   s = \frac{\imag\int P
-    \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV}{2\omega\mathcal{I}^2}. \label{vsi_check} 
+    \left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV}{2\omega\mathcal{I}^2}, \label{vsi_check} 
 \end{align} 
+again assuming $\omega\neq0$. 
 This is VSI caused by vertical shear arising from a radial

### 2017-01-06 summary.tex
minor edits on SGI stuff
tiny edits

@@
 
-\subsubsection{Gravitational instability} %vertical structure 
+\subsubsection{Classic gravitational instabilities} %vertical structure 
 The addition of dust enhances gravitational
@@
 explicitly show that the SI occurs because the evolution of gas
-pressure lags behind  dust density; a general property of \emph{all}
-instabilities driven by strong dust-gas drag. 
+pressure lags behind  dust density; a general property of 
+overstabilities driven by strong dust-gas drag. 
 It takes a finite time for the gas to respond to

### 2017-01-19 paper.tex
tiny intro edits

@@
 protoplanetary disks the
-disk temperature  $T(r)$ is time-independent but varies with radius
-$r$ from the star \citep{chiang97}. Since $P\propto \rho T$ for an ideal gas, the pressure of a fluid parcel varies with 
-position as well. There is no reason to expect $P$ and $\rho$ to
-be in phase. In fact, we will show that this is responsible for 
+disk temperature  $T(r)$ is time-independent but varies with the cylindrical radius
+$r$ from the star \citep{chiang97}. Since $P\propto \rhog T$ for an ideal gas, the pressure of a gas parcel varies with 
+position as well. There is no reason to expect $P$ and $\rhog$ to
+be in phase. In fact, we will show that this is a fundamental property of 
 the vertical shear instability \citep{lin15}.
@@
 and there is no work done. However, for finite dust-gas drag, $\rhod$ and
-$\rhog$ are not neccessarily in phase. If this causes $P$ to lag
+$\rhog$ are not neccessarily in phase, because dust particles can drift relative to the gas. 
+\emph{If} this causes $P$ to lag 
 $\rho$, then the positive work done would lead to growing

### 2017-04-24 model.tex
edits from discussion with andrew

@@
 Notice Eq. \ref{eos} resembles an ideal gas equation of state but with  
-effective temperature $T_\mathrm{eff} = 
+a reduced temperature $\tilde{T} = 
 T_\mathrm{ref}\left(1-\tepsilon\right)$. %sound-speed $c_{s,\mathrm{eff}} = c_s\sqrt{\left(1 -
@@
 
-In this work, we assume that $\tstop$  is sufficiently small so
-that settling occurs on much longer timescales  than that of interest; or
-consider equilibria with $\mathcal{C}\equiv 0$, namely   
+In this work, consider equilibria with $\mathcal{C}\equiv 0$, namely   
 \begin{inparaenum}[1)] 
@@
 \end{inparaenum} 
-Vertical settling is absent in both cases.
+Vertical settling is absent in both cases. Alternatively, we are 
+assuming that $\tstop$  is sufficiently small such 
+that settling occurs on much longer timescales than the instabilities of interest.
 
@@
 &= \int\left[ \rho
-  |\dd v_r|^2A + \rho  \dd v_z \dd v_r^* B + \rho \dd v_z^*\dd v_r C +
+  |\dd v_r|^2A + \rho  \left(\dd v_z \dd v_r^* + \dd v_z^*\dd v_r\right) B +
   \rho |\dd v_z|^2 D\phantom{\frac{1}{1}}\right. \notag\\
@@
 is the meridional kinetic energy, 
-and coefficients $A,B,C,D$ can be
+and coefficients $A,B,D$ be
 read off Eq. \ref{integral_ex}. %Note that $B=C$ from dynamical equilibrium
@@
 
-In this case, the last two integrals in Eq. \ref{int_rel} vanish. Then 
-requiring $\sigma^2>0$  gives the criteria 
-for the axisymmtric \emph{stability} of an isothermal gas perfectly 
-coupled to dust,  
+In this case, the last two integrals in Eq. \ref{int_rel} vanish. 
+The condition for axisymmtric \emph{stability}, $\sigma^2>0$, 
+is met if the integrand cofficients satisfy $A+D>0$ and $AD-B^2>0$
+\citep[][Section 11.6]{ogilvie16}, which becomes 
 \begin{align}
@@
 \subsubsection{Order-of-magnitude estimates} 
-We evaluate the above stability criteria for the vertically-Gaussian
-distributions for the gas density and the dust-to-gas ratio as 
-described in \S\ref{eqm}. 
+We evaluate the above stability criteria for the equilibria described in \S\ref{eqm}. 
 %However, here we permit 
@@
 %radius. 
-We assume the disk is approximately Keplerian so that
-$\kappa\simeq\OmK$. In terms of the dust-to-gas
+We assume the disk is approximately Keplerian so that $\kappa\simeq\OmK$, and that the vertical gas distribution is
+Gaussian. In terms of the dust-to-gas
 ratio $\epsilon$, Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2}

### 2017-04-24 summary.tex
typo in generalized PV
edits from discussion with andrew

@@
 
-\subsection{Higher order corrections}
-%{\bf notes for self. maybe exclude in final. }
-
-The one-fluid formulation, 
-Eqs. \ref{masseq}---\ref{tempeq}, can 
-produce spuriously-growing modes with frequency $|\sigma|\gtrsim
-\Omega$ (see Appendix \S\ref{spurious_epi}). This 
-results from neglecting terms of $O(|\widetilde{\bm{v}}|^2)$ after converting
-the full two-fluid equations to the one-fluid formulation. The full
-one-fluid equations 
-%is significantly more complex than our 
-%model \cite[e.g.][]{laibe14}. It 
- includes an evolutionary equation for $\widetilde{\bm{v}}$, which has no
- hydrodynamic counter-part \citep{laibe14}. However, it is
- still possible to improve upon 
- the first-order formulation whilst retaining similarilty to  
- standard hydrodynamics.  
-%without having the complexity of the full equations
-
-The full one-fluid equations of motion includes the acceleration term 
-\begin{align}
-%\left.\frac{D\bm{v}}{Dt}\right|_\mathrm{add} = 
-%\bm{a} =  
--\frac{1}{\rho}\nabla\cdot\left(\frac{\rhod\rhog}{\rho}\widetilde{\bm{v}}\otimes\widetilde{\bm{v}}\right) 
-\equiv  \frac{1}{\rho}\nabla\cdot \bm{T} 
-\end{align}
-\citep{youdin05a}, 
-where $\bm{T}$ is an effective stress tensor. 
-In the terminal velocity
-approximation with $\widetilde{\bm{v}} = \tstop 
-\nabla P/\rhog$, we have 
-\begin{align}
-\bm{T} = - \frac{\epsilon\tstop^2}{\rho}\nabla P \otimes \nabla P.
-\end{align}
-If we define 
-\begin{align}
-  \bm{g} \equiv \sqrt{4\pi G \epsilon \rho} \frac{\nabla P}{\rho}
-  \tstop,\label{effective_g} 
-\end{align}
-then
-\begin{align}
-\bm{T} = - \frac{1}{4\pi G} \bm{g}\otimes\bm{g}, \label{effective_gstress}
-\end{align}
-%{\bf tensor virial theorems?}
-which is just the anisotropic part of the  
-gravitational stress tensor in self-gravitating hydrodynamics 
-\citep{lynden-bell72}.
-
- However, unlike self-gravity, where $\bm{g}$ is
-determined through the Poisson equation; here $\bm{g}$ is directly
-related to $P$ (Eq. \ref{effective_g}), which itself obeys an 
-evolutionary equation (Eq. \ref{eff_energy}). 
-Nevertheless, Eq. \ref{effective_g}---\ref{effective_gstress}
-show that correlations in  pressure gradients lead to 
-momentum transport of the mixture as a whole. This is similar to Reynolds stresses in
-a turbulent disk \citep{balbus99}. In particular, 
-outwards angular momentum transport requires $\p_rP\p_\phi P > 0$.
-
-In a following paper we will include $\bm{T}$ in the one-fluid
-equations and futher explore the analogy between dusty-gas and viscous
-hydrodynamics.
-
 %In particular, we will explore whether this can surpress the suprious
@@
 %stratified case require more boundary conditions
-
-\subsection{Implications for numerical simulations}
-The dusty/adiabatic gas equivalence means that pure
-gas dynamics codes can be applied to simulate locally
-isothermal/polytropic dusty gas. One need not add any new 
-equations to evolve the dust component: 
-%In fact, no
-%modification is needed  
-%for strictly isothermal or polytropic gas perfectly-coupled to dust:  
- the standard energy equation substitutes for the dust continuity
-equation. When $c_s$ or $K$ is not constant and/or $\tstop\neq0$ one
-should add corresponding source terms in the energy equation
-(e.g. Eq. \ref{poly_energy}). 
-The source term associated with dust-gas drag, $\mathcal{C}$, is 
-analogous to radiative diffusion \citep{price15}, which is also
-featured in many hydrodynamics codes.
 
@@
 $\epsilon, Z\ll1$, this effect is unimportant. However, if $\epsilon$ is
-large (e.g. due to dust settling) then the effective temperature may 
+large (e.g. due to dust settling) then the reduced temperature $\tilde{T}$ may 
 be lowered to enable instability.
 
-As noted in \S\ref{loc_iso_eos}, dust settling causes the effective
-temperature to \emph{increase} away from the midplane. This contrasts
+As noted in \S\ref{loc_iso_eos}, dust settling causes 
+$\tilde{T}$ to \emph{increase} away from the midplane. This contrasts
 to previous studies of GI in vertically stratified disks
@@
 \begin{align}
- \mathcal{V} = \frac{\kappa^2}{2\Omega\Sigma}\mathcal{S}^{2/\gamma},  
+ \mathcal{V} = \frac{\kappa^2}{2\Omega\Sigma}\mathcal{S}^{-2/\gamma},  
 \end{align} 
@@
 
-\subsection{New dust-drag
-  instabilities}\label{dust_as_thermo}
+\subsection{Implications for numerical simulations}
+The dusty/adiabatic gas equivalence means that pure
+gas dynamics codes can be applied to simulate locally
+isothermal/polytropic dusty gas. One need not add any new 
+equations to evolve the dust component: 
+%In fact, no
+%modification is needed  
+%for strictly isothermal or polytropic gas perfectly-coupled to dust:  
+ the standard energy equation substitutes for the dust continuity
+equation. When $c_s$ or $K$ is not constant and/or $\tstop\neq0$ one
+should add corresponding source terms in the energy equation
+(e.g. Eq. \ref{poly_energy}). 
+The source term associated with dust-gas drag, $\mathcal{C}$, is 
+analogous to radiative diffusion \citep{price15}, which is also
+featured in many hydrodynamics codes.
 
-%{\bf vsi enabled by drag? slipping between dust and gas reduces
- % effective buoyancy}
+\subsection{Higher order corrections}
+%{\bf notes for self. maybe exclude in final. }
 
-As described in \S\ref{dust_work}, 
-the physical cause of dust-drag instabilities is a phase lag between
-the pressure and density response in a dusty gas\footnote{
-In fact, we expect a phase difference between pressure and the total
-density  whenever $\tstop\neq0$; not just $\tstop\OmK\ll 1$ as 
-considered in this work. This is simply because of the finite
-relaxation time required for dust to respond to gas, and vice versa. 
-An effective energy equation in the form of Eq. \ref{eff_energy} can
-still be obtained for general $\tstop$. The
-heat flux $\bm{F}$ in the cooling function $\mathcal{C}$ is determined
-through solution to the relative velocity $\Delta \bm{v} =
-\bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$ \citep{youdin05a,laibe14}.  We
-thus expect the `pressure-density lag' interpretation of dust-drag 
-instabilities to be generally applicable.}. 
-This may be used to re-interpret the streaming instability or to
-understand other dusty instabilities. For example,
-\cite{loren15,loren16} have recently discovered a new axisymmetric 
-instability in their two-fluid numerical simulations of dusty
-protoplanetary disks. They find it leads to the formation of
-`toroidal vortices'.
+The one-fluid formulation, 
+Eqs. \ref{masseq}---\ref{tempeq}, can 
+produce spuriously-growing modes with frequency $|\sigma|\gtrsim
+\Omega$ (see Appendix \S\ref{spurious_epi}). This 
+results from neglecting terms of $O(|\widetilde{\bm{v}}|^2)$ after converting
+the full two-fluid equations to the one-fluid formulation. The full
+one-fluid equations 
+%is significantly more complex than our 
+%model \cite[e.g.][]{laibe14}. It 
+ includes an evolutionary equation for $\widetilde{\bm{v}}$, which has no
+ hydrodynamic counter-part \citep{laibe14}. However, it is
+ still possible to improve upon 
+ the first-order formulation whilst retaining similarilty to  
+ standard hydrodynamics.  
+%without having the complexity of the full equations
 
-%give numerical example 
-We have applied our one-fluid stability analysis to find unstable modes  
-similar to that described by \cite{loren15}. We use the same
-parameters reported by them as follows. 
-%we set $p=q=0$, which eliminates VSI and minimizes streaming
-%instabilities as there is no radial pressure gradient at the disk
-%midplane.   
-The total metalicity is $Z=0.01$ and the dust scale-height $H_\mathrm{d} =
-0.4H_\mathrm{g}$, giving a maximum dust-to-gas ratio
-$\mathrm{max}(\epsilon) = 0.025$. The midplane stopping time is 
-$t_\mathrm{s0}=0.02\OmK^{-1}$. The gas disk parameters are $(p,q,
-\hgas) =(-1.5, -1,0.05)$.  We set $k_x = \pi/H_\mathrm{d}$.
+The full one-fluid equations of motion includes the acceleration term 
+\begin{align}
+%\left.\frac{D\bm{v}}{Dt}\right|_\mathrm{add} = 
+%\bm{a} =  
+-\frac{1}{\rho}\nabla\cdot\left(\frac{\rhod\rhog}{\rho}\widetilde{\bm{v}}\otimes\widetilde{\bm{v}}\right) 
+\equiv  \frac{1}{\rho}\nabla\cdot \bm{T} 
+\end{align}
+\citep{youdin05a}, 
+where $\bm{T}$ is an effective stress tensor. 
+In the terminal velocity
+approximation with $\widetilde{\bm{v}} = \tstop 
+\nabla P/\rhog$, we have 
+\begin{align}
+\bm{T} = - \frac{\epsilon\tstop^2}{\rho}\nabla P \otimes \nabla P.
+\end{align}
+If we define 
+\begin{align}
+  \bm{g} \equiv \sqrt{4\pi G \epsilon \rho} \frac{\nabla P}{\rho}
+  \tstop,\label{effective_g} 
+\end{align}
+then
+\begin{align}
+\bm{T} = - \frac{1}{4\pi G} \bm{g}\otimes\bm{g}, \label{effective_gstress}
+\end{align}
+%{\bf tensor virial theorems?}
+which is just the anisotropic part of the  
+gravitational stress tensor in self-gravitating hydrodynamics 
+\citep{lynden-bell72}.
 
-Fig. \ref{compare_eigenvals_ts0d02} shows the unstable modes found for
-this setup. We find the usual VSI modes, here with $\omega> 3\hgas\OmK,\, s<
-0.2\hgas\OmK$. However, we also find modes with $\omega<s$. One of these,
-marked by the blue asterisk, has 
-frequency 
-\begin{align*}
-\sigma = (0.3631\ii - 0.1251)\hgas\OmK. 
-\end{align*}
-This growth rate corresponds to $8.8$ orbits, comparable to the $12$
-orbits reported by \citeauthor{loren15}. We visualize this mode in
-Fig. \ref{result2d_loren}. We find this mode 
-persists when $q=0$, so it is not the VSI. 
-Using Eq. \ref{streaming_dispersion} with the above parameters,  
-we also estimate a much smaller growth 
-rate for the streaming instability of 
-$O(10^{-3}\hgas\OmK)$, We thus also rule out
-SI, as did \citeauthor{loren15}. However, the mode disappears with $\tstop=0$, so it \emph{is}
-associated with finite dust-gas drag.
+ However, unlike self-gravity, where $\bm{g}$ is
+determined through the Poisson equation; here $\bm{g}$ is directly
+related to $P$ (Eq. \ref{effective_g}), which itself obeys an 
+evolutionary equation (Eq. \ref{eff_energy}). 
+Nevertheless, Eq. \ref{effective_g}---\ref{effective_gstress}
+show that correlations in  pressure gradients lead to 
+momentum transport of the mixture as a whole. This is similar to Reynolds stresses in
+a turbulent disk \citep{balbus99}. In particular, 
+outwards angular momentum transport requires $\p_rP\p_\phi P > 0$.
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/compare_eigenvals_ts0d02}
-  \caption{Unstable modes in a locally isothermal, dusty disk
-    with stopping time  
-    $\tstop=0.02\OmK^{-1}$. Modes with $\omega>3$ are due to
-    VSI. The mode marked by the blue asterisk, visualized in
-    Fig. \protect\ref{result2d_loren}
-    , is associated with finite dust-gas
-    drag. 
-  }\label{compare_eigenvals_ts0d02}
-\end{figure}
+In a following paper we will include $\bm{T}$ in the one-fluid
+equations and futher explore the analogy between dusty-gas and viscous
+hydrodynamics.
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/result2d_loren.ps}
-  \caption{Visualization of the mode marked by the blue asterisk in
-    Fig. \protect\ref{compare_eigenvals_ts0d02}. 
-    This is a possible eigenmode of the dusty instability discovered by  
-    \cite{loren15} in their numerical simulations. The instability growth timescale and 
-    radial spacing of the `toroidal vortices' are similar to that 
-    reported by \citeauthor{loren15}. This instability is driven by
-    dust-gas drag. The color scale shows the 
-    perturbed  
-    dust-to-gas ratio, $\delta\epsilon$; and the arrows show
-    $\sqrt{\rho}\left(\dd v_x, \dd v_z\right)$. 
-  }\label{result2d_loren}
-\end{figure}
+% \subsection{New dust-drag
+%   instabilities}\label{dust_as_thermo}
 
-One possible interpretation of this mode is `drag-driven VSI', to be 
-compared with `cooling-driven VSI' studied by \cite{lin15}. In the
-latter case, buoyancy forces are eliminated by fast
-cooling, allowing vertical shear to destabilize the disk. Now, in an
-isothermal dusty disk if $\tstop=0$ then mixture is 
-effectively adiabatic: dust-induced buoyancy forces are
-strongest. However, having $\tstop\neq0$ allow dust particles to slip
-past the gas, reducing buoyancy. That is, dust-gas drag
-provies an effective cooling.
+% %{\bf vsi enabled by drag? slipping between dust and gas reduces
+%  % effective buoyancy}
 
-We stress that this example only serves to highlight
-the possibility of new axisymmetric dust-drag instabilities in
-stratified disks. A major caveat of the above result is that 
-%the base state is not an exact equilibrium, because dust-settling would occur
-%in practice. (
- the background dust settling timescale is comparable to the instability
-growth timescale. It is essential to construct
-exact equilibria for a meaningful stability analysis. We defer this
-to future work.
+% As described in \S\ref{dust_work}, 
+% the physical cause of dust-drag instabilities is a phase lag between
+% the pressure and density response in a dusty gas\footnote{
+% In fact, we expect a phase difference between pressure and the total
+% density  whenever $\tstop\neq0$; not just $\tstop\OmK\ll 1$ as 
+% considered in this work. This is simply because of the finite
+% relaxation time required for dust to respond to gas, and vice versa. 
+% An effective energy equation in the form of Eq. \ref{eff_energy} can
+% still be obtained for general $\tstop$. The
+% heat flux $\bm{F}$ in the cooling function $\mathcal{C}$ is determined
+% through solution to the relative velocity $\Delta \bm{v} =
+% \bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$ \citep{youdin05a,laibe14}.  We
+% thus expect the `pressure-density lag' interpretation of dust-drag 
+% instabilities to be generally applicable.}. 
+% This may be used to re-interpret the streaming instability or to
+% understand other dusty instabilities. For example,
+% \cite{loren15,loren16} have recently discovered a new axisymmetric 
+% instability in their two-fluid numerical simulations of dusty
+% protoplanetary disks. They find it leads to the formation of
+% `toroidal vortices'.
 
-%also unclear what are "proper" vertical bc for dusty disks
+% %give numerical example 
+% We have applied our one-fluid stability analysis to find unstable modes  
+% similar to that described by \cite{loren15}. We use the same
+% parameters reported by them as follows. 
+% %we set $p=q=0$, which eliminates VSI and minimizes streaming
+% %instabilities as there is no radial pressure gradient at the disk
+% %midplane.   
+% The total metalicity is $Z=0.01$ and the dust scale-height $H_\mathrm{d} =
+% 0.4H_\mathrm{g}$, giving a maximum dust-to-gas ratio
+% $\mathrm{max}(\epsilon) = 0.025$. The midplane stopping time is 
+% $t_\mathrm{s0}=0.02\OmK^{-1}$. The gas disk parameters are $(p,q,
+% \hgas) =(-1.5, -1,0.05)$.  We set $k_x = \pi/H_\mathrm{d}$.
+
+% Fig. \ref{compare_eigenvals_ts0d02} shows the unstable modes found for
+% this setup. We find the usual VSI modes, here with $\omega> 3\hgas\OmK,\, s<
+% 0.2\hgas\OmK$. However, we also find modes with $\omega<s$. One of these,
+% marked by the blue asterisk, has 
+% frequency 
+% \begin{align*}
+% \sigma = (0.3631\ii - 0.1251)\hgas\OmK. 
+% \end{align*}
+% This growth rate corresponds to $8.8$ orbits, comparable to the $12$
+% orbits reported by \citeauthor{loren15}. We visualize this mode in
+% Fig. \ref{result2d_loren}. We find this mode 
+% persists when $q=0$, so it is not the VSI. 
+% Using Eq. \ref{streaming_dispersion} with the above parameters,  
+% we also estimate a much smaller growth 
+% rate for the streaming instability of 
+% $O(10^{-3}\hgas\OmK)$, We thus also rule out
+% SI, as did \citeauthor{loren15}. However, the mode disappears with $\tstop=0$, so it \emph{is}
+% associated with finite dust-gas drag.
+
+% \begin{figure}
+%   \includegraphics[width=\linewidth]{figures/compare_eigenvals_ts0d02}
+%   \caption{Unstable modes in a locally isothermal, dusty disk
+%     with stopping time  
+%     $\tstop=0.02\OmK^{-1}$. Modes with $\omega>3$ are due to
+%     VSI. The mode marked by the blue asterisk, visualized in
+%     Fig. \protect\ref{result2d_loren}
+%     , is associated with finite dust-gas
+%     drag. 
+%   }\label{compare_eigenvals_ts0d02}
+% \end{figure}
+
+% \begin{figure}
+%   \includegraphics[width=\linewidth]{figures/result2d_loren.ps}
+%   \caption{Visualization of the mode marked by the blue asterisk in
+%     Fig. \protect\ref{compare_eigenvals_ts0d02}. 
+%     This is a possible eigenmode of the dusty instability discovered by  
+%     \cite{loren15} in their numerical simulations. The instability growth timescale and 
+%     radial spacing of the `toroidal vortices' are similar to that 
+%     reported by \citeauthor{loren15}. This instability is driven by
+%     dust-gas drag. The color scale shows the 
+%     perturbed  
+%     dust-to-gas ratio, $\delta\epsilon$; and the arrows show
+%     $\sqrt{\rho}\left(\dd v_x, \dd v_z\right)$. 
+%   }\label{result2d_loren}
+% \end{figure}
+
+% One possible interpretation of this mode is `drag-driven VSI', to be 
+% compared with `cooling-driven VSI' studied by \cite{lin15}. In the
+% latter case, buoyancy forces are eliminated by fast
+% cooling, allowing vertical shear to destabilize the disk. Now, in an
+% isothermal dusty disk if $\tstop=0$ then mixture is 
+% effectively adiabatic: dust-induced buoyancy forces are
+% strongest. However, having $\tstop\neq0$ allow dust particles to slip
+% past the gas, reducing buoyancy. That is, dust-gas drag
+% provies an effective cooling.
+
+% We stress that this example only serves to highlight
+% the possibility of new axisymmetric dust-drag instabilities in
+% stratified disks. A major caveat of the above result is that 
+% %the base state is not an exact equilibrium, because dust-settling would occur
+% %in practice. (
+%  the background dust settling timescale is comparable to the instability
+% growth timescale. It is essential to construct
+% exact equilibria for a meaningful stability analysis. We defer this
+% to future work.
+
+% %also unclear what are "proper" vertical bc for dusty disks
 

### 2017-04-26 model.tex
new fig to illustrate phase lag

@@
 \end{align}
-\citep{jacquet11}. This equation implies that dust particles
-drift towards pressure maxima.
+\citep{jacquet11}. This equation reflect the physical result that dust
+particles tend to drift towards pressure maxima.
 
@@
 
-We include the gravitational potential of the mixture 
-$\psi$ for completeness but will neglect it throughout most of this
-work ($\psi \equiv 0$) unless explicitly stated otherwise.
+The gravitational potential of the mixture 
+$\psi$ is included here for completeness, but we will mostly neglect
+it ($\psi \equiv 0$) unless explicitly stated otherwise.
 
@@
 Notice Eq. \ref{eos} resembles an ideal gas equation of state but with  
-a reduced temperature $\tilde{T} = 
+a reduced temperature $\widetilde{T} = 
 T_\mathrm{ref}\left(1-\tepsilon\right)$. %sound-speed $c_{s,\mathrm{eff}} = c_s\sqrt{\left(1 -
-This effective mixture  
-temperature is lowered with dust-loading. 
+This reduced temperature is lowered with dust-loading. 
 %   \tepsilon\right)}$. 
@@
 which is in the same form as cooling by radiative diffusion. In protoplanetary
-disks the heat flux, proportional to $\bm{F}$, is directly radially
-outwards and vertically upwards. This is simply a reflection of
-particle drift (inwards and downwards). Particle flux into a region contributes
-to `cooling' of that region because the effective temperature is
-lowered.
+disks the corresponding heat flux, proportional to $\bm{F}$, is
+directly radially outwards and vertically upwards. This is simply a
+reflection of particle drift towards increasing pressure (inwards and 
+downwards). Particle flux into a region contributes to `cooling' of
+that region because the reduced temperature is lowered.
 
@@
 The specific entropy of an ideal gas is 
-$S = C_p\ln{\left(P^{1/\gamma}/\rhog\right)}$, where $C_p$ is the heat 
+$S = C_P\ln{\left(P^{1/\gamma}/\rhog\right)}$, where $C_P$ is the heat 
 capacity at constant pressure.
 
-Since we have shown that a locally isothermal dusty gas effectively has $\gamma=1$ we can define an effective (dimensionless)   
+Since we have shown that a locally isothermal dusty gas effectively
+has $\gamma=1$ we can define an effective (dimensionless)    
 entropy for the \emph{mixture} as 
@@
 P\right)<0$.  This corresponds to pressure perturbations \emph{lagging}
-behind that in density (or displacements).%, since the phase of
+behind that in density (or displacements), see Fig. \ref{lag_cartoon}
+for the case of dusty gas.
+
+%, since the phase of
 %the assumed perturbations increase in time.  
 %$\imag\left(\Delta\rho\right)=0$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 5cm 0cm 0cm]{figures/lag}
+  \caption{Phase relation for overstable modes in isothermal dusty
+    gas. Such modes require (Lagrangian) oscillations in the gas
+    pressure, which is directly proportional to the gas density, to
+    lag behind that in dust density. 
+    \label{lag_cartoon}
+  }
+\end{figure}
 

### 2017-04-26 paper.tex
abstract slight edit
new fig to illustrate phase lag

@@
 \item 
- the settling of perfectly coupled dust cannot cause axisymmetric 
- instabilties, however thin the dust layer becomes; 
+ perfectly coupled dust cannot cause axisymmetric 
+ instabilties, however thin the dust layer is; 
 \item 
@@
 streaming instability 
-\citep{youdin05a,youdin07b,johansen07}, secular gravitational
+\citep[SI,][]{youdin05a,youdin07b,johansen07}, secular gravitational
 instabilities \citep[SGI,][]{ward00,youdin11,michikoshi12,takahashi14} and
@@
 layer is. We put forward a thermodynamical interpretation of
-dust-gas drag instabilities, and apply it to the
-streaming instability \citep[SI, ][]{youdin05a,youdin07b,jacquet11}. 
+dust-gas drag instabilities, and apply it to the streaming instability.   
 We also study the effect of dust-loading on the vertical shear
@@
 a pressure-density phase lag in fluids can lead to 
-instability. Thus finding an analogy between dusty gas and
-pure hydrodynamics allow us to interpret dusty-gas dynamics
+instability. Consequently, an analogy between dusty gas and  
+pure hydrodynamics allow us to interpret dusty-gas instabilities 
 within a thermodynamics framework.
@@
 \S\ref{linear_problem} we list the linearized equations and show that 
-they reproduce previous results on SGI and SI. In \S\ref{results} we
+they reproduce previous results for SGI and SI. In \S\ref{results} we
 extend the VSI to dusty disks. We   
@@
 From $A$ to $B$, a fluid parcel is expanding to return to 
-equilbrium, but pressure (i.e. restoring) forces is still 
-increasing. This over-compensation causes the parcel to expand
+equilbrium, but pressure is still 
+increasing. This over-compensation will cause the parcel to expand
 beyond the maximum volume of the previous cycle. 
@@
 equilibrium, but pressure 
-forces are still dropping. This allows the  
+is still dropping. This allows the  
 parcel's contraction to over-shoot the maximum density attained in the
@@
 disk temperature  $T(r)$ is time-independent but varies with the cylindrical radius
-$r$ from the star \citep{chiang97}. Since $P\propto \rhog T$ for an ideal gas, the pressure of a gas parcel varies with 
+$r$ from the star \citep{chiang97}. Since $P\propto \rhog T$ for an
+ideal gas, the pressure of a gas parcel varies with  
 position as well. There is no reason to expect $P$ and $\rhog$ to
-be in phase. In fact, we will show that this is a fundamental property of 
-the vertical shear instability \citep{lin15}.
+be in phase as a gas parcel oscillates between different radii. In
+fact, we will show that this is a fundamental property of the vertical
+shear instability \citep{lin15}.
 

### 2017-04-27 model.tex
new stuff on instability when dust is vertically well mixed. where to put in numerical example?

@@
 Since $\tepsilon$ typically decrease away from the midplane, we expect
-vertically isothermal dusty disks to in fact behave as if it had a
-temperture \emph{increasing} away from $z=0$.
+vertically isothermal dusty disks to behave as if it had a temperture
+\emph{increasing} away from $z=0$.
 
@@
 
-{\bf
+%{\bf
 \subsection{Physical disk conditions}
-physical disk conditions underwhich the above approx
-(small/well-coupled dust and isothermal gas) is applicable. smaller
-than mm at few tens of au. give dimensional numbers here
-}
+%physical disk conditions underwhich the above approx
+%(small/well-coupled dust and isothermal gas) is applicable. smaller
+%than mm at few tens of au. give dimensional numbers here
+%}
+
+The terminal velocity approximation applies to short
+stopping times, $\tstop\OmK\ll 1$; while the locally isothermal
+approximation applies to short cooling times, $t_\mathrm{cool}\OmK\ll
+1$. In typical protoplanetary disk
+models such as the Minimum Mass Solar Nebula, $\tstop$ decreases for
+smaller particles and/or at smaller radii \citep{youdin11}. However,
+$t_\mathrm{cool}$ is generally only small in the outer disk
+\citep{lin15,malygin17}. Combining these results, we estimate that our
+framework is best suited for mm (or smaller) sized particles at a few
+to 10s of AU in protoplanetary disks. However, note that the
+isothermal approximation may be relaxed to other fixed equations of
+state (see \S\ref{gen_poly}).
 
@@
 
+We expect Eq. \ref{dusty_solberg1} to be satisfied in typical
+protoplanetary disks where the dust-to-gas ratio increases in the same
+direction as the local pressure gradient, which is equivalent to the
+gas density gradient for isothermal gas. Thus to enable instability
+here we need to violate Eq. \ref{dusty_solberg2}. We explore this
+below.
+
+\subsubsection{Vertically well-mixed dust}
+Here we consider the special case where the dust-to-gas ratio is
+uniform in $z$. Then the LHS of Eq. \ref{dusty_solberg2} becomes
+\begin{align}
+-\frac{c_s^4}{(1+\tepsilon)^4}\left(\frac{\p\ln{\rhog}}{\p
+    z}\right)^2\left(\frac{\p\epsilon}{\p r}\right)^2<0.
+\end{align}
+Thus the disk is \emph{unstable}, since there is no effective vertical 
+buoyancy to stabilize vertical motions. In fact, this situation is
+identical to the pure gas, adiabtic simulations of \cite{nelson13} in
+which the gas entropy is vertically uniform but has a radial
+gradient. There, \citeauthor{nelson13} indeed find instability.
+
 \subsubsection{Order-of-magnitude estimates} 
-We evaluate the above stability criteria for the equilibria described in \S\ref{eqm}. 
-%However, here we permit 
-%$\epsilon_0=\epsilon_0(r)$ and $\Htilde=\Htilde(r)$ to vary with
-%radius. 
+In order to apply the stability criteria
+Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} to typical
+protoplanetary disks we conside the equilibria described in
+\S\ref{eqm}.
+
 We assume the disk is approximately Keplerian so that $\kappa\simeq\OmK$, and that the vertical gas distribution is
@@
 vary on an extremely short lengthscale. 
-%an extremely short  
-%radial length-scale in the dust-to-gas ratio. 
-This is unlikely to
-occur. On the other hand, if dust is well-mixed, say $\Hd\simeq
-0.99\Hg$, then instability only requires 
-$L_\epsilon\lesssim O(10\Hg)$, which is more realistic.
+% %an extremely short  
+% %radial length-scale in the dust-to-gas ratio. 
+% This is unlikely to
+% occur. On the other hand, if dust is well-mixed, say $\Hd\simeq
+% 0.99\Hg$, then instability only requires 
+% $L_\epsilon\lesssim O(10\Hg)$, which is more realistic.
 

### 2017-04-28 model.tex
added numerical example of unstable mode with vertically well mixed dust

@@
 
-\subsubsection{Vertically well-mixed dust}
+\subsubsection{Vertically well-mixed dust}\label{vert_mixed_theory}
 Here we consider the special case where the dust-to-gas ratio is
@@
 Thus the disk is \emph{unstable}, since there is no effective vertical 
-buoyancy to stabilize vertical motions. In fact, this situation is
-identical to the pure gas, adiabtic simulations of \cite{nelson13} in
-which the gas entropy is vertically uniform but has a radial
-gradient. There, \citeauthor{nelson13} indeed find instability.
+buoyancy to stabilize vertical motions. This situation is
+identical to the pure gas, adiabtic simulations of \cite{nelson13}
+where the gas entropy is vertically uniform but has a radial
+gradient. The authors indeed find instability. We give a numerical
+example in the dusty context in \S\ref{vert_mixed}.
 

### 2017-04-28 result.tex
new section 7 on secular GI. rest of paper might need adjusting. mostly doing non-SG, but one sec and one appendix has SG
added numerical example of unstable mode with vertically well mixed dust

@@
 
-We present numerical solutions for  
-vertically stratified disks assuming perfectly coupled dust 
-($\tstop=0$). This is so that the equilibria defined in \S\ref{eqm} are exact
-steady states. In this limit any growing modes must be associated with
-the imposed temperature gradient (\S\ref{limits}). We are thus  
-examining the effect of dust-loading on the VSI, previously studied in 
-pure gas disks \citep[][\citetalias{lin15} in this section]{lin15}.
+We present numerical solutions for vertically stratified,
+non-self-gravitating disks assuming perfectly coupled dust
+($\tstop=0$) so that the equilibria defined in \S\ref{eqm} 
+are exact steady states.
 
-Except in  \S\ref{varHd}, we consider constant midplane
-dust-to-gas ratios $\epsilon_0$ and its characteristic height
-$H_\epsilon$. The latter is specified via the dust layer thickness
-$\Hd(r_0)$. 
-These, together with the perturbation radial wavenumber $k_x$, comprises main the
-parameters of the linear problem. We choose a power-law 
-midplane gas density profile, $\rho_\mathrm{g0}\propto r^p$ with $p =
--1.5$. The fiducial power-law index for the 
-temperature profile is $q=-1$; and we set the gas disk aspect-ratio
-$h_\mathrm{g}=0.05$. These values were also used by 
+Except in \S\ref{varHd}---\ref{vert_mixed}, we consider constant midplane dust-to-gas
+ratios $\epsilon_0$ and characteristic thickness $H_\epsilon$. Then
+the dust-to-gas ratio $\epsilon=\epsilon(z)$. In this limit any
+growing modes must be associated with the imposed temperature gradient
+(\S\ref{limits}). We are thus examining the effect of dust-loading on
+the VSI, previously studied in pure gas disks
+\citep[][\citetalias{lin15} in this section]{lin15}.
+
+The parameters for this linear problem includes $\epsilon_0$, the dust
+layer thickness $\Hd(r_0)$, and the perturbation radial wavenumber
+$k_x$. We choose a power-law midplane gas density profile,
+$\rho_\mathrm{g0}\propto r^p$ with $p = -1.5$. The fiducial power-law
+index for the temperature profile is $q=-1$; and we set the gas disk
+aspect-ratio $h_\mathrm{g}=0.05$. These values were also used by  
 \citetalias{lin15}.
@@
 
+\subsection{Pure instability with vertically well-mixed
+  dust}\label{vert_mixed}
+
+In \S\ref{vert_mixed_theory} we found that for strictly isothermal
+disks, a vertically-uniform dust-to-gas ratio can be unstable
+if $d\epsilon/dr\neq0$. To demonstate this numerically we consider the
+same disk as above but with $q=0$ and choose $H_\mathrm{d}$ such
+that $H_\epsilon=10^3\Hg$. Then vertical shear $\p_z\Omega$ is attributed
+to the radially-varying dust-to-gas ratio (see Eq. \ref{vshear2}).
+
+We show unstable modes for this disk in
+Fig. \ref{vert_mixed_modes}. In order to obtain appreciable growth  
+rates, we needed to consider super-Solar metalicity $Z=0.03$ and high
+radial wavenumber $k_x\Hg=1800$. As expected from the discussion in
+\S\ref{vert_mixed_theory}, the disk admits purely growing modes with
+$\omega=0$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/vert_mixed_modes} 
+  \caption{Unstable modes in a strictly isothermal disk ($q=0$) with
+    vertically uniform dust-to-gas ratio. The instability is due to
+    vertical shear $\p_z\Omega\neq0$ arising from radial variations in
+    the dust-to-gas ratio, $d\epsilon/dr\neq 0$. \label{vert_mixed_modes}
+    }
+\end{figure}
+
 \subsection{Validity}

### 2017-05-02 linear.tex
misc edits from sect 4 to sect 8.3

@@
 the linearized equations for 
-locally isothermal dusty gas with the pressure
+vertically isothermal dusty gas with the pressure
 equation in place of the dust-fraction
@@
 where $^\prime \equiv \p_z$ and recall $\bm{F} \equiv -\nabla
-P/\rho$. Self-gravity has been restored temporarily for the next
-section (only). The linearized dust-duffusion function 
+P/\rho$. Here we restore self-gravity temporarily for SGI in the next
+section. The linearized dust-duffusion function 
 $\dd\mathcal{C}$ is given in  Appendix \ref{lin_dust}. 
@@
 %force, given in Appendix \ref{lin_press}. 
-Note that we have assumed a temperature profile that only depends on $r$.
+%Note that we have assumed a temperature profile that only depends on $r$.
 
@@
 Q$. The thin-disk solution to Eq. \ref{lin_sg} is $\dd\psi = -2\pi G
-\Sigma W/\left|k_x\right|$
+\Sigma W/\left|k_x\right|$, where $\Sigma$ is the total surface
+density.
 
@@
 
-This exercise shows that the one-fluid framework, futher simplified by
+This exercise shows that the one-fluid framework, further simplified by
 the terminal velocity approximation, is sufficient to capture SGI
@@
 density.
+
+In a future work we will perform a detailed comparison between the
+simplified one-fluid and full two-fluid frameworks in calculating SI.
 

### 2017-05-02 model.tex
lag cartoon update, minor text edits
misc edits from sect 4 to sect 8.3

@@
 
-The total gravitational potential is $\Phi_\mathrm{tot}=\Phi + \psi$,
-where $\Phi$ originates from a star of mass $M_*$, 
+The total gravitational potential is $\Phi_\mathrm{tot}=\Phi +
+\psi$. Here $\Phi$ originates from a star of mass $M_*$, 
 \begin{align}\label{thin_disk_potential}
@@
 \end{align}
-and $\psi$ is the disk potential satisfying
-\begin{align}
-  \nabla^2\psi = 4 \pi G \rho.
-\end{align}
-Here, $G$ is the gravitational constant, and $(r,\phi, z)$ are 
-cylindrical co-ordinates centered on the star.  
+where $(r,\phi, z)$ are cylindrical co-ordinates centered on the star.   
 The second equality in Eq. \ref{thin_disk_potential} is the 
@@
 
-The gravitational potential of the mixture 
-$\psi$ is included here for completeness, but we will mostly neglect
-it ($\psi \equiv 0$) unless explicitly stated otherwise.
+The disk's own gravitational potential $\psi$ is included for
+completeness, but we will mostly neglect it unless stated
+otherwise. In such cases $\psi$   
+satisfies  
+\begin{align}
+  \nabla^2\psi = 4 \pi G \rho,
+\end{align}
+where $G$ is the gravitational constant.
 
@@
 
-\subsection{Effective energy equation}
+\subsection{Effective energy equation}\label{energy_analogy}
 Although we have eliminated the need of an evolutionary energy
@@
 However, disk structures obtained this way generally do not satisfy 
-the steady state effective energy Eq. \ref{eff_energy} because
-$\mathcal{C}\neq0$ for general solutions to
-Eq. \ref{steady_momr}---\ref{steady_momz}.
+the steady state effective energy Eq. \ref{eff_energy} because typically 
+$\mathcal{C}\neq0$.  
+%for general solutions to 
+%Eq. \ref{steady_momr}---\ref{steady_momz}.
 
-In practice, dust particles settle to the disk midplane on a timescale 
-$t_\mathrm{settle}\sim 1/\OmK^2\tstop$, although for tightly-coupled dust
-particles this is much longer than that to establish vertical
-hydrostatic equilibrium in the gas component ($\sim 1/\OmK$).  
-True steady states may be calculated by invoking an underlying gas
-turbulence, which leads to dust diffusion counter-acting vertical
-settling \citep{takeuchi02, youdin07}. This complication is beyond the
-scope of this paper.
+% Dust particles settle to the disk midplane on a timescale 
+% $t_\mathrm{settle}\sim 1/\OmK^2\tstop$, although for tightly-coupled dust
+% particles this is much longer than that to establish vertical
+% hydrostatic equilibrium in the gas component ($\sim 1/\OmK$).  
+% True steady states may be calculated by invoking an underlying gas
+% turbulence, which leads to dust diffusion counter-acting vertical
+% settling \citep{takeuchi02, youdin07}. This complication is beyond the
+% scope of this paper.
 
-In this work, consider equilibria with $\mathcal{C}\equiv 0$, namely   
+In order to have a well-defined stability problem, we formally
+take equilibria with $\mathcal{C}\equiv 0$, namely      
 \begin{inparaenum}[1)] 
 \item 
-  unstratified disks with finite dust-gas drag; %(\S\ref{si});  
+  unstratified or 2D disks with finite dust-gas drag; %(\S\ref{si});   
 \item 
@@
 \end{inparaenum} 
-Vertical settling is absent in both cases. Alternatively, we are 
-assuming that $\tstop$  is sufficiently small such 
-that settling occurs on much longer timescales than the instabilities of interest.
+Alternatively, we are assuming that $\tstop$  is sufficiently small
+such that the evolution of the background disk (e.g. dust-settling) 
+occur on much longer timescales than the 
+instabilities of interest.
 
@@
 
-\subsection{Vertical buoyancy}\label{vbuoyancy}
+\subsection{Dusty vertical buoyancy}\label{vbuoyancy}
 Having identified the entropy of an isothermal dusty gas, we find the
@@
 dusty-gas mixture using Eqs.\ref{masseq}--\ref{momeq} with 
-Eq. \ref{eff_energy} in place of the dust continuity equation. Our
-discussion is based on the corresponding axisymmetric linearized
-equations. A variable $X$ is
+Eq. \ref{eff_energy} in place of the dust continuity
+equation. Self-gravity is neglected here. Our discussion is based on
+the corresponding axisymmetric linearized equations. A variable $X$ is
 subject to Eulerian perturbations of the form  
@@
 where $s$ and $\omega$ are the growth rate and real frequencies,
-respectively. Thus perturbations have time dependence $e^{st +
-  \ii\omega t}$.
+respectively. Then perturbations have time dependence $e^{st +
+  \ii\omega t}$.  Thus for $\omega>0$, perturbations rotate
+anti-clockwise in the complex plane.
 
@@
 gas if one inserts $C_p\ln{\left(1 -
-    \tepsilon\right)}$ for the entropy in the standard
+    \tepsilon\right)}$ for the entropy into the standard
 expression for said criteria \citep{tassoul78}.
@@
 direction as the local pressure gradient, which is equivalent to the
-gas density gradient for isothermal gas. Thus to enable instability
-here we need to violate Eq. \ref{dusty_solberg2}. We explore this
-below.
+gas density gradient for isothermal gas. {\bf ref for drift} Thus for
+instability we need to violate Eq. \ref{dusty_solberg2}. We explore
+this below.
 
 \subsubsection{Vertically well-mixed dust}\label{vert_mixed_theory}
-Here we consider the special case where the dust-to-gas ratio is
-uniform in $z$. Then the LHS of Eq. \ref{dusty_solberg2} becomes
+In the special case where the dust-to-gas ratio is
+uniform in $z$ the LHS of Eq. \ref{dusty_solberg2} becomes
 \begin{align}
@@
 \end{align}
-Thus the disk is \emph{unstable}, since there is no effective vertical 
-buoyancy to stabilize vertical motions. This situation is
+Thus the disk is \emph{unstable}, because there is no effective vertical 
+buoyancy to stabilize vertical motions ($N_z=0$). This situation is 
 identical to the pure gas, adiabtic simulations of \cite{nelson13}
@@
 
-\subsubsection{Order-of-magnitude estimates} 
-In order to apply the stability criteria
-Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} to typical
-protoplanetary disks we conside the equilibria described in
-\S\ref{eqm}.
-
-We assume the disk is approximately Keplerian so that $\kappa\simeq\OmK$, and that the vertical gas distribution is
+\subsubsection{Application to protoplanetary disks} 
+%In order to apply the stability criteria
+%Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} to typical
+%protoplanetary disks 
+We now consider the specific equilibria described in 
+\S\ref{eqm}. We assume the disk is approximately Keplerian so that
+$\kappa\simeq\OmK$, and that the vertical gas distribution is   
 Gaussian. In terms of the dust-to-gas
-ratio $\epsilon$, Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2}
-become   
+ratio $\epsilon$, Eq. \ref{dusty_solberg2}
+becomes 
 \begin{align}
-  &1 + \frac{\epsilon}{\left(1+
-    \epsilon\right)^2}\left(h_\mathrm{g}^2\frac{\p\ln{\rhog}}{\p\ln{r}}
-  \frac{\p\ln{\epsilon}}{\p\ln{r}} + 
-  \frac{z^2}{\Htilde^2}\right)>0 \label{stability_est1},   \\ 
+  % &1 + \frac{\epsilon}{\left(1+
+  %   \epsilon\right)^2}\left(h_\mathrm{g}^2\frac{\p\ln{\rhog}}{\p\ln{r}}
+  % \frac{\p\ln{\epsilon}}{\p\ln{r}} + 
+  % \frac{z^2}{\Htilde^2}\right)>0 \label{stability_est1},   \\ 
 &1 - \frac{\epsilon
@@
 \end{align}
-\emph{for stability}.
+for \emph{stability}. The second term is usually stabilizing in 
+PPDs. The third term is always destabilizing, but is small in smooth,
+thin disks with radial gradients $O(1/r)$,  $h_\mathrm{g}\ll 1$,
+provided that $H_\epsilon/\Hg$ is not large (e.g. some dust settling
+has occured). This means that typical dusty PPDs are stable to
+axisymmetric perturbations, a result valid for arbitrarily thin dust
+layers.
 
-Eq. \ref{stability_est1}---\ref{stability_est2} are easily satisfied
-in thin accretion disks where radial gradients are $O(1/r)$, 
-$h_\mathrm{g}\ll 1$ and $\Htilde/H_\mathrm{g}$ not too
-large.  Then the
-magnitude of the second term on the left-hand-side  of either
-inequality is much less than unity. Thus radially smooth, strictly
-isothermal disks perfectly coupled with dust are stable against
-axisymmetric perturbations. This result holds for arbitrarily thin
-dust layers.
-
-%Such disks are stable against
-%axisymmetric perturbations.
-
-\subsubsection{Instability at dust edges}
-Notice the left-hand-side of Eq. \ref{stability_est2} is a quadratic in
-$\p_r\epsilon$. It is possible to
-violate this inequality for sufficiently large (in magnitude) radial
-gradients  in the dust-to-gas ratio, 
+To violate Eq. \ref{stability_est2} and obtain instability, notice the 
+LHS is a quadratic in $\p_r\epsilon$. Thus instability is possible for
+sufficiently large (in magnitude) radial gradients  in the dust-to-gas
+ratio,  
 \begin{align}
@@
 \end{align}
-\emph{for instability}, where
+for \emph{instability}, where
 \begin{align}\label{spm}
@@
 \end{align} 
-In typical accretion disks where $\p_r\rhog<0$, it is easier to
-achieve instability at a given radius for increasing dust-to-gas
-ratios in outwards ($\p_r\epsilon > 0$), and vice versa.
+In typical accretion disks where $\p_r\rhog<0$, instability is easier
+for increasing dust-to-gas ratios ($\p_r\epsilon > 0$).
 
@@
 then the system is potentially unstable.  
-
 Taking $\epsilon\sim 0.01$,  we find that 
@@
 requires $L_\epsilon\ll O(0.1\Hg)$, i.e. the dust-to-gas ratio must 
-vary on an extremely short lengthscale. 
-% %an extremely short  
-% %radial length-scale in the dust-to-gas ratio. 
-% This is unlikely to
-% occur. On the other hand, if dust is well-mixed, say $\Hd\simeq
-% 0.99\Hg$, then instability only requires 
-% $L_\epsilon\lesssim O(10\Hg)$, which is more realistic.
-
-%This is responsible for 
-%the usual vertical shear instability in locally isothermal disks.
-%However, we may obtain an integral
-%relation for linear, axisymmetric waves with frequency $\sigma$ (see
-%Appendix \ref{var_prin}, where we also account for finite $\tstop$) in
-%the form:   
-%If $c_s$ is constant, then $\sigma^2$ is real, and 
-%the first integral leads to the previous Solberg-Hoiland criteria for 
-%axisymmetric stability. However, when $c_s$ is non-uniform,
+vary on an extremely short lengthscale. This might be achieved, for
+example, at sharp edges associated with gaps opened by giant planets. 
+{\bf mention this is equivalent to sharp entropy gradients?}
 
@@
 Eq. \ref{int_rel} indicates that $\sigma^2$ is generally complex, so
-we may have growing oscillations. %Writing $\sigma =
-%-\omega + \ii s$ where $\omega$ and $s$ are the real frequency and
-%growth rates, respectively, 
-The imaginary part of Eq. \ref{int_rel}
-gives  
+we may have growing oscillations, or overstability, due to dust-gas
+friction. This is seen by taking the imaginary part of
+Eq. \ref{int_rel} 
 \begin{align}
@@
 \end{align} 
-assuming $\omega\neq0$. This represents overstability due to dust-gas friction. 
+assuming $\omega\neq0$. %This represents overstability due to dust-gas friction. 
 Since drag ($\delta \mathcal{C}$) appears as a source term in our
@@
 
-\subsubsection{Physical requirement of dust-gas drag overstabilities}
+\subsubsection{Physical property of dust-gas drag overstabilities}
 
@@
 \end{align*}
-meaning that gas follows dust. Instabilities are thus not possible if the
-gas does not respond to dust-drag (i.e. no back-reaction). The
-pressure-density lag shown in Fig. \ref{pdv_cartoon} is achieved if, 
-after a parcel reaches maximum total density, its 
-gas content is still increasing ($A$ to $B$). This in turn implies a sufficiently
-large particle flux \emph{out} of the parcel.
+meaning that gas follows dust (Fig. \ref{lag_cartoon}). Instabilities
+are thus not possible if the gas does not respond to dust-drag
+(i.e. no back-reaction). The pressure-density lag shown in
+Fig. \ref{lag_cartoon} is achieved if, after a parcel reaches maximum
+total density, its  gas content is still increasing (see $A$ to $B$ in
+Fig. \ref{pdv_cartoon}). This in turn implies a sufficiently large
+particle flux \emph{out} of the parcel.
 
@@
 density, but shows 
-that this \emph{must} be the case for any growing modes associated the
+that this \emph{must} be the case for any growing oscillations associated the
 dust-gas drag. To rigorously understand how dust-drag causes this 
@@
 $\mathcal{C}$ (see Appendix \ref{lin_dust}), we might generally expect
- dusty disks to  support a range of stable and unstable modes, with the
+ dusty disks to  support a range of stable and overstable modes, with the
 latter being associated with pressure-density lag. 
-
-Indeed, \cite{jacquet11} explains the essence of the streaming 
-instability in dusty protoplanetary disks as dust accumulation
-(and hence compression) at a pressure bump, which then drags the 
-gas towards it (and hence heating) to strengthen the pressure
-bump. That is, heating occurs upon compression, as in stellar  
-pulsational instabilities \citep{cox67}. In \S\ref{si} we check that the
+%Indeed, \cite{jacquet11} explains the essence of the streaming 
+%instability in dusty protoplanetary disks as dust accumulation
+%(and hence compression) at a pressure bump, which then drags the 
+%gas towards it (and hence heating) to strengthen the pressure
+%bump. That is, heating occurs upon compression, as in stellar  
+%pulsational instabilities \citep{cox67}.
+ In \S\ref{si} we check that the
 streaming instability fits into this thermodynamic interpretation in

### 2017-05-02 paper.tex
misc edits from sect 4 to sect 8.3

@@
 %  isothermal dusty gas: application to protoplanetary disks}  
-\title{A thermodynamic view of dusty protoplanetary disks I:
-axisymmetric stability}
+\title{A thermodynamic view of dusty protoplanetary disks}
  % \title{Dust-free modeling of dusty protoplanetary disks I:
@@
 \item 
- perfectly coupled dust cannot cause axisymmetric 
- instabilties, however thin the dust layer is; 
+ perfectly coupled, thin dust layers cannot cause axisymmetric 
+ instabilties; 
 \item 
@@
   that lags dust density; 
-\item dust-loading generally stabilizes the vertical shear
-  instability, because dust-gas coupling introduces buoyancy. 
+\item dust-loading introduce buoyancy forces that generally 
+  stabilizes the vertical shear 
+  instability that arises from a radial temperature gradient. 
 \end{inparaenum}  

### 2017-05-02 result.tex
misc edits from sect 4 to sect 8.3

@@
 
-We present numerical solutions for vertically stratified,
+We now present numerical solutions for vertically stratified,
 non-self-gravitating disks assuming perfectly coupled dust
-($\tstop=0$) so that the equilibria defined in \S\ref{eqm} 
-are exact steady states.
+($\tstop=0$) so the equilibria
+defined in \S\ref{eqm} are exact steady states.
 
-Except in \S\ref{varHd}---\ref{vert_mixed}, we consider constant midplane dust-to-gas
-ratios $\epsilon_0$ and characteristic thickness $H_\epsilon$. Then
-the dust-to-gas ratio $\epsilon=\epsilon(z)$. In this limit any
-growing modes must be associated with the imposed temperature gradient
+Apart from \S\ref{varHd}---\ref{vert_mixed}, we consider constant
+midplane dust-to-gas ratios $\epsilon_0$ and characteristic thickness
+$H_\epsilon$. Then the dust-to-gas ratio $\epsilon=\epsilon(z)$. In 
+this limit any growing modes must be associated with the imposed temperature gradient
 (\S\ref{limits}). We are thus examining the effect of dust-loading on
-the VSI, previously studied in pure gas disks
+the VSI previously studied in pure gas disks 
 \citep[][\citetalias{lin15} in this section]{lin15}.
@@
 layer thickness $\Hd(r_0)$, and the perturbation radial wavenumber
-$k_x$. We choose a power-law midplane gas density profile,
-$\rho_\mathrm{g0}\propto r^p$ with $p = -1.5$. The fiducial power-law
+$k_x$. We choose a midplane gas density profile 
+$\rho_\mathrm{g0}\propto r^{-3/2}$. The fiducial power-law 
 index for the temperature profile is $q=-1$; and we set the gas disk
@@
 
-Two boundary conditions are needed for the $\tstop=0$
-problem \citep[e.g.][]{lubow93}. We impose solid boundaries so that
-$\delta v_z(\pm\zmax)=0$. We solve the linearized equations as a
-generalized eigenvalue problem using a pseudo-spectral code adapted
+%Two boundary conditions are needed for the $\tstop=0$
+%problem \citep[e.g.][]{lubow93}. 
+We impose solid vertical boundaries so that $\delta
+v_z(\pm\zmax)=0$. We solve the linearized equations as a generalized
+eigenvalue problem using a pseudo-spectral code adapted  
 from \citetalias{lin15}. Amplitudes are expanded in Chebyshev
-polynomials up to order $N_z=512$. We check results using
-Eq. \ref{vsi_check}.
+polynomials up to order $N_z=512$. %We check results using
+%Eq. \ref{vsi_check}.
 
@@
 \citetalias{lin15} found the appropriate way to compare 
-destabilizing vertical shear to stabilizing vertical buoyancy is
-$r\p_z\Omega/\OmK$ against $N_z^2/\OmK^2$. We see from \S\ref{vertshear}
-and \S\ref{vbuoyancy} that $|r\p_z\Omega|\sim q h_\mathrm{g}\OmK$ for a
+vertical shear (destabilizing) and  vertical buoyancy (stabilizing)
+is $r\p_z\Omega/\OmK$  against $N_z^2/\OmK^2$. From 
+\S\ref{vertshear} and \S\ref{vbuoyancy} we find $|r\p_z\Omega|\sim q
+h_\mathrm{g}\OmK$ for a 
 thin, gas dominated disk; while $N_z^2\sim \epsilon\OmK^2$. Thus we
@@
   \caption{Vertical shear rate (solid) compared to vertical buoyancy
-    (dotted) in a locally isothermal, dusty disl with midplane dust-to-gas ratio
+    (dotted) in a locally isothermal, dusty disk with midplane dust-to-gas ratio
     of $\epsilon_0=10^{-3}$ (black) and $\epsilon_0=1$ (blue). 
@@
 `surface modes' \citep[which are associated with the imposed vertical
-boundaries][]{barker15}.
+boundaries, ][]{barker15}.
 
@@
 but their oscillation frequency increases with
-$\epsilon_0$. %i.e. increases with buoyancing freq 
+$\epsilon_0$. i.e. it increases with the vertical buoyancy. 
 The total number of modes do not change 
 significantly. This is in contrast with the effect of increasing
-cooling times, where \citetalias{lin15} find fewer unstable modes.
+cooling times, for which \citetalias{lin15} find fewer unstable modes.
 %until the system is eventually completely stable.
@@
 function of $\epsilon_0$ for different perturbation wavenumbers
-$k_x$. We find dust-loading stabilizes the VSI more effectively for shorter
+$k_x$. Dust-loading stabilizes the VSI more effectively for shorter
 wavelength perturbations. This is because for high wavenumbers the
 dominant modes are surface modes, which are effectively stabilized by
-dust-loading since buoyancy forces are larges near the disk
+dust-loading as buoyancy forces are largest near the vertical 
 boundaries. The figure suggest that VSI becomes much less efficient
@@
 
-\subsection{Effect of dust thickness} 
-We now vary the dust layer thickness $\Hd$ but fix the metalicity 
-$Z \equiv \epsilon_0 \Hd/\Hg = 0.03$. This then gives the
-required midplane dust-to-gas ratio. Since we will consider thin dust
-layers, here we use a smaller 
+\subsection{Effect of dust layer thickness} 
+We now vary $\Hd$ but fix the metalicity 
+$Z \equiv \epsilon_0 \Hd/\Hg = 0.03$ to obtain $\epsilon_0$. Since we
+will consider thin dust layers, here we use a smaller   
 domain with $\zmax=2\Hg$ so that $\epsilon$ does not become
@@
 the $\Hd=0.1\Hg$ disk, as expected from the larger buoyancy frequency
-as compared to vertical shear there. This leads to a structure
+relative to vertical shear there. This leads to a structure
 analogous to PPD dead zones: a quiescent midplane between active
@@
 Thus a thin dust layer, however large its associated vertical shear,
-does not affect the VSI. The non-monotonic behavior for $\Hd\gtrsim
+does not affect VSI growth rates. The non-monotonic behavior for $\Hd\gtrsim
 0.5\Hg$ arises because the vertical buoyancy frequency 
@@
 \subsection{Axisymmetric stability of ultra-thin dust layers}
-In \S\ref{iso_perfect} we asserted that in the limit of perfectly
-coupled dust: radially smooth, strictly
-isothermal disks are stable against axisymmetric perturbations, no
-matter how thin the dust layer is. We now demonstrate this
-numerically.
+The discussion in \S\ref{iso_perfect} imply strictly  
+isothermal disks with a radially uniform dust-to-gas ratio are
+stable against axisymmetric perturbations, no matter how thin the dust
+layer is. We now demonstrate this numerically.
 
-For comparison with previous studies of dusty disks, we also use 
+To connect with previous studies, here  we also use 
 the Richardson number $\rich \equiv N_z^2/\left(r\p_z\Omega\right)^2$
-to label the calculations here \citep{youdin02}.   
+to label calculations \citep{youdin02}.   
 Numerical simulations show    
@@
 metalicity $Z=0.01$ so the midpane dust-to-gas ratio $\epsilon_0$ is 
-$O(1)$.  We consider radial wavenumbers such $k_x\Hd=1$.
+$O(1)$.  We consider radial wavenumbers with $k_x\Hd=1$.
 
@@
 \S\ref{vertshear}). However, we see that growth rates $s\propto
-|q|$, i.e .growth rates vanish if the disk was strictly isothermal. 
+|q|$, i.e .growth rates vanish in the strictly isothermal limit.  
 In particular, this holds for $\rich < 0.1$, the critical value 
@@
 
-Axisymmetric instability is associated with the thermal
+Axisymmetric instability here is associated with the thermal
 contribution to vertical shear: as $q\to0$,  $\p_z\Omega$ becomes
@@
 
-Here we allow the dust-to-gas ratio $\epsilon$ to depend on
-$r$. Specifically we let $\epsilon_0\propto r^{-1}$ and
-$H_\epsilon \propto \Hg$ (cf. constant values in the previous
-calculations). Then
+We now consider a radially-varying dust-to-gas ratio. Specifically we
+let $\epsilon_0\propto r^{-1}$ and $H_\epsilon \propto \Hg$
+(cf. constant values in the previous calculations). Then
 \begin{align*} 
@@
 \end{align*}
-We fix $Z=0.01$, $\Hd=0.8\Hg$ at $r=r_0$, and
-$k_x\Hg=30$.
+Here we fix $Z=0.01$, $\Hd=0.8\Hg$, and $k_x\Hg=30$.
 
@@
   \caption{Unstable VSI modes in the disk models of
-    Fig. \protect\ref{compare_vshear_varHd}.  
+    Fig. \protect\ref{compare_vshear_varHd}.  Black (blue) diamonds
+    are mode frequencies for a disk with radially uniform (varying)
+    dust-to-gas ratio. 
     \label{vsi_dust_varHd}
@@
 disks, a vertically-uniform dust-to-gas ratio can be unstable
-if $d\epsilon/dr\neq0$. To demonstate this numerically we consider the
+if $d\epsilon/dr\neq0$.  
+To demonstate this numerically we consider the
 same disk as above but with $q=0$ and choose $H_\mathrm{d}$ such
-that $H_\epsilon=10^3\Hg$. Then vertical shear $\p_z\Omega$ is attributed
-to the radially-varying dust-to-gas ratio (see Eq. \ref{vshear2}).
+that $H_\epsilon=10^3\Hg$. Then vertical shear $\p_z\Omega$ is then
+attributed to the radially-varying dust-to-gas ratio (see
+Eq. \ref{vshear2}).
 
@@
 \S\ref{vert_mixed_theory}, the disk admits purely growing modes with
-$\omega=0$.
+$\omega=0$. This is distinct from the growing oscillations associated
+with classic VSI discussed above.
 
@@
   \includegraphics[width=\linewidth]{figures/vert_mixed_modes} 
-  \caption{Unstable modes in a strictly isothermal disk ($q=0$) with
+  \caption{Purely growing modes in a strictly isothermal disk ($q=0$) with
     vertically uniform dust-to-gas ratio. The instability is due to 
@@
 \subsection{Validity}
+{\bf maybe incorporate into intro to this section?}
 We considered $\tstop=0$ in order to perform stability analysis on an

### 2017-05-02 summary.tex
misc edits from sect 4 to sect 8.3

@@
 The addition of dust enhances gravitational
-instability (GI). This is because dust particles contribute to the
+instability (GI) because dust particles contribute to the
 total disk mass but not thermal pressure, which effectively lowers the
@@
 $\epsilon, Z\ll1$, this effect is unimportant. However, if $\epsilon$ is
-large (e.g. due to dust settling) then the reduced temperature $\tilde{T}$ may 
-be lowered to enable instability.
+large (e.g. due to dust settling) then the reduced temperature
+$\widetilde{T}$ may be lowered to enable instability.
 
 As noted in \S\ref{loc_iso_eos}, dust settling causes 
-$\tilde{T}$ to \emph{increase} away from the midplane. This contrasts
+$\widetilde{T}$ to \emph{increase} away from the midplane. This contrasts
 to previous studies of GI in vertically stratified disks
@@
 characteristic temperatures are relevant to stability  
-\citep{toomre64}, a non-trivial effective vertical temperature
+\citep{toomre64}, a non-trivial vertical temperature
 structure, induced by dust, may modify the vertical structure of 3D
-waves and unstable modes. We will examine this linear problem in a following 
+waves and unstable modes. We will examine this problem in a following 
 study.
@@
 potential vorticity 
+%\begin{align}
+ $\mathcal{V}_\mathrm{g} =
+ \kappa^2\mathcal{S}_\mathrm{g}^{-2/\gamma}/2\Omega\Sigma_\mathrm{g}$, 
+%\end{align} 
+where $\mathcal{S}_\mathrm{g} = P/\Sigma_\mathrm{g}^\gamma$. Here $P$
+should be interpreted as the vertically integrated pressure. The
+non-linear result of the RWI is vortex formation \citep{li01}.
+
+The dusty/adiabatic gas analogy imply the condition for RWI in a
+polytropic dusty gas is given by extrema in  
 \begin{align}
- \mathcal{V} = \frac{\kappa^2}{2\Omega\Sigma}\mathcal{S}^{-2/\gamma},  
+  \mathcal{V} =
+  \frac{\kappa^2}{2\Omega\Sigma}\frac{1}{K^{2/\Gamma}}\left(1+
+    \frac{\Sigma_\mathrm{d}}{\Sigma_\mathrm{g}}\right)^2.    
 \end{align}
-where $\mathcal{S} = \Pi/\Sigma^\gamma$. Here, $\Sigma$ is the total
-surface density and $\Pi$ is the vertically integrated pressure.
-
-From the dusty/adiabatic gas equivalence, we deduce the corresponding
-condition for RWI in a (locally isothermal or polytroipc) dusty gas is
-also given by extrema in $\mathcal{V}$ with $\mathcal{S} = \exp{\left(\Gamma
-S_\mathrm{eff}\right)}$ and $\tepsilon = \Sigma_\mathrm{d}/\Sigma$. Since
-the effective entropy reflects the dust distribution, we expect that
-RWI may also be triggered by extrema in the dust-to-gas ratio, e.g. 
-narrow dust rings. This would lead to direction formation of dusty
-vortices.
+Thus we expect that RWI may also be triggered by extrema 
+in the dust-to-gas ratio, e.g. narrow dust rings/gaps. This may lead
+to direct formation of dusty vortices, as opposed to dust-trapping by  
+a pre-existing gas vortex \citep{barge95,lyra13}.
 
@@
 \subsubsection{Convective overstability}
-The `convective overstability' (ConO) operates in dust-free, non-adiabatic disks 
-when there is an unstable radial buoyancy frequency ($N_r^2 < 0 $)
-combined with finite thermal relaxation time, $t_\mathrm{cool}\sim
-\OmK^{-1}$. This lead to growing epicycles
+The `convective overstability' (ConO) operates in pure gas, non-adiabatic disks 
+when there is a radial buoyancy frequency such that $N_r^2\equiv
+F_r\p_rS/C_P  < 0 $ and cooling times $t_\mathrm{cool}\sim \OmK^{-1}$. This
+combination leads to growing epicycles
 \citep{klahr14,lyra14,latter16}.     
@@
 
-The conditions required for ConO might be realized in a locally
-isothermal dusty disk. Since $\p_rP < 0$ in typical disk models, 
-if the dust-induced radial entropy gradient is negative, then
-$N_r^2  =  F_r \p_r\seff < 0$. In 
-fact, the typical disk models used to study the streaming
-instability have zero background radial density gradient, in which
-case  $N_r^2= -\rho F_r^2/P$.
+%The disk structure required for ConO might be realized in a locally
+%isothermal dusty disk.
 
-Locally isothermal disks have $\tcool=0$, so the pure gas ConO cannot
+Now consider a strictly isothermal dusty disk where the relevant 
+entropy is dust-induced, $\seff =-\ln{(1+\rhod/\rhog)}$. 
+%The dusty/adiabatic gas
+%analogy implies the replacement $S\to\seff = -\ln{(1+\rhod/\rhog)}$ in
+%$N_r$. 
+Since $F_r\propto -\p_rP > 0$ in typical disk models, ConO would
+require $\p_r\seff<0$, implying the dust-to-gas ratio increases
+\emph{outwards}. This might be realized at special locations in
+protoplanetary disks.
+
+%fact, the typical disk models used to study the streaming
+%instability have zero background radial density gradient, in which
+%case  $N_r^2= -\rho F_r^2/P < 0$.
+
+Isothermal disks have $\tcool=0$, so the pure gas ConO cannot
 exist. However, we have shown that dust-gas drag provides an effective
-energy source/sink to the mixture. It would be
+energy source/sink to the mixture (\S\ref{energy_analogy}). It would be
 interesting to explore whether or not dust-gas drag can provide the 
@@
 The `zombie vortex instability' (ZVI) is a non-axisymmetric,
-non-linear (or finite-amplitude) instability in gas disks
-\citep{marcus15}. It operates by exciting `critical layers' which role
-up into vortices, exciting further critical layers. The result is an
-array of vortices. These critical
+non-linear instability discovered in pure gas disks
+\citep{marcus15,umurhan16d}. It operates by having finite-amplitude perturbations
+exciting `critical layers' which role 
+up into vortices, exciting further critical layers. The process
+repeats itself and the result is an array of vortices. These critical 
 layers are locations where the intrinsic wave frequency of a 
@@
 
-Since buoyancy is critical for the ZVI, it requires an almost
-purely adiabatic gas ($\tcool\OmK\gg 1$). This limits its
-applicability to the inner most parts of typical protoplanetary disks
-\citep{lesur16}.
+Since the necessary physical ingredient for the ZVI is vertical
+buoyancy, it requires an almost purely adiabatic gas ($\tcool\OmK\gg
+1$). This limits its applicability to typical protoplanetary disks at
+$\lesssim 1$AU \citep{lesur16, malygin17}. However, this consideration
+assumes a dust-free disk.
 
-However, we have show that dust-loading an isothermal disk also
-induces an effective buoyancy. This effect is maximized when dust is
+On the other hand, we have show that dust-loading an isothermal disk
+also induces an effective buoyancy. This effect is maximized when dust is
 perfectly coupled to the gas. It is thus natural to ask whether or not
 the required adiabatic conditions for ZVI can be realized through
-dust-loading, and thus produce a `dusty ZVI'.
+dust-loading, and thus produce a `dusty ZVI'. This may allow the ZVI
+to operate at $\gtrsim 1$AU.
 
@@
 The source term associated with dust-gas drag, $\mathcal{C}$, is 
-analogous to radiative diffusion \citep{price15}, which is also
-featured in many hydrodynamics codes.
+analogous to radiative diffusion \citep{price15} or thermal
+conduction, which is also common in modern hydrodynamics codes.
+
+We have taken advantage of the dusty/adiabatic gas equivalence to 
+convert the popular
+\textsc{Pluto}\footnote{\url{http://plutocode.ph.unito.it/}}      
+hydrodynamics code into a dusty gas dynamics code appropriate for
+simulating protoplanetary disks coupled to small dust. We will apply 
+our modified code --- d\textsc{Pluto} --- to the streaming instability
+and disk-planet interactions in an upcoming publication (Lin et al.,
+in preparation).
 
 \subsection{Higher order corrections}
-%{\bf notes for self. maybe exclude in final. }
+{\bf notes for self. maybe exclude in final. }
 
@@
 \section{Summary}\label{summary}
-In this paper we develop an analogy between isothermal 
+In this paper we develop a physical analogy between isothermal 
 dusty-gas and adiabatic, pure gas. The correspondence arises
@@
 the gas. Then the dust-to-gas ratio $\rhod/\rhog$ is conserved
-following the flow. This is analogous to entropy being conserved
+following the flow. This property is analogous to entropy being conserved
 following a adiabatic, pure gas.

### 2017-05-02 summary.tex
misc edits to summary

@@
 \citep{lovelace99,li00}. These studies consider adiabatic pure gas and 
-show instability is possible if there is an extrema in the generalized
+show instability is possible if there is an extremum in the generalized
 potential vorticity 
@@
 The dusty/adiabatic gas analogy imply the condition for RWI in a
-polytropic dusty gas is given by extrema in  
+polytropic dusty gas disk is an extremum   
 \begin{align}
@@
 \citep{marcus15,umurhan16d}. It operates by having finite-amplitude perturbations
-exciting `critical layers' which role 
+exciting `critical layers' which roll 
 up into vortices, exciting further critical layers. The process
@@
 buoyancy, it requires an almost purely adiabatic gas ($\tcool\OmK\gg
-1$). This limits its applicability to typical protoplanetary disks at
+1$). This limits its applicability in typical protoplanetary disks to 
 $\lesssim 1$AU \citep{lesur16, malygin17}. However, this consideration
@@
 On the other hand, we have show that dust-loading an isothermal disk
-also induces an effective buoyancy. This effect is maximized when dust is
-perfectly coupled to the gas. It is thus natural to ask whether or not
+also induces an effective buoyancy. % This effect is maximized when dust is
+% perfectly coupled to the gas
+. It is thus natural to ask whether or not
 the required adiabatic conditions for ZVI can be realized through
 dust-loading, and thus produce a `dusty ZVI'. This may allow the ZVI
-to operate at $\gtrsim 1$AU.
+to operate at $\gtrsim 1$AU in PPDs.
 
@@
 gas dynamics codes can be applied to simulate locally
-isothermal/polytropic dusty gas. One need not add any new 
-equations to evolve the dust component: 
+isothermal/polytropic dusty gas. One need not implement a separate 
+equation for the dust component: 
 %In fact, no
@@
  the standard energy equation substitutes for the dust continuity
-equation. When $c_s$ or $K$ is not constant and/or $\tstop\neq0$ one
+equation. When $c_s$ (or $K$) is not constant and/or $\tstop\neq0$ one
 should add corresponding source terms in the energy equation
@@
 \subsection{Higher order corrections}
-{\bf notes for self. maybe exclude in final. }
+{\bf notes for self. maybe exclude in final. maybe move into appendix}
 
-The one-fluid formulation, 
+The simplified one-fluid formulation, 
 Eqs. \ref{masseq}---\ref{tempeq}, can 
@@
 
-In a following paper we will include $\bm{T}$ in the one-fluid
-equations and futher explore the analogy between dusty-gas and viscous
-hydrodynamics.
+% In a following paper we will include $\bm{T}$ in the one-fluid
+% equations and futher explore the analogy between dusty-gas and viscous
+% hydrodynamics.
 
@@
 \section{Summary}\label{summary}
-In this paper we develop a physical analogy between isothermal 
+In this paper we develop the analogy between isothermal 
 dusty-gas and adiabatic, pure gas. The correspondence arises
@@
 
-For finite dust-gas drag, $\tstop\neq 0$, the dust content of a 
-parcel of the mixture is no longer conserved. The parcel 
+For finite drag, $\tstop\neq 0$, the dust content of a 
+parcel of the dusty-gas mixture is no longer conserved. The parcel 
 can exchange dust particles with neighboring parcels. % owing 
 %to dust-gas relative drift. 
-This is analogous to heat exchange between a parcel of pure gas and its surroundings.
+This is analogous to heat exchange between a parcel of pure gas and
+its surroundings.
 
 We explicitly show that for a fixed gas equation of state, the  
-evolutionary equation for $\rhod/\rhog$ may be replaced by an 
+evolutionary equation for the dust-to-gas ratio $\rhod/\rhog$ may be replaced by an 
 effective energy equation. This leads to a 
@@
 %dusty gas. 
-Finite dust-gas friction then appears as an energy
+The effect of finite dust-gas friction appears as an energy
 source term.  This analogy with standard 
-hydrodynamics with cooling/heating allow us to obtain dusty analogs of gaseous
+hydrodynamics with cooling/heating allow us to find dusty analogs of gaseous
 instabilities, and provide thermodynamical interpretations of  
@@
 
-We obtain the Solberg-Hoiland criteria for axisymmetric stability for
-strictly isothermal, perfectly-coupled dusty gas.  
+We obtain an analog of the Solberg-Hoiland criteria for
+axisymmetric stability for strictly isothermal, perfectly-coupled dusty gas.  
 Applying this to typical protoplanetary disks, we find that 
@@
   instabilities, however thin the dust layer is.    
-%This is consistent with the fact that only 
-%\emph{non-axisymmetric} Kelvin-Helmholtz instabilities have been 
-%observed in numerical simulations in the perfectly-coupled limit 
-%\citep{chiang08,barranco09,lee10}. 
 Instead, sharp radial  edges in the dust-to-gas ratio could destabilized the
 disk, as these imply sharp gradients in the disk's effective entropy
-profile. % {\bf presumably this limits how sharp dust rings can be}
+profile. Alternatively, if the dust is vertically well-mixed, then any
+radial gradient in $\rhod/\rhog$ can destabilize the disk.
+
+ % {\bf presumably this limits how sharp dust rings can be}
 
@@
 instability \citep[VSI, ][]{nelson13,lin15} to dusty disks. We
-explicitly show that the SI occurs because the evolution of gas
-pressure lags behind  dust density; a general property of 
+explicitly show in SI the evolution of gas
+pressure lags behind dust density; and is in fact a general property of 
 overstabilities driven by strong dust-gas drag. 
@@
 the dust motion. A lag implies there exists a time interval where 
-the gas pressure of a parcel of the mixture is increasing whilst dust is already being expelled.  
-The dusty gas then does positive work, which goes into amplifying
-oscillation amplitudes. This interpretation is analogous to stellar pulsational
+the gas pressure of a parcel of the dusty-gas mixture is increasing whilst dust is already being expelled.  
+The dusty gas then does positive work that amplifies 
+oscillations. This interpretation is analogous to stellar pulsational
 instabilities \citep{cox67}.
@@
 vertical buoyancy dominates over vertical shear, consistent with our
-previous study \citep{lin15}.
-Since the dust-induced buoyancy forces increase away from the midplane, we find dust-loading can stabilize `surface modes' of the VSI, that would otherwise have the largest growth rates.
+previous study \citep{lin15}. Since the dust-induced buoyancy forces
+increase away from the midplane, we find dust-loading can stabilize
+`surface modes'  of the VSI, that would otherwise have the largest
+growth rates.
 

### 2017-05-08 model.tex
extra intro section

@@
 \end{align}
-where $(r,\phi, z)$ are cylindrical co-ordinates centered on the star.   
-The second equality in Eq. \ref{thin_disk_potential} is the 
+where $(r,\phi, z)$ are cylindrical co-ordinates centered on the star, and 
+$G$ is the gravitational constant. The second equality is the  
 thin-disk approximation for the stellar potential, appropriate for
@@
 completeness, but we will mostly neglect it unless stated
-otherwise. In such cases $\psi$   
-satisfies  
+otherwise. In such cases $\psi$ satisfies the Poisson equation 
 \begin{align}
-  \nabla^2\psi = 4 \pi G \rho,
+  \nabla^2\psi = 4 \pi G \rho.
 \end{align}
-where $G$ is the gravitational constant.
 

### 2017-05-08 paper.tex
extra intro section

@@
 
-This paper is organized as follows. We first
-explain in \S\ref{grow_osc} a general result of fluid dynamics that
-underpins the rest of our study. We describe how 
-a pressure-density phase lag in fluids can lead to 
-instability. Consequently, an analogy between dusty gas and  
-pure hydrodynamics allow us to interpret dusty-gas instabilities 
-within a thermodynamics framework.
+This paper is organized as follows. In \S\ref{key_idea} we first
+describe, without derivation, the analogy between dusty gas and    
+ideal hydrodynamics. We then discuss in 
+\S\ref{grow_osc} a general result of fluid dynamics that
+underpins the bulk of our study: intabilities due to 
+pressure-density phase lags. This means that dusty-gas phenomena, such
+as SI, have thermodynamical interpretations.
 
@@
 \S\ref{discussion} before summarizing in \S\ref{summary}.
+
+\section{Dust-free modeling of dusty protoplanetary disks}\label{key_idea}
+{\bf is this section needed?} 
+The key concept we introduce in this work is the description
+of dusty protoplanetary disks \emph{without} having dust as an
+explicit variable. This is best illustrated by writing down our model  
+equations that describe a mixture of isothermal gas and small
+dust particles: 
+\begin{align}  
+  &\frac{\p \rho}{\p t} + \nabla\cdot\left(\rho\bm{v}\right)= 0, \label{mass0}\\ 
+  &\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} = -
+  \frac{1}{\rho}\nabla  P - \nabla \Phi_*,\label{mom0}\\ 
+ & \frac{\p P}{\p t} + \bm{v}\cdot\nabla P 
+  = - P \nabla\cdot\bm{v}
+                + \mathcal{C}(\rho, P; \tstop)\label{en0}, 
+\end{align}
+where $\rho=\rhod + \rhog$ is the total density, $P$ is the pressure,
+$\Phi_*$ is the stellar gravitational potential,   
+and $\mathcal{C}$ represent effects of dust-gas drag, which
+is paremterized by the stopping time $\tstop$.
+
+Eq. \ref{mass0}---\ref{en0} appear identical to adiabatic gas dynamics, even  
+though they physically represents a two-phase, dusty-gas. The main
+difference is the term $\mathcal{C}$. Here it does not represent
+physical cooling (as the system is isothermal). Instead, dust-gas drag
+causes particles to drift, hence by momentum conservation a gas drift
+in the opposite direction, which changes the pressure distribution.
+
+Since the above equations are similar to ideal gas dynamics, it 
+means that many of the results derived for ideal hydrodynamics can be  
+generalized to dusty disks without additional effort. This includes 
+the important effect of phase lags between density and pressure in
+fluids, discussed next.
 
@@
 and there is no work done. However, for finite dust-gas drag, $\rhod$ and
-$\rhog$ are not neccessarily in phase, because dust particles can drift relative to the gas. 
+$\rhog$ are not neccessarily in phase, because dust particles can
+drift relative to the gas.   
 \emph{If} this causes $P$ to lag 

### 2017-05-08 summary.tex
extra intro section

@@
 \subsubsection{Convective overstability}
-The `convective overstability' (ConO) operates in pure gas, non-adiabatic disks 
-when there is a radial buoyancy frequency such that $N_r^2\equiv
+The `convective overstability' (ConO) was discovered in pure gas, non-adiabatic, unstratified disk models    
+where the radial buoyancy frequency is such that $N_r^2\equiv
 F_r\p_rS/C_P  < 0 $ and cooling times $t_\mathrm{cool}\sim \OmK^{-1}$. This
@@
 require $\p_r\seff<0$, implying the dust-to-gas ratio increases
-\emph{outwards}. This might be realized at special locations in
-protoplanetary disks.
+\emph{outwards}. This might be realized at special radial locations in 
+protoplanetary disks, such as planet gaps. Dust-settling itself may
+also lead to the mid-plane dust-to-gas ratio to increase outwards
+\citep{takeuchi02}.
 
@@
 interesting to explore whether or not dust-gas drag can provide the 
-required `cooling' to enable a dusty version of the convective 
-overstability.
+required `cooling' to enable a `dusty convective 
+overstability'.
 
@@
 the required adiabatic conditions for ZVI can be realized through
-dust-loading, and thus produce a `dusty ZVI'. This may allow the ZVI
-to operate at $\gtrsim 1$AU in PPDs.
+dust-loading, and thus produce `dusty ZVI'. This may allow the zombie
+vortices to exist at $\gtrsim 1$AU in PPDs.
 

### 2017-05-09 linear.tex
home/work sync, typo correction in two-fluid frequency for SI
misc edits to the main text overall

@@
 term is then $-\delta\mathcal{C}/\rho = \tstop\tepsilon c_s^2k_x^2
-Q$. The thin-disk solution to Eq. \ref{lin_sg} is $\dd\psi = -2\pi G
-\Sigma W/\left|k_x\right|$, where $\Sigma$ is the total surface
-density.
+Q$. The thin-disk solution to Eq. \ref{lin_sg} is $\dd\psi(z=0) = -2\pi G
+\Sigma W/\left|k_x\right|$.%, where $\Sigma$ is the total surface
+%density.
 
@@
 %Eq. \ref{streaming_mass}---\ref{streaming_vz}. 
-Interestingly, we find that for cases with
+Interestingly, we find in dust-rich disks with 
 $\epsilon>1$, Eq. \ref{si_check} can be satisified with $\Delta
@@
 is responsible for growth. Conversely, for `linB' with $\epsilon
-< 1$, one can approximate $\Delta P \simeq \dd P$ in Eq. \ref{si_check}. This
-suggests that for SI in gas-dominated disks, the distinction between
-Eulerian and Lagragian pressure perturbations is unimportant. 
+< 1$, one can approximate $\Delta P \simeq \dd P$ in Eq. \ref{si_check}. %This
+%suggests that for SI in gas-dominated disks, the distinction between
+%Eulerian and Lagragian pressure perturbations is unimportant. 
 %{\bf but weak growth in this case}
@@
 %where we have applied the isothermal equation of state for the second equality.
-then $\varphi > 0 $ indicates pressure lagging behind density, which
-is true for all the cases. In particular, `linB' indicates that growth
-rates vanish as $\varphi \to 0$.  
-Thus the streaming instability indeed stems  
-from the gas pressure evolution lagging behind that of the dust 
-density.
+then $\varphi > 0 $ indicates gas density lagging behind dust density,
+which is true for all the cases. In particular, `linB' indicates that
+growth rates vanish as $\varphi \to 0$.  Thus the streaming
+instability is indeed associated with a phase lag between gas and dust
+density evolutions.
 
@@
 \startdata
- linA, \cite{youdin07b} &  $0.1$       & 3.0 & 30    & $-0.3480 +
+ linA, \cite{youdin07b} &  $0.1$       & 3.0 & 30    & $0.3480 +
  0.4190\ii$ & $0.3640 + 0.4249\ii$ & $0.90$ & $30\degr$\\
 
-linB, \cite{youdin07b} & $0.1$        &  0.2 & 6 & $0.4999 +
+linB, \cite{youdin07b} & $0.1$        &  0.2 & 6 & $-0.4999 +
 0.0155\ii$&   $-0.4981 + 0.0054\ii$  & $1.54$ & $1.2\degr$ \\

### 2017-05-09 model.tex
misc edits to the main text overall

@@
 \end{align}
-\citep{laibe14}, where $D/Dt \equiv \p_t + \bm{v}\cdot\nabla$ is the Lagragian
+\citep{laibe14,price15}, where $D/Dt \equiv \p_t + \bm{v}\cdot\nabla$ is the Lagragian
 derivative \emph{following the mixture} with center-of-mass velocity  
@@
   \Lambda = \frac{T -
-    T_\mathrm{ref}}{t_\mathrm{cool}}, \label{realenergy} 
+    T_\mathrm{ref}}{t_\mathrm{cool}}, \label{cooling} 
 \end{align}
@@
 $G$ is the gravitational constant. The second equality is the  
-thin-disk approximation for the stellar potential, appropriate for
+thin-disk approximation, appropriate for
 $|z|\ll r$. We shall adopt this approximate potential in order to
@@
   \frac{\p P}{\p t} + \bm{v}_\mathrm{g}\cdot\nabla P = - \gamma P
-  \nabla\cdot \bm{v}_\mathrm{g}  + \mathcal{H} 
+  \nabla\cdot \bm{v}_\mathrm{g}  + \mathcal{H}\label{real_pressure}
 %\frac{\rho \mathcal{R}}{\mu}\frac{T
@@
 T_\mathrm{ref}\left(1-\tepsilon\right)$. %sound-speed $c_{s,\mathrm{eff}} = c_s\sqrt{\left(1 -
-This reduced temperature is lowered with dust-loading. 
+This reduced temperature decreases with dust-loading. 
 %   \tepsilon\right)}$. 
@@
 \subsection{Effective energy equation}\label{energy_analogy}
-Although we have eliminated the need of an evolutionary energy
-equation (Eq. \ref{realenergy}) by choosing a locally isothermal
+Although we have eliminated the true energy
+equation (Eq. \ref{tempeq}) by fixing a locally isothermal
 equation of state (Eq. \ref{eos}), we show that the mixture 
-nevertheless obeys an effective evolutionary energy equation. This is
+nevertheless obeys an effective energy equation. This is
 because advection of the dust-fraction, described by
 Eq. \ref{dusteq}, can be transformed into an energy-like 
-equation. We 
+equation. 
 %We show that for a prescribed temperature distribution, the mixture
@@
 %dust-fraction. 
-eliminate $\tepsilon$ in favor of $P$, 
+Eliminating $\tepsilon$ in favor of $P$, 
 \begin{align*}
@@
 \end{align}
-Comparing with Eq. \ref{realenergy}, we see that 
+Comparing with Eq. \ref{real_pressure}, we see that 
 Eq. \ref{eff_energy} is the energy equation for an ideal gas of
@@
 role of radiative cooling ($\Lambda$ in 
-Eq. \ref{realenergy}).
+Eq. \ref{tempeq}).
 
@@
 
-An effective energy equation can also be derived for other 
-fixed gas equations of state, such as polytropic gas. We discuss this
-in \S\ref{gen_poly}.
+%An effective energy equation can also be derived for other 
+%fixed gas equations of state, such as polytropic gas. We discuss this
+%in \S\ref{gen_poly}.
 
@@
 %}
-
-The terminal velocity approximation applies to short
-stopping times, $\tstop\OmK\ll 1$; while the locally isothermal
-approximation applies to short cooling times, $t_\mathrm{cool}\OmK\ll
+The above correspondence between isothermal dusty gas and adiabatic
+pure gas is derived under the terminal velocity approximation, which applies
+to short stopping times, $\tstop\OmK\ll 1$; and the locally isothermal
+approximation, which applies to short cooling times, $t_\mathrm{cool}\OmK\ll
 1$. In typical protoplanetary disk
@@
 $t_\mathrm{cool}$ is generally only small in the outer disk
-\citep{lin15,malygin17}. Combining these results, we estimate that our
-framework is best suited for mm (or smaller) sized particles at a few
+\citep{lin15,malygin17}. Combining these results, we estimate the 
+correspondence applicable to particles less than mm in size at a few
 to 10s of AU in protoplanetary disks. However, note that the
@@
 In order to have a well-defined stability problem, we formally
-take equilibria with $\mathcal{C}\equiv 0$, namely      
+consider equilibria with $\mathcal{C}\equiv 0$. Namely      
 \begin{inparaenum}[1)] 
@@
 longer `weighed down' by the dust, and can slip past it. Thus finite
-drag diminishes the dust-induced buoyancy.
+drag diminishes the dust-induced buoyancy. This is similar to reducing
+gas buoyancy through cooling \citep{lin15}.
 
@@
 In this case, the last two integrals in Eq. \ref{int_rel} vanish. 
-The condition for axisymmtric \emph{stability}, $\sigma^2>0$, 
+Then the condition for axisymmtric \emph{stability}, $\sigma^2>0$, 
 is met if the integrand cofficients satisfy $A+D>0$ and $AD-B^2>0$
@@
 direction as the local pressure gradient, which is equivalent to the
-gas density gradient for isothermal gas. {\bf ref for drift} Thus for
+gas density gradient for isothermal gas. Thus for 
 instability we need to violate Eq. \ref{dusty_solberg2}. We explore
@@
 In the special case where the dust-to-gas ratio is
-uniform in $z$ the LHS of Eq. \ref{dusty_solberg2} becomes
+uniform in $z$ the left-hand-side of Eq. \ref{dusty_solberg2} becomes
 \begin{align}
@@
 for \emph{stability}. The second term is usually stabilizing in 
-PPDs. The third term is always destabilizing, but is small in smooth,
-thin disks with radial gradients $O(1/r)$,  $h_\mathrm{g}\ll 1$,
+PPDs for reasons given in \S\ref{iso_perfect}. The third term is
+always destabilizing, but is small in smooth, thin disks with radial
+gradients $O(1/r)$, $h_\mathrm{g}\ll 1$, 
 provided that $H_\epsilon/\Hg$ is not large (e.g. some dust settling
 has occured). This means that typical dusty PPDs are stable to
-axisymmetric perturbations, a result valid for arbitrarily thin dust
-layers.
+axisymmetric perturbations, a conclusion valid for arbitrarily thin
+dust layers.
+
+%%%%%%%%%%%%%%%%%%
 
@@
 P\right)<0$.  This corresponds to pressure perturbations \emph{lagging}
-behind that in density (or displacements), see Fig. \ref{lag_cartoon}
+behind that in density, see Fig. \ref{lag_cartoon}
 for the case of dusty gas.
@@
 meaning that gas follows dust (Fig. \ref{lag_cartoon}). Instabilities
-are thus not possible if the gas does not respond to dust-drag
-(i.e. no back-reaction). The pressure-density lag shown in
-Fig. \ref{lag_cartoon} is achieved if, after a parcel reaches maximum
-total density, its  gas content is still increasing (see $A$ to $B$ in
-Fig. \ref{pdv_cartoon}). This in turn implies a sufficiently large
-particle flux \emph{out} of the parcel.
+are thus not possible if the gas does not respond to dust 
+(i.e. no back-reaction).% {\bf can we say this means that any effective
+%  dust-drag instability require high dust to gas ratios?} 
+The pressure-density lag shown in
+Fig. \ref{lag_cartoon} is achieved if, just after the total density of a 
+parcel maximizes, its gas content
+is increasing, which requires a sufficiently large
+particle flux \emph{out} of the parcel. See $A$ to $B$ in 
+Fig. \ref{pdv_cartoon}.
 

### 2017-05-09 paper.tex
misc edits to the main text overall

@@
   This is analogous to entropy conservation
-  following an adiabatic gas.  
-  Motivated by this similarity, we develop a thermodynamical
+  following an ideal gas.   
+  Motivated by this similarity, we develop a thermodynamic 
   framework to study dusty gas dynamics in protoplanetary disks. We  
@@
   regarded as an effective 
-  heating/cooling. We exploit this equivalence 
+  heating/cooling. We exploit this correspondence 
   to deduce that   
@@
 \item 
-  the streaming instability neccessarily involves a pressure response
-  that lags dust density; 
+  the streaming instability neccessarily involves a gas pressure
+  response 
+  that lags behind dust density; 
+%  the streaming instability neccessarily involves a phase lag between
+%  gas pressure and dust densities; 
 \item dust-loading introduce buoyancy forces that generally 
   stabilizes the vertical shear 
-  instability that arises from a radial temperature gradient. 
+  instability associated with global radial temperature gradients. 
 \end{inparaenum}  
@@
 %dust settling does not lead to axisymmetric instability. 
-We also discuss future applications of the dusty/adiabatic gas
-correspondence in finding dusty analogs of other hydrodynamic
+We also discuss future applications of this thermodynamic framework in
+finding dusty analogs of other hydrodynamic 
 phenomena (e.g. Rossby wave instability, convective overstability, and 
-zombie vortices), and to simulate dusty protoplanetary disks
-using purely hydrodynamic codes. %We briefly describe extending this
+zombie vortices), in simulating dusty protoplanetary disks
+using purely hydrodynamic codes, as well as potential generalizations
+of the dusty/adiabatic gas equivalence principle.
+
+%We briefly describe extending this
 %equivalence principle to polytropic and adiabatic gas.
@@
 the momentum/energy equations to model dust-gas drag. This approach
-can take advantage of, and adapt, already-developed numerical methods/codes for
+can take advantage of existing numerical methods/codes for
 simulating pure gas dynamics.
@@
 This can be numerically challenging for two-fluid models and 
-following Lagragian particles; but is straight forward to simulate in the 
-one-fluid formulation \citep{price15}.  The single fluid formalism
+following Lagragian particles; but is straight forward with the 
+one-fluid formulation \citep{price15}.  The single fluid framework 
 also has the important advantage that it is 
@@
 1$, where $\Omega$ is the disk orbital frequency. We may then use the
-`terminal velocity approximation' to set $\Delta \bm{v} = \tstop\nabla
+`terminal velocity approximation' to set the dust-gas relative drift
+to $\Delta \bm{v} = \tstop\nabla
 P/\rhog$, where $P$ is the gas pressure \citep{youdin05a, 
-  jacquet11,laibe14}. %This eliminates the evolutionary equation for
-%$\Delta \bm{v}$. 
+  jacquet11,laibe14}. This approximation reflects the physical result
+that dust particles tend to drift towards pressure maximum.
+
  We also consider irradiated protoplanetary disks 
@@
 
-Under these approximations, we show that the equations of 
+Under these two approximations, we show that the equations of 
 isothermal dusty gas dynamics are equivalent to that of adiabatic, pure gas
@@
 entrained in the gas, which is analogous to the advection
-of entropy following an adiabatic fluid. We will find that the   
+of entropy following an adiabatic fluid. We show that the   
 effective entropy of isothermal dusty gas scales with its gas
@@
 
-By re-casting the dusty problem into identical form as standard  
-hydrodynamics, we can apply well-known methods to generalize previous stability 
-analyses for pure gas disks to the dusty case. %We explicitly demonstrate this for
-%perfectly coupled dust in strictly and locally isothermal disks. 
+By re-casting the dusty problem into standard  
+hydrodynamics, we can apply established methods to generalize previous stability 
+results for pure gas dynamics to dusty disks.  
 We show that strictly isothermal protoplanetary disks with
 perfectly-coupled dust are generally 
-stable to axisymmetric perturbations, regardless of how  thin the dust
-layer is. We put forward a thermodynamical interpretation of
+stable to axisymmetric perturbations.%, regardless of how  thin the dust
+%layer is. 
+{\bf should we mention vertically well mixed case here?} 
+We put forward a thermodynamical interpretation of 
 dust-gas drag instabilities, and apply it to the streaming instability.   
 We also study the effect of dust-loading on the vertical shear
-instability \citep[VSI,][]{nelson13,lin15,barker15}.
+instability previously considered in pure gas disks
+\citep[VSI,][]{nelson13,lin15,barker15}.
 
-This paper is organized as follows. In \S\ref{key_idea} we first
-describe, without derivation, the analogy between dusty gas and    
-ideal hydrodynamics. We then discuss in 
-\S\ref{grow_osc} a general result of fluid dynamics that
-underpins the bulk of our study: intabilities due to 
-pressure-density phase lags. This means that dusty-gas phenomena, such
-as SI, have thermodynamical interpretations.
+This paper is organized as follows. We first present, without
+derivation, the correspondence between dusty gas and     
+ideal hydrodynamics in \S\ref{key_idea}. In \S\ref{grow_osc} 
+we discuss a key idea in fluid dynamics, relevant to our 
+discussion on  SI and VSI: instabilities associated with
+pressure-density phase lags.   
+%This means that dusty-gas phenomena, such 
+%as SI and VSI, have thermodynamical interpretations.
 
@@
 The key concept we introduce in this work is the description
-of dusty protoplanetary disks \emph{without} having dust as an
-explicit variable. This is best illustrated by writing down our model  
-equations that describe a mixture of isothermal gas and small
+of dusty protoplanetary disks \emph{without} explicit dust
+equations. This is best illustrated by writing down our model   
+equations for a mixture of isothermal gas and small
 dust particles: 
@@
 and $\mathcal{C}$ represent effects of dust-gas drag, which
-is paremterized by the stopping time $\tstop$.
+is paremterized by the stopping time $\tstop$ ($\mathcal{C}=0$ for
+$\tstop=0$). Similar equations for dusty-gas are given by
+\cite{price15}.
 
-Eq. \ref{mass0}---\ref{en0} appear identical to adiabatic gas dynamics, even  
-though they physically represents a two-phase, dusty-gas. The main
-difference is the term $\mathcal{C}$. Here it does not represent
-physical cooling (as the system is isothermal). Instead, dust-gas drag
-causes particles to drift, hence by momentum conservation a gas drift
-in the opposite direction, which changes the pressure distribution.
+Eq. \ref{mass0}---\ref{en0} appear identical to adiabatic gas
+dynamics, even though they physically represents a two-phase,
+dusty-gas. The main difference is the term $\mathcal{C}$. Here it does
+not represent physical cooling, as the system is isothermal by
+assumption. Instead, dust-gas drag causes particles to drift, hence by
+momentum conservation a gas drift in the opposite direction, which
+changes the pressure distribution.
 
@@
 generalized to dusty disks without additional effort. This includes 
-the important effect of phase lags between density and pressure in
-fluids, discussed next.
+the important consequence of phase lags between density and pressure,
+discussed next.
 
 \section{Growing oscillations by doing work}\label{grow_osc}
-A general result in fluid dynamics is that work is done whenever the
-pressure and density evolution of the fluid are not in phase. If 
-the average work done is positive, then oscillation amplitudes 
-grow: the work done allows the system to `overshoot' beyond
+A general result in fluid dynamics is that work is done whenever
+oscillations in the pressure and density of the fluid are not in
+phase. If  the average work done is positive, then oscillation
+amplitudes grow: the work done allows the system to `overshoot' beyond 
 the amplitude of preceeding oscillations.
@@
     pressure to lag behind density, which results in a clockwise path in
-    the P-V plane, implying positive work done by the fluid and
-    hence would lead to growing oscillation amplitudes.  
+    the P-V plane, implying positive work done by the fluid, which 
+     would increase oscillation amplitudes.  
     \label{pdv_cartoon}
@@
 $r$ from the star \citep{chiang97}. Since $P\propto \rhog T$ for an
-ideal gas, the pressure of a gas parcel varies with  
-position as well. There is no reason to expect $P$ and $\rhog$ to
+ideal gas, %the pressure of a gas parcel varies with  
+%position as well. 
+there is no reason to expect $P$ and $\rhog$ to 
 be in phase as a gas parcel oscillates between different radii. In
@@
 In order to apply this thermodynamic interpretation of dust-gas drag
-instabilities, we need to develop an analogy between dusty-gas
+instabilities, we need to develop a formal analogy between dusty-gas
 and pure gas dynamics. We show this is possible in the limit of

### 2017-05-09 result.tex
extra intro section
misc edits to the main text overall

@@
 We now present numerical solutions for vertically stratified,
-non-self-gravitating disks assuming perfectly coupled dust
-($\tstop=0$) so the equilibria
-defined in \S\ref{eqm} are exact steady states.
+non-self-gravitating disks, assuming perfectly coupled dust. 
+We formally take $\tstop=0$ so the equilibria defined in \S\ref{eqm} 
+are exact steady states. In reality,   
+%\subsection{Validity}
+%{\bf maybe incorporate into intro to this section?}
+%We considered $\tstop=0$ in order to perform stability analysis on an
+%exact steady state. 
+%When $\tstop\neq0$, a stratified dusty 
+%disk initialized as in \S\ref{eqm} would evolve 
+%as 
+dust settles to the midplane on a timescale $t_\mathrm{settle}\sim 
+1/\tstop\OmK^2$ \citep{takeuchi02}. However, we expect the 
+perfectly-coupled limit to be valid  
+provided timescales of interest $t_\mathrm{grow}\ll
+t_\mathrm{settle}$. For the VSI this translates to 
+%Since VSI growth rates are $O(h_\mathrm{g}\OmK)$, 
+%we require  
+%\begin{align}
+ $ \tstop\OmK \ll h_\mathrm{g}. $
+%\end{align} 
+For thin PPDs, this is satisfied for $\tstop\OmK \ll O(10^{-2})$.
 
-Apart from \S\ref{varHd}---\ref{vert_mixed}, we consider constant
+We first consider constant
 midplane dust-to-gas ratios $\epsilon_0$ and characteristic thickness
 $H_\epsilon$. Then the dust-to-gas ratio $\epsilon=\epsilon(z)$. In 
-this limit any growing modes must be associated with the imposed temperature gradient
-(\S\ref{limits}). We are thus examining the effect of dust-loading on
-the VSI previously studied in pure gas disks 
-\citep[][\citetalias{lin15} in this section]{lin15}.
+this limit any growing modes must be associated with the imposed
+temperature gradient (see \S\ref{limits}). We are then studying the
+effect of dust-loading on the VSI previously studied in pure gas
+disks \citep[][\citetalias{lin15} in this section]{lin15}. In 
+\S\ref{varHd} we allow $\p_r\epsilon\neq 0$, and in 
+\S\ref{vert_mixed} we consider VSI driven entirely by radial
+gradients in the dust-to-gas ratio (as discussed in
+\S\ref{vert_mixed_theory}).
 
-The parameters for this linear problem includes $\epsilon_0$, the dust
-layer thickness $\Hd(r_0)$, and the perturbation radial wavenumber
+The parameters for the linear problem includes $\epsilon_0$, the dust
+layer thickness $\Hd$, and the perturbation radial wavenumber
 $k_x$. We choose a midplane gas density profile 
@@
 from \citetalias{lin15}. Amplitudes are expanded in Chebyshev
-polynomials up to order $N_z=512$. %We check results using
+polynomials up to order $512$. %We check results using
 %Eq. \ref{vsi_check}.
@@
 but their oscillation frequency increases with
-$\epsilon_0$. i.e. it increases with the vertical buoyancy. 
+dust-loading, i.e. it increases with the vertical buoyancy. 
 The total number of modes do not change 
 significantly. This is in contrast with the effect of increasing
-cooling times, for which \citetalias{lin15} find fewer unstable modes.
+cooling times in an adiabatic gas disk, for which \citetalias{lin15}
+find fewer unstable modes. 
 %until the system is eventually completely stable.
@@
 This is consistent with Fig. \ref{compare_vshear_fixHd} comparing the basic state vertical
-shear and buoyancy. Notice also maxium perturbtions to the dust-to-gas ratio,
-$\delta\epsilon$, shifts from disk boundaries 
-to the midplane.
+shear and buoyancy. % Notice also maxium perturbtions to the dust-to-gas ratio,
+% $\delta\epsilon$, shifts from disk boundaries 
+% to the midplane.
 
@@
 shear is dominated by that due to the dust layer (see
-\S\ref{vertshear}). However, we see that growth rates $s\propto
-|q|$, i.e .growth rates vanish in the strictly isothermal limit.  
+\S\ref{vertshear}). However, we see that $s\propto
+|q|$, i.e growth rates vanish in the strictly isothermal limit.  
 In particular, this holds for $\rich < 0.1$, the critical value 
@@
 contribution to vertical shear: as $q\to0$,  $\p_z\Omega$ becomes
-entirely due to dust-settling and there is no instability ($s\to
+entirely due to $\p_z\epsilon$ and there is no instability ($s\to
 0$). This result is independent of $\rich$, so the Richardson number
@@
 same disk as above but with $q=0$ and choose $H_\mathrm{d}$ such
-that $H_\epsilon=10^3\Hg$. Then vertical shear $\p_z\Omega$ is then
+that $H_\epsilon=10^3\Hg$. Then vertical shear $\p_z\Omega$ is 
 attributed to the radially-varying dust-to-gas ratio (see
@@
 \end{figure}
-
-\subsection{Validity}
-{\bf maybe incorporate into intro to this section?}
-We considered $\tstop=0$ in order to perform stability analysis on an
-exact steady state. When $\tstop\neq0$, a stratified dusty 
-disk initialized as in \S\ref{eqm} would evolve 
-%due to dust-gas relative drift. In particular, in the
-%absence of turbulent/diffusive processes, dust particles 
-as dust settles to the
-midplane on a timescale $t_\mathrm{settle}\sim 1/\tstop\OmK^2$
-\citep{takeuchi02}. However, we expect the above results to remain
-valid provided the VSI growth
-timescale $t_\mathrm{grow}\ll t_\mathrm{settle}$. Since 
-VSI growth rates are $O(h_\mathrm{g}\OmK)$, we require 
-\begin{align}
-  \tstop\OmK \ll h_\mathrm{g}. 
-\end{align} 
-For thin PPDs with $h_\mathrm{g}\sim 0.05$, this is satisfied 
-for $\tstop\OmK \ll O(10^{-2})$.

### 2017-05-09 summary.tex
misc edits to the main text overall

@@
 structure, induced by dust, may modify the vertical structure of 3D
-waves and unstable modes. We will examine this problem in a following 
-study.
+waves and unstable modes.
 
@@
 sustain a turbulent state where shock heating due to gravitational 
-instabilities are balanced by thermal losses
-\citep{gammie01}. Now, dust-gas drift behaves like thermal diffusion
-or cooling (Eq. \ref{eff_energy}), similar to how radiative cooling 
-appears in models of gravito-turbulent disks. Thus it may be
-conceivable that a massive dusty disk admit `dusty
-gravito-turbulence'.
+instabilities are balanced by radiative cooling 
+\citep{gammie01}. Since dust-gas drift appears as a diffusion or 
+cooling in our framework (Eq. \ref{eff_energy}), it may be conceivable
+to have `dusty gravito-turbulence'. As self-gravity increases the
+local density, the associated pressure maxima attract dust-particles,
+but the back-reaction onto the gas may try to flatten the pressure
+bump \citep{taki16}.
 
@@
 \end{align}
-Thus we expect that RWI may also be triggered by extrema 
+Thus RWI may also be triggered by extrema 
 in the dust-to-gas ratio, e.g. narrow dust rings/gaps. This may lead 
@@
 energy source/sink to the mixture (\S\ref{energy_analogy}). It would be
-interesting to explore whether or not dust-gas drag can provide the 
-required `cooling' to enable a `dusty convective 
+interesting to explore whether or not dust-gas drag can play the role
+of cooling to enable a `dusty convective 
 overstability'.
@@
 \citep{marcus15,umurhan16d}. It operates by having finite-amplitude perturbations
-exciting `critical layers' which roll 
+exciting `critical layers',  where the intrinsic wave frequency of the
+perturbation matches the local vertical buoyancy frequency 
+\citep{marcus13}. These critical layers roll 
 up into vortices, exciting further critical layers. The process
-repeats itself and the result is an array of vortices. These critical 
-layers are locations where the intrinsic wave frequency of a 
-disturbance matches the local vertical buoyancy frequency 
-\citep{marcus13}.
+repeats itself and the result is an array of vortices.
 
@@
 $\lesssim 1$AU \citep{lesur16, malygin17}. However, this consideration
-assumes a dust-free disk.
+assumes a dust-free disk as far as the dynamics is concerned.
 
-On the other hand, we have show that dust-loading an isothermal disk
-also induces an effective buoyancy. % This effect is maximized when dust is
-% perfectly coupled to the gas
-. It is thus natural to ask whether or not
+On the other hand, we have show that dust-loading induces an effective
+buoyancy even in isothermal disks. It is thus natural to ask whether or not
 the required adiabatic conditions for ZVI can be realized through
 dust-loading, and thus produce `dusty ZVI'. This may allow the zombie
-vortices to exist at $\gtrsim 1$AU in PPDs.
+vortices to develop $\gtrsim 1$AU in PPDs.
 
@@
 \textsc{Pluto}\footnote{\url{http://plutocode.ph.unito.it/}}      
-hydrodynamics code into a dusty gas dynamics code appropriate for
+hydrodynamics code \citep{mignone07} into a dusty gas dynamics code appropriate for
 simulating protoplanetary disks coupled to small dust. We will apply 
@@
 
-We obtain an analog of the Solberg-Hoiland criteria for
-axisymmetric stability for strictly isothermal, perfectly-coupled dusty gas.  
+We obtain the equivalent Solberg-Hoiland criteria for the 
+axisymmetric stability of strictly isothermal, perfectly-coupled dusty gas.  
 Applying this to typical protoplanetary disks, we find that 
@@
 explicitly show in SI the evolution of gas
-pressure lags behind dust density; and is in fact a general property of 
-overstabilities driven by strong dust-gas drag. 
+pressure lags behind dust density. This is in fact a general property of 
+instabilities driven by dust-gas drag. 
 It takes a finite time for the gas to respond to
@@
 as the Rossby Wave Instability \citep{li00} or the `Zombie Vortex
-Instability' \citep{marcus15} could have dusty counter-parts.   
+Instability' \citep{marcus15}, where entropy plays a role, could have dusty counter-parts.   
 Furthermore, hydrodynamic instabilities driven by thermal cooling, such as VSI in

### 2017-05-11 paper.tex
add flow chart in intro to summarize centreal idea

@@
 \usepackage{bm}
+\usepackage{tikz}
+\usetikzlibrary{shapes.geometric, arrows}
+
 %\usepackage{hyperref}
@@
 
+\begin{figure}
+\tikzstyle{process} = [rectangle, minimum width=2cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
+  \tikzstyle{line} = [draw, -triangle 45,thick]  
+  \begin{center}
+    \begin{tikzpicture}[node distance=5cm]
+      \node[process](ellip){isothermal gas + dust};
+      \node[process,right of=ellip, node distance = 6cm](gi){adiabatic gas};
+      \path [line](ellip) -- node [yshift=.8em] {\textcolor{blue}{equivalent}} (gi);
+      \path [line](gi) -- node [yshift=-.8em] {for small particles} (ellip);
+    \end{tikzpicture}
+ \end{center}
+\caption{The central concept of this work. A mixture of small particles
+  and isothermal gas behave similarly to an adiabatic, pure gas. This
+  permits one to use the equations of ideal hydrodynamics to study
+  dust-gas dynamics. }
+\end{figure}
+
 \section{Dust-free modeling of dusty protoplanetary disks}\label{key_idea}

### 2017-05-11 paper.tex
add flow chart in intro to summarize centreal idea

--- 2017-05-09 paper.tex
@@
 isothermal dusty gas dynamics are equivalent to that of adiabatic, pure gas
-dynamics. This is because tightly-coupled dust particles are almost
+dynamics (Fig. \ref{concept_chart}). This is because tightly-coupled dust particles are almost
 entrained in the gas, which is analogous to the advection
@@
 \S\ref{discussion} before summarizing in \S\ref{summary}.
+
+\begin{figure}
+\tikzstyle{process} = [rectangle, minimum width=2cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
+  \tikzstyle{line} = [draw, -triangle 45,thick]  
+  \begin{center}
+    \begin{tikzpicture}[node distance=5cm]
+      \node[process](ellip){isothermal gas+dust+drag};
+      \node[process,right of=ellip, node distance = 5.5cm](gi){pure 
+        gas+cooling};
+      \path [line](ellip) -- node [yshift=.8em]
+      {\textcolor{blue}{equivalent} for} (gi);
+      \path [line](gi) -- node [yshift=-.8em] {small particles} (ellip);
+    \end{tikzpicture}
+ \end{center}
+\caption{The central concept of this work. A mixture of small particles
+  coupled to isothermal gas behave similarly to an adiabatic, pure gas
+  subject to cooling.  This
+  permits one to use the equations for standard hydrodynamics to study
+  dust-gas dynamics. \label{concept_chart}}
+\end{figure}
 

### 2017-05-12 paper.tex
wording changes in section 2

@@
  & \frac{\p P}{\p t} + \bm{v}\cdot\nabla P 
-  = - P \nabla\cdot\bm{v}
-                + \mathcal{C}(\rho, P; \tstop)\label{en0}, 
+  = - P \nabla\cdot\bm{v} - \nabla\cdot\bm{\mathcal{F}}
+%                + \mathcal{C}(\rho, P; \tstop)
+\label{en0}, 
 \end{align}
@@
 $\Phi_*$ is the stellar gravitational potential,   
-and $\mathcal{C}$ represent effects of dust-gas drag, which
-is paremterized by the stopping time $\tstop$ ($\mathcal{C}=0$ for
-$\tstop=0$). Similar equations for dusty-gas are given by
+and $\bm{\mathcal{F}}$ model effects of dust-gas drag, explained below. 
+%$\mathcal{C}$ represent effects of dust-gas drag, which
+%is paremterized by the stopping time $\tstop$ ($\mathcal{C}=0$ for
+%$\tstop=0$). 
+Similar equations for dusty-gas are given by
 \cite{price15}.
 
-Eq. \ref{mass0}---\ref{en0} appear identical to adiabatic gas
-dynamics, even though they physically represents a two-phase,
-dusty-gas. The main difference is the term $\mathcal{C}$. Here it does
-not represent physical cooling, as the system is isothermal by
-assumption. Instead, dust-gas drag causes particles to drift, hence by
-momentum conservation a gas drift in the opposite direction, which
-changes the pressure distribution.
+Eq. \ref{mass0}---\ref{en0} appear identical to pure gas
+dynamics, but they physically represents a two-phase, 
+dusty gas. The mixture obeys mass and momentum conservation as usual,
+but also an energy-like equation, even though the system is
+isothermal by assumption. The term $\bm{\mathcal{F}}\propto -\tstop
+\nabla P$ is not a true heat flux. Instead, it reflects 
+particle flux towards pressure maximum due to dust-gas drag 
+\citep{weidenschilling77}. The back-reaction onto the gas then changes
+the pressure distribution.
+
+%weidenschilling7
 

### 2017-05-12 paper.tex
tiny wording changes to sec 2

@@
 $\Phi_*$ is the stellar gravitational potential,   
-and $\bm{\mathcal{F}}$ model effects of dust-gas drag, explained below. 
+and $\bm{\mathcal{F}}=~\bm{\mathcal{F}}(\rho, P; \tstop)$ model effects of dust-gas drag, explained below. 
 %$\mathcal{C}$ represent effects of dust-gas drag, which
@@
 but also an energy-like equation, even though the system is
-isothermal by assumption. The term $\bm{\mathcal{F}}\propto -\tstop
+isothermal by assumption. The term $-\bm{\mathcal{F}}\propto \tstop
 \nabla P$ is not a true heat flux. Instead, it reflects 

### 2017-05-15 paper.tex
updated SI table, now do full comparison

@@
 \usetikzlibrary{shapes.geometric, arrows}
+\usepackage{multirow}
 
@@
 %dust settling does not lead to axisymmetric instability. 
-We also discuss future applications of this thermodynamic framework in
-finding dusty analogs of other hydrodynamic 
-phenomena (e.g. Rossby wave instability, convective overstability, and 
+We also discuss %future applications of this thermodynamic framework in
+%finding 
+potential dusty analogs of other hydrodynamic 
+processes (e.g. Rossby wave instability, convective overstability, and 
 zombie vortices), in simulating dusty protoplanetary disks
-using purely hydrodynamic codes, as well as potential generalizations
-of the dusty/adiabatic gas equivalence principle.
+using purely hydrodynamic codes, as well as possible generalizations
+of the dusty/adiabatic gas analogy.
 

### 2017-05-17 linear.tex
updated SI table, now do full comparison
improved table 1 format
added bonus figure of comparing SI in two-fluid and one fluid

@@
 
+\begin{figure}
+\includegraphics[width=linewidth]{figures/si_2f_1f_compare}
+\caption{{\bf in case the paper isn't long enough} Comparison of the linear
+streaming instability obtained from a full two-fluid analysis and the current 
+one-fluid framework simplified by the terminal velocity approximation. The vertical
+wavenumber is fixed and growth rates are maximized over $K_x$. 
+\label{si_compare}}
+\end{figure}
 %"generalized stokes number" would include other problem parameters,
@@
 
-\begin{deluxetable*}{llrrrrrr}
+% \begin{deluxetable*}{llrrrrrr}
+%   \tablecolumns{8}
+%   \tablecaption{Selected eigenfrequencies for the streaming
+%     instability. \label{si_compare}
+%   }
+%   \tablehead{
+%     \colhead{Mode} & 
+%     \colhead{$\tau_\mathrm{s}\OmK$} &
+%     \colhead{$\epsilon$} &
+%      \colhead{$K_{x,z}$} &
+%       \colhead{$\sigma/\OmK$ (two-fluid)} &
+%     \colhead{$\sigma/\OmK$ (one-fluid)} &
+%     \colhead{$\mathcal{W}$ (arbitrary units)} &
+%       \colhead{$\Delta P$ lag}
+%   }
+% \startdata
+%  linA, \cite{youdin07b} &  $0.1$       & 3.0 & 30    & $0.3480 +
+%  0.4190\ii$ & $0.3640 + 0.4249\ii$ & $0.90$ & $30\degr$\\
+
+% linB, \cite{youdin07b} & $0.1$        &  0.2 & 6 & $-0.4999 +
+% 0.0155\ii$&   $-0.4981 + 0.0054\ii$  & $1.54$ & $1.2\degr$ \\
+
+% linC,  \cite{bai10b}  & $10^{-2}$   &  2.0 & 1500&   $0.1049 +
+% 0.5981\ii$   &  $0.1338 + 0.6650\ii$  & $0.15$& $11\degr$ \\
+
+% linD, \cite{bai10b} &  $10^{-3}$   &  2.0 & 2000 & $0.3225 + 
+% 0.3154\ii$& $0.3219 + 0.3154\ii$ &  $1.28$ & $22\degr$ 
+% \enddata
+% \end{deluxetable*}
+
+\begin{deluxetable*}{lrrrrrr}
   \tablecolumns{8}
-  \tablecaption{Selected eigenfrequencies for the streaming
+  \tablecaption{Selected modes of the streaming
     instability. \label{si_compare}
   }
-  \tablehead{
-    \colhead{Mode} & 
-    \colhead{$\tau_\mathrm{s}\OmK$} &
-    \colhead{$\epsilon$} &
-     \colhead{$K_{x,z}$} &
-      \colhead{$\sigma/\OmK$ (two-fluid)} &
-    \colhead{$\sigma/\OmK$ (one-fluid)} &
-    \colhead{$\mathcal{W}$ (arbitrary units)} &
-      \colhead{$\Delta P$ lag}
-  }
+%  \tablehead{
+%    \colhead{Mode} & 
+%    \colhead{$\tau_\mathrm{s}\OmK$} &
+%    \colhead{$\epsilon$} &
+%     \colhead{$K_{x,z}$} &
+%    \colhead{Complex frequency} &
+%    \colhead{$\sigma/\OmK$} &
+%    \colhead{Work done} &
+%    \colhead{$\mathcal{W}/\left|\Delta P\Delta\rho^*/\rho\right|$} &
+%    \colhead{Pressure-density lag} &
+%    \colhead{$\varphi$} 
+%  }
 \startdata
- linA, \cite{youdin07b} &  $0.1$       & 3.0 & 30    & $0.3480 +
- 0.4190\ii$ & $0.3640 + 0.4249\ii$ & $0.90$ & $30\degr$\\
-
-linB, \cite{youdin07b} & $0.1$        &  0.2 & 6 & $-0.4999 +
-0.0155\ii$&   $-0.4981 + 0.0054\ii$  & $1.54$ & $1.2\degr$ \\
-
-linC,  \cite{bai10b}  & $10^{-2}$   &  2.0 & 1500&   $0.1049 +
-0.5981\ii$   &  $0.1338 + 0.6650\ii$  & $0.15$& $11\degr$ \\
-
-linD, \cite{bai10b} &  $10^{-3}$   &  2.0 & 2000 & $0.3225 + 
-0.3154\ii$& $0.3219 + 0.3154\ii$ &  $1.28$ & $22\degr$ 
+\hline\hline
+\multicolumn{1}{c}{Mode ($\tau_\mathrm{s}\OmK$, $K_{x,z}, \rhog/\rhod$)} &
+\multicolumn{2}{c}{Complex frequency, $\sigma/\OmK$} &
+\multicolumn{2}{c}{Work done, $\mathcal{W}/\left|\Delta P\Delta\rho^*/\rho\right|$} &
+\multicolumn{2}{c}{Pressure-density lag, $\varphi$} \\
+\hline
+\multicolumn{1}{c}{} &
+ \multicolumn{1}{c}{two-fluid} &
+ \multicolumn{1}{c}{one-fluid} &
+ \multicolumn{1}{c}{two-fluid} &
+ \multicolumn{1}{r}{one-fluid} &
+ \multicolumn{1}{c}{two-fluid} &
+ \multicolumn{1}{r}{one-fluid} \\
+ \hline
+% \multicolumn{6}{c}{\textcolor{blue}{linA\tablenotemark{$\dagger$}: $\tau_\mathrm{s}\OmK=0.1$, $K_{x,z}=30$, $\rhog/\rhod = 3$}}\\
+% \hline
+linA\tablenotemark{$\dagger$} ($0.1, 30, 3$) &
+$0.3480 + 0.4190\ii$ & $0.3640 + 0.4249\ii$ & $0.078$ & $0.090$ & $27\degr$ & $30\degr$  \\
+% \hline
+% \multicolumn{6}{c}{\textcolor{blue}{linB\tablenotemark{$\dagger$}: $\tau_\mathrm{s}\OmK=0.1$, $K_{x,z}=6$, $\rhog/\rhod = 0.2$}} \\
+% \hline
+linB\tablenotemark{$\dagger$} ($0.1, 6,0.2$) &
+$-0.4999 + 0.0155\ii$&   $-0.4981 + 0.0054\ii$ & 0.025   & $0.0054$ & $5.8\degr$  &  $1.2\degr$  \\
+% \hline
+% \multicolumn{6}{c}{\textcolor{blue}{linC\tablenotemark{$\ddagger$}: $\tau_\mathrm{s}\OmK=10^{-2}$, $K_{x,z}=1500$, $\rhog/\rhod = 2$}} \\
+% \hline
+linC\tablenotemark{$\ddagger$} ($10^{-2}, 1500, 2$) &
+$0.1049 + 0.5981\ii$   &  $0.1338 + 0.6650\ii$ & 0.0076 & $0.013$ & $8.3\degr$ &$11\degr$  \\
+% \hline
+% \multicolumn{6}{c}{\textcolor{blue}{linD\tablenotemark{$\ddagger$}: $\tau_\mathrm{s}\OmK=10^{-3}$, $K_{x,z}=2000$, $\rhog/\rhod = 2$}} \\
+% \hline
+linD\tablenotemark{$\ddagger$} ($10^{-3}, 2000, 2$) &
+$0.3225 +  0.3154\ii$& $0.3219 + 0.3154\ii$ &  $0.061$ & $0.061$  & $22\degr$ & $22\degr$  
 \enddata
+\tablenotetext{$\dagger$}{\cite{youdin07b}}
+\tablenotetext{$\ddagger$}{\cite{bai10b}}
 \end{deluxetable*}

### 2017-05-22 appendix.tex
edits on the plane
fixed tiny typo in appendix
new section 7 on secular GI. rest of paper might need adjusting. mostly doing non-SG, but one sec and one appendix has SG
misc edits from sect 4 to sect 8.3

@@
 and $\Gamma$ is a constant. The effective energy equation is then
-Eq. \ref{poly_energy}. We also include the self-gravitational
-potential $\psi$ of the gas
-plus dust mixture, which satisfies 
-%\begin{align}
-$  \nabla^2 \psi  = 4 \pi G \rho. $
-%\end{align} 
+Eq. \ref{poly_energy}. % We also include the self-gravitational
+% potential $\psi$ of the gas
+% plus dust mixture, which satisfies 
+% %\begin{align}
+% $  \nabla^2 \psi  = 4 \pi G \rho. $
+% %\end{align} 
 The linearized equations for axisymmetric
@@
    \ii\sigma\frac{\delta P}{P} &= \Gamma \nabla\cdot\dd\bm{v} +
-  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
+  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{K} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
   -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
@@
 Appendix \ref{lin_press}. These equations do not assume the
-radially-local approximation used in the main text. From the
+radially-local approximation used in numerical computations. From the
 linearized meridional momentum equations, we find
@@
 -\frac{\dd \mathcal{C}}{\rho}  = \tstop c_s^2 \left[\ii k_x F_r\left( 1-2\tepsilon\right)
-  W + \tepsilon |\bm{k}|^2 Q\right]. 
+  W + \tepsilon |\bm{k}|^2 Q\right]. \label{lin_drag_si}
 \end{align}
@@
   \frac{P}{\rho}\sigma + k_x\tstop c_s^2 F_r\left(1 -
-  2\tepsilon\right)\right]W, \\
+  2\tepsilon\right)\right]W, \label{si_QW1}\\
    \sigma^2 \left( \kappa^2 - \sigma^2 - \ii k_x F_r\right)W &=
@@
 
+\subsection{Approximate solutions in the dust-rich limit}\label{si_dust_rich}
+Here we seek analytic solutions for the streaming instability by 
+examining limiting cases and with further approximations. We shall fix
+$k_z$ and $\fdust>1/2$ (the dust-rich case), and maximize growth rates
+over $k_x$. We consider small stopping times, $\tau_\mathrm{s}\to0$,
+and assume that the optimum $k_x\to\infty$ but that
+$k_x\tau_\mathrm{s}$ remain bounded.
+
+We begin with the incompressible dispersion
+relation, Eq. \ref{streaming_incompressible}. We assume a Keplerian
+disk and neglect the quartic  
+term because it is higher order in $\tau_\mathrm{s}$ (see also
+\S\ref{spurious_epi})  In dimensionless form: 
+\begin{align}
+  \nu^3 + \st\left[\ii\tepsilon + 2 K_x \fg \left(1-\fdust\right)\right]\nu^2
+  - \left(\frac{K_z}{K_x}\right)^2\nu - 2\st K_x \fg
+  \left(1-2\fdust\right)\left(\frac{K_z}{K_x}\right)^2 = 0,\label{dimless_cubic}
+\end{align}
+where $\nu = \sigma/\Omega$, $\st = \taus\Omega$, $\fg = 1-\fdust$ (the gas
+fraction), and recall $K_{x,z}=k_{x,z}\eta r$. We have used $F_r =
+2\eta \fg r \Omega^2$ and considered large $K_x$.
+
+Full analytic expressions for cubic roots are unwieldy. To proceed
+in an a tractable manner, let us \emph{assume} at this stage that the
+quadratic term can be neglected compare to the last term. This amounts
+to assuming a  sufficiently \emph{large} $K_z$ such that :
+\begin{align}
+\left|\frac{\ii\fdust}{2K_x\fg} + \left(1 -
+      \fdust\right)\right| \left|K_x\nu\right|^2 \ll
+  K_z^2\left(2\fdust - 1\right) \label{min_kz}
+\end{align}
+This will be checked later. (For fixed $\fdust$ and $K_x\to\infty$, the
+imaginary term can be neglected.)
+
+Eq. \ref{dimless_cubic} now becomes the depressed cubic 
+\begin{align}
+  \nu^3 - \calP \nu = \calQ, \label{depressed_cubic}
+\end{align}
+where $\calP$ and $\calQ$ can be read off Eq. \ref{dimless_cubic}. This can be
+solved with Vieta's substitution 
+\begin{align*}
+\nu = \mu + \frac{\calP}{3\mu},
+\end{align*}
+then Eq. \ref{depressed_cubic} becomes a quadratic for $\mu^3$,
+\begin{align}
+\mu^6 - \calQ\mu^3 + \frac{\calP^3}{27} = 0. 
+\end{align}
+The solution is
+\begin{align}
+  \mu^3 &= \frac{\calQ}{2}\left(1\pm \sqrt{1 -
+      \frac{4\calP^3}{27\calQ^2}}\right)\notag\\
+  & = \st \fg (1- 2\fdust)\frac{K_z^2}{K_x}\left\{1 \pm \sqrt{1 -
+      \frac{K_z^2}{27K_x^2\left[\st K_x \fg
+          \left(1-2\fdust\right)\right]^2}}\right\}.  
+\end{align}
+At this point we assume $K_z$ is sufficiently \emph{small} such that
+the term $\propto K_z^2$ inside the square root may be neglected. That
+is, 
+\begin{align}
+  K_z^2 \ll 27 \left(K_x^2\st\right)^2 \left[\fg
+          \left(1-2\fdust\right)\right]^2. \label{max_kz}
+%27 K_x^2 \left[\st K_x \fg
+%          \left(1-2\fdust\right)\right]^2.
+\end{align}
+This will be checked later. Remembering that we are considering the
+dust-rich case so that $1-2\fdust<0$, the solution for $\mu$ is   
+\begin{align}
+\mu = \left[2\st\fg \left(2\fdust
+    -1\right)\frac{K_z^2}{K_x}\right]^{1/3}e^{\ii\pi/3}. \label{mu_explicit} 
+\end{align}
+We have chosen the root for instability. We can now write down the
+solution for the eigenfrequency $\nu$:
+\begin{align}
+\real{(\nu)} &= \frac{1}{2}\left(|\mu| + \frac{K_z^2}{3
+    K_x^2|\mu|}\right)\notag\\
+&= \frac{1}{2}\left(\calA \st^{1/3}K_x^{-1/3} + \calB
+    \st^{-1/3}K_x^{-5/3}\right),\\
+\imag{(\nu)} &= \frac{\sqrt{3}}{2}\left(|\mu| - \frac{K_z^2}{3
+    K_x^2|\mu|}\right),\notag\\
+&= \frac{\sqrt{3}}{2}\left(\calA \st^{1/3}K_x^{-1/3} - \calB
+    \st^{-1/3}K_x^{-5/3}\right),
+\end{align}
+%Inserting the expression for $|\mu|$ into $\imag{\nu}$, the
+%explicit expression for the growth rate may be written as 
+%\begin{align}
+%  s = \frac{\sqrt{3}}{2}\left(\calA \st^{1/3}K_x^{-1/3} - \calB
+%    \st^{-1/3}K_x^{-5/3}\right)\OmK, 
+%\end{align}
+where
+\begin{align}
+  &\mathcal{A} = \left[2\fg \left(2\fdust
+      -1\right)K_z^2\right]^{1/3} ,\\
+  &\mathcal{B} = \frac{K_z^2}{3\calA}. 
+\end{align}
+
+Maximizing growth rates over $K_x$ by setting $\p \imag{(\nu)}/\p K_x
+= 0$, we find the optium wavenumber is given by
+\begin{align}
+  K_x = \left(\frac{5\calB}{\calA}\right)^{3/4}\st^{-1/2} =
+  \left(\frac{5}{3}\right)^{3/4}\sqrt{\frac{K_z}{2\fg\left(2\fdust-1\right)\st}}. 
+\end{align}
+The maximum growth rate is then 
+\begin{align}
+  \mathrm{max}\left(s\right) =
+  \frac{2\sqrt{3}\calA^{5/4}}{5^{5/4}\calB^{1/4}}\sqrt{\st}\OmK =
+  \frac{2\sqrt{2}\times3^{3/4}}{5^{5/4}}\sqrt{\fg\left(2\fdust-1\right)K_z\st}\OmK.
+\end{align}
+and the corresponding (dimensional) real frequency satisfies
+$\real{(\sigma)} =  \sqrt{3}s/2$.
+
+From the above result we see that as $\taus\to0$, the most unstable
+radial wavelength diverges, $K_x\propto \taus^{-1/2} \to \infty$ with
+a vanishing growth rate, $\mathrm{max}(s)\propto \sqrt{\taus}\to 0$.
+
+\subsubsection{Finite phase lag as $\taus\to 0$}  
+An interesting property of the special solutions described above is
+that there is always a phase lag between the Lagrangian pressure and
+density perturbations. Since $\real{(\nu)}>0$, the phase lag may be
+defined as $\varphi = \arg{\left(\Delta P \Delta \rho^*\right)}$. For
+SI we neglect the background density gradients, so $\Delta\rho = 
+\delta\rho$. It may be shown that in the limit of large $K_x$ 
+that $\Delta P \simeq \ii \delta v_x \p_rP/\sigma$. To obtain $\dd v_x$ we use Eq. \ref{streaming_vx} 
+assuming low frequency modes, $|\sigma|\ll \kappa^2$, valid for
+$\taus\to 0$ as shown above. Now, the expression for $\dd v_x$ involves
+the Eulerian pressure perturbation, $\delta P$. For this we use
+Eq. \ref{si_QW1} and assume incompressibility.
+
+With these simplifications (in addition to those already made in the
+previous section), the phase lag $\varphi$ is given via 
+\begin{align}
+  \tan{\varphi} = \frac{\imag{\left(\Delta
+        P\Delta\rho^*\right)}}{\real{\left(\Delta
+        P\Delta\rho^*\right)}}\simeq\frac{K_x\imag{(\nu)}}{2\fg^2\st
+      K_x^2 + K_x \real{(\nu)}}. 
+\end{align}
+From here it is clear for the solutions derived above, where
+$K_x\propto \st^{-1/2}$ and $\nu \propto \st^{1/2}$, that $\varphi$ is
+a constant. Explicitly inserting the solutions give 
+\begin{align}
+\tan{\varphi} = \frac{2\times 3^{3/2}\left(2\fdust-1\right)}{16 -
+  7\fdust}. 
+\end{align}
+For the case shown in Fig. \ref{si_compare_fig} with $\fdust=0.75$, we
+have $\varphi\simeq 26\degr$, comparable to the $\sim 30\degr$
+obtained from numerical solutions.
+
+\subsubsection{Validty of the approximations}
+The solution above require $K_z$ be sufficiently large for
+Eq. \ref{min_kz} but also sufficiently small for
+Eq. \ref{max_kz}. Inserting the above solutions, we find the
+requirement from Eq. \ref{min_kz} becomes 
+\begin{align}
+  \fdust \gg \frac{12}{17}\simeq 0.7.
+\end{align}
+However, since $\fdust \leq 1$, this requirement can only be marginally
+satisfied. We find that this mostly give errors in the real
+frequency (see Fig. \ref{si_compare_fig}), while growth rates are
+still captured. On the other hand, Eq. \ref{max_kz} becomes the trivial 
+inequality \begin{align}
+  1\ll \frac{27}{4}\left(\frac{5}{3}\right)^3\simeq 30,
+\end{align}
+so Eq. \ref{max_kz} is satisified, which justifies the solution for
+$\mu$ in Eq. \ref{mu_explicit}.
+
 \subsection{Spuriously growing epicycles}\label{spurious_epi}
@@
 \begin{align}      %streaming_incompressible
-\ii \sigma^2 - \frac{\sigma}{\epsilon \tstop}  - \left(\ii \kappa^2 +
-  \frac{k_xF_r}{\epsilon}\right) = 0. \label{kz_zero_si}
+%\ii \sigma^2 - \frac{\sigma}{\epsilon \tstop}  - \left(\ii \kappa^2 +
+%  \frac{k_xF_r}{\epsilon}\right) = 0. \label{kz_zero_si}
+\ii\nu^2 - \frac{\nu}{\fdust\st} - \left[\ii + 
+  2K_x\frac{\fg}{\fdust}\left(1-\fdust\right)\right]=0\label{kz_zero_si} 
 \end{align} 
-Neglecting the quadratic term leads to stability, $\imag{\sigma}<0$, 
-as found by \cite{youdin05a} in their full two-fluid analysis. 
+in dimensionless form. Neglecting the quadratic term leads to
+stability, $\imag{(\nu)}<0$, as found by \cite{youdin05a} in their full
+two-fluid analysis.  
 % However, the real and imaginary parts of Eq. \ref{kz_zero_si}  are
@@
 % \end{align*} 
-However, solving Eq. \ref{kz_zero_si} explicitly assuming $\epsilon s \tstop \ll 1$ would yield 
+However, solving Eq. \ref{kz_zero_si} explicitly assuming $\left|\fdust\taus
+\imag{(\nu)}\right| \ll 1$ would yield  
 \begin{align}
-  |\omega| &\simeq \left|k_xF_r\right| \tstop \sim \kappa, \\
-  s &\simeq \epsilon \tstop \left[\left(k_xF_r\tstop\right)^2 - \kappa^2\right],\label{epi_grow}
+%  |\omega| &\simeq \left|k_xF_r\right| \tstop \sim \kappa, \\
+%  s &\simeq \epsilon \tstop \left[\left(k_xF_r\tstop\right)^2 - \kappa^2\right],\label{epi_grow}
+\real{(\nu)}&\simeq 2K_x\left(1-\fdust\right)^2\st \sim O(1)\\
+\imag{(\nu)}&\simeq \fdust \st \left[4K_x^2\left(1-\fdust\right)^4\st^2 -
+1\right].\label{epi_grow}
 \end{align}
-Accordingly, given $k_xF_r$, growth is possible if 
+Accordingly, growth is possible if 
 \begin{align}
-\tstop > \frac{\kappa}{k_x\left|F_r\right|}\equiv t_{\mathrm{s}0}.\label{epi_grow_criterion}
+%\tstop > \frac{\kappa}{k_x\left|F_r\right|}\equiv t_{\mathrm{s}0}.\\%\label{epi_grow_criterion}
+\st > \frac{1}{2K_x\left(1-\fdust\right)^2}\label{epi_grow_criterion}
 \end{align}
-The mode is marginally stable when $\tstop =
-t_{\mathrm{s}0}$. Alternatively, for fixed $\tstop$ growth is enabled
-by a sufficiently large $\left|k_xF_r\right|$. Unlike the streaming instability,
-which is strongly surpressed when $\epsilon=1$, these modes can still
+%The mode is marginally stable when $\tstop =
+%t_{\mathrm{s}0}$. 
+Alternatively, for fixed $\taus$ growth is enabled
+by a sufficiently large $K_x$. Unlike the streaming instability,
+which is strongly surpressed when $\fdust=1/2$, these modes can still
 grow at equal dust-to-gas ratio.
@@
 full two-fluid models \citep{youdin05a}. The discrepency lies in the
-fact that the one-fluid equations, to first order in $\tstop$, are
+fact that the one-fluid equations, to first order in $\taus$, are
 only valid for low-frequency waves with $|\sigma|\lesssim 
@@
 rates are weakly depedent on $K_z$ and are well approximated by
-that in the $K_z=0$ limit, Eq. \ref{epi_grow}, which becomes 
-\begin{align*} 
-  s = \epsilon \left(1 - \tepsilon
-  \right)\taus\OmK^2\left[4K_x^2\left(1-\tepsilon\right)^4\left(\taus\OmK\right)^2
-  -1 \right]
-\end{align*}
-in the above notation. 
+that in the $K_z=0$ limit, Eq. \ref{epi_grow}. 
+%, which becomes 
+%\begin{align*} 
+%  s = \epsilon \left(1 - \tepsilon
+%  \right)\taus\OmK^2\left[4K_x^2\left(1-\tepsilon\right)^4\left(\taus\OmK\right)^2
+%  -1 \right]
+%\end{align*}
+%in the above notation. 
 For the case considerd in Fig. \ref{dusty_growth1} we find

### 2017-05-22 linear.tex
home/work sync

@@
 then $\varphi > 0 $ indicates gas density lagging behind dust density,
-which is true for all the cases. In particular, `linB' indicates that
-growth rates vanish as $\varphi \to 0$.  Thus the streaming
-instability is indeed associated with a phase lag between gas and dust
-density evolutions.
+which is true for all the cases. %In particular, `linB' indicates that
+%growth rates vanish as $\varphi \to 0$.  
+Thus the streaming instability is indeed associated with a phase lag
+between gas and dust density evolutions.
 
-In a future work we will perform a detailed comparison between the
-simplified one-fluid and full two-fluid frameworks in calculating SI.
+In Fig. \ref{si_compare_fig} we compare the most unstable SI 
+mode as a function of $\taus$ at fixed $K_z=30$ and $\epsilon =3$. The
+solid line is obtained from a full, two-fluid linear analysis 
+\citep[similar to][]{youdin05a, kowalik13}. The one-fluid results are
+obtained by solving the dispersion relation
+Eq. \ref{streaming_dispersion}, both numerically (green asterisks) and
+analytically with additional approximations and considerations
+(organge diamonds). The analytic model is described in Appendix 
+\ref{si_dust_rich}.
+
+Both results based on the one-fluid framework compares well with the 
+full, two-fluid analysis, only breaking down at a relatively large
+$\taus\OmK\gtrsim 0.1$. It is interesting to see for large $\taus$ the 
+two-fluid phase lag drops, along with the growth rate. Notcie also the
+phase lag remains finite as $\taus\to0$ because the optimum $K_x$
+diverges (this is explained in Appendix \ref{si_dust_rich}).
+
+In a future work we will perform a more detailed parameter survey to
+compare the simplified one-fluid and full two-fluid frameworks in
+calculating SI.
 
 \begin{figure}
-\includegraphics[width=linewidth]{figures/si_2f_1f_compare}
+\includegraphics[width=\linewidth]{figures/si_2f_1f_compare}
 \caption{{\bf in case the paper isn't long enough} Comparison of the linear
-streaming instability obtained from a full two-fluid analysis and the current 
-one-fluid framework simplified by the terminal velocity approximation. The vertical
+streaming instability between a full two-fluid analysis (solid
+line), the current one-fluid framework simplified by the terminal
+velocity approximation (green asterisks) and an analytic solution to
+the one-fluid dispersion relation in the dust-rich limit (organge
+diamonds, see also Appendix \ref{si_dust_rich}). The vertical 
 wavenumber is fixed and growth rates are maximized over $K_x$. 
-\label{si_compare}}
+\label{si_compare_fig}}
 \end{figure}

### 2017-05-23 linear.tex
wording edits in SI section (main text and new analytic model in appendix)

--- 2017-05-17 linear.tex
@@
 \subsection{Streaming instability}\label{si}
-We now consider 3D but non-self-gravitating disks.  
-In the special case of an unstratified disk (neglecting the vertical
-component of the stellar gravity), we may also Fourier analyze in $z$
-to obtain an algebraic dispersion relation of the form  
+We now consider 3D disks without self-gravity. We neglect the vertical     
+component of the stellar gravity, appropriate for studying regions
+near the disk midplane. This  
+allows us to Fourier analyze in $z$ to obtain 
+an algebraic dispersion relation of the form  
 $\sum_{j=0}^{5}c_j(k_x,k_z)\sigma^j = 0$, where $k_z$ is a real
@@
 
-We solve the full dispersion relation, Eq. \ref{streaming_dispersion} for selected cases where SI  
-growth rates have been 
-calculated analytically from full two-fluid disk models and confirmed
-with particle-gas numerical simulations 
-\citep[namely][]{youdin07b,bai10b}. These authors use normalized
+In Table \ref{si_compare} we solve the full dispersion relation
+(Eq. \ref{streaming_dispersion}) numerically for selected cases where
+analytic SI growth rates have been verified with particle-gas
+numerical simulations \citep[namely][]{youdin07b,bai10b}.  
+Following previous works on SI, we use normalized 
 wavenumbers $K_{x,z} = \eta r k_{x,z}$ where
@@
 \end{align} 
-measures the pressure offset of Keplerian rotation. We use
-$\eta=0.05c_s/r\OmK$. They also define the stopping time as
-$\tau_\mathrm{s}=\tstop/(1-\tepsilon)$.
+measures the pressure offset of Keplerian rotation. We fix 
+$\eta=0.05c_s/r\OmK$. In this section section we also quote the
+particle stopping time $\tau_\mathrm{s}=\tstop/(1-\tepsilon)$.
 
-Table \ref{si_compare} compares the eigenfrequencies obtained from the
-one-fluid dispersion relation to that from the above studies. As  
-expected eigenfrequencies agree better with decreasing $\tstop$ since
-in that limit the mixture behaves more like a single fluid. Most 
-importantly, we find the work done $\mathcal{W}>0$ in all cases, and
-hence find growing oscillations.
+The eigenfrequencies obtained from the one-fluid dispersion relation 
+are compared that from a full, two-fluid analysis \citep[similar 
+to][]{youdin05a, kowalik13}. As expected eigenfrequencies agree better
+with decreasing $\taus$ since in that limit the mixture behaves more 
+like a single fluid. Most importantly, we find the work done
+$\mathcal{W}>0$ in all cases, and hence find growing oscillations.
 
@@
 
-We also checked the growth rates $s$ satisfy 
-\begin{align} 
-  s = \frac{\left|\sigma\right|^2\imag\left(\Delta P
-    \Delta\rho^*/\rho\right)}{2\real\left(\sigma\right)\rho\left(\left|\dd
-  v_x\right|^2+\left|\dd
-  v_z\right|^2\right)}, \label{si_check}
-\end{align}
-as implied by Eq. \ref{thermal_instability} and
-Eq. \ref{pdv}. %, or directly from
-%Eq. \ref{streaming_mass}---\ref{streaming_vz}. 
-Interestingly, we find in dust-rich disks with 
-$\epsilon>1$, Eq. \ref{si_check} can be satisified with $\Delta
-P\simeq \ii\dd v_x \p_rP/\sigma$, i.e. the radial pressure gradient
-is responsible for growth. Conversely, for `linB' with $\epsilon
-< 1$, one can approximate $\Delta P \simeq \dd P$ in Eq. \ref{si_check}. %This
+% We also checked the growth rates $s$ satisfy 
+% \begin{align} 
+%   s = \frac{\left|\sigma\right|^2\imag\left(\Delta P
+%     \Delta\rho^*/\rho\right)}{2\real\left(\sigma\right)\rho\left(\left|\dd
+%   v_x\right|^2+\left|\dd
+%   v_z\right|^2\right)}, \label{si_check}
+% \end{align}
+% as implied by Eq. \ref{thermal_instability} and
+% Eq. \ref{pdv}. %, or directly from
+% %Eq. \ref{streaming_mass}---\ref{streaming_vz}. 
+% Interestingly, we find in dust-rich disks with 
+% $\epsilon>1$, Eq. \ref{si_check} can be satisified with $\Delta
+% P\simeq \ii\dd v_x \p_rP/\sigma$, i.e. the radial pressure gradient
+% is responsible for growth. Conversely, for `linB' with $\epsilon
+% < 1$, one can approximate $\Delta P \simeq \dd P$ in Eq. \ref{si_check}.
+% This
 %suggests that for SI in gas-dominated disks, the distinction between
@@
 \begin{align*} 
-\varphi \equiv -\sgn(\omega)\arg\left(\Delta P\Delta\rho^*\right)
-        = \sgn(\omega)\arg\left(\Delta\rhog^*\Delta\rhod\right),\notag
+\varphi&\equiv \sgn{\left[\real{(\sigma)}\right]}\arg\left(\Delta
+  P\Delta\rho^*\right).\notag 
+%\notag\\
+%       &= \sgn{\left[\real{(\sigma)}\right]}\arg\left(\Delta\rhog^*\Delta\rhod\right).\notag
 \end{align*}
-%where we have applied the isothermal equation of state for the second equality.
-then $\varphi > 0 $ indicates gas density lagging behind dust density,
-which is true for all the cases. In particular, `linB' indicates that
-growth rates vanish as $\varphi \to 0$.  Thus the streaming
-instability is indeed associated with a phase lag between gas and dust
-density evolutions.
-
-In a future work we will perform a detailed comparison between the
-simplified one-fluid and full two-fluid frameworks in calculating SI.
-
-\begin{figure}
-\includegraphics[width=linewidth]{figures/si_2f_1f_compare}
-\caption{{\bf in case the paper isn't long enough} Comparison of the linear
-streaming instability obtained from a full two-fluid analysis and the current 
-one-fluid framework simplified by the terminal velocity approximation. The vertical
-wavenumber is fixed and growth rates are maximized over $K_x$. 
-\label{si_compare}}
-\end{figure}
-%"generalized stokes number" would include other problem parameters,
-%e.g. wavenumber
+(Note that $\Delta P = \delta P + \ii \dd v_x \p_rP/\sigma$, and it is
+important to include the global radial pressure gradient here.) Then 
+$\varphi > 0 $ indicates gas pressure lagging behind total density, 
+which is true for all the cases. Thus SI is indeed associated with
+such a phase lag.
 
@@
 
+In Fig. \ref{si_compare_fig} we calculate the most unstable SI 
+mode as a function of $\taus$ at fixed $K_z=30$ and $\epsilon =3$. We
+compare results between the full, two-fluid linear analysis and the
+one-fluid framework. We also include an analytic model, described in Appendix 
+\ref{si_dust_rich}, based on the one-fluid dispersion relation with 
+additional approximations (orange diamonds).
+
+Both results based on the one-fluid framework compares well with the 
+full, two-fluid analysis, only breaking down at a relatively large
+$\taus\OmK\gtrsim 0.1$. For larger $\taus$ the two-fluid phase
+lag drops, along with the growth rate. This suggest that a
+non-vanishing phase lag is necessary for instability. However, the
+magnitude of the phase lag does not correlate with growth rates. In
+fact, $\varphi$ remains finite as $|\sigma|, \taus\to0$ (but
+non-zero). This arises because the optimum $K_x\propto \taus^{-1/2}$
+diverges (see Appendix \ref{si_dust_rich}).
+
+In a future work we will perform a more detailed parameter survey to
+compare the simplified one-fluid and full two-fluid frameworks in
+calculating SI.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{figures/si_2f_1f_compare}
+\caption{Comparison of the linear
+streaming instability between a full two-fluid analysis (solid
+line), the one-fluid framework simplified by the terminal
+velocity approximation (green asterisks) and an analytic solution to
+the one-fluid dispersion relation in the dust-rich limit (orange
+diamonds, see also Appendix \ref{si_dust_rich}). The vertical 
+wavenumber is fixed and growth rates are maximized over $K_x$. 
+\label{si_compare_fig}}
+\end{figure}
+
+%"generalized stokes number" would include other problem parameters,
+%e.g. wavenumber
+
 % \subsection{Eulerian interpretation}

### 2017-05-23 model.tex
wording edits in SI section (main text and new analytic model in appendix)

@@
 
-The two fluids interact via a drag force parameterized by a stopping
-time $\tstop$ 
-such that 
+The two fluids interact via a drag force parameterized by the relative stopping 
+time $\tstop$ such that 
 \begin{align}  
@@
 Note that $\tstop$ 
-differs slightly from the stopping time $\tau_\mathrm{s} =
+differs slightly from the \emph{particle} stopping time $\tau_\mathrm{s} =
 \tstop\rho/\rhog$ used in some studies
@@
 
-A parcel of the isothermal dusty-gas mixture does 
+A parcel of the strictly isothermal dusty-gas mixture does 
 postive work if  

### 2017-05-23 appendix.tex

@@
 Here we seek analytic solutions for the streaming instability by 
-examining limiting cases and with further approximations. We shall fix
-$k_z$ and $\fdust>1/2$ (the dust-rich case), and maximize growth rates
-over $k_x$. We consider small stopping times, $\tau_\mathrm{s}\to0$,
-and assume that the optimum $k_x\to\infty$ but that
-$k_x\tau_\mathrm{s}$ remain bounded.
+examining limiting cases and with additional approximations. We will
+fix $k_z, \fdust$ and maximize growth rates over $k_x$. It turns out
+simple solutions exist in the dust-rich case with 
+$\fdust$ near unity. We consider small stopping times, 
+$\tau_\mathrm{s}\to0$, and assume that the corresponding optimum
+$k_x\to\infty$.
 
@@
 relation, Eq. \ref{streaming_incompressible}. We assume a Keplerian
-disk and neglect the quartic  
-term because it is higher order in $\tau_\mathrm{s}$ (see also
-\S\ref{spurious_epi})  In dimensionless form: 
+disk ($\Omega=\kappa=\OmK$), and low frequency  modes 
+($|\sigma|\ll \OmK$), which leads to neglecting the quartic  
+term. (This is also necessary to avoid spurious modes, see
+\S\ref{spurious_epi}.)  In dimensionless form:   
 \begin{align}
@@
 \end{align}
-where $\nu = \sigma/\Omega$, $\st = \taus\Omega$, $\fg = 1-\fdust$ (the gas
-fraction), and recall $K_{x,z}=k_{x,z}\eta r$. We have used $F_r =
+where $\nu = \sigma/\OmK$, $\st = \taus\OmK$, $\fg = 1-\fdust$ (the gas
+fraction), and recall $K_{x,z}=\eta r k_{x,z}$. We have used $F_r =
 2\eta \fg r \Omega^2$ and considered large $K_x$.
 
-Full analytic expressions for cubic roots are unwieldy. To proceed
-in an a tractable manner, let us \emph{assume} at this stage that the
-quadratic term can be neglected compare to the last term. This amounts
-to assuming a  sufficiently \emph{large} $K_z$ such that :
+Analytic expressions for cubic roots are unwieldy. To keep the problem
+tractable, let us \emph{assume} at this stage that the quadratic term
+can be neglected compare to the last term. That is, 
+%This amounts 
+%to assuming a  sufficiently \emph{large} $K_z$ such that :
 \begin{align}
@@
       \fdust\right)\right| \left|K_x\nu\right|^2 \ll
-  K_z^2\left(2\fdust - 1\right) \label{min_kz}
+  K_z^2\left(2\fdust - 1\right). \label{min_kz}
 \end{align}
-This will be checked later. (For fixed $\fdust$ and $K_x\to\infty$, the
-imaginary term can be neglected.)
+(For fixed $\fdust$ but allowing $K_x\to\infty$, the imaginary term can be
+neglected.) We show in \S\ref{si_dust_rich_approx} that this
+simplification is valid for dust-rich disks.
 
@@
 solved with Vieta's substitution 
-\begin{align*}
-\nu = \mu + \frac{\calP}{3\mu},
-\end{align*}
+\begin{align}
+\nu = \mu + \frac{\calP}{3\mu},\label{si_vieta}
+\end{align}
 then Eq. \ref{depressed_cubic} becomes a quadratic for $\mu^3$,
@@
 \begin{align}
-  \mu^3 &= \frac{\calQ}{2}\left(1\pm \sqrt{1 -
-      \frac{4\calP^3}{27\calQ^2}}\right)\notag\\
-  & = \st \fg (1- 2\fdust)\frac{K_z^2}{K_x}\left\{1 \pm \sqrt{1 -
+  \mu^3 = \frac{\calQ}{2}\left(1\pm \sqrt{1 -
+      \frac{4\calP^3}{27\calQ^2}}\right) 
+   = \st \fg (1- 2\fdust)\frac{K_z^2}{K_x}\left\{1 \pm \sqrt{1 -
       \frac{K_z^2}{27K_x^2\left[\st K_x \fg
@@
 \end{align}
-At this point we assume $K_z$ is sufficiently \emph{small} such that
-the term $\propto K_z^2$ inside the square root may be neglected. That
-is, 
+At this point we assume the term $\propto K_z^2$ inside the square
+root may be neglected. That is, 
 \begin{align}
@@
 \end{align}
-This will be checked later. Remembering that we are considering the
-dust-rich case so that $1-2\fdust<0$, the solution for $\mu$ is   
+We show in \S\ref{si_dust_rich} that this is readily satisfied. Then
+$\mu \simeq \mathcal{Q}^{1/3}$.
+
+Now, remembering that we are considering the dust-rich limit with 
+$1-2\fdust<0$, the explicit solution for $\mu$ is    
 \begin{align}
@@
 \end{align}
-We have chosen the root for instability. We can now write down the
-solution for the eigenfrequency $\nu$:
+We have chosen the complex root for instability. Inserting this into
+Eq. \ref{si_vieta} gives the eigenfrequency $\nu$:
 \begin{align}
-\real{(\nu)} &= \frac{1}{2}\left(|\mu| + \frac{K_z^2}{3
-    K_x^2|\mu|}\right)\notag\\
+\real{(\nu)} %&= \frac{1}{2}\left(|\mu| + \frac{K_z^2}{3
+%    K_x^2|\mu|}\right)\notag\\
 &= \frac{1}{2}\left(\calA \st^{1/3}K_x^{-1/3} + \calB
     \st^{-1/3}K_x^{-5/3}\right),\\
-\imag{(\nu)} &= \frac{\sqrt{3}}{2}\left(|\mu| - \frac{K_z^2}{3
-    K_x^2|\mu|}\right),\notag\\
+\imag{(\nu)} %&= \frac{\sqrt{3}}{2}\left(|\mu| - \frac{K_z^2}{3
+%    K_x^2|\mu|}\right),\notag\\
 &= \frac{\sqrt{3}}{2}\left(\calA \st^{1/3}K_x^{-1/3} - \calB
@@
   K_x = \left(\frac{5\calB}{\calA}\right)^{3/4}\st^{-1/2} =
-  \left(\frac{5}{3}\right)^{3/4}\sqrt{\frac{K_z}{2\fg\left(2\fdust-1\right)\st}}. 
+  \left(\frac{5}{3}\right)^{3/4}\sqrt{\frac{K_z}{2\fg\left(2\fdust-1\right)\st}}. \label{si_dust_rich_optkx}
 \end{align}
@@
 \begin{align}
-  \mathrm{max}\left(s\right) =
-  \frac{2\sqrt{3}\calA^{5/4}}{5^{5/4}\calB^{1/4}}\sqrt{\st}\OmK =
-  \frac{2\sqrt{2}\times3^{3/4}}{5^{5/4}}\sqrt{\fg\left(2\fdust-1\right)K_z\st}\OmK.
+  \mathrm{max}\left[\imag{(\nu)}\right] =
+  \frac{2\sqrt{3}\calA^{5/4}}{5^{5/4}\calB^{1/4}}\sqrt{\st} =
+  \frac{2\sqrt{2}\times3^{3/4}}{5^{5/4}}\sqrt{\fg\left(2\fdust-1\right)K_z\st}.\label{si_dust_rich_grow}
 \end{align}
-and the corresponding (dimensional) real frequency satisfies
-$\real{(\sigma)} =  \sqrt{3}s/2$.
-
-From the above result we see that as $\taus\to0$, the most unstable
-radial wavelength diverges, $K_x\propto \taus^{-1/2} \to \infty$ with
-a vanishing growth rate, $\mathrm{max}(s)\propto \sqrt{\taus}\to 0$.
+and the real frequency $\real{(\nu)} =
+(\sqrt{3}/2)\imag{(\nu)}$. We see that as $\taus\to0$, the most unstable
+radial wavenumber diverges, $K_x\propto \taus^{-1/2} \to \infty$ with 
+a vanishing growth rate, $\mathrm{max}(s)\propto \sqrt{\taus}\to
+0$. This corresponds to instability at arbitrarily small radial
+lengthscales.
 
@@
 density perturbations. Since $\real{(\nu)}>0$, the phase lag may be
-defined as $\varphi = \arg{\left(\Delta P \Delta \rho^*\right)}$. For
-SI we neglect the background density gradients, so $\Delta\rho = 
-\delta\rho$. It may be shown that in the limit of large $K_x$ 
-that $\Delta P \simeq \ii \delta v_x \p_rP/\sigma$. To obtain $\dd v_x$ we use Eq. \ref{streaming_vx} 
-assuming low frequency modes, $|\sigma|\ll \kappa^2$, valid for
-$\taus\to 0$ as shown above. Now, the expression for $\dd v_x$ involves
-the Eulerian pressure perturbation, $\delta P$. For this we use
-Eq. \ref{si_QW1} and assume incompressibility.
+defined as $\varphi = \arg{\left(\Delta P \Delta \rho^*\right)}$. %For
+%SI we neglect the background density gradients, so $\Delta\rho
+%\simeq  \delta\rho$ but retain the background pressure gradient. 
+(Note that we may set $\Delta\rho = 1$ without loss of generality.) 
+It may be shown that in the limit of large $K_x$  that $\Delta P
+\simeq \ii \delta v_x \p_rP/\sigma$. To obtain $\dd v_x$ we use
+Eq. \ref{streaming_vx} in the low frequency limit. The expression for $\dd v_x$ involves
+the Eulerian pressure perturbation, $\delta P$, for which we use 
+Eq. \ref{si_QW1} and assume incompressibility ($\zeta\to0$).
 
-With these simplifications (in addition to those already made in the
-previous section), the phase lag $\varphi$ is given via 
+With these additional simplifications the phase lag $\varphi$ is given
+via  
 \begin{align}
@@
 \end{align}
-From here it is clear for the solutions derived above, where
+From here it is clear for the above solutions, where
 $K_x\propto \st^{-1/2}$ and $\nu \propto \st^{1/2}$, that $\varphi$ is
@@
 
-\subsubsection{Validty of the approximations}
-The solution above require $K_z$ be sufficiently large for
-Eq. \ref{min_kz} but also sufficiently small for
-Eq. \ref{max_kz}. Inserting the above solutions, we find the
-requirement from Eq. \ref{min_kz} becomes 
+\subsubsection{Validty of the approximations}\label{si_dust_rich_approx}
+Inserting the solutions for the most unstable mode 
+%(Eq. \ref{si_dust_rich_optkx}---\ref{si_dust_rich_grow} with ) 
+we find the assumption in Eq. \ref{min_kz} becomes  
 \begin{align}
@@
 \end{align}
-However, since $\fdust \leq 1$, this requirement can only be marginally
-satisfied. We find that this mostly give errors in the real
+Since $\fdust \leq 1$, this requirement can only be marginally
+satisfied. However, we find that this mostly give errors in the real
 frequency (see Fig. \ref{si_compare_fig}), while growth rates are
-still captured. On the other hand, Eq. \ref{max_kz} becomes the trivial 
+still captured correctly. 
+On the other hand, the assumption of Eq. \ref{max_kz} becomes the trivial
 inequality \begin{align} 
@@
 \end{align}
-so Eq. \ref{max_kz} is satisified, which justifies the solution for
+so Eq. \ref{max_kz} is satisified, which justifies the approximate solution for 
 $\mu$ in Eq. \ref{mu_explicit}.

### 2017-05-28 linear.tex
local copy

@@
 \end{align*}
+%where we have applied the isothermal equation of state for the second equality.
+then $\varphi > 0 $ indicates gas density lagging behind dust density,
+which is true for all the cases. In particular, `linB' indicates that
+growth rates vanish as $\varphi \to 0$.  Thus the streaming
+instability is indeed associated with a phase lag between gas and dust
+density evolutions.
+
+In a future work we will perform a detailed comparison between the
+simplified one-fluid and full two-fluid frameworks in calculating SI.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{figures/si_2f_1f_compare}
+\caption{{\bf in case the paper isn't long enough} Comparison of the linear
+streaming instability obtained from a full two-fluid analysis and the current 
+one-fluid framework simplified by the terminal velocity approximation. The vertical
+wavenumber is fixed and growth rates are maximized over $K_x$. 
+\label{si_compare_full}}
+\end{figure}
+%"generalized stokes number" would include other problem parameters,
+%e.g. wavenumber
+
+=======
 (Note that $\Delta P = \delta P + \ii \dd v_x \p_rP/\sigma$, and it is
@@
 such a phase lag. 
+>>>>>>> 138f926083725c30a16461d3f173ecb9280c9d0f
 

### 2017-05-28 linear.tex
local copy

--- 2017-05-23 linear.tex
@@
 \end{align*}
+
 (Note that $\Delta P = \delta P + \ii \dd v_x \p_rP/\sigma$, and it is
@@
 such a phase lag.
+
+In a future work we will perform a detailed comparison between the
+simplified one-fluid and full two-fluid frameworks in calculating SI.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{figures/si_2f_1f_compare}
+\caption{{\bf in case the paper isn't long enough} Comparison of the linear
+streaming instability obtained from a full two-fluid analysis and the current 
+one-fluid framework simplified by the terminal velocity approximation. The vertical
+wavenumber is fixed and growth rates are maximized over $K_x$. 
+\label{si_compare_full}}
+\end{figure}
+%"generalized stokes number" would include other problem parameters,
+%e.g. wavenumber
 

### 2017-06-12 linear.tex
minor aesthetic changes

--- 2017-05-23 linear.tex
@@
 \end{align*}
-(Note that $\Delta P = \delta P + \ii \dd v_x \p_rP/\sigma$, and it is
-important to include the global radial pressure gradient here.) Then 
+(Note that it is important to include 
+the global radial pressure gradient in $\Delta P = \delta P + \ii \dd v_x \p_rP/\sigma$.) Then 
 $\varphi > 0 $ indicates gas pressure lagging behind total density, 

### 2017-06-12 linear.tex
local save

--- 2017-05-28 linear.tex
@@
 \end{align*}
-
 (Note that $\Delta P = \delta P + \ii \dd v_x \p_rP/\sigma$, and it is

### 2017-06-12 linear.tex

--- 2017-06-12 linear.tex
@@
 
+In a future work we will perform a detailed comparison between the
+simplified one-fluid and full two-fluid frameworks in calculating SI.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{figures/si_2f_1f_compare}
+\caption{{\bf in case the paper isn't long enough} Comparison of the linear
+streaming instability obtained from a full two-fluid analysis and the current 
+one-fluid framework simplified by the terminal velocity approximation. The vertical
+wavenumber is fixed and growth rates are maximized over $K_x$. 
+\label{si_compare_full}}
+\end{figure}
+%"generalized stokes number" would include other problem parameters,
+%e.g. wavenumber
+
 %{\bf suppose large scale pressure grad drops outwards. 

### 2017-06-15 paper.tex
added analytic model in appendix for dust-rich limit
affliation re-order

@@
 \newcommand{\taus}{\tau_\mathrm{s}}
+\newcommand{\st}{\mathrm{St}}
 \newcommand{\w}{ \widetilde{W}}
+\newcommand{\fdust}{f_\mathrm{d}}
+\newcommand{\fg}{f_\mathrm{g}}
 \newcommand{\dphi}{\dd \Phi}
@@
 \newcommand{\Hd}{H_\mathrm{d}}
+\newcommand{\calP}{\mathcal{P}}
+\newcommand{\calQ}{\mathcal{Q}}
+\newcommand{\calA}{\mathcal{A}}
+\newcommand{\calB}{\mathcal{B}}
 \newcommand{\seff}{S_\mathrm{eff}}
@@
 \author{Min-Kai Lin\altaffilmark{1,2,3}}
-\author{Andrew N. Youdin\altaffilmark{1}}
+\author{Andrew N. Youdin\altaffilmark{2}}
 %\affil{Department of Astronomy and Steward Observatory, 
@@
 %Taipei 10617, Taiwan}
-\altaffiltext{1}{Department of Astronomy and Steward Observatory, 
-University of
-Arizona, Tucson, AZ 85721, USA}
-\altaffiltext{2}{Steward Theory Fellow}
-\altaffiltext{3}{Institute of Astronomy and Astrophysics, Academia Sinica, 
-Taipei 10617, Taiwan
-}
+\altaffiltext{1}{Institute of Astronomy and Astrophysics, Academia Sinica, Taipei 10617, Taiwan}
+\altaffiltext{2}{Department of Astronomy and Steward Observatory, University of Arizona, Tucson, AZ 85721, USA}
+\altaffiltext{3}{Steward Theory Fellow}
+
 \email{mklin@asiaa.sinica.edu.tw}

### 2017-06-29 paper.tex
minor tweaks to abstract

--- 2017-05-15 paper.tex
@@
   to strong dust-gas friction. Consequently, tightly-coupled dust
-  particles almost follows the gas flow. 
-  This is analogous to entropy conservation
-  following an ideal gas.   
-  Motivated by this similarity, we develop a thermodynamic 
+  particles almost follow the gas flow.
+  %AY: Edited b/c "This is" was too vague
+  This near conservation of dust-to-gas ratio along streamlines
+  is analogous to the near conservation of entropy along flows of (dust-free) gas with weak heating and cooling.   
+  We develop this thermodynamic analogy into a
   framework to study dusty gas dynamics in protoplanetary disks. We  
@@
 %  gas pressure and dust densities; 
-\item dust-loading introduce buoyancy forces that generally 
+\item dust-loading introduces buoyancy forces that generally 
   stabilizes the vertical shear 
@@
 %finding 
-potential dusty analogs of other hydrodynamic 
+ dusty analogs of other hydrodynamic 
 processes (e.g. Rossby wave instability, convective overstability, and 
-zombie vortices), in simulating dusty protoplanetary disks
-using purely hydrodynamic codes, as well as possible generalizations
-of the dusty/adiabatic gas analogy.
+zombie vortices), how to simulate dusty protoplanetary disks
+with minor tweaks to purely hydrodynamic codes, as well as possible generalizations
+of our analogy between dust-laden and nearly adiabatic flows.
 

### 2017-06-29 paper.tex
added GW73 as original source (I believe, but perhaps Safronov...) of dust KHI (before Weidenschilling 1980, etc.).  Treatment was not modern (Ekman layers etc) but was the basic idea and historically important.
Minor tweaks, switching J11 for J06 as a pioneering work with two way drag forces on lagrangian particles

@@
 Kelvin-Helmholtz instabilities
-\citep{chiang08,barranco09,lee10}. These lead to structure and
-planetesimal formation, as well as turbulence. 
+\citep{goldreich1973, chiang08,barranco09,lee10}. These instabilities 
+can trigger turbulence, and (in different regimes) promote or inhibit 
+planetesimal formation. 
 More recently, new 
-dusty instabilities have also been uncovered through numerical
-simulations \citep{loren15,loren16,lamb16}, but these are not yet well
+dusty instabilities have appeared in numerical
+simulations \citep{loren15,loren16,lamb16}, but their physical underpinnings are not yet well
 understood.
@@
 Lagragian dust particles
-\citep{nelson10,bai10,johansen11,yang14,zhu14,gibbons15,simon16,baruteau16}. 
+\citep{johansen2006, nelson10,bai10,yang14,zhu14,gibbons15,simon16,baruteau16}. 
 The equation of motion for each particle is solved 
@@
 
-Another approach is to model the dust population as a continous, pressureless
+Another approach is to model the dust population as a continuous, pressureless
 fluid 

### 2017-06-30 paper.tex
more mods to introduction.  need to proofread before MK judges...

--- 2017-06-29 paper.tex
@@
 Kelvin-Helmholtz instabilities
-\citep{goldreich1973, chiang08,barranco09,lee10}. These instabilities 
+\citep{goldreich73, chiang08,barranco09,lee10}. These instabilities 
 can trigger turbulence, and (in different regimes) promote or inhibit 
@@
 Lagragian dust particles
-\citep{johansen2006, nelson10,bai10,yang14,zhu14,gibbons15,simon16,baruteau16}. 
+\citep{johansen06, nelson10,bai10,yang14,zhu14,gibbons15,simon16,baruteau16}. 
 The equation of motion for each particle is solved 
@@
 can take advantage of existing numerical methods/codes for
-simulating pure gas dynamics.
+simulating pure gas dynamics. A practical difficulty with this approach is the need to numerical stabilize the pressureless dust fluid with (e.g.) artificial diffusion.
 
@@
 
-It is possible to re-formulate the two-fluid equations into that for a
-single fluid with total density $\rho$ and center-of-mass
-velocity $\bm{v}$ of the dusty gas mixture
-\citep{laibe14}. The standard hydrodynamic equations for
-$\rho$ and $\bm{v}$ are supplemented by evolutionary equations for
-the dust-to-gas ratio $\rhod/\rhog$ and the relative dust-gas velocity
-$\Delta\bm{v}\equiv \bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$. The latter also contributes to 
-additional source terms in the momentum and energy equations.
+%AY: reworded b/c (a) I didn't think it was clear enough that the one fluid simplification wasn't exact and (b) that this tight coiupling limit needed to be emphasized more.
 
-The one-fluid approach is appropriate for 
-simulating gas tightly coupled to dust particles with small $\tstop$.  
-This can be numerically challenging for two-fluid models and 
-following Lagragian particles; but is straight forward with the 
-one-fluid formulation \citep{price15}.  The single fluid framework 
-also has the important advantage that it is 
-amenable to standard analytical methods \citep{youdin05a,jacquet11}.  
-This is the necessary approach to gain physical insight.
+The two-fluid equations for dust and gas can be reformulated into 
+equivalent dynamical equations for the center of mass velocity $\bm{v}$
+and the relative dust-gas velocity  $\Delta\bm{v}\equiv \bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$ 
+\citep{youdin05a}. In this reformulation, continuity equations for dust and gas can be replaced 
+by the mass conservation of total density $\rho$ plus the evolution of the dust-to-gas ratio 
+(or dust fraction, \citealp{laibe14}).
 
-In this work, we are motivated to develop a general analytical
-framework to study the stability of dusty protoplanetary 
-disks by drawing on the physical and mathematical parallel between
-dusty gas and standard hydrodynamics. This is possible in certain limits of astrophysical
-interest, allowing us to interpret dusty phenomena through standard
-hydrodynamics.
+The advantage of this reformulation arises in the tight coupling limit of  $\tstop \ll 1/\Omega$, where the orbital frequency $\Omega$ sets the characteristic timescale in many disk dynamics problems.
+Models with two fluids or with Lagrangian solids are numerically difficult to evaluate in this regime, due to stiff drag forces.  
+In the reformulated equations, however, the center of mass motion does not experience drag forces and thus is not stiff.
+Moreover the relative motion satisfies the terminal-velocity approximation and can be eliminated from the equations.
+This single fluid framework for a well-coupled dust-gas mixture facilitates numerical calculations \citep{price15}.
+and also simplifies analytic calculations \citep{youdin05a,jacquet11}, which is useful for gaining insight.
 
-We consider tightly coupled dust particles with $\tstop\Omega\ll
-1$, where $\Omega$ is the disk orbital frequency. We may then use the
-`terminal velocity approximation' to set the dust-gas relative drift
-to $\Delta \bm{v} = \tstop\nabla
-P/\rhog$, where $P$ is the gas pressure \citep{youdin05a, 
-  jacquet11,laibe14}. This approximation reflects the physical result
-that dust particles tend to drift towards pressure maximum.
+%It is possible to re-formulate the two-fluid equations into that for a
+%single fluid with total density $\rho$ and center-of-mass
+%velocity $\bm{v}$ of the dusty gas mixture
+%\citep{laibe14}. The standard hydrodynamic equations for
+%$\rho$ and $\bm{v}$ are supplemented by evolutionary equations for
+%the dust-to-gas ratio $\rhod/\rhog$ and the relative dust-gas velocity
+%$\Delta\bm{v}\equiv \bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$. The latter also contributes to 
+%additional source terms in the momentum and energy equations.  
+%
+%The one-fluid approach is appropriate for 
+%simulating gas tightly coupled to dust particles with small $\tstop$.  
+%This can be numerically challenging for two-fluid models and 
+%following Lagragian particles; but is straight forward with the 
+%one-fluid formulation \citep{price15}.  The single fluid framework 
+%also has the important advantage that it is 
+%amenable to standard analytical methods \citep{youdin05a,jacquet11}.  
+%This is the necessary approach to gain physical insight.
 
- We also consider irradiated protoplanetary disks 
- in which the temperature is set externally
-\citep{chiang97,stam08}. This allows us to adopt a locally isothermal
-equation of state, $P = c_s^2\rhog$, where $c_s$ is a prescribed
-sound-speed profile. This physically represents short cooling times
-\citep{lin15}.
+%AY: Attempted to clarify a bit by separating more the analogy and it's applications.
 
-Under these two approximations, we show that the equations of 
-isothermal dusty gas dynamics are equivalent to that of adiabatic, pure gas
-dynamics (Fig. \ref{concept_chart}). This is because tightly-coupled dust particles are almost
-entrained in the gas, which is analogous to the advection
-of entropy following an adiabatic fluid. We show that the   
-effective entropy of isothermal dusty gas scales with its gas
-fraction. Finite dust-gas drag act as effective heating/cooling: the relative
+In this work, we develop a parallel between this single-fluid description of dust-gas mixtures and standard hydrodynamics.
+Specifically we recast the evolution of the dust-to-gas ratio as an energy equation for the thermal content of a fluid.
+This parallel is most precise when the gas in the dust/gas mixture obeys a locally isothermal equation of state for the gas, $P = c_s^2\rhog$, where $c_s$ is a prescribed sound-speed profile.
+Thus we make this approximation which holds when cooling times are short \citep{lin15}, as with protoplanetary disks
+ whose temperature is set by external irradiation
+\citep{chiang97,stam08}.
+
+%are motivated to develop a general analytical
+%framework to study the stability of dusty protoplanetary 
+%disks by drawing on the physical and mathematical parallel between
+%dusty gas and standard hydrodynamics. This is possible in certain limits of astrophysical
+%interest, allowing us to interpret dusty phenomena through standard
+%hydrodynamics.
+
+%AY: a bit detailed for an introduction, plus the idea of terminal velocity needed to be introduced earlier as the basis of the slngle fluid equations, I thought.
+%We consider tightly coupled dust particles with $\tstop\Omega\ll
+%1$, where $\Omega$ is the disk orbital frequency. We may then use the
+%`terminal velocity approximation' to set the dust-gas relative drift
+%to $\Delta \bm{v} = \tstop\nabla
+%P/\rhog$, where $P$ is the gas pressure \citep{youdin05a, 
+%  jacquet11,laibe14}. This approximation reflects the physical result
+%that dust particles tend to drift towards pressure maximum.
+
+% We also consider irradiated protoplanetary disks 
+% in which the temperature is set externally
+%\citep{chiang97,stam08}. This allows us to adopt a locally isothermal
+%equation of state, $P = c_s^2\rhog$, where $c_s$ is a prescribed
+%sound-speed profile. This physically represents short cooling times
+%\citep{lin15}.
+
+Under these two approximations of strong drag and isothermal gas, we show that the equations of 
+ dusty gas dynamics are equivalent to that of nearly adiabatic (i.e. slowly cooling) gas
+dynamics without dust (Fig. \ref{concept_chart}).  The basis of the analogy is that the entrainment 
+of tightly-coupled dust particles in gas flows is similar to the advection
+of entropy in an adiabatic fluid. %We show that the   
+%effective entropy of isothermal dusty gas scales with its gas
+%fraction.  %AY: too technical for introduction.
+Finite dust-gas drag act as effective heating/cooling: the relative
 drift between dust and gas causes a fluid parcel to exchange dust
 particles with its surroundings; an effect similar to heat exchange
-between a gas parcel and its surroundings.
+between a gas parcel and its surroundings.   Needless to say the effective heating term must have a 
+specific form, which we derive, for the equivalency to hold.
 
@@
 
-By re-casting the dusty problem into standard  
-hydrodynamics, we can apply established methods to generalize previous stability 
+The purpose of our physical and mathematical parallel between
+dusty gas and standard hydrodynamics is to gain a deeper understanding of the dynamics
+and stability of dusty disks.  
+We can apply established methods to generalize previous stability 
 results for pure gas dynamics to dusty disks.  
@@
 
-This paper is organized as follows. We first present, without
-derivation, the correspondence between dusty gas and     
-ideal hydrodynamics in \S\ref{key_idea}. In \S\ref{grow_osc} 
-we discuss a key idea in fluid dynamics, relevant to our 
-discussion on  SI and VSI: instabilities associated with
+This paper is organized as follows. 
+%AY: or not? %We first present, without
+%derivation, the correspondence between dusty gas and     
+%ideal hydrodynamics in \S\ref{key_idea}. 
+In \S\ref{grow_osc} 
+we describe a key insight of our study, that pressure (i.e. ``PdV") work explains the basis of several instabilities, including SI and VSI, in terms of
 pressure-density phase lags.   
@@
 %as SI and VSI, have thermodynamical interpretations.  
-
-We begin our formal discussion in \S\ref{setup} by listing the 
-basic equations describing tightly-coupled dusty gas. We show that for a
-fixed equation of state, these are in 
-fact the usual hydrodynamic equations in disguise. 
-We describe equilibrium solutions in \S\ref{eqm} that are used in
-explicit calculations. In \S\ref{limits} we discuss 
+We begin our formal discussion in \S\ref{setup} by transforming the
+basic equations describing tightly-coupled dusty gas into  hydrodynamic equations with a special cooling function. 
+We derive equilibrium solutions in \S\ref{eqm}, which we use in our explicit stability calculations. In \S\ref{limits} we analyze the 
 general stability properties of dusty disks. We then move onto
 explicit stability analyses. In  
-\S\ref{linear_problem} we list the linearized equations and show that 
-they reproduce previous results for SGI and SI. In \S\ref{results} we
+\S\ref{linear_problem} we show that our linearized equations 
+reproduce the SGI and SI in the tight coupling regime. In \S\ref{results} we
 extend the VSI to dusty disks. We   

### 2017-06-30 paper.tex
home backup

--- 2017-06-29 paper.tex
@@
 \author{Min-Kai Lin\altaffilmark{1,2,3}}
-\author{Andrew N. Youdin\altaffilmark{2}}
+\author{Andrew N. Youdin\altaffilmark{1}}
 %\affil{Department of Astronomy and Steward Observatory, 
@@
 %Taipei 10617, Taiwan}
-\altaffiltext{1}{Institute of Astronomy and Astrophysics, Academia Sinica, Taipei 10617, Taiwan}
-\altaffiltext{2}{Department of Astronomy and Steward Observatory, University of Arizona, Tucson, AZ 85721, USA}
-\altaffiltext{3}{Steward Theory Fellow}
-
+\altaffiltext{1}{Department of Astronomy and Steward Observatory, 
+University of
+Arizona, Tucson, AZ 85721, USA}
+\altaffiltext{2}{Steward Theory Fellow}
+\altaffiltext{3}{Institute of Astronomy and Astrophysics, Academia Sinica, 
+Taipei 10617, Taiwan
+}
 \email{mklin@asiaa.sinica.edu.tw}

### 2017-07-01 model.tex
deleted original section two; remixed original section 5 and 6 (now 4 and 5) so the stability discussion goes from general to special, from not needing explicit background to explicit dust distributions

@@
 \begin{align}
-  \nabla^2\psi = 4 \pi G \rho.
+  \nabla^2\psi = 4 \pi G \rho.\label{poisson}
 \end{align}
@@
 
-\subsection{Entropy of an isothermal dusty gas }
+\subsection{Entropy and buoyancy of isothermal dusty gas }\label{dusty_entropy}
 
@@
 
+%\subsection{Dusty buoyancy forces}
+Having identified the entropy of an isothermal dusty gas, we can also
+define the disk vertical buoyancy frequency $N_z$ as   
+%$N_z$ of the mixture is given by 
+\begin{align}
+  N_z^2 \equiv - \frac{1}{\rho}\frac{\p P}{\p z}\frac{\p \seff}{\p z}
+  =c_s^2(r) \frac{\p \ln{\rhog}}{\p z}\frac{\p \tepsilon}{\p z},
+%  \quad 
+ % N_z^2 \equiv - \frac{1}{\rho}\frac{\p P}{\p z}\frac{\p \seff}{\p z}.
+%&=
+  %\frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\frac{\p\ln\rhog}{\p 
+%  z}\frac{\p\epsilon}{\p z} \\ &
+  %                                =
+ % \frac{\epsilon}{\left(1+\epsilon\right)^2}\left(\frac{z}{\Htilde}\right)^2\OmK^2\notag,  
+\end{align}
+%Typically in a thin, smooth disk we have $|N_r|\ll |N_z|$. 
+where the second equality assumes a vertically isothermal disk. A
+similar definition exists for the radial buoyancy frequency
+$N_r$. Typically in a thin, smooth disk we have $|N_r|\ll |N_z|$.
+
+A vertical buoyancy force exists even in vertically isothermal
+dusty disks, because coupling gas to dust particles increases
+the fluid's inertia, but pressure (i.e. restoring) forces are
+unchanged. The increased weight of the fluid resists vertical
+oscillations and hence there is an associated buoyancy force. However,
+if $\tstop\neq0$ so the gas-dust coupling is imperfect, then gas is no
+longer `weighed down' by the dust, and can slip past it. Thus finite
+drag diminishes the dust-induced buoyancy. This is similar to reducing
+gas buoyancy through cooling \citep{lin15}.
+
 %{\bf
@@
 
-\section{Dynamical equilibria}\label{eqm}
+\section{General stability criteria for dusty gas}\label{limits}
 
-For a given distribution of the dust-fraction $\tepsilon$ (or
+In this section we discuss the stability properties of the 
+dusty-gas mixture using a variational principle. This method does not
+require explicit solutions to the equilibrium equations. Self-gravity is neglected here.
+
+\subsection{Steady states}\label{eqm}
+We consider axisymmetric steady states as background equilibria to
+perturb. For a given distribution of the dust-fraction $\tepsilon$ (or
 dust-to-gas ratio $\epsilon$), the 
@@
 (Eq. \ref{eos}). An explicit solution is presented in 
-\S\ref{steady_state}.  
+\S\ref{steady_state}, when we explicitly analyse the stability of
+protoplanetary disks.
+
 However, disk structures obtained this way generally do not satisfy 
 the steady state effective energy Eq. \ref{eff_energy} because typically 
-$\mathcal{C}\neq0$.  
-%for general solutions to 
-%Eq. \ref{steady_momr}---\ref{steady_momz}.
-
-% Dust particles settle to the disk midplane on a timescale 
-% $t_\mathrm{settle}\sim 1/\OmK^2\tstop$, although for tightly-coupled dust
-% particles this is much longer than that to establish vertical
-% hydrostatic equilibrium in the gas component ($\sim 1/\OmK$).  
-% True steady states may be calculated by invoking an underlying gas
-% turbulence, which leads to dust diffusion counter-acting vertical
-% settling \citep{takeuchi02, youdin07}. This complication is beyond the
-% scope of this paper.
-
-In order to have a well-defined stability problem, we formally
+$\mathcal{C}\neq0$.  In order to have a well-defined stability problem, we formally
 consider equilibria with $\mathcal{C}\equiv 0$. Namely      
@@
 such that the evolution of the background disk (e.g. dust-settling) 
-occur on much longer timescales than the 
-instabilities of interest.
-
-%Eq. \ref{steady_dust} makes it difficult to obtain equilibrium
-%solutions explicitly. However, if the diffusive process is slow and
-%can be neglected, then one may just solve
-%Eq. \ref{steady_momr}---\ref{steady_momz} with a prescribed (initial)
-%distribution of the dust fraction $\tepsilon(r,z)$.
-
-\subsection{Disk structure with a prescribed dust distribution}\label{steady_state}  
-For numerical computations, we assume a Gaussian profile in the
-dust-to-gas ratio,    
-\begin{align}\label{dust_gauss}
-  \epsilon(r,z) = \epsilon_0(r)
-  \exp{\left[-\frac{z^2}{2\Htilde^2(r)}\right]}. 
-\end{align}
-%For simplicity we assume the dust-to-gas ratio at the mid-plane
-%$\epsilon_0$, as well as its characteristic scale-height
-%$\widetilde{H}$, are both constant.
-% We take the mid-plane dust-to-gas ratio
-%to be a power-law in radius,  
-%\begin{align}
-%  \epsilon_0(r) = \epsilon_{00}\left(\frac{r}{r_0}\right)^{-d},  
-%\end{align}
-%where $\epsilon_{00}$ is the dust-to-gas ratio at the fiducial radius
-%$r_0$.  
-Inserting Eq. \ref{dust_gauss} into vertical hydrostatic equilibrium,
-Eq. \ref{steady_momz} and integrating with the approximate
-gravitational potential (Eq. \ref{thin_disk_potential}) we obtain the
-gas density as
-\begin{align}
-  &\rhog(r,z)= \notag\\
-&\rho_\mathrm{g0}(r)\exp{\left\{ - \frac{z^2}{2\Hgas^2}
-    -\epsilon_0\frac{\Htilde^2}{\Hgas^2}\left[1 -
-      \exp{\left(-\frac{z^2}{2\Htilde^2}\right)}\right] \right\}}, 
-\end{align}
-where
-\begin{align}
-  \Hgas = \frac{c_s}{\OmK}, \quad \OmK \equiv \sqrt{\frac{GM_*}{r^3}},   
-\end{align}
-is the gas scale-height in the dust-free limit and $\OmK$ is the
-Keplerian frequency, respectively.
-
-We usually consider gas-dominated disks with $\epsilon_0 < 1$.  
-Then $\rhog(r,z)$ is effectively Gaussian, as in the 
-dust-free case. The dust density is approximately 
-\begin{align}
-  \rhod \simeq \epsilon_0\rho_\mathrm{g0}(r) \exp
-        {\left(-\frac{z^2}{2H_\mathrm{d}^2}\right)}, 
-\end{align}
-with 
-\begin{align}
-  \frac{1}{H_\mathrm{d}^2} = \frac{1}{\Htilde^2} + \frac{1}{\Hgas^2}, 
-\end{align}
-and $H_\mathrm{d}$ is the dust-scale height. In numerical
-calculations, we  we specify $H_\mathrm{d}< \Hgas$ to obtain 
-$\Htilde$ for input.
-
-Finally, we define 
-\begin{align}
-  Z \equiv \epsilon_0\frac{\Hd}{\Hg} \simeq
-  \frac{\Sigma_\mathrm{d}}{\Sigma_\mathrm{g}} 
-\end{align}
-as a measure of the local metalicity, where $\Sigma_\mathrm{d}$ and
-$\Sigma_\mathrm{g}$ are the dust and gas surface densities,
-respectively. The second equality holds for $\epsilon_0\ll1$.
-
-%\Sigma_\mathrm{d}$ and the gas
-%surface density $\Sigma_\mathrm{g}$.
-
-\subsection{Orbital frequency} 
-In the thin-disk approximation, the disk orbital frequency is 
-\begin{align}
-  \Omega(r,z) = \OmK(r)\left[1 - \frac{3}{2}\frac{z^2}{r^2} +
-    \frac{h_\mathrm{g}^2}{\left(1+\epsilon\right)}\frac{\p}{\p\ln{r}}\ln{\left(c_s^2\rhog\right)}
-    \right]^{1/2}, 
-\end{align}
-where 
-\begin{align}
-  h_\mathrm{g} \equiv \frac{\Hgas}{r}
-\end{align}
-is the characteristic disk aspect-ratio.
-
-\subsection{Vertical shear}\label{vertshear}
-The mixture possess vertical shear. To see this, we eliminate $\Phi$
-between Eq. \ref{steady_momr}---\ref{steady_momz} to 
-obtain 
-\begin{align}\label{vshear}
-  r\frac{\p \Omega^2}{\p z} 
-%&= \frac{\p\ln{\rho}}{\p r}\frac{\p}{\p
-%    z}\left[c_s^2(1-\tepsilon)\right] - \frac{\p\ln{\rho}}{\p z}
-%  \frac{\p}{\p r} \left[c_s^2(1-\tepsilon)\right]\\  
-   = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p \seff}{\p z} -\frac{\p
-    P}{\p z}\frac{\p \seff}{\p r} \right). 
-\end{align}
-Writing Eq. \ref{vshear} in terms of the gas density and dust-to-gas
-ratio with a power-law temperature profile (Eq. \ref{power_temp}) gives 
-\begin{align}\label{vshear2}
-  &r\frac{\p \Omega^2}{\p z}  =
-  \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\left\{
-  \frac{\p\epsilon}{\p r}\frac{\p\ln{\rhog}}{\p z}
-%  -\frac{\p\epsilon}{\p z}\left[\frac{\p\ln{\rhog}}{\p r} + \frac{q}{r}\right]\right.\notag\\
- -\frac{\p\epsilon}{\p z}\frac{\p\ln{P}}{\p r}\right.\notag\\
-  &\phantom{ r\frac{\p \Omega^2}{\p z}  =
-    \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\left\{\right\} }
-  \left. -\frac{q}{r} \left(1+\epsilon\right)\frac{\p\ln{\rhog}}{\p z}
-  \right\}. 
-\end{align}
-The first two terms correspond to vertical shear caused by spatial
-variations in the dust-to-gas ratio. The third term
-proportional to $q$ corresponds to vertical shear due to the 
-radial temperature gradient, and survives in the dust-free limit.
-
-We can compare these sources by 
-evaluating them using the equilibrium
-solutions in \S\ref{steady_state}. We assume the disk is radially
-smooth so that $\p_r\sim 1/r$. % and the dust-to-gas ratio
-% $\epsilon\ll1$. 
-This gives
-
-\begin{align}\label{vshear_split}
-  \frac{\left|r\p_z\Omega\right|_{\text{
-        dust/gas gradient}}}{\left|r\p_z\Omega\right|_{\text{
-        temp. gradient}}} \sim
- % \epsilon \frac{\mathrm{max}\left(\delta^2,
-  %  1\right)}{\left|q\right|\left(\epsilon + \delta^2\right)},
- \epsilon \frac{\mathrm{max}\left(\delta^2,
-    1\right)}{\left|q\right|\left(1+\epsilon\right)\delta^2},
-\end{align}
-where $\delta\equiv \Htilde/\Hgas$. 
-Since $|q|=O(1)$ in PPDs, Eq. \ref{vshear_split} indicates that
-vertical shear due to variations in the dust-to-gas ratio dominates 
-over that due to the radial temperature gradient for thin dust layers
-such that $\delta^2\ll \epsilon$. Otherwise, vertical shear is
-associated with the radial temperature gradient. %However, axisymmtric
-%\emph{instability} always caused by $\p_rc_s^2$.
-
-\subsection{Dusty vertical buoyancy}\label{vbuoyancy}
-Having identified the entropy of an isothermal dusty gas, we find the
-vertical buoyancy frequency $N_z$ of the mixture is given by 
-\begin{align}
-  N_z^2 \equiv - \frac{1}{\rho}\frac{\p P}{\p z}\frac{\p \seff}{\p z} &=
-  \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\frac{\p\ln\rhog}{\p 
-  z}\frac{\p\epsilon}{\p z} \\ &
-                                  =
-  \frac{\epsilon}{\left(1+\epsilon\right)^2}\left(\frac{z}{\Htilde}\right)^2\OmK^2\notag,  
-\end{align}
-where the second equality assumes $c_s=c_s(r)$ and the final equality
-applies to the equilibria in \S\ref{eqm}. Thus,  
-\begin{align*}
-N_z\lesssim
-O\left(\sqrt{\epsilon_0}\OmK\right). 
-\end{align*}
-However, for well-mixed dust layers such that $H_\epsilon \gg \Hg$, 
-$\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.
-
-The physical origin for this vertical buoyancy, which exists even in vertically isothermal
-disks, is that coupling gas to dust particles increases
-the fluid's inertia, but pressure (i.e. restoring) forces are
-unchanged. The increased weight of the fluid resists vertical
-oscillations and hence there is an associated buoyancy force. However,
-if $\tstop\neq0$ so the gas-dust coupling is imperfect, then gas is no
-longer `weighed down' by the dust, and can slip past it. Thus finite
-drag diminishes the dust-induced buoyancy. This is similar to reducing
-gas buoyancy through cooling \citep{lin15}.
+occur on much longer timescales than the instabilities of interest.
 
@@
 
-\section{Limiting behaviours}\label{limits}
+%\section{Limiting behaviours}\label{limits}
 
-In this section we discuss the basic properties of the 
-dusty-gas mixture using Eqs.\ref{masseq}--\ref{momeq} with 
-Eq. \ref{eff_energy} in place of the dust continuity
-equation. Self-gravity is neglected here. Our discussion is based on
-the corresponding axisymmetric linearized equations. A variable $X$ is
-subject to Eulerian perturbations of the form  
+\subsection{Integral relation}
+
+%Our discussion is based on
+%the corresponding axisymmetric linearized equations. 
+We consider axisymmetric Eulerian perturbations to a variable $X$ of
+the form  
 \begin{align}
@@
 
-In Appendix
-\ref{var_prin} we derive the integral relation,  
+In Appendix \ref{var_prin} we derive the integral relation from the
+corresponding linearized equations (Eqs.\ref{masseq}--\ref{momeq}, and 
+Eq. \ref{eff_energy} in place of the dust continuity equation): 
 \begin{align}
@@
 expression for said criteria \citep{tassoul78}. 
-
 We expect Eq. \ref{dusty_solberg1} to be satisfied in typical
@@
 direction as the local pressure gradient, which is equivalent to the
-gas density gradient for isothermal gas. Thus for 
-instability we need to violate Eq. \ref{dusty_solberg2}. We explore
-this below.
+gas density gradient for isothermal gas.
 
-\subsubsection{Vertically well-mixed dust}\label{vert_mixed_theory}
-In the special case where the dust-to-gas ratio is
-uniform in $z$ the left-hand-side of Eq. \ref{dusty_solberg2} becomes
+On the other hand, Eq. \ref{dusty_solberg2} can be violated in
+disks if the dust is vertically well-mixed ($\p_z\epsilon =
+0$) but  radially-varying ($\p_r\epsilon\neq0$). In this case the
+left-hand-side of Eq. \ref{dusty_solberg2} becomes
 \begin{align}
@@
 \end{align}
-Thus the disk is \emph{unstable}, because there is no effective vertical 
+Such an isothermal dusty disk is \emph{unstable} because there is no effective vertical 
 buoyancy to stabilize vertical motions ($N_z=0$). This situation is 
@@
 example in the dusty context in \S\ref{vert_mixed}.
-
-\subsubsection{Application to protoplanetary disks} 
-%In order to apply the stability criteria
-%Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} to typical
-%protoplanetary disks 
-We now consider the specific equilibria described in 
-\S\ref{eqm}. We assume the disk is approximately Keplerian so that
-$\kappa\simeq\OmK$, and that the vertical gas distribution is   
-Gaussian. In terms of the dust-to-gas
-ratio $\epsilon$, Eq. \ref{dusty_solberg2}
-becomes 
-\begin{align}
-  % &1 + \frac{\epsilon}{\left(1+
-  %   \epsilon\right)^2}\left(h_\mathrm{g}^2\frac{\p\ln{\rhog}}{\p\ln{r}}
-  % \frac{\p\ln{\epsilon}}{\p\ln{r}} + 
-  % \frac{z^2}{\Htilde^2}\right)>0 \label{stability_est1},   \\ 
-&1 - \frac{\epsilon
-  h_\mathrm{g}^2}{\left(1+\epsilon\right)^2}
-  \frac{\p\ln{\epsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\Htilde^2}{H_\mathrm{g}^2}\frac{\p\ln{\epsilon}}{\p\ln{r}}\right)
-  > 0 \label{stability_est2},
-\end{align}
-for \emph{stability}. The second term is usually stabilizing in 
-PPDs for reasons given in \S\ref{iso_perfect}. The third term is
-always destabilizing, but is small in smooth, thin disks with radial
-gradients $O(1/r)$, $h_\mathrm{g}\ll 1$, 
-provided that $H_\epsilon/\Hg$ is not large (e.g. some dust settling
-has occured). This means that typical dusty PPDs are stable to
-axisymmetric perturbations, a conclusion valid for arbitrarily thin
-dust layers.
-
-%%%%%%%%%%%%%%%%%%
-
-To violate Eq. \ref{stability_est2} and obtain instability, notice the 
-LHS is a quadratic in $\p_r\epsilon$. Thus instability is possible for
-sufficiently large (in magnitude) radial gradients  in the dust-to-gas
-ratio,  
-\begin{align}
-  \frac{\p\ln{\epsilon}}{\p\ln{r}} > \chi_+ \quad \mathrm{or} \quad 
-  \frac{\p\ln{\epsilon}}{\p\ln{r}} < \chi_-,
-\end{align}
-for \emph{instability}, where
-\begin{align}\label{spm}
-\chi_\pm = \frac{1}{2}\frac{H_\mathrm{g}^2}{\Htilde^2} 
-  \left[
-  \frac{\p\ln{\rhog}}{\p\ln{r}} \pm 
-  \sqrt{
-  \left(\frac{\p\ln{\rhog}}{\p\ln{r}}\right)^2 + 
-  4 \frac{\Htilde^2}{H_\mathrm{g}^2}
-  \frac{\left(1+\epsilon\right)^2}{\epsilon h_\mathrm{g}^2}
-  }
-  \,\right]. 
-\end{align} 
-In typical accretion disks where $\p_r\rhog<0$, instability is easier
-for increasing dust-to-gas ratios ($\p_r\epsilon > 0$).
-
-We can neglect density gradients in Eq. \ref{spm} 
-if $r\p_r\ln{\rhog}\sim O(1)$ and $H_\epsilon/\Hgas\gg
-\sqrt{\epsilon}\hgas$. For example, if $\epsilon\simeq 0.01$ and
-$\hgas\simeq 0.05$, then we require $H_\epsilon/\Hg\gg
-5\times10^{-3}$. (So the dust layer thickness $\Hd\gg 5
-\times10^{-3}\Hgas$ as well.) Then 
-instability requires 
-\begin{align}\label{dust_edge}
-\left|\frac{\p\ln{\epsilon}}{\p r}\right| \gtrsim
-  \frac{1}{\Htilde}\frac{\left(1+\epsilon\right)}{\sqrt{\epsilon}}.
-%\simeq
- % \frac{1}{\sqrt{\epsilon}H_\epsilon}, 
-\end{align}
-%where the second equality applies to $\epsilon\ll 1$. 
-That is, if the radial lengthscale of the dust-to-gas ratio,
-$L_\epsilon\ll O(\Htilde)$, i.e. its vertical lengthscale,  
-then the system is potentially unstable.  
-Taking $\epsilon\sim 0.01$,  we find that 
-for thin dust layers with $H_\epsilon \simeq \Hd\ll \Hg$,  instability
-requires $L_\epsilon\ll O(0.1\Hg)$, i.e. the dust-to-gas ratio must 
-vary on an extremely short lengthscale. This might be achieved, for
-example, at sharp edges associated with gaps opened by giant planets. 
-{\bf mention this is equivalent to sharp entropy gradients?}
 

### 2017-07-01 paper.tex
deleted original section two; remixed original section 5 and 6 (now 4 and 5) so the stability discussion goes from general to special, from not needing explicit background to explicit dust distributions

--- 2017-06-30 paper.tex
@@
 Kelvin-Helmholtz instabilities
-\citep{goldreich73, chiang08,barranco09,lee10}. These instabilities 
+\citep{goldreich1973, chiang08,barranco09,lee10}. These instabilities 
 can trigger turbulence, and (in different regimes) promote or inhibit 
@@
 Lagragian dust particles
-\citep{johansen06, nelson10,bai10,yang14,zhu14,gibbons15,simon16,baruteau16}. 
+\citep{johansen2006, nelson10,bai10,yang14,zhu14,gibbons15,simon16,baruteau16}. 
 The equation of motion for each particle is solved 
@@
 Models with two fluids or with Lagrangian solids are numerically difficult to evaluate in this regime, due to stiff drag forces.  
-In the reformulated equations, however, the center of mass motion does not experience drag forces and thus is not stiff.
-Moreover the relative motion satisfies the terminal-velocity approximation and can be eliminated from the equations.
-This single fluid framework for a well-coupled dust-gas mixture facilitates numerical calculations \citep{price15}.
-and also simplifies analytic calculations \citep{youdin05a,jacquet11}, which is useful for gaining insight.
+In the reformulated equations, however, the center of mass motion does
+not experience drag forces and thus is not stiff.  Moreover,
+for $\tstop\Omega\ll 1$ the relative motion satisfies the terminal-velocity approximation and can be eliminated from the equations.
+This single fluid framework for a well-coupled dust-gas mixture facilitates numerical calculations \citep{price15}; and also simplifies analytic calculations \citep{youdin05a,jacquet11}, which is useful for gaining insight.
 
@@
 perfectly-coupled dust are generally 
-stable to axisymmetric perturbations.%, regardless of how  thin the dust
+stable to axisymmetric perturbations, unless the 
+dust-to-gas ratio varies more rapidly in radius than in height.  
+%, regardless of how  thin the dust
 %layer is. 
-{\bf should we mention vertically well mixed case here?} 
+%{\bf should we mention vertically well mixed case here?} 
 We put forward a thermodynamical interpretation of 
@@
 
-\section{Dust-free modeling of dusty protoplanetary disks}\label{key_idea}
-{\bf is this section needed?} 
-The key concept we introduce in this work is the description 
-of dusty protoplanetary disks \emph{without} explicit dust 
-equations. This is best illustrated by writing down our model   
-equations for a mixture of isothermal gas and small 
-dust particles: 
-\begin{align}  
-  &\frac{\p \rho}{\p t} + \nabla\cdot\left(\rho\bm{v}\right)= 0, \label{mass0}\\ 
-  &\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} = -
-  \frac{1}{\rho}\nabla  P - \nabla \Phi_*,\label{mom0}\\ 
- & \frac{\p P}{\p t} + \bm{v}\cdot\nabla P 
-  = - P \nabla\cdot\bm{v} - \nabla\cdot\bm{\mathcal{F}}
-%                + \mathcal{C}(\rho, P; \tstop)
-\label{en0}, 
-\end{align}
-where $\rho=\rhod + \rhog$ is the total density, $P$ is the pressure,
-$\Phi_*$ is the stellar gravitational potential,   
-and $\bm{\mathcal{F}}=~\bm{\mathcal{F}}(\rho, P; \tstop)$ model effects of dust-gas drag, explained below. 
-%$\mathcal{C}$ represent effects of dust-gas drag, which
-%is paremterized by the stopping time $\tstop$ ($\mathcal{C}=0$ for
-%$\tstop=0$). 
-Similar equations for dusty-gas are given by
-\cite{price15}.
+% \section{Dust-free modeling of dusty protoplanetary disks}\label{key_idea}
+% {\bf is this section needed?} 
+% The key concept we introduce in this work is the description 
+% of dusty protoplanetary disks \emph{without} explicit dust 
+% equations. This is best illustrated by writing down our model   
+% equations for a mixture of isothermal gas and small 
+% dust particles: 
+% \begin{align}  
+%   &\frac{\p \rho}{\p t} + \nabla\cdot\left(\rho\bm{v}\right)= 0, \label{mass0}\\ 
+%   &\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} = -
+%   \frac{1}{\rho}\nabla  P - \nabla \Phi_*,\label{mom0}\\ 
+%  & \frac{\p P}{\p t} + \bm{v}\cdot\nabla P 
+%   = - P \nabla\cdot\bm{v} - \nabla\cdot\bm{\mathcal{F}}
+% %                + \mathcal{C}(\rho, P; \tstop)
+% \label{en0}, 
+% \end{align}
+% where $\rho=\rhod + \rhog$ is the total density, $P$ is the pressure,
+% $\Phi_*$ is the stellar gravitational potential,   
+% and $\bm{\mathcal{F}}=~\bm{\mathcal{F}}(\rho, P; \tstop)$ model effects of dust-gas drag, explained below. 
+% %$\mathcal{C}$ represent effects of dust-gas drag, which
+% %is paremterized by the stopping time $\tstop$ ($\mathcal{C}=0$ for
+% %$\tstop=0$). 
+% Similar equations for dusty-gas are given by
+% \cite{price15}.
 
-Eq. \ref{mass0}---\ref{en0} appear identical to pure gas
-dynamics, but they physically represents a two-phase, 
-dusty gas. The mixture obeys mass and momentum conservation as usual,
-but also an energy-like equation, even though the system is
-isothermal by assumption. The term $-\bm{\mathcal{F}}\propto \tstop
-\nabla P$ is not a true heat flux. Instead, it reflects 
-particle flux towards pressure maximum due to dust-gas drag 
-\citep{weidenschilling77}. The back-reaction onto the gas then changes
-the pressure distribution.
+% Eq. \ref{mass0}---\ref{en0} appear identical to pure gas
+% dynamics, but they physically represents a two-phase, 
+% dusty gas. The mixture obeys mass and momentum conservation as usual,
+% but also an energy-like equation, even though the system is
+% isothermal by assumption. The term $-\bm{\mathcal{F}}\propto \tstop
+% \nabla P$ is not a true heat flux. Instead, it reflects 
+% particle flux towards pressure maximum due to dust-gas drag 
+% \citep{weidenschilling77}. The back-reaction onto the gas then changes
+% the pressure distribution.
 
-%weidenschilling7
+% %weidenschilling7
 
-Since the above equations are similar to ideal gas dynamics, it 
-means that many of the results derived for ideal hydrodynamics can be  
-generalized to dusty disks without additional effort. This includes 
-the important consequence of phase lags between density and pressure,
-discussed next.
+% Since the above equations are similar to ideal gas dynamics, it 
+% means that many of the results derived for ideal hydrodynamics can be  
+% generalized to dusty disks without additional effort. This includes 
+% the important consequence of phase lags between density and pressure,
+% discussed next.
 

### 2017-07-01 paper.tex
tweaked intro to fit restructure

@@
 We begin our formal discussion in \S\ref{setup} by transforming the
-basic equations describing tightly-coupled dusty gas into  hydrodynamic equations with a special cooling function. 
-We derive equilibrium solutions in \S\ref{eqm}, which we use in our explicit stability calculations. In \S\ref{limits} we analyze the 
-general stability properties of dusty disks. We then move onto
-explicit stability analyses. In  
-\S\ref{linear_problem} we show that our linearized equations 
-reproduce the SGI and SI in the tight coupling regime. In \S\ref{results} we
+basic equations describing tightly-coupled dusty gas into
+hydrodynamic equations with a special cooling function.  Here we also
+formalize the concept of dust-induced entropy and buoyancy. 
+%We derive equilibrium solutions in \S\ref{eqm}, which we use in our
+%explicit stability calculations.
+In \S\ref{limits} we analyze the general stability properties of dusty
+disks. We then move onto explicit stability analyses in
+\S\ref{linear_problem}, were we show that dusty edges are unstable,
+and revisit the SGI and SI in the tight
+coupling regime. In \S\ref{results} we
 extend the VSI to dusty disks. We   

### 2017-07-01 paper.tex
tweaked intro to fit restructure

@@
 %Taipei 10617, Taiwan}
-\altaffiltext{1}{Department of Astronomy and Steward Observatory, 
+\altaffiltext{1}{Institute of Astronomy and Astrophysics, Academia Sinica, 
+Taipei 10617, Taiwan
+}
+\altaffiltext{2}{Department of Astronomy and Steward Observatory, 
 University of
 Arizona, Tucson, AZ 85721, USA}
-\altaffiltext{2}{Steward Theory Fellow}
-\altaffiltext{3}{Institute of Astronomy and Astrophysics, Academia Sinica, 
-Taipei 10617, Taiwan
-}
+\altaffiltext{3}{Steward Theory Fellow}
 \email{mklin@asiaa.sinica.edu.tw}

### 2017-07-01 result.tex

@@
 gradients in the dust-to-gas ratio (as discussed in
-\S\ref{vert_mixed_theory}).
+\S\ref{iso_perfect}).
 
@@
 
-In \S\ref{vert_mixed_theory} we found that for strictly isothermal
+In \S\ref{iso_perfect} we found that for strictly isothermal
 disks, a vertically-uniform dust-to-gas ratio can be unstable
@@
 radial wavenumber $k_x\Hg=1800$. As expected from the discussion in
-\S\ref{vert_mixed_theory}, the disk admits purely growing modes with
+\S\ref{iso_perfect}, the disk admits purely growing modes with
 $\omega=0$. This is distinct from the growing oscillations associated

### 2017-07-02 linear.tex

--- 2017-06-12 linear.tex
@@
-\section{Linearized equations}\label{linear_problem}
-To compute explicit solutions to the linear problem, 
-we consider radially-localized axisymmetric disturbances of the form  
+\section{Application to protoplanetary disks}\label{linear_problem}
+We now examine instabilities in dusty protoplanetary disks
+based on explicit descriptions of the equilibrium state. We first show
+that sharp radial edges in the dust component are unstable. We then
+revisit two well-known instabilities in dusty gas using the
+hydrodynamic framework developed thusfar, namly secular graviational
+instabilities and the streaming instability.
+
+\subsection{Disk structure with a prescribed dust distribution}\label{steady_state}  
+We assume a Gaussian profile in the
+dust-to-gas ratio,    
+\begin{align}\label{dust_gauss}
+  \epsilon(r,z) = \epsilon_0(r)
+  \exp{\left[-\frac{z^2}{2\Htilde^2(r)}\right]}. 
+\end{align}
+%For simplicity we assume the dust-to-gas ratio at the mid-plane
+%$\epsilon_0$, as well as its characteristic scale-height
+%$\widetilde{H}$, are both constant.
+% We take the mid-plane dust-to-gas ratio
+%to be a power-law in radius,  
+%\begin{align}
+%  \epsilon_0(r) = \epsilon_{00}\left(\frac{r}{r_0}\right)^{-d},  
+%\end{align}
+%where $\epsilon_{00}$ is the dust-to-gas ratio at the fiducial radius
+%$r_0$.  
+Inserting Eq. \ref{dust_gauss} into vertical hydrostatic equilibrium,
+Eq. \ref{steady_momz} and integrating with the approximate
+gravitational potential (Eq. \ref{thin_disk_potential}) we obtain the
+gas density as
+\begin{align}
+  &\rhog(r,z)= \notag\\
+&\rho_\mathrm{g0}(r)\exp{\left\{ - \frac{z^2}{2\Hgas^2}
+    -\epsilon_0\frac{\Htilde^2}{\Hgas^2}\left[1 -
+      \exp{\left(-\frac{z^2}{2\Htilde^2}\right)}\right] \right\}}, 
+\end{align}
+where
+\begin{align}
+  \Hgas = \frac{c_s}{\OmK}, \quad \OmK \equiv \sqrt{\frac{GM_*}{r^3}},   
+\end{align}
+is the gas scale-height in the dust-free limit and $\OmK$ is the
+Keplerian frequency, respectively.
+
+We usually consider gas-dominated disks with $\epsilon_0 < 1$.  
+Then $\rhog(r,z)$ is effectively Gaussian, as in the 
+dust-free case. The dust density is approximately 
+\begin{align}
+  \rhod \simeq \epsilon_0\rho_\mathrm{g0}(r) \exp
+        {\left(-\frac{z^2}{2H_\mathrm{d}^2}\right)}, 
+\end{align}
+with 
+\begin{align}
+  \frac{1}{H_\mathrm{d}^2} = \frac{1}{\Htilde^2} + \frac{1}{\Hgas^2}, 
+\end{align}
+and $H_\mathrm{d}$ is the dust-scale height. In numerical
+calculations, we  we specify $H_\mathrm{d}< \Hgas$ to obtain 
+$\Htilde$ for input.
+
+Finally, we define 
+\begin{align}
+  Z \equiv \epsilon_0\frac{\Hd}{\Hg} \simeq
+  \frac{\Sigma_\mathrm{d}}{\Sigma_\mathrm{g}} 
+\end{align}
+as a measure of the local metalicity, where $\Sigma_\mathrm{d}$ and
+$\Sigma_\mathrm{g}$ are the dust and gas surface densities,
+respectively. The second equality holds for $\epsilon_0\ll1$.
+
+%\Sigma_\mathrm{d}$ and the gas
+%surface density $\Sigma_\mathrm{g}$.
+
+\subsubsection{Orbital frequency} 
+In the thin-disk approximation, the disk orbital frequency is 
+\begin{align}
+  \Omega(r,z) = \OmK(r)\left[1 - \frac{3}{2}\frac{z^2}{r^2} +
+    \frac{h_\mathrm{g}^2}{\left(1+\epsilon\right)}\frac{\p}{\p\ln{r}}\ln{\left(c_s^2\rhog\right)}
+    \right]^{1/2}, 
+\end{align}
+where 
+\begin{align}
+  h_\mathrm{g} \equiv \frac{\Hgas}{r}
+\end{align}
+is the characteristic disk aspect-ratio.
+
+\subsubsection{Vertical shear}\label{vertshear}
+The mixture possess vertical shear. To see this, we eliminate $\Phi$
+between Eq. \ref{steady_momr}---\ref{steady_momz} to 
+obtain 
+\begin{align}\label{vshear}
+  r\frac{\p \Omega^2}{\p z} 
+%&= \frac{\p\ln{\rho}}{\p r}\frac{\p}{\p
+%    z}\left[c_s^2(1-\tepsilon)\right] - \frac{\p\ln{\rho}}{\p z}
+%  \frac{\p}{\p r} \left[c_s^2(1-\tepsilon)\right]\\  
+   = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p \seff}{\p z} -\frac{\p
+    P}{\p z}\frac{\p \seff}{\p r} \right). 
+\end{align}
+Writing Eq. \ref{vshear} in terms of the gas density and dust-to-gas
+ratio with a power-law temperature profile (Eq. \ref{power_temp}) gives 
+\begin{align}\label{vshear2}
+  &r\frac{\p \Omega^2}{\p z}  =
+  \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\left\{
+  \frac{\p\epsilon}{\p r}\frac{\p\ln{\rhog}}{\p z}
+%  -\frac{\p\epsilon}{\p z}\left[\frac{\p\ln{\rhog}}{\p r} + \frac{q}{r}\right]\right.\notag\\
+ -\frac{\p\epsilon}{\p z}\frac{\p\ln{P}}{\p r}\right.\notag\\
+  &\phantom{ r\frac{\p \Omega^2}{\p z}  =
+    \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\left\{\right\} }
+  \left. -\frac{q}{r} \left(1+\epsilon\right)\frac{\p\ln{\rhog}}{\p z}
+  \right\}. 
+\end{align}
+The first two terms correspond to vertical shear caused by spatial
+variations in the dust-to-gas ratio. The third term
+proportional to $q$ corresponds to vertical shear due to the 
+radial temperature gradient, and survives in the dust-free limit.
+
+We can compare these sources by 
+evaluating them using the equilibrium
+solutions in \S\ref{steady_state}. We assume the disk is radially
+smooth so that $\p_r\sim 1/r$. % and the dust-to-gas ratio
+% $\epsilon\ll1$. 
+This gives
+
+\begin{align}\label{vshear_split}
+  \frac{\left|r\p_z\Omega\right|_{\text{
+        dust/gas gradient}}}{\left|r\p_z\Omega\right|_{\text{
+        temp. gradient}}} \sim
+ % \epsilon \frac{\mathrm{max}\left(\delta^2,
+  %  1\right)}{\left|q\right|\left(\epsilon + \delta^2\right)},
+ \epsilon \frac{\mathrm{max}\left(\delta^2,
+    1\right)}{\left|q\right|\left(1+\epsilon\right)\delta^2},
+\end{align}
+where $\delta\equiv \Htilde/\Hgas$. 
+Since $|q|=O(1)$ in PPDs, Eq. \ref{vshear_split} indicates that
+vertical shear due to variations in the dust-to-gas ratio dominates 
+over that due to the radial temperature gradient for thin dust layers
+such that $\delta^2\ll \epsilon$. Otherwise, vertical shear is
+associated with the radial temperature gradient. %However, axisymmtric
+%\emph{instability} always caused by $\p_rc_s^2$.
+
+\subsubsection{Dusty vertical buoyancy}\label{vbuoyancy}
+For the above equilibria the vertical buoyancy frequency is given
+explicitly as 
+\begin{align}
+  N_z^2 =
+  %\frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\frac{\p\ln\rhog}{\p 
+%  z}\frac{\p\epsilon}{\p z} \\ &
+  %                                =
+ \frac{\epsilon}{\left(1+\epsilon\right)^2}\left(\frac{z}{\Htilde}\right)^2\OmK^2\notag,  
+\end{align}
+where we have used $c_s=c_s(r)$. Then 
+\begin{align*}
+N_z\lesssim
+O\left(\sqrt{\epsilon_0}\OmK\right). 
+\end{align*}
+However, for well-mixed dust layers such that $H_\epsilon \gg \Hg$, 
+$\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.
+
+\subsection{Instability of dusty edges} 
+Here we apply the dusty analog of the Solberg-Hoiland criteria derived
+in \S\ref{iso_perfect}. As discussed there, the first criterion in
+generally satisifed. Thus we only consider the second condition,
+Eq. \ref{dusty_solberg2}.
+
+We assume the disk is approximately Keplerian so that
+$\kappa\simeq\OmK$, and that the vertical gas distribution is   
+Gaussian. In terms of the dust-to-gas
+ratio $\epsilon$, Eq. \ref{dusty_solberg2}
+becomes 
+\begin{align}
+  % &1 + \frac{\epsilon}{\left(1+
+  %   \epsilon\right)^2}\left(h_\mathrm{g}^2\frac{\p\ln{\rhog}}{\p\ln{r}}
+  % \frac{\p\ln{\epsilon}}{\p\ln{r}} + 
+  % \frac{z^2}{\Htilde^2}\right)>0 \label{stability_est1},   \\ 
+&1 - \frac{\epsilon
+  h_\mathrm{g}^2}{\left(1+\epsilon\right)^2}
+  \frac{\p\ln{\epsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\Htilde^2}{H_\mathrm{g}^2}\frac{\p\ln{\epsilon}}{\p\ln{r}}\right)
+  > 0 \label{stability_est2},
+\end{align}
+for \emph{stability}. The second term is usually stabilizing in 
+PPDs for reasons given in \S\ref{iso_perfect}. The third term is
+always destabilizing, but is small in smooth, thin disks with radial
+gradients $O(1/r)$, $h_\mathrm{g}\ll 1$, 
+provided that $H_\epsilon/\Hg$ is not large (e.g. some dust settling
+has occured). This means that typical dusty PPDs are stable to
+axisymmetric perturbations, a conclusion valid for arbitrarily thin
+dust layers.
+
+%%%%%%%%%%%%%%%%%%
+
+To violate Eq. \ref{stability_est2} and obtain instability, notice the 
+LHS is a quadratic in $\p_r\epsilon$. Thus instability is possible for
+sufficiently large (in magnitude) radial gradients  in the dust-to-gas
+ratio,  
+\begin{align}
+  \frac{\p\ln{\epsilon}}{\p\ln{r}} > \chi_+ \quad \mathrm{or} \quad 
+  \frac{\p\ln{\epsilon}}{\p\ln{r}} < \chi_-,
+\end{align}
+for \emph{instability}, where
+\begin{align}\label{spm}
+\chi_\pm = \frac{1}{2}\frac{H_\mathrm{g}^2}{\Htilde^2} 
+  \left[
+  \frac{\p\ln{\rhog}}{\p\ln{r}} \pm 
+  \sqrt{
+  \left(\frac{\p\ln{\rhog}}{\p\ln{r}}\right)^2 + 
+  4 \frac{\Htilde^2}{H_\mathrm{g}^2}
+  \frac{\left(1+\epsilon\right)^2}{\epsilon h_\mathrm{g}^2}
+  }
+  \,\right]. 
+\end{align} 
+In typical accretion disks where $\p_r\rhog<0$, instability is easier
+for increasing dust-to-gas ratios ($\p_r\epsilon > 0$).
+
+We can neglect density gradients in Eq. \ref{spm} 
+if $r\p_r\ln{\rhog}\sim O(1)$ and $H_\epsilon/\Hgas\gg
+\sqrt{\epsilon}\hgas$. For example, if $\epsilon\simeq 0.01$ and
+$\hgas\simeq 0.05$, then we require $H_\epsilon/\Hg\gg
+5\times10^{-3}$. (So the dust layer thickness $\Hd\gg 5
+\times10^{-3}\Hgas$ as well.) Then 
+instability requires 
+\begin{align}\label{dust_edge}
+\left|\frac{\p\ln{\epsilon}}{\p r}\right| \gtrsim
+  \frac{1}{\Htilde}\frac{\left(1+\epsilon\right)}{\sqrt{\epsilon}}.
+%\simeq
+ % \frac{1}{\sqrt{\epsilon}H_\epsilon}, 
+\end{align}
+%where the second equality applies to $\epsilon\ll 1$. 
+That is, if the radial lengthscale of the dust-to-gas ratio 
+is much less its vertical lengthscale, $L_\epsilon\ll O(\Htilde)$, 
+then the system is  unstable.  
+Taking $\epsilon\sim 0.01$,  we find that 
+for thin dust layers with $H_\epsilon \simeq \Hd\ll \Hg$,  instability
+requires $L_\epsilon\ll O(0.1\Hg)$, i.e. the dust-to-gas ratio must 
+vary on an extremely short lengthscale. This might be achieved, for
+example, at sharp edges associated with gaps opened by giant planets.
+
+Technically, the above discussion is only applicable to strictly
+isothermal disks perfectly coupled to dust (see \S\ref{iso_perfect}). 
+However, since dusty edges translate to sharp entropy gradients (see
+\S\ref{dusty_entropy}), we may generally expect sharp features in the
+dust distribution of protoplanetary disks to be unstable.
+
+%{\bf mention this is equivalent to sharp entropy gradients?}
+
+\subsection{Radially localized linear problem}
+We now specialize further and compute explicit solutions to the linear
+problem. We consider radially-localized axisymmetric disturbances of the form  
 \begin{align}
@@
 where $^\prime \equiv \p_z$ and recall $\bm{F} \equiv -\nabla
-P/\rho$. Here we restore self-gravity temporarily for SGI in the next
+P/\rho$. We have temporarily restored self-gravity  to
+discuss SGI in the next
 section. The linearized dust-duffusion function 
@@
 
-\subsection{Secular gravitational instability}\label{sgi}
+\subsubsection{Secular gravitational instability}\label{sgi}
 We first consider a razor-thin, self-gravitating disk so that $\rho =
@@
 
-\subsection{Streaming instability}\label{si}
+\subsubsection{Streaming instability}\label{si}
 We now consider 3D disks without self-gravity. We neglect the vertical     
@@
 such a phase lag.
+
+In a future work we will perform a detailed comparison between the
+simplified one-fluid and full two-fluid frameworks in calculating SI.
+
+%"generalized stokes number" would include other problem parameters,
+%e.g. wavenumber
 

### 2017-07-12 paper.tex
added reference to latter 17 and squires 17 in intro, need to add later hopkins 17

@@
 %Taipei 10617, Taiwan}
-\altaffiltext{1}{Institute of Astronomy and Astrophysics, Academia Sinica, 
-Taipei 10617, Taiwan
-}
-\altaffiltext{2}{Department of Astronomy and Steward Observatory, 
-University of
-Arizona, Tucson, AZ 85721, USA}
+\altaffiltext{1}{Institute of Astronomy and Astrophysics, Academia Sinica, Taipei 10617, Taiwan}
+\altaffiltext{2}{Department of Astronomy and Steward Observatory, University of Arizona, Tucson, AZ 85721, USA}
 \altaffiltext{3}{Steward Theory Fellow}
+
 \email{mklin@asiaa.sinica.edu.tw}
@@
 \citep[SI,][]{youdin05a,youdin07b,johansen07}, secular gravitational
-instabilities \citep[SGI,][]{ward00,youdin11,michikoshi12,takahashi14} and
+instabilities \citep[SGI,][]{ward00,youdin11,michikoshi12,takahashi14,latter17} and 
 Kelvin-Helmholtz instabilities
-\citep{goldreich1973, chiang08,barranco09,lee10}. These instabilities 
-can trigger turbulence, and (in different regimes) promote or inhibit 
+\citep{goldreich1973, chiang08,barranco09,lee10}.
+
+These instabilities can trigger turbulence, and (in different regimes) promote or inhibit  
 planetesimal formation. 
@@
 dusty instabilities have appeared in numerical  
-simulations \citep{loren15,loren16,lamb16}, but their physical underpinnings are not yet well
+simulations \citep{loren15,loren16,lamb16} and analytical calculations \citep{squire17},   
+but their physical underpinnings are not yet well 
 understood.

### 2017-07-12 paper.tex
added ref to hopkins 17

@@
 dusty instabilities have appeared in numerical  
-simulations \citep{loren15,loren16,lamb16} and analytical calculations \citep{squire17},   
+simulations \citep{loren15,loren16,lamb16} and analytical calculations \citep{squire17,hopkins17},   
 but their physical underpinnings are not yet well 

### 2017-07-13 paper.tex
minor tweaks to intro which I'm happy with.  Remaining concern is whether we should clarify what general vs. explicit stability analyses are (i.e. for a specific disk model or not).

@@
 
-The advantage of this reformulation arises in the tight coupling limit of  $\tstop \ll 1/\Omega$, where the orbital frequency $\Omega$ sets the characteristic timescale in many disk dynamics problems.
+This reformulation is particularly advantageous in the tight coupling limit of  $\tstop \ll 1/\Omega$, where the orbital frequency $\Omega$ sets the characteristic timescale in many disk dynamics problems.
 Models with two fluids or with Lagrangian solids are numerically difficult to evaluate in this regime, due to stiff drag forces.  
@@
 This parallel is most precise when the gas in the dust/gas mixture obeys a locally isothermal equation of state for the gas, $P = c_s^2\rhog$, where $c_s$ is a prescribed sound-speed profile.
-Thus we make this approximation which holds when cooling times are short \citep{lin15}, as with protoplanetary disks
+This isothermal approximation holds when cooling times are short \citep{lin15}, e.g.\ in protoplanetary disks
  whose temperature is set by external irradiation
@@
 This paper is organized as follows. 
-%AY: or not? %We first present, without
-%derivation, the correspondence between dusty gas and     
-%ideal hydrodynamics in \S\ref{key_idea}. 
 In \S\ref{grow_osc} 
-we describe a key insight of our study, that pressure (i.e. ``PdV") work explains the basis of several instabilities, including SI and VSI, in terms of
+we describe a key insight of our study, that pressure (i.e.\ ``PdV") work explains the basis of several instabilities, including SI and VSI, in terms of
 pressure-density phase lags.    
-%This means that dusty-gas phenomena, such 
-%as SI and VSI, have thermodynamical interpretations.  
-We begin our formal discussion in \S\ref{setup} by transforming the
-basic equations describing tightly-coupled dusty gas into
+We develop our formalism in \S\ref{setup} by transforming the
+two fluid equations of a tightly-coupled dusty gas into single fluid
 hydrodynamic equations with a special cooling function.  Here we also
-formalize the concept of dust-induced entropy and buoyancy. 
+define the effective entropy and buoyancy of our model fluid. 
 %We derive equilibrium solutions in \S\ref{eqm}, which we use in our

### 2017-07-14 paper.tex
tiny intro edits (this paper is organized...), remove repeated text in SI section

@@
  dusty gas dynamics are equivalent to that of nearly adiabatic (i.e. slowly cooling) gas
-dynamics without dust (Fig. \ref{concept_chart}).  The basis of the analogy is that the entrainment 
+dynamics without dust, see Fig.~\ref{concept_chart}.  The basis of the analogy is that the entrainment 
 of tightly-coupled dust particles in gas flows is similar to the advection
@@
 %explicit stability calculations.
-In \S\ref{limits} we analyze the general stability properties of dusty
-disks. We then move onto explicit stability analyses in
+In \S\ref{limits} we discuss general stability properties of dusty
+disks. We then analyze explicit disk models 
 \S\ref{linear_problem}, were we show that dusty edges are unstable,

### 2017-07-14 linear.tex
tiny intro edits (this paper is organized...), remove repeated text in SI section
moved general expression of vertical shear up into general stability section

@@
 \subsubsection{Vertical shear}\label{vertshear}
-The mixture possess vertical shear. To see this, we eliminate $\Phi$
-between Eq. \ref{steady_momr}---\ref{steady_momz} to 
-obtain 
-\begin{align}\label{vshear}
-  r\frac{\p \Omega^2}{\p z} 
-%&= \frac{\p\ln{\rho}}{\p r}\frac{\p}{\p
-%    z}\left[c_s^2(1-\tepsilon)\right] - \frac{\p\ln{\rho}}{\p z}
-%  \frac{\p}{\p r} \left[c_s^2(1-\tepsilon)\right]\\  
-   = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p \seff}{\p z} -\frac{\p
-    P}{\p z}\frac{\p \seff}{\p r} \right). 
-\end{align}
 Writing Eq. \ref{vshear} in terms of the gas density and dust-to-gas
-ratio with a power-law temperature profile (Eq. \ref{power_temp}) gives 
+ratio with a power-law temperature profile (Eq. \ref{power_temp})
+gives the disk's explicit vertical shear profile:
 \begin{align}\label{vshear2}
@@
 
-In a future work we will perform a detailed comparison between the
-simplified one-fluid and full two-fluid frameworks in calculating SI.
-
 %"generalized stokes number" would include other problem parameters,

### 2017-07-14 model.tex
moved general expression of vertical shear up into general stability section

@@
 similar definition exists for the radial buoyancy frequency
-$N_r$. Typically in a thin, smooth disk we have $|N_r|\ll |N_z|$.
+$N_r$. However, in thin, smooth disks we have $|N_r|\ll |N_z|$.
 
@@
 
-However, disk structures obtained this way generally do not satisfy 
+The mixture possess vertical shear. To see this, we eliminate $\Phi$
+between Eq. \ref{steady_momr}---\ref{steady_momz} to 
+obtain 
+\begin{align}\label{vshear}
+  r\frac{\p \Omega^2}{\p z} 
+%&= \frac{\p\ln{\rho}}{\p r}\frac{\p}{\p
+%    z}\left[c_s^2(1-\tepsilon)\right] - \frac{\p\ln{\rho}}{\p z}
+%  \frac{\p}{\p r} \left[c_s^2(1-\tepsilon)\right]\\  
+   = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p \seff}{\p z} -\frac{\p
+    P}{\p z}\frac{\p \seff}{\p r} \right). 
+\end{align}
+Since $\seff$ depends on $\tepsilon$, we see that radial gradients in
+the dust-to-gas ratio also drives vertical shear \citep[cf. that 
+driven by $\p_z\epsilon$ usually discussed in dusty disks,  
+e.g.][]{chiang08}.
+
+However, disk structures obtained from
+Eq. \ref{steady_momr}---\ref{steady_momz} generally do not satisfy  
 the steady state effective energy Eq. \ref{eff_energy} because typically 
@@
 \end{align}
-Such an isothermal dusty disk is \emph{unstable} because there is no effective vertical 
-buoyancy to stabilize vertical motions ($N_z=0$). This situation is 
-identical to the pure gas, adiabtic simulations of \cite{nelson13}
+Such an isothermal dusty disk is \emph{unstable} but there is no
+effective vertical buoyancy to stabilize vertical motions ($N_z=0$)
+that can tap into the free energy associated with  
+vertical shear due to the radial gradient in the dust-to-gas ratio. 
+This situation is identical to the pure gas, adiabtic simulations of \cite{nelson13}
 where the gas entropy is vertically uniform but has a radial

### 2017-08-02 model.tex
revising according to AY's comments, up to 3.4

@@
 velocity field as $(\rhog,\bm{v}_\mathrm{g})$ and
-$(\rhod,\bm{v}_\mathrm{d})$, respectively. The mixture has a single 
-temperature $T$, and its pressure $P$ arise solely from the 
-gas component. %neglect turbulence, viscosity, self-g and b fields.
+$(\rhod,\bm{v}_\mathrm{d})$, respectively. The 
+mixture has total density \begin{align}
+  \rho \equiv \rhog + \rhod,
+\end{align}
+and center of mass velocity, 
+\begin{align}
+  \bm{v} \equiv \frac{\rhog\bm{v}_\mathrm{g} + 
+    \rhod\bm{v}_\mathrm{d}}{\rho}, 
+\end{align}
+a single temperature $T$, and its pressure $P$ arise solely from the  
+gas component. Our goal is to obtain a set of equations
+describing the dust-gas mixture that resembles standard, single-phase
+hydrodynamics. 
+%This is possible by considering
+%strongly coupled dust particles (but not neccessarily perfectly
+%coupled) and an isothermal gas equation of state.
 
-The two fluids interact via a drag force parameterized by the relative stopping 
+%\subsection{Terminal velocity approximation}
+
+The dust and gas fluids interact via a drag force parameterized by the relative stopping 
 time $\tstop$ such that 
@@
   \rhog\left.\frac{\p \bm{v}_\mathrm{g}}{\p t}\right|_\mathrm{drag}=
-  - \frac{\rhog\rhod}{\rho}\frac{\widetilde{\bm{v}}}{\tstop}. 
+  - \frac{\rhog\rhod}{\rho}\frac{\left(\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g}\right)}{\tstop}. 
 \end{align}
-is the dust-gas friction force per unit volume, where
-\begin{align}
-  \widetilde{\bm{v}} \equiv\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} 
-\end{align}
-is the dust-gas velocity difference; and 
-\begin{align}
-  \rho \equiv \rhog + \rhod
-\end{align}
-is the total density. 
+is the dust-gas friction force per unit volume.%, where
+%\begin{align}
+%  \widetilde{\bm{v}} \equiv\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} 
+%\end{align}
+%is the dust-gas velocity difference. 
  Note that $\tstop$ 
@@
 \tstop\rho/\rhog$ used in some studies
-\citep[e.g.][]{youdin05a,jacquet11}.
+\citep[e.g.][]{youdin05a}.
 
-In general $\widetilde{\bm{v}}$ obeys a complicated evolutionary equation
-\citep[see, e.g.][]{youdin05a}. However, for tightly-coupled dust
+In general the relative velocity $\bm{v}_\mathrm{d} -
+\bm{v}_\mathrm{g}$ obeys a complicated evolutionary equation 
+\citep[see, e.g.][]{youdin05a}.% obtained from taking the difference
+%between evolutionary equations for $\bm{v}_\mathrm{g}$ and
+%$\bm{v}_\mathrm{d}$. 
+ However, for tightly-coupled dust
 particles with $\tstop\OmK\ll 1 $, where $\OmK$ is the Keplerian orbital
@@
 \begin{align}
-  \widetilde{\bm{v}} = \frac{\nabla
+  \bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} = \frac{\nabla
     P}{\rhog}\tstop 
 \end{align}
-\citep{jacquet11}. This equation reflect the physical result that dust
-particles tend to drift towards pressure maxima.
+\citep{jacquet11}. This equation reflects the well-known effect of 
+particle drift towards pressure maxima \citep{weidenschilling77}.
 
-In the limit of strong dust-gas drag, $\tstop\to0$, the mixture obeys 
-the first-order (in $\tstop$) single-fluid equations: 
+%\subsection{One-fluid equations for dusty gas}
+
+Under the terminal velocity approximation the dust-gas mixture obeys   
+the first-order (in $\tstop$) one-fluid equations: 
 \begin{align} 
   &\frac{D\rho}{Dt} = -\rho\nabla\cdot\bm{v}, \label{masseq}\\ 
-  &\frac{D\bm{v}}{Dt} = - \nabla\Phi_\mathrm{tot} - \frac{1}{\rho}\nabla  P, \label{momeq}\\ 
    &\frac{D\tepsilon}{Dt} = -\frac{1}{\rho} \nabla \cdot \left(\tepsilon 
      \tstop \nabla P \right),\label{dusteq}\\
+  &\frac{D\bm{v}}{Dt} = - \nabla\Phi_\mathrm{tot} - \frac{1}{\rho}\nabla  P, \label{momeq}\\ 
   &\frac{D T}{D t} = - \left(\gamma-1\right)T\nabla\cdot\bm{v} +
-  \mathcal{H}_\mathrm{drag} + \mathcal{H}_\mathrm{eff}  - \Lambda  \label{tempeq} 
+  \mathcal{H} + \mathcal{H}_\mathrm{eff}  - \Lambda  \label{tempeq} 
 \end{align} 
-\citep{laibe14,price15}, where $D/Dt \equiv \p_t + \bm{v}\cdot\nabla$ is the Lagragian
-derivative \emph{following the mixture} with center-of-mass velocity  
-\begin{align}
-  \bm{v} \equiv \frac{\rhog\bm{v}_\mathrm{g} + 
-    \rhod\bm{v}_\mathrm{d}}{\rho}. 
-\end{align}
+\citep[see ][for a detailed derivation from the two-fluid
+equations]{laibe14,price15}, where $D/Dt \equiv \p_t +
+\bm{v}\cdot\nabla$ is the Lagragian 
+derivative \emph{following the mixture}'s center-of-mass velocity $\bm{v}$. 
 For an ideal gas the pressure is given 
@@
 is the dust-fraction and $\epsilon=\rhod/\rhog$ is the usual
-dust-to-gas ratio. Note that Eq. \ref{dusteq} is 
-equivalent to Eq. 46 of \cite{jacquet11}.
-
-For the mixture's temperature evolution, Eq. \ref{tempeq}, $\gamma$ is
-the adiabatic index; $\mathcal{H}_\mathrm{drag}$ represents heating
-due to dust-gas drag and $\mathcal{H}_\mathrm{eff}$ is an effective
-energy source arising from transforming from the two-fluid to
-one-fluid description \citep[see][ for details]{laibe14}. We also
-include a simple model of radiative cooling,
-\begin{align}
-  \Lambda = \frac{T -
-    T_\mathrm{ref}}{t_\mathrm{cool}}, \label{cooling} 
-\end{align}
-which relaxes the mixture back to a prescribed temperature profile   
-$T_\mathrm{ref}$ on a timescale of $\tcool$. We will shortly simplify
-the problem by considering rapid cooling, $\tcool\to0$.
+dust-to-gas ratio. If $\tstop=0$ then the 
+dust-to-gas ratio is conserved following the fluid. Otherwise, 
+$\epsilon$ evolves due to particle drift in response to pressure
+gradients. Note that Eq. \ref{dusteq} is equivalent to Eq. 46 of 
+\cite{jacquet11}.
 
@@
 $|z|\ll r$. We shall adopt this approximate potential in order to
-obtain explicit expressions for disk equilibria (\S\ref{eqm}).
-
+obtain explicit expressions for disk equilibria (\S\ref{steady_state}). 
 The disk's own gravitational potential $\psi$ is included for
@@
 
+For the mixture's temperature evolution, Eq. \ref{tempeq}, $\gamma$ is
+the adiabatic index; $\mathcal{H}$ represents heating (which would
+include that due to dust-gas friction) and $\mathcal{H}_\mathrm{eff}$
+is an effective source term arising from transforming from the
+two-fluid to one-fluid formalism \citep[see][ for
+details]{laibe14}. We also include a simple model of radiative
+cooling, 
+\begin{align}
+  \Lambda = \frac{T -
+    T_\mathrm{ref}}{t_\mathrm{cool}}, \label{cooling} 
+\end{align}
+which relaxes the mixture back to a prescribed temperature profile   
+$T_\mathrm{ref}$ on a timescale of $\tcool$. We will shortly simplify
+the problem by considering rapid cooling, $\tcool\to0$.
+
+Eq. \ref{masseq}---\ref{tempeq} are not yet equivalent to standard 
+hydrodynamics, which typically evolves two scalar fields: the density and
+temperature (or pressure). By contrast, Eq. \ref{masseq}---\ref{tempeq}
+involves 3 scalars: $\rho$, $\tepsilon$, and $T$. 
+However, we can establish an  analogy by fixing the gas
+equation of state, thus eliminating Eq. \ref{tempeq}; but then reform 
+the dust-to-gas ratio evolution equation (\ref{dusteq}) into an
+energy-like equation.
+
 \subsection{Locally isothermal equation of state}\label{loc_iso_eos}
-For later comparison it is convenient to re-write the temperature
-evolution, Eq. \ref{tempeq}, to that for the pressure: 
-\begin{align}
-  \frac{\p P}{\p t} + \bm{v}_\mathrm{g}\cdot\nabla P = - \gamma P
-  \nabla\cdot \bm{v}_\mathrm{g}  + \mathcal{H}\label{real_pressure}
-%\frac{\rho \mathcal{R}}{\mu}\frac{T
-%  - T_\mathrm{ref}}{t_\mathrm{cool}}, \label{realenergy}
-\end{align}
-where $\mathcal{H}$ collects together dust-drag heating and radiative
-cooling.
-
-We now simplify the gas energy equation by considering short cooling 
+We simplify the gas energy equation by considering short cooling 
 times $\tcool\to 0$, appropriate for the outer parts of an irradiated
@@
 where $c_s(r,z)= \sqrt{\mathcal{R}T_\mathrm{ref}/\mu}$ is a prescribed
-sound-speed profile fixed in time. In practice we consider vertically 
+sound-speed profile fixed in time. In most applications we consider vertically 
 isothermal disks with \begin{align}\label{power_temp}
@@
 Since $\tepsilon$ typically decrease away from the midplane, we expect
-vertically isothermal dusty disks to behave as if it had a temperture
-\emph{increasing} away from $z=0$.
+vertically isothermal dusty disks to behave as if the temperture
+\emph{increased} away from $z=0$.
 
 \subsection{Effective energy equation}\label{energy_analogy}
-Although we have eliminated the true energy
+Although we have deleted the true energy
 equation (Eq. \ref{tempeq}) by fixing a locally isothermal
@@
 \end{align} 
-Comparing with Eq. \ref{real_pressure}, we see that 
-Eq. \ref{eff_energy} is the energy equation for an ideal gas of
-adiabatic index $\gamma=1$, but now with the imposed temperature
-gradient $\nabla c_s^2$ and dust-gas drag $\mathcal{C}$ playing the
-role of radiative cooling ($\Lambda$ in 
-Eq. \ref{tempeq}).
+For comparison with the standard energy equation in hydrodynamics,
+re-writing Eq. \ref{tempeq} for a pure ideal gas with $P \propto
+\rhog T$, and reverting back to gas velocities, gives  
+%For later comparison it is convenient to re-write the temperature
+%evolution, Eq. \ref{tempeq}, to that for the pressure: 
+\begin{align}
+  \frac{\p P}{\p t} + \bm{v}_\mathrm{g}\cdot\nabla P = - \gamma P
+  \nabla\cdot \bm{v}_\mathrm{g}  +
+  \frac{P}{T}\left(\mathcal{H} -
+    \Lambda\right).\label{real_pressure} 
+%\frac{\rho \mathcal{R}}{\mu}\frac{T
+%  - T_\mathrm{ref}}{t_\mathrm{cool}}, \label{realenergy}
+\end{align} 
+We can thus interpret Eq. \ref{eff_energy} as the energy equation for
+and ideal gas with adiabatic index $\gamma=1$, but now with the
+imposed temperature gradient $\nabla c_s^2$ and dust-gas drag $\mathcal{C}$ 
+acting as source terms.
 
+%playing the
+%role of radiative cooling ($\Lambda$ in 
+%Eq. \ref{tempeq}). 
 %In this form, the function $\mathcal{C}$ can be interpreted
 %as a cooling term.
+
 If we denote 
@@
 which is in the same form as cooling by radiative diffusion. In protoplanetary
-disks the corresponding heat flux, proportional to $\bm{F}$, is
+disks the corresponding `heat flux', proportional to $\bm{F}$, is
 directly radially outwards and vertically upwards. This is simply a
@@
 Since we have shown that a locally isothermal dusty gas effectively
-has $\gamma=1$ we can define an effective (dimensionless)    
+has $\gamma=1$ we can define an effective 
 entropy for the \emph{mixture} as 
@@
 \end{align} 
+The constant heat capacity has been absorbed into $\seff$. 
 %This follows naturally by combining the effective 
@@
 \end{align*}
-which is just the statement of entropy evolution in standard  
+which is equivalent to entropy evolution in standard  
 hydrodynamics, albeit with source terms. With an effective
@@
 %\subsection{Dusty buoyancy forces}
-Having identified the entropy of an isothermal dusty gas, we can also
-define the disk vertical buoyancy frequency $N_z$ as   
+We can now define the vertical buoyancy frequency $N_z$ of the mixture
+as    
 %$N_z$ of the mixture is given by 
@@
 %Typically in a thin, smooth disk we have $|N_r|\ll |N_z|$. 
-where the second equality assumes a vertically isothermal disk. A
-similar definition exists for the radial buoyancy frequency
-$N_r$. However, in thin, smooth disks we have $|N_r|\ll |N_z|$.
+where the second equality assumes a vertically isothermal disk. 
+Protoplanetary disks have $\p_z\rhog<0$, thus stability against
+convection ($N_z^2>0$) requires $\p_z\tepsilon <0 $. This is equivalent to
+entropy increasing away from the midplane. A similar expression exists
+for the radial buoyancy frequency $N_r$, but $|N_r|\ll |N_z|$ in thin,
+smooth disks.
 
@@
 \citep{lin15,malygin17}. Combining these results, we estimate the 
-correspondence applicable to particles less than mm in size at a few
+correspondence will be applicable to particles less than mm in size at a few
 to 10s of AU in protoplanetary disks. However, note that the
@@
 On the other hand, Eq. \ref{dusty_solberg2} can be violated in
-disks if the dust is vertically well-mixed ($\p_z\epsilon =
-0$) but  radially-varying ($\p_r\epsilon\neq0$). In this case the
+disks if the dust is vertically well-mixed ($\p_z\tepsilon =
+0$) but  radially-varying ($\p_r\tepsilon\neq0$). In this case the
 left-hand-side of Eq. \ref{dusty_solberg2} becomes
 \begin{align}
--\frac{c_s^4}{(1+\tepsilon)^4}\left(\frac{\p\ln{\rhog}}{\p
-    z}\right)^2\left(\frac{\p\epsilon}{\p r}\right)^2<0.
+-c_s^4\left(\frac{\p\ln{\rhog}}{\p
+    z}\right)^2\left(\frac{\p\tepsilon}{\p r}\right)^2<0.
 \end{align}

### 2017-08-02 paper.tex
revising according to AY's comments, up to 3.4

@@
 equivalent dynamical equations for the center of mass velocity $\bm{v}$
-and the relative dust-gas velocity  $\Delta\bm{v}\equiv \bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$ 
+and the relative dust-gas velocity  $\bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$ 
 \citep{youdin05a}. In this reformulation, continuity equations for dust and gas can be replaced 

### 2017-08-02 summary.tex
revising according to AY's comments, up to 3.4

@@
 \Omega$ (see Appendix \S\ref{spurious_epi}). This 
-results from neglecting terms of $O(|\widetilde{\bm{v}}|^2)$ after converting
+results from neglecting terms of $O(|\bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}|^2)$ after converting
 the full two-fluid equations to the one-fluid formulation. The full
@@
 %model \cite[e.g.][]{laibe14}. It 
- includes an evolutionary equation for $\widetilde{\bm{v}}$, which has no
+ includes an evolutionary equation for $\widetilde{\bm{v}}\equiv\bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$, which has no
  hydrodynamic counter-part \citep{laibe14}. However, it is

### 2017-08-02 linear.tex
revising in response to AY comments, up to section 5.1.1

@@
 We now examine instabilities in dusty protoplanetary disks
-based on explicit descriptions of the equilibrium state. We first show
+based on explicit descriptions of the equilibrium state. We first
+specify explicit disk structures to analyze. We show 
 that sharp radial edges in the dust component are unstable. We then
@@
 
-We usually consider gas-dominated disks with $\epsilon_0 < 1$.  
-Then $\rhog(r,z)$ is effectively Gaussian, as in the 
-dust-free case. The dust density is approximately 
+In gas-dominated disks with $\epsilon_0 < 1$ the gas distribution
+$\rhog(r,z)$ is close to Gaussian, as in the   
+dust-free case, and the dust density is approximately 
 \begin{align}
@@
 \end{align}
-and $H_\mathrm{d}$ is the dust-scale height. In numerical
-calculations, we  we specify $H_\mathrm{d}< \Hgas$ to obtain 
-$\Htilde$ for input.
+and $H_\mathrm{d}$ is the dust-scale height.% In numerical
+%calculations, we  we specify $H_\mathrm{d}< \Hgas$ to obtain 
+%$\Htilde$ for input.
 
@@
 \subsubsection{Orbital frequency} 
-In the thin-disk approximation, the disk orbital frequency is 
+From Eq. \ref{steady_momr}} the disk orbital frequency is 
 \begin{align}

### 2017-08-02 model.tex
revising in response to AY comments, up to section 5.1.1

@@
 \bm{v}_\mathrm{g}$ obeys a complicated evolutionary equation 
-\citep[see, e.g.][]{youdin05a}.% obtained from taking the difference
+\citep[see, e.g.][]{youdin05a}. % obtained from taking the difference
 %between evolutionary equations for $\bm{v}_\mathrm{g}$ and
@@
 In this section we discuss the stability properties of the 
-dusty-gas mixture using a variational principle. This method does not
-require explicit solutions to the equilibrium equations. Self-gravity is neglected here.
+dusty-gas mixture using a variational principle. This approach does not
+require explicit solutions (i.e. a specific distribution of the density and dust-to-gas ratio)  
+to the equilibrium equations. We consider axisymmetric systems and neglect self-gravity here.
 
 \subsection{Steady states}\label{eqm}
-We consider axisymmetric steady states as background equilibria to
-perturb. For a given distribution of the dust-fraction $\tepsilon$ (or
+%We consider axisymmetric steady states as background equilibria to
+%perturb.
+
+For a given distribution of the dust-fraction $\tepsilon$ (or
 dust-to-gas ratio $\epsilon$), the 
 mass and momentum Eqs. \ref{masseq}---\ref{momeq} admit     
-axisymmetric steady states with $\rho(r,z)$ and 
+solutions with $\rho(r,z)$ and 
 $\bm{v}=r\Omega(r,z)\hat{\bm{\phi}}$ where $\Omega = v_\phi/r$, which satisfy 
@@
 (Eq. \ref{eos}). An explicit solution is presented in 
-\S\ref{steady_state}, when we explicitly analyse the stability of
+\S\ref{steady_state}, when we analyse the stability of
 protoplanetary disks.
@@
 %  \frac{\p}{\p r} \left[c_s^2(1-\tepsilon)\right]\\  
-   = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p \seff}{\p z} -\frac{\p
-    P}{\p z}\frac{\p \seff}{\p r} \right). 
+   &= \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p \seff}{\p z} -\frac{\p
+    P}{\p z}\frac{\p \seff}{\p r} \right).% \\
+%& = \frac{1}{\rhog}\left(\frac{\p P}{\p r}\frac{\p \seff}{\p z} -\frac{\p
+%    P}{\p z}\frac{\p \seff}{\p r} \right)
 \end{align}
-Since $\seff$ depends on $\tepsilon$, we see that radial gradients in
-the dust-to-gas ratio also drives vertical shear \citep[cf. that 
-driven by $\p_z\epsilon$ usually discussed in dusty disks,  
-e.g.][]{chiang08}.
+and recall $\seff = \ln{\left[c_s^2\left(1-\tepsilon\right)\right]}$. It is well appreciated that 
+vertical stratification of the dust layer contributes to vertical shear
+\citep{chiang08}. Eq. \ref{vshear} shows that a radial dust
+stratification ($\p_r\tepsilon$) also contributes to vertical shear.
 
-However, disk structures obtained from
-Eq. \ref{steady_momr}---\ref{steady_momz} generally do not satisfy  
-the steady state effective energy Eq. \ref{eff_energy} because typically 
-$\mathcal{C}\neq0$.  In order to have a well-defined stability problem, we formally
-consider equilibria with $\mathcal{C}\equiv 0$. Namely      
-\begin{inparaenum}[1)] 
+%Since $\seff$ depends on $\tepsilon$, we see that radial gradients in
+%the dust-to-gas ratio also drives vertical shear \citep[cf. that 
+%driven by $\p_z\epsilon$ usually discussed in dusty disks,  
+%e.g.][]{chiang08}.
+
+Our equilibrium solutions satisfy the hydrostatic constraints of
+Eq. \ref{steady_momr}---\ref{steady_momz} but are not in general
+steady state solutions to the energy 
+equation (\ref{eff_energy}), because the source term $\mathcal{C} \ne0$ for realistic disks.  
+Limiting cases with $\mathcal{C} \equiv 0$ include \begin{inparaenum}[1)] 
 \item 
-  unstratified or 2D disks with finite dust-gas drag; %(\S\ref{si});   
+  unstratified or 2D, razor-thin disk models with finite dust-gas drag; %(\S\ref{si});   
 \item 
@@
 \end{inparaenum} 
-Alternatively, we are assuming that $\tstop$  is sufficiently small
-such that the evolution of the background disk (e.g. dust-settling) 
+The following analyses thus apply strictly to these limiting cases of
+$\mathcal{C} = 0$.  Our analysis is also a good approximation when
+$\tstop$ is sufficiently small such that the evolution of the background disk (e.g. dust-settling) 
 occur on much longer timescales than the instabilities of interest.
+
+%However, disk structures obtained from
+%Eq. \ref{steady_momr}---\ref{steady_momz} generally do not satisfy  
+%the steady state effective energy Eq. \ref{eff_energy} because typically 
+%$\mathcal{C}\neq0$.  In order to have a well-defined stability problem, we formally
+%consider equilibria with $\mathcal{C}\equiv 0$. Namely      
+%\begin{inparaenum}[1)] 
+%\item 
+%  unstratified or 2D disks with finite dust-gas drag; %(\S\ref{si});   
+%\item 
+%  perfectly-coupled dust with $\tstop=0$. %(\S\ref{results}). 
+%\end{inparaenum} 
+%Alternatively, we are assuming that $\tstop$  is sufficiently small
 
@@
 \end{align}
-where $\sigma$ is the complex mode frequency. For later convenience we
+where $\sigma$ is the complex mode frequency. We 
 write 
@@
 
-In Appendix \ref{var_prin} we derive the integral relation from the
-corresponding linearized equations (Eqs.\ref{masseq}--\ref{momeq}, and 
-Eq. \ref{eff_energy} in place of the dust continuity equation): 
+In Appendix \ref{var_prin} we derive linearize Eqs. \ref{masseq}, \ref{momeq}, and 
+Eq. \ref{eff_energy} to derive the following integral relation: 
 \begin{align}
@@
 \subsection{Strictly isothermal gas perfectly coupled to dust}\label{iso_perfect}
-When $c_s^2$ is a constant and $\tstop=0$, the dusty-gas equations are
-exactly equivalent to that for adiabatic hydrodyamics with unit adiabatic
+When $c_s^2$ is a constant and $\tstop=0$ (so that $\mathcal{C} = 0$), the dusty-gas equations are
+exactly equivalent to that for adiabatic hydrodyamics with unit
+adiabatic 
 index. Although the gas is strictly isothermal, the mixture behaves 
@@
 \begin{align}
-  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\tepsilon &> 0,\label{dusty_solberg1}  \\
-  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\tepsilon}{\p
+  \kappa^2 + \frac{1}{\rhog}\nabla P \cdot\nabla\tepsilon &> 0,\label{dusty_solberg1}  \\
+  -\frac{1}{\rhog}\frac{\p P}{\p z}\left(-\kappa^2\frac{\p\tepsilon}{\p
     z} + r\frac{\p\Omega^2}{\p z}\frac{\p \tepsilon}{\p r} \right) & > 0, \label{dusty_solberg2}
 \end{align} 
+%\begin{align}
+%  \kappa^2 + c_s^2 \nabla\ln{\rhog}\cdot\nabla\tepsilon &> 0,\label{dusty_solberg1}  \\
+%  -c_s^2\frac{\p\ln{\rhog}}{\p z}\left(-\kappa^2\frac{\p\tepsilon}{\p
+%    z} + r\frac{\p\Omega^2}{\p z}\frac{\p \tepsilon}{\p r} \right) & > 0, \label{dusty_solberg2}
+%\end{align} 
 %\begin{align}
@@
 $\kappa^2 \equiv r^{-3}\p_r\left(r^4\Omega^2\right)>0$ is the 
-square of the epicylic frequency. %We assume $\kappa^2>0$.  
+square of the epicylic frequency. Note that $\nabla P/\rhog =
+c_s^2\nabla \ln{\rhog}$ for strictly isothermal gas considered here.
+
 Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} are in fact the 
 the Solberg-Hoiland criteria for axisymmetric stability of adiabatic
-gas if one inserts $C_p\ln{\left(1 -
+gas if one inserts $C_P\ln{\left(1 -
     \tepsilon\right)}$ for the entropy into the standard
-expression for said criteria \citep{tassoul78}. 
+expression for said criteria \citep{tassoul78}. This substitution is
+consistent with our definition of the effective entropy $\seff$ since
+we are considering constant $c_s$.
+
 We expect Eq. \ref{dusty_solberg1} to be satisfied in typical
@@
 On the other hand, Eq. \ref{dusty_solberg2} can be violated in
-disks if the dust is vertically well-mixed ($\p_z\tepsilon =
-0$) but  radially-varying ($\p_r\tepsilon\neq0$). In this case the
+disks if the dust is vertically well-mixed but radially-stratified,  
+such that $\p_z\tepsilon = 0$ but $\p_r\tepsilon\neq0$). In this case the
 left-hand-side of Eq. \ref{dusty_solberg2} becomes
 \begin{align}
--c_s^4\left(\frac{\p\ln{\rhog}}{\p
+-\left(\frac{1}{\rhog}\frac{\p P}{\p
     z}\right)^2\left(\frac{\p\tepsilon}{\p r}\right)^2<0.
 \end{align}
-Such an isothermal dusty disk is \emph{unstable} but there is no
-effective vertical buoyancy to stabilize vertical motions ($N_z=0$)
-that can tap into the free energy associated with  
-vertical shear due to the radial gradient in the dust-to-gas ratio. 
+%\begin{align}
+%-c_s^4\left(\frac{\p\ln{\rhog}}{\p
+%    z}\right)^2\left(\frac{\p\tepsilon}{\p r}\right)^2<0.
+%\end{align}
+Such an isothermal dusty disk is \emph{unstable} because there is no
+effective vertical buoyancy to stabilize vertical motions ($N_z=0$),
+which can tap into the free energy associated with  
+vertical shear due to the radial gradient in the dust-fraction. 
 This situation is identical to the pure gas, adiabtic simulations of \cite{nelson13}
@@
 \subsection{Thermodynamics of dust-drag instabilities}\label{dust_work}
-Now consider constant $c_s$ but $\tstop\neq0$. 
-Eq. \ref{int_rel} indicates that $\sigma^2$ is generally complex, so
+The physical interpretation of \S\ref{grow_osc} is here derived in
+detail. Consider constant $c_s$ but $\tstop\neq0$ (so $\delta\mathcal{C}
+\neq 0$). Eq. \ref{int_rel} indicates that $\sigma^2$ is generally complex, so
 we may have growing oscillations, or overstability, due to dust-gas
@@
 $\imag\left[\left(\nabla\cdot\dd\bm{v}^*\right)\dd\mathcal{C}\right]$
-may be thought of as a correlation between compression/expansion and  
+represent correlations between compression/expansion and  
 heating/cooling.
@@
 \end{align}
-\citep{cox67}, 
+\citep[][Eq. 4.10]{cox67}, 
 where $\upsilon=1/\rho$ is the specific volume of the mixture, $D/Dt$
@@
 again assuming $\omega\neq0$. 
-This is VSI caused by vertical shear arising from a radial
+This instability represents VSI caused by vertical shear arising from a radial
 temperature gradient \citep{nelson13,barker15,lin15}. We present 

### 2017-08-03 appendix.tex
edited in response to AY comments, up to 5.2

@@
 and $\Gamma$ is a constant. The effective energy equation is then
-Eq. \ref{poly_energy}. % We also include the self-gravitational
+Eq. \ref{poly_energy}, which generalizes
+Eq. \ref{eff_energy}. Assuming axisymmetry throughout, we linearize this equation along with Eq. \ref{masseq},  
+\ref{momeq},  and \ref{poisson} gives:
+
+% We also include the self-gravitational
 % potential $\psi$ of the gas
@@
 % %\end{align}
-The linearized equations for axisymmetric
-disturbances are
 
 \begin{align}
-  \ii\sigma\frac{\delta\rho}{\rho} &= \nabla\cdot\dd\bm{v} +
-  \dd\bm{v}\cdot\nabla\ln{\rho},\label{lin_mass_full}\\
-%  \ii\sigma\frac{\delta P}{P} &= \nabla\cdot\dd\bm{v} +
-   \ii\sigma\frac{\delta P}{P} &= \Gamma \nabla\cdot\dd\bm{v} +
-  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{K} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
+%  \ii\sigma\frac{\delta\rho}{\rho} &= \nabla\cdot\dd\bm{v} +
+%  \dd\bm{v}\cdot\nabla\ln{\rho},\label{lin_mass_full}\\
+ \ii\sigma\frac{\Delta\rho}{\rho} &= \ii\sigma\frac{\delta\rho}{\rho} - 
+  \dd\bm{v}\cdot\nabla\ln{\rho} =  \nabla\cdot\dd\bm{v},\label{lin_mass_full}\\
+ %  \ii\sigma\frac{\delta P}{P} &= \Gamma \nabla\cdot\dd\bm{v} +
+ % \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{K} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
+  \ii\sigma\frac{\Delta P}{P} &= \ii\sigma\frac{\delta P}{P} -
+  \dd\bm{v}\cdot\nabla\ln{P} = 
+\Gamma \nabla\cdot\dd\bm{v}  - \dd\bm{v}\cdot\nabla\ln{K} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
   -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
@@
 where the linearizd pressure force $\delta \bm{F}$ is given in
-Appendix \ref{lin_press}. These equations do not assume the
-radially-local approximation used in numerical computations. From the
+Appendix \ref{lin_press}. Recall $\Delta = \delta +
+\bm{\xi}\cdot\nabla$ is the Lagrangian derivative and $\bm{\xi}$ is the
+Lagrangian displacement. For axisymmetric flow, we have
+$\ii\sigma\bm{\xi}\cdot\nabla = -\dd\bm{v}\cdot\nabla$.  These equations do not assume the
+radially-local approximation used in numerical computations. Note that
+Eq. \ref{lin_mass_full}---\ref{lin_energy_full} imply 
+\begin{align}
+\left|\sigma\right|^2\frac{\Delta P}{P}\frac{\Delta \rho^*}{\rho} =
+\Gamma\left|\nabla\cdot\dd\bm{v}\right|^2 -
+\nabla\cdot\dd\bm{v}^*\left(\dd\bm{v}\cdot\nabla\ln{K} + \frac{\dd\mathcal{C}}{P}\right),
+\end{align}
+so the first (real) term on the right hand side does not contribute to
+$\imag{\left(\Delta P\Delta\rho^*\right)}$. Setting $K$ to constant
+and taking the imaginary part gives Eq. \ref{pdv}.
+
+From the
 linearized meridional momentum equations, we find  
-
 \begin{align}
@@
 \begin{align}
-  \tstop  =  \frac{K}{\rho c_s}\label{epstein}
+  \tstop  \propto  \frac{1}{\rho c_s}\label{epstein}
 \end{align}
-\citep{price15}, where $K$ is a constant. Then the dust diffusion
+\citep{price15}. Then the dust diffusion
 function becomes

### 2017-08-03 linear.tex
edited in response to AY comments, up to 5.2

@@
 \subsubsection{Orbital frequency} 
-From Eq. \ref{steady_momr}} the disk orbital frequency is 
+From Eq. \ref{steady_momr} the disk orbital frequency is 
 \begin{align}
@@
 over that due to the radial temperature gradient for thin dust layers
-such that $\delta^2\ll \epsilon$. Otherwise, vertical shear is
-associated with the radial temperature gradient. %However, axisymmtric
+such that $\delta^2\ll \epsilon$. %Otherwise, vertical shear is
+%associated with the radial temperature gradient. %However, axisymmtric
 %\emph{instability} always caused by $\p_rc_s^2$.
@@
 Here we apply the dusty analog of the Solberg-Hoiland criteria derived
-in \S\ref{iso_perfect}. As discussed there, the first criterion in
-generally satisifed. Thus we only consider the second condition,
-Eq. \ref{dusty_solberg2}.
+in \S\ref{iso_perfect} assuming strictly isothermal gas. As discussed
+there, the first criterion in generally satisifed. Thus we only
+consider the second condition, Eq. \ref{dusty_solberg2}.
 
@@
   % \frac{z^2}{\Htilde^2}\right)>0 \label{stability_est1},   \\ 
+% &1 - \frac{\epsilon
+%   h_\mathrm{g}^2}{\left(1+\epsilon\right)^2}
+%   \frac{\p\ln{\epsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\Htilde^2}{H_\mathrm{g}^2}\frac{\p\ln{\epsilon}}{\p\ln{r}}\right)
+%   > 0 \label{stability_est2},
 &1 - \frac{\epsilon
   h_\mathrm{g}^2}{\left(1+\epsilon\right)^2}
-  \frac{\p\ln{\epsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\Htilde^2}{H_\mathrm{g}^2}\frac{\p\ln{\epsilon}}{\p\ln{r}}\right)
+  \frac{\p\ln{\epsilon}}{\p\ln{r}}\left(-\frac{\p\ln{P}}{\p\ln{r}}+\frac{\Htilde^2}{H_\mathrm{g}^2}\frac{\p\ln{\epsilon}}{\p\ln{r}}\right)
   > 0 \label{stability_est2},
@@
 has occured). This means that typical dusty PPDs are stable to
-axisymmetric perturbations, a conclusion valid for arbitrarily thin
+axisymmetric perturbations, even for arbitrarily thin
 dust layers.
@@
 \begin{align}\label{spm}
+% \chi_\pm = \frac{1}{2}\frac{H_\mathrm{g}^2}{\Htilde^2} 
+%   \left[
+%   \frac{\p\ln{\rhog}}{\p\ln{r}} \pm 
+%   \sqrt{
+%   \left(\frac{\p\ln{\rhog}}{\p\ln{r}}\right)^2 + 
+%   4 \frac{\Htilde^2}{H_\mathrm{g}^2}
+%   \frac{\left(1+\epsilon\right)^2}{\epsilon h_\mathrm{g}^2}
+%   }
+%   \,\right]. 
 \chi_\pm = \frac{1}{2}\frac{H_\mathrm{g}^2}{\Htilde^2} 
   \left[
-  \frac{\p\ln{\rhog}}{\p\ln{r}} \pm 
+  \frac{\p\ln{P}}{\p\ln{r}} \pm 
   \sqrt{
-  \left(\frac{\p\ln{\rhog}}{\p\ln{r}}\right)^2 + 
+  \left(\frac{\p\ln{P}}{\p\ln{r}}\right)^2 + 
   4 \frac{\Htilde^2}{H_\mathrm{g}^2}
@@
 \end{align} 
-In typical accretion disks where $\p_r\rhog<0$, instability is easier
+In typical accretion disks where $\p_rP<0$, instability is easier
 for increasing dust-to-gas ratios ($\p_r\epsilon > 0$).
 
-We can neglect density gradients in Eq. \ref{spm} 
-if $r\p_r\ln{\rhog}\sim O(1)$ and $H_\epsilon/\Hgas\gg
+We can neglect pressure gradients in Eq. \ref{spm} 
+if $r\p_r\ln{P}\sim O(1)$ and $H_\epsilon/\Hgas\gg
 \sqrt{\epsilon}\hgas$. For example, if $\epsilon\simeq 0.01$ and
@@
 
-Technically, the above discussion is only applicable to strictly
-isothermal disks perfectly coupled to dust (see \S\ref{iso_perfect}). 
-However, since dusty edges translate to sharp entropy gradients (see
+Technically, the above discussion is only applicable when $c_s$ is
+constant and $\tstop=0$ (see \S\ref{iso_perfect}). However, since
+dusty edges translate to sharp entropy gradients (
 \S\ref{dusty_entropy}), we may generally expect sharp features in the

### 2017-08-03 model.tex
edited in response to AY comments, up to 5.2

@@
 The following analyses thus apply strictly to these limiting cases of
-$\mathcal{C} = 0$.  Our analysis is also a good approximation when
+$\mathcal{C} = 0$ in equilibrium.  Our analysis is also a good approximation when 
 $\tstop$ is sufficiently small such that the evolution of the background disk (e.g. dust-settling) 
@@
 \end{align}
-\citep[][Eq. 4.10]{cox67}, 
+\citep[][see their Eq. 4.10 and related discussions]{cox67}, 
 where $\upsilon=1/\rho$ is the specific volume of the mixture, $D/Dt$
@@
   -|\sigma|^2\imag \left(\Delta P 
-  \frac{\Delta\rho^*}{\rho}\right). \label{pdv}
+  \frac{\Delta\rho^*}{\rho}\right), \label{pdv}
 \end{align}
+for constant $c_s$. 
 %This states that `PdV' work (the right-hand-side) is 

### 2017-08-03 model.tex
misc minor edits

@@
 where the second equality assumes a vertically isothermal disk. 
-Protoplanetary disks have $\p_z\rhog<0$, thus stability against
-convection ($N_z^2>0$) requires $\p_z\tepsilon <0 $. This is equivalent to
+Protoplanetary disks have $\p_z\rhog<0$, thus stability against vertical 
+convection ($N_z^2>0$) requires $\p_z\tepsilon <0 $, i.e. the dust density drops faster 
+than the gas density away from the midplane. 
+This is equivalent to 
 entropy increasing away from the midplane. A similar expression exists
@@
 $\mathcal{C} = 0$ in equilibrium.  Our analysis is also a good approximation when 
-$\tstop$ is sufficiently small such that the evolution of the background disk (e.g. dust-settling) 
-occur on much longer timescales than the instabilities of interest.
+$\tstop$ is sufficiently small such that the evolution of the background disk (e.g. dust-settling timescales) 
+occur on much longer timescales than the dust dynamics of interest (e.g. instability growth timescales).
 

### 2017-08-04 linear.tex
misc minor edits
removed higher order correction section for submission, misc edits

@@
 \end{align}
-is the characteristic disk aspect-ratio.
+is the characteristic disk aspect-ratio, with $h_\mathrm{g}\ll 1$ for protoplanetary disks.
 
@@
 if $r\p_r\ln{P}\sim O(1)$ and $H_\epsilon/\Hgas\gg
-\sqrt{\epsilon}\hgas$. For example, if $\epsilon\simeq 0.01$ and
+\sqrt{\epsilon}\hgas/2\left(1+\epsilon\right)$. For example, if $\epsilon\simeq 0.01$ and
 $\hgas\simeq 0.05$, then we require $H_\epsilon/\Hg\gg
-5\times10^{-3}$. (So the dust layer thickness $\Hd\gg 5
-\times10^{-3}\Hgas$ as well.) Then 
-instability requires 
+2\times10^{-3}$. This can be met if the dust is not well settled. 
+%(So the dust layer thickness $\Hd\gg 2
+%\times10^{-3}\Hgas$ as well.) 
+Then instability requires 
 \begin{align}\label{dust_edge}

### 2017-08-04 model.tex
removed higher order correction section for submission, misc edits

@@
 %The real integral $\mathcal{I}^2>0$ 
-where
+where $^*$ denotes the complex conjugate, 
 \begin{align}
@@
 is the meridional kinetic energy, 
-and coefficients $A,B,D$ be
+and coefficients $A,B,D$ can be 
 read off Eq. \ref{integral_ex}. %Note that $B=C$ from dynamical equilibrium
@@
 
-Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} are in fact the 
-the Solberg-Hoiland criteria for axisymmetric stability of adiabatic
-gas if one inserts $C_P\ln{\left(1 -
+Eq. \ref{dusty_solberg1}---\ref{dusty_solberg2} can in fact be obtained by
+inserting $C_P\ln{\left(1 - 
     \tepsilon\right)}$ for the entropy into the standard
-expression for said criteria \citep{tassoul78}. This substitution is
+expression for the Solberg-Hoiland criteria 
+\citep{tassoul78} for axisymmetric stability of adiabatic gas. 
+This substitution is 
 consistent with our definition of the effective entropy $\seff$ since

### 2017-08-04 paper.tex
removed higher order correction section for submission, misc edits

@@
 \item 
+ radial dust edges are unstable if dust is vertically well-mixed;  
+\item 
   the streaming instability neccessarily involves a gas pressure
@@
 processes (e.g. Rossby wave instability, convective overstability, and 
-zombie vortices), how to simulate dusty protoplanetary disks
-with minor tweaks to purely hydrodynamic codes, as well as possible generalizations
-of our analogy between dust-laden and nearly adiabatic flows.
+zombie vortices), and how to simulate dusty protoplanetary disks
+with minor tweaks to existing codes for pure gas dynamics.%, as well as possible generalizations
+%of our analogy between dust-laden and nearly adiabatic flows.
 
@@
 disks. We then analyze explicit disk models 
-\S\ref{linear_problem}, were we show that dusty edges are unstable,
+\S\ref{linear_problem}, were we show that radial dust edges are unstable,
 and revisit the SGI and SI in the tight 

### 2017-08-04 summary.tex
removed higher order correction section for submission, misc edits

--- 2017-05-09 summary.tex
@@
 simulating protoplanetary disks coupled to small dust. We will apply 
-our modified code --- d\textsc{Pluto} --- to the streaming instability
-and disk-planet interactions in an upcoming publication (Lin et al.,
-in preparation).
+this modified code ---d\textsc{Pluto}--- to study dusty disk-planet interaction (Lin et al., 
+in preparation). Preliminary results indicate the simplified one-fluid 
+approach can reproduce features such as dusty rings associated with 
+planet-induced gaps similar to that obtained from explicit two-fluid
+simulations \citep[e.g.][]{dong17}.
 
-\subsection{Higher order corrections}
-{\bf notes for self. maybe exclude in final. maybe move into appendix}
+% \subsection{Higher order corrections}
+% {\bf notes for self. maybe exclude in final. maybe move into appendix}
 
-The simplified one-fluid formulation, 
-Eqs. \ref{masseq}---\ref{tempeq}, can 
-produce spuriously-growing modes with frequency $|\sigma|\gtrsim
-\Omega$ (see Appendix \S\ref{spurious_epi}). This 
-results from neglecting terms of $O(|\widetilde{\bm{v}}|^2)$ after converting
-the full two-fluid equations to the one-fluid formulation. The full
-one-fluid equations 
-%is significantly more complex than our 
-%model \cite[e.g.][]{laibe14}. It 
- includes an evolutionary equation for $\widetilde{\bm{v}}$, which has no
- hydrodynamic counter-part \citep{laibe14}. However, it is
- still possible to improve upon 
- the first-order formulation whilst retaining similarilty to  
- standard hydrodynamics.  
-%without having the complexity of the full equations
+% The simplified one-fluid formulation, 
+% Eqs. \ref{masseq}---\ref{tempeq}, can 
+% produce spuriously-growing modes with frequency $|\sigma|\gtrsim
+% \Omega$ (see Appendix \S\ref{spurious_epi}). This 
+% results from neglecting terms of $O(|\bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}|^2)$ after converting
+% the full two-fluid equations to the one-fluid formulation. The full
+% one-fluid equations 
+% %is significantly more complex than our 
+% %model \cite[e.g.][]{laibe14}. It 
+%  includes an evolutionary equation for $\widetilde{\bm{v}}\equiv\bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$, which has no
+%  hydrodynamic counter-part \citep{laibe14}. However, it is
+%  still possible to improve upon 
+%  the first-order formulation whilst retaining similarilty to  
+%  standard hydrodynamics.  
+% %without having the complexity of the full equations
 
-The full one-fluid equations of motion includes the acceleration term 
-\begin{align}
-%\left.\frac{D\bm{v}}{Dt}\right|_\mathrm{add} = 
-%\bm{a} =  
--\frac{1}{\rho}\nabla\cdot\left(\frac{\rhod\rhog}{\rho}\widetilde{\bm{v}}\otimes\widetilde{\bm{v}}\right) 
-\equiv  \frac{1}{\rho}\nabla\cdot \bm{T} 
-\end{align}
-\citep{youdin05a}, 
-where $\bm{T}$ is an effective stress tensor. 
-In the terminal velocity
-approximation with $\widetilde{\bm{v}} = \tstop 
-\nabla P/\rhog$, we have 
-\begin{align}
-\bm{T} = - \frac{\epsilon\tstop^2}{\rho}\nabla P \otimes \nabla P.
-\end{align}
-If we define 
-\begin{align}
-  \bm{g} \equiv \sqrt{4\pi G \epsilon \rho} \frac{\nabla P}{\rho}
-  \tstop,\label{effective_g} 
-\end{align}
-then
-\begin{align}
-\bm{T} = - \frac{1}{4\pi G} \bm{g}\otimes\bm{g}, \label{effective_gstress}
-\end{align}
-%{\bf tensor virial theorems?}
-which is just the anisotropic part of the  
-gravitational stress tensor in self-gravitating hydrodynamics 
-\citep{lynden-bell72}.
+% The full one-fluid equations of motion includes the acceleration term 
+% \begin{align}
+% %\left.\frac{D\bm{v}}{Dt}\right|_\mathrm{add} = 
+% %\bm{a} =  
+% -\frac{1}{\rho}\nabla\cdot\left(\frac{\rhod\rhog}{\rho}\widetilde{\bm{v}}\otimes\widetilde{\bm{v}}\right) 
+% \equiv  \frac{1}{\rho}\nabla\cdot \bm{T} 
+% \end{align}
+% \citep{youdin05a}, 
+% where $\bm{T}$ is an effective stress tensor. 
+% In the terminal velocity
+% approximation with $\widetilde{\bm{v}} = \tstop 
+% \nabla P/\rhog$, we have 
+% \begin{align}
+% \bm{T} = - \frac{\epsilon\tstop^2}{\rho}\nabla P \otimes \nabla P.
+% \end{align}
+% If we define 
+% \begin{align}
+%   \bm{g} \equiv \sqrt{4\pi G \epsilon \rho} \frac{\nabla P}{\rho}
+%   \tstop,\label{effective_g} 
+% \end{align}
+% then
+% \begin{align}
+% \bm{T} = - \frac{1}{4\pi G} \bm{g}\otimes\bm{g}, \label{effective_gstress}
+% \end{align}
+% %{\bf tensor virial theorems?}
+% which is just the anisotropic part of the  
+% gravitational stress tensor in self-gravitating hydrodynamics 
+% \citep{lynden-bell72}.
 
- However, unlike self-gravity, where $\bm{g}$ is
-determined through the Poisson equation; here $\bm{g}$ is directly
-related to $P$ (Eq. \ref{effective_g}), which itself obeys an 
-evolutionary equation (Eq. \ref{eff_energy}). 
-Nevertheless, Eq. \ref{effective_g}---\ref{effective_gstress}
-show that correlations in  pressure gradients lead to 
-momentum transport of the mixture as a whole. This is similar to Reynolds stresses in
-a turbulent disk \citep{balbus99}. In particular, 
-outwards angular momentum transport requires $\p_rP\p_\phi P > 0$.
+%  However, unlike self-gravity, where $\bm{g}$ is
+% determined through the Poisson equation; here $\bm{g}$ is directly
+% related to $P$ (Eq. \ref{effective_g}), which itself obeys an 
+% evolutionary equation (Eq. \ref{eff_energy}). 
+% Nevertheless, Eq. \ref{effective_g}---\ref{effective_gstress}
+% show that correlations in  pressure gradients lead to 
+% momentum transport of the mixture as a whole. This is similar to Reynolds stresses in
+% a turbulent disk \citep{balbus99}. In particular, 
+% outwards angular momentum transport requires $\p_rP\p_\phi P > 0$.
 

### 2017-08-07 appendix.tex
resolve notation conflict in appendix C (now use curly K for the constant in drag law)
removed higher order correction section for submission, misc edits
misc home edits

@@
 Appendix \ref{lin_press}. Recall $\Delta = \delta +
-\bm{\xi}\cdot\nabla$ is the Lagrangian derivative and $\bm{\xi}$ is the
-Lagrangian displacement. For axisymmetric flow, we have
-$\ii\sigma\bm{\xi}\cdot\nabla = -\dd\bm{v}\cdot\nabla$.  These equations do not assume the
+\bm{\xi}\cdot\nabla$ is the Lagrangian perturbation and $\bm{\xi}$ is the
+Lagrangian displacement. In addition, 
+$\ii\sigma\bm{\xi}\cdot\nabla = -\dd\bm{v}\cdot\nabla$ for axisymmetric flow.  These equations do not assume the
 radially-local approximation used in numerical computations. Note that
@@
 \begin{align}
-  \tstop  \propto  \frac{1}{\rho c_s}\label{epstein}
+  \tstop   =   \frac{\mathcal{K}}{\rho c_s}\label{epstein}
 \end{align}
-\citep{price15}. Then the dust diffusion
+\citep{price15}, where $\mathcal{K}$ is a constant. Then the dust diffusion
 function becomes
@@
   \mathcal{C} \equiv c_s^2\nabla\cdot\left(\tepsilon\tstop\nabla
-  P\right) = - K c_s^2 \nabla\cdot
+  P\right) = - \mathcal{K} c_s^2 \nabla\cdot
   \left(\frac{\tepsilon}{c_s}\bm{F}\right) =
-  -Kc_s \left(\bm{F}\cdot\nabla\tepsilon + \tepsilon \nabla\cdot\bm{F}
+  -\mathcal{K}c_s \left(\bm{F}\cdot\nabla\tepsilon + \tepsilon \nabla\cdot\bm{F}
   - \frac{1}{2}\tepsilon\bm{F}\cdot\nabla\ln{c_s^2}\right).  
@@
 \begin{align}
-  - \frac{\dd\mathcal{C}}{Kc_s} = \bm{F}\cdot\nabla\dd\tepsilon +
+  - \frac{\dd\mathcal{C}}{\mathcal{K}c_s} = \bm{F}\cdot\nabla\dd\tepsilon +
   \dd\bm{F}\cdot\nabla\tepsilon + \dd\tepsilon\nabla\cdot\bm{F} +
@@
 The artifical factor $\zeta = 1$  is inserted to keep track of the
-left-hand-side of the energy equation. We note that the incompressible
-condition used by \cite{jacquet11} is obtained by setting $\zeta$ to zero. 
+left-hand-side of the energy equation.  Setting $\zeta$ to zero is
+equivalent to assuming incompressible gas \citep{jacquet11}. 
 The linearized dust diffusion
@@
 \end{align}
-(For fixed $\fdust$ but allowing $K_x\to\infty$, the imaginary term can be
-neglected.) We show in \S\ref{si_dust_rich_approx} that this
+(The imaginary term can be
+neglected for fixed $\fdust$ but allowing $K_x\to\infty$.) We show in \S\ref{si_dust_rich_approx} that this
 simplification is valid for dust-rich disks.

### 2017-08-07 model.tex
last head to toe edit before submitting, in progress

@@
 \end{align}
-is the dust-gas friction force per unit volume.%, where
+is the dust-gas friction force per unit volume. %, where
 %\begin{align}
@@
 particles with $\tstop\OmK\ll 1 $, where $\OmK$ is the Keplerian orbital
-frequency, we can use the 
+frequency (since we are interested in protoplanetary disks), we can use the 
 `terminal velocity approximation' to set %The dust-gas velocity difference 
@@
 
-Eq. \ref{dusteq} results from mass conservation of dust, where 
+Eq. \ref{dusteq} is obtained from the dust continuity equation,  
+$\p_t\rhod + \nabla\cdot\left(\rhod\bm{v}_\mathrm{d}\right)= 0$, 
+by writing 
+$\bm{v}_\mathrm{d}=\bm{v} + \tstop\nabla P/\rho$, and eliminating  
+$\rhod$ in favor of the dust fraction $\tepsilon$: 
 \begin{align}
-  \tepsilon \equiv \frac{\rhod}{\rho}  = \frac{\epsilon}{1+\epsilon} 
+  \tepsilon \equiv \frac{\rhod}{\rho}  = \frac{\epsilon}{1+\epsilon},
 \end{align}
-is the dust-fraction and $\epsilon=\rhod/\rhog$ is the usual
-dust-to-gas ratio. If $\tstop=0$ then the 
+where $\epsilon=\rhod/\rhog$ is the usual
+dust-to-gas ratio. 
+If $\tstop=0$ then the 
 dust-to-gas ratio is conserved following the fluid. Otherwise, 
@@
 
-The total gravitational potential is $\Phi_\mathrm{tot}=\Phi +
-\psi$. Here $\Phi$ originates from a star of mass $M_*$, 
+The total gravitational potential $\Phi_\mathrm{tot}=\Phi +
+\psi$ includes that from a central star of mass $M_*$ and the disk's own
+potential $\psi$. We adopt 
 \begin{align}\label{thin_disk_potential}
@@
 thin-disk approximation, appropriate for
-$|z|\ll r$. We shall adopt this approximate potential in order to
+$|z|\ll r$. We use this approximate form in order to
 obtain explicit expressions for disk equilibria (\S\ref{steady_state}). 
-The disk's own gravitational potential $\psi$ is included for
-completeness, but we will mostly neglect it unless stated
-otherwise. In such cases $\psi$ satisfies the Poisson equation 
+The disk potential $\psi$ satisfies the Poisson equation
 \begin{align}
@@
 \end{align}
+We include $\psi$ for
+completeness, but we will mostly neglect it unless stated
+otherwise.
 
-For the mixture's temperature evolution, Eq. \ref{tempeq}, $\gamma$ is
-the adiabatic index; $\mathcal{H}$ represents heating (which would
-include that due to dust-gas friction) and $\mathcal{H}_\mathrm{eff}$
+For the mixture's temperature evolution, Eq. \ref{tempeq}, reflects the gas energy equation. 
+Here $\gamma$ is 
+the adiabatic index; $\mathcal{H}$ represents gas heating and $\mathcal{H}_\mathrm{eff}$ 
 is an effective source term arising from transforming from the 
-two-fluid to one-fluid formalism \citep[see][ for
+two-fluid to one-fluid variables \citep[see][ for 
 details]{laibe14}. We also include a simple model of radiative
@@
 temperature (or pressure). By contrast, Eq. \ref{masseq}---\ref{tempeq}
-involves 3 scalars: $\rho$, $\tepsilon$, and $T$. 
+involves three scalars: $\rho$, $\tepsilon$, and $T$. 
 However, we can establish an  analogy by fixing the gas
@@
 \subsection{Locally isothermal equation of state}\label{loc_iso_eos}
-We simplify the gas energy equation by considering short cooling 
+We simplify the energy equation by considering short cooling 
 times $\tcool\to 0$, appropriate for the outer parts of an irradiated
@@
 %dust-fraction. 
-Eliminating $\tepsilon$ in favor of $P$, 
+Writing $\tepsilon=\tepsilon(P,\rho)$ using Eq. \ref{eos}
 \begin{align*}

### 2017-08-07 paper.tex
last head to toe edit before submitting, in progress

@@
 \item 
- radial dust edges are unstable if dust is vertically well-mixed;  
+ radial dust edges are unstable if the dust is vertically well-mixed;  
 \item 
@@
 \citep{youdin05a}. In this reformulation, continuity equations for dust and gas can be replaced 
-by the mass conservation of total density $\rho$ plus the evolution of the dust-to-gas ratio 
-(or dust fraction, \citealp{laibe14}).
+by the mass conservation of total density $\rho$ plus the evolution of the dust-to-gas ratio $\rhod/\rhog$ 
+(or dust fraction, $\rhod/\rho$, \citealp{laibe14}).
 
@@
 dusty gas and standard hydrodynamics is to gain a deeper understanding of the dynamics
-and stability of dusty disks.  
+and stability of dusty protoplanetary disks.  
 We can apply established methods to generalize previous stability 
 results for pure gas dynamics to dusty disks.  
-We show that strictly isothermal protoplanetary disks with
+We show that strictly isothermal disks with 
 perfectly-coupled dust are generally 
@@
 there is no reason to expect $P$ and $\rhog$ to 
-be in phase as a gas parcel oscillates between different radii. In
+be in phase as a gas parcel oscillates between different radii and adopt the corresponding local temperatures. 
+ In 
 fact, we will show that this is a fundamental property of the vertical
@@
 drift relative to the gas.   
-\emph{If} this causes $P$ to lag 
+If this causes $P$ to lag 
 $\rho$, then the positive work done would lead to growing
@@
 instabilities, we need to develop a formal analogy between dusty-gas
-and pure gas dynamics. We show this is possible in the limit of
+and pure hydrodynamics. We show this is possible in the limit of
 strong drag and a fixed gas equation of state.

### 2017-08-08 appendix.tex
last head to toe edit before submitting, in progress
final revision before submission, spell-checked

@@
 Eq. \ref{eff_energy}. Assuming axisymmetry throughout, we linearize this equation along with Eq. \ref{masseq},  
-\ref{momeq},  and \ref{poisson} gives:
+\ref{momeq},  and \ref{poisson} to give:
 
@@
   \dd\bm{v}\cdot\nabla\ln{P} = 
-\Gamma \nabla\cdot\dd\bm{v}  - \dd\bm{v}\cdot\nabla\ln{K} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
+\Gamma \nabla\cdot\dd\bm{v}  - \dd\bm{v}\cdot\nabla\ln{K} - \frac{\dd\mathcal{C}}{P},\label{lin_energy_full}\\
   -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
@@
 \end{align}  
-where the linearizd pressure force $\delta \bm{F}$ is given in
+where the linearized pressure force $\delta \bm{F}$ is given in
 Appendix \ref{lin_press}. Recall $\Delta = \delta +
@@
 The self-gravitational part of Eq. \ref{meridional} is, again
-nelgecting surface terms when integrating by parts, 
+neglecting surface terms when integrating by parts, 
 \begin{align}
@@
   &\dd F_z = - W F_z - \left[Q^\prime + Q
-    \left(\ln{\rho}\right)^\prime\right]  = - \ii\sigma \dd v_z, 
+    \left(\ln{\rho}\right)^\prime\right]  = - \ii\sigma \dd v_z + \dd\psi^\prime, 
 \end{align} 
@@
       F_z^\prime\right]W + \left(\ii k_x \p_r\ln{\rho} + k_x^2\right)Q  -
-      \ii\sigma \dd v_z^\prime.
+      \ii\sigma \dd v_z^\prime + \dd\psi^{\prime\prime}.
 \end{align}
@@
 \section{One-fluid dispersion relation for the streaming instability}\label{compressible_streaming}
-We consider an unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
+We consider an unstratified disk with $\Phi = \Phi (r)$ (by setting $z=0$ in Eq. \ref{thin_disk_potential})  
+so that $\p_zP = \p_z\rho = 
 0$. The background $\p_rP/\rho$, $\tepsilon$ are constant 
-input parameters. Then for the Epstein drag law, Eq. \ref{epstein}, we
-have $\mathcal{C}=0$ in the background state. 
+input parameters. %Then for the Epstein drag law, Eq. \ref{epstein}, we
+%have $\mathcal{C}=0$ in the background state. 
 We Fourier analyze in $r$ and $z$ so that $\p_z\to \ii k_z$ and
@@
 %  fix, but get more complicated dispersion}
-The artifical factor $\zeta = 1$  is inserted to keep track of the
+The artificial factor $\zeta = 1$  is inserted to keep track of the
 left-hand-side of the energy equation.  Setting $\zeta$ to zero is
@@
 disk ($\Omega=\kappa=\OmK$), and low frequency  modes 
-($|\sigma|\ll \OmK$), which leads to neglecting the quartic  
+($|\sigma|\ll \OmK$), which allows us to neglect the quartic  
 term. (This is also necessary to avoid spurious modes, see
-\S\ref{spurious_epi}.)  In dimensionless form:   
+\S\ref{spurious_epi}.)  In dimensionless form, the dispersion relation is 
 \begin{align}
@@
 Maximizing growth rates over $K_x$ by setting $\p \imag{(\nu)}/\p K_x
-= 0$, we find the optium wavenumber is given by
+= 0$, we find the optimum wavenumber is given by
 \begin{align}
@@
 0$. This corresponds to instability at arbitrarily small radial
-lengthscales.
+length scales.
 
@@
 
-\subsubsection{Validty of the approximations}\label{si_dust_rich_approx}
+\subsubsection{Validity of the approximations}\label{si_dust_rich_approx}
 Inserting the solutions for the most unstable mode 
@@
 \end{align}
-so Eq. \ref{max_kz} is satisified, which justifies the approximate solution for 
+so Eq. \ref{max_kz} is satisfied, which justifies the approximate solution for 
 $\mu$ in Eq. \ref{mu_explicit}.
@@
 Eq. \ref{streaming_dispersion} and \ref{streaming_incompressible} is 
-that they admit spurious unstable modes with $|\sigma|\gtrsim \Omega$
-that violates the one-fluid framework used in this
-paper. We demonstrate this below by considering the
+that they admit spurious unstable modes with $|\sigma|\gtrsim \Omega$. 
+This violates the one-fluid approximation to model dusty gas. 
+We demonstrate this below by considering the 
 incompressible dispersion relation. (We checked numerically that
@@
 in dimensionless form. Neglecting the quadratic term leads to
-stability, $\imag{(\nu)}<0$, as found by \cite{youdin05a} in their full
-two-fluid analysis.  
+stability, $\imag{(\nu)}<0$. This is consistent with a full two-fluid analysis 
+\citep{youdin05a}.
+
 % However, the real and imaginary parts of Eq. \ref{kz_zero_si}  are
@@
 by a sufficiently large $K_x$. Unlike the streaming instability,
-which is strongly surpressed when $\fdust=1/2$, these modes can still
+which is strongly suppressed when $\fdust=1/2$, these modes can still
 grow at equal dust-to-gas ratio.
@@
 These growing epicycles with $|\omega|\gtrsim \kappa$ are absent in
-full two-fluid models \citep{youdin05a}. The discrepency lies in the
+full two-fluid models \citep{youdin05a}. The discrepancy lies in the
 fact that the one-fluid equations, to first order in $\taus$, are
@@
 The (spurious) overstable dusty epicycles' growth
-rates are weakly depedent on $K_z$ and are well approximated by
+rates are weakly dependent on $K_z$ and are well approximated by
 that in the $K_z=0$ limit, Eq. \ref{epi_grow}. 
@@
 %in the above notation. 
-For the case considerd in Fig. \ref{dusty_growth1} we find
+For the case considered in Fig. \ref{dusty_growth1} we find
 $s\simeq0.067\OmK$, as observed. By contrast, the streaming
@@
 
-Caution must be taken if the first-order one-fluid equations are used
+Care must be taken if the first-order one-fluid equations are used
 to simulate dusty gas. 
@@
 Thus simulations based on the first-order one-fluid equations should
-be set up to surpress these spurious epicycles. This might be
+be set up to suppress these spurious epicycles. This might be
 achieved, for example, through physical or numerical viscosity to

### 2017-08-08 linear.tex
final revision before submission, spell-checked

@@
 based on explicit descriptions of the equilibrium state. We first
-specify explicit disk structures to analyze. We show 
-that sharp radial edges in the dust component are unstable. We then
+specify the disk structures to be analyzed. We show  
+that sharp radial edges in the dust-to-gas ratio can render the disk unstable. We then 
 revisit two well-known instabilities in dusty gas using the
-hydrodynamic framework developed thusfar, namly secular graviational
-instabilities and the streaming instability.
+hydrodynamic framework developed thus far, namely secular gravitational (SGI) 
+instabilities and the streaming instability (SI).
 
@@
 proportional to $q$ corresponds to vertical shear due to the 
-radial temperature gradient, and survives in the dust-free limit.
+radial temperature gradient, which survives in the dust-free limit.
 
@@
 \subsubsection{Dusty vertical buoyancy}\label{vbuoyancy}
-For the above equilibria the vertical buoyancy frequency is given
+For the above equilibrium the vertical buoyancy frequency is given
 explicitly as 
@@
 in \S\ref{iso_perfect} assuming strictly isothermal gas. As discussed
-there, the first criterion in generally satisifed. Thus we only
+there, the first criterion is generally satisfied. Thus we only
 consider the second condition, Eq. \ref{dusty_solberg2}.
@@
 provided that $H_\epsilon/\Hg$ is not large (e.g. some dust settling
-has occured). This means that typical dusty PPDs are stable to
+has occurred). This means that typical dusty PPDs are stable to
 axisymmetric perturbations, even for arbitrarily thin
@@
 To violate Eq. \ref{stability_est2} and obtain instability, notice the 
-LHS is a quadratic in $\p_r\epsilon$. Thus instability is possible for
+left hand side is a quadratic in $\p_r\epsilon$. Thus instability is possible for
 sufficiently large (in magnitude) radial gradients  in the dust-to-gas
@@
 requires $L_\epsilon\ll O(0.1\Hg)$, i.e. the dust-to-gas ratio must 
-vary on an extremely short lengthscale. This might be achieved, for
+vary on an extremely short radial lengthscale. This might be achieved, for
 example, at sharp edges associated with gaps opened by giant planets.
@@
 
-\subsection{Radially localized linear problem}
+\subsection{Radially local problem}
 We now specialize further and compute explicit solutions to the linear
@@
 equation in place of the dust-fraction
-(Eq. \ref{masseq}---\ref{momeq}, Eq. \ref{eff_energy}) are then:
+(Eq. \ref{masseq}, \ref{momeq}, \ref{poisson}, \ref{eff_energy}) are then:
 
@@
 discuss SGI in the next
-section. The linearized dust-duffusion function 
+section. The linearized dust-diffusion function 
 $\dd\mathcal{C}$ is given in  Appendix \ref{lin_dust}. 
@@
 Eq. \ref{lin_mass}---\ref{lin_sg} is a set of ordinary
-differntial equations in $z$. All coefficients and amplitudes are
+differential equations in $z$. All coefficients and amplitudes are
 evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
 is retained. We next discuss solutions to these equations.  We first
-show that the above equations yield the secular
-gravitational instability and streaming instability in the strong
+show that the above equations yield the SGI and SI in the strong
 drag regime in \S\ref{sgi} and \S\ref{si}, respectively. We then
-consider 3D, stratified disks in \S\ref{results} to study the dusty
-vertical shear instability.
+consider 3D, stratified disks in \S\ref{results} to study how the addition of  dust 
+affects the vertical shear instability.
 
@@
 \subsubsection{Secular gravitational instability}\label{sgi}
-We first consider a razor-thin, self-gravitating disk so that $\rho =
-\Sigma\delta(z)$. Here $\delta$ is the Dirac delta function and $\Sigma$ is
-the total surface density. Now $\epsilon$ and 
-$\tepsilon$ refers to the global dust-to-gas ratio and dust 
-fraction, respectively. The background disk is
+Consider a razor-thin, self-gravitating disk so that $\rho =
+\Sigma\delta(z)$, where $\delta$ is the Dirac delta function and $\Sigma$ is
+the total surface density. Here $\epsilon = \Sigma_\mathrm{d}/\Sigma_\mathrm{g}$ and 
+$\tepsilon=\Sigma_\mathrm{d}/\Sigma$, where $\Sigma_\mathrm{d,g}$ are the dust and gas surface densities. 
+The background disk is 
 uniform and we neglect the vertical dimension   
@@
 In Table \ref{si_compare} we also calculate the phase difference
-between the Lagragian pressure perturbation and density perturbations
+between the Lagrangian pressure perturbation and density perturbations
 as    
@@
 In Fig. \ref{si_compare_fig} we calculate the most unstable SI 
-mode as a function of $\taus$ at fixed $K_z=30$ and $\epsilon =3$. We
+mode as a function of $\taus$ at fixed $K_z=30$ and $\epsilon =3$. Growth rates are maximized over $K_x$. 
+We 
 compare results between the full, two-fluid linear analysis and the
-one-fluid framework. We also include an analytic model, described in Appendix 
+one-fluid framework. We also include an analytic model, developed in Appendix 
 \ref{si_dust_rich}, based on the one-fluid dispersion relation with 
@@
 lag drops, along with the growth rate. This suggest that a
-non-vanishing phase lag is necessary for instability. However, the
+non-vanishing phase lag is indeed necessary for instability. However, the
 magnitude of the phase lag does not correlate with growth rates. In

### 2017-08-08 model.tex
final revision before submission, spell-checked

@@
 \citep{jacquet11}. This equation reflects the well-known effect of 
-particle drift towards pressure maxima \citep{weidenschilling77}.
+particle drift towards pressure maximum  \citep{weidenschilling77}.
 
@@
 equations]{laibe14,price15}, where $D/Dt \equiv \p_t +
-\bm{v}\cdot\nabla$ is the Lagragian 
+\bm{v}\cdot\nabla$ is the Lagrangian 
 derivative \emph{following the mixture}'s center-of-mass velocity $\bm{v}$. 
@@
 
-For the mixture's temperature evolution, Eq. \ref{tempeq}, reflects the gas energy equation. 
-Here $\gamma$ is 
-the adiabatic index; $\mathcal{H}$ represents gas heating and $\mathcal{H}_\mathrm{eff}$ 
-is an effective source term arising from transforming from the 
+For the mixture's temperature evolution, Eq. \ref{tempeq}, 
+$\gamma$ is 
+the gas adiabatic index; $\mathcal{H}$ represents heating and $\mathcal{H}_\mathrm{eff}$ 
+is an effective source term arising from transforming the gas energy equation (Eq. \ref{real_pressure}) 
+from the 
 two-fluid to one-fluid variables \citep[see][ for 
@@
 \subsection{Locally isothermal equation of state}\label{loc_iso_eos}
-We simplify the energy equation by considering short cooling 
-times $\tcool\to 0$, appropriate for the outer parts of an irradiated
+We consider the limit of short cooling 
+times, $\tcool\to 0$, appropriate for the outer parts of an irradiated
 protoplanetary disk \citep{chiang97,lin15}. Then the disk temperature
@@
 Since $\tepsilon$ typically decrease away from the midplane, we expect
-vertically isothermal dusty disks to behave as if the temperture
+vertically isothermal dusty disks to behave as if the temperature
 \emph{increased} away from $z=0$.
@@
 Although we have deleted the true energy
-equation (Eq. \ref{tempeq}) by fixing a locally isothermal
-equation of state (Eq. \ref{eos}), we show that the mixture 
-nevertheless obeys an effective energy equation. This is
+equation by fixing a locally isothermal
+equation of state, we show that the mixture 
+nevertheless obeys an effective energy evolution equation. This is
 because advection of the dust-fraction, described by
@@
 %dust-fraction. 
-Writing $\tepsilon=\tepsilon(P,\rho)$ using Eq. \ref{eos}
+The equation of state, Eq. \ref{eos}, implies 
 \begin{align*}
-  \tepsilon = 1 - \frac{P}{c_s^2(r,z)\rho}, 
+  \tepsilon = 1 - \frac{P}{c_s^2(r,z)\rho}.  
 \end{align*}
-then Eq. \ref{dusteq} becomes
+Then Eq. \ref{dusteq} becomes
 \begin{align}
@@
 The specific entropy of an ideal gas is 
-$S = C_P\ln{\left(P^{1/\gamma}/\rhog\right)}$, where $C_P$ is the heat 
+$S~=~C_P\ln{\left(P^{1/\gamma}/\rhog\right)}$, where $C_P$ is the heat 
 capacity at constant pressure.
@@
 \begin{align}
-   \seff \equiv \ln \frac{P}{\rho} = \ln{\left[c_s^2(1-\tepsilon)\right]}.  
+   \seff \equiv \ln \frac{P}{\rho} = \ln{\left[c_s^2(1-\tepsilon)\right]},  
 \end{align} 
-The constant heat capacity has been absorbed into $\seff$. 
+where the constant heat capacity has been absorbed into $\seff$. 
 %This follows naturally by combining the effective 
 %energy Eq. \ref{eff_energy} and Eq. \ref{masseq}. 
-Then 
+Then combining Eq. \ref{eff_energy} and \ref{masseq} gives  
 \begin{align*}
@@
 Protoplanetary disks have $\p_z\rhog<0$, thus stability against vertical 
-convection ($N_z^2>0$) requires $\p_z\tepsilon <0 $, i.e. the dust density drops faster 
+convection ($N_z^2>0$) requires $\p_z\tepsilon <0 $, i.e. the dust density should drop faster 
 than the gas density away from the midplane. 
@@
 models such as the Minimum Mass Solar Nebula, $\tstop$ decreases for
-smaller particles and/or at smaller radii \citep{youdin11}. However,
+smaller particles and/or towards smaller radii \citep{youdin11}. However,
 $t_\mathrm{cool}$ is generally only small in the outer disk
@@
 (Eq. \ref{eos}). An explicit solution is presented in 
-\S\ref{steady_state}, when we analyse the stability of
+\S\ref{steady_state}, when we analyze the stability of
 protoplanetary disks.
@@
 $\mathcal{C} = 0$ in equilibrium.  Our analysis is also a good approximation when 
-$\tstop$ is sufficiently small such that the evolution of the background disk (e.g. dust-settling timescales) 
-occur on much longer timescales than the dust dynamics of interest (e.g. instability growth timescales).
+$\tstop$ is sufficiently small such that the evolution of the background disk (e.g. dust-settling) 
+occur on much longer timescales than the instability growth timescales.
 
@@
 
-In Appendix \ref{var_prin} we derive linearize Eqs. \ref{masseq}, \ref{momeq}, and 
+In Appendix \ref{var_prin} we linearize Eqs. \ref{masseq}, \ref{momeq}, and 
 Eq. \ref{eff_energy} to derive the following integral relation: 
@@
 When $c_s^2$ is a constant and $\tstop=0$ (so that $\mathcal{C} = 0$), the dusty-gas equations are
-exactly equivalent to that for adiabatic hydrodyamics with unit
+exactly equivalent to that for adiabatic hydrodynamics with unit
 adiabatic 
@@
 adiabatically because the dust fraction $\tepsilon$ is advected with 
-the gas. This is similar to entropy being conserved following an
-adiabatic gas when there is no heating or cooling.
+the gas. This is similar to entropy conservation following an 
+adiabatic gas without heating or cooling.
 
@@
 Then the condition for axisymmtric \emph{stability}, $\sigma^2>0$, 
-is met if the integrand cofficients satisfy $A+D>0$ and $AD-B^2>0$
+is met if the integrand coefficients satisfy $A+D>0$ and $AD-B^2>0$
 \citep[][Section 11.6]{ogilvie16}, which becomes 
@@
 disks if the dust is vertically well-mixed but radially-stratified,  
-such that $\p_z\tepsilon = 0$ but $\p_r\tepsilon\neq0$). In this case the
+such that $\p_z\tepsilon = 0$ but $\p_r\tepsilon\neq0$. In this case the
 left-hand-side of Eq. \ref{dusty_solberg2} becomes
@@
 vertical shear due to the radial gradient in the dust-fraction. 
-This situation is identical to the pure gas, adiabtic simulations of \cite{nelson13}
+This situation is identical to the pure gas, adiabatic simulations of \cite{nelson13}
 where the gas entropy is vertically uniform but has a radial
@@
 
-It is well-known that such correlations can lead to pulsational
+It is well-known that such correlations may lead to pulsational
 instabilities in stars \citep{cox67}. We thus interpret  
@@
 where $\upsilon=1/\rho$ is the specific volume of the mixture, $D/Dt$
-is the Lagragian derivative, and the 
+is the Lagrangian derivative, and the 
 spatial integral is taken over the total mass $M$ of the mixture and
@@
 %mixture; and 
- the Lagragian perturbation $\Delta$ of a variable $f$ is 
-$\Delta f = \delta f + \bm{\xi}\cdot\nabla f$ and $\bm{\xi}$ is the
-Lagragian displacement ($    \xi_{x,z} =  \ii \dd v_{x,z}/\sigma$).  
+ the Lagrangian perturbation $\Delta$ of a variable $X$ is 
+$\Delta X = \delta X + \bm{\xi}\cdot\nabla X$ 
+and $\bm{\xi}$ is the
+Lagrangian displacement ($    \xi_{x,z} =  \ii \dd v_{x,z}/\sigma$).  
 Eq. \ref{work_real} show that a phase difference between gas pressure and
@@
 
-Now, quite generally, Eq. \ref{lin_mass_full}---\ref{lin_energy_full} 
+Now, Eq. \ref{lin_mass_full}---\ref{lin_energy_full} 
 imply the integrand of the numerator in Eq. \ref{thermal_instability} 
@@
 Eq. \ref{pdv}. Eq. \ref{growth_work} states that if the average work
-done is positive during an oscillation, $\mathcal{W}>0$, then 
+done is positive during an oscillation, $\mathcal{W}>0$, 
 then its amplitude would actually grow ($s>0$).  %and not be periodic
@@
 requires $\imag\left(\Delta
-P\right)<0$.  This corresponds to pressure perturbations \emph{lagging}
+P\right)<0$.  This corresponds to Lagrangian pressure perturbations \emph{lagging}
 behind that in density, see Fig. \ref{lag_cartoon}
@@
     pressure, which is directly proportional to the gas density, to
-    lag behind that in dust density. 
+    lag behind that in dust density. The eigenvectors rotate anti-clockwise for $\omega>0$. 
     \label{lag_cartoon}
@@
 
-The discussion above applies to any single fluid with a pressure and
+The above discussion applies to any single fluid with a pressure and
 density. In the case of interest --- dusty gas --- the work done 
@@
 A parcel of the strictly isothermal dusty-gas mixture does 
-postive work if  
+positive work if  
 \begin{align*}
@@
 \end{align*}
-meaning that gas follows dust (Fig. \ref{lag_cartoon}). Instabilities
+meaning that gas follows dust (Fig. \ref{lag_cartoon}). Overstabilities 
 are thus not possible if the gas does not respond to dust 
@@
 This thermodynamic interpretation does not explain 
-\emph{why} drag forces causes gas pressure to lag behind the dust
+why drag forces causes gas pressure to lag behind the dust
 density, but shows 

### 2017-08-08 paper.tex
final revision before submission, spell-checked

@@
  perfectly coupled, thin dust layers cannot cause axisymmetric 
- instabilties; 
+ instabilities; 
 \item 
@@
 \item 
-  the streaming instability neccessarily involves a gas pressure
+  the streaming instability necessarily involves a gas pressure
   response 
@@
 disks directly simulate gas dynamics coupled to explicit
-Lagragian dust particles
+Lagrangian dust particles
 \citep{johansen2006, nelson10,bai10,yang14,zhu14,gibbons15,simon16,baruteau16}. 
@@
 \caption{The central concept of this work. A mixture of small particles
-  coupled to isothermal gas behave similarly to an adiabatic, pure gas
+  imperfectly coupled to isothermal gas behave similarly to an adiabatic, pure gas
   subject to cooling.  This
@@
 amplitudes grow: the work done allows the system to `overshoot' beyond 
-the amplitude of preceeding oscillations.
+the amplitude of preceding oscillations.
 
@@
 From $A$ to $B$, a fluid parcel is expanding to return to 
-equilbrium, but pressure is still 
+equilibrium, but pressure is still 
 increasing. This over-compensation will cause the parcel to expand
@@
     dust-drag causes  
-    pressure to lag behind density, which results in a clockwise path in
+    pressure (due to gas only) to lag behind total density. This results in a clockwise path in
     the P-V plane, implying positive work done by the fluid, which 
@@
 and there is no work done. However, for finite dust-gas drag, $\rhod$ and
-$\rhog$ are not neccessarily in phase, because dust particles can
+$\rhog$ are not necessarily in phase, because dust particles can
 drift relative to the gas.   
@@
 \input{summary}
-
-%{\bf acknowledgements. s. inutsuka for asking about kz=0 limit}
+\acknowledgements
+We thank P. Loren-Aguilar for initial discussions that motivated this study. We also thank 
+C. Baruteau, R. Dong, J. Fung, S. Inutsuka, W. Lyra, and M. Pessah 
+for comments during the course of this project. This work is supported by the Theoretical Institute for Advanced Research in Astrophysics 
+(TIARA) based in Academia Sinica Institute of Astronomy and Astrophysics (ASIAA) and NASA Astrophysics Theory Program grant NNX17AK59G.
 

### 2017-08-08 result.tex
final revision before submission, spell-checked

@@
 \section{
-  Dusty vertical shear instability}\label{results}
+  Vertical shear instability with dust}\label{results}
 
@@
 non-self-gravitating disks, assuming perfectly coupled dust. 
-We formally take $\tstop=0$ so the equilibria defined in \S\ref{eqm} 
+We formally take $\tstop=0$ so the equilibrium defined in \S\ref{eqm} 
 are exact steady states. In reality,   
@@
     dust-to-gas ratio $\epsilon$ is approximately constant with
-    height. 
+    height. Vertical buoyancy here is due to dust-loading. 
     \label{compare_vshear_fixHd}
@@
 rates. Notably, surface modes, which are typically fastest growing in
-the dust-free case, are surpressed in dusty disks for $\epsilon_0\geq
+the dust-free case, are suppressed in dusty disks for $\epsilon_0\geq
 0.1$ (i.e. $\epsilon_0> \hgas$).   
@@
 $\Hd=0.99\Hg$. Fig. \ref{compare_vshear_fixZ} compares the vertical
-vhear rate and buoyancy frequency. For $|z|\gtrsim 0.4\Hg$ the two
+shear  rate and buoyancy frequency. For $|z|\gtrsim 0.4\Hg$ the two
 disks have the same profile with vertical shear dominating over
@@
 since the vertical shear profile is similar throughout most of the
-disk. However, meridional motions are surpressed near the midplane of
+disk. However, meridional motions are suppressed near the midplane of
 the $\Hd=0.1\Hg$ disk, as expected from the larger buoyancy frequency
@@
 
-To connect with previous studies, here  we also use 
+To connect with similar studies, here  we also use 
 the Richardson number $\rich \equiv N_z^2/\left(r\p_z\Omega\right)^2$
@@
 by very thin dust layers \citep{chiang08, lee10}. We show that \emph{axisymmetric}
-instabilities never develop, however small $\rich$.
+instabilities never develop in radially uniform disks, however small $\rich$.
 
@@
   \caption{Maximum VSI growth rate for ultra-thin dust layers $\Hd 
-    \leq 0.01\Hg$. Vertical shear is dominated by that due to
-    dust-settling, but axisymmetric instability is associated with the
-    radial temperature gradient $q$. Here, $\rich_0$ is the minium
-    Richardson number in the domain in the limit $q\to0$. 
+    \leq 0.01\Hg$ (with radially uniform dust-to-gas  ratio). Vertical shear is dominated by that due to 
+    vertical dust stratification, but axisymmetric instability is still 
+    associated with the radial temperature gradient $q$. 
+ Here, $\rich_0$ is the minimum 
+    Richardson number in the domain in the limit $q\to0$. The disk is stable to axisymmetric perturbations 
+    in the strictly isothermal limit, regardless of the dust layer thickness. 
     \label{ultra_thin}
@@
 if $d\epsilon/dr\neq0$.  
-To demonstate this numerically we consider the
+To demonstrate this numerically we consider the
 same disk as above but with $q=0$ and choose $H_\mathrm{d}$ such
@@
 Fig. \ref{vert_mixed_modes}. In order to obtain appreciable growth   
-rates, we needed to consider super-Solar metalicity $Z=0.03$ and high
+rates, we consider super-Solar metalicity $Z=0.03$ and high
 radial wavenumber $k_x\Hg=1800$. As expected from the discussion in

### 2017-08-08 summary.tex
misc home edits
final revision before submission, spell-checked

@@
   \frac{D P}{D t} &= - \Gamma P\nabla\cdot\bm{v}  + P \bm{v}\cdot\nabla
-  \ln{K} + \mathcal{C},\\
-  \mathcal{C}& = \frac{\Gamma P}{\rhog}\nabla\cdot\left(\tepsilon\tstop\nabla
+  \ln{K} + \frac{\Gamma P}{\rhog}\nabla\cdot\left(\tepsilon\tstop\nabla
   P\right). 
+%+ \mathcal{C},\\
+%  \mathcal{C}& = \frac{\Gamma P}{\rhog}\nabla\cdot\left(\tepsilon\tstop\nabla
+%  P\right).
 \end{align}
@@
 disk temperature \citep[][]{thompson88,shi13}. For typical dust-loading 
-$\epsilon, Z\ll1$, this effect is unimportant. However, if $\epsilon$ is
+$\tepsilon, Z\ll1$, this effect is unimportant. However, if $\epsilon$ is
 large (e.g. due to dust settling) then the reduced temperature
-$\widetilde{T}$ may be lowered to enable instability.
+$\widetilde{T}~=~T(1 - \tepsilon)$ may be lowered to enable instability.
 
@@
 but the back-reaction onto the gas may try to flatten the pressure
-bump \citep{taki16}.
+bump \citep{taki16}, thus enabling a quasi-steady state.
 
@@
 %\end{align} 
-where $\mathcal{S}_\mathrm{g} = P/\Sigma_\mathrm{g}^\gamma$. Here $P$
+where $\mathcal{S}_\mathrm{g} = P/\Sigma_\mathrm{g}^\gamma$ is essentially
+the gas entropy. Here $P$
 should be interpreted as the vertically integrated pressure. The
@@
 Since the necessary physical ingredient for the ZVI is vertical
-buoyancy, it requires an almost purely adiabatic gas ($\tcool\OmK\gg
+buoyancy, it requires an almost purely adiabatic gas with long cooling times ($\tcool\OmK\gg
 1$). This limits its applicability in typical protoplanetary disks to 
@@
 the required adiabatic conditions for ZVI can be realized through
-dust-loading, and thus produce `dusty ZVI'. This may allow the zombie
+dust-loading (specifically a vertical gradient in the dust-to-gas ratio), 
+and thus produce `dusty ZVI'. This may allow the zombie 
 vortices to develop $\gtrsim 1$AU in PPDs.
@@
 In this paper we develop the analogy between isothermal 
-dusty-gas and adiabatic, pure gas. The correspondence arises
+dusty-gas and pure ideal gas. The correspondence arises
 because drag forces reduce the relative velocity between gas and
-dust. In the limit of perfect dust-gas coupling, $\tstop \to 0$,  
+dust. In the limit of perfect dust-gas coupling, with stopping time $\tstop \to 0$,  
  dust is entrained in 
 the gas. Then the dust-to-gas ratio $\rhod/\rhog$ is conserved
-following the flow. This property is analogous to entropy being conserved
-following a adiabatic, pure gas.
+following the flow. This property is analogous to entropy conservation 
+following an adiabatic, pure gas.
 
@@
 We explicitly show that for a fixed gas equation of state, the  
-evolutionary equation for the dust-to-gas ratio $\rhod/\rhog$ may be replaced by an 
+evolutionary equation for $\rhod/\rhog$ may be replaced by an 
 effective energy equation. This leads to a 
@@
 \begin{align*}
-  \seff  = \ln{\left(\frac{c_s^2\rhog}{\rhog+\rhod}\right)}.  
+  \seff  = \ln{\left(\frac{c_s^2\rhog}{\rhog+\rhod}\right)},  
 \end{align*}
@@
 %dusty gas. 
+which implies that a non-uniform dust-to-gas ratio induces buoyancy forces.  
 The effect of finite dust-gas friction appears as an energy
@@
 We  apply our thermodynamic framework to interpret
-the streaming instability \citep[SI, ][]{youdin05a,jacquet11} and to generalize the vertical shear
+the streaming instability \citep[SI, ][]{youdin05a,jacquet11} and to generalize the gaseous vertical shear
 instability \citep[VSI, ][]{nelson13,lin15} to dusty disks. We 
 explicitly show in SI the evolution of gas 
-pressure lags behind dust density. This is in fact a general property of 
-instabilities driven by dust-gas drag. 
+pressure lags behind dust density. In fact, this is a general property of 
+overstabilities driven by dust-gas drag. 
 It takes a finite time for the gas to respond to
@@
 
+The dusty/adiabatic gas equivalence also offers a simple way to simulate dusty protoplanetary disks 
+using purely hydrodynamic codes. All that is required is a re-interpretation of the fluid variables and additional 
+source terms in the usual energy equation. The latter is already available in many public codes. In a follow-up work
+we will apply this approach to study dusty disk-planet interaction.
+
 %In particular, we emphasize the 

### 2017-09-11 paper.tex
revising in response to report

@@
 \citep[SI,][]{youdin05a,youdin07b,johansen07}, secular gravitational
-instabilities \citep[SGI,][]{ward00,youdin11,michikoshi12,takahashi14,latter17} and 
+instabilities \citep[SGI,][]{ward00,youdin11,michikoshi12,takahashi14,taka16,latter17} and 
 Kelvin-Helmholtz instabilities
@@
 This paper is organized as follows.  
-In \S\ref{grow_osc} 
-we describe a key insight of our study, that pressure (i.e.\ ``PdV") work explains the basis of several instabilities, including SI and VSI, in terms of
-pressure-density phase lags.    
+In \S\ref{grow_osc} we describe a key insight of our study, that
+pressure (i.e.\ ``PdV")  work {\bf can} explain the basis of several
+instabilities, including SI and VSI, in terms of 
+pressure-density phase lags. {\bf This motivates us to seek an 
+  analogy between dust-gas mixtures and pure gas.}    
 We develop our formalism in \S\ref{setup} by transforming the
@@
 define the effective entropy and buoyancy of our model fluid. 
-%We derive equilibrium solutions in \S\ref{eqm}, which we use in our
-%explicit stability calculations.
 In \S\ref{limits} we discuss general stability properties of dusty
@@
 \section{Growing oscillations by doing work}\label{grow_osc}
+{\bf
+In this section we provide physical arguments to highlight the
+similarity between single phase fluids and dust-gas mixtures. It is
+well-known that in the limit of perfect coupling the addition of dust
+increases the gas inertia but not pressure. One can then regard the
+dusty gas as a single mixture with a reduced sound-speed. Here, we
+argue that when dust-gas coupling is imperfect, there exists another
+similarity with pure gas related the phase of pressure and density
+evolution. Thus it is still possible to regard
+partially-coupled dusty-gas as a single fluid. The mathematical
+description of this section is given in \S\ref{dust_work}.   
+}
+%dust_work
+
 A general result in fluid dynamics is that work is done whenever
@@
 We thank P. Loren-Aguilar for initial discussions that motivated this study. We also thank 
-C. Baruteau, R. Dong, J. Fung, S. Inutsuka, W. Lyra, and M. Pessah 
-for comments during the course of this project. This work is supported by the Theoretical Institute for Advanced Research in Astrophysics 
-(TIARA) based in Academia Sinica Institute of Astronomy and Astrophysics (ASIAA) and NASA Astrophysics Theory Program grant NNX17AK59G.
+C. Baruteau, R. Dong, J. Fung, S. Inutsuka, W. Lyra, M. Pessah, and O. Umurhan 
+for comments during the course of this project. This work is supported
+by the Theoretical Institute for Advanced Research in Astrophysics  
+(TIARA) based in Academia Sinica Institute of Astronomy and
+Astrophysics (ASIAA), NASA Astrophysics Theory Program grant
+NNX17AK59G, and  
+the Steward Theory Fellowship at the University of Arizona.
 

### 2017-09-11 result.tex
revising in response to report

@@
 \end{align*}
-Here we fix $Z=0.01$, $\Hd=0.8\Hg$, and $k_x\Hg=30$.
+Here we fix $Z=0.01$, $\Hd=0.8\Hg$.
 
@@
 dust-to-gas ratio increases the magnitude of the vertical shear rate
-away from the midplane (see also Eq. \ref{vshear2}). Thus we find
-higher VSI growth rates, as shown in Fig. \ref{vsi_dust_varHd}. 
-Surfaces modes are more effectively enhanced by the additional vertical 
-shear induced by $\p_r\epsilon$. However, all growth rates remain
+away from the midplane (see also Eq. \ref{vshear2}). Thus we {\bf
+  typically} find higher VSI growth rates, as shown in
+Fig. \ref{vsi_dust_varHd}. 
+Surfaces modes, {\bf appearing at high $k_x$}, are more effectively
+enhanced by the additional vertical shear induced by
+$\p_r\epsilon$. {\bf We find that low frequency body modes with small
+  $k_x$ are more affected by radial variations in the 
+  dust-to-gas ratio.}
+
+However, all growth rates remain
 $O(\hgas\OmK)$. Importantly, the increase in the growth rate of the
@@
   \caption{Unstable VSI modes in the disk models of
-    Fig. \protect\ref{compare_vshear_varHd}.  Black (blue) diamonds
+    Fig. \protect\ref{compare_vshear_varHd}. {\bf Diamonds (asterisks) 
     are mode frequencies for a disk with radially uniform (varying)
-    dust-to-gas ratio. 
+    dust-to-gas ratio. The radial wavenumbers are $k_xH_g = 10$
+    (green),$k_xH_g = 30$ (black) and $k_xH_g = 100$ (orange). }
     \label{vsi_dust_varHd}
@@
 if $d\epsilon/dr\neq0$.   
-To demonstrate this numerically we consider the
-same disk as above but with $q=0$ and choose $H_\mathrm{d}$ such
-that $H_\epsilon=10^3\Hg$. Then vertical shear $\p_z\Omega$ is 
+{\bf
+To demonstrate this numerically we set $q=0$ and $H_\mathrm{d}$ such
+that $H_\epsilon=10^3\Hg$. We let $\rhod/\rhog \propto r^{-d}$
+with different power indices $d$. The metalicity is fixed to $Z=0.03$.  
+In these cases vertical shear $\p_z\Omega$ is 
 attributed to the radially-varying dust-to-gas ratio (see
 Eq. \ref{vshear2}).    
+}
 
-We show unstable modes for this disk in
-Fig. \ref{vert_mixed_modes}. In order to obtain appreciable growth   
-rates, we consider super-Solar metalicity $Z=0.03$ and high
-radial wavenumber $k_x\Hg=1800$. As expected from the discussion in
+{\bf We show unstable modes in 
+Fig. \ref{vert_mixed_modes}.} 
+%In order to obtain appreciable growth   
+%rates, we consider super-Solar metalicity $Z=0.03$ and high
+%radial wavenumber $k_x\Hg=1800$. 
+As expected from the discussion in
 \S\ref{iso_perfect}, the disk admits purely growing modes with
@@
 with classic VSI discussed above.
+
+{\bf We find in smooth disks large 
+  $k_x$ is needed for appreciable growth rates, e.g. $k_xH_g=1800$
+  with $\rhod/\rhog\propto r^{-1}$. 
+  Local shearing box simulations \citep[e.g.][]{bai10b,yang16} may 
+  be required to study the non-linear evolution of such short 
+  wavelengths. Alternatively, as shown in Fig. \ref{vert_mixed_modes},
+  a rapidly varying dust-to-gas ratio permits dynamical instability at
+  longer radial wavelengths. 
+}
 
@@
 \end{figure}
+
+%{\bf
+%  \subsection{Implications for protoplanetary disks}
+%dust settling, dust-induced buoyancy 
+%The above calculations 
+%radial d/g -> radial temp profile 
+%}

### 2017-09-11 summary.tex
revising in response to report

@@
 \section{Discussion}\label{discussion}
+
+%pros and cons 
+%energy analogy not restricted to term vel approx 
+%spurious modes
 
@@
 vertical buoyancy dominates over vertical shear, consistent with our
-previous study \citep{lin15}. Since the dust-induced buoyancy forces
+previous study \citep{lin15}. 
+Since the dust-induced buoyancy forces
 increase away from the midplane, we find dust-loading can stabilize
 `surface modes'  of the VSI, that would otherwise have the largest
-growth rates.
+growth rates. {\bf We also showed that radial variations in
+  $\rhod/\rhog$ translates to an effectively non-uniform radial temperature
+  profile, and the resulting vertical shear can lead to instability,
+  even if the true sound-speed is constant. 
+}
+
+%implications for realistic disks
+{\bf In a realistic disk, dust particles settle on a timescale $
+  t_\mathrm{settle} \sim 1/\tstop\OmK^2$ \citep{takeuchi02}, compared 
+  with typical VSI growth timescales, $t_\mathrm{grow}\sim 
+  1/h_\mathrm{g}\OmK$. This suggest that particles with $\tstop\OmK\lesssim 
+  h_g$ cannot settle against the VSI. On the 
+  other hand, larger particles with $\tstop\OmK\gtrsim h_g$ should
+  settle to form a dusty midplane. In fact, our calculations suggest that
+  settling would stabilize the disk against VSI, and allow
+  further settling. The result may be a quiet, dusty midplane \citep[unless non-axisymmetric
+  instabilities develop][]{chiang08} with VSI-turbulent gaseous
+  atmospheres. 
+}
 

### 2017-09-11 summary.tex
revising, new section 7.1

@@
 
+{\bf
+
 %pros and cons 
@@
 %spurious modes
+
+\subsection{Applications and limitations}
+
+The main application we envision for the dusty/adiabatic gas analogy 
+to develop physical interpretations of dust-gas instabilities; and to find 
+dusty analogs of pure gas instabilities in protoplanetary disks, 
+which is discussed in \S\ref{dust_analogs}. One can also exploit the similarity to adapt existing
+hydrodynamic codes to simulate dusty protoplanetary disks (see \S\ref{dust_sims}).
+
+It is important to keep in mind the assumptions 
+used to develop our thermodynamic model of dusty gas. 
+The terminal velocity approximation, $\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} = \tstop \nabla P /\rhog$, was employed from the outset. 
+%It assumes  
+%the dust-gas relative drift $\bm{v}_mathrm{d} - \bm{v}_\mathrm{g} = \tstop \nabla P /\rhog$ 
+This is applicable to small particles with 
+$\tstop\OmK\ll 1$, but the precise critical value is problem-dependent. For the linear streaming instability,
+even $\tstop\OmK \sim 0.1$ can be captured by simplified one-fluid equations 
+(Fig. \ref{si_compare_fig}). On the other hand, the simplified equations contain spurious modes that
+may overwhelm physical instabilities in certain situations (see \S\ref{spurious_epi}).  %Ultimately, one must check against the full equations.
+
+However, our thermodynamic interpretation of dust-gas drift does \emph{not} in fact depend on the terminal velocity approximation.  
+Once the gas equation of state is fixed and the true energy equation deleted, 
+the dust continuity equation can be converted to a new effective energy equation. 
+%This source term reflects  
+%a heat flux proportional to $\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g}$, which must generally be solved via a separate 
+%evolutionary equation \citep{laibe14}. Nevertheless, this does not change the interpretation of dusty gas as a single ideal fluid subject to cooling.  
+As an example, for strictly isothermal dusty gas we obtain   
+\begin{align}
+\frac{DP}{Dt} = - P\nabla\cdot\bm{v} + c_s^2\nabla\cdot\left[\fdust\left(1-\fdust\right)\rho\left(\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g}\right)\right], \label{no_term_vel} 
+\end{align}
+\citep[see also][]{laibe14}. This equation does not use the terminal velocity approximation (which would give Eq. \ref{eff_energy}), but 
+ we see that dust-gas relative drift can be regarded as a heat flux within the mixture. Calculating this heat flux generally requires solving the evolutionary equation for 
+$\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g}$. Nevertheless, one can still consider the dusty-gas as a single fluid subject to cooling. 
+}
 
@@
 
-\subsection{Dusty analogs of other gaseous instabilities}
+\subsection{Dusty analogs of other gaseous instabilities} \label{dust_analogs}
 
@@
 
-\subsection{Implications for numerical simulations}
+\subsection{Implications for numerical simulations}\label{dust_sims}
 The dusty/adiabatic gas equivalence means that pure
@@
 `surface modes'  of the VSI, that would otherwise have the largest
-growth rates. {\bf We also showed that radial variations in
-  $\rhod/\rhog$ translates to an effectively non-uniform radial temperature
+growth rates. {\bf We also show that radial variations in
+  $\rhod/\rhog$ effectively translates to a non-uniform radial temperature
   profile, and the resulting vertical shear can lead to instability,

### 2017-09-12 response.txt
first draft of revision 1 and response to referee

We thank the referee for a prompt report, which helped to improve the
clarity of the manuscript. We have made revisions in bold, and further
explain them below.

Major points:

* The section 2 is interesting as it relates the work done by
  pressure, when not in phase with density, to the growth of
  oscillations. Although I understand that it is a link between the
  pure gas context and the dusty flow case (because the drag force
  acting on the gas can dephase density), I think the authors should
  approach the point later in the paper. In my opinion, it would be
  more adequate to approach it at the end of or after the actual
  section 3 (single fluid description of dusty gas). Then it is
  relevant to make the parallel with the pure gas dynamics. It is also
  a way to explain the analogy that appears in parag. 3.3 "The
  physical reason for this analogy..."

  We appreciate the referee’s suggestion that section 2 might make
  sense if moved.  With respect, our choice for the placement of this
  section is meant for readers who are lack the technical expertise of
  the reviewer. Specifically, our goal is to summarize a main result
  and a key physical insight, which motivates the technical
  derivation of the dusty/adiabatic gas analogy. We added a new
  paragraph at the beginning of section 2 to state its purpose. 
  We hope section 2 motivates further reading and of course any reader
  can revisit  section 2 (section 4.4 refers back to it).

While the referee is certainly correct that the single fluid analogy
needs to made, we have attempted to explain the concept in the
introduction. In any event, we hope the referee will agree to some
authorial prerogative in the order of presentation.

* In section 6 on the numerical results, in particular parag. 6.5 and
  6.6, I wonder which growth rates are obtained with different radial
  mode numbers. It would be an advantage to see other values as in Fig
  8. In 6.5 kx is 30, does the instability change with smaller, larger
  modes? It is worse in 6.6, where kx is very large, 1800. Is it
  relevant to have such a high radial mode, that could barely be
  resolved in non linear numerical simulations? Finally, in a general,
  I think that a more physical summary of the results of the section 6
  should be provided. It is almost limited now to a description of the
  results.

  For section 6.5 we have updated Fig. 14 to include cases with
  different kx. All growth rates remain the same order of
  magnitude. The effect of a radial varation in rhod/rhog is also
  similar between different kx. However, surface modes only appear at
  high kx (as found for pure gas VSI, Barker & Latter, 2017)

  For section 6.6 we also updated Fig. 15 with more kx values. 
  We find to obtain instability with smaller kx, one needs a more
  rapidly-varying dust-to-gas ratio. We do not dwell 
  on this case because we have already shown, using the more general
  analyses in section 4.3 and 5.2

  Note that while the original case with high kx=1800 may not be  
  resolved in global simulations, it can be resolved in local shearing
  boxes. Simulations of SI in fact consider even larger
  kx = O(10^3/h), where h= H/R is the gas aspect-ratio. See, Bai &
  Stone (2010) and Yang & Johansen (2017), for example.

  In section 8 we extend the summary on the results of section
  6. We discuss possible implications of our results in 
  realistic disks with finite stopping times. In particular, we
  suggest the criterion for particles to settle against the VSI.

* The section 7 'Discussion' is for the moment a list of common
  instabilities that could be triggered by the effect of the drag in
  the framework of the one-fluid approximation. Even if it is very
  interesting, it may be too speculative; I would not limit the
  discussion only to this. The authors may first discuss more about
  the limitations/assets of their approach based on the numerical
  results obtained in section 6, which will give arguments to the
  possible numerical implementation of the method (see parag. 7.3). In
  particular, what is the impact of the terminal velocity
  approximation? How much reliable the results and the accuracy of the
  approach are with a finite St number? For example, what do you
  expect for a typical value of 0.01-0.1 (dust or pebble), for which
  it is known that the terminal velocity is inappropriate?

  We have added a new section 7.1 on the application and limitations
  of the one-fluid framework.

  The terminal velocity approximation from the requires ts to be small
  compared to other timescales in the problem. Since 1/Omega is the
  typical `fast' timescale for disks, we generally require ts*Omega <<
  1, but the exact critical value is problem-dependent.

  Fig. 4 showsthat the one-fluid framework can gives accurate
  streaming instability growth rates in dust-rich disks even for
  ts*Omega = 0.1. However, the dust-poor `linB' mode in Table 1 (also
  with ts*Omega=0.1) is not accurately captured by the one-fluid framework. 
  This suggest that the terminal velocity approximation is not the
  only condition required for the one-fluid equations to give accurate
  results. In the case of the streaming instability, it seems that one
  also requires rhod > rhog.

* In the Appendix: parag. D.2.2, the authors check if two assumptions
  done during the calculations are valid (eq D11 and eq D16). But it
  seems that they compute the values with the results obtained after
  using these approximations. Please clarify

  We retitled this section to `Consistency check' since, as the
  referee pointed out, we simply checked that the assumptions made to
  obtain the explicit solutions are self-consistent. The
  approximations are ultimately justified by comparison to numerical
  solutions in Fig. 4.

* Last point: the authors should re read carefully the article, indeed
  some sentences are not completly made agree with their later
  modifications.

  We have proof-read the new manuscript.

Minor points:

* parag 4.4.1 Can you explain briefly in the manuscript how to obtain
  the eq 33. ?

  Additional description (Eq. 33) is now included.

* parag. 5.1.2 Can you clarify the text related to eq 47, one cannot
  see clearly which terms are the second and the third. Maybe one can
  say : "The first term in brackets is usually stabilizing ... "

  This is clarified.

### 2017-09-12 appendix.tex
first draft of revision 1 and response to referee

@@
 \end{align}
-We show in \S\ref{si_dust_rich} that this is readily satisfied. Then
+We show in \S\ref{si_dust_rich_approx} that this is readily satisfied. Then
 $\mu \simeq \mathcal{Q}^{1/3}$.
@@
 
-\subsubsection{Validity of the approximations}\label{si_dust_rich_approx}
+\subsubsection{Consistency check}\label{si_dust_rich_approx}
+{\bf Here we check if the above explicit solutions are
+  consistent with the assumptions used to obtain them.  
+} 
 Inserting the solutions for the most unstable mode 
 %(Eq. \ref{si_dust_rich_optkx}---\ref{si_dust_rich_grow} with ) 
-we find the assumption in Eq. \ref{min_kz} becomes  
+{\bf we find Eq. \ref{min_kz} becomes}
 \begin{align}
@@
 $\mu$ in Eq. \ref{mu_explicit}.
+
+{\bf Ultimately, the validity of these assumptions are justified
+   a posteriori by comparison with the solution to the full equations
+   in Fig. \ref{si_compare_fig}. 
+}
 

### 2017-09-12 linear.tex
first draft of revision 1 and response to referee

@@
 \end{align}
-for \emph{stability}. The second term is usually stabilizing in 
-PPDs for reasons given in \S\ref{iso_perfect}. The third term is
+for \emph{stability}. The {\bf first term in  brackets} is usually stabilizing in 
+PPDs for reasons given in \S\ref{iso_perfect}. The {\bf second term in
+brackets} is 
 always destabilizing, but is small in smooth, thin disks with radial

### 2017-09-12 model.tex
first draft of revision 1 and response to referee

@@
 
-Noting that only products of perturbations contribute to
-$\mathcal{W}$, we find that for periodic oscillations (time dependence 
-$e^{\ii\omega t}$ and real $\omega$): 
+{\bf We consider perturbations such that 
+  \begin{align}
+ P \to P + \real\left(\Delta P e^{-\ii\sigma t}\right)
+ \end{align}
+ and similarly for $\rho$. Here, the Lagrangian 
+ perturbation $\Delta$ of a variable $X$ is $\Delta X = \delta X + \bm{\xi}\cdot\nabla X$
+ and $\bm{\xi}$ is the Lagrangian displacement ($\xi_{x,z} =  \ii\dd
+ v_{x,z}/\sigma$). Inserting the above pressure and density fields
+ into Eq. \ref{work_def}, and noting that only products of
+ perturbations contribute to $\mathcal{W}$ after time-averaging, we
+ find for periodic oscillations (time dependence $e^{\ii\omega t}$ and
+ real $\omega$) that:  
+}
 \begin{align}
@@
 \end{align}
-where %the integral is taken over the volume of the
-%mixture; and 
- the Lagrangian perturbation $\Delta$ of a variable $X$ is 
-$\Delta X = \delta X + \bm{\xi}\cdot\nabla X$ 
-and $\bm{\xi}$ is the
-Lagrangian displacement ($    \xi_{x,z} =  \ii \dd v_{x,z}/\sigma$).  
 Eq. \ref{work_real} show that a phase difference between gas pressure and

### 2017-09-12 summary.tex
first draft of revision 1 and response to referee

@@
 used to develop our thermodynamic model of dusty gas. 
-The terminal velocity approximation, $\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} = \tstop \nabla P /\rhog$, was employed from the outset. 
+The terminal velocity approximation, $\bm{v}_\mathrm{d} -
+\bm{v}_\mathrm{g} = \tstop \nabla P /\rhog$, was employed from the
+outset. 
 %It assumes  
-%the dust-gas relative drift $\bm{v}_mathrm{d} - \bm{v}_\mathrm{g} = \tstop \nabla P /\rhog$ 
-This is applicable to small particles with 
-$\tstop\OmK\ll 1$, but the precise critical value is problem-dependent. For the linear streaming instability,
-even $\tstop\OmK \sim 0.1$ can be captured by simplified one-fluid equations 
-(Fig. \ref{si_compare_fig}). On the other hand, the simplified equations contain spurious modes that
-may overwhelm physical instabilities in certain situations (see \S\ref{spurious_epi}).  %Ultimately, one must check against the full equations.
+%the dust-gas relative drift $\bm{v}_mathrm{d} - \bm{v}_\mathrm{g} =
+%\tstop \nabla P /\rhog$ 
+This is applicable to small particles with short stopping times and
+strongly coupled to the gas. Generally we require $\tstop$ to be the
+shortest timescale in the physical problem. For  
+example, to study dust settling, we require $\tstop \ll
+t_\mathrm{settle}\sim 1/\tstop\OmK^2$, or 
+$\tstop\OmK \ll 1$.
 
-However, our thermodynamic interpretation of dust-gas drift does \emph{not} in fact depend on the terminal velocity approximation.  
+However, the validity of the simplified one-fluid equations may be
+problem-dependent, and not just on the terminal velocity 
+approximation. For example, Table \ref{si_compare} shows the dust-rich
+`linA' mode and dust-poor `linB' mode of the streaming instability
+have the same $\tstop$, but the former is accurately captured by the
+simplified equations, while the latter is not.  This suggests that for
+the streaming instability the simplified equations is better suited
+for $\rhod/\rhog>1$. The simplified equations may also contain spurious
+modes in certain situations (see \S\ref{spurious_epi}).
+
+%Ultimately,
+                                %one must 
+                                %c heck against the full equations.
+
+Note that our thermodynamic interpretation of dust-gas drift does
+\emph{not} in fact depend on the terminal velocity approximation.   
 Once the gas equation of state is fixed and the true energy equation deleted, 
 the dust continuity equation can be converted to a new effective energy equation. 
-%This source term reflects  
-%a heat flux proportional to $\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g}$, which must generally be solved via a separate 
-%evolutionary equation \citep{laibe14}. Nevertheless, this does not change the interpretation of dusty gas as a single ideal fluid subject to cooling.  
 As an example, for strictly isothermal dusty gas we obtain   
 \begin{align}
-\frac{DP}{Dt} = - P\nabla\cdot\bm{v} + c_s^2\nabla\cdot\left[\fdust\left(1-\fdust\right)\rho\left(\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g}\right)\right], \label{no_term_vel} 
+\frac{DP}{Dt} = - P\nabla\cdot\bm{v} +
+c_s^2\nabla\cdot\left[\fdust\left(1-\fdust\right)\rho\left(\bm{v}_\mathrm{d}
+    - \bm{v}_\mathrm{g}\right)\right]. \label{no_term_vel}  
 \end{align}
-\citep[see also][]{laibe14}. This equation does not use the terminal velocity approximation (which would give Eq. \ref{eff_energy}), but 
- we see that dust-gas relative drift can be regarded as a heat flux within the mixture. Calculating this heat flux generally requires solving the evolutionary equation for 
-$\bm{v}_\mathrm{d} - \bm{v}_\mathrm{g}$. Nevertheless, one can still consider the dusty-gas as a single fluid subject to cooling. 
+This equation does not use
+the terminal velocity approximation (which gives
+Eq. \ref{eff_energy}); evaluating the right-hand-side generally 
+requires solving an evolutionary equation for $\bm{v}_\mathrm{d} - 
+\bm{v}_\mathrm{g}$ \citep{laibe14}. Nevertheless,
+Eq. \ref{no_term_vel} shows that dust-gas relative drift can be
+interpreted as a heat flux within the dust-gas mixture. Thus, even 
+without the terminal velocity approximation we can 
+regard the dusty-gas as a single ideal fluid subject to cooling.   
 }
@@
   1/h_\mathrm{g}\OmK$. This suggest that particles with $\tstop\OmK\lesssim 
-  h_g$ cannot settle against the VSI. On the 
+  h_\mathrm{g}$ cannot settle against the VSI. On the 
   other hand, larger particles with $\tstop\OmK\gtrsim h_g$ should

### 2017-09-12 response.txt
first draft of revision 1 and response to referee

@@
   rapidly-varying dust-to-gas ratio. We do not dwell 
-  on this case because we have already shown, using the more general
-  analyses in section 4.3 and 5.2
+  on this section because we have already shown, using the more general
+  analyses in section 4.3 and 5.2, that disks with a radial variation in
+  rhod/rhog are unstable if dust is vertically well-mixed.
 

### 2017-09-15 linear.tex
additional edits in response to guillaume's comments

@@
 $\hgas\simeq 0.05$, then we require $H_\epsilon/\Hg\gg
-2\times10^{-3}$. This can be met if the dust is not well settled. 
-%(So the dust layer thickness $\Hd\gg 2
-%\times10^{-3}\Hgas$ as well.) 
+2\times10^{-3}$. This can be met if the dust is not well settled {\bf (e.g. due to a small amount of turbulence).}  
+%this turbulence cannot be too big because the analysis itself does not have diffusion/turbulence 
+%large diffusion is expected to stabilize 
 Then instability requires 
@@
 compare the simplified one-fluid and full two-fluid frameworks in
-calculating SI.
+calculating SI \citep[cf. comparisons using other problems, ][]{laibe14,price15}.
 

### 2017-09-15 model.tex
additional edits in response to guillaume's comments

@@
   \bm{v}_\mathrm{d} - \bm{v}_\mathrm{g} = \frac{\nabla
-    P}{\rhog}\tstop 
+    P}{\rhog}\tstop\label{term_vel_approx} 
 \end{align}
@@
 detail. Consider constant $c_s$ but $\tstop\neq0$ (so $\delta\mathcal{C}
-\neq 0$). Eq. \ref{int_rel} indicates that $\sigma^2$ is generally complex, so
+\neq 0$). Eq. \ref{int_rel} indicates that $\sigma^2$ is generally complex, {\bf unless
+the second integral on the right-hand-side is real. Thus } 
 we may have growing oscillations, or overstability, due to dust-gas
@@
 The physical interpretation of Eq. \ref{thermal_instability} is that 
-work done by pressure forces in the dusty gas leads to growth in
-oscillation amplitudes. To demonstrate this, we calculate the average
+work done by pressure forces in the dusty gas leads to growth {\bf ($s>0$) or decay ($s<0$)} 
+%{\bf work done by pressure forces in the dusty gas modifies
+in oscillation amplitudes. 
+To demonstrate this, we calculate the average
 work done assuming periodic oscillations, and show that if the average
-work done is positive, then the 
-oscillation amplitude would actually grow.
+work done is positive, then the oscillation amplitude would actually grow.
 
@@
 
-{\bf We consider perturbations such that 
+{\bf Here we consider Lagragian perturbations such that 
   \begin{align}
@@
  \end{align}
- and similarly for $\rho$. Here, the Lagrangian 
- perturbation $\Delta$ of a variable $X$ is $\Delta X = \delta X + \bm{\xi}\cdot\nabla X$
+is the pressure of a fluid element (and similarly for $\rho$). 
+ The Lagrangian perturbation $\Delta$ of a variable $X$ is $\Delta X = \delta X + \bm{\xi}\cdot\nabla X$
  and $\bm{\xi}$ is the Lagrangian displacement ($\xi_{x,z} =  \ii\dd
@@
 The relative 
-drift between gas and dust causes a phase difference between the two
+drift between gas and dust{\bf, which only exists if $\tstop\nabla P \neq 0$ (Eq. \ref{term_vel_approx}),} causes a phase difference between the two
 components, and hence between pressure and total density.

### 2017-09-15 paper.tex
additional edits in response to guillaume's comments

@@
 fluid 
-\citep{paardekooper06b,meheut12,laibe12,loren14,fu14b,surville16}. The 
+\citep{barriere-fouchet05,paardekooper06b,ayliffe12,meheut12,laibe12,loren14,fu14b,surville16}. The 
 hydrodynamic equations are evolved for two fluids: the gas and dust with density and
@@
 $\rhog$ are not necessarily in phase, because dust particles can
-drift relative to the gas.   
+drift relative to the gas{\bf, i.e. the dust-to-gas ratio evolves in time}. 
 If this causes $P$ to lag 
@@
 We thank P. Loren-Aguilar for initial discussions that motivated this study. We also thank 
-C. Baruteau, R. Dong, J. Fung, S. Inutsuka, W. Lyra, M. Pessah, and O. Umurhan 
-for comments during the course of this project. This work is supported
-by the Theoretical Institute for Advanced Research in Astrophysics  
+C. Baruteau, R. Dong, J. Fung, S. Inutsuka, G. Laibe, W. Lyra, S.-J. Paardekooper, 
+M. Pessah, and O. Umurhan for comments during the course of this project. 
+This work is supported by the Theoretical Institute for Advanced Research in Astrophysics   
 (TIARA) based in Academia Sinica Institute of Astronomy and

### 2017-09-15 summary.tex
additional edits in response to guillaume's comments

@@
 \subsection{Implications for numerical simulations}\label{dust_sims}
-The dusty/adiabatic gas equivalence means that pure
-gas dynamics codes can be applied to simulate locally
-isothermal/polytropic dusty gas. One need not implement a separate 
-equation for the dust component: 
-%In fact, no
-%modification is needed  
+{\bf The one-fluid model with the terminal velocity approximation, Eq. \ref{masseq}---\ref{momeq}, 
+has been applied to simulate dusty protoplanetary disks \citep{dipierro15,ragusa17}. These studies 
+explicitly evolve the dust density.  
+}
+
+{\bf However, the dusty/adiabatic gas equivalence identified in this paper
+means that one need not implement a separate equation for the dust component in
+locally isothermal/polytropic gas. The standard energy equation substitutes for the dust continuity 
+equation. Thus pure gas dynamic codes can be used to simulate protoplanetary dusty disks. 
+} %means that pure 
+%gas dynamics codes can be applied to simulate locally 
+%isothermal/polytropic dusty gas. One need not implement a separate 
+%equation for the dust component: 
 %for strictly isothermal or polytropic gas perfectly-coupled to dust:  
- the standard energy equation substitutes for the dust continuity
-equation. When $c_s$ (or $K$) is not constant and/or $\tstop\neq0$ one
+% the standard energy equation substitutes for the dust continuity
+%equation. 
+When {\bf the sound-speed} $c_s$ (or $K$) is not constant and/or the dust-gas coupling is imperfect, $\tstop\neq0$, one
 should add corresponding source terms in the energy equation
@@
 hydrodynamics code \citep{mignone07} into a dusty gas dynamics code appropriate for
-simulating protoplanetary disks coupled to small dust. We will apply 
+simulating protoplanetary disks coupled to small dust. {\bf In fact, our code is unaware it is modeling dusty gas.} 
+We will apply 
 this modified code ---d\textsc{Pluto}--- to study dusty disk-planet interaction (Lin et al., 
-in preparation). Preliminary results indicate the simplified one-fluid 
-approach can reproduce features such as dusty rings associated with 
+in preparation). 
+Preliminary results {\bf show our approach} %indicate the simplified one-fluid 
+reproduce features such as dusty rings associated with 
 planet-induced gaps similar to that obtained from explicit two-fluid
@@
   further settling. The result may be a quiet, dusty midplane \citep[unless non-axisymmetric
-  instabilities develop][]{chiang08} with VSI-turbulent gaseous
+  instabilities develop, ][]{chiang08} with VSI-turbulent gaseous
   atmospheres. 

### 2017-09-23 response.txt
minor mods to response (grammar only) and summary (a bit of content)

@@
   sense if moved.  With respect, our choice for the placement of this
-  section is meant for readers who are lack the technical expertise of
+  section is meant for readers who lack the technical expertise of
   the reviewer. Specifically, our goal is to summarize a main result
-  and a key physical insight, which motivates the technical
+  and a key physical insight, and thereby to motivate the technical
   derivation of the dusty/adiabatic gas analogy. We added a new
@@
   different kx. All growth rates remain the same order of
-  magnitude. The effect of a radial varation in rhod/rhog is also
+  magnitude. The effect of a radial variation in rhod/rhog is also
   similar between different kx. However, surface modes only appear at
@@
 
-  Fig. 4 showsthat the one-fluid framework can gives accurate
+  Fig. 4 shows that the one-fluid framework gives accurate
   streaming instability growth rates in dust-rich disks even for

### 2017-09-23 summary.tex
minor mods to response (grammar only) and summary (a bit of content)

@@
 growth rates. {\bf We also show that radial variations in
-  $\rhod/\rhog$ effectively translates to a non-uniform radial temperature
-  profile, and the resulting vertical shear can lead to instability,
-  even if the true sound-speed is constant. 
+  $\rhod/\rhog$ can trigger a type of VSI, even when the 
+  usual sources of vertical shear -- vertical dust gradients and radial 
+  temperature gradients -- are completely absent.  
+  %effectively translates to a non-uniform radial temperature
+  %profile, and the resulting vertical shear can lead to instability,
+  %even if the true sound-speed is constant. 
 }

### 2017-09-23 paper.tex
tiny edits before second submission

@@
 well-known that in the limit of perfect coupling the addition of dust
-increases the gas inertia but not pressure. One can then regard the
-dusty gas as a single mixture with a reduced sound-speed. Here, we
+increases the gas inertia but not pressure. One can then regard 
+dusty gas as a single fluid with a reduced sound-speed. Here, we
 argue that when dust-gas coupling is imperfect, there exists another
-similarity with pure gas related the phase of pressure and density
+similarity with pure gas; related the phase of pressure and density
 evolution. Thus it is still possible to regard

### 2017-09-23 summary.tex
tiny edits before second submission

@@
 The main application we envision for the dusty/adiabatic gas analogy 
-to develop physical interpretations of dust-gas instabilities; and to find 
+is to develop physical interpretations of dust-gas instabilities; and to find 
 dusty analogs of pure gas instabilities in protoplanetary disks, 
@@
 
-However, the validity of the simplified one-fluid equations may be
-problem-dependent, and not just on the terminal velocity 
-approximation. For example, Table \ref{si_compare} shows the dust-rich
+%However, the validity of the simplified one-fluid equations may be
+%problem-dependent, and not just on the terminal velocity 
+%approximation.
+
+However, the validity of the terminal velocity approximation depends
+not just on $\tstop$, but also on the value of the dust gas-ratio, and
+possibly the problem itself. 
+%You can decide if this is a relevant improvement, and whether to
+%include. 
+For example, Table \ref{si_compare} shows the dust-rich
 `linA' mode and dust-poor `linB' mode of the streaming instability
@@
 for $\rhod/\rhog>1$. The simplified equations may also contain spurious
-modes in certain situations (see \S\ref{spurious_epi}).
+modes in certain limits (see \S\ref{spurious_epi}).
 
@@
 \section{Summary}\label{summary}
-In this paper we develop the analogy between isothermal 
-dusty-gas and pure ideal gas. The correspondence arises
+{\bf
+In this paper, we examine the conditions under which the presence of
+dust can trigger instabilities in gaseous protoplanetary disks.  To this end, 
+we develop an analogy between isothermal  dusty-gas and pure ideal
+gas. 
+} 
+The correspondence arises
 because drag forces reduce the relative velocity between gas and
@@
   usual sources of vertical shear -- vertical dust gradients and radial 
-  temperature gradients -- are completely absent.  
+  temperature gradients -- are negligible.  
   %effectively translates to a non-uniform radial temperature
@@
   h_\mathrm{g}$ cannot settle against the VSI. On the 
-  other hand, larger particles with $\tstop\OmK\gtrsim h_g$ should
+  other hand, larger particles with $\tstop\OmK\gtrsim h_\mathrm{g}$ should
   settle to form a dusty midplane. In fact, our calculations suggest that

### 2017-09-23 model.tex
tiny edits before second submission
updated to revision 1 submitted version

@@
  \end{align}
-is the pressure of a fluid element (and similarly for $\rho$). 
+is the pressure following a fluid element of the mixture (and similarly for $\rho$). 
  The Lagrangian perturbation $\Delta$ of a variable $X$ is $\Delta X = \delta X + \bm{\xi}\cdot\nabla X$
- and $\bm{\xi}$ is the Lagrangian displacement ($\xi_{x,z} =  \ii\dd
- v_{x,z}/\sigma$). Inserting the above pressure and density fields
+ and $\bm{\xi}$ is the Lagrangian displacement, so $\xi_{x,z} =  \ii\dd
+ v_{x,z}/\sigma$. Inserting the above pressure and density fields
  into Eq. \ref{work_def}, and noting that only products of

### 2017-09-23 result.tex
updated to revision 1 submitted version

@@
 enhanced by the additional vertical shear induced by
-$\p_r\epsilon$. {\bf We find that low frequency body modes with small
+$\p_r\epsilon$. {\bf Low frequency body modes with small
   $k_x$ are more affected by radial variations in the 
-  dust-to-gas ratio.}
+  dust-to-gas ratio than high-$k_x$, high-frequency body modes. }
 
@@
   with $\rhod/\rhog\propto r^{-1}$. 
-  Local shearing box simulations \citep[e.g.][]{bai10b,yang16} may 
+  Local shearing box simulations may 
   be required to study the non-linear evolution of such short 
-  wavelengths. Alternatively, as shown in Fig. \ref{vert_mixed_modes},
+  wavelengths \citep[e.g.][]{bai10b,yang16}. Alternatively, as shown in Fig. \ref{vert_mixed_modes},
   a rapidly varying dust-to-gas ratio permits dynamical instability at
-  longer radial wavelengths. 
+  longer radial wavelengths, which might be resolvable in global disk 
+  simulations. 
 }

### 2017-09-23 summary.tex
updated to revision 1 submitted version

@@
 The main application we envision for the dusty/adiabatic gas analogy 
-is to develop physical interpretations of dust-gas instabilities; and to find 
+is to develop physical interpretations of dust-gas drag instabilities; and to find 
 dusty analogs of pure gas instabilities in protoplanetary disks, 
@@
 
-However, the validity of the terminal velocity approximation depends
-not just on $\tstop$, but also on the value of the dust gas-ratio, and
-possibly the problem itself. 
+However, the validity of the terminal velocity approximation may not
+only depend on $\tstop$. The value of the dust gas-ratio and the
+problem itself may also be important. 
 %You can decide if this is a relevant improvement, and whether to
@@
 the streaming instability the simplified equations is better suited
-for $\rhod/\rhog>1$. The simplified equations may also contain spurious
+for $\rhod/\rhog>1$. The simplified equations also contain spurious
 modes in certain limits (see \S\ref{spurious_epi}).
@@
 Note that our thermodynamic interpretation of dust-gas drift does
-\emph{not} in fact depend on the terminal velocity approximation.   
+\emph{not} actually depend on the terminal velocity approximation.   
 Once the gas equation of state is fixed and the true energy equation deleted, 
@@
 the terminal velocity approximation (which gives
-Eq. \ref{eff_energy}); evaluating the right-hand-side generally 
+Eq. \ref{eff_energy}). Evaluating the right-hand-side generally 
 requires solving an evolutionary equation for $\bm{v}_\mathrm{d} - 
@@
 {\bf However, the dusty/adiabatic gas equivalence identified in this paper
-means that one need not implement a separate equation for the dust component in
+means that one need not implement a separate module for simulating the dust component in
 locally isothermal/polytropic gas. The standard energy equation substitutes for the dust continuity 
@@
 hydrodynamics code \citep{mignone07} into a dusty gas dynamics code appropriate for
-simulating protoplanetary disks coupled to small dust. {\bf In fact, our code is unaware it is modeling dusty gas.} 
+simulating protoplanetary disks coupled to small dust. {\bf In fact,
+  our code is unaware of the fact that it is modeling dusty gas.} 
 We will apply 

### 2017-09-23 appendix.tex
additional edits in response to guillaume's comments
update to re-submitted version, unbolded

@@
 
-\subsection{Incompressible limit}
+\subsection{Incompressible { gas} limit}
 
  We can set $\zeta = 0$ or consider $c_s^2\to \infty$ to
-obtain the incompressible limit. Using $\tau_\mathrm{s} \equiv
+obtain the incompressible { gas} limit. Using $\tau_\mathrm{s} \equiv
 \tstop/\left(1 - \tepsilon\right)$, we obtain: 
@@
 \subsubsection{Consistency check}\label{si_dust_rich_approx}
-{\bf Here we check if the above explicit solutions are
+{ Here we check if the above explicit solutions are
   consistent with the assumptions used to obtain them.  
@@
 %(Eq. \ref{si_dust_rich_optkx}---\ref{si_dust_rich_grow} with ) 
-{\bf we find Eq. \ref{min_kz} becomes}
+{ we find Eq. \ref{min_kz} becomes}
 \begin{align}
@@
 
-{\bf Ultimately, the validity of these assumptions are justified
+{ Ultimately, the validity of these assumptions are justified
    a posteriori by comparison with the solution to the full equations

### 2017-09-23 linear.tex
update to re-submitted version, unbolded

@@
 \end{align}
-for \emph{stability}. The {\bf first term in  brackets} is usually stabilizing in 
-PPDs for reasons given in \S\ref{iso_perfect}. The {\bf second term in
+for \emph{stability}. The { first term in  brackets} is usually stabilizing in 
+PPDs for reasons given in \S\ref{iso_perfect}. The { second term in
 brackets} is 
@@
 $\hgas\simeq 0.05$, then we require $H_\epsilon/\Hg\gg
-2\times10^{-3}$. This can be met if the dust is not well settled {\bf (e.g. due to a small amount of turbulence).}  
+2\times10^{-3}$. This can be met if the dust is not well settled { (e.g. due to a small amount of turbulence).}  
 %this turbulence cannot be too big because the analysis itself does not have diffusion/turbulence 

### 2017-09-23 model.tex
update to re-submitted version, unbolded

@@
 detail. Consider constant $c_s$ but $\tstop\neq0$ (so $\delta\mathcal{C}
-\neq 0$). Eq. \ref{int_rel} indicates that $\sigma^2$ is generally complex, {\bf unless
+\neq 0$). Eq. \ref{int_rel} indicates that $\sigma^2$ is generally complex, { unless
 the second integral on the right-hand-side is real. Thus } 
@@
 The physical interpretation of Eq. \ref{thermal_instability} is that 
-work done by pressure forces in the dusty gas leads to growth {\bf ($s>0$) or decay ($s<0$)} 
+work done by pressure forces in the dusty gas leads to growth { ($s>0$) or decay ($s<0$)} 
 %{\bf work done by pressure forces in the dusty gas modifies
@@
 
-{\bf Here we consider Lagragian perturbations such that 
+{ Here we consider Lagragian perturbations such that 
   \begin{align}
@@
 The relative 
-drift between gas and dust{\bf, which only exists if $\tstop\nabla P \neq 0$ (Eq. \ref{term_vel_approx}),} causes a phase difference between the two
+drift between gas and dust{, which only exists if $\tstop\nabla P \neq 0$ (Eq. \ref{term_vel_approx}),} causes a phase difference between the two
 components, and hence between pressure and total density.

### 2017-09-23 paper.tex
updated to revision 1 submitted version
update to re-submitted version, unbolded

@@
 In \S\ref{grow_osc} we describe a key insight of our study, that
-pressure (i.e.\ ``PdV")  work {\bf can} explain the basis of several
+pressure (i.e.\ ``PdV")  work {can} explain the basis of several
 instabilities, including SI and VSI, in terms of 
-pressure-density phase lags. {\bf This motivates us to seek an 
+pressure-density phase lags. { This motivates us to seek an 
   analogy between dust-gas mixtures and pure gas.}    
@@
 \section{Growing oscillations by doing work}\label{grow_osc}
-{\bf
+{
 In this section we provide physical arguments to highlight the
-similarity between single phase fluids and dust-gas mixtures. It is
+similarity between single phase fluids and dust-gas mixtures. The mathematical
+description of this section is deferred to \S\ref{dust_work}.
+
+It is
 well-known that in the limit of perfect coupling the addition of dust
@@
 evolution. Thus it is still possible to regard
-partially-coupled dusty-gas as a single fluid. The mathematical
-description of this section is given in \S\ref{dust_work}.   
+partially-coupled dusty-gas as a single fluid.  
 }
@@
 $\rhog$ are not necessarily in phase, because dust particles can
-drift relative to the gas{\bf, i.e. the dust-to-gas ratio evolves in time}. 
+drift relative to the gas{, i.e. the dust-to-gas ratio evolves in time}. 
 If this causes $P$ to lag 

### 2017-09-23 result.tex
update to re-submitted version, unbolded

@@
 dust-to-gas ratio increases the magnitude of the vertical shear rate
-away from the midplane (see also Eq. \ref{vshear2}). Thus we {\bf
+away from the midplane (see also Eq. \ref{vshear2}). Thus we {
   typically} find higher VSI growth rates, as shown in
 Fig. \ref{vsi_dust_varHd}. 
-Surfaces modes, {\bf appearing at high $k_x$}, are more effectively
+Surfaces modes, { appearing at high $k_x$}, are more effectively
 enhanced by the additional vertical shear induced by
-$\p_r\epsilon$. {\bf Low frequency body modes with small
+$\p_r\epsilon$. { Low frequency body modes with small
   $k_x$ are more affected by radial variations in the 
@@
   \caption{Unstable VSI modes in the disk models of
-    Fig. \protect\ref{compare_vshear_varHd}. {\bf Diamonds (asterisks) 
+    Fig. \protect\ref{compare_vshear_varHd}. { Diamonds (asterisks) 
     are mode frequencies for a disk with radially uniform (varying)
@@
 if $d\epsilon/dr\neq0$.   
-{\bf
+{
 To demonstrate this numerically we set $q=0$ and $H_\mathrm{d}$ such
@@
 
-{\bf We show unstable modes in 
+{ We show unstable modes in 
 Fig. \ref{vert_mixed_modes}.} 
@@
 
-{\bf We find in smooth disks large 
+{ We find in smooth disks large 
   $k_x$ is needed for appreciable growth rates, e.g. $k_xH_g=1800$

### 2017-09-23 summary.tex
update to re-submitted version, unbolded

@@
 
-{\bf
+{
 
@@
 \subsection{Implications for numerical simulations}\label{dust_sims}
-{\bf The one-fluid model with the terminal velocity approximation, Eq. \ref{masseq}---\ref{momeq}, 
+{ The one-fluid model with the terminal velocity approximation, Eq. \ref{masseq}---\ref{momeq}, 
 has been applied to simulate dusty protoplanetary disks \citep{dipierro15,ragusa17}. These studies 
@@
 
-{\bf However, the dusty/adiabatic gas equivalence identified in this paper
+{ However, the dusty/adiabatic gas equivalence identified in this paper
 means that one need not implement a separate module for simulating the dust component in
@@
 %equation. 
-When {\bf the sound-speed} $c_s$ (or $K$) is not constant and/or the dust-gas coupling is imperfect, $\tstop\neq0$, one
+When { the sound-speed} $c_s$ (or $K$) is not constant and/or the dust-gas coupling is imperfect, $\tstop\neq0$, one
 should add corresponding source terms in the energy equation
@@
 hydrodynamics code \citep{mignone07} into a dusty gas dynamics code appropriate for
-simulating protoplanetary disks coupled to small dust. {\bf In fact,
+simulating protoplanetary disks coupled to small dust. { In fact,
   our code is unaware of the fact that it is modeling dusty gas.} 
@@
 in preparation). 
-Preliminary results {\bf show our approach} %indicate the simplified one-fluid 
+Preliminary results { show our approach} %indicate the simplified one-fluid 
 reproduce features such as dusty rings associated with 
@@
 \section{Summary}\label{summary}
-{\bf
+{
 In this paper, we examine the conditions under which the presence of
@@
 `surface modes'  of the VSI, that would otherwise have the largest
-growth rates. {\bf We also show that radial variations in
+growth rates. { We also show that radial variations in
   $\rhod/\rhog$ can trigger a type of VSI, even when the 
@@
 %implications for realistic disks
-{\bf In a realistic disk, dust particles settle on a timescale $
+{ In a realistic disk, dust particles settle on a timescale $
   t_\mathrm{settle} \sim 1/\tstop\OmK^2$ \citep{takeuchi02}, compared 
